### 2016-07-16 Chapters/Chapter02.tex
First commit

%*****************************************
\chapter{Examples}\label{ch:examples}
%*****************************************
%\setcounter{figure}{10}
% \NoCaseChange{Homo Sapiens}
Ei choro aeterno antiopam mea, labitur bonorum pri no
\citeauthor{taleb:2012} \citep{taleb:2012}. His no decore
nemore graecis. In eos meis nominavi, liber soluta vim cu. Sea commune
suavitate interpretaris eu, vix eu libris efficiantur.

\section{A New Section}
Illo principalmente su nos. Non message \emph{occidental} angloromanic
da. Debitas effortio simplificate sia se, auxiliar summarios da que,
se avantiate publicationes via. Pan in terra summarios, capital
interlingua se que. Al via multo esser specimen, campo responder que
da. Le usate medical addresses pro, europa origine sanctificate nos
se.

Examples: \textit{Italics}, \spacedallcaps{All Caps}, \textsc{Small
Caps}, \spacedlowsmallcaps{Low Small Caps}.

Acronym testing: \ac{UML} -- \acs{UML} -- \acf{UML} -- \acp{UML}

\subsection{Test for a Subsection}
\graffito{Note: The content of this chapter is just some dummy text.
It is not a real language.}
Lorem ipsum at nusquam appellantur his, ut eos erant homero
concludaturque. Albucius appellantur deterruisset id eam, vivendum
partiendo dissentiet ei ius. Vis melius facilisis ea, sea id convenire
referrentur, takimata adolescens ex duo. Ei harum argumentum per. Eam
vidit exerci appetere ad, ut vel zzril intellegam interpretaris.

Errem omnium ea per, pro \ac{UML} con populo ornatus cu, ex qui
dicant nemore melius. No pri diam iriure euismod. Graecis eleifend
appellantur quo id. Id corpora inimicus nam, facer nonummy ne pro,
kasd repudiandae ei mei. Mea menandri mediocrem dissentiet cu, ex
nominati imperdiet nec, sea odio duis vocent ei. Tempor everti
appareat cu ius, ridens audiam an qui, aliquid admodum conceptam ne
qui. Vis ea melius nostrum, mel alienum euripidis eu.

Ei choro aeterno antiopam mea, labitur bonorum pri no. His no decore
nemore graecis. In eos meis nominavi, liber soluta vim cu.

\subsection{Autem Timeam}
Nulla fastidii ea ius, exerci suscipit instructior te nam, in ullum
postulant quo. Congue quaestio philosophia his at, sea odio autem
vulputate ex. Cu usu mucius iisque voluptua. Sit maiorum propriae at,
ea cum \ac{API} primis intellegat. Hinc cotidieque reprehendunt eu
nec. Autem timeam deleniti usu id, in nec nibh altera.

%Equidem detraxit cu nam, vix eu delenit periculis. Eos ut vero
%constituto, no vidit propriae complectitur sea. Diceret nonummy in
%has, no qui eligendi recteque consetetur. Mel eu dictas suscipiantur,
%et sed placerat oporteat. At ipsum electram mei, ad aeque atomorum
%mea.
%
%Ei solet nemore consectetuer nam. Ad eam porro impetus, te choro omnes
%evertitur mel. Molestie conclusionemque vel at.

\section{Another Section in This Chapter} % \ensuremath{\NoCaseChange{\mathbb{ZNR}}}
Non vices medical da. Se qui peano distinguer demonstrate, personas
internet in nos. Con ma presenta instruction initialmente, non le toto
gymnasios, clave effortio primarimente su del.\footnote{Uno il nomine
integre, lo tote tempore anglo-romanic per, ma sed practic philologos
historiettas.}

Sia ma sine svedese americas. Asia \citeauthor{bentley:1999}
\citep{bentley:1999} representantes un nos, un altere membros
qui.\footnote{De web nostre historia angloromanic.} Medical
representantes al uso, con lo unic vocabulos, tu peano essentialmente
qui. Lo malo laborava anteriormente uso.

\begin{description}
  \item[Description-Label Test:] Illo secundo continentes sia il, sia
  russo distinguer se. Contos resultato preparation que se, uno
  national historiettas lo, ma sed etiam parolas latente. Ma unic
  quales sia. Pan in patre altere summario, le pro latino resultato.
    \item[basate americano sia:] Lo vista ample programma pro, uno
    europee addresses ma, abstracte intention al pan. Nos duce infra
    publicava le. Es que historia encyclopedia, sed terra celos
    avantiate in. Su pro effortio appellate, o.
\end{description}

Tu uno veni americano sanctificate. Pan e union linguistic
\citeauthor{cormen:2001} \citep{cormen:2001} simplificate, traducite
linguistic del le, del un apprende denomination.

\subsection{Personas Initialmente}
Uno pote summario methodicamente al, uso debe nomina hereditage ma.
Iala rapide ha del, ma nos esser parlar. Maximo dictionario sed al.

\subsubsection{A Subsubsection}
Deler utilitate methodicamente con se. Technic scriber uso in, via
appellate instruite sanctificate da, sed le texto inter encyclopedia.
Ha iste americas que, qui ma tempore capital. \citeauthor{dueck:trio} \citep{dueck:trio}

\begin{aenumerate}
    \item Enumeration with small caps (alpha)
    \item Second item
\end{aenumerate}

\paragraph{A Paragraph Example} Uno de membros summario preparation,
es inter disuso qualcunque que. Del hodie philologos occidental al,
como publicate litteratura in web. Veni americano \citeauthor{knuth:1976}
\citep{knuth:1976} es con, non internet millennios secundarimente ha.
Titulo utilitate tentation duo ha, il via tres secundarimente, uso
americano initialmente ma. De duo deler personas initialmente. Se
duce facite westeuropee web, \autoref{tab:example} nos clave
articulos ha.

Medio integre lo per, non \citeauthor{sommerville:1992}
\citep{sommerville:1992} es linguas integre. Al web altere integre
periodicos, in nos hodie basate. Uno es rapide tentation, usos human
synonymo con ma, parola extrahite greco-latin ma web. Veni signo
rapide nos da.

%Se russo proposito anglo-romanic pro, es celos westeuropee
%incorporate uno. Il web unic periodicos. Que usate scientia ma, sed
%tres unidirectional al, asia personas duo de. De sed russo nomina
%anteriormente, toto resultato anteriormente uno ma. Non se signo
%romanic technologia, un medio millennios con.

%Major facto sia es, con o titulo maximo international. Inviar
%publicationes con in, uno le parola tentation, pan de studio romanic
%greco-latin. Tu duo titulo debitas latente, que vista programma ma.
%Non tote tres germano se, lo parola periodicos non.

\begin{table}
    \myfloatalign
  \begin{tabularx}{\textwidth}{Xll} \toprule
    \tableheadline{labitur bonorum pri no} & \tableheadline{que vista}
    & \tableheadline{human} \\ \midrule
    fastidii ea ius & germano &  demonstratea \\
    suscipit instructior & titulo & personas \\
    %postulant quo & westeuropee & sanctificatec \\
    \midrule
    quaestio philosophia & facto & demonstrated \citeauthor{knuth:1976} \\
    %autem vulputate ex & parola & romanic \\
    %usu mucius iisque & studio & sanctificatef \\
    \bottomrule
  \end{tabularx}
  \caption[Autem timeam deleniti usu id]{Autem timeam deleniti usu
  id. \citeauthor{knuth:1976}}  \label{tab:example}
\end{table}

\enlargethispage{2cm}
\subsection{Linguistic Registrate}
Veni introduction es pro, qui finalmente demonstrate il. E tamben
anglese programma uno. Sed le debitas demonstrate. Non russo existe o,
facite linguistic registrate se nos. Gymnasios, \eg, sanctificate sia
le, publicate \autoref{fig:example} methodicamente e qui.

Lo sed apprende instruite. Que altere responder su, pan ma, \ie, signo
studio. \autoref{fig:example-b} Instruite preparation le duo, asia
altere tentation web su. Via unic facto rapide de, iste questiones
methodicamente o uno, nos al.

%*****************************************
%*****************************************
%*****************************************
%*****************************************
%*****************************************

### 2016-07-17 FrontBackmatter/Publications.tex
Clean up style configuration

%*******************************************************
% Publications
%*******************************************************
\pdfbookmark[1]{Publications}{publications}
\chapter*{Publications}%\graffito{This is just an early --~and currently ugly~-- test!}

This thesis is based on the following scientific papers:

%\noindent Put your publications from the thesis here. The packages
%\texttt{multibib} or \texttt{bibtopic} etc. can be used to handle
%multiple different bibliographies in your document.

\begin{refsection}[ownpubs]
    \nocite{*} % is local to to the enclosing refsection
    \printbibliography[env=nolabelbib, notcategory=cited, heading=none]
\end{refsection}

### 2016-07-17 Chapters/Chapter02.tex
Adjusted inclusion of papers.

@@
 qui. Lo malo laborava anteriormente uso.
-
-\begin{description}
-  \item[Description-Label Test:] Illo secundo continentes sia il, sia
-  russo distinguer se. Contos resultato preparation que se, uno
-  national historiettas lo, ma sed etiam parolas latente. Ma unic
-  quales sia. Pan in patre altere summario, le pro latino resultato.
-    \item[basate americano sia:] Lo vista ample programma pro, uno
-    europee addresses ma, abstracte intention al pan. Nos duce infra
-    publicava le. Es que historia encyclopedia, sed terra celos
-    avantiate in. Su pro effortio appellate, o.
-\end{description}
 

### 2016-07-17 FrontBackmatter/Publications.tex
Adjusted inclusion of papers.

@@
 %*******************************************************
-\pdfbookmark[1]{Publications}{publications}
+\thispagestyle{empty}
+\refstepcounter{dummy}
+\pdfbookmark[1]{Publications}{Publications}
 \chapter*{Publications}%\graffito{This is just an early --~and currently ugly~-- test!}
@@
 
-%\noindent Put your publications from the thesis here. The packages
-%\texttt{multibib} or \texttt{bibtopic} etc. can be used to handle
-%multiple different bibliographies in your document.
+\begin{enumerate}[label=\textbf{\Roman{*} },ref=\Roman{*}]
 
-\begin{refsection}[ownpubs]
-    \nocite{*} % is local to to the enclosing refsection
-    \printbibliography[env=nolabelbib, notcategory=cited, heading=none]
-\end{refsection}
+\item
+  \textbf{R. Di Remigio}, R. Bast, L. Frediani, and T. Saue:
+  "Four-Component Relativistic Calculations in Solution with the
+  Polarizable Continuum Model of Solvation: Theory,
+  Implementation, and Application to the Group 16 Dihydrides
+  \ce{H2X} (\ce{X} = \ce{O}, \ce{S}, \ce{Se}, \ce{Te},
+  \ce{Po})",
+\textit{J. Phys. Chem. A}, \textrm{2015}, \textbf{119}, 5061--5077
+\label{relapcm}
+
+\item
+  M. Bugeanu, \textbf{R. Di Remigio}, K. Mozgawa, S. S. Reine, H.
+  Harbrecht,  and L. Frediani:
+  "Wavelet Formulation of the Polarizable Continuum Model. II. Use of
+  Piecewise Bilinear Boundary Elements",
+  \textit{Phys. Chem. Chem. Phys.}, \textrm{2015}, \textbf{17},
+  31566--31581
+\label{wemlin}
+
+\item \textbf{R. Di Remigio}, K. Mozgawa, H. Cao, V. Weijo, and L.
+  Frediani:
+  "A Polarizable Continuum Model for Molecules at Spherical
+  Diffuse Interfaces",
+  \textit{J. Chem. Phys.}, \textrm{2016}, \textbf{144}, 124103
+\label{spherical}
+
+\end{enumerate}

### 2016-07-17 Chapters/Chapter02.tex
Add notation chapter

@@
 %*****************************************
-\chapter{Examples}\label{ch:examples}
+\chapter{Response Theory and Molecular Properties}\label{ch:molprop}
 %*****************************************
-%\setcounter{figure}{10}
-% \NoCaseChange{Homo Sapiens}
-Ei choro aeterno antiopam mea, labitur bonorum pri no
-\citeauthor{taleb:2012} \citep{taleb:2012}. His no decore
-nemore graecis. In eos meis nominavi, liber soluta vim cu. Sea commune
-suavitate interpretaris eu, vix eu libris efficiantur.
-
-\section{A New Section}
-Illo principalmente su nos. Non message \emph{occidental} angloromanic
-da. Debitas effortio simplificate sia se, auxiliar summarios da que,
-se avantiate publicationes via. Pan in terra summarios, capital
-interlingua se que. Al via multo esser specimen, campo responder que
-da. Le usate medical addresses pro, europa origine sanctificate nos
-se.
-
-Examples: \textit{Italics}, \spacedallcaps{All Caps}, \textsc{Small
-Caps}, \spacedlowsmallcaps{Low Small Caps}.
-
-Acronym testing: \ac{UML} -- \acs{UML} -- \acf{UML} -- \acp{UML}
-
-\subsection{Test for a Subsection}
-\graffito{Note: The content of this chapter is just some dummy text.
-It is not a real language.}
-Lorem ipsum at nusquam appellantur his, ut eos erant homero
-concludaturque. Albucius appellantur deterruisset id eam, vivendum
-partiendo dissentiet ei ius. Vis melius facilisis ea, sea id convenire
-referrentur, takimata adolescens ex duo. Ei harum argumentum per. Eam
-vidit exerci appetere ad, ut vel zzril intellegam interpretaris.
-
-Errem omnium ea per, pro \ac{UML} con populo ornatus cu, ex qui
-dicant nemore melius. No pri diam iriure euismod. Graecis eleifend
-appellantur quo id. Id corpora inimicus nam, facer nonummy ne pro,
-kasd repudiandae ei mei. Mea menandri mediocrem dissentiet cu, ex
-nominati imperdiet nec, sea odio duis vocent ei. Tempor everti
-appareat cu ius, ridens audiam an qui, aliquid admodum conceptam ne
-qui. Vis ea melius nostrum, mel alienum euripidis eu.
-
-Ei choro aeterno antiopam mea, labitur bonorum pri no. His no decore
-nemore graecis. In eos meis nominavi, liber soluta vim cu.
-
-\subsection{Autem Timeam}
-Nulla fastidii ea ius, exerci suscipit instructior te nam, in ullum
-postulant quo. Congue quaestio philosophia his at, sea odio autem
-vulputate ex. Cu usu mucius iisque voluptua. Sit maiorum propriae at,
-ea cum \ac{API} primis intellegat. Hinc cotidieque reprehendunt eu
-nec. Autem timeam deleniti usu id, in nec nibh altera.
-
-%Equidem detraxit cu nam, vix eu delenit periculis. Eos ut vero
-%constituto, no vidit propriae complectitur sea. Diceret nonummy in
-%has, no qui eligendi recteque consetetur. Mel eu dictas suscipiantur,
-%et sed placerat oporteat. At ipsum electram mei, ad aeque atomorum
-%mea.
-%
-%Ei solet nemore consectetuer nam. Ad eam porro impetus, te choro omnes
-%evertitur mel. Molestie conclusionemque vel at.
-
-\section{Another Section in This Chapter} % \ensuremath{\NoCaseChange{\mathbb{ZNR}}}
-Non vices medical da. Se qui peano distinguer demonstrate, personas
-internet in nos. Con ma presenta instruction initialmente, non le toto
-gymnasios, clave effortio primarimente su del.\footnote{Uno il nomine
-integre, lo tote tempore anglo-romanic per, ma sed practic philologos
-historiettas.}
-
-Sia ma sine svedese americas. Asia \citeauthor{bentley:1999}
-\citep{bentley:1999} representantes un nos, un altere membros
-qui.\footnote{De web nostre historia angloromanic.} Medical
-representantes al uso, con lo unic vocabulos, tu peano essentialmente
-qui. Lo malo laborava anteriormente uso.
-
-Tu uno veni americano sanctificate. Pan e union linguistic
-\citeauthor{cormen:2001} \citep{cormen:2001} simplificate, traducite
-linguistic del le, del un apprende denomination.
-
-\subsection{Personas Initialmente}
-Uno pote summario methodicamente al, uso debe nomina hereditage ma.
-Iala rapide ha del, ma nos esser parlar. Maximo dictionario sed al.
-
-\subsubsection{A Subsubsection}
-Deler utilitate methodicamente con se. Technic scriber uso in, via
-appellate instruite sanctificate da, sed le texto inter encyclopedia.
-Ha iste americas que, qui ma tempore capital. \citeauthor{dueck:trio} \citep{dueck:trio}
-
-\begin{aenumerate}
-    \item Enumeration with small caps (alpha)
-    \item Second item
-\end{aenumerate}
-
-\paragraph{A Paragraph Example} Uno de membros summario preparation,
-es inter disuso qualcunque que. Del hodie philologos occidental al,
-como publicate litteratura in web. Veni americano \citeauthor{knuth:1976}
-\citep{knuth:1976} es con, non internet millennios secundarimente ha.
-Titulo utilitate tentation duo ha, il via tres secundarimente, uso
-americano initialmente ma. De duo deler personas initialmente. Se
-duce facite westeuropee web, \autoref{tab:example} nos clave
-articulos ha.
-
-Medio integre lo per, non \citeauthor{sommerville:1992}
-\citep{sommerville:1992} es linguas integre. Al web altere integre
-periodicos, in nos hodie basate. Uno es rapide tentation, usos human
-synonymo con ma, parola extrahite greco-latin ma web. Veni signo
-rapide nos da.
-
-%Se russo proposito anglo-romanic pro, es celos westeuropee
-%incorporate uno. Il web unic periodicos. Que usate scientia ma, sed
-%tres unidirectional al, asia personas duo de. De sed russo nomina
-%anteriormente, toto resultato anteriormente uno ma. Non se signo
-%romanic technologia, un medio millennios con.
-
-%Major facto sia es, con o titulo maximo international. Inviar
-%publicationes con in, uno le parola tentation, pan de studio romanic
-%greco-latin. Tu duo titulo debitas latente, que vista programma ma.
-%Non tote tres germano se, lo parola periodicos non.
-
-\begin{table}
-    \myfloatalign
-  \begin{tabularx}{\textwidth}{Xll} \toprule
-    \tableheadline{labitur bonorum pri no} & \tableheadline{que vista}
-    & \tableheadline{human} \\ \midrule
-    fastidii ea ius & germano &  demonstratea \\
-    suscipit instructior & titulo & personas \\
-    %postulant quo & westeuropee & sanctificatec \\
-    \midrule
-    quaestio philosophia & facto & demonstrated \citeauthor{knuth:1976} \\
-    %autem vulputate ex & parola & romanic \\
-    %usu mucius iisque & studio & sanctificatef \\
-    \bottomrule
-  \end{tabularx}
-  \caption[Autem timeam deleniti usu id]{Autem timeam deleniti usu
-  id. \citeauthor{knuth:1976}}  \label{tab:example}
-\end{table}
-
-\enlargethispage{2cm}
-\subsection{Linguistic Registrate}
-Veni introduction es pro, qui finalmente demonstrate il. E tamben
-anglese programma uno. Sed le debitas demonstrate. Non russo existe o,
-facite linguistic registrate se nos. Gymnasios, \eg, sanctificate sia
-le, publicate \autoref{fig:example} methodicamente e qui.
-
-Lo sed apprende instruite. Que altere responder su, pan ma, \ie, signo
-studio. \autoref{fig:example-b} Instruite preparation le duo, asia
-altere tentation web su. Via unic facto rapide de, iste questiones
-methodicamente o uno, nos al.
+\epigraph{What's the frequency, Kenneth?}{--- \textsc{R.E.M.}}
 

### 2016-07-17 FrontBackmatter/Introduction.tex
Add notation chapter

\chapter*{Introduction}
\addtocontents{toc}{\protect\vspace{\beforebibskip}} % to have the bib a bit from the rest in the toc
\addcontentsline{toc}{section}{\tocEntry{Introduction}}

\epigraph{How does newness come into world? How is it born?

    Of what fusions, translations, conjoinings is it made?

    How does it survive, extreme and dangerous as it is? What compromises, what deals,
    what betrayals of its secret nature must it take to stave off the wrecking crew, the
    exterminating angel, the guillotine?

    Is birth always a fall?

    Do angels have wings? Can men fly?}{
    --- \textsc{Salman Rushdie}, \textit{The Satanic Verses}}

### 2016-07-17 FrontBackmatter/Notation.tex
Add notation chapter

\chapter*{Notations, conventions and units}
\addcontentsline{toc}{section}{\tocEntry{Notations, conventions and units}}

%A number of notations and typographic conventions has been adopted in order to maintain consistency throughout. We summarize them here:
%\begin{table*}[!h]
%\begin{tabular}{c l}
% $\mathcal{A}, \mathcal{D}, \ldots$ & integral operators over a closed subset of $\mathbb{R}^2$ \\
% $H, \varPhi, \ldots$ & first or second quantized $N$-electron operators \\
% $\vect{r}, \vect{r}^\prime, \ldots$ & position vectors in $\mathbb{R}^3$ \\
% $\vect{s}, \vect{s}^\prime, \ldots$ & position vectors in a closed subset of $\mathbb{R}^2$ \\
% $\vect{v}, \vect{q}, \ldots$ & vectors in an arbitrary vector space \\
% $\mat{K}, \mat{F}, \ldots$ & matrices in an arbitrary vector space \\
% $0_N, I_N$ & the zero and the identity in an $N$-dimensional vector space
%\end{tabular}
%\end{table*}
%
%Summation over repeated indices, both discrete and continuous, is always to be understood:
%\begin{alignat*}{2}
% f_r = C_{\kappa r}g_{\kappa} \equiv \sum_{\kappa}C_{\kappa r}g_{\kappa}, \quad&\quad
% f(\vect{r}) = U(\vect{r},\vect{r}^\prime)g(\vect{r}^\prime) \equiv \int\diff\vect{r}^\prime U(\vect{r},\vect{r}^\prime)g(\vect{r}^\prime).
%\end{alignat*}
%
%Further conventions have been adopted for the indices of functions. Lower case Latin letters are used for MO 4-spinors, while lower case
%Greek letters ($\kappa, \lambda, \mu, \ldots$) are reserved for one-electron basis functions in 2-spinor or scalar form.
%Specific ranges of letters are used as follows:
%\begin{table*}[!h]
%\begin{tabular}{c l}
% $r, s, t, \ldots$ & general MO 4-spinor indices \\
% $i, j, k, \ldots$ & occupied MO 4-spinor indices \\
% $a, b, c, \ldots$ & virtual MO 4-spinor indices
%\end{tabular}
%\end{table*}
%
%Upper case Latin letters always refer to functions in vector spaces other than the two listed above. Complex conjugation will always be shown using
%a dagger ($\dagger$) instead of a star ($*$).
%
%SI-based atomic units have been used \cite{Whiffen1978}:
%\begin{equation*}
% \si{\electronmass} = \si{\elementarycharge} = \hslash = \frac{1}{4\pi\diel_0} = 1
%\end{equation*}
%the unit of length is the Bohr \si{\bohr}, while that of energy is the Hartree \si{\hartree}.
%The speed of light is then:
%\begin{equation*}
% c = \SI{137.035999074}{\bohr\hartree\per\planckbar}
%\end{equation*}
%
%
%Some useful conversion constants to SI units are here listed \cite{CODATA2010}:
%\begin{subequations}
% \begin{align*}
%  \SI{1}{\bohr} &= \SI{0.52917721092}{\angstrom} = \SI{5.2917721092e-11}{\meter} \\
%  \SI{1}{\hartree} &= \SI{27.21138505}{\electronvolt} = \SI{2625.4996404}{\kilo\joule\per\mole} \\
%  \SI{1}{\debye} &= \SI{3.3356410E-30}{\coulomb\meter} \\
%  \SI{1}{\angstrom\cubed} &= \SI{1.11264984E-40}{\coulomb\squared\meter\squared\per\joule}
% \end{align*}
%\end{subequations}

### 2016-07-17 Chapters/Ch01-ContinuumSolvationModels.tex
Start on papers summaries and contributions

%************************************************
\chapter{Continuum solvation models}\label{ch:CSM}
%************************************************
%*****************************************
%*****************************************
%*****************************************
%*****************************************
%*****************************************

### 2016-07-17 Chapters/PaperSummary.tex
Start on papers summaries and contributions

%************************************************
\chapter{Summary of the papers}\label{ch:papers-summary}

The codes used in the papers are all interfaced to the \PCMSolver library,
of which I am the main developer and maintainer.

\section{Paper I. Continuum solvation in the relativistic regime}

I contributed the theoretical derivation of the quantum/classical polarizable
terms in a four-component \acs{SCF} framework, for energies and linear response
properties. I devised the coupling of the four-component program \DIRAC with
\PCMSolver by providingi the implementation and testing of:
\begin{itemize}
  \item \acs{MEP} integrals for four-component wave functions,
  \item the additional Fock matrix contributions,
  \item the additional terms in the response equations.
\end{itemize}
I performed all the calculations and large part of the data analysis
for the results reported in the paper.
Finally, I wrote the first draft of the paper and coordinated editing of
all subsequent versions.

\section{Paper II. The wavelet boundary element method for \acs{PCM}}
\section{Paper III: Continuum modelling of non homogeneous environments}

%************************************************
%*****************************************
%*****************************************
%*****************************************
%*****************************************
%*****************************************

### 2016-07-17 Chapters/Ch04-PaperSummary.tex
Added acronyms and some commands

--- 2016-07-17 Chapters/PaperSummary.tex
@@
 
-The codes used in the papers are all interfaced to the \PCMSolver library,
+The codes used in the papers are all interfaced to the \pcmsolver library,
 of which I am the main developer and maintainer.
@@
 properties. I devised the coupling of the four-component program \DIRAC with
-\PCMSolver by providingi the implementation and testing of:
+\pcmsolver by providing the implementation and testing of:
 \begin{itemize}
@@
 
-\section{Paper II. The wavelet boundary element method for \acs{PCM}}
+\section{Paper II. The wavelet boundary element method for PCM}
 \section{Paper III: Continuum modelling of non homogeneous environments}

### 2016-07-17 acronyms.tex
Added acronyms and some commands

\begin{acronym}
\acro{CC}{coupled cluster}
\acro{CCSD}{coupled cluster with single and double substitutions}
\acro{CI}{configuration interaction}
\acro{MO}{molecular orbital}
\acro{LR}{linear response}
\acro{DFT}{density-functional theory}
\acro{PCM}{polarizable continuum model}
\acro{MM}{molecular mechanics}
\acro{MP}{M{\o}ller--Plesset}
\acro{HF}{Hartree--Fock}
\acro{ASC}{apparent surface charge}
\acro{PTE}{perturbation to the energy}
\acro{PTED}{perturbation to the energy and the density}
\acro{QM/MM}{Quantum Mechanics/Molecular Mechanics}
\acro{DC}{Dielectric Continuum}
\acro{GEPOL}{GEPOL}
\acro{IEF}{Integral Equation Formalism}
\acro{LHS}{left-hand side}
\acro{MD}{Molecular Dynamics}
\acro{MM}{Molecular Mechanics}
\acro{BVP}{boundary value problem}
\acro{PDE}{partial differential equation}
\acro{PE}{Polarizable Embedding}
\acro{QC}{Quantum Chemistry}
\acro{RHS}{right-hand side}
\acro{SAS}{Solvent-Accessible Surface}
\acro{SES}{Solvent-Excluded Surface}
\acro{MEP}{molecular electrostatic potential}
\acro{TDSCF}{time-dependent SCF}
\acro{KS}{Kohn--Sham}
\end{acronym}

### 2016-07-17 FrontBackmatter/Publications.tex
Add notation chapter
Change formatting of list of papers

@@
 \item
-  \textbf{R. Di Remigio}, R. Bast, L. Frediani, and T. Saue:
-  "Four-Component Relativistic Calculations in Solution with the
+  \textbf{R. Di Remigio}, R. Bast, L. Frediani, and T. Saue
+  \\
+  Four-Component Relativistic Calculations in Solution with the
   Polarizable Continuum Model of Solvation: Theory,
@@
   \ce{H2X} (\ce{X} = \ce{O}, \ce{S}, \ce{Se}, \ce{Te},
-  \ce{Po})",
+  \ce{Po})
+  \\
 \textit{J. Phys. Chem. A}, \textrm{2015}, \textbf{119}, 5061--5077
@@
   M. Bugeanu, \textbf{R. Di Remigio}, K. Mozgawa, S. S. Reine, H.
-  Harbrecht,  and L. Frediani:
-  "Wavelet Formulation of the Polarizable Continuum Model. II. Use of
-  Piecewise Bilinear Boundary Elements",
+  Harbrecht,  and L. Frediani
+  \\
+  Wavelet Formulation of the Polarizable Continuum Model. II. Use of
+  Piecewise Bilinear Boundary Elements
+  \\
   \textit{Phys. Chem. Chem. Phys.}, \textrm{2015}, \textbf{17},
@@
 
-\item \textbf{R. Di Remigio}, K. Mozgawa, H. Cao, V. Weijo, and L.
-  Frediani:
-  "A Polarizable Continuum Model for Molecules at Spherical
-  Diffuse Interfaces",
+\item
+  \textbf{R. Di Remigio}, K. Mozgawa, H. Cao, V. Weijo, and L.
+  Frediani
+  \\
+  A Polarizable Continuum Model for Molecules at Spherical
+  Diffuse Interfaces
+  \\
   \textit{J. Chem. Phys.}, \textrm{2016}, \textbf{144}, 124103

### 2016-07-27 Chapters/Ch04-PaperSummary.tex
Add commands, fix notations

@@
+% mainfile: ../RobertoDiRemigioPhDThesis.tex
 %************************************************
 \chapter{Summary of the papers}\label{ch:papers-summary}
+
+In the following, I provide a brief summary of the motivations, contents and
+conclusions of the papers this thesis is based on.
+In addition, I give a description of my contributions to the work leading to
+the publication of each of the papers.
+\todo[inline]{Write something along these lines: "My coauthors have all read and approved."}
+
+\section*{Software}
 
@@
 
-\section{Paper I. Continuum solvation in the relativistic regime}
+\section*{Paper I. Continuum solvation in the relativistic regime}
+
+Systems containing heavy elements are notoriously challenging for quantum chemistry.
+In addition to their significantly large sizes, one has to properly include the
+effect of special relativity in the quantum chemical description in order to
+achieve at least qualitative agreement with experiment.
+Additional complications arise if one needs to also include environment effects.
+Continuum models arguably represent a cost-effective strategy to achieve a first
+approximation of these effects.
+In this paper, we presented the first derivation and implementation of the \acs{PCM}
+coupled to a \acs{SCF} description of the solute in the fully relativistic, four-component
+regime.
 
@@
 
-\section{Paper II. The wavelet boundary element method for PCM}
-\section{Paper III: Continuum modelling of non homogeneous environments}
+\section*{Paper II. The wavelet Galerkin boundary element method for PCM}
+
+Continuum solvation models are inherently parametrized. Apart from solvent permittivities,
+the atomic radii and molecular surface definition play a crucial role
+in determining the performance of the models.
+However, the numerical accuracy of the \acs{BEM} procedure used to numerically solve the
+underlying \acs{BIE} is a not-so-often studied aspect of these models.
+Traditionally, collocation methods have been used, but these require parametrization of some of the necessary surface integrals.
+Galerkin methods do not suffer from such a limitation and additionally preserve symmetry of the underlying
+boundary integral operators.
+The use of biorthogonal wavelet bases as finite elements achieves sparsity in
+the \acs{BEM} procedure, due to their intrinsically hierarchical structure and
+the existence of \emph{a priori} and \emph{a posteriori} matrix compression
+estimates.
+Thus, wavelet Galerkin \acs{BEM} represents a valid alternative to traditional collocation methods,
+both to achieve a better computational scaling and to provide accurate, benchmark results.
+
+This work builds upon and extends previous work by~\citeauthor{Weijo2010-hy}
+\citeauthor{Weijo2010-hy} used \ac{PWC} bases to show that superior accuracy
+and convergence could be obtained for the calculation of molecular energies.
+In this work we showed that faster convergence can be achieved when \ac{PWL} bases are used instead.
+Moreover, the same holds for the calculation of static electric properties.
+Notably, the traditional collocation solver cannot guarantee the same accuracy,
+even for very large finite element bases. This suggests that, in some cases,
+\acs{BEM} collocation methodologies might slow down or even prevent the
+convergence of the quantum mechanical response equations solvers.
+
+For this paper, I provided template interface and test sets for the cavity
+generator\autocite{Harbrecht2009-no, Harbrecht2011-dk}
+and wavelet Galerkin \acs{BEM} solver\autocite{Harbrecht2004-uo, Harbrecht2006-ug, Dahmen2006-pj}
+with \pcmsolver.
+These were used to interface with the new C++ implementation of the wavelet
+Galerkin solvers of Monica Bugeanu.
+I implemented the interface between the \LSDALTON quantum chemistry software
+package and the \pcmsolver software library. The interface allows to run \acs{HF} and
+\acs{KS}-\acs{DFT} single-point and linear response calculations.
+I, together with coauthor Krzysztof Mozgawa, performed the benchmark quantum
+chemical calculations presented in the paper.
+Finally, I coordinated the editing of all manuscript drafts. I wrote Sections 2.1 and 2.3,
+while Sections 3 and 4 were co-written with the first author, Monica Bugeanu.
+I performed most of the data analysis and produced tables and graphs.
+
+The interface to \LSDALTON was later released in version ...
+providing \acs{PCM} capabilities to the software package.
+
+\section*{Paper III: Continuum modelling of non homogeneous environments}
 

### 2016-07-27 FrontBackmatter/Abstract.tex
Add commands, fix notations

%*******************************************************
% Abstract
%*******************************************************
%\renewcommand{\abstractname}{Abstract}
\pdfbookmark[1]{Abstract}{Abstract}
\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Abstract}

\vfill

\endgroup

\vfill

### 2016-07-27 FrontBackmatter/Notation.tex
Add commands, fix notations

@@
 
-%A number of notations and typographic conventions has been adopted in order to maintain consistency throughout. We summarize them here:
-%\begin{table*}[!h]
-%\begin{tabular}{c l}
-% $\mathcal{A}, \mathcal{D}, \ldots$ & integral operators over a closed subset of $\mathbb{R}^2$ \\
-% $H, \varPhi, \ldots$ & first or second quantized $N$-electron operators \\
-% $\vect{r}, \vect{r}^\prime, \ldots$ & position vectors in $\mathbb{R}^3$ \\
-% $\vect{s}, \vect{s}^\prime, \ldots$ & position vectors in a closed subset of $\mathbb{R}^2$ \\
-% $\vect{v}, \vect{q}, \ldots$ & vectors in an arbitrary vector space \\
-% $\mat{K}, \mat{F}, \ldots$ & matrices in an arbitrary vector space \\
-% $0_N, I_N$ & the zero and the identity in an $N$-dimensional vector space
-%\end{tabular}
-%\end{table*}
-%
-%Summation over repeated indices, both discrete and continuous, is always to be understood:
-%\begin{alignat*}{2}
-% f_r = C_{\kappa r}g_{\kappa} \equiv \sum_{\kappa}C_{\kappa r}g_{\kappa}, \quad&\quad
-% f(\vect{r}) = U(\vect{r},\vect{r}^\prime)g(\vect{r}^\prime) \equiv \int\diff\vect{r}^\prime U(\vect{r},\vect{r}^\prime)g(\vect{r}^\prime).
-%\end{alignat*}
-%
-%Further conventions have been adopted for the indices of functions. Lower case Latin letters are used for MO 4-spinors, while lower case
-%Greek letters ($\kappa, \lambda, \mu, \ldots$) are reserved for one-electron basis functions in 2-spinor or scalar form.
-%Specific ranges of letters are used as follows:
-%\begin{table*}[!h]
-%\begin{tabular}{c l}
-% $r, s, t, \ldots$ & general MO 4-spinor indices \\
-% $i, j, k, \ldots$ & occupied MO 4-spinor indices \\
-% $a, b, c, \ldots$ & virtual MO 4-spinor indices
-%\end{tabular}
-%\end{table*}
-%
-%Upper case Latin letters always refer to functions in vector spaces other than the two listed above. Complex conjugation will always be shown using
-%a dagger ($\dagger$) instead of a star ($*$).
-%
-%SI-based atomic units have been used \cite{Whiffen1978}:
-%\begin{equation*}
-% \si{\electronmass} = \si{\elementarycharge} = \hslash = \frac{1}{4\pi\diel_0} = 1
-%\end{equation*}
-%the unit of length is the Bohr \si{\bohr}, while that of energy is the Hartree \si{\hartree}.
-%The speed of light is then:
-%\begin{equation*}
-% c = \SI{137.035999074}{\bohr\hartree\per\planckbar}
-%\end{equation*}
-%
-%
-%Some useful conversion constants to SI units are here listed \cite{CODATA2010}:
-%\begin{subequations}
-% \begin{align*}
-%  \SI{1}{\bohr} &= \SI{0.52917721092}{\angstrom} = \SI{5.2917721092e-11}{\meter} \\
-%  \SI{1}{\hartree} &= \SI{27.21138505}{\electronvolt} = \SI{2625.4996404}{\kilo\joule\per\mole} \\
-%  \SI{1}{\debye} &= \SI{3.3356410E-30}{\coulomb\meter} \\
-%  \SI{1}{\angstrom\cubed} &= \SI{1.11264984E-40}{\coulomb\squared\meter\squared\per\joule}
-% \end{align*}
-%\end{subequations}
+A number of notations and typographic conventions has been adopted in
+order to maintain consistency throughout. We summarize them here:
+\begin{table*}[!h]
+\begin{tabular}{c l}
+ $\bi{A}, \bi{D}, \ldots$ & integral operators over a closed subset of $\mathbb{R}^2$ \\
+ $H, \varPhi, \ldots$ & first or second quantized $N$-electron operators \\
+ $\vect{r}, \vect{r}^\prime, \ldots$ & position vectors in $\mathbb{R}^3$ \\
+ $\vect{s}, \vect{s}^\prime, \ldots$ & position vectors in a closed subset of $\mathbb{R}^2$ \\
+ $\vect{v}, \vect{q}, \ldots$ & vectors in an arbitrary vector space \\
+ $\mat{K}, \mat{F}, \ldots$ & matrices in an arbitrary vector space \\
+ $0_N, I_N$ & the zero and the identity in an $N$-dimensional vector space
+\end{tabular}
+\end{table*}
+
+Further conventions have been adopted for the indices of functions.
+Lower case Latin letters are used for MO 4-spinors, while lower case
+Greek letters ($\kappa, \lambda, \mu, \ldots$) are reserved for
+one-electron basis functions in 2-spinor or scalar form.
+Specific ranges of letters are used as follows:
+\begin{table*}[!h]
+\begin{tabular}{c l}
+ $r, s, t, \ldots$ & general MO 4-spinor indices \\
+ $i, j, k, \ldots$ & occupied MO 4-spinor indices \\
+ $a, b, c, \ldots$ & virtual MO 4-spinor indices
+\end{tabular}
+\end{table*}
+
+Upper case Latin letters always refer to functions in vector spaces
+other than the two listed above. Complex conjugation will always be
+shown using a dagger ($\dagger$) instead of a star ($*$).
+
+Hartree atomic units have been used:~\autocite{Whiffen1978}
+\begin{equation*}
+ \si{\electronmass} = \si{\elementarycharge} = \hslash = 4\pi\diel_0 = 1
+\end{equation*}
+the unit of length is the Bohr \si{\bohr}, while that of energy is the Hartree \si{\hartree}.
+The speed of light is then:
+\begin{equation*}
+ c = \SI{137.035999074}{\bohr\hartree\per\planckbar}
+\end{equation*}
+
+Some useful conversion constants to SI units are here
+listed:~\autocite{Mohr2012-zc}
+\begin{subequations}
+ \begin{align*}
+  \SI{1}{\bohr} &= \SI{0.52917721092}{\angstrom} = \SI{5.2917721092e-11}{\meter} \\
+  \SI{1}{\hartree} &= \SI{27.21138505}{\electronvolt} = \SI{2625.4996404}{\kilo\joule\per\mole} \\
+  \SI{1}{\debye} &= \SI{3.3356410E-30}{\coulomb\meter} \\
+  \SI{1}{\angstrom\cubed} &= \SI{1.11264984E-40}{\coulomb\squared\meter\squared\per\joule}
+ \end{align*}
+\end{subequations}

### 2016-07-27 Chapters/Ch01-ContinuumSolvationModels.tex
Some more work

@@
-%************************************************
+% mainfile: ../RobertoDiRemigioPhDThesis.tex
 \chapter{Continuum solvation models}\label{ch:CSM}
-%************************************************
-%*****************************************
-%*****************************************
-%*****************************************
-%*****************************************
-%*****************************************
+
+\section{A unifying view of classical polarizable models}
+
+Variational functional for implicit models, \acs{PCM} in this case:
+\begin{equation}
+ U_\mathrm{PCM} = \frac{1}{2}\scalprod{\sigma}{\PCM\sigma} + \scalprod{\esp}{\sigma}
+\end{equation}
+
+Variational functional for explicit models, could be MMpol or PE or FQ:
+\begin{equation}
+  U_\mathrm{MM} = \frac{1}{2}\kappa\MM\kappa + \kappa\zeta
+\end{equation}
+
+The PCM equations will be written in the ``complete basis'' meaning that
+we will introduce the usual boundary-element method (BEM) discretization
+at the very end of the derivation. In other words, we will be working
+with the exact integral equation and not with its discretized
+counterpart. As a consequence, the apparent surface charge $\sigma$ and
+the electrostatic potential $\esp$ will have a \emph{continuous}
+dependence on a ``cavity surface'' index $\vect{s}$. Whenever a
+charge-potential product is present it is then to be interpreted as the
+\emph{surface integral}, i.e. the scalar product in the suitable,
+infinite-dimensional vector space on the cavity boundary $\Gamma$. The
+following shorthand notation will be adopted:
+\begin{equation}
+ \sigma\esp = \scalprod{\sigma}{\esp}
+\end{equation}
+
+Coupled implicit/explicit polarization energy functional:
+\begin{equation}
+  U_\mathrm{pol} =
+   \frac{1}{2}\sigma\PCM\sigma + \sigma\esp
+ + \frac{1}{2}\kappa\MM\kappa + \kappa\zeta
+ + \sigma\bi{X}\kappa
+\end{equation}
+where one has:
+\begin{equation}
+  \sigma\MMPCM\kappa = \kappa\MMPCM^\dagger\sigma
+\end{equation}
+and $\MMPCM$ is the implicit/explicit interaction kernel. This is
+charge/dipole or charge/charge interaction kernel for the MMpol and PE models
+or the FQ model, respectively.
+The global minimum of the convex functional is found by differentiating
+it with respect to the variational degrees of freedom, \ie{} the
+\acl{ASC} (\acs{ASC}) density $\sigma$ and the classical point
+charges/dipoles $\kappa$. This leads to the coupled equations:
+\begin{align}
+  \PCM\sigma + \esp + \MMPCM\kappa &= 0 \\
+  \MM\kappa  + \zeta + \MMPCM^\dagger\sigma &= 0
+\end{align}
+or in the commonly found matrix form:
+\begin{equation}
+  \begin{pmatrix}
+    \PCM & \MMPCM^\dagger \\
+    \MMPCM & \MM
+  \end{pmatrix}
+  \begin{pmatrix}
+   \sigma \\
+   \kappa
+  \end{pmatrix}
+  =
+  -
+  \begin{pmatrix}
+   \esp \\
+   \zeta
+  \end{pmatrix}
+\end{equation}
+Finally, let us re-express the equations above in a "supermatrix"
+formalism:
+\begin{equation}
+  U_\mathrm{pol} =
+  \frac{1}{2}\p^\dagger\V\p + \p^\dagger\s
+\end{equation}
+where:
+\begin{alignat}{3}
+  \p =
+  \begin{pmatrix}
+    \sigma \\
+    \kappa
+  \end{pmatrix},
+  \quad&
+  \s =
+  \begin{pmatrix}
+   \esp \\
+   \zeta
+  \end{pmatrix},
+  \quad&
+  \V =
+  \begin{pmatrix}
+    \PCM & \MMPCM^\dagger \\
+    \MMPCM & \MM
+  \end{pmatrix}
+\end{alignat}
+The supermatrix form will prove particularly useful in the following
+derivations.

### 2016-07-27 Chapters/Ch03-PCMCoupledCluster.tex
Start on papers summaries and contributions
Some more work

--- 2016-07-17 Chapters/Ch01-ContinuumSolvationModels.tex
@@
-%************************************************
-\chapter{Continuum solvation models}\label{ch:CSM}
-%************************************************
-%*****************************************
-%*****************************************
-%*****************************************
-%*****************************************
-%*****************************************
+% mainfile: ../RobertoDiRemigioPhDThesis.tex
+\chapter{Coupled cluster}\label{ch:PCMCC}
+
+\section{Notation}
+Cluster operator:
+\begin{equation}
+  T = \sum_{i=1}^\mathcal{M}\sum_{\mu_i}\tamp{\mu_i}\cluster{\mu_i}
+\end{equation}
+$\mu_i$ is the $\mu$-th excitation at the $i$-th
+excitation level and $\mathcal{M}$ is the truncation level.
+
+The term at the $j$-th excitation level in the left CC wavefunction:
+\begin{equation}
+  \bra{\tbar{j}} = \sum_{\mu_j}\tbar{\mu_j}\bra{\mu_j}
+\end{equation}
+
+\acl{MP} partitioning:
+\begin{equation}
+ H_0 = F + \Phi = \sum_P\epsilon_P\cons{P}\anni{P} + (g- V_\mathrm{HF})
+\end{equation}
+where $F$ if the Fock matrix and $\Phi$ is the fluctuation potential.
+We assume to deal with an \acl{HF} (\acs{HF}) reference function.
+
+Similarity transformed operator:
+\begin{equation}
+  \overline{O} =\expo{-T}O\expo{T}
+\end{equation}
+Since the Fock operator is diagonal, its similarity transformation
+truncates after the second term:
+\begin{equation}
+  \overline{F} = F + \sum_{i}^4\sum_{\mu_i}\tamp{\mu_i}\cluster{\mu_i}\denom{\mu_i}
+\end{equation}
+where $\denom{\mu_i}$ is the difference in orbital energy between the
+virtual and occupied spin-orbitals of excitation $i$.
+For the fluctuation potential the similarity transformation truncates
+after the fifth term, \ie{} includes up to four-fold nested commutators
+with the cluster operator.
+
+\acs{CC} expectation values are formed using the left and
+right \acs{CC} states and are denoted as:
+\begin{equation}
+  O(\tamp{}, \tbar{}) = \braket{\mathrm{HF} | \overline{O} | \mathrm{HF} }
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{j} | \overline{O} | \mathrm{HF} }
+\end{equation}
+We will also use the notation $O^\mathcal{M}(\tamp{}, \tbar{})$ to
+denote a \acs{CC} expectation value truncated at the $\mathcal{M}$-th
+excitation level.
+
+\section{Effective Coupled Cluster Lagrangian}
+
+The effective CC Lagrangian can be readily obtained by coupling the
+usual CC Lagrangian with the classical polarization functional.
+Care must however be taken in the definition of the classical
+variational degrees of freedom as to not erroneously include
+contributions from the reference state in the subsequent correlated
+treatment.
+
+Let us assume that the reference state is the optimized single \acs{HF}
+determinant in the presence of the implicit/explicit environment.
+Then the corresponding variational degrees of freedom satisfy:
+\begin{equation}
+  \begin{pmatrix}
+    \PCM & \MMPCM^\dagger \\
+    \MMPCM & \MM
+  \end{pmatrix}
+  \begin{pmatrix}
+   \sigma^\mathrm{HF} \\
+   \kappa^\mathrm{HF}
+  \end{pmatrix}
+  =
+  -
+  \begin{pmatrix}
+   \esp^\mathrm{HF} \\
+   \zeta^\mathrm{HF}
+  \end{pmatrix}
+\end{equation}
+and correspond to the \emph{reference reaction field}.
+In the subsequent correlated treatment one can separate the reaction
+field into reference and correlated components:
+\begin{equation}
+  \begin{pmatrix}
+   \sigma \\
+   \kappa
+  \end{pmatrix}
+  =
+  \begin{pmatrix}
+   \sigma^\mathrm{HF} \\
+   \kappa^\mathrm{HF}
+  \end{pmatrix}
+  +
+  \begin{pmatrix}
+   \sigma^\mathrm{CC} \\
+   \kappa^\mathrm{CC}
+  \end{pmatrix}
+\end{equation}
+and similarly for the sources:
+\begin{equation}
+  \begin{pmatrix}
+   \esp(\tamp{}, \tbar{}) \\
+   \zeta(\tamp{}, \tbar{})
+  \end{pmatrix}
+  =
+  \begin{pmatrix}
+   \esp^\mathrm{HF} \\
+   \zeta^\mathrm{HF}
+  \end{pmatrix}
+  +
+  \begin{pmatrix}
+    \tilde{\esp}(\tamp{}, \tbar{}) \\
+   \tilde{\zeta}(\tamp{}, \tbar{})
+  \end{pmatrix}
+  =
+  \begin{pmatrix}
+  \braket{\mathrm{HF} | \esp  | \mathrm{HF}} \\
+  \braket{\mathrm{HF} | \zeta | \mathrm{HF}}
+  \end{pmatrix}
+  +
+  \begin{pmatrix}
+   \braket{\mathrm{HF} | \tilde{\esp} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} | \mathrm{HF}} \\
+   \braket{\mathrm{HF} | \tilde{\zeta} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} | \mathrm{HF}}
+  \end{pmatrix}
+\end{equation}
+The Baker--Campbell--Hausdorff expansions of the source terms have been
+rewritten as:
+\begin{equation}
+  \overline{\s} = \s + \tilde{\s} = \s + \commutator{\s}{T}
+  + \frac{1}{2}\commutator{\commutator{\s}{T}}{T}
+\end{equation}
+taking advantage of the fact that $\s$ is a one-electron operator and
+hence its commutator expansion truncates at the third term.
+We apply a shift to the correlation part of the source term
+one-electron operator to remove the reference source terms:
+\begin{equation}
+  \begin{aligned}
+  \begin{pmatrix}
+   \braket{\mathrm{HF} | \tilde{\esp} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \tilde{\esp} | \mathrm{HF}} \\
+   \braket{\mathrm{HF} | \tilde{\zeta} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \tilde{\zeta} | \mathrm{HF}}
+  \end{pmatrix}
+  &=
+  \begin{pmatrix}
+    \braket{\mathrm{HF} | \overline{\esp} - \esp^\mathrm{HF} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} - \esp^\mathrm{HF} | \mathrm{HF}} \\
+  \braket{\mathrm{HF} | \overline{\zeta} - \zeta^\mathrm{HF} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} -
+  \zeta^\mathrm{HF}| \mathrm{HF}}
+  \end{pmatrix}
+  \\
+  &=
+  \begin{pmatrix}
+   \braket{\mathrm{HF} | \tilde{\esp} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} | \mathrm{HF}} \\
+   \braket{\mathrm{HF} | \tilde{\zeta} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} | \mathrm{HF}}
+  \end{pmatrix}
+  =
+  \begin{pmatrix}
+    \tilde{\esp}(\tamp{}, \tbar{}) \\
+    \tilde{\zeta}(\tamp{}, \tbar{})
+  \end{pmatrix}
+\end{aligned}
+\end{equation}
+which is equivalent to imposing normal ordering of the
+operators.~\cite{Shavitt}
+The total functional will include both reference and correlation
+components:
+\begin{equation}
+  \begin{aligned}
+  U_\mathrm{pol} &=
+  \frac{1}{2}(\sigma^\mathrm{HF}+\sigma^\mathrm{CC})\PCM(\sigma^\mathrm{HF}+\sigma^\mathrm{CC})
++ (\sigma^\mathrm{HF}+\sigma^\mathrm{CC})(\esp^\mathrm{HF}+\tilde{\esp}(\tamp{},\tbar{})) \\
+&+ \frac{1}{2}(\kappa^\mathrm{HF}+\kappa^\mathrm{CC})\MM(\kappa^\mathrm{HF}+\kappa^\mathrm{CC})
++ (\kappa^\mathrm{HF}+\kappa^\mathrm{CC})(\zeta^\mathrm{HF}+\tilde{\zeta}(\tamp{},\tbar{})) \\
+&+ (\sigma^\mathrm{HF}+\sigma^\mathrm{CC})\bi{X}(\kappa^\mathrm{HF}+\kappa^\mathrm{CC}) \\
+&=
+    \frac{1}{2}\sigma^\mathrm{HF}\PCM\sigma^\mathrm{HF}
+  + \sigma^\mathrm{HF}\esp^\mathrm{HF}
+  + \frac{1}{2}\kappa^\mathrm{HF}\MM\kappa^\mathrm{HF}
+  + \kappa^\mathrm{HF}\zeta^\mathrm{HF}
+  + \sigma^\mathrm{HF}\bi{X}\kappa^\mathrm{HF} \\
+&+  \frac{1}{2}\sigma^\mathrm{CC}\PCM\sigma^\mathrm{CC}
+  + \sigma^\mathrm{CC}\tilde{\esp}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa^\mathrm{CC}\MM\kappa^\mathrm{CC}
+  + \kappa^\mathrm{CC}\tilde{\zeta}(\tamp{},\tbar{})
+  + \sigma^\mathrm{CC}\bi{X}\kappa^\mathrm{CC} \\
+&+  \frac{1}{2}\sigma^\mathrm{HF}\PCM\sigma^\mathrm{CC}
+  + \sigma^\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa^\mathrm{HF}\MM\kappa^\mathrm{CC}
+  + \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
+  + \sigma^\mathrm{HF}\bi{X}\kappa^\mathrm{CC} \\
+&+  \frac{1}{2}\sigma^\mathrm{CC}\PCM\sigma^\mathrm{HF}
+  + \sigma^\mathrm{CC}\esp^\mathrm{HF}
+  + \frac{1}{2}\kappa^\mathrm{CC}\MM\kappa^\mathrm{HF}
+  + \kappa^\mathrm{CC}\zeta^\mathrm{HF}
+  + \sigma^\mathrm{CC}\bi{X}\kappa^\mathrm{HF} \\
+&=
+    \frac{1}{2}\sigma^\mathrm{HF}\esp^\mathrm{HF}
+  + \frac{1}{2}\kappa^\mathrm{HF}\zeta^\mathrm{HF} \\
+&+  \frac{1}{2}\sigma^\mathrm{CC}\PCM\sigma^\mathrm{CC}
+  + \sigma^\mathrm{CC}\tilde{\esp}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa^\mathrm{CC}\MM\kappa^\mathrm{CC}
+  + \kappa^\mathrm{CC}\tilde{\zeta}(\tamp{},\tbar{})
+  + \sigma^\mathrm{CC}\bi{X}\kappa^\mathrm{CC} \\
+&+ \sigma^\mathrm{CC}(\bi{Y}\sigma^\mathrm{HF} + \bi{X}\kappa^\mathrm{HF} + \esp^\mathrm{HF}) \\
+&+ \kappa^\mathrm{CC}(\bi{W}\kappa^\mathrm{HF} + \bi{X}^\dagger\sigma^\mathrm{HF} + \zeta^\mathrm{HF}) \\
+&+ \sigma^\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{}) + \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{}) \\
+&=
+    \frac{1}{2}\sigma^\mathrm{HF}\esp^\mathrm{HF}
+  + \frac{1}{2}\kappa^\mathrm{HF}\zeta^\mathrm{HF} \\
+&+  \frac{1}{2}\sigma^\mathrm{CC}\PCM\sigma^\mathrm{CC}
+  + \sigma^\mathrm{CC}\tilde{\esp}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa^\mathrm{CC}\MM\kappa^\mathrm{CC}
+  + \kappa^\mathrm{CC}\tilde{\zeta}(\tamp{},\tbar{})
+  + \sigma^\mathrm{CC}\bi{X}\kappa^\mathrm{CC} \\
+&+ \sigma^\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{}) + \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
+  \end{aligned}
+\end{equation}
+where we made use of the stationarity condition for the reference
+reaction field to eliminate a number of terms. Notice that the
+reference reaction energy still appears among the terms in the
+functional. However, these terms will not enter in the optimization of
+the \acs{CC} wave function and we will therefore drop them in the
+following, out of notational convenience:
+\begin{equation}
+  \begin{aligned}
+  U_\mathrm{pol} &=
+    \frac{1}{2}\sigma\PCM\sigma
+    + \sigma\tilde{\esp}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa\MM\kappa
+  + \kappa\tilde{\zeta}(\tamp{},\tbar{})
+  + \sigma\bi{X}\kappa
+  + \sigma^\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{}) +
+  \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{}) \\
+  &= \frac{1}{2}\p^\dagger\V\p + \p^\dagger\tilde{\s}(\tamp{},\tbar{})
+  + \p^\mathrm{HF}\tilde{\s}(\tamp{},\tbar{})
+\end{aligned}
+\end{equation}
+
+One can then write the effective Lagrangian as:
+\begin{equation}
+  \begin{aligned}
+  \lag{M}(\tamp{}, \tbar{}, \sigma) &=
+  \braket{\mathrm{HF} | \overline{H_0} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tamp{i} | \overline{H_0} | \mathrm{HF}} \\
+  &+ \frac{1}{2}\sigma\PCM\sigma
+  + \sigma\tilde{\esp}(\tamp{}, \tbar{})
+  + \frac{1}{2}\kappa\MM\kappa
+  + \kappa\tilde{\zeta}(\tamp{}, \tbar{})
+  + \sigma\bi{X}\kappa
+  + \sigma^\mathrm{HF}\tilde{\esp}(\tamp{}, \tbar{})
+  + \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{}, \tbar{})
+  \end{aligned}
+\end{equation}
+In accordance with previous work~\cite{Cammi2009-gu, Caricato2011-tx}, the
+Hamiltonian $H_0$ is augmented with the terms
+$\sigma^\mathrm{HF}\tilde{\esp}(\tamp{}, \tbar{})$ and
+$\kappa^\mathrm{HF}\tilde{\zeta}(\tamp{}, \tbar{})$
+part appearing in the previous functional yielding the \acl{PTE} (\acs{PTE}) Hamiltonian --
+$H$ -- and
+eventually the \acl{PTED} (\acs{PTED}) Lagrangian:
+\begin{equation}\label{eq:pted-cc}
+  \begin{aligned}
+  \lag{M}(\tamp{}, \tbar{}, \sigma) &=
+  \braket{\mathrm{HF} | \overline{H} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}} \\
+  &+ \frac{1}{2}\sigma\PCM\sigma
+  + \sigma\tilde{\esp}(\tamp{}, \tbar{})
+  + \frac{1}{2}\kappa\MM\kappa
+  + \kappa\tilde{\zeta}(\tamp{}, \tbar{})
+  + \sigma\bi{X}\kappa \\
+  &=
+  \braket{\mathrm{HF} | \overline{H} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}}
+  + \frac{1}{2}\p^\dagger\V\p + \p^\dagger\tilde{\s}(\tamp{},\tbar{})
+  \end{aligned}
+\end{equation}
+As noted by \citeauthor{Cammi2009-gu} for the implicit \acs{PCM} model,
+the \acs{PTE} Hamiltonian is computed in \acs{MP}-partitioned form by
+employing the "solvated" orbitals and Fock matrix:
+\begin{equation}
+  H = F + \Phi.
+\end{equation}
+$F$ is a diagonal one-electron operator and $\Phi$ is the
+fluctuation potential.
+The \acs{PTE} effective Lagrangian ${}^\mathrm{PTE}\lag{M}$ can be
+recovered by dropping the classical terms in the Lagrangian
+\eqref{eq:pted-cc}:
+\begin{equation}
+{}^\mathrm{PTE}\lag{M}
+  =
+  \braket{\mathrm{HF} | \overline{H} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}}
+\end{equation}
+in accordance with the derivation of \citeauthor{Cammi2009-gu}.
+As shown by \citeauthor{Caricato2011-tx}, it is possible to achieve less
+drastic approximations of the \acs{PTED} Lagrangian. Such approximations
+can be easily derived within the current variational framework. We
+postpone this analysis to a later stage of development, \emph{vide
+infra}.
+
+\section{Source terms in the classical energy functional}
+
+The variational formulation offers a series of advantages.
+The classical energy functional:
+\begin{equation}
+  U_\mathrm{pol} =
+   \frac{1}{2}\sigma\PCM\sigma + \sigma\esp
+ + \frac{1}{2}\kappa\MM\kappa + \kappa\zeta
+ + \sigma\bi{X}\kappa
+\end{equation}
+renders itself to a straightforward physical interpretation. In the
+three bilinear terms, \ie{} the ones mediated by an interaction
+operator, we can identify the self-interaction of the classical
+polarization with itself, be it purely implicit:
+$\frac{1}{2}\sigma\PCM\sigma$, purely explicit
+$\frac{1}{2}\kappa\MM\kappa$ or mixed $\sigma\MMPCM\kappa$.
+These terms are positive definite and give rise to an unfavorable
+contribution, which is counterbalanced by the linear terms. These terms
+mediate the interaction between the induced polarization and the
+inducing fields: $\esp$ and $\zeta$. While $\esp$ is quite clearly the
+\acl{MEP} (\acs{MEP}), $\zeta$ can either be the molecular electric
+field (MMpol and PE models) or again the \acs{MEP} (FQ model).
+In any case, both will be determined by the quantum mechanical molecular
+charge density and can thus be formulated as expectation values of
+one-electron operators. Eventually, this achieves the coupling between
+the classical -- $\sigma$ and $\kappa$ -- and
+the quantum mechanical variational degrees of freedom.
+
+The \acs{MEP} operator is given as:
+\begin{equation}
+  \esp = \sum_{PQ}\esp_{PQ}\cons{P}\anni{Q}
+\end{equation}
+and similarly for the \acs{MM} source operator:
+\begin{equation}
+  \zeta = \sum_{PQ}\zeta_{PQ}\cons{P}\anni{Q}
+\end{equation}
+both are nondiagonal in the MO basis. However,
+by virtue of the \emph{cluster commutation condition}, their similarity
+transformations will truncate after two
+terms \cite{Helgaker2000-tz}:
+\begin{equation}
+  \overline{\esp} = \esp + \tilde{\esp}  = \esp + \commutator{\esp}{T} +
+  \frac{1}{2}\commutator{\commutator{\esp}{T}}{T}
+\end{equation}
+and
+\begin{equation}
+  \overline{\zeta} = \zeta + \tilde{\zeta}  = \zeta + \commutator{\zeta}{T} +
+  \frac{1}{2}\commutator{\commutator{\zeta}{T}}{T}
+\end{equation}
+
+In the supermatrix formalism, the similarity-transformed terms will be
+written as:
+\begin{equation}
+  \overline{\s} = \s + \tilde{\s} = \s +
+  \commutator{\s}{T} +
+  \frac{1}{2}\commutator{\commutator{\s}{T}}{T}
+\end{equation}
+It is important to single out which terms contribute in the various
+expectation values where the similarity-transformed sources will be
+involved.
+We develop this analysis in terms of excitation ranks of the
+operators involved, based on the following:
+\begin{equation}
+  \braket{\mu_n |
+  \commutator{\commutator{\commutator{O}{T_{n_1}}}{\ldots}}{T_{n_k}} | \mathrm{HF}} \neq 0
+  \Longleftrightarrow
+  n - m_O + k \leq \sum_{i=1}^k n_i \leq n + m_O
+\end{equation}
+where $n$ is the excitation rank of the $\ket{\mu_n}$ determinant, $k$
+is the total number of cluster operators appearing in the nested
+commutator, the $n_i$ are their respective excitations ranks and $m_O$
+is the \emph{particle rank} of the operator $O$ \cite{Helgaker2000-tz}.
+This follows from the realization that:
+\begin{equation}
+  \commutator{\commutator{\commutator{O}{T_{n_1}}}{\ldots}}{T_{n_k}} \ket{\mathrm{HF}}
+\end{equation}
+is a linear combination of determinants with excitation ranks $v$ in the
+range:
+\begin{equation}
+  \sum_{i=1}^k n_i - m_O \leq v \leq \sum_{i=1}^k n_i + m_O -k
+\end{equation}
+The source operator is a nondiagonal, one-electron
+operator and hence has particle rank $m_\s = 1$.
+
+Eventually, one can cast the source terms entering the Lagrangian as
+follows:
+\begin{equation}
+  \tilde{\s}(\tamp{}, \tbar{}) =
+  \braket{\mathrm{HF} | \BCHfirst{\s}{T} + \BCHsecond{\s}{T} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \s + \BCHfirst{\s}{T} + \BCHsecond{\esp}{T} | \mathrm{HF}}
+\end{equation}
+
+\begin{align}
+  \braket{\mathrm{HF} |
+  \BCHfirst{\s}{T_t}
+  | \mathrm{HF}}
+  &\quad\quad\text{nonzero contributions} \quad 0 \leq t \leq 1
+  \Longleftrightarrow t = 1 \\
+  \braket{\mathrm{HF} |
+  \commutator{\commutator{\s}{T_t}}{T_u}
+  | \mathrm{HF}}
+  &\quad\quad\text{nonzero contributions} \quad t + u = 1 \\
+  \braket{\tbar{i} |
+  \esp
+  | \mathrm{HF} }
+  &\quad\quad\text{nonzero contributions} \quad i = 1 \\
+  \braket{\tbar{i} |
+  \BCHfirst{\s}{T_t}
+  | \mathrm{HF} }
+  &\quad\quad\text{nonzero contributions} \quad i \leq t \leq i + 1 \\
+  \braket{\tbar{i} |
+  \commutator{\commutator{\s}{T_t}}{T_u}
+  | \mathrm{HF} }
+  &\quad\quad\text{nonzero contributions} \quad t + u = i + 1
+\end{align}
+
+For a general truncation level we accordingly obtain:
+\begin{equation}\label{eq:source-term}
+  \tilde{\s}(\tamp{}, \tbar{}) =
+  \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
+  + \braket{\tbar{1} | \s | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \BCHfirst{\s}{T} + \BCHsecond{\s}{T} | \mathrm{HF}}
+\end{equation}
+which shows the unique role the singles amplitudes assume when a
+semiclassical environment model is introduced, \emph{vide infra}.
+Furthermore, for the \acs{CCS} model ($\mathcal{M} = 1$):
+\begin{equation}
+  \tilde{\s}^1(\tamp{}, \tbar{}) =
+  \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
+  + \braket{\tbar{1} | \s | \mathrm{HF}}
+  + \braket{\tbar{1} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
+  + \frac{1}{2}\braket{\tbar{1} | \commutator{\commutator{\s}{T_1}}{T_1} | \mathrm{HF}}
+\end{equation}
+while for the \acs{CCSD} model ($\mathcal{M} = 2$):
+\begin{equation}
+  \tilde{\s}^2(\tamp{}, \tbar{}) =
+  \tilde{\s}^1(\tamp{}, \tbar{})
+  + \braket{\tbar{1} | \BCHfirst{\s}{T_2} | \mathrm{HF}}
+  + \braket{\tbar{2} | \BCHfirst{\s}{T_2} | \mathrm{HF}}
+  + \braket{\tbar{2} | \commutator{\commutator{\s}{T_1}}{T_2} | \mathrm{HF}}
+\end{equation}
+Eventually, within the \acs{CCSDT} model ($\mathcal{M} = 3$) four more
+terms are added:
+\begin{equation}
+  \begin{aligned}
+  \tilde{\s}^3(\tamp{}, \tbar{}) &=
+  \tilde{\s}^2(\tamp{}, \tbar{})
+  + \braket{\tbar{2} | \BCHfirst{\s}{T_3} | \mathrm{HF}} \\
+  &+ \braket{\tbar{3} | \BCHfirst{\s}{T_3} | \mathrm{HF}}
+  + \frac{1}{2}
+  \braket{\tbar{3} | \commutator{\commutator{\s}{T_2}}{T_2} |
+  \mathrm{HF}}
+  + \braket{\tbar{3} | \commutator{\commutator{\s}{T_1}}{T_3} |
+  \mathrm{HF}}
+  \end{aligned}
+\end{equation}
+We note that:
+\begin{itemize}
+    \item double commutators involving the highest order cluster
+      operators can only lead to nonzero terms when $T_1$ is also
+      involved and by projection onto the highest order left \acs{CC}
+      state, \emph{e.g.}
+      $\braket{\tbar{2} | \commutator{\commutator{\s}{T_1}}{T_2} | \mathrm{HF}}$
+      and
+      $\braket{\tbar{3} | \commutator{\commutator{\s}{T_1}}{T_3} | \mathrm{HF}}$
+\end{itemize}
+
+\section{The PTED-CC equations}\label{sec:pted-cc-equations}
+
+Differentiation of the effective Lagrangian in Eq.~\eqref{eq:pted-cc}
+with respect to the variational parameters $\tamp{\mu_i}$,
+$\tbar{\mu_i}$, $\sigma$ and $\kappa$ yields the \acs{PTED}-\acs{CC}
+equations:
+\begin{equation}
+  \begin{aligned}
+   \pderiv{\lag{M}}{\tbar{\mu_i}} &=
+   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
+    + \sigma\braket{\mu_i | \overline{\esp} | \mathrm{HF}}
+    + \kappa\braket{\mu_i | \overline{\zeta} | \mathrm{HF}}
+    = \tampEq{\mu_i}(\tamp{}, \tbar{}, \sigma, \kappa) \\
+    \pderiv{\lag{M}}{\tamp{\mu_i}} &=
+    \denom{\mu_i}\tbar{\mu_i} +
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
+    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} \\
+    &+ \sigma\lbrace
+    \braket{\mathrm{HF} | \commutator{\esp}{\cluster{\mu_i}} +
+    \commutator{\commutator{\esp}{T}}{\cluster{\mu_i}}| \mathrm{HF}} +
+    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\overline{\esp}}{\cluster{\mu_i}} | \mathrm{HF}}
+             \rbrace
+             \\
+    &+ \kappa\lbrace
+    \braket{\mathrm{HF} | \commutator{\zeta}{\cluster{\mu_i}} +
+    \commutator{\commutator{\zeta}{T}}{\cluster{\mu_i}}| \mathrm{HF}} +
+    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\overline{\zeta}}{\cluster{\mu_i}} | \mathrm{HF}}
+             \rbrace
+    = \tbarEq{\mu_i}(\tamp{}, \tbar{}, \sigma, \kappa) \\
+    \pderiv{\lag{M}}{\sigma} &= \PCM\sigma + \MMPCM\kappa
+    + \tilde{\esp}(\tamp{},\tbar{}) =
+    \Omega_\sigma(\tamp{}, \tbar{}, \sigma, \kappa) \\
+    \pderiv{\lag{M}}{\kappa} &= \MM\kappa + \MMPCM^\dagger\sigma + \tilde{\zeta}(\tamp{},\tbar{}) =
+    \Omega_\kappa(\tamp{}, \tbar{}, \sigma, \kappa)
+  \end{aligned}
+\end{equation}
+Given the evident similarities between the implicit and explicit terms
+in the amplitudes and multipliers equations, we adopt the supermatrix
+formalism for the classical polarizable terms. The equations are thus
+recast as:
+\begin{equation}
+  \begin{aligned}
+   \pderiv{\lag{M}}{\tbar{\mu_i}} &=
+   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
+   + \p^\dagger\braket{\mu_i | \overline{\s} | \mathrm{HF}}
+    = \tampEq{\mu_i}(\tamp{}, \tbar{}, \p) \\
+    \pderiv{\lag{M}}{\tamp{\mu_i}} &=
+    \denom{\mu_i}\tbar{\mu_i} +
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
+    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} \\
+    &+ \p^\dagger\lbrace
+    \braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_i}} +
+    \commutator{\commutator{\s}{T}}{\cluster{\mu_i}}| \mathrm{HF}} +
+    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\overline{\s}}{\cluster{\mu_i}} | \mathrm{HF}}
+             \rbrace
+    = \tbarEq{\mu_i}(\tamp{}, \tbar{}, \p) \\
+    \pderiv{\lag{M}}{{\p^{\dagger}}} &=
+    \V\p + \tilde{\s}(\tamp{}, \tbar{}) =
+    \Omega_\p(\tamp{}, \tbar{}, \p)
+  \end{aligned}
+\end{equation}
+
+\section{Approximate Effective Lagrangians}\label{sec:approximate-lagrangians}
+
+We have already noted that the leading terms in the correlated source operator expectation
+value, Eq.~\eqref{eq:source-term}, prominently involve the singles
+cluster operator and the singles left \acs{CC} state.
+This simple fact led \citeauthor{Caricato2011-tx} to propose two
+intermediate approximations in between the \acs{PTE} and \acs{PTED}
+schemes: the \acs{PTES} and \acs{PTE(S)} schemes.
+The \acs{PTE} scheme is advantageous since no coupling between the
+\acs{CC} amplitude and multiplier equations is introduced. However, the
+polarization included in the correlation treatment is the same as for
+the reference determinant, an approximation which appears questionable
+from the point of view of classical electrostatics. On the contrary, the
+\acs{PTED} scheme describes the fully equilibrated electrostatic
+picture, at the expense of coupling the amplitude and multiplier
+equations.
+However, if we truncate the correlated source operator expectation value
+to the first term in Eq.~\eqref{eq:source-term}, the coupling between
+the \acs{CC} right and left states is broken, while still improving upon the
+reference reaction field with correlated contributions.
+This is the essence of the \acs{PTE(S)} approximation. The effective
+Lagrangian would then read as:
+\begin{equation}\label{eq:ptes-cc}
+  ^{\mathrm{PTE(S)}}\lag{M}(\tamp{}, \tbar{}, \p) =
+  \braket{\mathrm{HF} | \overline{H} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}}
+  + \frac{1}{2}\p^\dagger\V\p + \p^\dagger
+  \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
+\end{equation}
+which uncouples all equations:
+\begin{equation}
+  \begin{aligned}
+   \pderiv{\lag{M}}{\tbar{\mu_i}} &=
+   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
+   =
+   ^{\mathrm{PTE(S)}}\tampEq{\mu_i}(\tamp{}, \tbar{}, \p) \\
+    \pderiv{\lag{M}}{\tamp{\mu_1}} &=
+    \denom{\mu_1}\tbar{\mu_1} +
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}} +
+    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} |
+    \commutator{\overline{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}}
+    + \p^\dagger
+    \braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_1}} | \mathrm{HF}}
+    =
+    ^{\mathrm{PTE(S)}}\tbarEq{\mu_1}(\tamp{}, \tbar{}, \p) \\
+    \pderiv{\lag{M}}{\tamp{\mu_i}} &=
+    \denom{\mu_i}\tbar{\mu_i} +
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
+    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} |
+    \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}}
+    =
+    ^{\mathrm{PTE(S)}}\tbarEq{\mu_i}(\tamp{}, \tbar{}, \p) \quad \forall i \neq 1 \\
+    \pderiv{\lag{M}}{{\p^{\dagger}}} &=
+    \V\p + \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}} =
+    ^{\mathrm{PTE(S)}}\Omega_\p(\tamp{}, \tbar{}, \p)
+  \end{aligned}
+\end{equation}
+The amplitude equations are formally equivalent to those derived in the
+\acs{PTE} scheme. The equations for multipliers are also formally
+equivalent to the \acs{PTE} equations, apart from the equation
+determining the singles multipliers which is augmented by an additional term:
+$\p^\dagger\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_1}} |
+\mathrm{HF}}$
+This term is not, however, coupled to the other equations given that
+$\p$ is fixed once the singles amplitudes have been determined.
+Practical implementation of the \acs{PTE(S)} scheme will solve the
+\acs{PTE}-like amplitude equation, form the source term
+$\braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}}$, solve for
+the polarization degrees of freedom and calculate the polarization
+energy as a correction to the \acs{CC} energy.
+Only when a first-order molecular property, such as the molecular
+gradient, is needed will the multipliers equations need to be solved.
+
+Summarizing:
+In the \acs{PTE} scheme, the effective Lagrangian has the same form as
+the \emph{in vacuo} case. However, the Hamiltonian includes the
+reference reaction field. The amplitudes and multipliers equations
+are formally identical to the \emph{in vacuo} case. Neither
+computational prefactor nor scaling are affected, since there is no
+coupling in between the two and the reaction field is fixed.
+In the \acs{PTES(S)} scheme, the source term in the effective Lagrangian
+is restricted to the $T_1$-dependent term only. The amplitudes
+equations are formally identical to the \acs{PTE} case and are
+uncoupled from the multipliers and the polarization equations.
+The multipliers equations are formally equivalent to the \acs{PTE}
+case, but for the singles multipliers, which feature a
+polarization-dependent term.
+In other words, only the $T_1$-dependent part of the \acs{CC} density is
+used to define the classical sources.
+In the \acs{PTES} scheme, the multipliers equations are formally
+equivalent to the \acs{PTED} case. However, the
+polarization-dependent term is calculated by only retaining the
+$T_1$-dependent term in the source. This retains the coupling between the
+amplitudes and polarization equations, but avoid coupling the
+multipliers equation.
+As for the \acs{PTE(S)} scheme, only the $T_1$-dependent part of the
+\acs{CC} density is in fact used to define the classical sources.
+In the \acs{PTED} scheme, \ldots
+
+\todo[inline]{Can't really figure out how the \acs{PTES} scheme can be
+devised in this framework. The idea seems to be to preserve the form of
+the \acs{PTED} equations, but dropping the dependence on the Lagrange
+multipliers. This can be achieved by decoupling the \acs{PCM} equation
+from the multipliers equation, \ie{} by using only the part of the
+density matrix that depends on the $T_1$ amplitudes to define the source
+term. This seems to miss, however, part of the terms in the \acs{PTES}
+multipliers equations, that are, in Caricato's treatment, still
+coupled with the \acs{PCM} equation.}
+
+\section{Coupled cluster perturbation theory for a quantum/classical
+  polarizable Hamiltonian}\label{sec:ccpt-polarizable}
+
+We will now develop a perturbative expansion of the effective \acs{CC}
+Lagrangian and derive the \acs{CC}\acs{PT} series when a
+quantum/classical polarizable Hamiltonian is used.
+The fluctuation potential $\Phi$ is the perturbation. As such $\Phi =
+O(1)$, while the cluster operators will be of first, at least, and higher
+order in the fluctuation potential.
+Orders will be counted assuming an Hartree--Fock closed-shell reference:
+we will use square brackets $[i]$ for an $i$-th order contributions.
+
+The \acs{PTED}-\acs{CC} equations are expanded in orders of the
+perturbation and terms are collected order by order. The use of an
+effective, variational Lagrangian implies the validity of the $2n+1$
+rule for the amplitudes $\tamp{\mu_i}$ and polarization $\p$ and of the
+$2n+2$ rule for the multipliers $\tbar{\mu_i}$.
+We can thus derive energy corrections up to fifth order by means of the
+amplitudes and polarization up to and including second order terms and
+the multipliers up to and including second order.
+
+The polarization equation couples to the multipliers and amplitudes
+equations \emph{via} the source term, see Eq.~\eqref{eq:source-term}.
+Its perturbative expansion will be given as:
+\begin{equation}\label{eq:source-term-pt}
+  \tilde{\s}(\tamp{},\tbar{}) =
+  \tilde{\s}^{[0]}
+  + \tilde{\s}^{[1]}
+  + \tilde{\s}^{[2]}
+  + \ldots
+\end{equation}
+and correspondingly for the polarization:
+\begin{equation}\label{eq:pol-eq-pt}
+  \V\p^{[i]} + \tilde{\s}^{[i]} = 0
+\end{equation}
+
+The source operator $\s$ will
+be considered as \emph{zeroth-order} in the perturbation. The order of
+the contributions to $\tilde{\s}(\tamp{},\tbar{})$ will be solely
+determined by the amplitudes and multipliers.
+It is already clear that $\tilde{\s}(\tamp{},\tbar{})$ is at least first
+order in the fluctuation potential:
+\begin{equation}
+  \tilde{\s}(\tamp{},\tbar{}) =
+  \tilde{\s}^{[1]}
+  + \tilde{\s}^{[2]}
+  + \ldots
+\end{equation}
+which also implies:
+\begin{equation}
+  \V\p^{[0]} + \tilde{\s}^{[0]} = 0 \Rightarrow  \p^{[0]} = 0
+\end{equation}
+
+\paragraph*{Perturbative expansion of $\tilde{\s}$}
+
+\begin{equation}\label{eq:source-1st-order}
+  \tilde{\s}^{[1]}(\tamp{}, \tbar{}) =
+  \braket{\mathrm{HF} | \BCHfirst{\s}{T^{[1]}_1} | \mathrm{HF}}
+  + \braket{\tbar{1}^{[1]} | \s | \mathrm{HF}}
+\end{equation}
+In the free energy corrections we will use the $2n+1$ and $2n+2$ complying
+expression:
+\begin{equation}\label{eq:source-1st-order-rules}
+  \tilde{\s}^{[1]}(\tamp{}, \tbar{}) =
+  \braket{\mathrm{HF} | \BCHfirst{\s}{T^{[1]}_1} | \mathrm{HF}}
+\end{equation}
+
+\begin{equation}\label{eq:source-2nd-order}
+  \tilde{\s}^{[1]}(\tamp{}, \tbar{}) =
+  \braket{\mathrm{HF} | \BCHfirst{\s}{T^{[2]}_1} | \mathrm{HF}}
+  + \braket{\tbar{1}^{[2]} | \s | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i}^{[1]} | \BCHfirst{\s}{T^{[1]}} | \mathrm{HF}}
+\end{equation}
+In the free energy corrections we will use the $2n+1$ and $2n+2$ complying
+expression:
+\begin{equation}\label{eq:source-2nd-order-rules}
+  \tilde{\s}^{[2]}(\tamp{}, \tbar{}) =
+  \braket{\mathrm{HF} | \BCHfirst{\s}{T^{[2]}_1} | \mathrm{HF}}
+  + \braket{\tbar{i}^{[1]} | \BCHfirst{\s}{T^{[1]}} | \mathrm{HF}}
+\end{equation}
+
+\subsection{First order equations}\label{sec:first-order-pt}
+
+\paragraph*{Amplitudes}
+\begin{equation}
+  \Omega_{\mu_i}^{[1]} = \denom{\mu_i}\tamp{\mu_i}^{[1]}
+  + \braket{\mu_i | \Phi | \mathrm{HF}}
+  + \p^{\dagger\,[1]}\braket{\mu_i | \bar{\s}^{[0]} | \mathrm{HF}}
+  + \p^{\dagger\,[0]}\braket{\mu_i | \bar{\s}^{[1]}  | \mathrm{HF}}
+  = 0
+\end{equation}
+And since $\p^{[0]} = 0$ and $\bar{\s}^{[0]} = \s$:
+\begin{equation}\label{eq:1st-order-amp}
+  \Omega_{\mu_i}^{[1]} = \denom{\mu_i}\tamp{\mu_i}^{[1]}
+  + \braket{\mu_i | \Phi | \mathrm{HF}}
+  + \p^{\dagger\,[1]}\braket{\mu_i | \s | \mathrm{HF}}
+  = 0
+\end{equation}
+Considering the singles and doubles excitation manifolds:
+\begin{subequations}
+  \begin{align}
+  \Omega_{\mu_1}^{[1]} &= \denom{\mu_1}\tamp{\mu_1}^{[1]}
+  + \p^{\dagger\,[1]}\braket{\mu_1 | \s | \mathrm{HF}}
+  = 0 \\
+  \Omega_{\mu_2}^{[1]} &= \denom{\mu_2}\tamp{\mu_2}^{[1]}
+  + \braket{\mu_2 | \Phi | \mathrm{HF}}
+  = 0
+  \end{align}
+\end{subequations}
+Thus, despite the closed-shell \acs{HF} reference, the singles
+amplitudes will already appear in first order, due to the
+quantum/classical coupling.
+
+\paragraph*{Multipliers}
+  \begin{equation}\label{eq:1st-order-mult}
+    \bar{\Omega}_{\mu_i}^{[1]} =
+      \denom{\mu_i}\tbar{\mu_i}^{[1]}
+      + \braket{\mathrm{HF} | \commutator{\Phi}{\cluster{\mu_i}} | \mathrm{HF}}
+      + \p^{\dagger\,[1]}\braket{\mathrm{HF} |
+      \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}} =0
+  \end{equation}
+Considering the singles and doubles excitation manifolds:
+\begin{subequations}
+  \begin{align}
+    \bar{\Omega}_{\mu_1}^{[1]} &= \denom{\mu_1}\tbar{\mu_1}^{[1]}
+      + \p^{\dagger\,[1]}\braket{\mathrm{HF} |
+      \commutator{\s}{\cluster{\mu_1}} | \mathrm{HF}}
+  = 0 \\
+    \bar{\Omega}_{\mu_2}^{[1]} &= \denom{\mu_2}\tbar{\mu_2}^{[1]}
+      + \braket{\mathrm{HF} | \commutator{\Phi}{\cluster{\mu_2}} | \mathrm{HF}}
+  = 0
+  \end{align}
+\end{subequations}
+Thus, despite the closed-shell \acs{HF} reference, the singles
+multipliers will already appear in first order, due to the
+quantum/classical coupling.
+\paragraph*{Polarization}
+  \begin{equation}\label{eq:1st-order-pol}
+    \V\p^{[1]} + \braket{\mathrm{HF} | \commutator{\s}{T_1^{[1]}}| \mathrm{HF}}
+    + \braket{\tbar{1}^{[1]} | \s | \mathrm{HF} } = 0
+  \end{equation}
+
+\subsection{Second order equations}\label{sec:second-order-pt}
+
+\paragraph*{Amplitudes}
+\begin{equation}\label{eq:2nd-order-amp}
+  \Omega_{\mu_i}^{[2]} = \denom{\mu_i}\tamp{\mu_i}^{[2]}
+  + \braket{\mu_i | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
+  + \p^{\dagger\,[2]}\braket{\mu_i | \s | \mathrm{HF}}
+  + \p^{\dagger\,[1]}\braket{\mu_i | \commutator{\s}{T^{[1]}} |
+  \mathrm{HF}}
+  = 0
+\end{equation}
+Considering the singles, doubles and triples excitation manifolds:
+\begin{subequations}
+  \begin{align}
+  \Omega_{\mu_1}^{[2]} &= \denom{\mu_1}\tamp{\mu_1}^{[2]}
+  + \braket{\mu_1 | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
+  + \p^{\dagger\,[2]}\braket{\mu_1 | \s | \mathrm{HF}}
+  + \p^{\dagger\,[1]}\braket{\mu_1 | \commutator{\s}{T^{[1]}} |
+  \mathrm{HF}}
+  = 0 \\
+  \Omega_{\mu_2}^{[2]} &= \denom{\mu_2}\tamp{\mu_2}^{[2]}
+  + \braket{\mu_2 | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
+  + \p^{\dagger\,[1]}\braket{\mu_2 | \commutator{\s}{T_2^{[1]}} |
+  \mathrm{HF}}
+  = 0 \\
+  \Omega_{\mu_3}^{[2]} &= \denom{\mu_3}\tamp{\mu_3}^{[2]}
+  + \braket{\mu_3 | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
+  = 0
+  \end{align}
+\end{subequations}
+The triples first appear to second order and are not \emph{directly}
+affected by the quantum/classical coupling.
+\paragraph*{Multipliers}
+  \begin{equation}\label{eq:2nd-order-mult}
+    \begin{aligned}
+    \bar{\Omega}_{\mu_i}^{[2]} &=
+      \denom{\mu_i}\tbar{\mu_i}^{[2]}
+      + \braket{\mathrm{HF} | \commutator{\commutator{\Phi}{T^{[1]}}}{\cluster{\mu_i}} | \mathrm{HF}}
+      + \sum_j\braket{\tbar{j}^{[1]} |
+      \commutator{\Phi}{\cluster{\mu_i}} | \mathrm{HF}} \\
+      &+ \p^{\dagger\,[2]}\braket{\mathrm{HF} |
+      \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}
+      + \p^{\dagger\,[1]}
+      \sum_j\braket{\tbar{j}^{[1]} |
+      \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}
+      =0
+    \end{aligned}
+  \end{equation}
+Considering the singles, doubles and triples excitation manifolds:
+\begin{subequations}
+  \begin{align}
+    \bar{\Omega}_{\mu_1}^{[2]} &=
+      \denom{\mu_1}\tbar{\mu_1}^{[2]}
+      + \braket{\mathrm{HF} | \commutator{\commutator{\Phi}{T^{[1]}}}{\cluster{\mu_1}} | \mathrm{HF}}
+      + \sum_j\braket{\tbar{j}^{[1]} |
+      \commutator{\Phi}{\cluster{\mu_1}} | \mathrm{HF}} \\
+      &+ \p^{\dagger\,[2]}\braket{\mathrm{HF} |
+      \commutator{\s}{\cluster{\mu_1}} | \mathrm{HF}}
+      + \p^{\dagger\,[1]}
+      \braket{\tbar{1}^{[1]} |
+      \commutator{\s}{\cluster{\mu_1}} | \mathrm{HF}}
+      =0 \\
+    \bar{\Omega}_{\mu_2}^{[2]} &=
+      \denom{\mu_2}\tbar{\mu_2}^{[2]}
+      + \braket{\mathrm{HF} | \commutator{\commutator{\Phi}{T^{[1]}}}{\cluster{\mu_2}} | \mathrm{HF}}
+      + \sum_j\braket{\tbar{j}^{[1]} |
+      \commutator{\Phi}{\cluster{\mu_2}} | \mathrm{HF}} \\
+      &+ \p^{\dagger\,[1]}
+      \braket{\tbar{1}^{[1]} + \tbar{2}^{[1]} |
+      \commutator{\s}{\cluster{\mu_2}} | \mathrm{HF}}
+      =0 \\
+    \bar{\Omega}_{\mu_3}^{[2]} &=
+      \denom{\mu_3}\tbar{\mu_3}^{[2]}
+      + \braket{\mathrm{HF} | \commutator{\commutator{\Phi}{T^{[1]}}}{\cluster{\mu_3}} | \mathrm{HF}}
+      + \sum_j\braket{\tbar{j}^{[1]} |
+      \commutator{\Phi}{\cluster{\mu_3}} | \mathrm{HF}}
+      =0
+  \end{align}
+\end{subequations}
+The triples first appear to second order and are not \emph{directly}
+affected by the quantum/classical coupling.
+
+\paragraph*{Polarization}
+  \begin{equation}\label{eq:2nd-order-pol}
+    \begin{aligned}
+    \V\p^{[2]} &+
+    \braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}}| \mathrm{HF}}
+    + \braket{\tbar{1}^{[2]} | \s | \mathrm{HF} }
+    + \braket{\tbar{1}^{[1]} | \commutator{\s}{T_1^{[1]}} | \mathrm{HF}}
+    \\
+    &+
+    \braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[1]}} | \mathrm{HF}}
+    = 0
+  \end{aligned}
+  \end{equation}
+No triples contributions enter directly up to second order.
+
+\subsection{Free energies up to fifth order}\label{sec:energies-pt}
+
+From the expansion of the effective Lagrangian we can obtain free energy
+corrections up to the desired order. Given the variational nature of the
+Lagrangian, only terms fulfilling the $2n+1$ and $2n+2$ rules will
+appear in the free energy corrections.
+To zeroth-order all variational parameters are zero. To first order we
+thus have:
+\begin{equation}\label{eq:first-order-G}
+  G^{[1]} = \braket{\mathrm{HF} | \Phi | \mathrm{HF}}
+\end{equation}
+The reference energy can thus be recovered as usual:
+\begin{equation}
+  G_\mathrm{HF} = G^{[0]} + G^{[1]}
+\end{equation}
+where $G^{[0]}$ includes the quantum/classical coupling.
+\todo[inline]{Check this statement, the Lagrangian has a zeroth-order
+term!}
+
+\begin{equation}\label{eq:second-order-G}
+  \begin{aligned}
+  G^{[2]} &= \braket{\mathrm{HF} | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
+  + \frac{1}{2}\p^{\dagger\,[1]}\V\p^{[1]} +
+  \p^{\dagger\,[1]}\tilde{\s}^{[1]} \\
+  &= E^{[2]} + \frac{1}{2}\p^{\dagger\,[1]}\V\p^{[1]} +
+  \p^{\dagger\,[1]}
+  \braket{\mathrm{HF} | \commutator{\s}{T^{[1]}_1} | \mathrm{HF}}
+  \end{aligned}
+\end{equation}
+Notice that the source term is here assembled in accordance to the $2n+1$
+and $2n+2$ rules, while this is not the case in the equations for the
+polarization.
+
+\begin{equation}\label{eq:third-order-G}
+  \begin{aligned}
+    G^{[3]} &= \sum_j\braket{\tbar{j}^{[1]} | \commutator{\Phi}{T^{[1]}} |
+    \mathrm{HF}} + \p^{\dagger\, [1]}\tilde{\s}^{[2]} \\
+    &= E^{[3]}
+    + \p^{\dagger\, [1]}
+    \lbrace
+      \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}_1} | \mathrm{HF}}
+    + \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}_2} | \mathrm{HF}}
+    + \braket{\tbar{2}^{[1]} | \commutator{\s}{T^{[1]}_2} | \mathrm{HF}}
+    \rbrace
+  \end{aligned}
+\end{equation}
+
+\begin{equation}\label{eq:fourth-order-G}
+  \begin{aligned}
+    G^{[4]} &= \sum_j\braket{\tbar{j}^{[1]} | \commutator{\Phi}{T^{[2]}} |
+    \mathrm{HF}} +
+    \frac{1}{2}\sum_j\braket{\tbar{j}^{[1]} |
+    \commutator{\commutator{\Phi}{T^{[1]}}}{T^{[1]}} |
+    \mathrm{HF}} \\
+    &+\frac{1}{2}\p^{\dagger\,[2]}\V\p^{[2]}
+    + \p^{\dagger\,[2]}\tilde{\s}^{[2]}
+    + \p^{\dagger\,[1]}\tilde{\s}^{[3]} \\
+    &= E^{[4]} + \frac{1}{2}\p^{\dagger\,[2]}\V\p^{[2]}
+    +\p^{\dagger\,[2]}
+    \lbrace
+      \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}_1} | \mathrm{HF}}
+    + \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}_2} | \mathrm{HF}}
+    + \braket{\tbar{2}^{[1]} | \commutator{\s}{T^{[1]}_2} | \mathrm{HF}}
+    \rbrace \\
+    &+\p^{\dagger\,[1]}
+    \lbrace
+    \frac{1}{2}\braket{\tbar{1}^{[1]} |
+    \commutator{\commutator{\s}{T^{[1]}_1}}{T^{[1]}_1}
+    |
+    \mathrm{HF}}
+    + \braket{\tbar{2}^{[1]} |
+    \commutator{\commutator{\s}{T^{[1]}_1}}{T^{[1]}_2}
+    |
+    \mathrm{HF}}
+    \rbrace
+  \end{aligned}
+\end{equation}
+
+\begin{equation}\label{eq:fifth-order-G}
+  \begin{aligned}
+    G^{[5]} &=
+    \frac{1}{2}\braket{\mathrm{HF} |
+    \commutator{\commutator{\Phi}{T^{[2]}}}{T^{[2]}}
+    | \mathrm{HF}}
+    +
+    \sum_j \braket{\tbar{j}^{[1]} |
+    \commutator{\commutator{\Phi}{T^{[2]}}}{T^{[1]}}
+    | \mathrm{HF}} \\
+    &+ \sum_j \braket{\tbar{j}^{[2]} |
+      \commutator{\Phi}{T^{[2]}}
+    | \mathrm{HF}}
+    +
+    \frac{1}{2}\sum_j\braket{\tbar{j}^{[2]} |
+    \commutator{\commutator{\Phi}{T^{[1]}}}{T^{[1]}}
+    | \mathrm{HF}}
+    + \p^{\dagger\,[2]}\tilde{\s}^{[3]}
+    + \p^{\dagger\,[1]}\tilde{\s}^{[4]} \\
+    &= E^{[5]}
+    + \p^{\dagger\,[2]}
+    \lbrace
+    \frac{1}{2}\braket{\tbar{1}^{[1]} |
+    \commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[1]}}
+    | \mathrm{HF}}
+    + \braket{\tbar{2}^{[1]} |
+    \commutator{\commutator{\s}{T_1^{[1]}}}{T_2^{[1]}}
+    | \mathrm{HF}}
+    \rbrace \\
+    &+ \p^{\dagger\,[1]}
+    \lbrace
+    \braket{\tbar{1}^{[1]} |
+    \commutator{\commutator{\s}{T_1^{[2]}}}{T_1^{[1]}}
+    | \mathrm{HF}}
+    + \braket{\tbar{2}^{[1]} |
+    \commutator{\commutator{\s}{T_1^{[2]}}}{T_2^{[1]}}
+    | \mathrm{HF}}
+    + \braket{\tbar{2}^{[1]} |
+    \commutator{\commutator{\s}{T_1^{[1]}}}{T_2^{[2]}}
+    | \mathrm{HF}}
+    \rbrace
+  \end{aligned}
+\end{equation}

### 2016-07-27 Chapters/Ch04-PaperSummary.tex
Some more work

@@
 %************************************************
-\chapter{Summary of the papers}\label{ch:papers-summary}
+\chapter{Summary of the publications}\label{ch:publications-summary}
 
@@
 
+\begin{figure}[tb]
+ \centering
+  \missingfigure{Take figure from paper!}
+  \caption[Convergence of $\alpha_\mathrm{iso}$ with respect to the number of molecular
+  electrostatic potential (MEP) evaluation points on the cavity surface.]
+  {Convergence of $\alpha_\mathrm{iso}$ with respect to the number of
+   molecular electrostatic potential (MEP) evaluation points on the cavity
+   surface.
+   The values reported are in $a_0^3$ and refer to benzene.
+   The lower axis reports the average area for the collocation
+   tesselation, while the upper axis refers to the patch level in the
+   wavelet Galerkin discretization. The annotation report the number of
+   MEP evaluation points.
+   The compression parameter triplet was set to $a = 2.5,\ d^\prime =
+   2.0,\ b = 0.01$ for the piecewise constant wavelets and to $a = 2.5,\
+   d^\prime = 3.0,\ b = 0.01$ for piecewise bilinear wavelets.
+   The limit value is extrapolated from the results obtained from the
+   piecewise bilinear wavelet Galerkin scheme.}
+  \label{fig:alpha_convergence}
+\end{figure}
+
 The interface to \LSDALTON was later released in version ...

### 2016-07-27 FrontBackmatter/Notation.tex
Some more work

@@
 
+\todo[inline]{Clarify the notation for the 4-component basis in
+accordance with the EPR paper.}
 Further conventions have been adopted for the indices of functions.
-Lower case Latin letters are used for MO 4-spinors, while lower case
+Lower case Latin letters are used for \ac{MO} 4-spinors, while lower case
 Greek letters ($\kappa, \lambda, \mu, \ldots$) are reserved for
@@
 
-Hartree atomic units have been used:~\autocite{Whiffen1978}
+Hartree atomic units have been used:~\autocite{Whiffen1978-xx}
 \begin{equation*}

### 2016-07-27 Chapters/Ch03-PCMCoupledCluster.tex
Beef up summary of papers

@@
 which is equivalent to imposing normal ordering of the
-operators.~\cite{Shavitt}
+operators.~\autocite{Shavitt}
 The total functional will include both reference and correlation
@@
 \end{equation}
-In accordance with previous work~\cite{Cammi2009-gu, Caricato2011-tx}, the
+In accordance with previous work,\autocite{Cammi2009-gu, Caricato2011-tx} the
 Hamiltonian $H_0$ is augmented with the terms
@@
 transformations will truncate after two
-terms \cite{Helgaker2000-tz}:
+terms:\autocite{Helgaker2000-tz}
 \begin{equation}
@@
 commutator, the $n_i$ are their respective excitations ranks and $m_O$
-is the \emph{particle rank} of the operator $O$ \cite{Helgaker2000-tz}.
+is the \emph{particle rank} of the operator $O$.\autocite{Helgaker2000-tz}
 This follows from the realization that:

### 2016-07-27 Chapters/Ch04-PaperSummary.tex
Beef up summary of papers

@@
 conclusions of the papers this thesis is based on.
+Figures and tables here reported are taken \emph{verbatim} from the published work.
 In addition, I give a description of my contributions to the work leading to
@@
 regime.
+
+\begin{figure}[!htb]
+\centering
+\scalebox{0.8}{\input{gfx/algorithm.tikz}}
+\caption{
+  Schematic view of the implemented SCF algorithm. Computations/data in
+  \textcolor{Blue}{blue} are on the \pcmsolver side, in
+  \textcolor{Green}{green} on the \DIRAC side.
+  }
+\label{fig:algorithm}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{subfigure}[b]{0.5\textwidth}
+\missingfigure{Reobtain this figure: H2O\_MEP-diff-pcmdc-dc}
+\caption{\ce{H2O}: $\mathrm{MEP}_\mathrm{PCM-DC} - \mathrm{MEP}_\mathrm{DC}$}
+\label{fig:H2Odiff_pcmdc-dc}
+\end{subfigure}%
+\begin{subfigure}[b]{0.5\textwidth}
+\missingfigure{Reobtain this figure: H2Po\_MEP-diff-pcmdc-dc}
+\caption{\ce{H2Po}: $\mathrm{MEP}_\mathrm{PCM-DC} - \mathrm{MEP}_\mathrm{DC}$}
+\label{fig:H2Podiff_pcmdc-dc}
+\end{subfigure}
+\caption{
+  \todo[inline]{Recalculate MEP maps and reobtain these plots using ParaView.}
+  Molecular electrostatic potential (\acs{MEP}) maps for \ce{H2O} and \ce{H2Po} at
+  the \acs{DFT}/PBE0 level of theory.
+  The isocontours are in the range $[-0.02, 0.02] \si{\hartree\per\elementarycharge}$ and are
+  color-coded from red to blue.
+  The values plotted are differences between the values obtained for the
+  Hamiltonians referred to in the subcaptions. The geometry optimized \emph{in
+  vacuo}, using the \acl{DC} Hamiltonian as reference.
+  \acs{PCM}-\acs{DC}: \acl{DC} in water, \acs{DC}: \acl{DC} \emph{in vacuo}.}
+\label{fig:MEP_maps}
+\end{figure}
 
@@
  \centering
-  \missingfigure{Take figure from paper!}
+  \missingfigure{Reobtain plot from paper using a style similar to that in MPA paper}
   \caption[Convergence of $\alpha_\mathrm{iso}$ with respect to the number of molecular

### 2016-07-28 Chapters/Ch04-PaperSummary.tex
Not much...

@@
 \chapter{Summary of the publications}\label{ch:publications-summary}
+
+\epigraph{$\frac{1}{2}$ research and $\frac{1}{2}\,\,T\acute{\epsilon}\chi\nu\eta$
+
+          $\frac{1}{2}$ observation, $\frac{1}{2}\,\,T\acute{\epsilon}\chi\nu\eta$
+
+          $\frac{1}{2}$ training, $\frac{1}{2}\,\,T\acute{\epsilon}\chi\nu\eta$}{
+  --- \textsc{Ezra Pound}, \textit{Canto LXXXV}}
 
@@
 \scalebox{0.8}{\input{gfx/algorithm.tikz}}
-\caption{
+\caption[Modular approach to programming a \acs{PCM} functionality into an existing \acs{SCF} code.]{
   Schematic view of the implemented SCF algorithm. Computations/data in
@@
 
+\todo[inline]{Recalculate MEP maps and reobtain these plots using ParaView.}
 \begin{figure}
@@
 \end{subfigure}
-\caption{
-  \todo[inline]{Recalculate MEP maps and reobtain these plots using ParaView.}
+\caption[Molecular electrostatic potential (\acs{MEP}) maps for \ce{H2O} and \ce{H2Po} at
+  the \acs{DFT}/PBE0 level of theory.]{
   Molecular electrostatic potential (\acs{MEP}) maps for \ce{H2O} and \ce{H2Po} at

### 2016-07-28 FrontBackmatter/Acknowledgments.tex
Add commands, fix notations
Not much...

--- 2016-07-17 FrontBackmatter/Introduction.tex
@@
-\chapter*{Introduction}
-\addtocontents{toc}{\protect\vspace{\beforebibskip}} % to have the bib a bit from the rest in the toc
-\addcontentsline{toc}{section}{\tocEntry{Introduction}}
+%*******************************************************
+% Acknowledgments
+%*******************************************************
+\pdfbookmark[1]{Acknowledgments}{acknowledgments}
 
-\epigraph{How does newness come into world? How is it born?
+\bigskip
 
-    Of what fusions, translations, conjoinings is it made?
+\begingroup
+\let\clearpage\relax
+\let\cleardoublepage\relax
+\let\cleardoublepage\relax
+\chapter*{Acknowledgments}
 
-    How does it survive, extreme and dangerous as it is? What compromises, what deals,
-    what betrayals of its secret nature must it take to stave off the wrecking crew, the
-    exterminating angel, the guillotine?
+\epigraph{
+Just a perfect day
 
-    Is birth always a fall?
+You made me forget myself
 
-    Do angels have wings? Can men fly?}{
-    --- \textsc{Salman Rushdie}, \textit{The Satanic Verses}}
+I thought I was
+
+Someone else, someone good}{
+--- \textsc{Lou Reed}}
+
+\endgroup

### 2016-07-28 FrontBackmatter/Introduction.tex
Not much...

@@
 
-\epigraph{How does newness come into world? How is it born?
-
-    Of what fusions, translations, conjoinings is it made?
-
-    How does it survive, extreme and dangerous as it is? What compromises, what deals,
-    what betrayals of its secret nature must it take to stave off the wrecking crew, the
-    exterminating angel, the guillotine?
-
-    Is birth always a fall?
-
-    Do angels have wings? Can men fly?}{
-    --- \textsc{Salman Rushdie}, \textit{The Satanic Verses}}
+\epigraph{
+pity this busy monster, manunkind,
+    }{
+    --- \textsc{E. E. Cummings}}

### 2016-08-10 gfx/wemlin/alpha_convergence.py
Add convergence of isotropic static polarizability WRT finite elements picture

@@ -1,0 +1,100 @@
+#!/usr/bin/python2
+# -*- python -*-
+# -*- coding: utf-8 -*-
+# vim:filetype=python:
+# (c) Roberto Di Remigio  <roberto.d.remigio@uit.no>
+# licensed under the GNU Lesser General Public License
+
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+import numpy as np
+
+params = {'text.usetex': False, 'mathtext.fontset': 'stixsans'}
+mpl.rcParams.update(params)
+mpl.style.use('ggplot')
+# The colors are from the ggplot style, consistent with grey background
+ggplot = ['#E24A33', '#348ABD', '#988ED5', '#777777', '#FBC15E', '#8EBA42', '#FFB5B8']
+
+limit = 76.678052
+# Get the data
+fname = 'ief_convergence.dat'
+data = np.loadtxt(fname, usecols=(0, 4))
+ief_x = data[:, 0].astype(int)
+ief_y = data[:, 1]
+
+fname = 'pwc_convergence.dat'
+data = np.loadtxt(fname, usecols=(0, 4))
+pwc_x = data[:, 0].astype(int)
+pwc_y = data[:, 1]
+
+fname = 'pwl_convergence.dat'
+data = np.loadtxt(fname, usecols=(0, 4))
+pwl_x = data[:, 0].astype(int)
+pwl_y = data[:, 1]
+
+y_formatter = mpl.ticker.ScalarFormatter(useOffset=False)
+
+fig, ax = plt.subplots(figsize=(12, 8))
+
+ax.spines['right'].set_color('none')
+ax.yaxis.set_ticks_position('left')
+
+ax.tick_params(axis='x')
+ax.tick_params(axis='y')
+
+l3 = ax.plot(ief_x, ief_y, label=r'Collocation', lw=1, ls='-', marker='d', color=ggplot[0])
+dof = [str(x) for x in ief_x]
+coords = zip(ief_x, ief_y)
+ax.annotate('{}'.format(dof[0]), xy=(coords[0]), xytext=(0, 10),  va='bottom', ha='right',  textcoords='offset points', color=ggplot[0], weight='demi')
+ax.annotate('{}'.format(dof[1]), xy=(coords[1]), xytext=(0, 10),  va='bottom', ha='right',  textcoords='offset points', color=ggplot[0], weight='demi')
[…]

### 2016-08-10 Chapters/Ch04-PaperSummary.tex
wemlin paper summary: add picture to text.

@@
 conclusions of the papers this thesis is based on.
-Figures and tables here reported are taken \emph{verbatim} from the published work.
+Figures and tables here reported are minimally modified versions taken from the published work,
+captions are instead copied \emph{verbatim}.
 In addition, I give a description of my contributions to the work leading to
@@
  \centering
-  \missingfigure{Reobtain plot from paper using a style similar to that in MPA paper}
+  \includegraphics[width=\textwidth]{gfx/wemlin/alpha_convergence.png}
   \caption[Convergence of $\alpha_\mathrm{iso}$ with respect to the number of molecular

### 2016-08-10 gfx/wemlin/alpha_convergence.py

@@ -10,11 +10,15 @@ import matplotlib.pyplot as plt
 
 import numpy as np
 
-params = {'text.usetex': False, 'mathtext.fontset': 'stixsans'}
-mpl.rcParams.update(params)
 mpl.style.use('ggplot')
 # The colors are from the ggplot style, consistent with grey background
 ggplot = ['#E24A33', '#348ABD', '#988ED5', '#777777', '#FBC15E', '#8EBA42', '#FFB5B8']
+
+params = {'text.usetex': False,
+          'mathtext.fontset': 'stixsans',
+          'font.size' : 14
+         }
+mpl.rcParams.update(params)
 
 limit = 76.678052
 # Get the data
@@ -80,13 +84,8 @@ ax2.set_xticklabels(labels)
 
 l4 = ax2.plot((1.5, 5.5), (limit, limit), label='Limit', ls='-.', color=ggplot[3])
 
-#ax.set_ylim([-5.26, -5.16])
 ax.yaxis.set_major_formatter(y_formatter)
 ax.set_ylabel(r'Isotropic polarizability $[a_0^3]$')
-ax.yaxis.set_ticks_position('none')
-
-ax.xaxis.set_ticks_position('none')
-ax2.xaxis.set_ticks_position('none')
 
 ax.grid(False)
 ax2.grid(False)

### 2016-08-10 Chapters/Ch02-ResponseTheory.tex
Start on papers summaries and contributions
Add commands, fix notations
Write summary for paper III

--- 2016-07-17 Chapters/Chapter02.tex
@@
 %*****************************************
-\chapter{Response Theory and Molecular Properties}\label{ch:molprop}
+\chapter{Response theory and molecular properties}\label{ch:molprop}
 %*****************************************
-\epigraph{What's the frequency, Kenneth?}{--- \textsc{R.E.M.}}
+
+\epigraph{
+White light, white light goin' messin' up my brain
+
+White light, aww white light it's gonna drive me insane
+}{
+--- \textsc{The Velvet Underground}
+}
+
+\epigraph{What's the frequency, Kenneth? [...]
+
+          I never understood the frequency, uh-huh}{
+          --- \textsc{R.E.M.}}
 

### 2016-08-10 Chapters/Ch04-PaperSummary.tex
Write summary for paper III

@@
 regime.
+Our preliminary calculations on the group 16 dihydrides \ce{H2X} (\ce{X} =
+\ce{O}, \ce{S}, \ce{Se}, \ce{Te}, \ce{Po}) have shown that the method predicts
+a noticeable interplay of relativistic and solvent effects when heavier
+elements are involved.
+The main point of the paper was, however, the adoption of a fully modular
+programming strategy. We showed that it is entirely possible to adopt the same
+\acs{PCM} code and implementation "checkpoints" across altogether different
+problem domains.
 
@@
 
+As a by-product of the 4-component implementation, we were able to obtain and visualize \acs{MEP} maps
+from 4-component \acs{SCF} wave functions. These add yet another interpretive tool to the toolbox available
+in the 4-component relativistic regime.
 \todo[inline]{Recalculate MEP maps and reobtain these plots using ParaView.}
@@
 \end{figure}
+The interface to \DIRAC was later released in the 2014 and later versions,
+providing \acs{PCM} capabilities to the software package.
 
@@
 estimates.
-Thus, wavelet Galerkin \acs{BEM} represents a valid alternative to traditional collocation methods,
-both to achieve a better computational scaling and to provide accurate, benchmark results.
+Thus, wavelet Galerkin \acs{BEM} represents a valid alternative to traditional
+collocation methods, both to achieve a better computational scaling and to
+provide accurate, benchmark results.\autocite{Harbrecht2004-uo,
+Harbrecht2006-ug, Dahmen2006-pj}
 
@@
 For this paper, I provided template interface and test sets for the cavity
-generator\autocite{Harbrecht2009-no, Harbrecht2011-dk}
-and wavelet Galerkin \acs{BEM} solver\autocite{Harbrecht2004-uo, Harbrecht2006-ug, Dahmen2006-pj}
-with \pcmsolver.
+generator\autocite{Harbrecht2009-no, Harbrecht2011-dk} and wavelet Galerkin
+\acs{BEM} solver with \pcmsolver.
 These were used to interface with the new C++ implementation of the wavelet
@@
 
-The interface to \LSDALTON was later released in version ...
+Finally, the interface to \LSDALTON was later released in the 2016 version,
 providing \acs{PCM} capabilities to the software package.
@@
 
-%************************************************
-%*****************************************
-%*****************************************
-%*****************************************
-%*****************************************
-%*****************************************
+Continuum solvation models offer a simple route to the treatment of non
+homogeneous environments.
+The general integral equation formulation exposed in Chapter~\ref{ch:CSM} is in
+fact transparent with respect to the definition of the Green's function for the
+space portion exterior to the cavity.
+As already shown by \citeauthor{Cances1998-og}, the same mathematical formalism can easily accommodate
+environments characterized by distance-dependent or anisotropic Green's functions, such as those for
+ionic liquids (in the weakly interacting regime described by the linearized Poisson--Boltzmann equation)
+and liquid crystals, respectively.
+\citeauthor{Frediani2004-er} showed that a numerical representation of the
+Green's function is sufficient to obtain the boundary integral operators in the
+\acs{PCM} integral equation.
+The authors introduced a numerical integration procedure to calculate the
+Green's function for an environment characterized by spatially varying, yet
+cilindrically symmetric, permittivity functions: a model for planar diffuse interfaces.
+
+In this work, a similar procedure was introduced to tackle diffuse interfaces
+in spherical symmetry.
+In contrast to previously existing work, our implementation offer a more robust
+treatment of thin interfaces, with a rather generic functional form for the
+permittivity profile.
+We thoroughly analyzed the necessity for the \emph{a posteriori} removal of the
+Coulomb singularity from the computed Green's function and its efficient
+implementation.
+Transfer of ions and molecules across spherically symmetric interfaces exhibits
+peculiar properties. These arise as consequence of both interface width and
+interface curvature.
+In this work, we analyzed both effect on the water-vapor and oil-water transfer
+of \ce{Li+}, \ce{Br-}, acetone, \emph{para}-nitroaniline and the L0 dye.
+
+\begin{figure}[tb]
+  \centering
+  \begin{subfigure}[t]{.49\linewidth}
+  \centering
+  %\includegraphics[width=.49\linewidth]{placeholder}
+  % Figure8a <--> dipole_acetone_perpendicular
+\missingfigure{Reobtain this figure}
+  \caption{Width effect}
+  \label{fig:profile_dipole_acetone_width}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[t]{.49\linewidth}
+  \centering
+  %\includegraphics[width=.49\linewidth]{placeholder}
+  % Figure8b <--> curv_fullprofile_dip_acetone
+\missingfigure{Reobtain this figure}
+  \caption{Radius effect}
+  \label{fig:profile_dipole_acetone_fullprofile}
+  \end{subfigure}
+
+  \begin{subfigure}[t]{.49\linewidth}
+  \centering
+  %\includegraphics[width=.49\linewidth]{placeholder}
+  % Figure8c <--> curv_dip_acetone
+\missingfigure{Reobtain this figure}
+  \caption{Curvature effect}
+  \label{fig:profile_dipole_acetone_curvature}
+  \end{subfigure}
+  \caption[Width, radius and curvature effect on the dipole moment of acetone for water-air droplet phase transfer.]
+  {Dipole moment profiles for acetone (with dipole moment
+  perpendicular to the interface) transfer at water droplet in air, for
+  varying interface widths, values of $R_0$ and the curvature effect on
+  dipole moment, respectively.}
+  \label{fig:profile_dipole_acetone}
+\end{figure}

### 2016-08-11 Chapters/Ch04-PaperSummary.tex
Write summary of third paper

@@
 
+\refstepcounter{dummy}
+\addcontentsline{toc}{section}{\tocEntry{Continuum solvation in the relativistic regime}}
 \section*{Paper I. Continuum solvation in the relativistic regime}
@@
 \centering
-\scalebox{0.8}{\input{gfx/algorithm.tikz}}
+\scalebox{0.8}{\input{gfx/relapcm/algorithm.tikz}}
 \caption[Modular approach to programming a \acs{PCM} functionality into an existing \acs{SCF} code.]{
@@
 
+\refstepcounter{dummy}
+\addcontentsline{toc}{section}{\tocEntry{The wavelet Galerkin boundary element method for PCM}}
 \section*{Paper II. The wavelet Galerkin boundary element method for PCM}
@@
 
-This work builds upon and extends previous work by~\citeauthor{Weijo2010-hy}
-\citeauthor{Weijo2010-hy} used \ac{PWC} bases to show that superior accuracy
-and convergence could be obtained for the calculation of molecular energies.
-In this work we showed that faster convergence can be achieved when \ac{PWL} bases are used instead.
+Already~\citeauthor{Weijo2010-hy} had shown that using \ac{PWC} bases can lead
+to superior accuracy and convergence in the calculation of quantum mechanical
+molecular solvation energies.
+In this work we showed that even faster convergence can be achieved when
+\ac{PWL} bases are used instead.
 Moreover, the same holds for the calculation of static electric properties.
@@
 
+\refstepcounter{dummy}
+\addcontentsline{toc}{section}{\tocEntry{Continuum modelling of non homogeneous environments}}
 \section*{Paper III: Continuum modelling of non homogeneous environments}
@@
 implementation.
-Transfer of ions and molecules across spherically symmetric interfaces exhibits
-peculiar properties. These arise as consequence of both interface width and
-interface curvature.
+Interface width and curvature influence the transfer of ions and molecules across
+spherically symmetric interfaces and peculiar properties may arise.
 In this work, we analyzed both effect on the water-vapor and oil-water transfer
 of \ce{Li+}, \ce{Br-}, acetone, \emph{para}-nitroaniline and the L0 dye.
+Nonelectrostatic interactions were not included in our implementation, although
+they have been proved to be crucial for non homogeneous
+environments.~\autocite{Mozgawa2014-ad}
+Nevertheless our implementation represents a first significant step in the continuum treatment
+of such nontrivial environments.
 
@@
   \centering
-  %\includegraphics[width=.49\linewidth]{placeholder}
   % Figure8a <--> dipole_acetone_perpendicular
-\missingfigure{Reobtain this figure}
+  \includegraphics[width=\textwidth]{gfx/spherical/width.png}
   \caption{Width effect}
@@
   \centering
-  %\includegraphics[width=.49\linewidth]{placeholder}
   % Figure8b <--> curv_fullprofile_dip_acetone
-\missingfigure{Reobtain this figure}
+  \includegraphics[width=\textwidth]{gfx/spherical/radius.png}
   \caption{Radius effect}
@@
   \centering
-  %\includegraphics[width=.49\linewidth]{placeholder}
   % Figure8c <--> curv_dip_acetone
-\missingfigure{Reobtain this figure}
+  \includegraphics[width=\textwidth]{gfx/spherical/curvature.png}
   \caption{Curvature effect}
@@
 \end{figure}
+
+I contributed most of the theoretical work for this paper, based on early drafts from coauthors
+Ville Weijo and Hui Cao. In particular, I derived the separation of
+the Coulomb singularity in its final form.
+Moreover, I contributed the implementation and testing of the Green's function code.
+The interface to the \LSDALTON program package, developed within paper II, was
+also used for this paper.
+I wrote the first draft of the paper and coordinated all subsequent editing stages.
+
+\refstepcounter{dummy}
+\addcontentsline{toc}{section}{\tocEntry{Relativistic calculation of EPR and pNMR parameters in solution}}
+\section*{Paper IV: Relativistic calculation of EPR and pNMR parameters in solution}
+
+\refstepcounter{dummy}
+\addcontentsline{toc}{section}{\tocEntry{Open-ended self-consistent field response theory in solution}}
+\section*{Paper V: Open-ended self-consistent field response theory in solution}

### 2016-08-11 FrontBackmatter/Publications.tex
Write summary of third paper

@@
 \pdfbookmark[1]{Publications}{Publications}
-\chapter*{Publications}%\graffito{This is just an early --~and currently ugly~-- test!}
+\chapter*{Publications}
 
@@
 
+\item
+  \textbf{R. Di Remigio}, M. Repisky, S. Komorovsky, P. Hrobarik, L.
+  Frediani, and K. Ruud
+  \\
+  Four-Component Relativistic Density Functional Theory with the
+  Polarizable Continuum Model: Application to EPR Parameters
+  and Paramagnetic NMR Shifts
+  \\
+  submitted for publication to \textit{Mol. Phys.}
+\label{pcmepr}
+
+\item
+  \textbf{R. Di Remigio}, M. T. P. Beerepoot, Y. Cornaton, M. Ringholm,
+  A. H. S. Steindal, K. Ruud, and L. Frediani
+  \\
+  Open-ended formulation of self-consistent field response theory with
+  the polarizable continuum model for solvation
+  \\
+  In preparation for submission to \textit{Phys. Chem. Chem. Phys.}
+\label{pcmopenrsp}
 \end{enumerate}
+
+The following proceeding is also part of the thesis:
+\begin{enumerate}
+\item
+  \textbf{R. Di Remigio}, T. D. Crawford, and L. Frediani
+  \\
+  PCMSolver: an Application Programming Interface
+  for the Polarizable Continuum Model
+  \\
+  Accepted for publication in the proceedings of the
+  \emph{Producing High Performance and Sustainable Software for
+  Molecular Simulation} workshop held at the 2015 Supercomputing
+  Conference
+\label{SC15}
+\end{enumerate}

### 2016-08-11 gfx/spherical/curvature.py

@@ -1,0 +1,36 @@
+#!/usr/bin/env python
+
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+import numpy as np
+
+mpl.style.use('ggplot')
+# The colors are from the ggplot style, consistent with grey background
+ggplot = ['#E24A33', '#348ABD', '#988ED5', '#777777', '#FBC15E', '#8EBA42', '#FFB5B8']
+# List of line styles
+lines = ['-', '--', ':', '-.', '-', '--', ':']
+# List of labels
+labs = [r'$R_0 = 40$', r'$R_0 = 20$', r'$R_0 = 10$', r'$R_0 = 5$']
+params = {'text.usetex': False,
+          'mathtext.fontset': 'stixsans',
+          'font.size' : 14
+         }
+mpl.rcParams.update(params)
+
+fig, ax = plt.subplots(figsize=(10, 5), facecolor='w', edgecolor='k')
+
+fname='curv_dip_acetone_perpen.dat'
+
+# Read columns
+data = np.loadtxt(fname, usecols=(0, 1, 2, 3, 4))
+for col in range(1, 5):
+    ax.plot(data[:, 0], data[:, col], color=ggplot[col-1], marker='+', ls=lines[col-1], lw=2, ms=7, mew=1.5, label=labs[col-1])
+    ax.set_ylabel(r'$\mu\,\,[\mathrm{D}]$')
+    ax.legend(loc='upper right', ncol=1)
+
+ax.set_xlim(-5.0, 10.0)
+ax.set_ylim(0.0, 1.4)
+ax.set_xlabel(r'$z\,\,[\AA]$')
+
+fig.savefig('curvature.png', transparent=False, bbox_inches='tight')

### 2016-08-11 gfx/spherical/radius.py

@@ -1,0 +1,39 @@
+#!/usr/bin/env python
+
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+import numpy as np
+
+mpl.style.use('ggplot')
+# The colors are from the ggplot style, consistent with grey background
+ggplot = ['#E24A33', '#348ABD', '#988ED5', '#777777', '#FBC15E', '#8EBA42', '#FFB5B8']
+# List of line styles
+lines = ['-', '--', ':', '-.', '-', '--', ':']
+params = {'text.usetex': False,
+          'mathtext.fontset': 'stixsans',
+          'font.size' : 14
+         }
+mpl.rcParams.update(params)
+
+fig, ax = plt.subplots(figsize=(10, 5), facecolor='w', edgecolor='k')
+
+R0_5  = np.loadtxt(fname='R0_5.dat',  usecols=(0, 1))
+ax.plot(R0_5[:, 0] , R0_5[:, 1] , color=ggplot[4], marker='+', ls=lines[4], lw=2, ms=7, mew=1.5, label=r'$R_0 = 5$')
+R0_10 = np.loadtxt(fname='R0_10.dat', usecols=(0, 1))
+ax.plot(R0_10[:, 0], R0_10[:, 1], color=ggplot[3], marker='+', ls=lines[3], lw=2, ms=7, mew=1.5, label=r'$R_0 = 10$')
+R0_20 = np.loadtxt(fname='R0_20.dat', usecols=(0, 1))
+ax.plot(R0_20[:, 0], R0_20[:, 1], color=ggplot[2], marker='+', ls=lines[2], lw=2, ms=7, mew=1.5, label=r'$R_0 = 20$')
+R0_40 = np.loadtxt(fname='R0_40.dat', usecols=(0, 1))
+ax.plot(R0_40[:, 0], R0_40[:, 1], color=ggplot[1], marker='+', ls=lines[1], lw=2, ms=7, mew=1.5, label=r'$R_0 = 40$')
+R0_80 = np.loadtxt(fname='R0_80.dat', usecols=(0, 1))
+ax.plot(R0_80[:, 0], R0_80[:, 1], color=ggplot[0], marker='+', ls=lines[0], lw=2, ms=7, mew=1.5, label=r'$R_0 = 80$')
+
+ax.set_ylabel(r'$\mu\,\,[\mathrm{D}]$')
+ax.legend(loc='upper right', ncol=1)
+
[…]

### 2016-08-11 gfx/spherical/width.py

@@ -1,0 +1,37 @@
+#!/usr/bin/env python
+
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+import numpy as np
+
+mpl.style.use('ggplot')
+# The colors are from the ggplot style, consistent with grey background
+ggplot = ['#E24A33', '#348ABD', '#988ED5', '#777777', '#FBC15E', '#8EBA42', '#FFB5B8']
+# List of line styles
+lines = ['-', '--', ':', '-.', '-', '--', ':']
+params = {'text.usetex': False,
+          'mathtext.fontset': 'stixsans',
+          'font.size' : 14
+         }
+mpl.rcParams.update(params)
+
+fig, ax = plt.subplots(figsize=(10, 5), facecolor='w', edgecolor='k')
+
+# Read in data
+w5_c20  = np.loadtxt(fname='dipole_acetone_perpendicular_w5_c20.dat',  usecols=(0, 1))
+w10_c20 = np.loadtxt(fname='dipole_acetone_perpendicular_w10_c20.dat', usecols=(1,))
+w20_c20 = np.loadtxt(fname='dipole_acetone_perpendicular_w20_c20.dat', usecols=(1,))
+z = w5_c20[:, 0] - 20.0
+
+ax.plot(z, w5_c20[:, 1], color=ggplot[0], marker='+', ls=lines[0], lw=2, ms=7, mew=1.5, label=r'$w = 5$')
+ax.plot(z, w10_c20[:], color=ggplot[1], marker='+', ls=lines[1], lw=2, ms=7, mew=1.5, label=r'$w = 10$')
+ax.plot(z, w20_c20[:], color=ggplot[2], marker='+', ls=lines[2], lw=2, ms=7, mew=1.5, label=r'$w = 20$')
+ax.set_ylabel(r'$\mu\,\,[\mathrm{D}]$')
+ax.legend(loc='upper right', ncol=1)
+
+ax.set_xlim(-10.0, 25.0)
+ax.set_ylim(3.0, 4.4)
+ax.set_xlabel(r'$z\,\,[\AA]$')
+
+fig.savefig('width.png', transparent=False, bbox_inches='tight')

### 2016-08-11 gfx/wemlin/alpha_convergence.py

@@ -85,10 +85,10 @@ ax2.set_xticklabels(labels)
 l4 = ax2.plot((1.5, 5.5), (limit, limit), label='Limit', ls='-.', color=ggplot[3])
 
 ax.yaxis.set_major_formatter(y_formatter)
-ax.set_ylabel(r'Isotropic polarizability $[a_0^3]$')
-
 ax.grid(False)
 ax2.grid(False)
+ax.yaxis.grid(True)
+ax.set_ylabel(r'Isotropic polarizability $[a_0^3]$')
 
 lines = l1 + l2 + l3 + l4
 labels = [l.get_label() for l in lines]

### 2016-08-15 Chapters/Ch03-PCMCoupledCluster.tex
Not much...
Save work on CC chapter

@@
 % mainfile: ../RobertoDiRemigioPhDThesis.tex
-\chapter{Coupled cluster}\label{ch:PCMCC}
+%************************************************
+\chapter{Treating solvation within correlated wave function models}\label{ch:PCMCC}
+
+\epigraph{Non temerai i terrori della notte non temerai il terrore}{--- \textsc{CCCP}}
+
+In this Chapter, I will summarize previous and current work on the
+formulation of quantum/classical polarizable models in the context of
+the highly accurate \acl{CC} model.
+This is an ongoing and yet unpublished collaboration with prof.~T.~Daniel Crawford
+(Virginia Tech) and Dr.~Andrew C.~Simmonett (NIH)
+The first part of this Chapter summarizes existing work, mainly by Cammi
+and Caricato, by recasting it in language of variational
+quantum/classical polarizable Hamiltonians.
+The second part of this Chapter illustrates how the variational language
+facilitates the formulation of a \acl{CC}-based \acl{PT} (\acs{CC}-\acs{PT})
+including a classical polarizable environment.
+I will finally show how the perturbation series can be successfully
+exploited to derive iterative and noniterative approximate \acs{CC}
+schemes.
+
+\pagebreak
 
@@
   \begin{pmatrix}
+   \sigma^\mathrm{tot} \\
+   \kappa^\mathrm{tot}
+  \end{pmatrix}
+  =
+  \begin{pmatrix}
+   \sigma^\mathrm{HF} \\
+   \kappa^\mathrm{HF}
+  \end{pmatrix}
+  +
+  \begin{pmatrix}
    \sigma \\
    \kappa
-  \end{pmatrix}
-  =
-  \begin{pmatrix}
-   \sigma^\mathrm{HF} \\
-   \kappa^\mathrm{HF}
-  \end{pmatrix}
-  +
-  \begin{pmatrix}
-   \sigma^\mathrm{CC} \\
-   \kappa^\mathrm{CC}
   \end{pmatrix}
@@
   U_\mathrm{pol} &=
-  \frac{1}{2}(\sigma^\mathrm{HF}+\sigma^\mathrm{CC})\PCM(\sigma^\mathrm{HF}+\sigma^\mathrm{CC})
-+ (\sigma^\mathrm{HF}+\sigma^\mathrm{CC})(\esp^\mathrm{HF}+\tilde{\esp}(\tamp{},\tbar{})) \\
-&+ \frac{1}{2}(\kappa^\mathrm{HF}+\kappa^\mathrm{CC})\MM(\kappa^\mathrm{HF}+\kappa^\mathrm{CC})
-+ (\kappa^\mathrm{HF}+\kappa^\mathrm{CC})(\zeta^\mathrm{HF}+\tilde{\zeta}(\tamp{},\tbar{})) \\
-&+ (\sigma^\mathrm{HF}+\sigma^\mathrm{CC})\bi{X}(\kappa^\mathrm{HF}+\kappa^\mathrm{CC}) \\
+  \frac{1}{2}(\sigma^\mathrm{HF}+\sigma)\PCM(\sigma^\mathrm{HF}+\sigma)
++ (\sigma^\mathrm{HF}+\sigma)(\esp^\mathrm{HF}+\tilde{\esp}(\tamp{},\tbar{})) \\
+&+ \frac{1}{2}(\kappa^\mathrm{HF}+\kappa)\MM(\kappa^\mathrm{HF}+\kappa)
++ (\kappa^\mathrm{HF}+\kappa)(\zeta^\mathrm{HF}+\tilde{\zeta}(\tamp{},\tbar{})) \\
+&+ (\sigma^\mathrm{HF}+\sigma)\bi{X}(\kappa^\mathrm{HF}+\kappa)
+  %%%% Steps
+= \left/ \text{Expand and collect} \right/ \\
 &=
@@
   + \sigma^\mathrm{HF}\bi{X}\kappa^\mathrm{HF} \\
-&+  \frac{1}{2}\sigma^\mathrm{CC}\PCM\sigma^\mathrm{CC}
-  + \sigma^\mathrm{CC}\tilde{\esp}(\tamp{},\tbar{})
-  + \frac{1}{2}\kappa^\mathrm{CC}\MM\kappa^\mathrm{CC}
-  + \kappa^\mathrm{CC}\tilde{\zeta}(\tamp{},\tbar{})
-  + \sigma^\mathrm{CC}\bi{X}\kappa^\mathrm{CC} \\
-&+  \frac{1}{2}\sigma^\mathrm{HF}\PCM\sigma^\mathrm{CC}
+&+  \frac{1}{2}\sigma\PCM\sigma
+  + \sigma\tilde{\esp}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa\MM\kappa
+  + \kappa\tilde{\zeta}(\tamp{},\tbar{})
+  + \sigma\bi{X}\kappa \\
+&+  \frac{1}{2}\sigma^\mathrm{HF}\PCM\sigma
   + \sigma^\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{})
-  + \frac{1}{2}\kappa^\mathrm{HF}\MM\kappa^\mathrm{CC}
+  + \frac{1}{2}\kappa^\mathrm{HF}\MM\kappa
   + \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
-  + \sigma^\mathrm{HF}\bi{X}\kappa^\mathrm{CC} \\
-&+  \frac{1}{2}\sigma^\mathrm{CC}\PCM\sigma^\mathrm{HF}
-  + \sigma^\mathrm{CC}\esp^\mathrm{HF}
-  + \frac{1}{2}\kappa^\mathrm{CC}\MM\kappa^\mathrm{HF}
-  + \kappa^\mathrm{CC}\zeta^\mathrm{HF}
-  + \sigma^\mathrm{CC}\bi{X}\kappa^\mathrm{HF} \\
+  + \sigma^\mathrm{HF}\bi{X}\kappa \\
+&+  \frac{1}{2}\sigma\PCM\sigma^\mathrm{HF}
+  + \sigma\esp^\mathrm{HF}
+  + \frac{1}{2}\kappa\MM\kappa^\mathrm{HF}
+  + \kappa\zeta^\mathrm{HF}
+  + \sigma\bi{X}\kappa^\mathrm{HF} \\
+  %%%% Steps
+&= \left/
+  \frac{1}{2}\sigma^\mathrm{HF}\PCM\sigma^\mathrm{HF}
+  +
+  \frac{1}{2}\kappa^\mathrm{HF}\MM\kappa^\mathrm{HF}
+  +
+  \sigma^\mathrm{HF}\esp^\mathrm{HF}
+  +
+  \kappa^\mathrm{HF}\zeta^\mathrm{HF}
+  +
+  \sigma^\mathrm{HF}\bi{X}\kappa^\mathrm{HF}
+  =
+  \frac{1}{2}\sigma^\mathrm{HF}\esp^\mathrm{HF}
+  +
+  \frac{1}{2}\kappa^\mathrm{HF}\zeta^\mathrm{HF}
+  \right/ \\
 &=
@@
   + \frac{1}{2}\kappa^\mathrm{HF}\zeta^\mathrm{HF} \\
-&+  \frac{1}{2}\sigma^\mathrm{CC}\PCM\sigma^\mathrm{CC}
-  + \sigma^\mathrm{CC}\tilde{\esp}(\tamp{},\tbar{})
-  + \frac{1}{2}\kappa^\mathrm{CC}\MM\kappa^\mathrm{CC}
-  + \kappa^\mathrm{CC}\tilde{\zeta}(\tamp{},\tbar{})
-  + \sigma^\mathrm{CC}\bi{X}\kappa^\mathrm{CC} \\
-&+ \sigma^\mathrm{CC}(\bi{Y}\sigma^\mathrm{HF} + \bi{X}\kappa^\mathrm{HF} + \esp^\mathrm{HF}) \\
-&+ \kappa^\mathrm{CC}(\bi{W}\kappa^\mathrm{HF} + \bi{X}^\dagger\sigma^\mathrm{HF} + \zeta^\mathrm{HF}) \\
-&+ \sigma^\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{}) + \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{}) \\
+&+  \frac{1}{2}\sigma\PCM\sigma
+  + \sigma\tilde{\esp}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa\MM\kappa
+  + \kappa\tilde{\zeta}(\tamp{},\tbar{})
+  + \sigma\bi{X}\kappa \\
+&+  \frac{1}{2}\sigma^\mathrm{HF}\PCM\sigma
+  + \sigma^\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa^\mathrm{HF}\MM\kappa
+  + \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
+  + \sigma^\mathrm{HF}\bi{X}\kappa \\
+&+  \frac{1}{2}\sigma\PCM\sigma^\mathrm{HF}
+  + \sigma\esp^\mathrm{HF}
+  + \frac{1}{2}\kappa\MM\kappa^\mathrm{HF}
+  + \kappa\zeta^\mathrm{HF}
+  + \sigma\bi{X}\kappa^\mathrm{HF} \\
+  %%%% Steps
+&= \left/
+  \frac{1}{2}\sigma\PCM\sigma^\mathrm{HF}
+  =
+  \frac{1}{2}\sigma^\mathrm{HF}\PCM\sigma;\quad
+  \frac{1}{2}\kappa\MM\kappa^\mathrm{HF}
+  =
+  \frac{1}{2}\kappa^\mathrm{HF}\MM\kappa; \quad
+  \sigma^\mathrm{HF}\bi{X}\kappa
+  =
+  \kappa\bi{X}^\dagger\sigma^\mathrm{HF}
+  \right/ \\
 &=
@@
   + \frac{1}{2}\kappa^\mathrm{HF}\zeta^\mathrm{HF} \\
-&+  \frac{1}{2}\sigma^\mathrm{CC}\PCM\sigma^\mathrm{CC}
-  + \sigma^\mathrm{CC}\tilde{\esp}(\tamp{},\tbar{})
-  + \frac{1}{2}\kappa^\mathrm{CC}\MM\kappa^\mathrm{CC}
-  + \kappa^\mathrm{CC}\tilde{\zeta}(\tamp{},\tbar{})
-  + \sigma^\mathrm{CC}\bi{X}\kappa^\mathrm{CC} \\
+&+  \frac{1}{2}\sigma\PCM\sigma
+  + \sigma\tilde{\esp}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa\MM\kappa
+  + \kappa\tilde{\zeta}(\tamp{},\tbar{})
+  + \sigma\bi{X}\kappa \\
+  &+ \sigma(\cancelto{0}{\bi{Y}\sigma^\mathrm{HF} + \bi{X}\kappa^\mathrm{HF} + \esp^\mathrm{HF}})
+  + \kappa(\cancelto{0}{\bi{W}\kappa^\mathrm{HF} + \bi{X}^\dagger\sigma^\mathrm{HF} + \zeta^\mathrm{HF}}) \\
+&+ \sigma^\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{}) + \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{}) \\
+  %%%% Steps
+&= \left/
+    \frac{1}{2}\sigma^\mathrm{HF}\esp^\mathrm{HF}
+  + \frac{1}{2}\kappa^\mathrm{HF}\zeta^\mathrm{HF}
+  \coloneqq U_\mathrm{pol}^\mathrm{ref}
+  \right/ \\
+&=
+  U_\mathrm{pol}^\mathrm{ref}
+  +  \frac{1}{2}\sigma\PCM\sigma
+  + \sigma\tilde{\esp}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa\MM\kappa
+  + \kappa\tilde{\zeta}(\tamp{},\tbar{})
+  + \sigma\bi{X}\kappa \\
 &+ \sigma^\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{}) + \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
@@
 \end{equation}
-where we made use of the stationarity condition for the reference
-reaction field to eliminate a number of terms. Notice that the
-reference reaction energy still appears among the terms in the
-functional. However, these terms will not enter in the optimization of
-the \acs{CC} wave function and we will therefore drop them in the
-following, out of notational convenience:
+Notice that the reference reaction energy still appears among the terms
+in the functional. Obviously, these terms will not enter in the optimization of
+the \acs{CC} wave function.
+Introducing the supermatrix formalism, further simplifies the
+polarization functional:
 \begin{equation}
@@
   + \sigma^\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{}) +
-  \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{}) \\
-  &= \frac{1}{2}\p^\dagger\V\p + \p^\dagger\tilde{\s}(\tamp{},\tbar{})
-  + \p^\mathrm{HF}\tilde{\s}(\tamp{},\tbar{})
+  \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
+  + U_\mathrm{pol}^\mathrm{ref}
+  \\
+  &=
+  \frac{1}{2}\p^\dagger\V\p + \p^\dagger\tilde{\s}(\tamp{},\tbar{})
+  + \p^{\mathrm{HF}, \dagger}\tilde{\s}(\tamp{},\tbar{})
+  + U_\mathrm{pol}^\mathrm{ref}
 \end{aligned}
@@
   \begin{aligned}
-  \lag{M}(\tamp{}, \tbar{}, \sigma) &=
+  \lag{M}(\tamp{}, \tbar{}, \p) &=
   \braket{\mathrm{HF} | \overline{H_0} | \mathrm{HF}}
   + \sum_{i=1}^{\mathcal{M}}\braket{\tamp{i} | \overline{H_0} | \mathrm{HF}} \\
-  &+ \frac{1}{2}\sigma\PCM\sigma
-  + \sigma\tilde{\esp}(\tamp{}, \tbar{})
-  + \frac{1}{2}\kappa\MM\kappa
-  + \kappa\tilde{\zeta}(\tamp{}, \tbar{})
-  + \sigma\bi{X}\kappa
-  + \sigma^\mathrm{HF}\tilde{\esp}(\tamp{}, \tbar{})
-  + \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{}, \tbar{})
+  &+
+  \frac{1}{2}\p^\dagger\V\p + \p^\dagger\tilde{\s}(\tamp{},\tbar{})
+  + \p^{\mathrm{HF}, \dagger}\tilde{\s}(\tamp{},\tbar{})
+  + U_\mathrm{pol}^\mathrm{ref}
   \end{aligned}
@@
 In accordance with previous work,\autocite{Cammi2009-gu, Caricato2011-tx} the
-Hamiltonian $H_0$ is augmented with the terms
-$\sigma^\mathrm{HF}\tilde{\esp}(\tamp{}, \tbar{})$ and
-$\kappa^\mathrm{HF}\tilde{\zeta}(\tamp{}, \tbar{})$
-part appearing in the previous functional yielding the \acl{PTE} (\acs{PTE}) Hamiltonian --
-$H$ -- and
+Hamiltonian $H_0$ is augmented with the
+$\p^{\mathrm{HF},\dagger}\tilde{\s}(\tamp{},\tbar{})$ term
+to yield the \acl{PTE} (\acs{PTE}) Hamiltonian -- $H$ -- and the
 eventually the \acl{PTED} (\acs{PTED}) Lagrangian:
 \begin{equation}\label{eq:pted-cc}
-  \begin{aligned}
-  \lag{M}(\tamp{}, \tbar{}, \sigma) &=
-  \braket{\mathrm{HF} | \overline{H} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}} \\
-  &+ \frac{1}{2}\sigma\PCM\sigma
-  + \sigma\tilde{\esp}(\tamp{}, \tbar{})
-  + \frac{1}{2}\kappa\MM\kappa
-  + \kappa\tilde{\zeta}(\tamp{}, \tbar{})
-  + \sigma\bi{X}\kappa \\
-  &=
-  \braket{\mathrm{HF} | \overline{H} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}}
+  \lag{M}(\tamp{}, \tbar{}, \p) =
+  {}^\mathrm{PTE}\lag{M}
   + \frac{1}{2}\p^\dagger\V\p + \p^\dagger\tilde{\s}(\tamp{},\tbar{})
-  \end{aligned}
-\end{equation}
-As noted by \citeauthor{Cammi2009-gu} for the implicit \acs{PCM} model,
-the \acs{PTE} Hamiltonian is computed in \acs{MP}-partitioned form by
-employing the "solvated" orbitals and Fock matrix:
-\begin{equation}
-  H = F + \Phi.
-\end{equation}
-$F$ is a diagonal one-electron operator and $\Phi$ is the
-fluctuation potential.
-The \acs{PTE} effective Lagrangian ${}^\mathrm{PTE}\lag{M}$ can be
-recovered by dropping the classical terms in the Lagrangian
-\eqref{eq:pted-cc}:
+\end{equation}
+where the \acs{PTE} Lagrangian has been introduced:
 \begin{equation}
@@
   + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}}
-\end{equation}
-in accordance with the derivation of \citeauthor{Cammi2009-gu}.
+  + U_\mathrm{pol}^\mathrm{ref}
+\end{equation}
+
+As noted by \citeauthor{Cammi2009-gu} for the implicit \acs{PCM} model,
+the \acs{PTE} Hamiltonian is computed in \acs{MP}-partitioned form by
+employing the ``solvated'' orbitals and Fock matrix:
+\begin{equation}
+  H = F + \Phi.
+\end{equation}
+$F$ is a diagonal one-electron operator and $\Phi$ is the
+fluctuation potential.
+The simple nature of the Fock operator lets us re-express the \acs{PTE}
+Lagrangian as follows:
+\begin{equation}
+{}^\mathrm{PTE}\lag{M}
+  =
+  G_0
+  + \sum_{i=1}^{\mathcal{M}}\tbar{\mu_i}\denom{\mu_i}\tamp{\mu_i}
+  + \braket{\mathrm{HF} | \overline{\Phi} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\Phi} | \mathrm{HF}}
+\end{equation}
+where $G_0 = E_0 + U_\mathrm{pol}^\mathrm{ref}$ is the reference free
+energy.
+
+Use of the \acs{PTE} Lagrangian yields \acs{CC} equations that are
+completely equivalent in form to those obtained \emph{in vacuo}.
+The model is also easily implemented, as it only requires access to a
+reference state optimized in the presence of the classical medium.
+The \acs{PTE} model thus offers an efficient approximation to the full
+\acs{PTED} model with a computational cost comparable to that of
+\emph{in vacuo} \acs{CC} theory.
 As shown by \citeauthor{Caricato2011-tx}, it is possible to achieve less
@@
 \end{equation}
-The source operator is a nondiagonal, one-electron
-operator and hence has particle rank $m_\s = 1$.
-
-Eventually, one can cast the source terms entering the Lagrangian as
-follows:
+
+We first expand the source terms entering the Lagrangian as follows:
 \begin{equation}
@@
   \braket{\mathrm{HF} | \BCHfirst{\s}{T} + \BCHsecond{\s}{T} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \s + \BCHfirst{\s}{T} + \BCHsecond{\esp}{T} | \mathrm{HF}}
-\end{equation}
-
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \s + \BCHfirst{\s}{T} +
+  \BCHsecond{\esp}{T} | \mathrm{HF}},
+\end{equation}
+and since $\s$ has particle rank $m_\s = 1$ we can conclude that the
+following relationships hold:
+\begin{subequations}
 \begin{align}
@@
   \braket{\tbar{i} |
-  \esp
+  \s
   | \mathrm{HF} }
@@
 \end{align}
-
+\end{subequations}
 For a general truncation level we accordingly obtain:
@@
 \end{equation}
-which shows the unique role the singles amplitudes assume when a
-semiclassical environment model is introduced, \emph{vide infra}.
 Furthermore, for the \acs{CCS} model ($\mathcal{M} = 1$):
@@
 \begin{itemize}
+  \item singles amplitudes assume a unique role when a quantum/classical
+    Hamiltonian is introduced, a role that will be more evident in the
+    perturbative treatment,
   \item double commutators involving the highest order cluster
@@
     state, \emph{e.g.}
-      $\braket{\tbar{2} | \commutator{\commutator{\s}{T_1}}{T_2} | \mathrm{HF}}$
+    $\braket{\tbar{2} | \commutator{\commutator{\s}{T_1}}{T_2} |
+    \mathrm{HF}} \neq 0$
     and
-      $\braket{\tbar{3} | \commutator{\commutator{\s}{T_1}}{T_3} | \mathrm{HF}}$
+    $\braket{\tbar{3} | \commutator{\commutator{\s}{T_1}}{T_3} |
+    \mathrm{HF}} \neq 0$
 \end{itemize}
 
-\section{The PTED-CC equations}\label{sec:pted-cc-equations}
+\section{A family of PCM-CC models}\label{sec:pcm-cc-models}
 
@@
 with respect to the variational parameters $\tamp{\mu_i}$,
-$\tbar{\mu_i}$, $\sigma$ and $\kappa$ yields the \acs{PTED}-\acs{CC}
+$\tbar{\mu_i}$ and $\p$ yields the \acs{PTED}-\acs{CC}
 equations:
-\begin{equation}
-  \begin{aligned}
-   \pderiv{\lag{M}}{\tbar{\mu_i}} &=
+\begin{subequations}
+  \begin{align}
+   \tampEq{\mu_i}(\tamp{}, \tbar{}, \p)  &=
    \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
-    + \sigma\braket{\mu_i | \overline{\esp} | \mathrm{HF}}
-    + \kappa\braket{\mu_i | \overline{\zeta} | \mathrm{HF}}
-    = \tampEq{\mu_i}(\tamp{}, \tbar{}, \sigma, \kappa) \\
-    \pderiv{\lag{M}}{\tamp{\mu_i}} &=
+   + \p^\dagger\braket{\mu_i | \overline{\s} | \mathrm{HF}}
+             = 0 \label{eq:pted-cc-amplitudes}\\
+   \tbarEq{\mu_i}(\tamp{}, \tbar{}, \p)
+    &=
     \denom{\mu_i}\tbar{\mu_i} +
     \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
-    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} \\
-    &+ \sigma\lbrace
-    \braket{\mathrm{HF} | \commutator{\esp}{\cluster{\mu_i}} +
-    \commutator{\commutator{\esp}{T}}{\cluster{\mu_i}}| \mathrm{HF}} +
-    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\overline{\esp}}{\cluster{\mu_i}} | \mathrm{HF}}
-             \rbrace
-             \\
-    &+ \kappa\lbrace
-    \braket{\mathrm{HF} | \commutator{\zeta}{\cluster{\mu_i}} +
-    \commutator{\commutator{\zeta}{T}}{\cluster{\mu_i}}| \mathrm{HF}} +
-    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\overline{\zeta}}{\cluster{\mu_i}} | \mathrm{HF}}
-             \rbrace
-    = \tbarEq{\mu_i}(\tamp{}, \tbar{}, \sigma, \kappa) \\
-    \pderiv{\lag{M}}{\sigma} &= \PCM\sigma + \MMPCM\kappa
-    + \tilde{\esp}(\tamp{},\tbar{}) =
-    \Omega_\sigma(\tamp{}, \tbar{}, \sigma, \kappa) \\
-    \pderiv{\lag{M}}{\kappa} &= \MM\kappa + \MMPCM^\dagger\sigma + \tilde{\zeta}(\tamp{},\tbar{}) =
-    \Omega_\kappa(\tamp{}, \tbar{}, \sigma, \kappa)
-  \end{aligned}
-\end{equation}
-Given the evident similarities between the implicit and explicit terms
-in the amplitudes and multipliers equations, we adopt the supermatrix
-formalism for the classical polarizable terms. The equations are thus
-recast as:
-\begin{equation}
-  \begin{aligned}
-   \pderiv{\lag{M}}{\tbar{\mu_i}} &=
-   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
-   + \p^\dagger\braket{\mu_i | \overline{\s} | \mathrm{HF}}
-    = \tampEq{\mu_i}(\tamp{}, \tbar{}, \p) \\
-    \pderiv{\lag{M}}{\tamp{\mu_i}} &=
-    \denom{\mu_i}\tbar{\mu_i} +
-    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
-    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} \\
+    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} |
+    \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}}
+    \nonumber \\
     &+ \p^\dagger\lbrace
@@
              \rbrace
-    = \tbarEq{\mu_i}(\tamp{}, \tbar{}, \p) \\
-    \pderiv{\lag{M}}{{\p^{\dagger}}} &=
-    \V\p + \tilde{\s}(\tamp{}, \tbar{}) =
+             = 0 \label{eq:pted-cc-multipliers}\\
     \Omega_\p(\tamp{}, \tbar{}, \p)
-  \end{aligned}
-\end{equation}
-
-\section{Approximate Effective Lagrangians}\label{sec:approximate-lagrangians}
-
-We have already noted that the leading terms in the correlated source operator expectation
+    &=
+    \V\p + \tilde{\s}(\tamp{}, \tbar{}) = 0
+    \label{eq:pted-cc-polarization}
+  \end{align}
+\end{subequations}
+These equations clearly show:
+\begin{enumerate}
+  \item The $\tilde{\s}(\tamp{}, \tbar{})$ expectation value term in the
+    polarization equation couples the amplitudes, multipliers and
+    polarization equations.
+    This requires a proper macroiteration/microiteration
+    self-consistency scheme for its implementation.~\autocite{Cammi2009-gu, Caricato2010-hx}
+  \item As a consequence, a single-point \acs{PTED}-\acs{CC} will
+    suffer from a $2\times$ prefactor in its computational cost with
+    respect to an \emph{in vacuo} \acs{CC} calculation.
+  \item Eq.~\eqref{eq:pted-cc-polarization} is \emph{directly} coupled to
+    the multipliers equation, Eq.~\eqref{eq:pted-cc-multipliers}, since
+    the source terms directly depends on amplitudes and multipliers, but
+    only \emph{indirectly} coupled to the amplitudes equation,
+    Eq.~\eqref{eq:pted-cc-amplitudes}.
+\end{enumerate}
+
+One might exploit point 3 above to reduce the prefactor of the method to
+that of an \emph{in vacuo} \acs{CC} calculation.
+The simplest approach is to neglect altogether the polarization equation
+and the polarization degree of freedom. This is equivalent to
+differentiating the polarization-independent \acs{PTE} effective
+Lagrangian:
+\begin{subequations}
+  \begin{align}
+   \tampEq{\mu_i}(\tamp{}, \tbar{}, \p)  &=
+   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
+    = 0 \\
+   \tbarEq{\mu_i}(\tamp{}, \tbar{}, \p)
+    &=
+    \denom{\mu_i}\tbar{\mu_i} +
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
+    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} |
+    \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}}
+    = 0
+  \end{align}
+\end{subequations}
+The \acs{PTE} scheme naturally preserves the scaling of the underlying \acs{CC}
+method.
+The \acs{PTE} scheme is advantageous since no coupling between the
+\acs{CC} amplitude and multiplier equations is introduced. However, the
+polarization included in the correlation treatment is the same as for
+the reference determinant, an approximation which appears questionable
+from the point of view of classical electrostatics.
+There is, in fact, no relaxation of the reference reaction
+field due to the correlated description of the electronic density.
+
+The leading terms in the correlated source operator expectation
 value, Eq.~\eqref{eq:source-term}, prominently involve the singles
@@
 schemes: the \acs{PTES} and \acs{PTE(S)} schemes.
-The \acs{PTE} scheme is advantageous since no coupling between the
-\acs{CC} amplitude and multiplier equations is introduced. However, the
-polarization included in the correlation treatment is the same as for
-the reference determinant, an approximation which appears questionable
-from the point of view of classical electrostatics. On the contrary, the
-\acs{PTED} scheme describes the fully equilibrated electrostatic
-picture, at the expense of coupling the amplitude and multiplier
-equations.
-However, if we truncate the correlated source operator expectation value
+If we truncate the correlated source operator expectation value
 to the first term in Eq.~\eqref{eq:source-term}, the coupling between
@@
 which uncouples all equations:
-\begin{equation}
-  \begin{aligned}
-   \pderiv{\lag{M}}{\tbar{\mu_i}} &=
-   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
-   =
-   ^{\mathrm{PTE(S)}}\tampEq{\mu_i}(\tamp{}, \tbar{}, \p) \\
-    \pderiv{\lag{M}}{\tamp{\mu_1}} &=
-    \denom{\mu_1}\tbar{\mu_1} +
+\begin{subequations}
+  \begin{align}
+   &\denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
+   = 0 \\
+    &\denom{\mu_1}\tbar{\mu_1} +
     \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}} +
@@
     \braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_1}} | \mathrm{HF}}
-    =
-    ^{\mathrm{PTE(S)}}\tbarEq{\mu_1}(\tamp{}, \tbar{}, \p) \\
-    \pderiv{\lag{M}}{\tamp{\mu_i}} &=
-    \denom{\mu_i}\tbar{\mu_i} +
+    = 0 \\
+    &\denom{\mu_i}\tbar{\mu_i} +
     \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
@@
     \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}}
-    =
-    ^{\mathrm{PTE(S)}}\tbarEq{\mu_i}(\tamp{}, \tbar{}, \p) \quad \forall i \neq 1 \\
-    \pderiv{\lag{M}}{{\p^{\dagger}}} &=
-    \V\p + \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}} =
-    ^{\mathrm{PTE(S)}}\Omega_\p(\tamp{}, \tbar{}, \p)
-  \end{aligned}
-\end{equation}
+    = 0  \quad \forall i \neq 1 \\
+    &\V\p + \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}} = 0
+  \end{align}
+\end{subequations}
 The amplitude equations are formally equivalent to those derived in the
@@
 Only when a first-order molecular property, such as the molecular
-gradient, is needed will the multipliers equations need to be solved.
-
-Summarizing:
-In the \acs{PTE} scheme, the effective Lagrangian has the same form as
-the \emph{in vacuo} case. However, the Hamiltonian includes the
-reference reaction field. The amplitudes and multipliers equations
-are formally identical to the \emph{in vacuo} case. Neither
-computational prefactor nor scaling are affected, since there is no
-coupling in between the two and the reaction field is fixed.
-In the \acs{PTES(S)} scheme, the source term in the effective Lagrangian
-is restricted to the $T_1$-dependent term only. The amplitudes
-equations are formally identical to the \acs{PTE} case and are
-uncoupled from the multipliers and the polarization equations.
-The multipliers equations are formally equivalent to the \acs{PTE}
-case, but for the singles multipliers, which feature a
-polarization-dependent term.
+gradient, is needed, will the multipliers equations need to be solved.
 In other words, only the $T_1$-dependent part of the \acs{CC} density is
 used to define the classical sources.
-In the \acs{PTES} scheme, the multipliers equations are formally
-equivalent to the \acs{PTED} case. However, the
-polarization-dependent term is calculated by only retaining the
-$T_1$-dependent term in the source. This retains the coupling between the
-amplitudes and polarization equations, but avoid coupling the
-multipliers equation.
-As for the \acs{PTE(S)} scheme, only the $T_1$-dependent part of the
-\acs{CC} density is in fact used to define the classical sources.
-In the \acs{PTED} scheme, \ldots
 
@@
 
-\section{Coupled cluster perturbation theory for a quantum/classical
-  polarizable Hamiltonian}\label{sec:ccpt-polarizable}
+\section[Coupled Cluster Perturbation Theory]{
+Coupled Cluster Perturbation Theory for a Quantum/Classical Polarizable Hamiltonian}\label{sec:ccpt-polarizable}
 
@@
 quantum/classical polarizable Hamiltonian is used.
-The fluctuation potential $\Phi$ is the perturbation. As such $\Phi =
-O(1)$, while the cluster operators will be of first, at least, and higher
-order in the fluctuation potential.
-Orders will be counted assuming an Hartree--Fock closed-shell reference:
+The fluctuation potential $\Phi$ is assumed as the perturbation, hence
+it will be considered of order 1: $\Phi = O(1)$.
+Orders will be counted assuming closed-shell \acs{HF} reference:
 we will use square brackets $[i]$ for an $i$-th order contributions.
+
+The structure of the stationarity conditions
+Eqs.~\eqref{eq:pted-cc-amplitudes} and \eqref{eq:pted-cc-multipliers},
+already shows that $\tamp{\mu_i}^{[0]} = 0, \quad \forall i$ and
+$\tbar{\mu_i}^{[0]} = 0, \quad \forall i$.
 
@@
 
-The source operator $\s$ will
-be considered as \emph{zeroth-order} in the perturbation. The order of
-the contributions to $\tilde{\s}(\tamp{},\tbar{})$ will be solely
-determined by the amplitudes and multipliers.
-It is already clear that $\tilde{\s}(\tamp{},\tbar{})$ is at least first
-order in the fluctuation potential:
+The source operator $\s$ will be considered as \emph{zeroth-order} in
+the perturbation.
+The order of the contributions to $\tilde{\s}(\tamp{},\tbar{})$ will be
+solely determined by the amplitudes and multipliers.
+From the structure of \eqref{eq:pted-cc-polarization}, it is already
+clear that $\tilde{\s}(\tamp{},\tbar{})$ is at least first order in the
+fluctuation potential:
 \begin{equation}

### 2016-08-15 Chapters/Ch04-PaperSummary.tex
Save work on CC chapter

@@
 \refstepcounter{dummy}
-\addcontentsline{toc}{section}{\tocEntry{Continuum solvation in the relativistic regime}}
+\addcontentsline{toc}{section}{Continuum solvation in the relativistic regime}
 \section*{Paper I. Continuum solvation in the relativistic regime}
@@
 \refstepcounter{dummy}
-\addcontentsline{toc}{section}{\tocEntry{The wavelet Galerkin boundary element method for PCM}}
+\addcontentsline{toc}{section}{The wavelet Galerkin boundary element method for PCM}
 \section*{Paper II. The wavelet Galerkin boundary element method for PCM}
@@
 \refstepcounter{dummy}
-\addcontentsline{toc}{section}{\tocEntry{Continuum modelling of non homogeneous environments}}
-\section*{Paper III: Continuum modelling of non homogeneous environments}
+\addcontentsline{toc}{section}{Non homogeneous environments}
+\section*{Paper III: Non homogeneous environments}
 
@@
 \refstepcounter{dummy}
-\addcontentsline{toc}{section}{\tocEntry{Relativistic calculation of EPR and pNMR parameters in solution}}
+\addcontentsline{toc}{section}{Relativistic calculation of EPR and pNMR parameters in solution}
 \section*{Paper IV: Relativistic calculation of EPR and pNMR parameters in solution}
@@
 \refstepcounter{dummy}
-\addcontentsline{toc}{section}{\tocEntry{Open-ended self-consistent field response theory in solution}}
+\addcontentsline{toc}{section}{Open-ended self-consistent field response theory in solution}
 \section*{Paper V: Open-ended self-consistent field response theory in solution}

### 2016-08-18 Chapters/Ch03-PCMCoupledCluster.tex
Save the day's work. Energy corrections up to fifth order re-checked and typed in

@@
 (Virginia Tech) and Dr.~Andrew C.~Simmonett (NIH)
+
 The first part of this Chapter summarizes existing work, mainly by Cammi
@@
 quantum/classical polarizable Hamiltonians.
+
 The second part of this Chapter illustrates how the variational language
@@
 including a classical polarizable environment.
+
 I will finally show how the perturbation series can be successfully
@@
   \begin{pmatrix}
-    \PCM & \MMPCM^\dagger \\
-    \MMPCM & \MM
-  \end{pmatrix}
-  \begin{pmatrix}
-   \sigma^\mathrm{HF} \\
-   \kappa^\mathrm{HF}
+    \PCM & \MMPCM \\
+    \MMPCM^\dagger & \MM
+  \end{pmatrix}
+  \begin{pmatrix}
+   \sigma_\mathrm{HF} \\
+   \kappa_\mathrm{HF}
   \end{pmatrix}
@@
   \begin{pmatrix}
-   \esp^\mathrm{HF} \\
-   \zeta^\mathrm{HF}
+   \esp_\mathrm{HF} \\
+   \zeta_\mathrm{HF}
   \end{pmatrix}
@@
   \begin{pmatrix}
-   \sigma^\mathrm{HF} \\
-   \kappa^\mathrm{HF}
+   \sigma_\mathrm{HF} \\
+   \kappa_\mathrm{HF}
   \end{pmatrix}
@@
   \begin{pmatrix}
-    \braket{\mathrm{HF} | \overline{\esp} - \esp^\mathrm{HF} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} - \esp^\mathrm{HF} | \mathrm{HF}} \\
-  \braket{\mathrm{HF} | \overline{\zeta} - \zeta^\mathrm{HF} | \mathrm{HF}}
+    \braket{\mathrm{HF} | \overline{\esp} - \esp_\mathrm{HF} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} - \esp_\mathrm{HF} | \mathrm{HF}} \\
+  \braket{\mathrm{HF} | \overline{\zeta} - \zeta_\mathrm{HF} | \mathrm{HF}}
   + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} -
-  \zeta^\mathrm{HF}| \mathrm{HF}}
+  \zeta_\mathrm{HF}| \mathrm{HF}}
   \end{pmatrix}
@@
   U_\mathrm{pol} &=
-  \frac{1}{2}(\sigma^\mathrm{HF}+\sigma)\PCM(\sigma^\mathrm{HF}+\sigma)
-+ (\sigma^\mathrm{HF}+\sigma)(\esp^\mathrm{HF}+\tilde{\esp}(\tamp{},\tbar{})) \\
-&+ \frac{1}{2}(\kappa^\mathrm{HF}+\kappa)\MM(\kappa^\mathrm{HF}+\kappa)
-+ (\kappa^\mathrm{HF}+\kappa)(\zeta^\mathrm{HF}+\tilde{\zeta}(\tamp{},\tbar{})) \\
-&+ (\sigma^\mathrm{HF}+\sigma)\bi{X}(\kappa^\mathrm{HF}+\kappa)
+  \frac{1}{2}(\sigma_\mathrm{HF}+\sigma)\PCM(\sigma_\mathrm{HF}+\sigma)
++ (\sigma_\mathrm{HF}+\sigma)(\esp_\mathrm{HF}+\tilde{\esp}(\tamp{},\tbar{})) \\
+&+ \frac{1}{2}(\kappa_\mathrm{HF}+\kappa)\MM(\kappa_\mathrm{HF}+\kappa)
++ (\kappa_\mathrm{HF}+\kappa)(\zeta_\mathrm{HF}+\tilde{\zeta}(\tamp{},\tbar{}))
++ (\sigma_\mathrm{HF}+\sigma)\bi{X}(\kappa_\mathrm{HF}+\kappa) \\
   %%%% Steps
-= \left/ \text{Expand and collect} \right/ \\
+&= \left/ \text{Expand and collect} \right/ \\
 &=
-    \frac{1}{2}\sigma^\mathrm{HF}\PCM\sigma^\mathrm{HF}
-  + \sigma^\mathrm{HF}\esp^\mathrm{HF}
-  + \frac{1}{2}\kappa^\mathrm{HF}\MM\kappa^\mathrm{HF}
-  + \kappa^\mathrm{HF}\zeta^\mathrm{HF}
-  + \sigma^\mathrm{HF}\bi{X}\kappa^\mathrm{HF} \\
+    \frac{1}{2}\sigma_\mathrm{HF}\PCM\sigma_\mathrm{HF}
+  + \sigma_\mathrm{HF}\esp_\mathrm{HF}
+  + \frac{1}{2}\kappa_\mathrm{HF}\MM\kappa_\mathrm{HF}
+  + \kappa_\mathrm{HF}\zeta_\mathrm{HF}
+  + \sigma_\mathrm{HF}\bi{X}\kappa_\mathrm{HF} \\
 &+  \frac{1}{2}\sigma\PCM\sigma
@@
   + \sigma\bi{X}\kappa \\
-&+  \frac{1}{2}\sigma^\mathrm{HF}\PCM\sigma
-  + \sigma^\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{})
-  + \frac{1}{2}\kappa^\mathrm{HF}\MM\kappa
-  + \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
-  + \sigma^\mathrm{HF}\bi{X}\kappa \\
-&+  \frac{1}{2}\sigma\PCM\sigma^\mathrm{HF}
-  + \sigma\esp^\mathrm{HF}
-  + \frac{1}{2}\kappa\MM\kappa^\mathrm{HF}
-  + \kappa\zeta^\mathrm{HF}
-  + \sigma\bi{X}\kappa^\mathrm{HF} \\
+&+  \frac{1}{2}\sigma_\mathrm{HF}\PCM\sigma
+  + \sigma_\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa_\mathrm{HF}\MM\kappa
+  + \kappa_\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
+  + \sigma_\mathrm{HF}\bi{X}\kappa \\
+&+  \frac{1}{2}\sigma\PCM\sigma_\mathrm{HF}
+  + \sigma\esp_\mathrm{HF}
+  + \frac{1}{2}\kappa\MM\kappa_\mathrm{HF}
+  + \kappa\zeta_\mathrm{HF}
+  + \sigma\bi{X}\kappa_\mathrm{HF} \\
   %%%% Steps
 &= \left/
-  \frac{1}{2}\sigma^\mathrm{HF}\PCM\sigma^\mathrm{HF}
-  +
-  \frac{1}{2}\kappa^\mathrm{HF}\MM\kappa^\mathrm{HF}
-  +
-  \sigma^\mathrm{HF}\esp^\mathrm{HF}
-  +
-  \kappa^\mathrm{HF}\zeta^\mathrm{HF}
-  +
-  \sigma^\mathrm{HF}\bi{X}\kappa^\mathrm{HF}
-  =
-  \frac{1}{2}\sigma^\mathrm{HF}\esp^\mathrm{HF}
-  +
-  \frac{1}{2}\kappa^\mathrm{HF}\zeta^\mathrm{HF}
+  \frac{1}{2}\sigma_\mathrm{HF}\PCM\sigma_\mathrm{HF}
+  +
+  \frac{1}{2}\kappa_\mathrm{HF}\MM\kappa_\mathrm{HF}
+  +
+  \sigma_\mathrm{HF}\esp_\mathrm{HF}
+  +
+  \kappa_\mathrm{HF}\zeta_\mathrm{HF}
+  +
+  \sigma_\mathrm{HF}\bi{X}\kappa_\mathrm{HF}
+  =
+  \frac{1}{2}\sigma_\mathrm{HF}\esp_\mathrm{HF}
+  +
+  \frac{1}{2}\kappa_\mathrm{HF}\zeta_\mathrm{HF}
   \right/ \\
 &=
-    \frac{1}{2}\sigma^\mathrm{HF}\esp^\mathrm{HF}
-  + \frac{1}{2}\kappa^\mathrm{HF}\zeta^\mathrm{HF} \\
+    \frac{1}{2}\sigma_\mathrm{HF}\esp_\mathrm{HF}
+  + \frac{1}{2}\kappa_\mathrm{HF}\zeta_\mathrm{HF} \\
 &+  \frac{1}{2}\sigma\PCM\sigma
@@
   + \sigma\bi{X}\kappa \\
-&+  \frac{1}{2}\sigma^\mathrm{HF}\PCM\sigma
-  + \sigma^\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{})
-  + \frac{1}{2}\kappa^\mathrm{HF}\MM\kappa
-  + \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
-  + \sigma^\mathrm{HF}\bi{X}\kappa \\
-&+  \frac{1}{2}\sigma\PCM\sigma^\mathrm{HF}
-  + \sigma\esp^\mathrm{HF}
-  + \frac{1}{2}\kappa\MM\kappa^\mathrm{HF}
-  + \kappa\zeta^\mathrm{HF}
-  + \sigma\bi{X}\kappa^\mathrm{HF} \\
+&+  \frac{1}{2}\sigma_\mathrm{HF}\PCM\sigma
+  + \sigma_\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa_\mathrm{HF}\MM\kappa
+  + \kappa_\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
+  + \sigma_\mathrm{HF}\bi{X}\kappa \\
+&+  \frac{1}{2}\sigma\PCM\sigma_\mathrm{HF}
+  + \sigma\esp_\mathrm{HF}
+  + \frac{1}{2}\kappa\MM\kappa_\mathrm{HF}
+  + \kappa\zeta_\mathrm{HF}
+  + \sigma\bi{X}\kappa_\mathrm{HF} \\
   %%%% Steps
 &= \left/
-  \frac{1}{2}\sigma\PCM\sigma^\mathrm{HF}
-  =
-  \frac{1}{2}\sigma^\mathrm{HF}\PCM\sigma;\quad
-  \frac{1}{2}\kappa\MM\kappa^\mathrm{HF}
-  =
-  \frac{1}{2}\kappa^\mathrm{HF}\MM\kappa; \quad
-  \sigma^\mathrm{HF}\bi{X}\kappa
-  =
-  \kappa\bi{X}^\dagger\sigma^\mathrm{HF}
+  \frac{1}{2}\sigma\PCM\sigma_\mathrm{HF}
+  =
+  \frac{1}{2}\sigma_\mathrm{HF}\PCM\sigma;\quad
+  \frac{1}{2}\kappa\MM\kappa_\mathrm{HF}
+  =
+  \frac{1}{2}\kappa_\mathrm{HF}\MM\kappa; \quad
+  \sigma_\mathrm{HF}\bi{X}\kappa
+  =
+  \kappa\bi{X}^\dagger\sigma_\mathrm{HF}
   \right/ \\
 &=
-    \frac{1}{2}\sigma^\mathrm{HF}\esp^\mathrm{HF}
-  + \frac{1}{2}\kappa^\mathrm{HF}\zeta^\mathrm{HF} \\
+    \frac{1}{2}\sigma_\mathrm{HF}\esp_\mathrm{HF}
+  + \frac{1}{2}\kappa_\mathrm{HF}\zeta_\mathrm{HF} \\
 &+  \frac{1}{2}\sigma\PCM\sigma
@@
   + \sigma\bi{X}\kappa \\
-  &+ \sigma(\cancelto{0}{\bi{Y}\sigma^\mathrm{HF} + \bi{X}\kappa^\mathrm{HF} + \esp^\mathrm{HF}})
-  + \kappa(\cancelto{0}{\bi{W}\kappa^\mathrm{HF} + \bi{X}^\dagger\sigma^\mathrm{HF} + \zeta^\mathrm{HF}}) \\
-&+ \sigma^\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{}) + \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{}) \\
+  &+ \sigma(\cancelto{0}{\bi{Y}\sigma_\mathrm{HF} + \bi{X}\kappa_\mathrm{HF} + \esp_\mathrm{HF}})
+  + \kappa(\cancelto{0}{\bi{W}\kappa_\mathrm{HF} + \bi{X}^\dagger\sigma_\mathrm{HF} + \zeta_\mathrm{HF}}) \\
+&+ \sigma_\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{}) + \kappa_\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{}) \\
   %%%% Steps
 &= \left/
-    \frac{1}{2}\sigma^\mathrm{HF}\esp^\mathrm{HF}
-  + \frac{1}{2}\kappa^\mathrm{HF}\zeta^\mathrm{HF}
+    \frac{1}{2}\sigma_\mathrm{HF}\esp_\mathrm{HF}
+  + \frac{1}{2}\kappa_\mathrm{HF}\zeta_\mathrm{HF}
   \coloneqq U_\mathrm{pol}^\mathrm{ref}
@@
   + \sigma\bi{X}\kappa \\
-&+ \sigma^\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{}) + \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
+&+ \sigma_\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{}) + \kappa_\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
   \end{aligned}
@@
   + \sigma\bi{X}\kappa
-  + \sigma^\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{}) +
-  \kappa^\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
+  + \sigma_\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{}) +
+  \kappa_\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
   + U_\mathrm{pol}^\mathrm{ref}
@@
   &=
-  \frac{1}{2}\p^\dagger\V\p + \p^\dagger\tilde{\s}(\tamp{},\tbar{})
-  + \p^{\mathrm{HF}, \dagger}\tilde{\s}(\tamp{},\tbar{})
+  \frac{1}{2}{}^t\p\V\p + {}^t\p\tilde{\s}(\tamp{},\tbar{})
+  + {}^t\p_\mathrm{HF}\tilde{\s}(\tamp{},\tbar{})
   + U_\mathrm{pol}^\mathrm{ref}
@@
   &+
-  \frac{1}{2}\p^\dagger\V\p + \p^\dagger\tilde{\s}(\tamp{},\tbar{})
-  + \p^{\mathrm{HF}, \dagger}\tilde{\s}(\tamp{},\tbar{})
+  \frac{1}{2}{}^t\p\V\p + {}^t\p\tilde{\s}(\tamp{},\tbar{})
+  + {}^t\p_\mathrm{HF}\tilde{\s}(\tamp{},\tbar{})
   + U_\mathrm{pol}^\mathrm{ref}
@@
 Hamiltonian $H_0$ is augmented with the
-$\p^{\mathrm{HF},\dagger}\tilde{\s}(\tamp{},\tbar{})$ term
+${}^t\p_\mathrm{HF}\tilde{\s}(\tamp{},\tbar{})$ term
 to yield the \acl{PTE} (\acs{PTE}) Hamiltonian -- $H$ -- and the
@@
   {}^\mathrm{PTE}\lag{M}
-  + \frac{1}{2}\p^\dagger\V\p + \p^\dagger\tilde{\s}(\tamp{},\tbar{})
+  + \frac{1}{2}{}^t\p\V\p + {}^t\p\tilde{\s}(\tamp{},\tbar{})
 \end{equation}
@@
   | \mathrm{HF}}
-  &\quad\quad\text{nonzero contributions} \quad t + u = 1 \\
+  &\quad\quad\text{nonzero contributions} \quad t + u = 1
+  \label{eq:HF-double-commutator} \\
   \braket{\tbar{i} |
@@
 \end{equation}
+By virtue of Eq.~\eqref{eq:HF-double-commutator} the \acs{HF}
+expectation value of the double commutator was dropped.
+
 Furthermore, for the \acs{CCS} model ($\mathcal{M} = 1$):
@@
    \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
-   + \p^\dagger\braket{\mu_i | \overline{\s} | \mathrm{HF}}
+   + {}^t\p\braket{\mu_i | \s | \mathrm{HF}}\delta_{\mu_i\mu_1} \nonumber \\
+   &+ {}^t\p\braket{\mu_i | \BCHfirst{\s}{T} + \BCHsecond{\s}{T} | \mathrm{HF}}
              = 0 \label{eq:pted-cc-amplitudes}\\
@@
     \nonumber \\
-    &+ \p^\dagger\lbrace
-    \braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_i}} +
-    \commutator{\commutator{\s}{T}}{\cluster{\mu_i}}| \mathrm{HF}} +
-    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\overline{\s}}{\cluster{\mu_i}} | \mathrm{HF}}
-             \rbrace
+    &+
+    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}\delta_{\mu_i\mu_1} \nonumber \\
+    &+
+    {}^t\p\sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}
+    +
+    {}^t\p\sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\commutator{\s}{T}}{\cluster{\mu_i}} | \mathrm{HF}}
              = 0 \label{eq:pted-cc-multipliers}\\
@@
   + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}}
-  + \frac{1}{2}\p^\dagger\V\p + \p^\dagger
+  + \frac{1}{2}{}^t\p\V\p + {}^t\p
   \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
@@
     \commutator{\overline{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}}
-    + \p^\dagger
+    + {}^t\p
     \braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_1}} | \mathrm{HF}}
@@
 determining the singles multipliers which is augmented by an additional term:
-$\p^\dagger\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_1}} |
+${}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_1}} |
 \mathrm{HF}}$
@@
 it will be considered of order 1: $\Phi = O(1)$.
-Orders will be counted assuming closed-shell \acs{HF} reference:
+Orders will be counted assuming a closed-shell \acs{HF} reference:
 we will use square brackets $[i]$ for an $i$-th order contributions.
+We further assume that real wave functions are used.
 
@@
 \paragraph*{Amplitudes}
-\begin{equation}
-  \Omega_{\mu_i}^{[1]} = \denom{\mu_i}\tamp{\mu_i}^{[1]}
-  + \braket{\mu_i | \Phi | \mathrm{HF}}
-  + \p^{\dagger\,[1]}\braket{\mu_i | \bar{\s}^{[0]} | \mathrm{HF}}
-  + \p^{\dagger\,[0]}\braket{\mu_i | \bar{\s}^{[1]}  | \mathrm{HF}}
-  = 0
-\end{equation}
-And since $\p^{[0]} = 0$ and $\bar{\s}^{[0]} = \s$:
 \begin{equation}\label{eq:1st-order-amp}
@@
   + \braket{\mu_i | \Phi | \mathrm{HF}}
-  + \p^{\dagger\,[1]}\braket{\mu_i | \s | \mathrm{HF}}
+  + {}^t\p^{[1]}\braket{\mu_i | \s | \mathrm{HF}}\delta_{\mu_i\mu_1}
   = 0
 \end{equation}
-Considering the singles and doubles excitation manifolds:
+Considering the singles, doubles and triples excitation manifolds:
 \begin{subequations}
@@
   \Omega_{\mu_1}^{[1]} &= \denom{\mu_1}\tamp{\mu_1}^{[1]}
-  + \p^{\dagger\,[1]}\braket{\mu_1 | \s | \mathrm{HF}}
+  + {}^t\p^{[1]}\braket{\mu_1 | \s | \mathrm{HF}}
   = 0 \\
@@
   + \braket{\mu_2 | \Phi | \mathrm{HF}}
-  = 0
+  = 0 \\
+  \Omega_{\mu_3}^{[1]} &= \denom{\mu_3}\tamp{\mu_3}^{[1]} = 0
   \end{align}
@@
 \paragraph*{Multipliers}
+\begin{equation}%\label{eq:1st-order-mult}
+  \bar{\Omega}_{\mu_i}^{[1]} =
+    \denom{\mu_i}\tbar{\mu_i}^{[1]}
+    + \braket{\mathrm{HF} | \commutator{\Phi}{\cluster{\mu_i}} | \mathrm{HF}}
+    + {}^t\p^{[1]}\braket{\mathrm{HF} |
+    \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}\delta_{\mu_i\mu_1} =0
+\end{equation}
+and expanding the commutators:
 \begin{equation}\label{eq:1st-order-mult}
@@
     \denom{\mu_i}\tbar{\mu_i}^{[1]}
-      + \braket{\mathrm{HF} | \commutator{\Phi}{\cluster{\mu_i}} | \mathrm{HF}}
-      + \p^{\dagger\,[1]}\braket{\mathrm{HF} |
-      \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}} =0
-  \end{equation}
-Considering the singles and doubles excitation manifolds:
-\begin{subequations}
-  \begin{align}
-    \bar{\Omega}_{\mu_1}^{[1]} &= \denom{\mu_1}\tbar{\mu_1}^{[1]}
-      + \p^{\dagger\,[1]}\braket{\mathrm{HF} |
-      \commutator{\s}{\cluster{\mu_1}} | \mathrm{HF}}
-  = 0 \\
-    \bar{\Omega}_{\mu_2}^{[1]} &= \denom{\mu_2}\tbar{\mu_2}^{[1]}
-      + \braket{\mathrm{HF} | \commutator{\Phi}{\cluster{\mu_2}} | \mathrm{HF}}
-  = 0
-  \end{align}
-\end{subequations}
-Thus, despite the closed-shell \acs{HF} reference, the singles
+    + \braket{\mathrm{HF} | \Phi | \mu_i}
+    + {}^t\p^{[1]}\braket{\mathrm{HF} | \s | \mu_i}\delta_{\mu_i\mu_1} =0
+\end{equation}
+This clearly shows that Eq.~\eqref{eq:1st-order-amp}
+and Eq.~\eqref{eq:1st-order-mult} are complex conjugates. Under the
+assumption of real wave functions we can also conclude that:
+\begin{equation}
+  \tamp{\mu_i}^{[1]} = \tbar{\mu_i}^{[1]},\,\,\forall i
+\end{equation}
+Thus, as for the amplitudes, despite the closed-shell \acs{HF} reference, the singles
 multipliers will already appear in first order, due to the
 quantum/classical coupling.
+Moreover, as is the case \emph{in vacuo}, there are no triples contributions to
+the first order amplitudes and multipliers: $\tamp{\mu_3}^{[1]} = \tbar{\mu_3}^{[1]} = 0,\,\,\forall \mu$.
+
 \paragraph*{Polarization}
-  \begin{equation}\label{eq:1st-order-pol}
+To first order, the polarization degree of freedom will be given as:
+\begin{equation}%\label{eq:1st-order-pol}
   \V\p^{[1]} + \braket{\mathrm{HF} | \commutator{\s}{T_1^{[1]}}| \mathrm{HF}}
   + \braket{\tbar{1}^{[1]} | \s | \mathrm{HF} } = 0
+\end{equation}
+Under the assumption of real wave functions, the equivalence of
+first order amplitudes and multipliers yields:
+\begin{equation}
+  \begin{aligned}
+  \braket{\mathrm{HF} | \commutator{\s}{T_1^{[1]}}| \mathrm{HF}}
+  %%%% Steps
+  &= \left/ T_1^{[1]} = \sum_{\mu_1}\ket{\mu_1}\tamp{\mu_1} \right/
+  =
+  \sum_{\mu_1}\braket{\mathrm{HF}|\s|\mu_1}\tamp{\mu_1} \\
+  %%%% Steps
+  &= \left/ \text{Complex conjugation and real wave functions} \right/ \\
+  &= \sum_{\mu_1}\braket{\mu_1 | \s | \mathrm{HF}}\tbar{\mu_1}
+  = \braket{\tbar{\mu_1}^{[1]} | \s | \mathrm{HF}}
+ \end{aligned}
+\end{equation}
+so that the first order polarization equation becomes:
+\begin{equation}\label{eq:1st-order-pol}
+  \V\p^{[1]} + 2\braket{\mathrm{HF} | \commutator{\s}{T_1^{[1]}}| \mathrm{HF}} = 0
 \end{equation}
@@
   + \braket{\mu_i | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
-  + \p^{\dagger\,[2]}\braket{\mu_i | \s | \mathrm{HF}}
-  + \p^{\dagger\,[1]}\braket{\mu_i | \commutator{\s}{T^{[1]}} |
+  + {}^t\p^{[2]}\braket{\mu_i | \s | \mathrm{HF}}\delta_{\mu_i\mu_1}
+  + {}^t\p^{[1]}\braket{\mu_i | \commutator{\s}{T^{[1]}} |
   \mathrm{HF}}
@@
   + \braket{\mu_1 | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
-  + \p^{\dagger\,[2]}\braket{\mu_1 | \s | \mathrm{HF}}
-  + \p^{\dagger\,[1]}\braket{\mu_1 | \commutator{\s}{T^{[1]}} |
+  + {}^t\p^{[2]}\braket{\mu_1 | \s | \mathrm{HF}}
+  + {}^t\p^{[1]}\braket{\mu_1 | \commutator{\s}{T^{[1]}} |
   \mathrm{HF}}
@@
   + \braket{\mu_2 | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
-  + \p^{\dagger\,[1]}\braket{\mu_2 | \commutator{\s}{T_2^{[1]}} |
+  + {}^t\p^{[1]}\braket{\mu_2 | \commutator{\s}{T_2^{[1]}} |
   \mathrm{HF}}
@@
 \end{subequations}
-The triples first appear to second order and are not \emph{directly}
+As already noted elsewhere~\autocite{Koch1997-nm, Helgaker2000-tz},
+the second term in Eq.~\eqref{eq:2nd-order-amp} can involve no higher
+than triple excitations.
+Moreover, the triples first appear to second order and are not \emph{directly}
 affected by the quantum/classical coupling.
+
 \paragraph*{Multipliers}
-  \begin{equation}\label{eq:2nd-order-mult}
+\begin{equation}
   \begin{aligned}
@@
     \commutator{\Phi}{\cluster{\mu_i}} | \mathrm{HF}} \\
-      &+ \p^{\dagger\,[2]}\braket{\mathrm{HF} |
-      \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}
-      + \p^{\dagger\,[1]}
+    &+ {}^t\p^{[2]}\braket{\mathrm{HF} |
+    \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}\delta_{\mu_i\mu_1}
+    + {}^t\p^{[1]}
     \sum_j\braket{\tbar{j}^{[1]} |
@@
 \end{equation}
+The double commutator term is null thus:~\autocite{Koch1997-nm, Helgaker2000-tz}
+\begin{equation}\label{eq:2nd-order-mult}
+  \begin{aligned}
+  \bar{\Omega}_{\mu_i}^{[2]} &=
+    \denom{\mu_i}\tbar{\mu_i}^{[2]}
+    + \sum_j\braket{\tbar{j}^{[1]} |
+    \commutator{\Phi}{\cluster{\mu_i}} | \mathrm{HF}} \\
+    &+ {}^t\p^{[2]}\braket{\mathrm{HF} |
+    \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}\delta_{\mu_i\mu_1}
+    + {}^t\p^{[1]}
+    \sum_j\braket{\tbar{j}^{[1]} |
+    \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}
+    =0
+  \end{aligned}
+\end{equation}
+\citeauthor{Koch1997-nm} showed that the \emph{in vacuo} second order
+amplitudes and multipliers equations are the complex conjugates of each
+other for the singles, doubles and tripled manifold. Under the further
+assumption of real wave functions thus $\tamp{\mu_i}^{[2]} =
+\tbar{\mu_i}^{[2]},\,\,\forall i = 1 - 3$.
+A similar conclusion holds also for the quantum/classical polarizable
+Hamiltonians here considered.
+First of all, the terms involving the second order polarization degree of
+freedom are identical:
+\begin{equation}
+    {}^t\p^{[2]}\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}\delta_{\mu_i\mu_1}
+    =
+    {}^t\p^{[2]}\braket{\mathrm{HF} | \s|\mu_i}\delta_{\mu_i\mu_1}
+    =
+    {}^t\p^{[2]}\braket{\mu_i| \s|\mathrm{HF}}\delta_{\mu_i\mu_1}
+\end{equation}
+since they are complex conjugates and real wave functions are assumed.
+Under the same set of assumptions, the first order terms are also identical:
+\begin{equation}
+  \begin{aligned}
+  {}^t\p^{[1]} \sum_j\braket{\tbar{j}^{[1]} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}
+  %%%% Steps
+  &= \left/ \sum_j\bra{\tbar{j}^{[1]}} = \sum_j\sum_{\mu_j}\tbar{\mu_j}\bra{\mu_j}  \right/ \\
+  &=
+  {}^t\p^{[1]} \sum_j\sum_{\mu_j}\tbar{\mu_j}^{[1]} \braket{\mu_j| \s | \mu_i}
+  %%%% Steps
+  = \left/ \text{Complex conjugation and real wave functions} \right/ \\
+  &=
+  {}^t\p^{[1]} \sum_j\sum_{\mu_j} \braket{\mu_i| \s | \mu_j} \tamp{\mu_j}^{[1]}
+  %%%% Steps
+  = \left/ \sum_j\sum_{\mu_j}\tamp{\mu_j}\ket{\mu_j}  = T^{[1]}\right/ \\
+  &=
+  {}^t\p^{[1]}\braket{\mu_i | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
+ \end{aligned}
+\end{equation}
+
 Considering the singles, doubles and triples excitation manifolds:
@@
       \commutator{\Phi}{\cluster{\mu_1}} | \mathrm{HF}} \\
-      &+ \p^{\dagger\,[2]}\braket{\mathrm{HF} |
+      &+ {}^t\p^{[2]}\braket{\mathrm{HF} |
       \commutator{\s}{\cluster{\mu_1}} | \mathrm{HF}}
-      + \p^{\dagger\,[1]}
+      + {}^t\p^{[1]}
       \braket{\tbar{1}^{[1]} |
@@
       \commutator{\Phi}{\cluster{\mu_2}} | \mathrm{HF}} \\
-      &+ \p^{\dagger\,[1]}
+      &+ {}^t\p^{[1]}
       \braket{\tbar{1}^{[1]} + \tbar{2}^{[1]} |
@@
 \paragraph*{Polarization}
-  \begin{equation}\label{eq:2nd-order-pol}
-    \begin{aligned}
-    \V\p^{[2]} &+
+\begin{equation}
+  \V\p^{[2]} +
   \braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}}| \mathrm{HF}}
   + \braket{\tbar{1}^{[2]} | \s | \mathrm{HF} }
-    + \braket{\tbar{1}^{[1]} | \commutator{\s}{T_1^{[1]}} | \mathrm{HF}}
-    \\
-    &+
-    \braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[1]}} | \mathrm{HF}}
+  + \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
+  + \braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[1]}} | \mathrm{HF}}
   = 0
-  \end{aligned}
-  \end{equation}
-No triples contributions enter directly up to second order.
+\end{equation}
+There are no explicit contributions from triples.
+Moreover, since the second order amplitudes and multipliers are
+identical up to the triples one has $\braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}}| \mathrm{HF}} = \braket{\tbar{1}^{[2]} | \s | \mathrm{HF} }$
+and thus:
+\begin{equation}\label{eq:2nd-order-pol}
+  \V\p^{[2]} +
+  2\braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}}| \mathrm{HF}}
+  + \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
+  + \braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[1]}} | \mathrm{HF}}
+  = 0
+\end{equation}
 
@@
 where $G^{[0]}$ includes the quantum/classical coupling.
-\todo[inline]{Check this statement, the Lagrangian has a zeroth-order
-term!}
-
+
+We introduce the following notation:
+\begin{equation}
+  \tilde{\s}^{[m]}(\tamp{}^{[p]}, \tbar{}^{[q]})
+\end{equation}
+for a source term of $m$-th order formed by amplitudes up to and
+including $p$-th order and by multipliers up to and including $q$-th order.
+
+The second order energy correction will be formed including first order
+amplitudes and polarization, while using zeroth order multipliers:
 \begin{equation}\label{eq:second-order-G}
@@
   G^{[2]} &= \braket{\mathrm{HF} | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
-  + \frac{1}{2}\p^{\dagger\,[1]}\V\p^{[1]} +
-  \p^{\dagger\,[1]}\tilde{\s}^{[1]} \\
-  &= E^{[2]} + \frac{1}{2}\p^{\dagger\,[1]}\V\p^{[1]} +
-  \p^{\dagger\,[1]}
-  \braket{\mathrm{HF} | \commutator{\s}{T^{[1]}_1} | \mathrm{HF}}
+  + \frac{1}{2}{}^t\p^{[1]}\V\p^{[1]} +
+  {}^t\p^{[1]}\tilde{\s}^{[1]}(\tamp{}^{[1]}, \tbar{}^{[0]}) \\
+  &= E^{[2]} +
+  \frac{1}{2}{}^t\p^{[1]}\V\p^{[1]}
+  +
+  {}^t\p^{[1]}\braket{\mathrm{HF} | \commutator{\s}{T^{[1]}_1} | \mathrm{HF}} \\
+  &= \left/
+  \V\p^{[1]} + 2\braket{\mathrm{HF} | \commutator{\s}{T_1^{[1]}}| \mathrm{HF}} = 0
+  \right/ \\
+  &= E^{[2]}
   \end{aligned}
 \end{equation}
-Notice that the source term is here assembled in accordance to the $2n+1$
-and $2n+2$ rules, while this is not the case in the equations for the
-polarization.
-
+\todo[inline]{I feel like there is some important lesson/conclusion to
+draw from this.}
+
+The third order energy correction will be formed including first order
+amplitudes, multipliers and polarization:
 \begin{equation}\label{eq:third-order-G}
@@
     G^{[3]} &= \sum_j\braket{\tbar{j}^{[1]} | \commutator{\Phi}{T^{[1]}} |
-    \mathrm{HF}} + \p^{\dagger\, [1]}\tilde{\s}^{[2]} \\
+    \mathrm{HF}} + {}^t\p^{[1]}\tilde{\s}^{[2]}(\tamp{}^{[1]}, \tbar{}^{[1]}) \\
     &= E^{[3]}
-    + \p^{\dagger\, [1]}
+    + {}^t\p^{[1]}
     \lbrace
-      \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}_1} | \mathrm{HF}}
-    + \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}_2} | \mathrm{HF}}
+      \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
     + \braket{\tbar{2}^{[1]} | \commutator{\s}{T^{[1]}_2} | \mathrm{HF}}
@@
 
+The fourth order energy corrections will be formed including second
+order amplitudes and polarization, while using first order multipliers:
 \begin{equation}\label{eq:fourth-order-G}
@@
     \mathrm{HF}} \\
-    &+\frac{1}{2}\p^{\dagger\,[2]}\V\p^{[2]}
-    + \p^{\dagger\,[2]}\tilde{\s}^{[2]}
-    + \p^{\dagger\,[1]}\tilde{\s}^{[3]} \\
-    &= E^{[4]} + \frac{1}{2}\p^{\dagger\,[2]}\V\p^{[2]}
-    +\p^{\dagger\,[2]}
+    &+\frac{1}{2}{}^t\p^{[2]}\V\p^{[2]}
+    + {}^t\p^{[2]}\tilde{\s}^{[2]}(\tamp{}^{[2]}, \tbar{}^{[1]})
+    + {}^t\p^{[1]}\tilde{\s}^{[3]}(\tamp{}^{[2]}, \tbar{}^{[1]}) \\
+    &= E^{[4]} + \frac{1}{2}{}^t\p^{[2]}\V\p^{[2]}
+    +{}^t\p^{[2]}
     \lbrace
-      \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}_1} | \mathrm{HF}}
-    + \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}_2} | \mathrm{HF}}
-    + \braket{\tbar{2}^{[1]} | \commutator{\s}{T^{[1]}_2} | \mathrm{HF}}
+    \braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}} | \mathrm{HF}}
+    + \sum_i\braket{\tbar{i}^{[1]}| \commutator{\s}{T^{[1]}} | \mathrm{HF} }
     \rbrace \\
-    &+\p^{\dagger\,[1]}
+    &+{}^t\p^{[1]}
+    \lbrace
+    \sum_i\braket{\tbar{i}^{[1]} | \commutator{\s}{T^{[2]}} | \mathrm{HF}}
+    +
+    \sum_i\frac{1}{2}\braket{\tbar{i}^{[1]} |
+    \commutator{\commutator{\s}{T^{[1]}}}{T^{[1]}}
+    |
+    \mathrm{HF}}
+    \rbrace \\
+  %%%% Steps
+  &= \left/ \text{Remove null terms} \right/ \\
+    &= E^{[4]} \\
+    &+ \frac{1}{2}{}^t\p^{[2]}\V\p^{[2]}
+    +{}^t\p^{[2]}
+    \lbrace
+    \braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}} | \mathrm{HF}}
+    + \braket{\tbar{1}^{[1]}| \commutator{\s}{T^{[1]}} | \mathrm{HF} }
+    + \braket{\tbar{2}^{[1]}| \commutator{\s}{T_2^{[1]}} | \mathrm{HF} }
+    \rbrace \\
+    &+
+    {}^t\p^{[1]}
+    \lbrace
+    \braket{\tbar{1}^{[1]} | \commutator{\s}{T_1^{[2]} + T_2^{[2]}} | \mathrm{HF}}
+    +
+    \braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[2]} + T_3^{[2]}} | \mathrm{HF}}
+    \rbrace \\
+    &+
+    {}^t\p^{[1]}
     \lbrace
     \frac{1}{2}\braket{\tbar{1}^{[1]} |
-    \commutator{\commutator{\s}{T^{[1]}_1}}{T^{[1]}_1}
+    \commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[1]}}
     |
     \mathrm{HF}}
-    + \braket{\tbar{2}^{[1]} |
-    \commutator{\commutator{\s}{T^{[1]}_1}}{T^{[1]}_2}
+    +
+    \braket{\tbar{2}^{[1]} |
+    \commutator{\commutator{\s}{T_1^{[1]}}}{T_2^{[1]}}
     |
     \mathrm{HF}}
-    \rbrace
+    \rbrace \\
+  %%%% Steps
+  &= \left/
+  \V\p^{[2]} +
+  2\braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}}| \mathrm{HF}}
+  + \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
+  + \braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[1]}} | \mathrm{HF}}
+  = 0
+  \right/ \\
+    &= E^{[4]}
+    + \frac{1}{2}{}^t\p^{[2]}\braket{\tbar{1}^{[1]}| \commutator{\s}{T^{[1]}} | \mathrm{HF} }
+    + \frac{1}{2}{}^t\p^{[2]}\braket{\tbar{2}^{[1]}| \commutator{\s}{T_2^{[1]}} | \mathrm{HF} } \\
+    &+
+    {}^t\p^{[1]}
+    \braket{\tbar{1}^{[1]} | \commutator{\s}{T_1^{[2]} + T_2^{[2]}} | \mathrm{HF}}
+    +
+    \highlight{
+    {}^t\p^{[1]}
+    \braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[2]} + T_3^{[2]}} | \mathrm{HF}}
+    }
+    \\
+    &+
+    {}^t\p^{[1]}
+    \lbrace
+    \frac{1}{2}\braket{\tbar{1}^{[1]} |
+    \commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[1]}}
+    |
+    \mathrm{HF}}
+    +
+    \braket{\tbar{2}^{[1]} |
+    \commutator{\commutator{\s}{T_1^{[1]}}}{T_2^{[1]}}
+    |
+    \mathrm{HF}}
+    \rbrace \\
   \end{aligned}
 \end{equation}
-
+\todo[inline]{There is one term involving second order triples and first
+order polarization}
+
+Eventually, the fifth order energy correction is formed including second order
+amplitudes, multipliers and polarization:
 \begin{equation}\label{eq:fifth-order-G}
@@
     | \mathrm{HF}}
-    + \p^{\dagger\,[2]}\tilde{\s}^{[3]}
-    + \p^{\dagger\,[1]}\tilde{\s}^{[4]} \\
+    + {}^t\p^{[2]}\tilde{\s}^{[3]}(\tamp{}^{[2]}, \tbar{}^{[2]})
+    + {}^t\p^{[1]}\tilde{\s}^{[4]}(\tamp{}^{[2]}, \tbar{}^{[2]}) \\
     &= E^{[5]}
-    + \p^{\dagger\,[2]}
+    + {}^t\p^{[2]}
     \lbrace
-    \frac{1}{2}\braket{\tbar{1}^{[1]} |
-    \commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[1]}}
-    | \mathrm{HF}}
-    + \braket{\tbar{2}^{[1]} |
+    \sum_i\braket{\tbar{i}^{[1]} | \commutator{\s}{T^{[2]}} | \mathrm{HF}}
+    +
+    \sum_i\braket{\tbar{i}^{[1]} |
+    \frac{1}{2}\commutator{\commutator{\s}{T^{[1]}}}{T^{[1]}}
+    | \mathrm{HF}}
+    +
+    \sum_i\braket{\tbar{i}^{[2]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
+    \rbrace \\
+    &+ {}^t\p^{[1]}
+    \lbrace
+    \sum_i\braket{\tbar{i}^{[2]} | \commutator{\s}{T^{[2]}} | \mathrm{HF}}
+    +
+    \sum_i\braket{\tbar{i}^{[2]} |
+    \frac{1}{2}\commutator{\commutator{\s}{T^{[1]}}}{T^{[1]}}
+    | \mathrm{HF}}
+    +
+    \sum_i\braket{\tbar{i}^{[1]} |
+    \commutator{\commutator{\s}{T^{[1]}}}{T^{[2]}}
+    | \mathrm{HF}}
+    \rbrace \\
+  %%%% Steps
+  &= \left/ \text{Remove null terms} \right/ \\
+    &= E^{[5]}
+    + {}^t\p^{[2]}\braket{\tbar{1}^{[1]} | \commutator{\s}{T_1^{[2]} + T_2^{[2]}} | \mathrm{HF}}
+    + \highlight{
+    {}^t\p^{[2]}\braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[2]} + T_3^{[2]}} | \mathrm{HF}}
+    } \\
+    &+ {}^t\p^{[2]}\braket{\tbar{1}^{[2]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
+    + {}^t\p^{[2]}
+    \braket{\tbar{1}^{[1]} |
+    \frac{1}{2}\commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[1]}}
+    | \mathrm{HF}}
+    + {}^t\p^{[2]}
+    \braket{\tbar{2}^{[1]} |
     \commutator{\commutator{\s}{T_1^{[1]}}}{T_2^{[1]}}
     | \mathrm{HF}}
-    \rbrace \\
-    &+ \p^{\dagger\,[1]}
-    \lbrace
+    \\
+    &+
+    {}^t\p^{[1]}
+    \braket{\tbar{1}^{[2]} | \commutator{\s}{T_1^{[2]} + T_2^{[2]}} | \mathrm{HF}}
+    +
+    \highlight{
+    {}^t\p^{[1]}
+    \braket{\tbar{2}^{[2]} | \commutator{\s}{T_2^{[2]} + T_3^{[2]}} | \mathrm{HF}}
+    }
+    +
+    \highlight{
+    {}^t\p^{[1]}
+    \braket{\tbar{3}^{[2]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
+    } \\
+    &+
+    {}^t\p^{[1]}
+    \braket{\tbar{1}^{[2]} |
+    \frac{1}{2}\commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[1]}}
+    | \mathrm{HF}}
+    +
+    {}^t\p^{[1]}
+    \braket{\tbar{2}^{[2]} |
+    \commutator{\commutator{\s}{T_1^{[1]}}}{T_2^{[1]}}
+    | \mathrm{HF}} \\
+    &+
+    \highlight{
+    {}^t\p^{[1]}
+    \braket{\tbar{3}^{[2]} |
+    \frac{1}{2}\commutator{\commutator{\s}{T_2^{[1]}}}{T_2^{[1]}}
+    | \mathrm{HF}}
+    } \\
+    &+ {}^t\p^{[1]}
     \braket{\tbar{1}^{[1]} |
-    \commutator{\commutator{\s}{T_1^{[2]}}}{T_1^{[1]}}
-    | \mathrm{HF}}
-    + \braket{\tbar{2}^{[1]} |
-    \commutator{\commutator{\s}{T_1^{[2]}}}{T_2^{[1]}}
-    | \mathrm{HF}}
-    + \braket{\tbar{2}^{[1]} |
+    \commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[2]}}
+    | \mathrm{HF}}
+    + {}^t\p^{[1]}
+    \braket{\tbar{2}^{[1]} |
     \commutator{\commutator{\s}{T_1^{[1]}}}{T_2^{[2]}}
-    | \mathrm{HF}}
-    \rbrace
+    | \mathrm{HF}} \\
+    &+ {}^t\p^{[1]}
+    \braket{\tbar{2}^{[1]} |
+    \commutator{\commutator{\s}{T_2^{[1]}}}{T_1^{[2]}}
+    | \mathrm{HF}}
   \end{aligned}

### 2016-08-19 Chapters/Ch03-PCMCoupledCluster.tex
Save today's work

@@
 \chapter{Treating solvation within correlated wave function models}\label{ch:PCMCC}
+
+\epigraph{Potius sero quam numquam}{--- \textsc{Titus Livius}, \textit{Ab Urbe Condita}}
 
@@
   + \frac{1}{2}\kappa_\mathrm{HF}\zeta_\mathrm{HF}
-  \coloneqq U_\mathrm{pol}^\mathrm{ref}
+  \coloneq U_\mathrm{pol}^\mathrm{ref}
   \right/ \\
@@
 \end{equation}
-There are no explicit contributions from triples.
+There are no explicit contributions from triples. This means that the
+\acs{CCSD} polarization is correct through fourth order in the
+fluctuation potential.
 Moreover, since the second order amplitudes and multipliers are
@@
 including $p$-th order and by multipliers up to and including $q$-th order.
+Green boxes will be put around terms that are formally equivalent to the
+vacuum expression for the energy corrections. Red boxes will appear
+around quantum/classical coupling terms that involve the triples
+manifold.
 
@@
   \begin{aligned}
-  G^{[2]} &= \braket{\mathrm{HF} | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
+  G^{[2]} &=
+  \highlight{green}{
+  \braket{\mathrm{HF} | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
+  }
   + \frac{1}{2}{}^t\p^{[1]}\V\p^{[1]} +
   {}^t\p^{[1]}\tilde{\s}^{[1]}(\tamp{}^{[1]}, \tbar{}^{[0]}) \\
-  &= E^{[2]} +
+  &= \highlight{green}{E^{[2]}}
+  +
   \frac{1}{2}{}^t\p^{[1]}\V\p^{[1]}
@@
   \right/ \\
-  &= E^{[2]}
+  &= \highlight{green}{E^{[2]}}
   \end{aligned}
@@
   \begin{aligned}
-    G^{[3]} &= \sum_j\braket{\tbar{j}^{[1]} | \commutator{\Phi}{T^{[1]}} |
-    \mathrm{HF}} + {}^t\p^{[1]}\tilde{\s}^{[2]}(\tamp{}^{[1]}, \tbar{}^{[1]}) \\
-    &= E^{[3]}
+    G^{[3]} &=
+    \highlight{green}{\sum_j\braket{\tbar{j}^{[1]} | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}}
+    + {}^t\p^{[1]}\tilde{\s}^{[2]}(\tamp{}^{[1]}, \tbar{}^{[1]}) \\
+    &=
+    \highlight{green}{E^{[3]}}
     + {}^t\p^{[1]}
@@
   \begin{aligned}
-    G^{[4]} &= \sum_j\braket{\tbar{j}^{[1]} | \commutator{\Phi}{T^{[2]}} |
-    \mathrm{HF}} +
-    \frac{1}{2}\sum_j\braket{\tbar{j}^{[1]} |
-    \commutator{\commutator{\Phi}{T^{[1]}}}{T^{[1]}} |
-    \mathrm{HF}} \\
+    G^{[4]} &=
+    \highlight{green}
+    {
+    \sum_j\braket{\tbar{j}^{[1]} | \commutator{\Phi}{T^{[2]}} | \mathrm{HF}}
+    +
+    \sum_j\braket{\tbar{j}^{[1]} |
+    \frac{1}{2}\commutator{\commutator{\Phi}{T^{[1]}}}{T^{[1]}} | \mathrm{HF}}
+    }
+    \\
     &+\frac{1}{2}{}^t\p^{[2]}\V\p^{[2]}
@@
     + {}^t\p^{[1]}\tilde{\s}^{[3]}(\tamp{}^{[2]}, \tbar{}^{[1]}) \\
-    &= E^{[4]} + \frac{1}{2}{}^t\p^{[2]}\V\p^{[2]}
+    &= \highlight{green}{E^{[4]}} + \frac{1}{2}{}^t\p^{[2]}\V\p^{[2]}
     +{}^t\p^{[2]}
@@
   &= \left/ \text{Remove null terms} \right/ \\
-    &= E^{[4]} \\
+  &= \highlight{green}{E^{[4]}} \\
     &+ \frac{1}{2}{}^t\p^{[2]}\V\p^{[2]}
@@
   \right/ \\
-    &= E^{[4]}
+  &= \highlight{green}{E^{[4]}}
     + \frac{1}{2}{}^t\p^{[2]}\braket{\tbar{1}^{[1]}| \commutator{\s}{T^{[1]}} | \mathrm{HF} }
@@
     +
-    \highlight{
+    \highlight{red}{
     {}^t\p^{[1]}
@@
     G^{[5]} &=
+    \highlight{green}
+    {
     \frac{1}{2}\braket{\mathrm{HF} |
@@
     \commutator{\commutator{\Phi}{T^{[2]}}}{T^{[1]}}
-    | \mathrm{HF}} \\
-    &+ \sum_j \braket{\tbar{j}^{[2]} |
+    | \mathrm{HF}}
+    }
+    \\
+    &+ \highlight{green}{\sum_j \braket{\tbar{j}^{[2]} |
       \commutator{\Phi}{T^{[2]}}
@@
     | \mathrm{HF}}
+    }
     + {}^t\p^{[2]}\tilde{\s}^{[3]}(\tamp{}^{[2]}, \tbar{}^{[2]})
     + {}^t\p^{[1]}\tilde{\s}^{[4]}(\tamp{}^{[2]}, \tbar{}^{[2]}) \\
-    &= E^{[5]}
+    &= \highlight{green}{E^{[5]}}
     + {}^t\p^{[2]}
@@
   &= \left/ \text{Remove null terms} \right/ \\
-    &= E^{[5]}
+  &= \highlight{green}{E^{[5]}}
     + {}^t\p^{[2]}\braket{\tbar{1}^{[1]} | \commutator{\s}{T_1^{[2]} + T_2^{[2]}} | \mathrm{HF}}
-    + \highlight{
+    + \highlight{red}{
     {}^t\p^{[2]}\braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[2]} + T_3^{[2]}} | \mathrm{HF}}
@@
     &+
-    \highlight{
+    \highlight{red}{
     {}^t\p^{[1]}
@@
 \end{equation}
+
+\section[Noniterative Triples Corrections to PCM-CCSD]{
+Noniterative Perturbative Triples Corrections to PCM-CCSD}\label{sec:ccsd-t-noniterative}
+
+The \acs{PTED}-\acs{CCSD} scheme for the quantum/classical coupling is
+only includes singles and doubles in its excitation manifold and is thus
+correct up to third order in perturbation theory:
+\begin{subequations}
+  \begin{align}
+    \tamp{\mu_i}^{*} &= \tamp{\mu_i}^{[1]} + \tamp{\mu_i}^{[2]} + O(3) \\
+    \tbar{\mu_i}^{*} &= \tbar{\mu_i}^{[1]} + \tbar{\mu_i}^{[2]} + O(3) \\
+    \p^{*} &= \p^{[1]} + \p^{[2]} + \p^{[3]} + O(4)
+  \end{align}
+\end{subequations}
+We will denote the \acs{CCSD} converged parameters with a $*$ superscript.
+
+Already in fourth order connected triples make their appearance and any
+model going beyond third order must then take their effect into account.
+We first select the fourth and fifth order contributions to the energy
+involving connected triples.
+\begin{equation}\label{eq:triples-fourth-order}
+  G_T^{[4]} =
+  \highlight{red}{\braket{\tbar{1}^{[1]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}}
+  + \braket{\tbar{2}^{[1]}| \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+  + {}^t\p^{[1]}\braket{\tbar{2}^{[1]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
+\end{equation}
+\begin{equation}\label{eq:triples-fifth-order}
+  \begin{aligned}
+  G_T^{[5]} &=
+    \highlight{red}{
+    \braket{\tbar{2}^{[1]} |
+    \commutator{\commutator{\Phi}{T_3^{[2]}}}{T_1^{[1]}}
+    | \mathrm{HF}}
+    } \\
+    &+
+    \braket{\tbar{1}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+  + \braket{\tbar{2}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+  + \braket{\tbar{3}^{[2]} | \commutator{\Phi}{T_2^{[2]}} | \mathrm{HF}}
+  + \braket{\tbar{3}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+  \\
+    &+
+    \frac{1}{2}\braket{\tbar{3}^{[2]} |
+    \commutator{\commutator{\Phi}{T_2^{[1]}}}{T_2^{[1]}}
+    | \mathrm{HF}}
+  +
+  \highlight{red}{
+    \braket{\tbar{3}^{[2]} |
+    \commutator{\commutator{\Phi}{T_1^{[1]}}}{T_2^{[1]}}
+    | \mathrm{HF}}
+   }
+   \\
+   &+
+    {}^t\p^{[2]}\braket{\tbar{2}^{[1]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
+  + {}^t\p^{[1]}\braket{\tbar{2}^{[2]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}} \\
+  &+ {}^t\p^{[1]}\braket{\tbar{3}^{[2]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
+  + {}^t\p^{[1]}\braket{\tbar{3}^{[2]} |
+    \frac{1}{2}\commutator{\commutator{\s}{T_2^{[1]}}}{T_2^{[1]}}
+    | \mathrm{HF}}
+  \end{aligned}
+\end{equation}
+Terms in red boxes are vacuum-like terms that appear due to nonzero
+first order singles amplitudes and multipliers.
+
+\subsection{The \texorpdfstring{$\Lambda\text{CCSD}[\text{T}]$}{CCSD[aT]} and \texorpdfstring{$\text{CCSD}[\text{T}]$}{CCSD[T]} models}
+
+\todo[inline]{References for the a-CCSD[T] and CCSD[T] models!!!}
+
+The [T] noniterative triples corrections are formed by adding the fourth
+order \acs{CC}\acs{PT} terms arising from connected triples on top of
+the \acs{CCSD} energy. The \emph{converged} \acs{CCSD} amplitudes,
+multipliers and polarization are however used instead of the
+perturbative quantities in Eq.~\eqref{eq:triples-fourth-order}.
+The correction term is thus improved, since the \acs{CCSD} quantities
+are determined to infinite order, within the singles and doubles space.
+The asymmetric $\Lambda\text{CCSD}[\text{T}]$ correction is then
+obtained as:
+\begin{equation}\label{eq:4-lambda-bracket-t}
+  G_{\Lambda[\text{T}]} =
+  {}^*G_{\Lambda[\text{T}]}^{[4]} =
+  \highlight{red}{\braket{\tbar{1}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}}
+  + \braket{\tbar{2}^{*}| \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
+  + {}^t\p^{*}\braket{\tbar{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
+\end{equation}
+Replacing the perturbation multipliers by converged \acs{CCSD}
+amplitudes leads to the symmetric CCSD[T] model:
+\begin{equation}\label{eq:4-bracket-t}
+  G_{[\text{T}]} =
+  {}^*G_{[\text{T}]}^{[4]} =
+  \highlight{red}{\braket{\tamp{1}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}}
+  + \braket{\tamp{2}^{*}| \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
+  + {}^t\p^{*}\braket{\tamp{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
+\end{equation}
+
+\subsection{The
+\texorpdfstring{$\Lambda\text{CCSD}(\text{T})$}{CCSD(aT)} and CCSD(T) models}
+
+\todo[inline]{References for the a-CCSD(T) and CCSD(T) models!!!}
+
+The (T) noniterative triples corrections are formed by adding some fifth
+order terms to the fourth order [T] correction given above.
+Also for the (T) corrections, the \emph{converged} \acs{CCSD} amplitudes, multipliers and polarization
+are used instead of the perturbative quantities.
+The fifth order terms involving projection of
+$\commutator{\Phi}{T_3^{*}}$ on the singles and doubles manifold are
+chosen to form the additional perturbative correction.
+
+The correction term is thus improved, since the \acs{CCSD} quantities
+are determined to infinite order, within the singles and doubles space.
+The asymmetric $\Lambda\text{CCSD}(\text{T})$ correction is then
+obtained as:
+\begin{equation}\label{eq:5-lambda-parens-t}
+ \begin{aligned}
+  G_{\Lambda(\text{T})} =
+  {}^*G_{\Lambda[\text{T}]}^{[4]} +
+  {}^*G_{\Lambda(\text{T})}^{[5]} &=
+    \braket{\tbar{1}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
+  + \braket{\tbar{2}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}} \\
+  &+ {}^t\p^{*}\braket{\tbar{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
+  + {}^t\p^{*}\braket{\tbar{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
+ \end{aligned}
+\end{equation}
+Replacing the perturbation multipliers by converged \acs{CCSD}
+amplitudes leads to the symmetric CCSD(T) model:
+\begin{equation}\label{eq:5-parens-t}
+ \begin{aligned}
+  G_{(\text{T})} = {}^*G_{[\text{T}]}^{[4]} + {}^*G_{(\text{T})}^{[5]}
+  &=
+    \braket{\tamp{1}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
+  + \braket{\tamp{2}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}} \\
+  &+ {}^t\p^{*}\braket{\tamp{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
+  + {}^t\p^{*}\braket{\tamp{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
+ \end{aligned}
+\end{equation}
+
+\subsection{MBPT Terms in the
+\texorpdfstring{$\Lambda\text{CCSD}(\text{T})$}{CCSD(aT)} and CCSD(T) models}
+
+Triples excitations enter the \acs{MBPT} equations for the singles and
+doubles amplitudes and multipliers at third order. Thus the \acs{CCSD}
+amplitudes and multipliers are correct through second order in the
+fluctuation potential, while the polarization is correct through third
+order. Moreover, amplitudes and multipliers are equal to one another.
+
+We can rework the energy correction given above so that the perturbative
+counterparts clearly appear.
+For the fourth order terms one has:
+\begin{equation}
+  {}^*G_{\Lambda[\text{T}]}^{[4]} =
+  \highlight{red}{\braket{\tbar{1}^{[1]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}}
+  + \braket{\tbar{2}^{*}| \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
+  + {}^t\p^{*}\braket{\tbar{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
+\end{equation}

### 2016-08-20 Chapters/Ch03-PCMCoupledCluster.tex
More on triples corrections...

@@
 \begin{equation}\label{eq:ptes-cc}
-  ^{\mathrm{PTE(S)}}\lag{M}(\tamp{}, \tbar{}, \p) =
+  {}^{\mathrm{PTE(S)}}\lag{M}(\tamp{}, \tbar{}, \p) =
   \braket{\mathrm{HF} | \overline{H} | \mathrm{HF}}
@@
   \begin{align}
-   &\denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
+  {}^{\mathrm{PTE(S)}}\tampEq{\mu_i}(\tamp{}, \tbar{})  &=
+   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
    = 0 \\
-    &\denom{\mu_1}\tbar{\mu_1} +
-    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}} +
-    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} |
-    \commutator{\overline{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}}
-    + {}^t\p
-    \braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_1}} | \mathrm{HF}}
-    = 0 \\
-    &\denom{\mu_i}\tbar{\mu_i} +
+   {}^{\mathrm{PTE(S)}}\tbarEq{\mu_i}(\tamp{}, \tbar{}, \p) &=
+    \denom{\mu_i}\tbar{\mu_i} +
     \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
@@
     \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}}
-    = 0  \quad \forall i \neq 1 \\
-    &\V\p + \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}} = 0
+    \nonumber \\
+    &+
+    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}\delta_{\mu_i\mu_1}
+    = 0 \\
+    {}^{\mathrm{PTE(S)}}\Omega_\p(\tamp{}, \tbar{}, \p)
+    &=
+    \V\p + \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}} = 0
   \end{align}
@@
   \Omega_{\mu_3}^{[2]} &= \denom{\mu_3}\tamp{\mu_3}^{[2]}
-  + \braket{\mu_3 | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
+  + \braket{\mu_3 | \commutator{\Phi}{T_2^{[1]}} | \mathrm{HF}}
   = 0
@@
 The double commutator term is null thus:~\autocite{Koch1997-nm, Helgaker2000-tz}
+\todo[inline]{Check this statement... I am not 100\% sure and I have
+found no derivation...}
 \begin{equation}\label{eq:2nd-order-mult}
@@
       \denom{\mu_1}\tbar{\mu_1}^{[2]}
-      + \braket{\mathrm{HF} | \commutator{\commutator{\Phi}{T^{[1]}}}{\cluster{\mu_1}} | \mathrm{HF}}
       + \sum_j\braket{\tbar{j}^{[1]} |
-      \commutator{\Phi}{\cluster{\mu_1}} | \mathrm{HF}} \\
+      \commutator{\Phi}{\cluster{\mu_1}} | \mathrm{HF}} \nonumber \\
       &+ {}^t\p^{[2]}\braket{\mathrm{HF} |
@@
       \denom{\mu_2}\tbar{\mu_2}^{[2]}
-      + \braket{\mathrm{HF} | \commutator{\commutator{\Phi}{T^{[1]}}}{\cluster{\mu_2}} | \mathrm{HF}}
       + \sum_j\braket{\tbar{j}^{[1]} |
-      \commutator{\Phi}{\cluster{\mu_2}} | \mathrm{HF}} \\
-      &+ {}^t\p^{[1]}
+      \commutator{\Phi}{\cluster{\mu_2}} | \mathrm{HF}}
+      + {}^t\p^{[1]}
       \braket{\tbar{1}^{[1]} + \tbar{2}^{[1]} |
@@
       \denom{\mu_3}\tbar{\mu_3}^{[2]}
-      + \braket{\mathrm{HF} | \commutator{\commutator{\Phi}{T^{[1]}}}{\cluster{\mu_3}} | \mathrm{HF}}
       + \sum_j\braket{\tbar{j}^{[1]} |
@@
   \begin{align}
-    \tamp{\mu_i}^{*} &= \tamp{\mu_i}^{[1]} + \tamp{\mu_i}^{[2]} + O(3) \\
-    \tbar{\mu_i}^{*} &= \tbar{\mu_i}^{[1]} + \tbar{\mu_i}^{[2]} + O(3) \\
+    \tamp{\mu_i}^{*} &= \tamp{\mu_i}^{[1]} + \tamp{\mu_i}^{[2]} + O(3)
+    \label{eq:tamp-star} \\
+    \tbar{\mu_i}^{*} &= \tbar{\mu_i}^{[1]} + \tbar{\mu_i}^{[2]} + O(3)
+    \label{eq:tbar-star} \\
     \p^{*} &= \p^{[1]} + \p^{[2]} + \p^{[3]} + O(4)
+    \label{eq:p-star}
   \end{align}
@@
 We will denote the \acs{CCSD} converged parameters with a $*$ superscript.
+\todo[inline]{Check the statement above about $\p^{*}$.}
 
@@
 first order singles amplitudes and multipliers.
+
+\todo[inline]{WARNING. I am not 100\% sure that the multipliers
+equations are complex conjugate to the amplitudes equations also in
+second order!}
 
@@
   {}^*G_{\Lambda[\text{T}]}^{[4]} +
-  {}^*G_{\Lambda(\text{T})}^{[5]} &=
-    \braket{\tbar{1}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
-  + \braket{\tbar{2}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}} \\
-  &+ {}^t\p^{*}\braket{\tbar{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
+  {}^*G_{\Lambda(\text{T})}^{[5]} =
+  \sum_{j=1}^2
+    \braket{\tbar{j}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
   + {}^t\p^{*}\braket{\tbar{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
@@
  \begin{aligned}
-  G_{(\text{T})} = {}^*G_{[\text{T}]}^{[4]} + {}^*G_{(\text{T})}^{[5]}
-  &=
-    \braket{\tamp{1}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
-  + \braket{\tamp{2}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}} \\
-  &+ {}^t\p^{*}\braket{\tamp{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
+  G_{(\text{T})} =
+  {}^*G_{[\text{T}]}^{[4]} +
+  {}^*G_{(\text{T})}^{[5]} =
+  \sum_{j=1}^2
+    \braket{\tamp{j}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
   + {}^t\p^{*}\braket{\tamp{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
@@
 counterparts clearly appear.
-For the fourth order terms one has:
-\begin{equation}
-  {}^*G_{\Lambda[\text{T}]}^{[4]} =
-  \highlight{red}{\braket{\tbar{1}^{[1]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}}
-  + \braket{\tbar{2}^{*}| \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
-  + {}^t\p^{*}\braket{\tbar{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
-\end{equation}
+First of all, we notice that $T_3^{*} = T_3^{[2]} + \tilde{T}_3$ contains higher than second order
+contributions where $\tilde{T}_3$ contains third and higher order contributions to the
+connected triples. In particular, the following holds:
+\begin{equation}
+  \denom{\mu_3}\tamp{\mu_3}^{*} =
+  - \braket{\mu_3 |\commutator{\Phi}{T_2^{1}} | \mathrm{HF}}
+  - \braket{\mu_3 | \commutator{\Phi}{T_2^[2]} | \mathrm{HF}}
+  + O(3)
+\end{equation}
+such that:
+\begin{equation}
+  \tilde{T}_3
+  = \sum_{\mu_3}\tilde{t}_{\mu_3}\cluster{\mu_3}
+  \coloneq \sum_{\mu_3}
+  \left(-\denom{\mu_3}^{-1}\braket{\mu_3 |
+  \commutator{\Phi}{T_3^{[2]}}
+  | \mathrm{HF}}\right)
+  \cluster{\mu_3}
+\end{equation}
+
+Using Eqs.~\eqref{eq:tamp-star}--\eqref{eq:p-star} one can expand the
+energy correction as follows:
+\begin{equation}
+  \begin{aligned}
+  G_{\Lambda(\text{T})} &=
+  \sum_{j=1}^2 \braket{\tbar{j}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
+  + {}^t\p^{*}\braket{\tbar{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}} \\
+  &= \left/ \text{Insert perturbative expansions of parameters} \right/ \\
+  &=
+  \sum_{j=1}^2 \braket{\tbar{j}^{[1]}+\tbar{j}^{[2]} |
+  \commutator{\Phi}{T_3^{[2]} + \tilde{T}_3} | \mathrm{HF}} \\
+  &+ ({}^t\p^{[1]} + {}^t\p^{[2]})
+  \braket{\tbar{2}^{[1]} + \tbar{2}^{[2]} | \commutator{\s}{T_3^{[2]} + \tilde{T}_3} | \mathrm{HF}} \\
+  &= \left/ \text{Expand} \right/ \\
+  &\highlight{green}{
+  \sum_{j=1}^2 \braket{\tbar{j}^{[1]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+  }
+  + \highlight{blue}{
+  \sum_{j=1}^2 \braket{\tbar{j}^{[1]} | \commutator{\Phi}{\tilde{T}_3} | \mathrm{HF}}
+  }
+  + \highlight{blue}{
+  \sum_{j=1}^2 \braket{\tbar{j}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+  } \\
+  &+
+  \highlight{green}{
+  {}^t\p^{[1]} \braket{\tbar{2}^{[1]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
+  }
+  + \highlight{blue}{
+  {}^t\p^{[1]} \braket{\tbar{2}^{[2]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
+  } \\
+  &+ \highlight{blue}{
+  {}^t\p^{[1]} \braket{\tbar{2}^{[1]} | \commutator{\s}{\tilde{T}_3} | \mathrm{HF}}
+  }
+  + \highlight{blue}{
+  {}^t\p^{[2]} \braket{\tbar{2}^{[1]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
+  }
+  + O(6) \\
+  &= \left/ \text{Use definition of $\tilde{T}_3$ and triples multipliers equation} \right/ \\
+  &\highlight{green}{
+  \sum_{j=1}^2 \braket{\tbar{j}^{[1]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+  }
+  + \highlight{blue}{
+  \braket{\tbar{3}^{[2]} | \commutator{\Phi}{T_2^{[2]}} | \mathrm{HF}}
+  }
+  + \highlight{blue}{
+  \sum_{j=1}^2 \braket{\tbar{j}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+  } \\
+  &+
+  \highlight{green}{
+  {}^t\p^{[1]} \braket{\tbar{2}^{[1]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
+  }
+  + \highlight{blue}{
+  {}^t\p^{[1]} \braket{\tbar{2}^{[2]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
+  } \\
+  &+ \highlight{blue}{
+  {}^t\p^{[1]} \braket{\tbar{2}^{[1]} | \commutator{\s}{\tilde{T}_3} | \mathrm{HF}}
+  }
+  + \highlight{blue}{
+  {}^t\p^{[2]} \braket{\tbar{2}^{[1]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
+  }
+  + O(6) \\
+  \end{aligned}
+\end{equation}
+
+\todo[inline]{Don't really know what to do with the
+${}^t\p^{[1]} \braket{\tbar{2}^{[1]} | \commutator{\s}{\tilde{T}_3} | \mathrm{HF}}$ term.
+I think it is correctly identified as $O(5)$, but that it's among the
+discarded terms.}

### 2016-08-23 Chapters/Ch01-ContinuumSolvationModels.tex
Save the day's work. Energy corrections up to fifth order re-checked and typed in
Minor additions
Save the day's work

@@
 
-\section{A unifying view of classical polarizable models}
+\section{A unifying view of classical polarizable
+models}\label{sec:variational}
 
@@
 Coupled implicit/explicit polarization energy functional:
-\begin{equation}
+\begin{equation}\label{eq:pcm-mm-functional}
   U_\mathrm{pol} =
@@
   \begin{pmatrix}
-    \PCM & \MMPCM^\dagger \\
-    \MMPCM & \MM
+    \PCM & \MMPCM \\
+    \MMPCM^\dagger & \MM
   \end{pmatrix}
@@
   U_\mathrm{pol} =
-  \frac{1}{2}\p^\dagger\V\p + \p^\dagger\s
+  \frac{1}{2}{}^t\p\V\p + {}^t\p\s
 \end{equation}
@@
   \begin{pmatrix}
-    \PCM & \MMPCM^\dagger \\
-    \MMPCM & \MM
+    \PCM & \MMPCM \\
+    \MMPCM^\dagger & \MM
   \end{pmatrix}
 \end{alignat}
+and the ${}^t\p$ symbol denotes the transposed supervector $\p$.
 The supermatrix form will prove particularly useful in the following
 derivations.
+
+\section{Perché è figo}
+
+\todo[inline]{PUT REFERENCES HERE!!!}
+The variational formulation offers a series of advantages.
+The classical energy functional:
+\begin{equation}
+  U_\mathrm{pol} =
+   \frac{1}{2}\sigma\PCM\sigma + \sigma\esp
+ + \frac{1}{2}\kappa\MM\kappa + \kappa\zeta
+ + \sigma\bi{X}\kappa
+\end{equation}
+renders itself to a straightforward physical interpretation. In the
+three bilinear terms, \ie{} the ones mediated by an interaction
+operator, we can identify the self-interaction of the classical
+polarization with itself, be it purely implicit:
+$\frac{1}{2}\sigma\PCM\sigma$, purely explicit
+$\frac{1}{2}\kappa\MM\kappa$ or mixed $\sigma\MMPCM\kappa$.
+These terms are positive definite and give rise to an unfavorable
+contribution, which is counterbalanced by the linear terms. These terms
+mediate the interaction between the induced polarization and the
+inducing fields: $\esp$ and $\zeta$. While $\esp$ is quite clearly the
+\acl{MEP} (\acs{MEP}), $\zeta$ can either be the molecular electric
+field (MMpol and PE models) or again the \acs{MEP} (FQ model).
+In any case, both will be determined by the quantum mechanical molecular
+charge density and can thus be formulated as expectation values of
+one-electron operators. Eventually, this achieves the coupling between
+the classical -- $\sigma$ and $\kappa$ -- and
+the quantum mechanical variational degrees of freedom.

### 2016-08-23 Chapters/Ch03-PCMCoupledCluster.tex
Minor additions
Save the day's work

@@
 
-In this Chapter, I will summarize previous and current work on the
-formulation of quantum/classical polarizable models in the context of
-the highly accurate \acl{CC} model.
-This is an ongoing and yet unpublished collaboration with prof.~T.~Daniel Crawford
-(Virginia Tech) and Dr.~Andrew C.~Simmonett (NIH)
-
-The first part of this Chapter summarizes existing work, mainly by Cammi
-and Caricato, by recasting it in language of variational
-quantum/classical polarizable Hamiltonians.
-
-The second part of this Chapter illustrates how the variational language
-facilitates the formulation of a \acl{CC}-based \acl{PT} (\acs{CC}-\acs{PT})
-including a classical polarizable environment.
-
-I will finally show how the perturbation series can be successfully
-exploited to derive iterative and noniterative approximate \acs{CC}
-schemes.
+I will present results on the formulation of quantum/classical
+polarizable models when the quantum part of the system includes a
+many-body description of electron correlation either by a \ac{MBPT},
+\ac{CC} or hybrid approach.
+These results leverage the variational formulation of quantum/classical
+polarizable Hamiltonians described in Section \ref{sec:variational}.
+This is an element of novelty with respect to the existing literature: I
+will show how a fully variational framework enables transparent
+derivations of \acs{MBPT} and \acs{CC} theory, encompassing previous
+work~\cite{Cammi, Olivares del Valle, Caricato, Sneskov}
+and solving some inconsistencies.~\cite{Angyan1992}
+
+The contents of this Chapter are the result of an ongoing collaboration
+with prof.~T.~Daniel Crawford (Virginia Tech) and Dr.~Andrew
+C.~Simmonett (NIH)
+
+After introducing the notation, I will derive the form of the \acs{CC}
+effective Lagrangian and proceed to show how the work of
+\citeauthor{Cammi2009-gu} and \citeauthor{Caricato2011-tx} on
+\acs{PCM}-\acs{CC} and \citeauthor{Sneskov} on \acs{PE}-\acs{CC}
+can easily be derived in the variational framework.
+
+The second part of this Chapter will be devoted to the derivation of
+\acs{MBPT} for quantum/classical polarizable Hamiltonians.
+I will exploit the \acs{CC} Ansatz for the wave function and formulate
+\acs{MBPT} in terms of the effective \acs{CC} Lagrangian, an approach
+guaranteeing term-by-term size extensivity.
+
+Iterative approximations to the \acs{CCSD} and \acs{CCSDT} models will
+be the subject of Section \ref{sec:cc2-and-cc3} where the equations
+governing the \acs{CC2} and \acs{CC3} models will be derived.
+Eventually, I will describe schemes for the noniterative inclusion of
+the effect of connected triples excitations on top of
+the \acs{CCSD} wave function. The symmetric and asymmetric [T] and (T)
+corrections will be derived.
 
@@
 
-\section{Notation}
-Cluster operator:
-\begin{equation}
-  T = \sum_{i=1}^\mathcal{M}\sum_{\mu_i}\tamp{\mu_i}\cluster{\mu_i}
-\end{equation}
-$\mu_i$ is the $\mu$-th excitation at the $i$-th
-excitation level and $\mathcal{M}$ is the truncation level.
-
-The term at the $j$-th excitation level in the left CC wavefunction:
-\begin{equation}
-  \bra{\tbar{j}} = \sum_{\mu_j}\tbar{\mu_j}\bra{\mu_j}
-\end{equation}
-
-\acl{MP} partitioning:
+\section{A Brief Summary of Coupled Cluster Theory}
+
+The analyses and derivations in this Chapter are preeminently based
+on the \acl{CC} wave function Ansatz.
+We will assume a closed-shell, \acs{HF} single reference function and
+construct our correlated wave function as an exponential mapping:
+\begin{equation}
+  \ket{\mathrm{CC}} = \expo{T}\ket{\mathrm{HF}}.
+\end{equation}
+The \emph{cluster} operator appearing in the exponential is given as:
+\begin{equation}
+  T = \sum_{p=1}^\mathcal{M} T_p = \sum_{p=1}^\mathcal{M}\sum_{\mu_p}\tamp{\mu_p}\cluster{\mu_p}
+\end{equation}
+that is, as a truncated sum of excitation operators, $\cluster{\mu_p}$,
+times the corresponding cluster amplitudes, $\tamp{\mu_p}$.
+Here $\mu_p$ is the $\mu$-th excitation at the $p$-th excitation level and
+$\mathcal{M}$ is the truncation level.
+One of the main strengths of the \acs{CC} model is its size extensivity
+which stems directly from the exponential nature of the wave
+operator.~\autocite{Shavitt, Helgaker2000-tz}
+
+The \acs{CC} method is a \emph{projective} method. Chosen the truncation
+level $\mathcal{M}$, one projects the nonrelativistic Schrödinger
+equation for this Ansatz on the excitation manifold which comprises the
+reference function and all possible excited determinants, up to the
+chosen truncation order:
+\begin{subequations}
+  \begin{align}
+    \braket{\mathrm{HF} | H_0\expo{T} | \mathrm{HF}} &= E_\mathrm{CC} \\
+    \braket{\mathrm{exc} | H_0\expo{T} | \mathrm{HF}} &=
+    E_\mathrm{CC}\braket{\mathrm{exc} | \expo{T} | \mathrm{HF}}
+  \end{align}
+\end{subequations}
+where $\bra{\mathrm{exc}}$ means any of the excited determinants in Fock
+space compatible with the excitation level truncation.
+This is the \emph{unlinked} form of the \acs{CC}
+equations.~\autocite{Helgaker2000-tz} It is usually more convenient to
+perform a \emph{similarity transformation} of the Hamiltonian operator,
+leading to the \emph{linked} form of the \acs{CC} equations:
+\begin{subequations}
+  \begin{align}
+    \braket{\mathrm{HF} | \expo{-T}H_0\expo{T} | \mathrm{HF}} &= E_\mathrm{CC} \\
+    \braket{\mathrm{exc} | \expo{-T}H_0\expo{T} | \mathrm{HF}} &= 0
+  \end{align}
+\end{subequations}
+The linked and unlinked formulations are completely equivalent. However,
+the former is size extensive term-by-term.
+Use of similarity transformed operators is ubiquitous when dealing with
+\acs{CC} theory and we introduce the following notation for it:
+\begin{equation}
+  \overline{O} =\expo{-T}O\expo{T}
+\end{equation}
+The cluster operator is not Hermitian. Hence, the similarity transformation
+will not, in general, preserve any of the symmetries, such as
+hermiticity, that the bare operators might have possessed.
+Similarity-transformed operators can be expanded in a commutator series,
+also known as the \ac{BCH} expansion:
+\begin{equation}
+  \overline{O} =
+  \sum_{n=0}^{\infty}\frac{1}{n!}[A, B]^n
+\end{equation}
+where $[A,B]^n$ is the $n$-fold nested commutator. It can be shown that
+the \acs{BCH} expansion of the similarity-transformed Hamiltonian
+\emph{exactly} truncates after the four-fold nested commutator, greatly
+simplifying algebraic derivations and manipulations.
+
+Introducing the \acl{MP} partitioning of the Hamiltonian leads to
+further insight:
 \begin{equation}
@@
 where $F$ if the Fock matrix and $\Phi$ is the fluctuation potential.
-We assume to deal with an \acl{HF} (\acs{HF}) reference function.
-
-Similarity transformed operator:
-\begin{equation}
-  \overline{O} =\expo{-T}O\expo{T}
-\end{equation}
-Since the Fock operator is diagonal, its similarity transformation
-truncates after the second term:
+Given our initial assumption on the reference function, the Fock
+operator is diagonal and expressed in terms of spin-orbital energies and
+number operators.
+Its similarity transformation truncates after the second term and has a
+relatively compact form:
 \begin{equation}
@@
 \end{equation}
-where $\denom{\mu_i}$ is the difference in orbital energy between the
+where $\denom{\mu_i}$ is the difference in orbital energies between the
 virtual and occupied spin-orbitals of excitation $i$.
 For the fluctuation potential the similarity transformation truncates
-after the fifth term, \ie{} includes up to four-fold nested commutators
-with the cluster operator.
-
-\acs{CC} expectation values are formed using the left and
-right \acs{CC} states and are denoted as:
+after the fifth term, including up to four-fold nested commutator.
+
+As formulated so far, the \acs{CC} method is not variational. It is
+however possible to introduce a fully variational Lagrangian that
+leads to the same amplitudes equations. The Lagrangian is built by
+treating the amplitude equations as constraints in the optimization of
+the \acs{CC} energy:
+\begin{equation}\label{eq:cc-lagrangian}
+  \begin{aligned}
+    \mathcal{L}^\mathcal{M}(\tamp{}, \tbar{})
+    &=
+    \braket{\mathrm{HF} | \overline{H_0} | \mathrm{HF}}
+    + \sum_{i=1}^{\mathcal{M}}\sum_{\mu_i}
+    \tbar{\mu_i}\braket{\mu_i | \overline{H_0} | \mathrm{HF}} \\
+    &=
+  E_0
+  + \sum_{i=1}^{\mathcal{M}}\tbar{\mu_i}\denom{\mu_i}\tamp{\mu_i}
+  + \braket{\mathrm{HF} | \overline{\Phi} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\Phi} | \mathrm{HF}}
+  \end{aligned}
+\end{equation}
+where the shorthand notation for the Lagrangian multiplier state was
+introduced:
+\begin{equation}
+  \bra{\tbar{j}} = \sum_{\mu_j}\tbar{\mu_j}\bra{\mu_j}
+\end{equation}
+and the \acs{MP}-partitioned form of the Hamiltonian was also exploited.
+Differentiation of the Lagrangian with respect to the multipliers
+correctly yields the amplitudes equations, while differentiation with
+respect to the amplitudes leads to the governing equations for the
+multipliers. While these are not needed for the evaluation of the
+\acs{CC} energy, their calculation is mandatory when expectation values
+and molecular properties in general are sought.
+The \acs{CC} expectation values are formed using the left and
+right \acs{CC} states and we introduce the following notation:
 \begin{equation}
   O(\tamp{}, \tbar{}) = \braket{\mathrm{HF} | \overline{O} | \mathrm{HF} }
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{j} | \overline{O} | \mathrm{HF} }
-\end{equation}
-We will also use the notation $O^\mathcal{M}(\tamp{}, \tbar{})$ to
-denote a \acs{CC} expectation value truncated at the $\mathcal{M}$-th
-excitation level.
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{O} | \mathrm{HF} }
+\end{equation}
+When required by the context, we will explicitly write at which
+truncation level in the cluster operator the expectation was evaluated:
+$O^\mathcal{M}(\tamp{}, \tbar{})$.
+
+Having summarized the essential points of \acs{CC} theory and having
+introduced the relevant notation, we can proceed to analyze its coupling
+with quantum/classical polarizable Hamiltonians.
 
@@
 
-The effective CC Lagrangian can be readily obtained by coupling the
-usual CC Lagrangian with the classical polarization functional.
+Our purpose in this Section is to derive an \emph{effective} \acs{CC}
+Lagrangian that takes into account the coupling of the correlated
+electronic state with the classical polarizable
+environment.
+This can be readily obtained~\autocite{Lipparini2016-mo} by coupling the
+usual Lagrangian in Eq.~\eqref{eq:cc-lagrangian} with the classical
+polarization functional in Eq.~\eqref{eq:pcm-mm-functional}.
 Care must however be taken in the definition of the classical
@@
 
-Let us assume that the reference state is the optimized single \acs{HF}
-determinant in the presence of the implicit/explicit environment.
-Then the corresponding variational degrees of freedom satisfy:
+Once again, let us assume that the reference state is a closed-shell
+\acs{HF} determinant. Further, we assume that the reference determinant
+was optimized in the presence of the classical polarizable environment.
+Then the variational degrees of freedom corresponding to the environment
+satisfy:
 \begin{equation}
@@
 \begin{equation}
+\begin{aligned}
   \begin{pmatrix}
@@
   \end{pmatrix}
-  =
+  &=
   \begin{pmatrix}
@@
    \tilde{\zeta}(\tamp{}, \tbar{})
-  \end{pmatrix}
-  =
+  \end{pmatrix} \\
+  &=
   \begin{pmatrix}
@@
   \end{pmatrix}
-\end{equation}
-The Baker--Campbell--Hausdorff expansions of the source terms have been
+\end{aligned}
+\end{equation}
+The \acs{BCH} expansions of the source terms have been
 rewritten as:
-\begin{equation}
-  \overline{\s} = \s + \tilde{\s} = \s + \commutator{\s}{T}
-  + \frac{1}{2}\commutator{\commutator{\s}{T}}{T}
-\end{equation}
-taking advantage of the fact that $\s$ is a one-electron operator and
-hence its commutator expansion truncates at the third term.
-We apply a shift to the correlation part of the source term
-one-electron operator to remove the reference source terms:
+\begin{subequations}
+  \begin{align}
+  \overline{\esp} &= \esp + \tilde{\esp} = \esp + \commutator{\esp}{T}
+  + \frac{1}{2}\commutator{\commutator{\esp}{T}}{T} \\
+  \overline{\kappa} &= \kappa + \tilde{\kappa} = \kappa + \commutator{\kappa}{T}
+  + \frac{1}{2}\commutator{\commutator{\kappa}{T}}{T}
+  \end{align}
+\end{subequations}
+taking advantage of the fact that both $\esp$ and $\kappa$ are
+nondiagonal, one-electron operators and hence their commutator
+expansions truncate at the third term.
+Next, we apply a shift to the correlation part of the source term one-electron
+operators to remove the reference source terms:
 \begin{equation}
@@
   \\
-  &=
+  =
   \begin{pmatrix}
@@
 operators.~\autocite{Shavitt}
-The total functional will include both reference and correlation
-components:
+
+Eventually, the polarization functional can be rewritten as:
 \begin{equation}
@@
 \end{equation}
-Notice that the reference reaction energy still appears among the terms
-in the functional. Obviously, these terms will not enter in the optimization of
-the \acs{CC} wave function.
-Introducing the supermatrix formalism, further simplifies the
-polarization functional:
+
+The reference polarization energy $U_\mathrm{pol}^\mathrm{ref}$ still
+appears among the terms in the functional but will obviously not enter
+in the optimization of the \acs{CC} wave function.
+Recasting the polarization functional in the supermatrix formalism (see
+Section \ref{sec:variational}) will simplify subsequent algebraic
+manipulations:
 \begin{equation}
@@
 \end{equation}
-In accordance with previous work,\autocite{Cammi2009-gu, Caricato2011-tx} the
-Hamiltonian $H_0$ is augmented with the
-${}^t\p_\mathrm{HF}\tilde{\s}(\tamp{},\tbar{})$ term
-to yield the \acl{PTE} (\acs{PTE}) Hamiltonian -- $H$ -- and the
-eventually the \acl{PTED} (\acs{PTED}) Lagrangian:
+In accordance with previous work,\autocite{Cammi2009-gu,
+Caricato2011-tx} the Hamiltonian $H_0$ is augmented with the
+${}^t\p_\mathrm{HF}\tilde{\s}(\tamp{},\tbar{})$ term to yield the
+\acl{PTE} (\acs{PTE}) Hamiltonian -- $H$ -- and eventually the
+\acl{PTED} (\acs{PTED}) Lagrangian:~\autocite{Olivares del Valle,
+Lipparini}
 \begin{equation}\label{eq:pted-cc}
@@
 
-Use of the \acs{PTE} Lagrangian yields \acs{CC} equations that are
-completely equivalent in form to those obtained \emph{in vacuo}.
-The model is also easily implemented, as it only requires access to a
-reference state optimized in the presence of the classical medium.
-The \acs{PTE} model thus offers an efficient approximation to the full
-\acs{PTED} model with a computational cost comparable to that of
-\emph{in vacuo} \acs{CC} theory.
-As shown by \citeauthor{Caricato2011-tx}, it is possible to achieve less
-drastic approximations of the \acs{PTED} Lagrangian. Such approximations
-can be easily derived within the current variational framework. We
-postpone this analysis to a later stage of development, \emph{vide
-infra}.
+Before delving into the details of the governing equations for
+amplitudes, multipliers and polarization, we need to take a closer look
+at the source term in classical energy functional. This is the purpose
+of the next Section.
 
 \section{Source terms in the classical energy functional}
-
-The variational formulation offers a series of advantages.
-The classical energy functional:
-\begin{equation}
-  U_\mathrm{pol} =
-   \frac{1}{2}\sigma\PCM\sigma + \sigma\esp
- + \frac{1}{2}\kappa\MM\kappa + \kappa\zeta
- + \sigma\bi{X}\kappa
-\end{equation}
-renders itself to a straightforward physical interpretation. In the
-three bilinear terms, \ie{} the ones mediated by an interaction
-operator, we can identify the self-interaction of the classical
-polarization with itself, be it purely implicit:
-$\frac{1}{2}\sigma\PCM\sigma$, purely explicit
-$\frac{1}{2}\kappa\MM\kappa$ or mixed $\sigma\MMPCM\kappa$.
-These terms are positive definite and give rise to an unfavorable
-contribution, which is counterbalanced by the linear terms. These terms
-mediate the interaction between the induced polarization and the
-inducing fields: $\esp$ and $\zeta$. While $\esp$ is quite clearly the
-\acl{MEP} (\acs{MEP}), $\zeta$ can either be the molecular electric
-field (MMpol and PE models) or again the \acs{MEP} (FQ model).
-In any case, both will be determined by the quantum mechanical molecular
-charge density and can thus be formulated as expectation values of
-one-electron operators. Eventually, this achieves the coupling between
-the classical -- $\sigma$ and $\kappa$ -- and
-the quantum mechanical variational degrees of freedom.
-
-The \acs{MEP} operator is given as:
-\begin{equation}
-  \esp = \sum_{PQ}\esp_{PQ}\cons{P}\anni{Q}
-\end{equation}
-and similarly for the \acs{MM} source operator:
-\begin{equation}
-  \zeta = \sum_{PQ}\zeta_{PQ}\cons{P}\anni{Q}
-\end{equation}
-both are nondiagonal in the MO basis. However,
-by virtue of the \emph{cluster commutation condition}, their similarity
-transformations will truncate after two
-terms:\autocite{Helgaker2000-tz}
-\begin{equation}
-  \overline{\esp} = \esp + \tilde{\esp}  = \esp + \commutator{\esp}{T} +
-  \frac{1}{2}\commutator{\commutator{\esp}{T}}{T}
-\end{equation}
-and
-\begin{equation}
-  \overline{\zeta} = \zeta + \tilde{\zeta}  = \zeta + \commutator{\zeta}{T} +
-  \frac{1}{2}\commutator{\commutator{\zeta}{T}}{T}
-\end{equation}
 
@@
 involved.
-We develop this analysis in terms of excitation ranks of the
-operators involved, based on the following:
+We develop this analysis in terms of excitation ranks of the operators
+involved, based on the following:
 \begin{equation}
@@
 is the total number of cluster operators appearing in the nested
-commutator, the $n_i$ are their respective excitations ranks and $m_O$
+commutator, $n_i$ are their respective excitations ranks and $m_O$
 is the \emph{particle rank} of the operator $O$.\autocite{Helgaker2000-tz}
@@
   + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \s + \BCHfirst{\s}{T} +
-  \BCHsecond{\esp}{T} | \mathrm{HF}},
+  \BCHsecond{\s}{T} | \mathrm{HF}},
 \end{equation}
@@
 \end{subequations}
-For a general truncation level we accordingly obtain:
+By virtue of Eq.~\eqref{eq:HF-double-commutator} the \acs{HF}
+expectation value of the double commutator can be dropped. Thus, for a
+general truncation level we obtain:
 \begin{equation}\label{eq:source-term}
@@
 \end{equation}
-By virtue of Eq.~\eqref{eq:HF-double-commutator} the \acs{HF}
-expectation value of the double commutator was dropped.
-
-Furthermore, for the \acs{CCS} model ($\mathcal{M} = 1$):
+Note that the singles amplitudes assume a unique role in the \acs{CC}
+expression of the source term.
+In addition, double commutators involving the highest order cluster
+operators can only lead to nonzero terms when $T_1$ is also involved and
+by projection onto the highest order multiplier state, \eg
+$\braket{\tbar{2} | \commutator{\commutator{\s}{T_1}}{T_2} |
+\mathrm{HF}} \neq 0$ and $\braket{\tbar{3} |
+\commutator{\commutator{\s}{T_1}}{T_3} | \mathrm{HF}} \neq 0$
+More explicitly, for the \ac{CCS} model ($\mathcal{M} = 1$) one has:
 \begin{equation}
@@
 \end{equation}
-while for the \acs{CCSD} model ($\mathcal{M} = 2$):
+while for the \ac{CCSD} model ($\mathcal{M} = 2$):
 \begin{equation}
@@
 \end{equation}
-Eventually, within the \acs{CCSDT} model ($\mathcal{M} = 3$) four more
+Eventually, within the \ac{CCSDT} model ($\mathcal{M} = 3$) four more
 terms are added:
@@
 \end{equation}
-We note that:
-\begin{itemize}
-  \item singles amplitudes assume a unique role when a quantum/classical
-    Hamiltonian is introduced, a role that will be more evident in the
-    perturbative treatment,
-  \item double commutators involving the highest order cluster
-    operators can only lead to nonzero terms when $T_1$ is also
-    involved and by projection onto the highest order left \acs{CC}
-    state, \emph{e.g.}
-    $\braket{\tbar{2} | \commutator{\commutator{\s}{T_1}}{T_2} |
-    \mathrm{HF}} \neq 0$
-    and
-    $\braket{\tbar{3} | \commutator{\commutator{\s}{T_1}}{T_3} |
-    \mathrm{HF}} \neq 0$
-\end{itemize}
-
-\section{A family of PCM-CC models}\label{sec:pcm-cc-models}
+
+\section{Governing Equations and Their Approximations}\label{sec:pcm-cc-models}
 
@@
    \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
-   + {}^t\p\braket{\mu_i | \s | \mathrm{HF}}\delta_{\mu_i\mu_1} \nonumber \\
-   &+ {}^t\p\braket{\mu_i | \BCHfirst{\s}{T} + \BCHsecond{\s}{T} | \mathrm{HF}}
+%   + {}^t\p\braket{\mu_i | \s | \mathrm{HF}}\delta_{\mu_i\mu_1} \nonumber \\
++ {}^t\p\braket{\mu_i | \overline{\s} | \mathrm{HF}}
              = 0 \label{eq:pted-cc-amplitudes}\\
@@
 \end{subequations}
-These equations clearly show:
-\begin{enumerate}
-  \item The $\tilde{\s}(\tamp{}, \tbar{})$ expectation value term in the
-    polarization equation couples the amplitudes, multipliers and
-    polarization equations.
-    This requires a proper macroiteration/microiteration
-    self-consistency scheme for its implementation.~\autocite{Cammi2009-gu, Caricato2010-hx}
-  \item As a consequence, a single-point \acs{PTED}-\acs{CC} will
-    suffer from a $2\times$ prefactor in its computational cost with
-    respect to an \emph{in vacuo} \acs{CC} calculation.
-  \item Eq.~\eqref{eq:pted-cc-polarization} is \emph{directly} coupled to
-    the multipliers equation, Eq.~\eqref{eq:pted-cc-multipliers}, since
-    the source terms directly depends on amplitudes and multipliers, but
-    only \emph{indirectly} coupled to the amplitudes equation,
-    Eq.~\eqref{eq:pted-cc-amplitudes}.
-\end{enumerate}
-
-One might exploit point 3 above to reduce the prefactor of the method to
-that of an \emph{in vacuo} \acs{CC} calculation.
-The simplest approach is to neglect altogether the polarization equation
-and the polarization degree of freedom. This is equivalent to
-differentiating the polarization-independent \acs{PTE} effective
-Lagrangian:
-\begin{subequations}
-  \begin{align}
-   \tampEq{\mu_i}(\tamp{}, \tbar{}, \p)  &=
-   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
-    = 0 \\
-   \tbarEq{\mu_i}(\tamp{}, \tbar{}, \p)
-    &=
-    \denom{\mu_i}\tbar{\mu_i} +
-    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
-    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} |
-    \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}}
-    = 0
-  \end{align}
-\end{subequations}
-The \acs{PTE} scheme naturally preserves the scaling of the underlying \acs{CC}
-method.
-The \acs{PTE} scheme is advantageous since no coupling between the
-\acs{CC} amplitude and multiplier equations is introduced. However, the
-polarization included in the correlation treatment is the same as for
-the reference determinant, an approximation which appears questionable
-from the point of view of classical electrostatics.
-There is, in fact, no relaxation of the reference reaction
-field due to the correlated description of the electronic density.
-
-The leading terms in the correlated source operator expectation
-value, Eq.~\eqref{eq:source-term}, prominently involve the singles
-cluster operator and the singles left \acs{CC} state.
-This simple fact led \citeauthor{Caricato2011-tx} to propose two
-intermediate approximations in between the \acs{PTE} and \acs{PTED}
-schemes: the \acs{PTES} and \acs{PTE(S)} schemes.
-If we truncate the correlated source operator expectation value
-to the first term in Eq.~\eqref{eq:source-term}, the coupling between
-the \acs{CC} right and left states is broken, while still improving upon the
+A coupling of the amplitudes and multipliers equations is introduced by
+the presence of the $\tilde{\s}(\tamp{}, \tbar{})$ expectation value
+in the polarization equation.
+This requires a proper macroiteration/microiteration self-consistency
+scheme for its implementation.~\autocite{Cammi2009-gu, Caricato2010-hx}
+As a consequence, a single-point \acs{PTED}-\acs{CC} will suffer from a
+$2\times$ prefactor in its computational cost with respect to an
+\emph{in vacuo} \acs{CC} calculation.
+
+It is thus convenient to devise approximations that are able to simplify the
+Lagrangian and the governing equations by breaking or weakening the coupling.
+While Eq.~\eqref{eq:pted-cc-polarization} is \emph{directly} coupled to
+the multipliers equation, since the source terms directly depends on
+amplitudes and multipliers, it is only \emph{indirectly} coupled to the
+amplitudes equation, where only the polarization enters.
+Moreover, the leading terms in the correlated source operator
+expectation value, Eq.~\eqref{eq:source-term}, prominently involve the
+singles cluster operator and the singles multiplier state.
+Three approximation have currently been proposed: \acs{PTES}, \acs{PTE(S)}
+and \acs{PTE}.
+
+\subsection*{PTES Scheme}
+
+\todo[inline]{MY CLAIM
+The first approximate scheme truncates the source term to its \acs{CCS}
+expectation value.}
+
+\subsection*{PTE(S) Scheme}
+
+If we further truncate the correlated source operator expectation value to the
+first term in Eq.~\eqref{eq:source-term}, the coupling between the
+\acs{CC} right and left states is broken, while still improving upon the
 reference reaction field with correlated contributions.
-This is the essence of the \acs{PTE(S)} approximation. The effective
-Lagrangian would then read as:
+This is the essence of the \acs{PTE(S)} approximation.~\autocite{Caricato2011-tx}
+The effective Lagrangian would then read as:
 \begin{equation}\label{eq:ptes-cc}
@@
 
-\todo[inline]{Can't really figure out how the \acs{PTES} scheme can be
-devised in this framework. The idea seems to be to preserve the form of
-the \acs{PTED} equations, but dropping the dependence on the Lagrange
-multipliers. This can be achieved by decoupling the \acs{PCM} equation
-from the multipliers equation, \ie{} by using only the part of the
-density matrix that depends on the $T_1$ amplitudes to define the source
-term. This seems to miss, however, part of the terms in the \acs{PTES}
-multipliers equations, that are, in Caricato's treatment, still
-coupled with the \acs{PCM} equation.}
+\subsection*{PTE Scheme}
+
+The simplest approach is to neglect altogether the polarization equation
+and the polarization degree of freedom. This is equivalent to
+differentiating the polarization-independent \acs{PTE} effective
+Lagrangian:
+\begin{subequations}
+  \begin{align}
+   \tampEq{\mu_i}(\tamp{}, \tbar{}, \p)  &=
+   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
+    = 0 \\
+   \tbarEq{\mu_i}(\tamp{}, \tbar{}, \p)
+    &=
+    \denom{\mu_i}\tbar{\mu_i} +
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
+    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} |
+    \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}}
+    = 0
+  \end{align}
+\end{subequations}
+The \acs{PTE} scheme naturally preserves the scaling of the underlying
+\acs{CC} method.
+The \acs{PTE} scheme is advantageous since no coupling between the
+\acs{CC} amplitude and multiplier equations is introduced.
+The model is also easily implemented, as it only requires access to a
+reference state optimized in the presence of the classical medium.
+The \acs{PTE} model thus offers an efficient approximation to the full
+\acs{PTED} model with a computational cost comparable to that of
+\emph{in vacuo} \acs{CC} theory.
+
+However, the polarization included in the correlation treatment is the
+same as for the reference determinant, an approximation which appears
+questionable from the point of view of classical electrostatics. There
+is, in fact, no relaxation of the reference reaction field due to the
+correlated description of the electronic density.
 

### 2016-08-23 acronyms.tex
Add commands, fix notations
Some more work
Beef up summary of papers
Save work on CC chapter
Save the day's work

@@
 \acro{CC}{coupled cluster}
+\acro{CCS}{coupled cluster with single substitutions}
 \acro{CCSD}{coupled cluster with single and double substitutions}
+\acro{CCSDT}{coupled cluster with single, double and triple substitutions}
+\acro{CC2}{approximate coupled cluster singles and doubles}
+\acro{CC3}{approximate coupled cluster singles, doubles and triples}
+\acro{PT}{perturbation theory}
+\acro{MBPT}{many-body perturbation theory}
 \acro{CI}{configuration interaction}
 \acro{MO}{molecular orbital}
+\acro{AO}{atomic orbital}
 \acro{LR}{linear response}
@@
 \acro{ASC}{apparent surface charge}
-\acro{PTE}{perturbation to the energy}
-\acro{PTED}{perturbation to the energy and the density}
+\acro{PTE}{perturbation-to-the-energy}
+\acro{PTES}{perturbation to the energy}
+\acro{PTE(S)}{perturbation to the energy}
+\acro{PTED}{perturbation-to-the-energy-and-the-density}
 \acro{QM/MM}{Quantum Mechanics/Molecular Mechanics}
@@
 \acro{BVP}{boundary value problem}
+\acro{BEM}{boundary element method}
+\acro{BIE}{boundary integral equation}
 \acro{PDE}{partial differential equation}
@@
 \acro{MEP}{molecular electrostatic potential}
+\acro{SCF}{self-consistent field}
 \acro{TDSCF}{time-dependent SCF}
 \acro{KS}{Kohn--Sham}
+\acro{PWC}{piecewise constant wavelet}
+\acro{PWL}{piecewise constant linear}
+\acro{DC}{Dirac--Coulomb}
+\acro{BCH}{Baker--Campbell--Hausdorff}
 \end{acronym}

### 2016-08-24 Chapters/Ch03-SolvationAndCorrelation.tex
Some organizational work

--- 2016-08-23 Chapters/Ch03-PCMCoupledCluster.tex
@@
 %************************************************
-\chapter{Treating solvation within correlated wave function models}\label{ch:PCMCC}
+\chapter{A New Perspective on Electron Correlation and Solvation}\label{ch:solvation-correlation}
 
 \epigraph{Potius sero quam numquam}{--- \textsc{Titus Livius}, \textit{Ab Urbe Condita}}
-
 \epigraph{Non temerai i terrori della notte non temerai il terrore}{--- \textsc{CCCP}}
@@
   \begin{pmatrix}
-    \tilde{\esp}(\tamp{}, \tbar{}) \\
-   \tilde{\zeta}(\tamp{}, \tbar{})
+    \esp_{\mathrm{N}}(\tamp{}, \tbar{}) \\
+    \zeta_{\mathrm{N}}(\tamp{}, \tbar{})
   \end{pmatrix} \\
@@
 nondiagonal, one-electron operators and hence their commutator
-expansions truncate at the third term.
-Next, we apply a shift to the correlation part of the source term one-electron
-operators to remove the reference source terms:
-\begin{equation}
-  \begin{aligned}
+expansions truncate at the third term. Moreover, normal-ordering of the
+operators has been introduced.~\autocite{Crawford2000-ey, Shavitt}
+This is equivalent to applying a shift to the correlation part of the
+source term one-electron operators to remove the reference source terms:
+\begin{equation}
+  \begin{aligned}
+  \begin{pmatrix}
+    \esp_{\mathrm{N}}(\tamp{}, \tbar{}) \\
+    \zeta_{\mathrm{N}}(\tamp{}, \tbar{})
+  \end{pmatrix}
+  &=
   \begin{pmatrix}
    \braket{\mathrm{HF} | \tilde{\esp} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \tilde{\esp} | \mathrm{HF}} \\
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} | \mathrm{HF}} \\
    \braket{\mathrm{HF} | \tilde{\zeta} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \tilde{\zeta} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} | \mathrm{HF}}
   \end{pmatrix}
+  \\
   &=
@@
   \end{pmatrix}
-  \\
-  =
-  \begin{pmatrix}
-   \braket{\mathrm{HF} | \tilde{\esp} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} | \mathrm{HF}} \\
-   \braket{\mathrm{HF} | \tilde{\zeta} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} | \mathrm{HF}}
-  \end{pmatrix}
-  &=
-  \begin{pmatrix}
-    \tilde{\esp}(\tamp{}, \tbar{}) \\
-    \tilde{\zeta}(\tamp{}, \tbar{})
-  \end{pmatrix}
-\end{aligned}
-\end{equation}
-which is equivalent to imposing normal ordering of the
-operators.~\autocite{Shavitt}
+\end{aligned}
+\end{equation}
 
@@
   \frac{1}{2}(\sigma_\mathrm{HF}+\sigma)\PCM(\sigma_\mathrm{HF}+\sigma)
-+ (\sigma_\mathrm{HF}+\sigma)(\esp_\mathrm{HF}+\tilde{\esp}(\tamp{},\tbar{})) \\
++ (\sigma_\mathrm{HF}+\sigma)(\esp_\mathrm{HF}+\esp_{\mathrm{N}}(\tamp{},\tbar{})) \\
 &+ \frac{1}{2}(\kappa_\mathrm{HF}+\kappa)\MM(\kappa_\mathrm{HF}+\kappa)
-+ (\kappa_\mathrm{HF}+\kappa)(\zeta_\mathrm{HF}+\tilde{\zeta}(\tamp{},\tbar{}))
++ (\kappa_\mathrm{HF}+\kappa)(\zeta_\mathrm{HF}+\zeta_{\mathrm{N}}(\tamp{},\tbar{}))
 + (\sigma_\mathrm{HF}+\sigma)\bi{X}(\kappa_\mathrm{HF}+\kappa) \\
@@
 &+  \frac{1}{2}\sigma\PCM\sigma
-  + \sigma\tilde{\esp}(\tamp{},\tbar{})
+  + \sigma\esp_{\mathrm{N}}(\tamp{},\tbar{})
   + \frac{1}{2}\kappa\MM\kappa
-  + \kappa\tilde{\zeta}(\tamp{},\tbar{})
+  + \kappa\zeta_{\mathrm{N}}(\tamp{},\tbar{})
   + \sigma\bi{X}\kappa \\
 &+  \frac{1}{2}\sigma_\mathrm{HF}\PCM\sigma
-  + \sigma_\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{})
+  + \sigma_\mathrm{HF}\esp_{\mathrm{N}}(\tamp{},\tbar{})
   + \frac{1}{2}\kappa_\mathrm{HF}\MM\kappa
-  + \kappa_\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
+  + \kappa_\mathrm{HF}\zeta_{\mathrm{N}}(\tamp{},\tbar{})
   + \sigma_\mathrm{HF}\bi{X}\kappa \\
@@
 &+  \frac{1}{2}\sigma\PCM\sigma
-  + \sigma\tilde{\esp}(\tamp{},\tbar{})
+  + \sigma\esp_{\mathrm{N}}(\tamp{},\tbar{})
   + \frac{1}{2}\kappa\MM\kappa
-  + \kappa\tilde{\zeta}(\tamp{},\tbar{})
+  + \kappa\zeta_{\mathrm{N}}(\tamp{},\tbar{})
   + \sigma\bi{X}\kappa \\
@@
   + \kappa(\cancelto{0}{\bi{W}\kappa_\mathrm{HF} + \bi{X}^\dagger\sigma_\mathrm{HF} + \zeta_\mathrm{HF}}) \\
-&+ \sigma_\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{}) + \kappa_\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{}) \\
+&+ \sigma_\mathrm{HF}\esp_{\mathrm{N}}(\tamp{},\tbar{}) + \kappa_\mathrm{HF}\zeta_{\mathrm{N}}(\tamp{},\tbar{}) \\
   %%%% Steps
@@
   +  \frac{1}{2}\sigma\PCM\sigma
-  + \sigma\tilde{\esp}(\tamp{},\tbar{})
+  + \sigma\esp_{\mathrm{N}}(\tamp{},\tbar{})
   + \frac{1}{2}\kappa\MM\kappa
-  + \kappa\tilde{\zeta}(\tamp{},\tbar{})
+  + \kappa\zeta_{\mathrm{N}}(\tamp{},\tbar{})
   + \sigma\bi{X}\kappa \\
-&+ \sigma_\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{}) + \kappa_\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
+&+ \sigma_\mathrm{HF}\esp_{\mathrm{N}}(\tamp{},\tbar{}) + \kappa_\mathrm{HF}\zeta_{\mathrm{N}}(\tamp{},\tbar{})
   \end{aligned}
@@
     \frac{1}{2}\sigma\PCM\sigma
-    + \sigma\tilde{\esp}(\tamp{},\tbar{})
+    + \sigma\esp_{\mathrm{N}}(\tamp{},\tbar{})
   + \frac{1}{2}\kappa\MM\kappa
-  + \kappa\tilde{\zeta}(\tamp{},\tbar{})
+  + \kappa\zeta_{\mathrm{N}}(\tamp{},\tbar{})
   + \sigma\bi{X}\kappa
-  + \sigma_\mathrm{HF}\tilde{\esp}(\tamp{},\tbar{}) +
-  \kappa_\mathrm{HF}\tilde{\zeta}(\tamp{},\tbar{})
+  + \sigma_\mathrm{HF}\esp_{\mathrm{N}}(\tamp{},\tbar{}) +
+  \kappa_\mathrm{HF}\zeta_{\mathrm{N}}(\tamp{},\tbar{})
   + U_\mathrm{pol}^\mathrm{ref}
@@
   &=
-  \frac{1}{2}{}^t\p\V\p + {}^t\p\tilde{\s}(\tamp{},\tbar{})
-  + {}^t\p_\mathrm{HF}\tilde{\s}(\tamp{},\tbar{})
+  \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})
+  + {}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})
   + U_\mathrm{pol}^\mathrm{ref}
@@
   &+
-  \frac{1}{2}{}^t\p\V\p + {}^t\p\tilde{\s}(\tamp{},\tbar{})
-  + {}^t\p_\mathrm{HF}\tilde{\s}(\tamp{},\tbar{})
+  \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})
+  + {}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})
   + U_\mathrm{pol}^\mathrm{ref}
@@
 Caricato2011-tx} the Hamiltonian $H_0$ is augmented with the
-${}^t\p_\mathrm{HF}\tilde{\s}(\tamp{},\tbar{})$ term to yield the
+${}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})$ term to yield the
 \acl{PTE} (\acs{PTE}) Hamiltonian -- $H$ -- and eventually the
@@
   {}^\mathrm{PTE}\lag{M}
-  + \frac{1}{2}{}^t\p\V\p + {}^t\p\tilde{\s}(\tamp{},\tbar{})
+  + \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})
 \end{equation}
@@
 \begin{equation}
-  \tilde{\s}(\tamp{}, \tbar{}) =
+  \s_{\mathrm{N}}(\tamp{}, \tbar{}) =
   \braket{\mathrm{HF} | \BCHfirst{\s}{T} + \BCHsecond{\s}{T} | \mathrm{HF}}
@@
 \begin{equation}\label{eq:source-term}
-  \tilde{\s}(\tamp{}, \tbar{}) =
+  \s_{\mathrm{N}}(\tamp{}, \tbar{}) =
   \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
@@
 \begin{equation}
-  \tilde{\s}^1(\tamp{}, \tbar{}) =
+  \begin{aligned}
+  \s_{\mathrm{N}}^1(\tamp{}, \tbar{}) &=
+  \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
+  + \braket{\tbar{1} | \check{\s} | \mathrm{HF}} \\
+  &=
   \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
@@
   + \frac{1}{2}\braket{\tbar{1} | \commutator{\commutator{\s}{T_1}}{T_1} | \mathrm{HF}}
+  \end{aligned}
 \end{equation}
@@
 \begin{equation}
-  \tilde{\s}^2(\tamp{}, \tbar{}) =
-  \tilde{\s}^1(\tamp{}, \tbar{})
+  \s_{\mathrm{N}}^2(\tamp{}, \tbar{}) =
+  \s_{\mathrm{N}}^1(\tamp{}, \tbar{})
   + \braket{\tbar{1} | \BCHfirst{\s}{T_2} | \mathrm{HF}}
@@
   \begin{aligned}
-  \tilde{\s}^3(\tamp{}, \tbar{}) &=
-  \tilde{\s}^2(\tamp{}, \tbar{})
+  \s_{\mathrm{N}}^3(\tamp{}, \tbar{}) &=
+  \s_{\mathrm{N}}^2(\tamp{}, \tbar{})
   + \braket{\tbar{2} | \BCHfirst{\s}{T_3} | \mathrm{HF}} \\
@@
    \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
-%   + {}^t\p\braket{\mu_i | \s | \mathrm{HF}}\delta_{\mu_i\mu_1} \nonumber \\
    + {}^t\p\braket{\mu_i | \overline{\s} | \mathrm{HF}}
@@
     &=
-    \V\p + \tilde{\s}(\tamp{}, \tbar{}) = 0
+    \V\p + \s_{\mathrm{N}}(\tamp{}, \tbar{}) = 0
     \label{eq:pted-cc-polarization}
@@
 A coupling of the amplitudes and multipliers equations is introduced by
-the presence of the $\tilde{\s}(\tamp{}, \tbar{})$ expectation value
+the presence of the $\s_{\mathrm{N}}(\tamp{}, \tbar{})$ expectation value
 in the polarization equation.
@@
 
-\todo[inline]{MY CLAIM
+\todo[inline]{This is my claim. I am waiting on an answer from MC to
+clarify this section.}
+
 The first approximate scheme truncates the source term to its \acs{CCS}
-expectation value.}
+expectation value, yielding the effective Lagrangian:
+\begin{equation}\label{eq:ptes-cc}
+  {}^{\mathrm{PTES}}\lag{M}(\tamp{}, \tbar{}, \p) =
+  \braket{\mathrm{HF} | \overline{H} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}}
+  + \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}^1(\tamp{}, \tbar{})
+\end{equation}
+and the corresponding governing equations:
+\begin{subequations}
+  \begin{align}
+   \tampEq{\mu_i}(\tamp{}, \tbar{}, \p)  &=
+   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
+   + {}^t\p\braket{\mu_i | \check{\s} | \mathrm{HF}}
+             = 0 \\
+   \tbarEq{\mu_i}(\tamp{}, \tbar{}, \p)
+    &=
+    \denom{\mu_i}\tbar{\mu_i} +
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
+    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} |
+    \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}}
+    \nonumber \\
+    &+
+    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}\delta_{\mu_i\mu_1} \nonumber \\
+    &+
+    {}^t\p\sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}
+    +
+    {}^t\p\sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\commutator{\s}{T}}{\cluster{\mu_i}} | \mathrm{HF}}
+             = 0 \\
+    \Omega_\p(\tamp{}, \tbar{}, \p)
+    &=
+    \V\p + \s_{\mathrm{N}}^1(\tamp{}, \tbar{}) = 0
+  \end{align}
+\end{subequations}
+These equations are, however, still coupled due to the presence of the
+multipliers in the \acs{CCS} source term in the polarization equation.
+To break the coupling, we further approximate $\s_{\mathrm{N}}^1(\tamp{}, \tbar{})$
+with its multiplier-independent part:
+\begin{equation}
+  \s_{\mathrm{N}}^1(\tamp{}, \tbar{})
+  \simeq
+  \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
+  =
+  \braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}}
+\end{equation}
+so that the governing equations for the \acs{PTES} scheme become:
+\begin{subequations}
+  \begin{align}
+   {}^{\mathrm{PTES}}\tampEq{\mu_i}(\tamp{}, \tbar{}, \p)  &=
+   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
+   + {}^t\p\braket{\mu_i | \check{\s} | \mathrm{HF}}
+             = 0 \label{eq:ptes-cc-amplitudes}\\
+   {}^{\mathrm{PTES}}\tbarEq{\mu_i}(\tamp{}, \tbar{}, \p)
+    &=
+    \denom{\mu_i}\tbar{\mu_i} +
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
+    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} |
+    \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}}
+    \nonumber \\
+    &+
+    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}\delta_{\mu_i\mu_1} \nonumber \\
+    &+
+    {}^t\p\sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}
+    +
+    {}^t\p\sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\commutator{\s}{T}}{\cluster{\mu_i}} | \mathrm{HF}}
+             = 0 \label{eq:ptes-cc-multipliers}\\
+    {}^{\mathrm{PTES}}\Omega_\p(\tamp{}, \tbar{}, \p)
+    &=
+    \V\p + \braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}} = 0
+    \label{eq:ptes-cc-polarization}
+  \end{align}
+\end{subequations}
+
+\todo[inline]{This thing looks very inconsistent with electrostatics...
+Moreover, it doesn't add up with Caricato's derivation. Maybe there are
+\emph{two different} polarization equations that are used? One when
+solving for the amplitudes and one when solving for the multipliers?}
 
@@
 
-If we further truncate the correlated source operator expectation value to the
-first term in Eq.~\eqref{eq:source-term}, the coupling between the
-\acs{CC} right and left states is broken, while still improving upon the
-reference reaction field with correlated contributions.
-This is the essence of the \acs{PTE(S)} approximation.~\autocite{Caricato2011-tx}
+The \acs{PTES} approximation can be brought one step further, by
+truncating the \acs{CCS} expectation value to its multiplier-independent
+term:
+\begin{equation}
+  \s_{\mathrm{N}}^1(\tamp{}, \tbar{})
+  \simeq
+  \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
+  =
+  \braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}}
+\end{equation}
+the coupling between the amplitudes and multipliers is broken, while
+still improving upon the reference reaction field with correlated
+contributions.~\autocite{Caricato2011-tx}
 The effective Lagrangian would then read as:
-\begin{equation}\label{eq:ptes-cc}
+\begin{equation}\label{eq:pte-s-cc}
   {}^{\mathrm{PTE(S)}}\lag{M}(\tamp{}, \tbar{}, \p) =
@@
   + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}}
-  + \frac{1}{2}{}^t\p\V\p + {}^t\p
-  \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
+  + \frac{1}{2}{}^t\p\V\p + {}^t\p\braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}}
 \end{equation}
@@
     &=
-    \V\p + \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}} = 0
+    \V\p + \braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}} = 0
   \end{align}
@@
 The amplitude equations are formally equivalent to those derived in the
-\acs{PTE} scheme. The equations for multipliers are also formally
+\acs{PTE} scheme, \emph{vide infra}. The equations for multipliers are also formally
 equivalent to the \acs{PTE} equations, apart from the equation
-determining the singles multipliers which is augmented by an additional term:
-${}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_1}} |
-\mathrm{HF}}$
+determining the singles multipliers which features an additional term.
 This term is not, however, coupled to the other equations given that
 $\p$ is fixed once the singles amplitudes have been determined.
+
 Practical implementation of the \acs{PTE(S)} scheme will solve the
 \acs{PTE}-like amplitude equation, form the source term
-$\braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}}$, solve for
+$\braket{\mathrm{HF} | \check{\s}_\mathrm{N} | \mathrm{HF}}$, solve for
 the polarization degrees of freedom and calculate the polarization
@@
 
-The simplest approach is to neglect altogether the polarization equation
-and the polarization degree of freedom. This is equivalent to
-differentiating the polarization-independent \acs{PTE} effective
-Lagrangian:
+Complete neglect of the polarization functional leads to the most
+drastic approximation to the full quantum/classical polarizable
+effective Lagrangian, the \acs{PTE} scheme.
+This is equivalent to differentiating the polarization-independent
+\acs{PTE} effective Lagrangian:
 \begin{subequations}
   \begin{align}
-   \tampEq{\mu_i}(\tamp{}, \tbar{}, \p)  &=
+ {}^{\mathrm{PTE}}\tampEq{\mu_i}(\tamp{}, \tbar{}, \p)  &=
    \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
     = 0 \\
-   \tbarEq{\mu_i}(\tamp{}, \tbar{}, \p)
+   {}^{\mathrm{PTE}}\tbarEq{\mu_i}(\tamp{}, \tbar{}, \p)
     &=
@@
 The \acs{PTE} scheme naturally preserves the scaling of the underlying
-\acs{CC} method.
-The \acs{PTE} scheme is advantageous since no coupling between the
-\acs{CC} amplitude and multiplier equations is introduced.
-The model is also easily implemented, as it only requires access to a
-reference state optimized in the presence of the classical medium.
-The \acs{PTE} model thus offers an efficient approximation to the full
+\acs{CC} method, since no coupling between the \acs{CC} amplitudes and
+multipliers equations is introduced.
+The scheme only requires access to a
+reference state optimized in the presence of the classical medium and is
+thus readily implemented.
+The \acs{PTE} model offers an efficient approximation to the full
 \acs{PTED} model with a computational cost comparable to that of
@@
 
-\section[Coupled Cluster Perturbation Theory]{
-Coupled Cluster Perturbation Theory for a Quantum/Classical Polarizable Hamiltonian}\label{sec:ccpt-polarizable}
-
-We will now develop a perturbative expansion of the effective \acs{CC}
-Lagrangian and derive the \acs{CC}\acs{PT} series when a
-quantum/classical polarizable Hamiltonian is used.
+\section[Coupled Cluster Many-Body Perturbation Theory]{
+Coupled Cluster Many-Body Perturbation Theory}\label{sec:mbpt-polarizable}
+
+We will now develop a perturbative many-body expansion of the effective
+\acs{CC} Lagrangian when a quantum/classical polarizable Hamiltonian is
+used.
+The use of a \acs{CC} Ansatz for the wave function ensures term-by-term
+size extensivity of the energy at all orders.~\autocite{Helgaker2000-tz}
 The fluctuation potential $\Phi$ is assumed as the perturbation, hence
 it will be considered of order 1: $\Phi = O(1)$.
-Orders will be counted assuming a closed-shell \acs{HF} reference:
-we will use square brackets $[i]$ for an $i$-th order contributions.
+We still assume a closed-shell \acs{HF} reference function and
+perturbation orders will be counted accordingly.
+A superscript index in square brackets - $[i]$ -  will denote an $i$-th
+order contribution.
 We further assume that real wave functions are used.
-
-The structure of the stationarity conditions
+\todo[inline]{Is this a necessary assumption??}
+
+The structure of the stationarity conditions,
 Eqs.~\eqref{eq:pted-cc-amplitudes} and \eqref{eq:pted-cc-multipliers},
-already shows that $\tamp{\mu_i}^{[0]} = 0, \quad \forall i$ and
-$\tbar{\mu_i}^{[0]} = 0, \quad \forall i$.
+already shows that $\tamp{\mu_i}^{[0]} = 0, \,\forall i$ and
+$\tbar{\mu_i}^{[0]} = 0, \, \forall i$.
+Thus cluster operators can be of order 1 and higher. We introduce the
+following notation for the $i$-th order cluster operators and
+multipliers states:
+\begin{alignat}{2}
+  T_p^{[i]} = \sum_{\mu_p}\tamp{\mu_p}^{[i]}\cluster{\mu_p},
+  \quad&
+  \bra{\tbar{p}^{[i]}} = \sum_{\mu_p}\tbar{\mu_p}^{[i]}\bra{\mu_p}
+\end{alignat}
 
@@
 \begin{equation}\label{eq:source-term-pt}
-  \tilde{\s}(\tamp{},\tbar{}) =
-  \tilde{\s}^{[0]}
-  + \tilde{\s}^{[1]}
-  + \tilde{\s}^{[2]}
+  \s_{\mathrm{N}}(\tamp{},\tbar{}) =
+  \s_{\mathrm{N}}^{[0]}
+  + \s_{\mathrm{N}}^{[1]}
+  + \s_{\mathrm{N}}^{[2]}
   + \ldots
@@
 \begin{equation}\label{eq:pol-eq-pt}
-  \V\p^{[i]} + \tilde{\s}^{[i]} = 0
+  \V\p^{[i]} + \s_{\mathrm{N}}^{[i]} = 0
 \end{equation}
@@
 the perturbation.
-The order of the contributions to $\tilde{\s}(\tamp{},\tbar{})$ will be
+The order of the contributions to $\s_{\mathrm{N}}(\tamp{},\tbar{})$ will be
 solely determined by the amplitudes and multipliers.
 From the structure of \eqref{eq:pted-cc-polarization}, it is already
-clear that $\tilde{\s}(\tamp{},\tbar{})$ is at least first order in the
+clear that $\s_{\mathrm{N}}(\tamp{},\tbar{})$ is at least first order in the
 fluctuation potential:
 \begin{equation}
-  \tilde{\s}(\tamp{},\tbar{}) =
-  \tilde{\s}^{[1]}
-  + \tilde{\s}^{[2]}
+  \s_{\mathrm{N}}(\tamp{},\tbar{}) =
+  \s_{\mathrm{N}}^{[1]}
+  + \s_{\mathrm{N}}^{[2]}
   + \ldots
@@
 \begin{equation}
-  \V\p^{[0]} + \tilde{\s}^{[0]} = 0 \Rightarrow  \p^{[0]} = 0
-\end{equation}
-
-\paragraph*{Perturbative expansion of $\tilde{\s}$}
-
-\begin{equation}\label{eq:source-1st-order}
-  \tilde{\s}^{[1]}(\tamp{}, \tbar{}) =
-  \braket{\mathrm{HF} | \BCHfirst{\s}{T^{[1]}_1} | \mathrm{HF}}
-  + \braket{\tbar{1}^{[1]} | \s | \mathrm{HF}}
-\end{equation}
-In the free energy corrections we will use the $2n+1$ and $2n+2$ complying
-expression:
-\begin{equation}\label{eq:source-1st-order-rules}
-  \tilde{\s}^{[1]}(\tamp{}, \tbar{}) =
-  \braket{\mathrm{HF} | \BCHfirst{\s}{T^{[1]}_1} | \mathrm{HF}}
-\end{equation}
-
-\begin{equation}\label{eq:source-2nd-order}
-  \tilde{\s}^{[1]}(\tamp{}, \tbar{}) =
-  \braket{\mathrm{HF} | \BCHfirst{\s}{T^{[2]}_1} | \mathrm{HF}}
-  + \braket{\tbar{1}^{[2]} | \s | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i}^{[1]} | \BCHfirst{\s}{T^{[1]}} | \mathrm{HF}}
-\end{equation}
-In the free energy corrections we will use the $2n+1$ and $2n+2$ complying
-expression:
-\begin{equation}\label{eq:source-2nd-order-rules}
-  \tilde{\s}^{[2]}(\tamp{}, \tbar{}) =
-  \braket{\mathrm{HF} | \BCHfirst{\s}{T^{[2]}_1} | \mathrm{HF}}
-  + \braket{\tbar{i}^{[1]} | \BCHfirst{\s}{T^{[1]}} | \mathrm{HF}}
+  \V\p^{[0]} + \s_{\mathrm{N}}^{[0]} = 0 \Rightarrow  \p^{[0]} = 0
 \end{equation}
@@
 \end{equation}
-Considering the singles, doubles and triples excitation manifolds:
+Explicitly considering the singles, doubles and triples excitation manifolds:
 \begin{subequations}
@@
 Moreover, as is the case \emph{in vacuo}, there are no triples contributions to
-the first order amplitudes and multipliers: $\tamp{\mu_3}^{[1]} = \tbar{\mu_3}^{[1]} = 0,\,\,\forall \mu$.
+the first order amplitudes and multipliers: $\tamp{\mu_3}^{[1]} = \tbar{\mu_3}^{[1]} = 0$.
 
@@
   %%%% Steps
-  &= \left/ T_1^{[1]} = \sum_{\mu_1}\ket{\mu_1}\tamp{\mu_1} \right/
+  &= \left/ \text{Expand commutator}, T_1^{[1]} = \sum_{\mu_1}\ket{\mu_1}\tamp{\mu_1} \right/
   =
-  \sum_{\mu_1}\braket{\mathrm{HF}|\s|\mu_1}\tamp{\mu_1} \\
+  \sum_{\mu_1}\braket{\mathrm{HF}|\s|\mu_1}\tamp{\mu_1}^{[1]} \\
   %%%% Steps
   &= \left/ \text{Complex conjugation and real wave functions} \right/ \\
-  &= \sum_{\mu_1}\braket{\mu_1 | \s | \mathrm{HF}}\tbar{\mu_1}
-  = \braket{\tbar{\mu_1}^{[1]} | \s | \mathrm{HF}}
+  &= \sum_{\mu_1}\tbar{\mu_1}^{[1]}\braket{\mu_1 | \s | \mathrm{HF}} \\
+  %%%% Steps
+  &= \left/
+  \bra{\tbar{1}^{[1]}} = \sum_{\mu_1}\tbar{\mu_1}^{[1]}\bra{\mu_1} \right/
+  = \braket{\tbar{1}^{[1]} | \s | \mathrm{HF}}
  \end{aligned}
@@
 \begin{equation}
-  \tilde{\s}^{[m]}(\tamp{}^{[p]}, \tbar{}^{[q]})
+  \s_{\mathrm{N}}^{[m]}(\tamp{}^{[p]}, \tbar{}^{[q]})
 \end{equation}
@@
   + \frac{1}{2}{}^t\p^{[1]}\V\p^{[1]} +
-  {}^t\p^{[1]}\tilde{\s}^{[1]}(\tamp{}^{[1]}, \tbar{}^{[0]}) \\
+  {}^t\p^{[1]}\s_{\mathrm{N}}^{[1]}(\tamp{}^{[1]}, \tbar{}^{[0]}) \\
   &= \highlight{green}{E^{[2]}}
@@
     \highlight{green}{\sum_j\braket{\tbar{j}^{[1]} | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}}
-    + {}^t\p^{[1]}\tilde{\s}^{[2]}(\tamp{}^{[1]}, \tbar{}^{[1]}) \\
+    + {}^t\p^{[1]}\s_{\mathrm{N}}^{[2]}(\tamp{}^{[1]}, \tbar{}^{[1]}) \\
     &=
@@
     &+\frac{1}{2}{}^t\p^{[2]}\V\p^{[2]}
-    + {}^t\p^{[2]}\tilde{\s}^{[2]}(\tamp{}^{[2]}, \tbar{}^{[1]})
-    + {}^t\p^{[1]}\tilde{\s}^{[3]}(\tamp{}^{[2]}, \tbar{}^{[1]}) \\
+    + {}^t\p^{[2]}\s_{\mathrm{N}}^{[2]}(\tamp{}^{[2]}, \tbar{}^{[1]})
+    + {}^t\p^{[1]}\s_{\mathrm{N}}^{[3]}(\tamp{}^{[2]}, \tbar{}^{[1]}) \\
     &= \highlight{green}{E^{[4]}} + \frac{1}{2}{}^t\p^{[2]}\V\p^{[2]}
@@
     }
-    + {}^t\p^{[2]}\tilde{\s}^{[3]}(\tamp{}^{[2]}, \tbar{}^{[2]})
-    + {}^t\p^{[1]}\tilde{\s}^{[4]}(\tamp{}^{[2]}, \tbar{}^{[2]}) \\
+    + {}^t\p^{[2]}\s_{\mathrm{N}}^{[3]}(\tamp{}^{[2]}, \tbar{}^{[2]})
+    + {}^t\p^{[1]}\s_{\mathrm{N}}^{[4]}(\tamp{}^{[2]}, \tbar{}^{[2]}) \\
     &= \highlight{green}{E^{[5]}}
@@
 \end{equation}
+
+\section[Iterative Approximations: CC2 and CC3]{Iterative Approximations to CCSD and CCSDT: the CC2 and CC3 Models}\label{sec:cc2-and-cc3}
 
@@
   \denom{\mu_3}\tamp{\mu_3}^{*} =
-  - \braket{\mu_3 |\commutator{\Phi}{T_2^{1}} | \mathrm{HF}}
-  - \braket{\mu_3 | \commutator{\Phi}{T_2^[2]} | \mathrm{HF}}
+  - \braket{\mu_3 |\commutator{\Phi}{T_2^{[1]}} | \mathrm{HF}}
+  - \braket{\mu_3 | \commutator{\Phi}{T_2^{[2]}} | \mathrm{HF}}
   + O(3)

### 2016-08-24 Chapters/ContinuumSolvationModels.tex
Clean up and add one chapter (to be written)

--- 2016-08-23 Chapters/Ch01-ContinuumSolvationModels.tex
@@
 
-\section{A unifying view of classical polarizable
-models}\label{sec:variational}
+\begin{itemize}
+    \item why and how continuum solvation models
+    \item derivation of \acs{IEF}-\acs{PCM}, importance of Green's
+      functions.
+    \item \ac{BEM} and wavelet \acs{BEM} I think it's necessary to
+      mention also ddCOSMO and ddPCM as alternative strategies (put a
+      reference)
+    \item variational formulation (no derivations, just mention the
+      relevant literature)
+\end{itemize}
+
+\pagebreak
+
+\section[Variational Formulation of Classical Polarizable Models]{
+A Unifying View of Classical Polarizable Models within a Variational Formulation}
+\label{sec:variational}
 
@@
 
-\section{Perché è figo}
-
 \todo[inline]{PUT REFERENCES HERE!!!}
@@
 the quantum mechanical variational degrees of freedom.
+
+\todo[inline]{Mention the classical equivalent of the Hellmann--Feynman
+theorem and why it is particularly useful in formulating response
+theory.}

### 2016-08-24 Chapters/PaperSummary.tex
Clean up and add one chapter (to be written)

@@
+% mainfile: ../RobertoDiRemigioPhDThesis.tex
 %************************************************
-\chapter{Summary of the papers}\label{ch:papers-summary}
+\chapter{Summary of the publications}\label{ch:publications-summary}
 
-The codes used in the papers are all interfaced to the \PCMSolver library,
+\epigraph{$\frac{1}{2}$ research and $\frac{1}{2}\,\,T\acute{\epsilon}\chi\nu\eta$
+
+          $\frac{1}{2}$ observation, $\frac{1}{2}\,\,T\acute{\epsilon}\chi\nu\eta$
+
+          $\frac{1}{2}$ training, $\frac{1}{2}\,\,T\acute{\epsilon}\chi\nu\eta$}{
+  --- \textsc{Ezra Pound}, \textit{Canto LXXXV}}
+
+In the following, I provide a brief summary of the motivations, contents and
+conclusions of the papers this thesis is based on.
+Figures and tables here reported are minimally modified versions taken from the published work,
+captions are instead copied \emph{verbatim}.
+In addition, I give a description of my contributions to the work leading to
+the publication of each of the papers.
+\todo[inline]{Write something along these lines: "My coauthors have all read and approved."}
+
+\pagebreak
+
+\section*{Software}
+
+The codes used in the papers are all interfaced to the \pcmsolver library,
 of which I am the main developer and maintainer.
 
-\section{Paper I. Continuum solvation in the relativistic regime}
+\refstepcounter{dummy}
+\addcontentsline{toc}{section}{Continuum solvation in the relativistic regime}
+\section*{Paper I. Continuum solvation in the relativistic regime}
+
+Systems containing heavy elements are notoriously challenging for quantum chemistry.
+In addition to their significantly large sizes, one has to properly include the
+effect of special relativity in the quantum chemical description in order to
+achieve at least qualitative agreement with experiment.
+Additional complications arise if one needs to also include environment effects.
+Continuum models arguably represent a cost-effective strategy to achieve a first
+approximation of these effects.
+In this paper, we presented the first derivation and implementation of the \acs{PCM}
+coupled to a \acs{SCF} description of the solute in the fully relativistic, four-component
+regime.
+Our preliminary calculations on the group 16 dihydrides \ce{H2X} (\ce{X} =
+\ce{O}, \ce{S}, \ce{Se}, \ce{Te}, \ce{Po}) have shown that the method predicts
+a noticeable interplay of relativistic and solvent effects when heavier
+elements are involved.
+The main point of the paper was, however, the adoption of a fully modular
+programming strategy. We showed that it is entirely possible to adopt the same
+\acs{PCM} code and implementation "checkpoints" across altogether different
+problem domains.
+
+\begin{figure}[!htb]
+\centering
+\scalebox{0.8}{\input{gfx/relapcm/algorithm.tikz}}
+\caption[Modular approach to programming a \acs{PCM} functionality into an existing \acs{SCF} code.]{
+  Schematic view of the implemented SCF algorithm. Computations/data in
+  \textcolor{Blue}{blue} are on the \pcmsolver side, in
+  \textcolor{Green}{green} on the \DIRAC side.
+  }
+\label{fig:algorithm}
+\end{figure}
+
+As a by-product of the 4-component implementation, we were able to obtain and visualize \acs{MEP} maps
+from 4-component \acs{SCF} wave functions. These add yet another interpretive tool to the toolbox available
+in the 4-component relativistic regime.
+\todo[inline]{Recalculate MEP maps and reobtain these plots using ParaView.}
+\begin{figure}
+\centering
+\begin{subfigure}[b]{0.5\textwidth}
+\missingfigure{Reobtain this figure: H2O\_MEP-diff-pcmdc-dc}
+\caption{\ce{H2O}: $\mathrm{MEP}_\mathrm{PCM-DC} - \mathrm{MEP}_\mathrm{DC}$}
+\label{fig:H2Odiff_pcmdc-dc}
+\end{subfigure}%
+\begin{subfigure}[b]{0.5\textwidth}
+\missingfigure{Reobtain this figure: H2Po\_MEP-diff-pcmdc-dc}
+\caption{\ce{H2Po}: $\mathrm{MEP}_\mathrm{PCM-DC} - \mathrm{MEP}_\mathrm{DC}$}
+\label{fig:H2Podiff_pcmdc-dc}
+\end{subfigure}
+\caption[Molecular electrostatic potential (\acs{MEP}) maps for \ce{H2O} and \ce{H2Po} at
+  the \acs{DFT}/PBE0 level of theory.]{
+  Molecular electrostatic potential (\acs{MEP}) maps for \ce{H2O} and \ce{H2Po} at
+  the \acs{DFT}/PBE0 level of theory.
+  The isocontours are in the range $[-0.02, 0.02] \si{\hartree\per\elementarycharge}$ and are
+  color-coded from red to blue.
+  The values plotted are differences between the values obtained for the
+  Hamiltonians referred to in the subcaptions. The geometry optimized \emph{in
+  vacuo}, using the \acl{DC} Hamiltonian as reference.
+  \acs{PCM}-\acs{DC}: \acl{DC} in water, \acs{DC}: \acl{DC} \emph{in vacuo}.}
+\label{fig:MEP_maps}
+\end{figure}
+The interface to \DIRAC was later released in the 2014 and later versions,
+providing \acs{PCM} capabilities to the software package.
 
@@
 
-\section{Paper II. The wavelet boundary element method for \acs{PCM}}
-\section{Paper III: Continuum modelling of non homogeneous environments}
+\refstepcounter{dummy}
+\addcontentsline{toc}{section}{The wavelet Galerkin boundary element method for PCM}
+\section*{Paper II. The wavelet Galerkin boundary element method for PCM}
 
-%************************************************
-%*****************************************
-%*****************************************
-%*****************************************
-%*****************************************
-%*****************************************
+Continuum solvation models are inherently parametrized. Apart from solvent permittivities,
+the atomic radii and molecular surface definition play a crucial role
+in determining the performance of the models.
+However, the numerical accuracy of the \acs{BEM} procedure used to numerically solve the
+underlying \acs{BIE} is a not-so-often studied aspect of these models.
+Traditionally, collocation methods have been used, but these require parametrization of some of the necessary surface integrals.
+Galerkin methods do not suffer from such a limitation and additionally preserve symmetry of the underlying
+boundary integral operators.
+The use of biorthogonal wavelet bases as finite elements achieves sparsity in
+the \acs{BEM} procedure, due to their intrinsically hierarchical structure and
+the existence of \emph{a priori} and \emph{a posteriori} matrix compression
+estimates.
+Thus, wavelet Galerkin \acs{BEM} represents a valid alternative to traditional
+collocation methods, both to achieve a better computational scaling and to
+provide accurate, benchmark results.\autocite{Harbrecht2004-uo,
+Harbrecht2006-ug, Dahmen2006-pj}
+
+Already~\citeauthor{Weijo2010-hy} had shown that using \ac{PWC} bases can lead
+to superior accuracy and convergence in the calculation of quantum mechanical
+molecular solvation energies.
+In this work we showed that even faster convergence can be achieved when
+\ac{PWL} bases are used instead.
+Moreover, the same holds for the calculation of static electric properties.
+Notably, the traditional collocation solver cannot guarantee the same accuracy,
+even for very large finite element bases. This suggests that, in some cases,
+\acs{BEM} collocation methodologies might slow down or even prevent the
+convergence of the quantum mechanical response equations solvers.
+
+For this paper, I provided template interface and test sets for the cavity
+generator\autocite{Harbrecht2009-no, Harbrecht2011-dk} and wavelet Galerkin
+\acs{BEM} solver with \pcmsolver.
+These were used to interface with the new C++ implementation of the wavelet
+Galerkin solvers of Monica Bugeanu.
+I implemented the interface between the \LSDALTON quantum chemistry software
+package and the \pcmsolver software library. The interface allows to run \acs{HF} and
+\acs{KS}-\acs{DFT} single-point and linear response calculations.
+I, together with coauthor Krzysztof Mozgawa, performed the benchmark quantum
+chemical calculations presented in the paper.
+Finally, I coordinated the editing of all manuscript drafts. I wrote Sections 2.1 and 2.3,
+while Sections 3 and 4 were co-written with the first author, Monica Bugeanu.
+I performed most of the data analysis and produced tables and graphs.
+
+\begin{figure}[tb]
+ \centering
+  \includegraphics[width=\textwidth]{gfx/wemlin/alpha_convergence.png}
+  \caption[Convergence of $\alpha_\mathrm{iso}$ with respect to the number of molecular
+  electrostatic potential (MEP) evaluation points on the cavity surface.]
+  {Convergence of $\alpha_\mathrm{iso}$ with respect to the number of
+   molecular electrostatic potential (MEP) evaluation points on the cavity
+   surface.
+   The values reported are in $a_0^3$ and refer to benzene.
+   The lower axis reports the average area for the collocation
+   tesselation, while the upper axis refers to the patch level in the
+   wavelet Galerkin discretization. The annotation report the number of
+   MEP evaluation points.
+   The compression parameter triplet was set to $a = 2.5,\ d^\prime =
+   2.0,\ b = 0.01$ for the piecewise constant wavelets and to $a = 2.5,\
+   d^\prime = 3.0,\ b = 0.01$ for piecewise bilinear wavelets.
+   The limit value is extrapolated from the results obtained from the
+   piecewise bilinear wavelet Galerkin scheme.}
+  \label{fig:alpha_convergence}
+\end{figure}
+
+Finally, the interface to \LSDALTON was later released in the 2016 version,
+providing \acs{PCM} capabilities to the software package.
+
+\refstepcounter{dummy}
+\addcontentsline{toc}{section}{Non homogeneous environments}
+\section*{Paper III: Non homogeneous environments}
+
+Continuum solvation models offer a simple route to the treatment of non
+homogeneous environments.
+The general integral equation formulation exposed in Chapter~\ref{ch:CSM} is in
+fact transparent with respect to the definition of the Green's function for the
+space portion exterior to the cavity.
+As already shown by \citeauthor{Cances1998-og}, the same mathematical formalism can easily accommodate
+environments characterized by distance-dependent or anisotropic Green's functions, such as those for
+ionic liquids (in the weakly interacting regime described by the linearized Poisson--Boltzmann equation)
+and liquid crystals, respectively.
+\citeauthor{Frediani2004-er} showed that a numerical representation of the
+Green's function is sufficient to obtain the boundary integral operators in the
+\acs{PCM} integral equation.
+The authors introduced a numerical integration procedure to calculate the
+Green's function for an environment characterized by spatially varying, yet
+cilindrically symmetric, permittivity functions: a model for planar diffuse interfaces.
+
+In this work, a similar procedure was introduced to tackle diffuse interfaces
+in spherical symmetry.
+In contrast to previously existing work, our implementation offer a more robust
+treatment of thin interfaces, with a rather generic functional form for the
+permittivity profile.
+We thoroughly analyzed the necessity for the \emph{a posteriori} removal of the
+Coulomb singularity from the computed Green's function and its efficient
+implementation.
+Interface width and curvature influence the transfer of ions and molecules across
+spherically symmetric interfaces and peculiar properties may arise.
+In this work, we analyzed both effect on the water-vapor and oil-water transfer
+of \ce{Li+}, \ce{Br-}, acetone, \emph{para}-nitroaniline and the L0 dye.
+Nonelectrostatic interactions were not included in our implementation, although
+they have been proved to be crucial for non homogeneous
+environments.~\autocite{Mozgawa2014-ad}
+Nevertheless our implementation represents a first significant step in the continuum treatment
+of such nontrivial environments.
+
+\begin{figure}[tb]
+  \centering
+  \begin{subfigure}[t]{.49\linewidth}
+  \centering
+  % Figure8a <--> dipole_acetone_perpendicular
+  \includegraphics[width=\textwidth]{gfx/spherical/width.png}
+  \caption{Width effect}
+  \label{fig:profile_dipole_acetone_width}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[t]{.49\linewidth}
+  \centering
+  % Figure8b <--> curv_fullprofile_dip_acetone
+  \includegraphics[width=\textwidth]{gfx/spherical/radius.png}
+  \caption{Radius effect}
+  \label{fig:profile_dipole_acetone_fullprofile}
+  \end{subfigure}
+
+  \begin{subfigure}[t]{.49\linewidth}
+  \centering
+  % Figure8c <--> curv_dip_acetone
+  \includegraphics[width=\textwidth]{gfx/spherical/curvature.png}
+  \caption{Curvature effect}
+  \label{fig:profile_dipole_acetone_curvature}
+  \end{subfigure}
+  \caption[Width, radius and curvature effect on the dipole moment of acetone for water-air droplet phase transfer.]
+  {Dipole moment profiles for acetone (with dipole moment
+  perpendicular to the interface) transfer at water droplet in air, for
+  varying interface widths, values of $R_0$ and the curvature effect on
+  dipole moment, respectively.}
+  \label{fig:profile_dipole_acetone}
+\end{figure}
+
+I contributed most of the theoretical work for this paper, based on early drafts from coauthors
+Ville Weijo and Hui Cao. In particular, I derived the separation of
+the Coulomb singularity in its final form.
+Moreover, I contributed the implementation and testing of the Green's function code.
+The interface to the \LSDALTON program package, developed within paper II, was
+also used for this paper.
+I wrote the first draft of the paper and coordinated all subsequent editing stages.
+
+\refstepcounter{dummy}
+\addcontentsline{toc}{section}{Relativistic calculation of EPR and pNMR parameters in solution}
+\section*{Paper IV: Relativistic calculation of EPR and pNMR parameters in solution}
+
+\refstepcounter{dummy}
+\addcontentsline{toc}{section}{Open-ended self-consistent field response theory in solution}
+\section*{Paper V: Open-ended self-consistent field response theory in solution}

### 2016-08-24 Chapters/QuantumMechanics.tex
Clean up and add one chapter (to be written)

% mainfile: ../RobertoDiRemigioPhDThesis.tex
\chapter{A Quick Tour of Molecular Quantum Mechanics}\label{ch:QM}

\begin{itemize}
  \item nonrelativistic and relativistic Hamiltonians use in molecular
    quantum mechanics
  \item Fock space
  \item single-reference methods: \ac{HF} and \ac{KS} \ac{DFT}
\end{itemize}

### 2016-08-24 FrontBackmatter/Colophon.tex
Clean up and add one chapter (to be written)

\pagestyle{empty}

\hfill

\vfill

\pdfbookmark[0]{Colophon}{colophon}
\section*{Colophon}
This document was typeset using the typographical look-and-feel
\texttt{classicthesis} developed by Andr\'e Miede. The style was
inspired by Robert Bringhurst's seminal book on typography ``\emph{The
Elements of Typographic Style}''. \texttt{classicthesis} is available
for both \LaTeX\ and \mLyX:
\begin{center}
\url{https://bitbucket.org/amiede/classicthesis/}
\end{center}

\bigskip

\noindent\finalVersionString

\emph{Garamond} and \emph{XITS} type faces are used.

%\paragraph{note:} The custom size of the textblock was calculated
%using the directions given by Mr. Bringhurst (pages 26--29 and
%175/176). 10~pt Palatino needs  133.21~pt for the string
%``abcdefghijklmnopqrstuvwxyz''. This yields a good line length between
%24--26~pc (288--312~pt). Using a ``\emph{double square textblock}''
%with a 1:2 ratio this results in a textblock of 312:624~pt (which
%includes the headline in this design). A good alternative would be the
%``\emph{golden section textblock}'' with a ratio of 1:1.62, here
%312:505.44~pt. For comparison, \texttt{DIV9} of the \texttt{typearea}
%package results in a line length of 389~pt (32.4~pc), which is by far
%too long. However, this information will only be of interest for
%hardcore pseudo-typographers like me.%
%
%To make your own calculations, use the following commands and look up
%the corresponding lengths in the book:
%\begin{verbatim}
%    \settowidth{\abcd}{abcdefghijklmnopqrstuvwxyz}
%    \the\abcd\ % prints the value of the length
%\end{verbatim}
%Please see the file \texttt{classicthesis.sty} for some precalculated
%values for Palatino and Minion.
%
%    \settowidth{\abcd}{abcdefghijklmnopqrstuvwxyz}
%    \the\abcd\ % prints the value of the length

### 2016-08-24 FrontBackmatter/Introduction.tex
Some organizational work
Clean up and add one chapter (to be written)

@@
     --- \textsc{E. E. Cummings}}
+
+This will be the subject of Chapter \ref{ch:QM}
+
+\section*{The Problem of Solvation or Taming Complexity with Models}
+
+\autocite{Anderson1972-ai}
+
+This will be the subject of Chapter \ref{ch:CSM}
+
+\section*{The Road to Reality or Molecular Response Properties}
+
+This will be the subject of Chapter \ref{ch:molprop}
+
+\section*{Accurate Methods for Accurate Properties}
+
+This will be the subject of Chapter \ref{ch:solvation-correlation}

### 2016-08-24 Chapters/PaperSummary.tex
Remove useless pictures

@@
 
+\begin{figure}[!htb]
+\centering
+\scalebox{0.8}{\input{gfx/relapcm/algorithm.tikz}}
+\caption[Modular approach to programming a \acs{PCM} functionality into an existing \acs{SCF} code.]{
+  Schematic view of the implemented SCF algorithm. Computations/data in
+  \textcolor{Blue}{blue} are on the \pcmsolver side, in
+  \textcolor{Green}{green} on the \DIRAC side.
+  }
+\label{fig:algorithm}
+\end{figure}
+
 \refstepcounter{dummy}
@@
 
-\begin{figure}[!htb]
-\centering
-\scalebox{0.8}{\input{gfx/relapcm/algorithm.tikz}}
-\caption[Modular approach to programming a \acs{PCM} functionality into an existing \acs{SCF} code.]{
-  Schematic view of the implemented SCF algorithm. Computations/data in
-  \textcolor{Blue}{blue} are on the \pcmsolver side, in
-  \textcolor{Green}{green} on the \DIRAC side.
-  }
-\label{fig:algorithm}
-\end{figure}
-
 As a by-product of the 4-component implementation, we were able to obtain and visualize \acs{MEP} maps
@@
 in the 4-component relativistic regime.
-\todo[inline]{Recalculate MEP maps and reobtain these plots using ParaView.}
-\begin{figure}
-\centering
-\begin{subfigure}[b]{0.5\textwidth}
-\missingfigure{Reobtain this figure: H2O\_MEP-diff-pcmdc-dc}
-\caption{\ce{H2O}: $\mathrm{MEP}_\mathrm{PCM-DC} - \mathrm{MEP}_\mathrm{DC}$}
-\label{fig:H2Odiff_pcmdc-dc}
-\end{subfigure}%
-\begin{subfigure}[b]{0.5\textwidth}
-\missingfigure{Reobtain this figure: H2Po\_MEP-diff-pcmdc-dc}
-\caption{\ce{H2Po}: $\mathrm{MEP}_\mathrm{PCM-DC} - \mathrm{MEP}_\mathrm{DC}$}
-\label{fig:H2Podiff_pcmdc-dc}
-\end{subfigure}
-\caption[Molecular electrostatic potential (\acs{MEP}) maps for \ce{H2O} and \ce{H2Po} at
-  the \acs{DFT}/PBE0 level of theory.]{
-  Molecular electrostatic potential (\acs{MEP}) maps for \ce{H2O} and \ce{H2Po} at
-  the \acs{DFT}/PBE0 level of theory.
-  The isocontours are in the range $[-0.02, 0.02] \si{\hartree\per\elementarycharge}$ and are
-  color-coded from red to blue.
-  The values plotted are differences between the values obtained for the
-  Hamiltonians referred to in the subcaptions. The geometry optimized \emph{in
-  vacuo}, using the \acl{DC} Hamiltonian as reference.
-  \acs{PCM}-\acs{DC}: \acl{DC} in water, \acs{DC}: \acl{DC} \emph{in vacuo}.}
-\label{fig:MEP_maps}
-\end{figure}
 The interface to \DIRAC was later released in the 2014 and later versions,
@@
 
-\begin{figure}[tb]
- \centering
-  \includegraphics[width=\textwidth]{gfx/wemlin/alpha_convergence.png}
-  \caption[Convergence of $\alpha_\mathrm{iso}$ with respect to the number of molecular
-  electrostatic potential (MEP) evaluation points on the cavity surface.]
-  {Convergence of $\alpha_\mathrm{iso}$ with respect to the number of
-   molecular electrostatic potential (MEP) evaluation points on the cavity
-   surface.
-   The values reported are in $a_0^3$ and refer to benzene.
-   The lower axis reports the average area for the collocation
-   tesselation, while the upper axis refers to the patch level in the
-   wavelet Galerkin discretization. The annotation report the number of
-   MEP evaluation points.
-   The compression parameter triplet was set to $a = 2.5,\ d^\prime =
-   2.0,\ b = 0.01$ for the piecewise constant wavelets and to $a = 2.5,\
-   d^\prime = 3.0,\ b = 0.01$ for piecewise bilinear wavelets.
-   The limit value is extrapolated from the results obtained from the
-   piecewise bilinear wavelet Galerkin scheme.}
-  \label{fig:alpha_convergence}
-\end{figure}
-
 Finally, the interface to \LSDALTON was later released in the 2016 version,
@@
 
-\begin{figure}[tb]
-  \centering
-  \begin{subfigure}[t]{.49\linewidth}
-  \centering
-  % Figure8a <--> dipole_acetone_perpendicular
-  \includegraphics[width=\textwidth]{gfx/spherical/width.png}
-  \caption{Width effect}
-  \label{fig:profile_dipole_acetone_width}
-  \end{subfigure}
-  ~
-  \begin{subfigure}[t]{.49\linewidth}
-  \centering
-  % Figure8b <--> curv_fullprofile_dip_acetone
-  \includegraphics[width=\textwidth]{gfx/spherical/radius.png}
-  \caption{Radius effect}
-  \label{fig:profile_dipole_acetone_fullprofile}
-  \end{subfigure}
-
-  \begin{subfigure}[t]{.49\linewidth}
-  \centering
-  % Figure8c <--> curv_dip_acetone
-  \includegraphics[width=\textwidth]{gfx/spherical/curvature.png}
-  \caption{Curvature effect}
-  \label{fig:profile_dipole_acetone_curvature}
-  \end{subfigure}
-  \caption[Width, radius and curvature effect on the dipole moment of acetone for water-air droplet phase transfer.]
-  {Dipole moment profiles for acetone (with dipole moment
-  perpendicular to the interface) transfer at water droplet in air, for
-  varying interface widths, values of $R_0$ and the curvature effect on
-  dipole moment, respectively.}
-  \label{fig:profile_dipole_acetone}
-\end{figure}
-
 I contributed most of the theoretical work for this paper, based on early drafts from coauthors

### 2016-08-24 gfx/spherical/curvature.py

@@ -1,36 +1,0 @@
-#!/usr/bin/env python
-
-import matplotlib as mpl
-import matplotlib.pyplot as plt
-
-import numpy as np
-
-mpl.style.use('ggplot')
-# The colors are from the ggplot style, consistent with grey background
-ggplot = ['#E24A33', '#348ABD', '#988ED5', '#777777', '#FBC15E', '#8EBA42', '#FFB5B8']
-# List of line styles
-lines = ['-', '--', ':', '-.', '-', '--', ':']
-# List of labels
-labs = [r'$R_0 = 40$', r'$R_0 = 20$', r'$R_0 = 10$', r'$R_0 = 5$']
-params = {'text.usetex': False,
-          'mathtext.fontset': 'stixsans',
-          'font.size' : 14
-         }
-mpl.rcParams.update(params)
-
-fig, ax = plt.subplots(figsize=(10, 5), facecolor='w', edgecolor='k')
-
-fname='curv_dip_acetone_perpen.dat'
-
-# Read columns
-data = np.loadtxt(fname, usecols=(0, 1, 2, 3, 4))
-for col in range(1, 5):
-    ax.plot(data[:, 0], data[:, col], color=ggplot[col-1], marker='+', ls=lines[col-1], lw=2, ms=7, mew=1.5, label=labs[col-1])
-    ax.set_ylabel(r'$\mu\,\,[\mathrm{D}]$')
-    ax.legend(loc='upper right', ncol=1)
-
-ax.set_xlim(-5.0, 10.0)
-ax.set_ylim(0.0, 1.4)
-ax.set_xlabel(r'$z\,\,[\AA]$')
-
-fig.savefig('curvature.png', transparent=False, bbox_inches='tight')

### 2016-08-24 gfx/spherical/radius.py

@@ -1,39 +1,0 @@
-#!/usr/bin/env python
-
-import matplotlib as mpl
-import matplotlib.pyplot as plt
-
-import numpy as np
-
-mpl.style.use('ggplot')
-# The colors are from the ggplot style, consistent with grey background
-ggplot = ['#E24A33', '#348ABD', '#988ED5', '#777777', '#FBC15E', '#8EBA42', '#FFB5B8']
-# List of line styles
-lines = ['-', '--', ':', '-.', '-', '--', ':']
-params = {'text.usetex': False,
-          'mathtext.fontset': 'stixsans',
-          'font.size' : 14
-         }
-mpl.rcParams.update(params)
-
-fig, ax = plt.subplots(figsize=(10, 5), facecolor='w', edgecolor='k')
-
-R0_5  = np.loadtxt(fname='R0_5.dat',  usecols=(0, 1))
-ax.plot(R0_5[:, 0] , R0_5[:, 1] , color=ggplot[4], marker='+', ls=lines[4], lw=2, ms=7, mew=1.5, label=r'$R_0 = 5$')
-R0_10 = np.loadtxt(fname='R0_10.dat', usecols=(0, 1))
-ax.plot(R0_10[:, 0], R0_10[:, 1], color=ggplot[3], marker='+', ls=lines[3], lw=2, ms=7, mew=1.5, label=r'$R_0 = 10$')
-R0_20 = np.loadtxt(fname='R0_20.dat', usecols=(0, 1))
-ax.plot(R0_20[:, 0], R0_20[:, 1], color=ggplot[2], marker='+', ls=lines[2], lw=2, ms=7, mew=1.5, label=r'$R_0 = 20$')
-R0_40 = np.loadtxt(fname='R0_40.dat', usecols=(0, 1))
-ax.plot(R0_40[:, 0], R0_40[:, 1], color=ggplot[1], marker='+', ls=lines[1], lw=2, ms=7, mew=1.5, label=r'$R_0 = 40$')
-R0_80 = np.loadtxt(fname='R0_80.dat', usecols=(0, 1))
-ax.plot(R0_80[:, 0], R0_80[:, 1], color=ggplot[0], marker='+', ls=lines[0], lw=2, ms=7, mew=1.5, label=r'$R_0 = 80$')
-
-ax.set_ylabel(r'$\mu\,\,[\mathrm{D}]$')
-ax.legend(loc='upper right', ncol=1)
-
[…]

### 2016-08-24 gfx/spherical/width.py

@@ -1,37 +1,0 @@
-#!/usr/bin/env python
-
-import matplotlib as mpl
-import matplotlib.pyplot as plt
-
-import numpy as np
-
-mpl.style.use('ggplot')
-# The colors are from the ggplot style, consistent with grey background
-ggplot = ['#E24A33', '#348ABD', '#988ED5', '#777777', '#FBC15E', '#8EBA42', '#FFB5B8']
-# List of line styles
-lines = ['-', '--', ':', '-.', '-', '--', ':']
-params = {'text.usetex': False,
-          'mathtext.fontset': 'stixsans',
-          'font.size' : 14
-         }
-mpl.rcParams.update(params)
-
-fig, ax = plt.subplots(figsize=(10, 5), facecolor='w', edgecolor='k')
-
-# Read in data
-w5_c20  = np.loadtxt(fname='dipole_acetone_perpendicular_w5_c20.dat',  usecols=(0, 1))
-w10_c20 = np.loadtxt(fname='dipole_acetone_perpendicular_w10_c20.dat', usecols=(1,))
-w20_c20 = np.loadtxt(fname='dipole_acetone_perpendicular_w20_c20.dat', usecols=(1,))
-z = w5_c20[:, 0] - 20.0
-
-ax.plot(z, w5_c20[:, 1], color=ggplot[0], marker='+', ls=lines[0], lw=2, ms=7, mew=1.5, label=r'$w = 5$')
-ax.plot(z, w10_c20[:], color=ggplot[1], marker='+', ls=lines[1], lw=2, ms=7, mew=1.5, label=r'$w = 10$')
-ax.plot(z, w20_c20[:], color=ggplot[2], marker='+', ls=lines[2], lw=2, ms=7, mew=1.5, label=r'$w = 20$')
-ax.set_ylabel(r'$\mu\,\,[\mathrm{D}]$')
-ax.legend(loc='upper right', ncol=1)
-
-ax.set_xlim(-10.0, 25.0)
-ax.set_ylim(3.0, 4.4)
-ax.set_xlabel(r'$z\,\,[\AA]$')
-
-fig.savefig('width.png', transparent=False, bbox_inches='tight')

### 2016-08-24 gfx/wemlin/alpha_convergence.py

@@ -1,99 +1,0 @@
-#!/usr/bin/python2
-# -*- python -*-
-# -*- coding: utf-8 -*-
-# vim:filetype=python:
-# (c) Roberto Di Remigio  <roberto.d.remigio@uit.no>
-# licensed under the GNU Lesser General Public License
-
-import matplotlib as mpl
-import matplotlib.pyplot as plt
-
-import numpy as np
-
-mpl.style.use('ggplot')
-# The colors are from the ggplot style, consistent with grey background
-ggplot = ['#E24A33', '#348ABD', '#988ED5', '#777777', '#FBC15E', '#8EBA42', '#FFB5B8']
-
-params = {'text.usetex': False,
-          'mathtext.fontset': 'stixsans',
-          'font.size' : 14
-         }
-mpl.rcParams.update(params)
-
-limit = 76.678052
-# Get the data
-fname = 'ief_convergence.dat'
-data = np.loadtxt(fname, usecols=(0, 4))
-ief_x = data[:, 0].astype(int)
-ief_y = data[:, 1]
-
-fname = 'pwc_convergence.dat'
-data = np.loadtxt(fname, usecols=(0, 4))
-pwc_x = data[:, 0].astype(int)
-pwc_y = data[:, 1]
-
-fname = 'pwl_convergence.dat'
-data = np.loadtxt(fname, usecols=(0, 4))
-pwl_x = data[:, 0].astype(int)
-pwl_y = data[:, 1]
-
-y_formatter = mpl.ticker.ScalarFormatter(useOffset=False)
-
-fig, ax = plt.subplots(figsize=(12, 8))
-
-ax.spines['right'].set_color('none')
-ax.yaxis.set_ticks_position('left')
-
-ax.tick_params(axis='x')
-ax.tick_params(axis='y')
-
-l3 = ax.plot(ief_x, ief_y, label=r'Collocation', lw=1, ls='-', marker='d', color=ggplot[0])
-dof = [str(x) for x in ief_x]
-coords = zip(ief_x, ief_y)
-ax.annotate('{}'.format(dof[0]), xy=(coords[0]), xytext=(0, 10),  va='bottom', ha='right',  textcoords='offset points', color=ggplot[0], weight='demi')
[…]

### 2016-08-24 Chapters/ContinuumSolvationModels.tex
Add some more stuff from notes scattered around

@@
 \pagebreak
+
+\section{Boundary integral equations}\label{sec:BIE}
+
+\subsection{Transmission problem}
+We assume Euclidean space $\mathbb{R}^3$ to be partitioned into two
+subdomains $\Omegai$ and $\Omegae$ sharing a boundary $\Gamma$.
+We further assume that $\Omegai$ is a closed domain, entirely contained inside $\Omegae$.
+The transmission problem is posed as follows:
+\todo[inline]{Some consideration on suitable functional spaces and the metric thereof.}
+\begin{subequations}\label{eq:transmission}
+\begin{align}
+  \Li u &= f_\mathrm{i} \,\, \text{in}\,\, \Omegai \label{eq:internal} \\
+  \Le u &= f_\mathrm{e} \,\, \text{in}\,\, \Omegae \label{eq:external} \\
+  [u] &= \ue - \ui = g_\mathrm{D} \,\, \text{on}\,\, \Gamma
+  \label{eq:trace-jump} \\
+[\partial_L u] &= \partiale u - \partiali u = g_\mathrm{N} \,\,
+\text{on}\,\, \Gamma \label{eq:conormal-jump} \\
+|u| &\leq C \|x \|^{-1} \,\,\text{for}\,\,\| x \|\rightarrow\infty
+\label{eq:radiation}
+\end{align}
+\end{subequations}
+where the jump conditions (eqs. \eqref{eq:trace-jump} and
+\eqref{eq:conormal-jump}) are given in terms of Dirichlet and Neumann
+data for the solution $u$. The jump conditions are expressed in terms of
+trace operators for the solution $u$ and its conormal derivative. For
+notational simplicity, will use the symbols $\partiale$ and $\partiali$
+for the latter and only give it in explicit form when needed. Please
+refer to the book by \citeauthor{Sauter2011} for further details. The
+fundamental solutions, or Green's functions, for the elliptic
+differential operators $\Li$ and $\Le$ will be denoted by $\Gi$ and
+$\Ge$, respectively.
+
+\subsection{Boundary integral operators}
+We introduce the relevant boundary integral operators for the abovementioned
+transmission problem. They are three of the four components of the Calder\'on projector.
+For $v \in L^2(\Gamma)$, $x, y \in \Gamma$:\footnote{We can assume a less restrictive functional space, \ie a Sobolev space of fractional order.}
+\begin{equation}
+\begin{aligned}
+(\bi{S}_\star v)(x) &= \int_\Gamma G_\star(x, y)v(y)\diff y \\
+(\bi{D}_\star v)(x) &= \int_\Gamma \partial_{L_\star,y}G_\star(x, y)v(y)\diff y \\
+(\bi{D}^\dagger_\star v)(x) &= \int_\Gamma \partial_{L_\star,x}G_\star(x, y)v(y)\diff y \\
+\end{aligned}
+\end{equation}
+where $\star$ is a placeholder for the $\mathrm{i}$ or $\mathrm{e}$ subscript.
+
+\begin{lemma}[Properties of the boundary integral operators]
+  The integral operators introduced above enjoy the following
+  properties~\autocite{Hsiao2008, Sauter2011}:
+  \begin{enumerate}
+      \item on $L^2(\Gamma)$ $\bi{S}_\star$ is self-adjoint,
+        $\bi{D}^\dagger_\star$ is the adjoint operator of
+        $\bi{D}_\star$.
+      \item The commutation relations:
+        \begin{alignat}{2}
+          \bi{D}_\star\bi{S}_\star = \bi{S}_\star\bi{D}^\dagger_\star, \quad&
+          \bi{S}_\star\bi{D}_\star = \bi{D}^\dagger_\star\bi{S}_\star
+        \end{alignat}
+        hold
+      \item The operators:
+        \begin{subequations}
+          \begin{align}
+   \bi{S}_\star &: H^{-\frac{1}{2}}(\Gamma) \rightarrow H^{\frac{1}{2}}(\Gamma) \\
+   \bi{D}_\star &: H^{\frac{1}{2}}(\Gamma) \rightarrow H^{\frac{1}{2}}(\Gamma) \\
+   \bi{D}^\dagger_\star &: H^{-\frac{1}{2}}(\Gamma) \rightarrow H^{-\frac{1}{2}}(\Gamma)
+          \end{align}
+        \end{subequations}
+      \item The operator $\bi{S}_\star$ is coervice and admits a
+        continuous inverse in the aforementioned Sobolev spaces.
+      \item The operators $\lambda - \bi{D}_\star$ and $\lambda -
+        \bi{D}^\dagger_\star$ with $\lambda \in (-2\pi, +\infty)$
+        admit a continuous inverse in the aforementioned
+        Sobolev spaces.
+  \end{enumerate}
+\end{lemma}
+
+\subsection{Lemma of Integral Representation}
+For the transmission problem \ref{eq:transmission} there holds:
+\begin{itemize}
+\item $\forall x \in \Omegai$
+\begin{equation}
+u = \bi{S}_\mathrm{i}(\partiali u)
+- \bi{D}_\mathrm{i}(\ui) + \int_{\Omegai} \Gi f_\mathrm{i}\diff y
+\end{equation}
+\item $\forall x \in \Omegae$
+\begin{equation}
+u = -\bi{S}_\mathrm{e}(\partiale u)
++ \bi{D}_\mathrm{e}(\ue) + \int_{\Omegae}\Ge f_\mathrm{e}\diff y
+\end{equation}
+\item $\forall x \in \Gamma$
+  \begin{equation}\label{eq:stat3}
+    \frac{1}{2}\ui = \bi{S}_\mathrm{i}(\partiali u)
+    - \bi{D}_\mathrm{i}(\ui) + \int_{\Omegai}\Gi f_\mathrm{i}\diff y
+  \end{equation}
+\item $\forall x \in \Gamma$
+  \begin{equation}\label{eq:stat4}
+    \frac{1}{2}\ue = -\bi{S}_\mathrm{e}(\partiale u)
+    + \bi{D}_\mathrm{e}(\ue) + \int_{\Omegae}\Ge f_\mathrm{e}\diff y
+  \end{equation}
+\end{itemize}
+
+\pagebreak
+\section{IEFPCM}\label{sec:IEF}
+
+Note that the figures refer to cases where the dielectric is supposed to
+be homogeneous and isotropic. The derivation are, however, general and
+do not assume homogeneity and isotropy.
+The source charge densities are not assumed to be continuous, the
+hypothesis that they are entirely supported in either of the internal or
+external subdomains is however made:
+\begin{alignat}{2}
+  \Supp(\rhoi) \subseteq \Omegai, \quad& \Supp(\rhoe) \subseteq \Omegae
+\end{alignat}
+
+%%%% CASE 0
+
+\subsection{Case 0: sources in a vacuum cavity inside a dielectric}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.65\textwidth]{case0}
+  \caption{Vacuum Cavity in dielectric with permittivity $\diel_\mathrm{e}>1$.
+  Source charge density $\rho_\mathrm{i}$ inside the cavity
+  ($\diel_\mathrm{i}=1$). \textcolor{red}{Regular IEF}}
+  \label{fig:case0}
+\end{figure}
+
+This is the regular IEF as derived in:~\autocite{Cances1998}
+\begin{subequations}
+\begin{align}
+  \nabla^2 u &= \rhoi \,\, \text{in}\,\, \Omegai \\
+  \Le u &= 0 \,\, \text{in}\,\, \Omegae \\
+  [u] &= \ue - \ui = 0 \,\, \text{on}\,\, \Gamma \\
+[\partial_L u] &= \partiale u - \partiali u = 0 \,\,
+\text{on}\,\, \Gamma \\
+|u| &\leq C \|x \|^{-1} \,\,\text{for}\,\,\| x \|\rightarrow\infty
+\end{align}
+\end{subequations}
+
+We introduce the following auxiliary potential:
+\begin{equation}
+  h(x) =
+  \begin{cases}
+    &\int_{\mathbb{R}^3}\Gi\rhoi(y)\diff y  \quad x \in \Omegai \\
+    &\int_{\mathbb{R}^3}\Ge\rhoi(y)\diff y  \quad x \in \Omegae
+  \end{cases}
+\end{equation}
+for which we have:\footnote{This can be checked by applying the
+differential operators to the definition of $h$ and then swapping
+differentiation and integration.}
+\begin{equation}
+  \begin{cases}
+    \nabla^2 h &= \rhoi \,\, \text{in}\,\, \Omegai \\
+    \Le h &= 0 \,\, \text{in}\,\, \Omegae
+  \end{cases}
+\end{equation}
+We then define the \emph{reaction potential} as:
+\begin{equation}
+  \xi = u - h
+\end{equation}
+such that:
+\begin{subequations}
+\begin{align}
+  \nabla^2 \xi &= 0 \,\, \text{in}\,\, \Omegai \\
+  \Le \xi &= 0 \,\, \text{in}\,\, \Omegae \\
+  -[\xi] &= [h] \,\, \text{on}\,\, \Gamma \\
+  -[\partial_L \xi] &= [\partial_L h] \,\, \text{on}\,\, \Gamma
+\end{align}
+\end{subequations}
+inside the cavity, the reaction potential can be represented by a
+\emph{single layer potential}, \ie by an apparent surface charge (ASC):
+\begin{equation}
+  \xi_\mathrm{i} = \bi{S}_\mathrm{i}\sigma
+\end{equation}
+
+Applying Eqs. \eqref{eq:stat3} and \eqref{eq:stat4} of the Lemma of Integral
+Representation to the reaction potential yields:
+\begin{align}
+  \frac{1}{2}\xi_\mathrm{i} &= \bi{S}_\mathrm{i}(\partiali \xi) -
+  \bi{D}_\mathrm{i}(\xi_\mathrm{i}) \\
+  -\frac{1}{2}\xi_\mathrm{e} &= \bi{S}_\mathrm{e}(\partiale \xi) -
+  \bi{D}_\mathrm{e}(\xi_\mathrm{e})
+\end{align}
+in both cases the last term, including the integration over the entire
+subdomain volume, drops out ($f_\mathrm{i} = 0 = f_\mathrm{e}$).
+While use of Eq. \eqref{eq:stat4} on the auxiliary potential $h$ yields:
+\begin{equation}
+  -\frac{1}{2}h_\mathrm{e} =
+  \bi{S}_\mathrm{e}(\partiale h) -
+  \bi{D}_\mathrm{e}(h_\mathrm{e})
+\end{equation}
+Together with the jump conditions one obtains:
+\begin{subequations}
+  \begin{align}
+    &\bi{S}_\mathrm{i}(\partiali \xi) - \bi{D}_\mathrm{i}(\xi_\mathrm{i})
+    = \frac{1}{2}\xi_\mathrm{i} \label{eq:first} \\
+    &\bi{S}_\mathrm{e}(\partiale \xi) - \bi{D}_\mathrm{e}(\xi_\mathrm{e})
+    = -\frac{1}{2}\xi_\mathrm{e} \label{eq:second} \\
+  &\bi{S}_\mathrm{e}(\partiale h) - \bi{D}_\mathrm{e}(h_\mathrm{e})
+  = -\frac{1}{2}h_\mathrm{e} \label{eq:third} \\
+  &\xi_\mathrm{i} - \xi_\mathrm{e} = h_\mathrm{e} - h_\mathrm{i}
+  \label{eq:fourth} \\
+  &\partiali \xi - \partiale \xi = \partiale h - \partiali h
+  \label{eq:fifth}
+  \end{align}
+\end{subequations}
+
+\begin{enumerate}
+  \item Add Eqs. \eqref{eq:second} and \eqref{eq:third}:
+    \begin{equation}
+    \bi{S}_\mathrm{e}(\partiale \xi + \partiale h) -
+    \bi{D}_\mathrm{e}(\xi_\mathrm{e} + h_\mathrm{e})
+    = -\frac{1}{2}(\xi_\mathrm{e} + h_\mathrm{e})
+    \end{equation}
+  \item Exploit Eqs. \eqref{eq:fourth} and \eqref{eq:fifth} to
+    eliminate the external traces and conormal derivatives:
+    \begin{equation}
+      \bi{S}_\mathrm{e}(\partiali \xi) + \left(\frac{1}{2} -
+      \bi{D}_\mathrm{e}\right)\xi_\mathrm{i} =
+      - \bi{S}_\mathrm{e}(\partiali h) - \left(\frac{1}{2} -
+      \bi{D}_\mathrm{e}\right)h_\mathrm{i}
+    \end{equation}
+  \item Introduce the integral representation of
+    $\xi_\mathrm{i}$:
+    \begin{equation}
+      \bi{S}_\mathrm{e}(\partiali \xi) + \left(\frac{1}{2} -
+      \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}\sigma =
+      - \bi{S}_\mathrm{e}(\partiali h) - \left(\frac{1}{2} -
+      \bi{D}_\mathrm{e}\right)h_\mathrm{i}
+    \end{equation}
+  \item Reorganize Eq. \eqref{eq:first}, using the commutation
+    relation and the integral representation of $\xi_\mathrm{i}$:
+    \begin{equation}
+    \partiali \xi = \left(\frac{1}{2} + \bi{D}^\dagger_\mathrm{i}\right)\sigma
+    \end{equation}
+    such that:
+    \begin{equation}
+      \left[ \bi{S}_\mathrm{e}\left(\frac{1}{2} + \bi{D}^\dagger_\mathrm{i}\right)
+      +
+      \left(\frac{1}{2} - \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}
+      \right]\sigma =
+      - \bi{S}_\mathrm{e}(\partiali h) - \left(\frac{1}{2} -
+      \bi{D}_\mathrm{e}\right)h_\mathrm{i}
+    \end{equation}
+  \item Employ the Dirichlet-to-Neumann (DtN) map to simplify
+    the RHS. The DtN map is derived by applying Eq. \eqref{eq:stat3} to
+    the Newton potential:
+    \begin{equation}
+      \phi(x) = (\bi{N}\rhoi)(x) = \int_{\mathbb{R}^3}\Gi\rhoi(y)\diff y
+      = \left.h(x)\right|_{\Omegai}
+    \end{equation}
+    which is equal, in $\Omegai$, to the auxiliary potential $h(x)$.
+    With this definition one has:
+    \begin{equation}
+      \frac{1}{2}\phi_\mathrm{i} =
+      \bi{S}_\mathrm{i}(\partiali \phi) -
+      \bi{D}_\mathrm{i}(\phi_\mathrm{i}) + \int_{\Omegai}\Gi\rhoi(y)\diff y
+      =
+      \bi{S}_\mathrm{i}(\partiali \phi) -
+      \bi{D}_\mathrm{i}(\phi_\mathrm{i}) + \phi_\mathrm{i}
+    \end{equation}
+    eventually leading to the DtN map:
+    \begin{equation}\label{eq:regular-DtN}
+      \left(\frac{1}{2} - \bi{D}_\mathrm{i}\right)\phi_\mathrm{i}
+      +\bi{S}_\mathrm{i}(\partiali \phi) = 0
+    \end{equation}
+  \item Having identified $\phi_\mathrm{i} \equiv h_\mathrm{i}$
+    and $\partiali \phi \equiv \partiali h$ and employing the DtN map,
+    Eq.\eqref{eq:regular-DtN}, one obtains the IEF-PCM equation:
+    \begin{equation}\label{eq:regular-IEF}
+      \left[ \bi{S}_\mathrm{e}\left(\frac{1}{2} + \bi{D}^\dagger_\mathrm{i}\right)
+      +
+      \left(\frac{1}{2} - \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}
+      \right]\sigma =
+      -\left[\left(\frac{1}{2}-\bi{D}_\mathrm{e}\right)
+      -\bi{S}_\mathrm{e}\bi{S}_\mathrm{i}^{-1}
+       \left(\frac{1}{2}-\bi{D}_\mathrm{i}\right)
+      \right]\phi
+    \end{equation}
+    where the subscript $\mathrm{i}$ for the Newton potential $\phi$ has
+    been dropped, given that it is continuous across the boundary
+    $\Gamma$.
+\end{enumerate}
+The above proves the existence of a representation of the reaction
+potential valid \emph{inside} the cavity in terms of an apparent surface
+charge density. We do not prove the uniqueness of this solution and
+refer the interested reader to the original paper.~\autocite{Cances1998}
+
+In the code, the Green's functions are defined \emph{without} the $4\pi$
+factors in the denominator. Instead of $\frac{1}{2}$ as coefficients to
+the identity operator, one will then have $2\pi$:
+\begin{equation}\label{eq:regular-IEF-2pi}
+  \left[ \bi{S}_\mathrm{e}\left(2\pi + \bi{D}^\dagger_\mathrm{i}\right)
+  +
+  \left(2\pi - \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}
+  \right]\sigma =
+  -\left[\left(2\pi-\bi{D}_\mathrm{e}\right)
+  -\bi{S}_\mathrm{e}\bi{S}_\mathrm{i}^{-1}
+  \left(2\pi-\bi{D}_\mathrm{i}\right)
+  \right]\phi
+\end{equation}
+
+\paragraph{Isotropic IEF}
+From Eq. \eqref{eq:regular-IEF-2pi} we can derive the isotropic IEF,
+the DPCM and the CPCM (by letting $\diel\rightarrow\infty$) equations.
+For a homogeneous, isotropic dielectric with permittivity $\diel$ the
+Green's function outside the cavity is simply given by scaling the one
+inside:
+\begin{equation}
+  \Ge = \frac{1}{\diel|x-y|} = \frac{1}{\diel} \Gi = \frac{1}{|x-y|}
+\end{equation}
+leading to the boundary integral operators:
+\begin{alignat}{2}
+  \bi{S}_\mathrm{i} = \diel\bi{S}_\mathrm{e} = \bi{S}, \quad&
+  \bi{D}_\mathrm{i} = \bi{D}_\mathrm{e} = \bi{D}
+\end{alignat}
+Eventually, one has:
+\begin{equation}
+  \left[ 2\pi \left(\frac{\diel+1}{\diel-1}\right)
+  - \bi{D} \right]\bi{S}\sigma
+  =
+  -\left[2\pi-\bi{D}\right]\phi
+\end{equation}
 

### 2016-08-24 Chapters/PaperSummary.tex
Add some more stuff from notes scattered around

@@
 \centering
-\scalebox{0.8}{\input{gfx/relapcm/algorithm.tikz}}
+\scalebox{0.8}{\input{gfx/algorithm.tikz}}
 \caption[Modular approach to programming a \acs{PCM} functionality into an existing \acs{SCF} code.]{
@@
 \end{figure}
+
+\todo[inline]{Have a closer look at this. I think it needs to be updated for the discussion
+on open source software that went down on JPC Lett. some time ago.
+Also, acronyms should be properly set up and reference to stuff not included in the thesis
+(time-dependent) removed.}
+It should also be mentioned that this has been accepted (in some partial/modified form)
+for publication as a proceeding:
+\begin{enumerate}
+\item
+  \textbf{R. Di Remigio}, T. D. Crawford, and L. Frediani
+  \\
+  PCMSolver: an Application Programming Interface
+  for the Polarizable Continuum Model
+  \\
+  Accepted for publication in the proceedings of the
+  \emph{Producing High Performance and Sustainable Software for
+  Molecular Simulation} workshop held at the 2015 Supercomputing
+  Conference
+\label{SC15}
+\end{enumerate}
+
+The growing complexity of quantum chemical program packages requires
+that an appropriate strategy be devised to implement new features.
+Scalability is of paramount importance, but it has become clear
+that maintainability and extensibility of the code play an equally
+important role in managing this growing complexity \cite{Wilson2014}.
+
+The use of a \emph{modular programming paradigm} has been recognized as
+beneficial in many other scientific computing contexts. New features are
+isolated into libraries that can be accessed by host programs through a
+well-defined \emph{application programming interface} (API).
+Computational tasks are thus implemented into separate, independent and
+interchangeable modules. These are \emph{developed} and \emph{tested}
+independently of the particular host program in which they will be used
+\cite{Dijkstra1968, Parnas1972}.  A well-defined API clearly delimits
+the boundaries of the functionality offered, thus forcing a programming
+style and standardisation of the functionality, eventually.
+
+When coupled with \emph{open-source} licensing, modularity is a proper
+step towards ensuring reproducibility of results from scientific
+simulation software \cite{Gezelter2015}.
+Open-source development can fully leverage the benefits
+of widespread, cloud-based, free, code development services, such as
+hosted distributed version control systems (DVCSs), continuous integration, code
+coverage, static and dynamic code analyses, nightly regression testing,
+public issue tracking, code review and so forth:
+adoption of a modern code development workflow is easily within reach.
+
+It is, of course, true that the above mentioned services are not
+exclusive prerogatives of open-source projects.
+However, an open review process of scientific software can often help
+to establish
+reproducibility, extensibility and sustainability of the software
+ecosystem.
+Open-source software, modular development of new functionalities
+and full-fledged exploitation of DVCSs ensure a much larger scientific
+impact. Third-parties can easily contribute to
+the project: by improving the documentation, by reporting bugs or
+by actively extending the codebase with new functionality.
+
+The description of chemical phenomena in complex environments, such as
+solutions or proteins, is as challenging as it is important
+for an accurate prediction of a wide range of properties
+\cite{Reichardt2010}. Quantum chemistry is nowadays a valuable
+tool: its simulations can provide significant insight into complex
+chemical phenomena \cite{Nobel1998, Nobel2013}.
+The major challenge lies in providing a faithful, realistic, yet cost-effective
+description of the environment of the molecule:
+a molecule in solution can be represented by a quantum mechanical model
+comprising a great number of atoms, thousands or more. Such a simulation would,
+however, be impossible to carry out with the current state-of-the-art
+computational resources.
+Models must be devised to overcome the dimensionality ``disease'' and
+the idea of introducing different approximation levels for different
+parts of the system under study is particularly appealing, as it helps
+our chemical understanding. This is the idea at the basis of
+\emph{multiscale} models.
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=.45\textwidth]{multiscale.jpg}
+\end{figure}
+
+\emph{Continuum} (or \emph{implicit}) models, deal explicitly
+only with the degrees of freedom associated with the solute, while
+replacing the solvent with a structureless continuum characterized by
+its bulk properties \cite{Onsager1936}. Among these models, the \acl{PCM} (\acs{PCM})
+is the most widespread and successful \cite{Tomasi2005, mennucci}.
+
+The \acs{PCM} is also an ideal candidate for the creation of a solvation API.
+Its input from and output to \emph{any} quantum chemistry code is
+limited and well-defined.
+This provides a natural API design: the API functions can be compared
+\emph{vis-à-vis} with the working equations derived for the different
+quantum chemical methods.
+
+We will elucidate the design principles adopted in the development of our
+\pcmsolver{} library: an \emph{open-source} API for the inclusion of the
+PCM in \emph{any} quantum chemistry code \cite{PCMSolver}.  Use of our
+API \emph{significantly} limits coding effort on the side of the host:
+continuum solvation at the self-consistent field
+(SCF) level of theory can be implemented in the host program almost out-of-the-box.  The module is
+agnostic of the host code internals and is tested separately from it.
+
+We will emphasize the role of the open-source licensing model and the
+importance of the modern workflows adopted in our project.  The adoption
+of \git as distributed version control system, \cmake for cross-platform builds and automatic
+documentation deployment on \readthedocs simplify extensibility of the
+module and promote third-party contributions to the code base, while
+continuous integration and nightly testing offer an invaluable level of
+confidence in the code.
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=.5\textwidth]{pcmsolver-scheme}
+\end{figure}
+
+\pcmsolver currently offers the basic PCM functionality already
+available in many widespread software packages, such as traditional
+collocation solvers for isotropic media \cite{Tomasi2005}.  Some unique
+functionalities are the wavelet Galerkin solvers on smooth molecular
+surfaces \cite{Weijo2010, Bugeanu2015}, the Green’s
+functions for spherical diffuse interfaces \cite{Mozgawa2015}
+and the delayed, time-dependent (TD) solvers for real-time TDSCF
+simulations \cite{Corni2015}.
+
+Careful usage of the C++ object-oriented paradigm is the key to this
+wide spectrum of functionalities. The library is in itself made of
+modules, communicating by means of composition at the outermost
+level of the design.
+This allows us to mix languages internally: we currently have Fortran
+and C modules alongside the main C++ structure.
+Library internal classes balance the dynamic polymorphism offered by
+inheritance and the static polymorphism offered by template programming
+\cite{Alexandrescu2001, Sutter2004}.
+Preeminent use of composition over inheritance keeps the coupling
+between different modules as low as possible \cite{Gamma1994}.
+
+As a final example of the power of our modular strategy, we will present
+our recent work on different quantum chemistry program packages,
+tackling both uncorrelated and correlated wave functions \cite{DALTON,
+PSI4} and also including relativistic effects \cite{DIRAC14brief,
+DiRemigio2014}.  In particular, our recent work-in-progess on a
+variational reformulation of the polarizable continuum model--coupled
+cluster (PCM-CC) algorithm \cite{Cammi2009, Lipparini2011, Caricato2010a} within the
+\psicode{} program package will be presented with emphasis on possible
+applications to the prediction of the optical properties of chiral
+molecules in solution.
 

### 2016-08-24 FrontBackmatter/Notation.tex
Add some more stuff from notes scattered around

@@
  $\vect{r}, \vect{r}^\prime, \ldots$ & position vectors in $\mathbb{R}^3$ \\
- $\vect{s}, \vect{s}^\prime, \ldots$ & position vectors in a closed subset of $\mathbb{R}^2$ \\
+ $\vect{s}, \vect{s}^\prime, \ldots$ & position vectors on a closed surface \\
  $\vect{v}, \vect{q}, \ldots$ & vectors in an arbitrary vector space \\

### 2016-08-24 FrontBackmatter/Publications.tex
Add some more stuff from notes scattered around

@@
 
-\begin{enumerate}[label=\textbf{\Roman{*} },ref=\Roman{*}]
+\begin{enumerate}[label=\textbf{\Roman{*} }, ref=\Roman{*}]
 
@@
 \end{enumerate}
-
-The following proceeding is also part of the thesis:
-\begin{enumerate}
-\item
-  \textbf{R. Di Remigio}, T. D. Crawford, and L. Frediani
-  \\
-  PCMSolver: an Application Programming Interface
-  for the Polarizable Continuum Model
-  \\
-  Accepted for publication in the proceedings of the
-  \emph{Producing High Performance and Sustainable Software for
-  Molecular Simulation} workshop held at the 2015 Supercomputing
-  Conference
-\label{SC15}
-\end{enumerate}

### 2016-08-26 Chapters/ContinuumSolvationModels.tex
Update notations

@@
     \item \ac{BEM} and wavelet \acs{BEM} I think it's necessary to
-      mention also ddCOSMO and ddPCM as alternative strategies (put a
-      reference)
+      mention also ddCOSMO and ddPCM as alternative strategies
+      ~\autocite{Cances2013-jg, Lipparini2013-cy, Lipparini2014-to,
+      Lipparini2014-fq, Stamm2016-fv}
     \item variational formulation (no derivations, just mention the
@@
 
-\section{Boundary integral equations}\label{sec:BIE}
+\section{Continuum Electrostatics as a Boundary Integral Problem}\label{sec:IEF}
 
-\subsection{Transmission problem}
 We assume Euclidean space $\mathbb{R}^3$ to be partitioned into two
-subdomains $\Omegai$ and $\Omegae$ sharing a boundary $\Gamma$.
-We further assume that $\Omegai$ is a closed domain, entirely contained inside $\Omegae$.
+subdomains $\Omegai$ and $\Omegae$ sharing a boundary
+$\Gamma\equiv\partial\Omegai$.
+We further assume that $\Omegai$ is a closed domain, entirely contained
+inside $\Omegae$. The boundary $\Gamma$ is thus a closed surface, \ie~a
+closed 2-manifold in $\mathbb{R}^3$.
 The transmission problem is posed as follows:
@@
 \begin{subequations}\label{eq:transmission}
-\begin{align}
+  \begin{align}[left={\empheqlbrace}]
   \Li u &= f_\mathrm{i} \,\, \text{in}\,\, \Omegai \label{eq:internal} \\
@@
 notational simplicity, will use the symbols $\partiale$ and $\partiali$
-for the latter and only give it in explicit form when needed. Please
-refer to the book by \citeauthor{Sauter2011} for further details. The
+for the latter and only give it in explicit form when needed.
+Further mathematical details and technical results on the definition of
+function traces and their use in setting up the proper normal and
+conormal derivatives can be found in the excellent book by
+\citeauthor{Sauter2011-an}.~\autocite{Sauter2011-an}.
+The
 fundamental solutions, or Green's functions, for the elliptic
@@
 
-\subsection{Boundary integral operators}
-We introduce the relevant boundary integral operators for the abovementioned
-transmission problem. They are three of the four components of the Calder\'on projector.
-For $v \in L^2(\Gamma)$, $x, y \in \Gamma$:\footnote{We can assume a less restrictive functional space, \ie a Sobolev space of fractional order.}
+We introduce the relevant boundary integral operators for the
+abovementioned transmission problem. They are three of the four
+components of the Calder\'on projector.
+For $v \in L^2(\Gamma)$, $x, y \in \Gamma$:
+\footnote{We can assume a less restrictive functional space, \ie a
+Sobolev space of fractional order.}
 \begin{equation}
@@
   The integral operators introduced above enjoy the following
-  properties~\autocite{Hsiao2008, Sauter2011}:
+  properties~\autocite{Hsiao2008-xb, Sauter2011-an}:
   \begin{enumerate}
@@
 
-\pagebreak
-\section{IEFPCM}\label{sec:IEF}
-
 Note that the figures refer to cases where the dielectric is supposed to
@@
 
-%%%% CASE 0
-
-\subsection{Case 0: sources in a vacuum cavity inside a dielectric}
-
 \begin{figure}[h!]
   \centering
-  \includegraphics[width=.65\textwidth]{case0}
+  \includegraphics[width=.65\textwidth]{IEF}
   \caption{Vacuum Cavity in dielectric with permittivity $\diel_\mathrm{e}>1$.
   Source charge density $\rho_\mathrm{i}$ inside the cavity
-  ($\diel_\mathrm{i}=1$). \textcolor{red}{Regular IEF}}
+  ($\diel_\mathrm{i}=1$). \textcolor{red}{Regular \acs{IEF}}}
   \label{fig:case0}
@@
 
-This is the regular IEF as derived in:~\autocite{Cances1998}
+This is the regular \acs{IEF} as derived in:~\autocite{Cances1998-og}
 \begin{subequations}
@@
     \end{equation}
-  \item Employ the Dirichlet-to-Neumann (DtN) map to simplify
-    the RHS. The DtN map is derived by applying Eq. \eqref{eq:stat3} to
+  \item Employ the \ac{DtN} map to simplify
+    the RHS. The \acs{DtN} map is derived by applying Eq. \eqref{eq:stat3} to
     the Newton potential:
@@
     \end{equation}
-    eventually leading to the DtN map:
+    eventually leading to the \acs{DtN} map:
     \begin{equation}\label{eq:regular-DtN}
@@
   \item Having identified $\phi_\mathrm{i} \equiv h_\mathrm{i}$
-    and $\partiali \phi \equiv \partiali h$ and employing the DtN map,
-    Eq.\eqref{eq:regular-DtN}, one obtains the IEF-PCM equation:
-    \begin{equation}\label{eq:regular-IEF}
+    and $\partiali \phi \equiv \partiali h$ and employing the \acs{DtN} map,
+    Eq.\eqref{eq:regular-DtN}, one obtains the \acs{IEF}-\acs{PCM} equation:
+    \begin{equation}\label{eq:full-IEF}
       \left[ \bi{S}_\mathrm{e}\left(\frac{1}{2} + \bi{D}^\dagger_\mathrm{i}\right)
@@
 charge density. We do not prove the uniqueness of this solution and
-refer the interested reader to the original paper.~\autocite{Cances1998}
+refer the interested reader to the original
+paper.~\autocite{Cances1998-og}
 
@@
 the identity operator, one will then have $2\pi$:
-\begin{equation}\label{eq:regular-IEF-2pi}
+\begin{equation}\label{eq:full-IEF-2pi}
   \left[ \bi{S}_\mathrm{e}\left(2\pi + \bi{D}^\dagger_\mathrm{i}\right)
@@
 
-\paragraph{Isotropic IEF}
-From Eq. \eqref{eq:regular-IEF-2pi} we can derive the isotropic IEF,
-the DPCM and the CPCM (by letting $\diel\rightarrow\infty$) equations.
+From Eq. \eqref{eq:full-IEF-2pi} we can derive the isotropic
+\acs{IEF}, the \acs{DPCM} and the \acs{CPCM} (by letting
+$\diel\rightarrow\infty$) equations.
 For a homogeneous, isotropic dielectric with permittivity $\diel$ the
@@
 Eventually, one has:
-\begin{equation}
+\begin{equation}\label{eq:isotropic-IEF}
   \left[ 2\pi \left(\frac{\diel+1}{\diel-1}\right)
@@
 
+\section{Detour and Teaser: Weak Formulation of Partial Differential Equations}\label{sec:weak}
+
+From the strong to the weak formulation of \acp{PDE}.\autocite{Ern2004-oo}
+
+\begin{itemize}
+  \item Poisson's equation with conductor boundary conditions
+        % This is COSMO!
+        % This equation makes sense iff \Psi \in \mathcal{C}^2_0(C) (at least)
+        % \Psi = 0 \quad \forall \vect{r}\in\partial C
+        \begin{equation*}
+            \nabla^2\Psi = -4\pi\rho
+            % Dirichlet boundary conditions for a conductor
+          \quad \Psi \in \mathcolor{ForestGreen}{\mathcal{C}^2_0(C)}
+        \end{equation*}
+  \item Project onto the space of test functions
+        % Project onto the space of test functions
+        \begin{equation*}
+          \left\{
+          \begin{aligned}
+            &\text{Seek $\Psi\in\mathcolor{ForestGreen}{V_0}$ such that:}\\
+            &(\zeta, \nabla^2\Psi) =
+            -4\pi(\zeta, \rho) \quad
+            \forall
+            \zeta\in\mathcolor{BurntOrange}{\mathcal{C}^\infty_0(C)}
+          \end{aligned}
+          \right.
+        \end{equation*}
+  \item Use
+    $\eta\nabla^2\gamma = \nabla\cdot(\eta\nabla\gamma) - \nabla\eta\cdot\nabla\gamma$
+        % The integral of the first term on the RHS vanishes.
+        % It can be shown using Gauss' Theorem and the conductor
+        % boundary conditions.
+        \begin{equation*}
+          \left\{
+          \begin{aligned}
+            &\text{Seek $\Psi \in \mathcolor{ForestGreen}{H_0^1(C)}$ such that:}\\
+            &\mathcolor{Red}{(\nabla\zeta, \nabla\Psi)} =
+            \mathcolor{Blue}{-4\pi(\zeta, \rho)} \quad
+            \forall \zeta\in \mathcolor{BurntOrange}{H_0^1(C)}
+          \end{aligned}
+          \right.
+        \end{equation*}
+\end{itemize}
+\begin{equation*}
+  \Psi\in \mathcolor{BurntOrange}{H_0^1(C)} = \lbrace f : C \rightarrow
+  \mathbb{R} | f \in L^2(C), \nabla f \in L^2(C), f = 0 \forall \vect{s}
+  \in\partial C \rbrace
+\end{equation*}
+\begin{equation}
+  \left\{
+  \begin{aligned}
+    &\text{Seek $\Psi \in V$ such that:}\\
+    &\forall \zeta \in V \,\,\,\,\,
+    \mathcolor{Red}{a(\zeta, \Psi)} = \mathcolor{Blue}{b(\zeta)}
+  \end{aligned}
+  \right.
+  \label{eq:weak}
+\end{equation}
+
+% Hadamard well-posedness definition:
+% Problem \eqref{eq:weak} is well-posed if it admits one and only
+% one solution and the solution is bounded by the initial data, by
+% an a priori estimate (see Ern&Guermond, p.82, Def. 2.1)
+\begin{defin}[Well-posedness]\label{def:hadamard}
+  The weak problem Eq. \eqref{eq:weak} is \emph{well-posed} if it
+  admits one and only one solution and the solution is bounded by
+  the initial data by an \emph{a priori} estimate.
+\end{defin}
+\begin{defin}[Continuity]\label{def:continuity}
+  A bilinear form on a normed vector space $V$ is \emph{bounded}, or
+  \emph{continuous}, if there is a constant $C$ such that
+  $\forall u, v \in V$:
+  \[
+  a(u, v) \leq C\lVert u\rVert\lVert v\rVert
+  \]
+\end{defin}
+\begin{defin}[Coercivity]\label{def:coercivity}
+  A bilinear form on a normed vector space $V$ is \emph{coercive},
+  or \emph{elliptic}, if there is a constant $\alpha >0$ such that
+  $\forall u \in V$:
+  \[
+  a(u, u) \geq \alpha\lVert u\rVert^2
+  \]
+\end{defin}
+This implies that no eigenvalue of the linear operator associated
+to the bilinear form can be zero, hence that the operator is
+invertible!!~\autocite{Ern2004-oo}
+
+% Maybe present connection with Steepest Descent and Conjugate
+% Gradient?
+% Introduce a suitable vector space, equipped with a norm
+% derived from a scalar product.
+% The solution belongs to such a vector space, which is
+% bigger that in the strong formulation (less constraints on
+% the regularity of the solution)
+\begin{equation}
+  \left\{
+  \begin{aligned}
+    &\text{Seek $\Psi \in V$ such that:}\\
+    &\forall \zeta \in V \,\,\,\,\,
+    \mathcolor{Red}{a(\zeta, \Psi)} = \mathcolor{Blue}{b(\zeta)}
+  \end{aligned}
+  \right.
+\end{equation}
+\begin{lemma}[Lax--Milgram]
+  If the bilinear form $a$ is \emph{continous} and \emph{coercive}
+  in $V$, then, for any continuous linear form $b$, Problem~\eqref{eq:weak} is
+  \emph{well-posed}.
+\end{lemma}
+\begin{corollary}[Variational property]
+  If the bilinear form is \emph{symmetric} and \emph{positive}
+  the unique solution to Problem \eqref{eq:weak} is
+  the \emph{unique minimum} on $V$ of the functional:
+  \[
+  \mathcal{F}(\zeta) = \frac{1}{2}a(\zeta, \zeta) - b(\zeta)
+  \]
+\end{corollary}
+
+Weak formulation of \acs{BIE}s.\autocite{Hsiao2008-xb}
+
+\section{Numerical Approaches to Boundary Integral Equations}\label{sec:BEM}
+
+Approximability, conformity, consistency and asymptotic consistency
+of the approximation setting, $\inf$-$\sup$ condition, Céa's Lemma
+and generalized $\inf$-$\sup$ condition.~\autocite{Ern2004-oo,
+TesiFilippo}
+Additional cavity generators
+TsLess~\autocite{Pomelli2004-lb}
+DefPol~\autocite{Pomelli1998-ob, Pomelli1999-dc}
+FIXPVA~\autocite{Su2009-em}
+Other discretization schemes
+CSC~\autocite{York1999-xy, Scalmani2010-tw}
+SWIG~\autocite{Lange2010-jp, Lange2010-qo, Lange2011-eu}
+
+\todo[inline]{CURRENTLY COPIED VERBATIM FROM THE SPHERICAL GREEN'S
+FUNCTION PAPER}
+
+The numerical solution of boundary integral equations, such as the
+\acs{IEF}-\acs{PCM}
+Eqs.~\eqref{eq:full-IEF-2pi}-\eqref{eq:isotropic-IEF}, can be achieved
+by the application of the \ac{BEM}.
+A discretization of the molecular cavity
+boundary is necessary.~\autocite{Pomelli2007-wq}
+The discretization of the
+molecular surface into finite elements leads to a corresponding
+discretization of both the integral operators and the functions in
+Eqs.~\eqref{eq:full-IEF-2pi}-\eqref{eq:isotropic-IEF}.
+
+Different strategies are available to perform the
+discretization: the \emph{collocation} or the \emph{Galerkin}
+approach.~\cite{Hackbusch1995-uq, Ern2004-oo}
+The final result is the transformation of
+the original integral equation to a system of linear equations whose
+dimension relates to the number of finite elements generated by
+discretizing the boundary.  For general integral equations and
+discretization schemes, a dense system matrix is obtained. Moreover, the
+discretization scheme might not, in general, preserve the symmetry
+property of the original integral equation.
+
+In this paper we apply the traditional centroid collocation
+method~\autocite{Tomasi2005-vm, Pomelli2007-wq} on the cavity boundary
+discretization obtained \emph{via} the GePol
+algorithm.~\autocite{Pascual-Ahuir1987-uo, Pascual-Ahuir1990-lp,
+Pomelli1998-qp, Pomelli2001-sj, Frediani2004-ua, Pomelli2007-wq}
+In this algorithm, the cavity is made of interlocking spheres, not all
+of them necessarily centered on the atoms. The surface is then
+discretized by means of $N_\mathrm{ts}$ small triangles $\Delta_i$ whose
+centroid defines the \emph{collocation point}. Once the centroid is
+properly defined, functions and operators are discretized by applying a
+one-point quadrature rule in the finite element centroid.
+The \acs{IEF}-\acs{PCM} equation is thus recast as:
+\begin{equation}
+  \vect{q}
+  =
+  \mat{T}^{-1}\mat{R}\vect{v}.
+\end{equation}
+$\vect{q}$ and $\vect{v}$ are vectors of dimension $N_\mathrm{ts}$
+collecting the values of the \acs{ASC} and \ac{MEP}
+at the collocation points.
+The matrices representing the boundary integral operators are:
+\begin{align}
+  \mat{T} &=
+  \left(2\pi\mat{I} - \mat{D}_\mathrm{e}\mat{A}\right)\mat{S}_\mathrm{i}
+  +\mat{S}_\mathrm{e}\left(2\pi\mat{I} +
+  \mat{A}\mat{D}_\mathrm{i}^\dagger\right) \\
+  \mat{R} &=
+  \left(2\pi\mat{I} - \mat{D}_\mathrm{e}\mat{A}\right) -
+  \mat{S}_\mathrm{e}\mat{S}^{-1}_\mathrm{i}\left(2\pi\mat{I}-\mat{D}_\mathrm{i}\mat{A}\right)
+\end{align}
+$\mat{I}$ is the $N_\mathrm{ts}$-dimensional identity matrix while the
+other matrices are defined as:
+\begin{alignat}{2}
+  &\mat{A} = \mathrm{diag}(\cdots\,\, a_i = \mathrm{area}(\Delta_i)\,\, \cdots) \\
+  &S_{ii, \mathrm{i}} = k\sqrt{\frac{4\pi}{a_i}}, \quad
+  S_{ij, \mathrm{i}} = \frac{1}{|\vect{s}_i-\vect{s}_j|}\quad (i \neq j) \\
+  &D_{ii, \mathrm{i}} = -k\sqrt{\frac{\pi}{a_i}}\frac{1}{R_i}, \quad
+  D_{ij, \mathrm{i}} =
+  \frac{(\vect{s}_i-\vect{s}_j)\cdot\vect{n}_j}{|\vect{s}_i-\vect{s}_j|^3}\quad (i \neq j) \\
+  &S_{ii, \mathrm{e}} = \frac{S_{ii, \mathrm{i}}}{C(\vect{s}_i, \vect{s}_i)} +
+        G_\mathrm{img}(\vect{s}_i, \vect{s}_i), \quad
+  S_{ij, \mathrm{e}} =
+  \frac{1}{C(\vect{s}_i, \vect{s}_j)|\vect{s}_i-\vect{s}_j|} +
+  G_\mathrm{img}(\vect{s}_i, \vect{s}_j) \quad (i \neq j) \\
+  &D_{ii, \mathrm{e}} = \diel(\vect{s}_i)\left[
+  \frac{D_{ii, \mathrm{i}}}{C(\vect{s}_i, \vect{s}_i)}
+  -S_{ii, \mathrm{i}}
+  \frac{\nabla C(\vect{s}_i,
+  \vect{s}_i)\cdot\vect{n}_i}{C^2(\vect{s}_i, \vect{s}_i)}
+  +
+  \nabla G_\mathrm{img}(\vect{s}_i, \vect{s}_i)\cdot\vect{n}_i
+  \right] \\
+  &D_{ij, \mathrm{e}} =
+  \diel(\vect{s}_j)\left[
+  \frac{1}{C(\vect{s}_i, \vect{s}_j)}
+  \frac{(\vect{s}_i-\vect{s}_j)\cdot\vect{n}_j}{|\vect{s}_i-\vect{s}_j|^3}
+  -\frac{\nabla C(\vect{s}_i,
+  \vect{s}_j)\cdot\vect{n}_j}{C^2(\vect{s}_i, \vect{s}_j)}
+  \frac{1}{|\vect{s}_i-\vect{s}_j|}
+  + \nabla G_\mathrm{img}(\vect{s}_i, \vect{s}_j)\cdot\vect{n}_j
+  \right]\quad (i \neq j)
+\end{alignat}
+in terms of the centroids $\vect{s}_i, \vect{s}_j$, the finite elements
+areas and their curvatures $R_i$. The factor $k=1.07$ is introduced to achieve a better precision on the diagonal elements~\cite{Tomasi2005}.
+
+\todo[inline]{CURRENTLY COPIED VERBATIM FROM THE WAVELET PAPER}
+
+Boundary integral equations, such as the \acs{IEF}-\acs{PCM}
+Eqs.~\eqref{eq:full-IEF-2pi}-\eqref{eq:isotropic-IEF},
+can conveniently be solved numerically by the application of the
+\ac{BEM}.
+Both the integral operators and the functions on which these act are
+defined solely on the boundary of the cavity.
+
+The application of the boundary element method requires that the boundary of
+the molecular cavity is discretized by a number of suitable finite elements.
+The discretization of the boundary leads to a discretization of the integral
+operators. This discretization can be carried out by using either the
+\emph{collocation} or the \emph{Galerkin}
+approach.~\autocite{Hackbusch1995-uq, Ern2004-oo}
+In both cases, the integral equation is transformed to a system of
+linear equations whose dimension is related to the number of finite
+elements used in the discretization of the boundary.
+The resulting system matrix is, in general, a dense matrix. The
+boundary element method thus suffers from limitations imposed by the
+number of matrix elements to be stored and the memory and time
+requirements of solving the resulting linear system.
+
+The use of a wavelet basis in the Galerkin approach has been proven
+beneficial in this respect.~\autocite{Harbrecht2004-uo, Dahmen2006-pj,
+Harbrecht2006-ug}.
+The resulting system matrices are quasi-sparse
+and can be further reduced to a sparse form by discarding
+negligible entries without considerable loss of precision.
+
+The wavelet boundary element method starts by defining
+a sequence of hierarchical trial spaces,
+spanned by standard finite element ansatz functions:
+\begin{equation}
+  \{0\} = V_{-1} \subset V_0 \subset V_1 \subset \ldots
+  \subset V_J, \qquad V_j = \Span{\phi_{j,k} : k\in \Delta_{j}}.
+\end{equation}
+Here, $\Delta_j$ is an index set for the
+single-scale basis of the space $V_j$. In the wavelet method,
+the trial space $V_j$ is split into the direct sum $V_j = V_{j-1}
+\oplus W_j$. The resulting complementary space $W_j$ is called the
+wavelet space and is not necessarily orthogonal to $V_{j-1}$.
+% It suffices the equation $W_j \cap V_{j-1} = \{0\}$,
+Recursive splitting of the trial spaces leads to the wavelet
+decomposition $V_j = \bigoplus_{l = 0}^{j} W_l$.
+
+The complementary space is spanned by the wavelet basis:
+\begin{equation} W_j =
+  \Span{\psi_{j,k}: k \in \Delta_{j+1}
+\setminus \Delta_{j}}.
+\end{equation}
+The choice of this basis turns
+out to be very convenient, since we can exploit the compression
+technique described by \citeauthor{Dahmen2006-pj} to build up the sparse
+system matrix in the wavelet basis directly, avoiding the computation of
+non-relevant matrix entries.~\autocite{Dahmen2006-pj}
+The \emph{a priori} compression is carried out in two steps. In the
+first step, contributions from basis functions sufficiently far away
+from each other are discarded. In the following, elements of the matrix
+where one wavelet is in the smooth part of the other one are ignored.
+The number of relevant matrix entries scales proportional to $O(N_J)$ with
+$N_J$ the number of degrees of freedom for a refinement of the geometry
+up to level $J$.
+
+In order to understand the compression steps, let us introduce some
+notation. Let the \emph{convex hull} of the support of the wavelet $\psi_{j,k}$
+be:
+\begin{equation}
+  \Omega_{j,k} = \conv\hull\left(\supp \psi_{j,k}\right).
+\end{equation}
+Moreover, let the \emph{singular support} of the wavelet $\psi_{j,k}$ be:
+\begin{equation}
+  \Omega'_{j,k} = \sing\supp\psi_{j,k}
+\end{equation}
+It contains all the points in the support where the wavelet is not smooth.
+The compression steps are then governed by the following set of
+equations
+\begin{equation}
+  \left[A_J\right]_{\left(j,k\right)\left(j',k'\right)} =
+  \begin{cases}
+    \quad\quad 0 & \dist(\Omega_{j,k}, \Omega_{j',k'}) > B_{j,j'} \text{ and }j,j' \geq 0, \\
+    \quad\quad 0 & \dist(\Omega_{j,k}, \Omega_{j',k'}) \lesssim 2^{-\min\left\{j,j'\right\}} \,\text{and} \\
+      & \quad \dist(\Omega'_{j,k}, \Omega_{j',k'}) > B'_{j,j'} \text{ if }j' > j \geq 0, \\
+      & \quad \dist(\Omega_{j,k}, \Omega'_{j',k'}) > B'_{j,j'} \text{ if }j > j' \geq 0, \\
+    \langle A\psi_{j',k'},\psi_{j,k}\rangle, & \text{otherwise}
+  \end{cases}
+\label{eq:compression}
+\end{equation}
+where $\dist(\cdot, \cdot)$ denotes the distance, either between
+the bounding boxes of the wavelets or between the singular support and the
+bounding box. The first and second conditions represent the first and second
+compression step, respectively. The parameters $j,j'$ are the
+levels of the wavelets under consideration and the
+level-dependent cut-off parameters $B_{j,j'}$ and $B'_{j,j'}$ are given
+by
+\begin{align*}
+  B_{j,j'} &= a \max
+  \left\{2^{-\min\left\{j,j'\right\}},
+  2^{\frac{J\left(2d^\prime-op\right) -
+  \left(j+j'\right)\left(d^\prime+\tilde{d}\right)}{\left(2\tilde{d}+op\right)}}\right\}
+  \text{ and }\\ B'_{j,j'} &= a \max
+  \left\{2^{-\min\left\{j,j'\right\}},
+  2^{\frac{J\left(2d^\prime-op\right) -
+\left(j+j'\right)d^\prime-\max\left\{j,j'\right\}\tilde{d}}{\left(\tilde{d}+op\right)}}\right\}
+\end{align*}
+with $op$ being the order of the integral operator under consideration.
+For the first kind integral equation $op = -1$, while $op = 0$ for the second
+kind integral equation. The integer $\tilde{d}$ is related to the vanishing
+moments of the wavelet basis:
+\begin{equation}
+  \int \diff\vect{x}\,\,\vect{x}^r \psi_{j,k}(\vect{x}) = 0,
+\quad r = 0, \ldots \tilde{d}-1
+\end{equation}
+In the particular implementation, $\tilde{d}=3$ for the piecewise
+constant wavelet basis,  while $\tilde{d}=4$ for the piecewise bilinear
+parametrization.~\autocite{Harbrecht2013-hb}
+The compression can thus be adjusted by the parameters $a$ and
+$d^\prime$:
+\begin{equation}\label{eq:param}
+  a \geq 1, \quad d < d^\prime < \tilde{d} + op
+\end{equation}
+where $d$ is the approximation order of the trial spaces $V_j$:
+$d=1, 2$ for piecewise constant and
+piecewise bilinear ansatz functions, respectively.
+Note that, if the interaction between two wavelets, $\psi_{j,k}$ and $\psi_{j',k'}$, on
+level $j$ and $j'$ is neglected in the system matrix, all other
+interactions between wavelets resulting from the refinement of
+$\psi_{j,k}$ and $\psi_{j',k'}$ can also be ignored. Thus, the
+compression pattern of the system matrix is calculated
+hierarchically starting from the coarsest level.
+
+Once the compressed system matrix is assembled, we arrive at a sparse
+system matrix in the wavelet basis which can be compressed further by
+leaving out the sufficiently small elements. This post-processing step
+is governed by the rule
+\begin{equation}
+  \left[A_J\right]_{\left(j,k\right)\left(j',k'\right)}
+  \begin{cases}
+    \quad\quad 0 & \text{if } |\left[ A_J \right]_{\left(j,k\right)\left(j',k'\right)}| \leq \epsilon_{j,j'}\\
+    \left[A_J\right]_{\left(j,k\right)\left(j',k'\right)}, & \text{otherwise,}
+  \end{cases}
+  \label{eq:aposteriori}
+\end{equation}
+where the coefficients $\epsilon_{j,j'}$ are given by
+\begin{equation}
+%\epsilon_{j,j'} = b\min \left\{2^{-|j-j'|}, 2^{-\left(2J-\left(j+j'\right)\right)\frac{2d^\prime-op}{2\tilde{d}+op}}\right\} 2^{J\,op}2^{-2d^\prime\left(J-\frac{j+j'}{2}\right)},
+  \epsilon_{j,j'} = b\min \left\{2^{-|j-j'|}, 2^{-\left(2J-\left(j+j'\right)\right)\frac{2d^\prime-op}{2\tilde{d}+op}}\right\} 2^{-2d^\prime\left(J-\frac{j+j'}{2}\right)}
+\end{equation}
+with the \emph{a posteriori} compression parameter $b < 1$.
+
 \section[Variational Formulation of Classical Polarizable Models]{
-A Unifying View of Classical Polarizable Models within a Variational Formulation}
+Classical Polarizable Models within a Variational Formulation}
 \label{sec:variational}
 
-Variational functional for implicit models, \acs{PCM} in this case:
+Variational functional for implicit models, \acs{PCM} in this
+case:~\autocite{Lipparini2010-be, Lipparini2011-aj, Lipparini2016-mo}
 \begin{equation}
@@
 
-Variational functional for explicit models, could be MMpol or PE or FQ:
+Variational functional for explicit models,~\autocite{Lipparini} could be
+MMpol~\autocite{Curutchet} or PE~\autocite{Steindal, Kongsted, Olsen} or
+FQ~\autocite{Rick1994-mn, Rick1995-wu, Rick1996-om, Lipparini}:
 \begin{equation}
@@
 The PCM equations will be written in the ``complete basis'' meaning that
-we will introduce the usual boundary-element method (BEM) discretization
+we will introduce the usual \ac{BEM} discretization
 at the very end of the derivation. In other words, we will be working
@@
 
-Coupled implicit/explicit polarization energy functional:
+Coupled implicit/explicit polarization energy
+functional:~\autocite{Lipparini}
 \begin{equation}\label{eq:pcm-mm-functional}
@@
 and $\MMPCM$ is the implicit/explicit interaction kernel. This is
-charge/dipole or charge/charge interaction kernel for the MMpol and PE models
-or the FQ model, respectively.
+charge/dipole or charge/charge interaction kernel for the MMpol and
+\ac{PE} models
+or the \ac{FQ} model, respectively.
 The global minimum of the convex functional is found by differentiating
@@
 \acl{MEP} (\acs{MEP}), $\zeta$ can either be the molecular electric
-field (MMpol and PE models) or again the \acs{MEP} (FQ model).
+field (MMpol and \acs{PE} models) or again the \acs{MEP} (\acs{FQ} model).
 In any case, both will be determined by the quantum mechanical molecular

### 2016-08-26 Chapters/PaperSummary.tex
Update notations

@@
 step towards ensuring reproducibility of results from scientific
-simulation software \cite{Gezelter2015}.
+simulation software \cite{Gezelter2015-gz, Krylov2015-fs, Jacob2016-oq}.
 Open-source development can fully leverage the benefits

### 2016-08-26 FrontBackmatter/Notation.tex
Update notations

@@
 order to maintain consistency throughout. We summarize them here:
+
+\section*{Basic Notation}
 \begin{table*}[!h]
-\begin{tabular}{c l}
- $\bi{A}, \bi{D}, \ldots$ & integral operators over a closed subset of $\mathbb{R}^2$ \\
- $H, \varPhi, \ldots$ & first or second quantized $N$-electron operators \\
- $\vect{r}, \vect{r}^\prime, \ldots$ & position vectors in $\mathbb{R}^3$ \\
- $\vect{s}, \vect{s}^\prime, \ldots$ & position vectors on a closed surface \\
- $\vect{v}, \vect{q}, \ldots$ & vectors in an arbitrary vector space \\
- $\mat{K}, \mat{F}, \ldots$ & matrices in an arbitrary vector space \\
+\begin{tabular}{l l}
+  $u_E$ & Restriction (trace) of the function $u$ to the set $E$ \\
+  $\Span{\vect{v}_1, \ldots, \vect{v}_n}$ & Vector space spanned by the
+  vectors $\vect{v}_1, \ldots, \vect{v}_n$ \\
+  $\delta_{ij}$ & Kronecker symbol: $\delta_{ij}=1i$ if $i=j$ and 0
+  otherwise
+\end{tabular}
+\end{table*}
+
+\section*{Vectors and Matrices}
+\begin{tabular}{l l}
+ $\vect{v}, \vect{q}$ & Vectors \\
+ $\mat{K}, \mat{F}$ & Matrices \\
+ $\mathbb{R}^{m, n}$ & Vector space of real-valued $m\times n$ matrices
+ \\
+ $\Tr\mat{A}$ & Trace of $\mat{A}$. For $\mat{A}\in\mathbb{R}^{n, n}$,
+ $\Tr\mat{A}=\sum_{i=1}^n A_{ii}$ \\
  $0_N, I_N$ & the zero and the identity in an $N$-dimensional vector space
 \end{tabular}
-\end{table*}
+
+\section*{Operators}
+\begin{tabular}{c l}
+ $\bi{A}, \bi{D}$ & integral operators \\
+ $H, \varPhi$ & first or second quantized $N$-electron operators
+\end{tabular}
+
+\section*{Function Spaces}
+\begin{tabular}{c l}
+ $\norm{u}{X}$ & Norm of $u$ in the normed space $X$ \\
+ $\mathcal{C}^0(\Omega)$ & Space of continuous
+ functions on $\Omega\subset\mathbb{R}^d$ \\
+ $\mathcal{C}^k(\Omega)$ & Space of $k$ times
+ continuously differentiable functions on $\Omega$ \\
+ $L^p(\Omega)$ & Space of functions whose $p$-th power is Lebesgue
+ integrable on $\Omega$ \\
+ $\norm{u}{0, p, \Omega}$ & Norm in $L^p(\Omega)$: $\norm{u}{0, p,
+ \Omega} = (\int_\Omega\diff\vect{r} |u(\vect{r})|^p)^\frac{1}{p}$ \\
+ $\norm{u}{\Omega}$ & Norm in $L^2(\Omega)$:
+ $\norm{u}{\Omega} = \norm{u}{0, 2,\Omega} = \sqrt{\int_\Omega\diff\vect{r} |u(\vect{r})|^2}$ \\
+ $H^s(\Omega)$ & Sobolev space of order $s$ of functions whose
+ derivative up to order $s$ are in $L^2(\Omega)$ \\
+ $H_0^s(\Omega)$ & Space of functions in $H^2(\Omega)$ with null trace on the boundary of $\Omega$ \\
+ $|u|_{s, \Omega}$ & Seminorm in $H^s(\Omega)$:
+ $|u|_{s, \Omega} \coloneq \sum_{|\alpha|=s}\norm{\partial^\alpha u}{\Omega}$ \\
+ $\norm{u}{s, \Omega}$ & Norm in $H^s(\Omega)$:
+ $\norm{u}{s, \Omega} \coloneq \sum_{l \leq s} |u|_{l, \Omega}$
+\end{tabular}
 

### 2016-08-26 FrontBackmatter/Introduction.tex
Add some more snippets

@@
 
+\todo[inline]{CURRENTLY VERBATIM FROM MASTER THESIS}
+A similar situation in encountered when quantum chemistry deals with the
+description of complex systems, such as solutions or proteins. Unlike
+relativity, the inclusion of the environment in the description of a
+chemical system has been considered from the early stages in the
+evolution of quantum chemistry.
+Already in 1936 \citeauthor{Onsager1936-wf} had proposed a model to
+account for the presence of a liquid on point-like electric
+moments.~\autocite{Onsager1936-wf}
+
+As the development of our knowledge of solutions reflects to some extent
+the development of chemistry itself, this interest is very well
+justified.
+Despite the large number of reactions known to happen in the solid state
+and in gas phase, we can state that almost always chemistry happens in
+solution.~\autocite{Reichardt2010-le}
+
+The theoretical treatment of environment effects suffers from a
+\emph{dimensionality} problem, since even the most simplified picture of
+the system under consideration would require taking into account
+500--1000 atoms, at least. Application of the most accurate methods in
+quantum chemistry is therefore impossible and not even desirable: as
+always in problems with a high dimensionality, the microscopic detail in
+the physical description cannot account for the macroscopic behaviour of
+the system. This is an important point that will be discussed in greater
+detail later.
+Models must be devised to overcome the dimensionality ``disease'': the
+ability to account, at least qualitatively, for environment effects is
+rather fundamental in all experimental branches of
+chemistry.~\autocite{Anderson1972-ai, Winsberg2010-sy, Kovac2011-ew}
+
 \section*{The Problem of Solvation or Taming Complexity with Models}
+
+\todo[inline]{CURRENTLY VERBATIM FROM MASTER THESIS}
+
+We should state more precisely what we mean by solution, here and in the
+following we will always refer to a system were the number of solvent
+molecules exceeds by far the number of solute molecules. From an
+experimental point of view, it is an \emph{extremely} dilute solution.
+We will follow the expositions given by Tomasi in refs.~
+\noparcite{Tomasi2004-dc} and \noparcite{Tomasi2007-es}.
+
+Our observables of interest are the same as \emph{in vacuo}, \ie
+relative energies, spectroscopic parameters and so forth. These
+observables are influenced by the presence of the solvent and this is
+what is commonly called \emph{solvent effect}.
+Solvent effects are, customarily but conveniently, classified as:
+\begin{itemize}
+\item[.] \emph{direct}. The presence of the solvent modifies the solute
+  electronic density, all observables depending on the density are
+  accordingly perturbed;
+\item[.] \emph{indirect}. In the Born--Oppenheimer picture the
+  electronic energy acts as \ac{PES} for the
+  nuclear motion. The most striking solvent effect is thus the
+  modification of the equilibrium geometry of a molecular solute.
+  Aminoacids are notable examples of this behaviour, exhibiting
+  different predominant conformers in different solvents;
+\item[.] \emph{local field}. The presence of the solvent can alter the
+  response of the system to applied fields, locally modifying the
+  applied field itself.~\autocite{Cammi1998-jp, Pipolo2014-sd}
+\item[.] \emph{dynamic}. When considering electronic excitations we
+  always work in a Franck--Condon framework. This is also true for
+  excitation processes in solution. In addition, we will have to
+  consider not only the geometric relaxation of the solute, but also the
+  relaxation of the surrounding medium;
+\item[.] \emph{specific}. A brief description of specific solvent
+  effects is not possible, due to their extremely varied nature. The
+  name specific comes from the fact that their existence depends on the
+  particular solute-solvent pair, rather than only on one of the
+  partners. It appears thus safe to say that all the effects that must
+  be described considering the atomistic nature of the solute-solvent
+  interaction are specific.
+\end{itemize}
+
+It is, again, customary and convenient to divide the models proposed in
+two classes, according to the microscopic description of the solvent
+these models give. Hence we distinguish between:
+\begin{itemize}
+ \item[.] \emph{discrete} (or \emph{explicit}). The degrees of freedom
+   associated with the solvent molecules are dealt with explicitly as
+   are those associated with the solute. The two sets of coordinates can
+   be treated at different levels of theory;
+ \item[.] \emph{continuum} (or \emph{implicit}). Only the degrees of
+   freedom associated with the solute are dealt with explicitly, the
+   solvent being replaced with a structureless continuum characterized
+   by its bulk properties.
+\end{itemize}
+
+Continuum models arise as limits, in a proper sense, to the discrete
+models. To stress this point we introduce the \emph{solution}
+Hamiltonian in the Born--Oppenheimer approximation as:
+\begin{equation}\label{eq:solution-ham}
+ H(\vect{r}_\mathrm{M}, \vect{r}_\mathrm{S}) =
+  H_\mathrm{M}(\vect{r}_\mathrm{M}) +  H_\mathrm{S}(\vect{r}_\mathrm{S})
++ H_\mathrm{MS}(\vect{r}_\mathrm{M}, \vect{r}_\mathrm{S})
+\end{equation}
+where the subscript M refers to the solute while S to the solvent. The
+coordinates $(\vect{r}_\mathrm{M}, \vect{r}_\mathrm{S})$ refer to
+\emph{both} nuclei and electrons, the interaction term is given by the
+usual Coulombic electrostatic Hamiltonian.
+
+It is important to remark that nuclear coordinates must be considered
+explicitly. Liquid phases are, by definition, characterized by the high
+mobility of their molecular constituents. Changes in the internal
+geometry of solvent molecules, triggered by intermolecular interactions
+between themselves and with the solute, always play a role in the
+solvation process. Whether this role is prominent or not cannot be
+decided \emph{a priori}.
+
+From a computational point of view, using the Hamiltonian in Eq.~\eqref{eq:solution-ham} is simply not feasible. We would have to
+determine a \acs{PES} of enormous dimensionality and then explore the full
+dynamics of the system. Even assuming that the minimum energy
+conformation on the \acs{PES} can be fully characterized, its knowledge will
+not lead us to a correct description of the physics of solvation. In
+fact, macroscopic observables of systems with a large number of degrees
+of freedom arise as \emph{average} values of the microscopic behaviour
+of every single component, the average being either on the phase space
+trajectory or on the appropriate ensemble.~\autocite{Hill1960-ql}. It is always
+necessary to perform some kind of average upon the energies
+corresponding to all the accessible conformations of the whole
+solute+solvent system.
+
+Of course, this formidable task may be accomplished using statistical
+thermodynamics and the idea of \emph{statistical ensembles} developed by
+Gibbs. To apply the idea of an ensemble average we must resort to a
+molecular dynamics or Monte Carlo simulation of the solution, in order
+to obtain the appropriate partition function, $Z$. This introduces a
+number of points that will be central in the development of continuum
+models:
+\begin{enumerate}
+ \item statistical simulations rely on \emph{macroscopic} parameters
+   such as the temperature or the density. These parameters are not
+   present in the Hamiltonian \eqref{eq:solution-ham}, but are necessary
+   for the description of molecular systems in the condensed phase. This
+   is true whether one considers discrete or continuum approaches;
+ \item the basic energetic quantity is the thermodynamic potential
+   characterizing the chosen ensemble. If a direct comparison with the
+   experiment is sought, one should run a simulation in the $(NpT)$
+   ensemble whose potential is the Gibbs free energy $G$. This point is
+   of paramount importance in continuum models and we will have the
+   opportunity to stress it in this Chapter as well as in Chapter
+   \ref{chap:relpcmscf};
+ \item even starting from a completely atomistic point of view, our
+   final results will always be in the form of an average over solvent
+   conformations. We have explicitly replaced a discrete description
+   with a continuous distribution function. Atomistic details can of
+   course be recovered by further analysis, but the point remains:
+   properties of solutions (the solvation energy among them) have an
+   intrinsic average nature.
+\end{enumerate}
+
+The fact that an averaging procedure is essential in studying solvation
+processes should convince the reader that, at least from a purely
+mathematical point of view, we can replace the full Hamiltonian
+\eqref{eq:solution-ham} with an \emph{effective} Hamiltonian, averaging
+the solvent degrees of freedom in a preliminary
+step.~\autocite{Angyan1992-vo, Tapia1992-pu}
+\todo[inline]{Some pieces missing here, FILL IN}
+Alternatively, an approximate decoupling can be obtained replacing the
+mean-field operator in Eq.~\eqref{eq:ave1} by its classical average over
+a suitable ensemble:
+\begin{equation}\label{eq:ave_classic}
+ \Braket{H_\mathrm{MS}}_\mathrm{S} = \frac{\int\diff\Gamma H_\mathrm{MS}\exp
+\left(-\beta H_\mathrm{S}\right)}{\int\diff\Gamma\exp\left(-\beta H_\mathrm{S}\right)}.
+\end{equation}
+This classical average can be written in terms of the $g(\vect{r})$ distribution functions. These functions are defined
+for the electrons and the nuclei of the solvent with respect to the fixed frame of reference of the solute molecule, see \noparcite[ref.][]{hansen}.
+This amounts to considering the solvent as a structureless continuum, hence defining continuum models.
+
+Continuum models are then characterized, by a solute-only
+\emph{effective} Hamiltonian:\footnote{We will use the symbol
+$H_\mathrm{MS}$ for the interaction Hamiltonian even if it is not the
+full interaction term of Eq.~\eqref{eq:solution-ham}.}
+\begin{equation}\label{eq:effective-ham}
+ H(\vect{r}_\mathrm{M}) = H_\mathrm{M}(\vect{r}_\mathrm{M})
++ H_\mathrm{MS}(\vect{r}_\mathrm{M}),
+\end{equation}
+the explicit form of the averaged interaction in
+Eq.~\eqref{eq:effective-ham} is in terms of solvent \emph{response
+functions}, $Q_x(\vect{r},\vect{r}^\prime)$.
+As will be apparent in the following, this term may depend on the solute
+density, making the effective Hamiltonian \emph{nonlinear}.
+
+An hierarchy of approximations for the resolution of the effective
+Schr\"odinger equation can be devised, as was done for the discrete
+models. The distinguishing feature is now the \emph{order} of the
+solvent response functions used in the definition of the interaction
+term.
+
+If we limit ourselves to a linear response of the solvent we obtain the
+class of models commonly called continuum solvation models, the
+Polarizable Continuum Model among them. Computer simulations can be used
+to determine these functions, but almost always one uses bulk properties
+of the pure solvent obtained from experiment.
+
+This, at first sight, crude approximation is rather well justified by
+the following argument: from a physical point of view, our solution is
+extremely dilute. The presence of the solute is an extremely small
+perturbation for the bulk of the solvent and it is safe to assume that
+the macroscopic properties of the solution are not far from those of the
+pure solvent. Furthermore, the use of macroscopic experimental values
+implicitly takes into account the averaging procedure one would carry
+out in atomistic simulations of the solution.
+
+As pictorially represented in Figure \ref{fig:discretetocontinuum} the
+solvent is now a \emph{structureless polarizable} continuum and the
+solute is accommodated inside a cavity formed in this continuum.
+\begin{figure}[!h]
+ \centering
+\includegraphics[width=.5\textwidth]{discretetocontinuum}
+\caption{Pictorial representation of the transition from a discrete to a
+continuum model of a solution.}
+\label{fig:discretetocontinuum}
+\end{figure}
 

### 2016-08-26 Chapters/PaperSummary.tex
Work on the summary of papers

--- 2016-08-24 Chapters/PaperSummary.tex
@@
 
-\epigraph{$\frac{1}{2}$ research and $\frac{1}{2}\,\,T\acute{\epsilon}\chi\nu\eta$
+% Very ugly
+\epigraph{\textonehalf\,\,research\,\,\,\,\,\,\,\,\,\,\,\,and \textonehalf\,\,\,\,\,\,\,\,Tεχνη
 
-          $\frac{1}{2}$ observation, $\frac{1}{2}\,\,T\acute{\epsilon}\chi\nu\eta$
+          \textonehalf\,\,observation,\,\,\,\,\,\,\,\,\,\,\,\,\textonehalf\,\,\,\,\,\,\,\,Tεχνη
 
-          $\frac{1}{2}$ training, $\frac{1}{2}\,\,T\acute{\epsilon}\chi\nu\eta$}{
+          \textonehalf\,\,training,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\textonehalf\,\,\,\,\,\,\,\,Tεχνη}{
   --- \textsc{Ezra Pound}, \textit{Canto LXXXV}}
 
-In the following, I provide a brief summary of the motivations, contents and
+This final Chapter provides a brief overview of the motivations, results and
 conclusions of the papers this thesis is based on.
-Figures and tables here reported are minimally modified versions taken from the published work,
-captions are instead copied \emph{verbatim}.
-In addition, I give a description of my contributions to the work leading to
-the publication of each of the papers.
+All publications included in this thesis have involved some programming effort.
+I will also describe the principles we have striven to follow in developing our
+\acs{PCM} software library and its management philosophy.
+I will also give some details on the interfaces we have developed to different
+quantum chemistry codes.
+Finally, for each paper, I will give a description of my contributions to the
+work leading to its publication.
 \todo[inline]{Write something along these lines: "My coauthors have all read and approved."}
@@
 
+\refstepcounter{dummy}
+\addcontentsline{toc}{section}{Software}
 \section*{Software}
+
+\todo[inline]{Have a closer look at this. I think it needs to be updated for the discussion
+on open-source software that went down on JPC Lett. some time ago.
+Also, acronyms should be properly set up.}
+
+A shorter version of this discussion has been accepted for publication in the
+proceedings of the \emph{Producing High Performance and Sustainable Software
+for Molecular Simulation} workshop held at the 2015 Supercomputing Conference.
+
+We have created an open-source, standalone library to provide \acs{PCM} functionality across
+altogether different quantum chemical program packages.
+The library serves as our development platform for the \acs{PCM}.
+Moreover, the interfaces developed to different quantum chemistry codes allow the exploration
+of new methodologies, significantly cutting down development times.
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=.8\textwidth]{notpretty.png}
+\end{figure}
 
@@
 \end{figure}
+
+The growing complexity of quantum chemical program packages requires that an
+appropriate strategy be devised to implement new features.
+Scalability is of paramount importance, but it has become clear that
+maintainability and extensibility of the code play an equally important role in
+managing this growing complexity.~\autocite{Hatton1994-us, Hatton1997-ml,
+Hatton1997-bb, Ioannidis2005-aj, Merali2010-uy, Prinz2011-kz, Wilson2014-vh}
+
+The use of a \emph{modular programming paradigm} has been recognized as
+beneficial in many other scientific computing contexts. New features are
+isolated into libraries that can be accessed by host programs through a
+well-defined \ac{API}.
+Computational tasks are thus implemented into separate, independent and
+interchangeable modules. These are \emph{developed} and \emph{tested}
+independently of the particular host program in which they will be
+used.~\cite{Dijkstra1968-vy, Parnas1972-im}
+A well-defined \ac{API} clearly delimits the boundaries of the functionality
+offered, thus forcing a programming style and standardisation of the
+functionality, eventually.~\autocite{Reddy2011-sd}
+
+When coupled with \emph{open-source} licensing, modularity is a proper
+step towards ensuring reproducibility of results from scientific
+simulation software \cite{Gezelter2015-gz, Krylov2015-fs, Jacob2016-oq}.
+Open-source development can fully leverage the benefits
+of widespread, cloud-based, free, code development services, such as
+hosted \acp{DVCS}, \ac{CI}, code
+coverage, static and dynamic code analyses, nightly regression testing,
+public issue tracking, code review and so forth:
+adoption of a modern code development workflow is easily within reach.
+
+It is, of course, true that the above mentioned services are not
+exclusive prerogatives of open-source projects.
+However, an open review process of scientific software can often help
+to establish
+reproducibility, extensibility and sustainability of the software
+ecosystem.
+Open-source software, modular development of new functionalities
+and full-fledged exploitation of \acsp{DVCS} ensure a much larger scientific
+impact. Third-parties can easily contribute to
+the project: by improving the documentation, by reporting bugs or
+by actively extending the codebase with new functionality.
+
+The description of chemical phenomena in complex environments, such as
+solutions or proteins, is as challenging as it is important
+for an accurate prediction of a wide range of properties.~\autocite{Reichardt2010-le}
+  Quantum chemistry is nowadays a valuable
+tool: its simulations can provide significant insight into complex
+chemical phenomena.~\autocite{Nobel1998, Nobel2013}.
+The major challenge lies in providing a faithful, realistic, yet cost-effective
+description of the environment of the molecule:
+a molecule in solution can be represented by a quantum mechanical model
+comprising a great number of atoms, thousands or more. Such a simulation would,
+however, be impossible to carry out with the current state-of-the-art
+computational resources.
+Models must be devised to overcome the dimensionality ``disease'' and
+the idea of introducing different approximation levels for different
+parts of the system under study is particularly appealing, as it helps
+our chemical understanding. This is the idea at the basis of
+\emph{multiscale} models.
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=.45\textwidth]{multiscale.jpg}
+\end{figure}
+
+\emph{Continuum} (or \emph{implicit}) models, deal explicitly
+only with the degrees of freedom associated with the solute, while
+replacing the solvent with a structureless continuum characterized by
+its bulk properties \cite{Onsager1936-wf}. Among these models, the \acl{PCM} (\acs{PCM})
+is the most widespread and successful \cite{Tomasi2005-vm, Mennucci2008-xl}.
+
+The \acs{PCM} is also an ideal candidate for the creation of a solvation \acs{API}.
+Its input from and output to \emph{any} quantum chemistry code is
+limited and well-defined.
+This provides a natural \acs{API} design: the \acs{API} functions can be compared
+\emph{vis-à-vis} with the working equations derived for the different
+quantum chemical methods.
+
+We will elucidate the design principles adopted in the development of our
+\pcmsolver library: an \emph{open-source} \acs{API} for the inclusion of the
+PCM in \emph{any} quantum chemistry code \cite{PCMSolver}.  Use of our
+\acs{API} \emph{significantly} limits coding effort on the side of the host:
+continuum solvation at the \acs{SCF}
+level of theory can be implemented in the host program almost out-of-the-box.  The module is
+agnostic of the host code internals and is tested separately from it.
+
+We will emphasize the role of the open-source licensing model and the
+importance of the modern workflows adopted in our project.  The adoption
+of \git as \acs{DVCS}, \cmake for cross-platform builds and automatic
+documentation deployment on \readthedocs simplify extensibility of the
+module and promote third-party contributions to the code base, while
+continuous integration and nightly testing offer an invaluable level of
+confidence in the code.
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=.5\textwidth]{pcmsolver-scheme}
+\end{figure}
+
+\pcmsolver currently offers the basic PCM functionality already
+available in many widespread software packages, such as traditional
+collocation solvers for isotropic media.~\autocite{Tomasi2005-vm}
+Some unique functionalities are the wavelet Galerkin solvers on smooth
+molecular surfaces~\autocite{Weijo2010-hy, Bugeanu2015-tp}, the Green’s
+functions for spherical diffuse interfaces~\autocite{DiRemigio2016-nn} and
+the delayed solvers for real-time \acl{TDSCF}
+simulations.~\autocite{Corni2015-pe}
+
+Careful usage of the C++ object-oriented paradigm is the key to this
+wide spectrum of functionalities. The library is in itself made of
+modules, communicating by means of composition at the outermost
+level of the design.
+This allows us to mix languages internally: we currently have Fortran
+and C modules alongside the main C++ structure.
+Library internal classes balance the dynamic polymorphism offered by
+inheritance and the static polymorphism offered by template programming.
+~\autocite{Alexandrescu2001-bp, Sutter2004-nt, Langr2012-js}
+Preeminent use of composition over inheritance keeps the coupling
+between different modules as low as possible.~\autocite{Gamma1995-fd}
+
+References for the various codes we have created interfaces to.
+\psicode~\autocite{Turney2012-de}
+\DIRAC~\autocite{DIRAC15}
+\DALTON~\autocite{DALTON16, Aidas2013-rp}
+\LSDALTON~\autocite{LSDALTON16, Aidas2013-rp}
+\ReSpect~\autocite{ReSpect citation}
+KOALA~\autocite{Hofener2014-ex, Hofener2016-qz}
 

### 2016-08-26 Chapters/ResponseTheory.tex
Clean up and add one chapter (to be written)
Work on the summary of papers

--- 2016-08-10 Chapters/Ch02-ResponseTheory.tex
@@
-%*****************************************
 \chapter{Response theory and molecular properties}\label{ch:molprop}
-%*****************************************
 
 \epigraph{
-White light, white light goin' messin' up my brain
+White light goin' messin' up my brain
 
-White light, aww white light it's gonna drive me insane
+White light it's gonna drive me insane
 }{
@@
 }
-
 \epigraph{What's the frequency, Kenneth? [...]
@@
 
-%*****************************************
-%*****************************************
-%*****************************************
-%*****************************************
-%*****************************************
+\begin{itemize}
+    \item Something about exact state response theory: quasienergy,
+      response functions, residues of response functions.~\autocite{}
+    \item  Derivation of the linear response function with
+      quantum/classical polarizable Hamiltonian and then refer to the
+      OpenRSP paper for the open-ended treatment.~\autocite{pcm-openrsp}
+\end{itemize}

### 2016-08-26 Chapters/SolvationAndCorrelation.tex
Clean up and add one chapter (to be written)
Work on the summary of papers

--- 2016-08-24 Chapters/Ch03-SolvationAndCorrelation.tex
@@
 %************************************************
-\chapter{A New Perspective on Electron Correlation and Solvation}\label{ch:solvation-correlation}
+\chapter{Electron Correlation and Solvation}\label{ch:solvation-correlation}
 
@@
 which stems directly from the exponential nature of the wave
-operator.~\autocite{Shavitt, Helgaker2000-tz}
+operator.~\autocite{Shavitt2009-mr, Helgaker2000-tz}
 

### 2016-08-26 FrontBackmatter/Notation.tex
Cosmetics of the notation

--- 2016-08-24 FrontBackmatter/Notation.tex
@@
-\chapter*{Notations, conventions and units}
-\addcontentsline{toc}{section}{\tocEntry{Notations, conventions and units}}
+\renewcommand\nomgroup[1]{%
+  \item[\bfseries
+  \ifstrequal{#1}{A}{Basic Notation}{%
+  \ifstrequal{#1}{B}{Vectors and Matrices}{%
+  \ifstrequal{#1}{C}{Operators}{}}{
+  \ifstrequal{#1}{D}{Function Spaces}{}}{
+  \ifstrequal{#1}{E}{Molecular Quantum Mechanics}{}}}%
+]}
 
+\renewcommand{\nomname}{
+Notations, Conventions and Units
+}
+\renewcommand{\nompreamble}{
 A number of notations and typographic conventions has been adopted in
-order to maintain consistency throughout. We summarize them here:
-\begin{table*}[!h]
-\begin{tabular}{c l}
- $\bi{A}, \bi{D}, \ldots$ & integral operators over a closed subset of $\mathbb{R}^2$ \\
- $H, \varPhi, \ldots$ & first or second quantized $N$-electron operators \\
- $\vect{r}, \vect{r}^\prime, \ldots$ & position vectors in $\mathbb{R}^3$ \\
- $\vect{s}, \vect{s}^\prime, \ldots$ & position vectors on a closed surface \\
- $\vect{v}, \vect{q}, \ldots$ & vectors in an arbitrary vector space \\
- $\mat{K}, \mat{F}, \ldots$ & matrices in an arbitrary vector space \\
- $0_N, I_N$ & the zero and the identity in an $N$-dimensional vector space
-\end{tabular}
-\end{table*}
-
-\todo[inline]{Clarify the notation for the 4-component basis in
-accordance with the EPR paper.}
-Further conventions have been adopted for the indices of functions.
-Lower case Latin letters are used for \ac{MO} 4-spinors, while lower case
-Greek letters ($\kappa, \lambda, \mu, \ldots$) are reserved for
-one-electron basis functions in 2-spinor or scalar form.
-Specific ranges of letters are used as follows:
-\begin{table*}[!h]
-\begin{tabular}{c l}
- $r, s, t, \ldots$ & general MO 4-spinor indices \\
- $i, j, k, \ldots$ & occupied MO 4-spinor indices \\
- $a, b, c, \ldots$ & virtual MO 4-spinor indices
-\end{tabular}
-\end{table*}
-
-Upper case Latin letters always refer to functions in vector spaces
-other than the two listed above. Complex conjugation will always be
-shown using a dagger ($\dagger$) instead of a star ($*$).
-
-Hartree atomic units have been used:~\autocite{Whiffen1978-xx}
+order to maintain consistency throughout. We summarize them here.
+Hartree atomic units are used throughout:~\autocite{Whiffen1978-xx, Mohr2012-zc}
 \begin{equation*}
@@
 \end{equation*}
+}
 
-Some useful conversion constants to SI units are here
-listed:~\autocite{Mohr2012-zc}
-\begin{subequations}
- \begin{align*}
-  \SI{1}{\bohr} &= \SI{0.52917721092}{\angstrom} = \SI{5.2917721092e-11}{\meter} \\
-  \SI{1}{\hartree} &= \SI{27.21138505}{\electronvolt} = \SI{2625.4996404}{\kilo\joule\per\mole} \\
-  \SI{1}{\debye} &= \SI{3.3356410E-30}{\coulomb\meter} \\
-  \SI{1}{\angstrom\cubed} &= \SI{1.11264984E-40}{\coulomb\squared\meter\squared\per\joule}
- \end{align*}
-\end{subequations}
+% Basic Notation
+\nomenclature[A, 01]{$u_E$}{Restriction (trace) of the function $u$ to the set $E$}
+\nomenclature[A, 02]{$\Span{\vect{v}_1, \ldots, \vect{v}_n}$}{Vector space spanned by the vectors $\vect{v}_1, \ldots, \vect{v}_n$}
+\nomenclature[A, 03]{$\delta_{ij}$}{Kronecker symbol: $\delta_{ij}=1i$ if $i=j$ and 0 otherwise}
+% Vectors and Matrices
+\nomenclature[B, 01]{$\vect{v}, \vect{q}$}{Vectors}
+\nomenclature[B, 02]{$\mat{K}, \mat{F}$}{Matrices}
+\nomenclature[B, 03]{$\mathbb{R}^{m, n}$}{Vector space of real-valued $m\times n$ matrices}
+\nomenclature[B, 04]{$\Tr\mat{A}$}{Trace of $\mat{A}$. For $\mat{A}\in\mathbb{R}^{n, n}$, $\Tr\mat{A}=\sum_{i=1}^n A_{ii}$}
+\nomenclature[B, 05]{$0_N, I_N$}{The zero and the identity in an $N$-dimensional vector space}
+% Operators
+\nomenclature[C, 01]{$\bi{A}, \bi{D}$}{Integral operators}
+\nomenclature[C, 02]{$H, \varPhi$}{First or second quantized $N$-electron operators}
+% Function Spaces
+\nomenclature[D, 01]{$\norm{u}{X}$}{Norm of $u$ in the normed space $X$}
+\nomenclature[D, 02]{$\mathcal{C}^0(\Omega)$}{Space of continuous functions on $\Omega\subset\mathbb{R}^d$}
+\nomenclature[D, 03]{$\mathcal{C}^k(\Omega)$}{Space of $k$ times continuously differentiable functions on $\Omega$}
+\nomenclature[D, 04]{$L^p(\Omega)$}{Space of functions whose $p$-th power is Lebesgue integrable on $\Omega$}
+\nomenclature[D, 05]{$\norm{u}{0, p, \Omega}$}{Norm in $L^p(\Omega)$: $\norm{u}{0, p, \Omega} = (\int_\Omega\diff\vect{r} |u(\vect{r})|^p)^\frac{1}{p}$}
+\nomenclature[D, 06]{$\norm{u}{\Omega}$}{Norm in $L^2(\Omega)$: $\norm{u}{\Omega} = \norm{u}{0, 2,\Omega} = \sqrt{\int_\Omega\diff\vect{r} |u(\vect{r})|^2}$}
+\nomenclature[D, 07]{$H^s(\Omega)$}{Sobolev space of order $s$ of functions whose derivative up to order $s$ are in $L^2(\Omega)$}
+\nomenclature[D, 08]{$H_0^s(\Omega)$}{Space of functions in $H^2(\Omega)$ with null trace on the boundary of $\Omega$}
+\nomenclature[D, 09]{$|u|_{s, \Omega}$}{Seminorm in $H^s(\Omega)$: $|u|_{s, \Omega} \coloneq \sum_{|\alpha|=s}\norm{\partial^\alpha u}{\Omega}$}
+\nomenclature[D, 10]{$\norm{u}{s, \Omega}$}{Norm in $H^s(\Omega)$: $\norm{u}{s, \Omega} \coloneq \sum_{l \leq s} |u|_{l, \Omega}$}
+% Molecular Quantum Mechanics
+\nomenclature[E, 01]{$r, s, t$}{General \acl{MO} indices}
+\nomenclature[E, 02]{$i, j, k$}{Occupied \acl{MO} indices}
+\nomenclature[E, 03]{$a, b, c$}{Virtual \acl{MO} indices}
+\nomenclature[E, 04]{$\kappa, \lambda, \mu$}{One-electron basis functions indices}
+
+\printnomenclature

### 2016-08-26 FrontBackmatter/Abstract.tex
Clean up and add one chapter (to be written)
Write some of the abstract

@@
 
+The vast majority of chemical phenomena happens in complex environments
+where the molecule of interest can interact with a large number of other
+moieties, be they solvent molecules or a structured environment, such as
+a protein.
+Spectroscopic experiments measure the response of a chemical system to
+an applied perturbing electromagnetic field and can probe the
+microscopic motions of electrons and nuclei in molecules.
+This achieves unique insight into the structural and physical behaviour
+of complex systems.
+Characterizing, explaining and predicting the effect of the chemical
+environment on a large number of measurable properties requires,
+however, a synergistic experimental and theoretical approach.
+
+In this thesis, I will present our efforts in the development of the
+polarizable continuum model for solvation.
+
 \vfill
+
+% Anything in Norwegian?
+%\pdfbookmark[1]{Zusammenfassung}{Zusammenfassung}
+%\chapter*{Zusammenfassung}
+%Kurze Zusammenfassung des Inhaltes in deutscher Sprache\dots
 
@@
 \vfill
+
+\begin{flushright}
+    \begin{tabular}{m{5cm}}
+      \textit{\myLocation, \myTime} \\
+      \includegraphics[width=0.40\textwidth]{gfx/signature.png} \\
+        \hline
+        \centering\myName \\
+    \end{tabular}
+\end{flushright}

### 2016-08-27 FrontBackmatter/Introduction.tex
Cosmetics of the notation
Massage the introduction

@@
 \addtocontents{toc}{\protect\vspace{\beforebibskip}} % to have the bib a bit from the rest in the toc
-\addcontentsline{toc}{section}{\tocEntry{Introduction}}
+\addcontentsline{toc}{chapter}{\tocEntry{Introduction}}
+
+\renewcommand{\thefigure}{\Alph{figure}}
 
@@
 
+\epigraph{
+[...] og fordi jeg alltid har hatt en dragning mot det skjulte og hemmelige.
+    }{
+    --- \textsc{K. O. Knausgård}, \textit{Om Høsten}}
+
+\begin{figure}[tb]
+\centering
+\scalebox{0.8}{\input{gfx/RQC-axes.tikz}}
+\caption{The three dimensions of Relativistic Quantum Chemistry. The computational scaling of a relativistic model chemistry is
+$ S=\textcolor{red}{y}(\textcolor{blue}{x})f(\textcolor{blue}{x})\textcolor{green}{M}^{\textcolor{blue}{x}} $. The expense of relativity is in
+the appearance of a new prefactor, not in the $M$-dependence of the particular method.}
+\label{fig:RQC-axis}
+\end{figure}
+and look at different axes.
+
+\section*{What Are the Electrons Really Doing in Molecules?}
+
+\todo[inline]{Write something about the scope of molecular quantum
+mechanics.}
+
+This question was posed by R.~S.~Mulliken over a half-century ago, as
+lecture title for the acceptance of the Gilbert N.~Lewis award in 1960.
+
 This will be the subject of Chapter \ref{ch:QM}
-
-\todo[inline]{CURRENTLY VERBATIM FROM MASTER THESIS}
-A similar situation in encountered when quantum chemistry deals with the
-description of complex systems, such as solutions or proteins. Unlike
-relativity, the inclusion of the environment in the description of a
-chemical system has been considered from the early stages in the
-evolution of quantum chemistry.
-Already in 1936 \citeauthor{Onsager1936-wf} had proposed a model to
-account for the presence of a liquid on point-like electric
-moments.~\autocite{Onsager1936-wf}
-
-As the development of our knowledge of solutions reflects to some extent
-the development of chemistry itself, this interest is very well
-justified.
-Despite the large number of reactions known to happen in the solid state
-and in gas phase, we can state that almost always chemistry happens in
-solution.~\autocite{Reichardt2010-le}
-
-The theoretical treatment of environment effects suffers from a
-\emph{dimensionality} problem, since even the most simplified picture of
-the system under consideration would require taking into account
-500--1000 atoms, at least. Application of the most accurate methods in
-quantum chemistry is therefore impossible and not even desirable: as
-always in problems with a high dimensionality, the microscopic detail in
-the physical description cannot account for the macroscopic behaviour of
-the system. This is an important point that will be discussed in greater
-detail later.
-Models must be devised to overcome the dimensionality ``disease'': the
-ability to account, at least qualitatively, for environment effects is
-rather fundamental in all experimental branches of
-chemistry.~\autocite{Anderson1972-ai, Winsberg2010-sy, Kovac2011-ew}
 
@@
 
-\todo[inline]{CURRENTLY VERBATIM FROM MASTER THESIS}
+Chemistry can be largely considered a \emph{wet} science: almost always
+chemical phenomena happen in a liquid environment.~\autocite{Reichardt2010-le}
+We hereby define a "solution", or more generally an "environment", as
+a system were the number of solvent molecules exceeds by far the number
+of solute molecules.~\autocite{Tomasi2004-dc, Tomasi2007-es}
+It is then clear that theoretical and computational approaches to such a
+problem will necessarily suffer from a \emph{dimensionality disease}.
+The number of degrees of freedom to be taken into account is, in
+principle, so large, that even the most advanced computing machines
+would have a hard time computing the desired observables.
+Moreover, on an interpretive level, it would not even be desirable to
+have such a detailed insight.
+As is well known from statistical mechanics, microscopic detail cannot
+account for the macroscopic behaviour.~\autocite{Hill1960-ql,
+Hansen2013-io}
+To tame this complexity and cure the disease, one must devise
+\emph{models} that simplify physical reality, while offering tools for
+understanding reality and predicting new and exciting
+phenomena.~\autocite{Anderson1972-ai, Winsberg2010-sy, Kovac2011-ew}
+One of the earlier attempts at tackling the problem of solvation is due
+to \citeauthor{Onsager1936-wf}. His was a rather crude model, but one
+that has had a lasting impact and informs much of the developments that
+will be presented in this thesis.~\autocite{Onsager1936-wf}
 
-We should state more precisely what we mean by solution, here and in the
-following we will always refer to a system were the number of solvent
-molecules exceeds by far the number of solute molecules. From an
-experimental point of view, it is an \emph{extremely} dilute solution.
-We will follow the expositions given by Tomasi in refs.~
-\noparcite{Tomasi2004-dc} and \noparcite{Tomasi2007-es}.
-
-Our observables of interest are the same as \emph{in vacuo}, \ie
-relative energies, spectroscopic parameters and so forth. These
-observables are influenced by the presence of the solvent and this is
-what is commonly called \emph{solvent effect}.
-Solvent effects are, customarily but conveniently, classified as:
+Before introducing our model of choice, let us consider how an
+environment might affect molecular observables of interest.
+Environment effects are usually classified as:
 \begin{itemize}
-\item[.] \emph{direct}. The presence of the solvent modifies the solute
-  electronic density, all observables depending on the density are
-  accordingly perturbed;
-\item[.] \emph{indirect}. In the Born--Oppenheimer picture the
-  electronic energy acts as \ac{PES} for the
-  nuclear motion. The most striking solvent effect is thus the
-  modification of the equilibrium geometry of a molecular solute.
-  Aminoacids are notable examples of this behaviour, exhibiting
-  different predominant conformers in different solvents;
-\item[.] \emph{local field}. The presence of the solvent can alter the
-  response of the system to applied fields, locally modifying the
-  applied field itself.~\autocite{Cammi1998-jp, Pipolo2014-sd}
-\item[.] \emph{dynamic}. When considering electronic excitations we
-  always work in a Franck--Condon framework. This is also true for
-  excitation processes in solution. In addition, we will have to
-  consider not only the geometric relaxation of the solute, but also the
-  relaxation of the surrounding medium;
-\item[.] \emph{specific}. A brief description of specific solvent
-  effects is not possible, due to their extremely varied nature. The
-  name specific comes from the fact that their existence depends on the
-  particular solute-solvent pair, rather than only on one of the
-  partners. It appears thus safe to say that all the effects that must
-  be described considering the atomistic nature of the solute-solvent
-  interaction are specific.
+\item \emph{direct}.
+  These effect stem straightforwardly from the modification underwent by
+  the solute electronic density when interacting with the environment.
+\item \emph{Indirect}.
+  It is common for solutes to exhibit different minimum-energy
+  conformations in different environments. These effects are commonly
+  labelled as indirect.
+\item \emph{Local field}.
+  Light-matter interactions are also affected by the environment. Local
+  modifications of externally applied fields subtly influence molecular
+  responses.~\autocite{Cammi1998-jp, Pipolo2014-sd}
+\item \emph{Dynamic}.
+  Probing the nature of molecular excited states has enormous
+  technological impact. The presence of the environment radically
+  influences excited states, since relaxation processes in the medium
+  become important.
+\item \emph{Specific}. This catch-all category includes all effects
+  stemming from the peculiar solute-solvent pair interactions that
+  cannot be fully described under any of the previous labels.
+  In general, modelling such effects demands an atomistic level of
+  detail.
 \end{itemize}
 
-It is, again, customary and convenient to divide the models proposed in
-two classes, according to the microscopic description of the solvent
-these models give. Hence we distinguish between:
+Faced with the problem of describing such a diverse array of effects,
+two main models have emerged in the past decades, each with its
+strengths and weaknesses.
+Both can be classified as \emph{multiscale} (or \emph{focused})
+models~\autocite{Nobel2013} and hinge on the same idea: treat different
+parts of the system with different methods and couple these methods by
+bridging "scales" at the boundary, \emph{vide infra}.
+While both models treat the molecular degrees of freedom at the quantum
+mechanical level, their approach to the microscopic description of the
+degrees of freedom of the environment differs:
 \begin{itemize}
- \item[.] \emph{discrete} (or \emph{explicit}). The degrees of freedom
-   associated with the solvent molecules are dealt with explicitly as
-   are those associated with the solute. The two sets of coordinates can
-   be treated at different levels of theory;
- \item[.] \emph{continuum} (or \emph{implicit}). Only the degrees of
-   freedom associated with the solute are dealt with explicitly, the
-   solvent being replaced with a structureless continuum characterized
-   by its bulk properties.
+ \item
+   \emph{Discrete} (or \emph{explicit}) models explicitly treat those
+   degrees of freedom.
+   This is either achieved by a cheaper quantum mechanical
+   method~\autocite{Vreven2006-gx} or by \ac{MM}.~\autocite{Senn2009-sk}
+   In the latter approach, commonly dubbed \acs{QM}/\acs{MM}, the \acs{MM}
+   region can either be polarizable~\autocite{MMPol} or non
+   polarizable.~\autocite{MM}
+ \item
+   \emph{Continuum} (or \emph{implicit}) completely remove the degrees
+   of freedom of the environment from the picture, replacing it with a
+   structureless continuum.
+   The effect of this continuum is described, classically, \emph{via}
+   its bulk properties.~\autocite{Onsager1936-wf, Miertus1981-mm}
 \end{itemize}
+\acs{QM}/\acs{MM} models can capture, albeit approximately, the effect
+of the atomistic nature of the environment on the active part of the
+system.
+However, they demand a statistical average of environment configurations
+to yield results of any significance. Moreover, a large number
+\emph{cutoff} for the \acs{MM} region is usually required to converge
+long-range electrostatic interactions.~\autocite{Something here??}
+Continuum models avoid both problems at once. Statistical averaging is
+built into the model \emph{via} their parametrization by
+means of the environment's bulk properties, such as the permittivity.
+In addition, long-range electrostatics is treated exactly.
+Unfortunately, atomistic detail is lost and it is then impossible to
+recover a satisfactory description of specific effects.
+To partly alleviate these sources of error, the \acs{QM}/\acs{MM} and
+\acs{QM}/Continuum methods can and have been successfully combined to
+yield the three-layer \acs{QM}/\acs{MM}/Continuum method.~\autocite{Curutchet, Steindal Lipparini}
+Figure \ref{fig:qm-to-multiscale} schematically portrays the transition
+from a full \acs{QM} model of the relevant system to its multiscale
+representations.
+
+\begin{figure}[tb]
+\missingfigure{Add picture with multiscale models}
+\caption{From quantum mechanical models to multiscale models.}
+\label{fig:qm-to-multiscale}
+\end{figure}
+
+Notice that we have deliberately ruled out so-called \emph{cluster}
+models from the above discussion.~\autocite{ClusterModels}
+These approaches replace the actual physical setting with a suitable
+truncation of the whole solute+solvent system, the \emph{model system}
+and treat it within a chosen quantum mechanical level of theory.
+Cluster models can be used to benchmark more approximate multiscale
+models, but their description is outside the scope of this thesis.
+
+Chapter \ref{ch:CSM} will present an overview of the \ac{PCM} for
+solvation.
+I will present a nontechnical discussion of the mathematical details of
+the model and an outline of current methodologies for the solution of
+the associated governing equations.
+Borrowing from the work of \citeauthor{Lipparini2010-be},\autocite{Lipparini2010-be,
+Lipparini2015-lq} I will introduce a unifying theoretical formalism for
+\acs{QM}/Continuum, \acs{QM}/\acs{MM} and \acs{QM}/\acs{MM}/Continuum models that will
+be extensively used throughout the thesis.
+
+\todo[inline]{CURRENTLY VERBATIM FROM MASTER THESIS}
 
@@
 
-\autocite{Anderson1972-ai}
+\section*{The Road to Reality or Molecular Response Properties}
 
-This will be the subject of Chapter \ref{ch:CSM}
-
-\section*{The Road to Reality or Molecular Response Properties}
+Characterizing, explaining and predicting the effect of the chemical
+environment on a large number of measurable properties requires,
+however, a synergistic experimental and theoretical approach.
 

### 2016-08-28 Chapters/QuantumMechanics.tex
Work on the summary of papers
Massage the introduction
Work on introduction

@@
 % mainfile: ../RobertoDiRemigioPhDThesis.tex
-\chapter{A Quick Tour of Molecular Quantum Mechanics}\label{ch:QM}
+
+\renewcommand{\thefigure}{\arabic{chapter}.\arabic{figure}}
+
+\chapter{A Quick Tour of Quantum Chemistry}\label{ch:QM}
 
 \begin{itemize}
-  \item nonrelativistic and relativistic Hamiltonians use in molecular
+  \item nonrelativistic and relativistic Hamiltonians used in molecular
     quantum mechanics
+  \item \ac{BO} approximation and the concept of \ac{PES}
   \item Fock space

### 2016-08-28 FrontBackmatter/Introduction.tex
Work on introduction

@@
 
-\begin{figure}[tb]
-\centering
-\scalebox{0.8}{\input{gfx/RQC-axes.tikz}}
-\caption{The three dimensions of Relativistic Quantum Chemistry. The computational scaling of a relativistic model chemistry is
-$ S=\textcolor{red}{y}(\textcolor{blue}{x})f(\textcolor{blue}{x})\textcolor{green}{M}^{\textcolor{blue}{x}} $. The expense of relativity is in
-the appearance of a new prefactor, not in the $M$-dependence of the particular method.}
-\label{fig:RQC-axis}
-\end{figure}
-and look at different axes.
-
 \section*{What Are the Electrons Really Doing in Molecules?}
 
-\todo[inline]{Write something about the scope of molecular quantum
-mechanics.}
+\todo[inline]{Quantum chemistry as they study of
+molecular properties by means of \emph{ab initio} physico-mathematical
+theories and computer programs.
+Introduce the concept of model chemistries, but then mention that we are
+interested in different aspects and will look at different "axes".}
 
@@
 lecture title for the acceptance of the Gilbert N.~Lewis award in 1960.
+
+\begin{figure}[!h]
+\centering
+\scalebox{0.8}{\input{gfx/RQC-axes.tikz}}
+\caption{
+Pictorial depiction of the concept of \emph{model chemistries} as the
+three dimensions of quantum chemistry.~\autocite{Pople1999-gt, Saue2011-qg}
+The computational scaling of a model chemistry is
+$S=\textcolor{PMS1797}{y}(\textcolor{PMS2229}{x})f(\textcolor{PMS2229}{x})\textcolor{PMS130}{M}^{\textcolor{PMS2229}{x}}$.
+}
+\label{fig:RQC-axis}
+\end{figure}
 

### 2016-08-28 Chapters/ContinuumSolvationModels.tex
Work on introduction
Reshuffle some sections and dump some reference around

@@
 \pagebreak
+
+\section{Continuum Solvation Models: Why and How}\label{sec:CSM-why-how}
+\todo[inline]{CURRENTLY VERBATIM FROM MASTER THESIS}
+
+Continuum models arise as limits, in a proper sense, to the discrete
+models. To stress this point we introduce the \emph{solution}
+Hamiltonian in the Born--Oppenheimer approximation as:
+\begin{equation}\label{eq:solution-ham}
+ H(\vect{r}_\mathrm{M}, \vect{r}_\mathrm{S}) =
+  H_\mathrm{M}(\vect{r}_\mathrm{M}) +  H_\mathrm{S}(\vect{r}_\mathrm{S})
++ H_\mathrm{MS}(\vect{r}_\mathrm{M}, \vect{r}_\mathrm{S})
+\end{equation}
+where the subscript M refers to the solute while S to the solvent. The
+coordinates $(\vect{r}_\mathrm{M}, \vect{r}_\mathrm{S})$ refer to
+\emph{both} nuclei and electrons, the interaction term is given by the
+usual Coulombic electrostatic Hamiltonian.
+
+It is important to remark that nuclear coordinates must be considered
+explicitly. Liquid phases are, by definition, characterized by the high
+mobility of their molecular constituents. Changes in the internal
+geometry of solvent molecules, triggered by intermolecular interactions
+between themselves and with the solute, always play a role in the
+solvation process. Whether this role is prominent or not cannot be
+decided \emph{a priori}.
+
+From a computational point of view, using the Hamiltonian in Eq.~\eqref{eq:solution-ham} is simply not feasible. We would have to
+determine a \acs{PES} of enormous dimensionality and then explore the full
+dynamics of the system. Even assuming that the minimum energy
+conformation on the \acs{PES} can be fully characterized, its knowledge will
+not lead us to a correct description of the physics of solvation. In
+fact, macroscopic observables of systems with a large number of degrees
+of freedom arise as \emph{average} values of the microscopic behaviour
+of every single component, the average being either on the phase space
+trajectory or on the appropriate ensemble.~\autocite{Hill1960-ql}. It is always
+necessary to perform some kind of average upon the energies
+corresponding to all the accessible conformations of the whole
+solute+solvent system.
+
+Of course, this formidable task may be accomplished using statistical
+thermodynamics and the idea of \emph{statistical ensembles} developed by
+Gibbs. To apply the idea of an ensemble average we must resort to a
+molecular dynamics or Monte Carlo simulation of the solution, in order
+to obtain the appropriate partition function, $Z$. This introduces a
+number of points that will be central in the development of continuum
+models:
+\begin{enumerate}
+ \item statistical simulations rely on \emph{macroscopic} parameters
+   such as the temperature or the density. These parameters are not
+   present in the Hamiltonian \eqref{eq:solution-ham}, but are necessary
+   for the description of molecular systems in the condensed phase. This
+   is true whether one considers discrete or continuum approaches;
+ \item the basic energetic quantity is the thermodynamic potential
+   characterizing the chosen ensemble. If a direct comparison with the
+   experiment is sought, one should run a simulation in the $(NpT)$
+   ensemble whose potential is the Gibbs free energy $G$. This point is
+   of paramount importance in continuum models and we will have the
+   opportunity to stress it in this Chapter as well as in Chapter
+   \ref{chap:relpcmscf};
+ \item even starting from a completely atomistic point of view, our
+   final results will always be in the form of an average over solvent
+   conformations. We have explicitly replaced a discrete description
+   with a continuous distribution function. Atomistic details can of
+   course be recovered by further analysis, but the point remains:
+   properties of solutions (the solvation energy among them) have an
+   intrinsic average nature.
+\end{enumerate}
+
+The fact that an averaging procedure is essential in studying solvation
+processes should convince the reader that, at least from a purely
+mathematical point of view, we can replace the full Hamiltonian
+\eqref{eq:solution-ham} with an \emph{effective} Hamiltonian, averaging
+the solvent degrees of freedom in a preliminary
+step.~\autocite{Angyan1992-vo, Tapia1992-pu}
+\todo[inline]{Some pieces missing here, FILL IN}
+Alternatively, an approximate decoupling can be obtained replacing the
+mean-field operator in Eq.~\eqref{eq:ave1} by its classical average over
+a suitable ensemble:
+\begin{equation}\label{eq:ave_classic}
+ \Braket{H_\mathrm{MS}}_\mathrm{S} = \frac{\int\diff\Gamma H_\mathrm{MS}\exp
+\left(-\beta H_\mathrm{S}\right)}{\int\diff\Gamma\exp\left(-\beta H_\mathrm{S}\right)}.
+\end{equation}
+This classical average can be written in terms of the $g(\vect{r})$ distribution functions. These functions are defined
+for the electrons and the nuclei of the solvent with respect to the fixed frame of reference of the solute molecule, see \noparcite[ref.][]{hansen}.
+This amounts to considering the solvent as a structureless continuum, hence defining continuum models.
+
+Continuum models are then characterized, by a solute-only
+\emph{effective} Hamiltonian:\footnote{We will use the symbol
+$H_\mathrm{MS}$ for the interaction Hamiltonian even if it is not the
+full interaction term of Eq.~\eqref{eq:solution-ham}.}
+\begin{equation}\label{eq:effective-ham}
+ H(\vect{r}_\mathrm{M}) = H_\mathrm{M}(\vect{r}_\mathrm{M})
++ H_\mathrm{MS}(\vect{r}_\mathrm{M}),
+\end{equation}
+the explicit form of the averaged interaction in
+Eq.~\eqref{eq:effective-ham} is in terms of solvent \emph{response
+functions}, $Q_x(\vect{r},\vect{r}^\prime)$.
+As will be apparent in the following, this term may depend on the solute
+density, making the effective Hamiltonian \emph{nonlinear}.
+
+An hierarchy of approximations for the resolution of the effective
+Schr\"odinger equation can be devised, as was done for the discrete
+models. The distinguishing feature is now the \emph{order} of the
+solvent response functions used in the definition of the interaction
+term.
+
+If we limit ourselves to a linear response of the solvent we obtain the
+class of models commonly called continuum solvation models, the
+Polarizable Continuum Model among them. Computer simulations can be used
+to determine these functions, but almost always one uses bulk properties
+of the pure solvent obtained from experiment.
+
+This, at first sight, crude approximation is rather well justified by
+the following argument: from a physical point of view, our solution is
+extremely dilute. The presence of the solute is an extremely small
+perturbation for the bulk of the solvent and it is safe to assume that
+the macroscopic properties of the solution are not far from those of the
+pure solvent. Furthermore, the use of macroscopic experimental values
+implicitly takes into account the averaging procedure one would carry
+out in atomistic simulations of the solution.
+
+As pictorially represented in Figure \ref{fig:discretetocontinuum} the
+solvent is now a \emph{structureless polarizable} continuum and the
+solute is accommodated inside a cavity formed in this continuum.
+\begin{figure}[!h]
+ \centering
+\includegraphics[width=.5\textwidth]{discretetocontinuum}
+\caption{Pictorial representation of the transition from a discrete to a
+continuum model of a solution.}
+\label{fig:discretetocontinuum}
+\end{figure}
 
@@
 \section[Variational Formulation of Classical Polarizable Models]{
-Classical Polarizable Models within a Variational Formulation}
+Above and Beyond: Classical Polarizable Models within a Variational Formulation}
 \label{sec:variational}

### 2016-08-28 Chapters/QuantumMechanics.tex
Reshuffle some sections and dump some reference around

@@
   \item single-reference methods: \ac{HF} and \ac{KS} \ac{DFT}
+  \item the \ac{CC}  model
 \end{itemize}
+
+\pagebreak
+
+\section{A Brief Summary of Coupled Cluster Theory}
+
+\todo[inline]{MOVED FROM CHAPTER 4 HERE: READJUST!}
+
+The analyses and derivations in this Chapter are preeminently based
+on the \acl{CC} wave function Ansatz.
+We will assume a closed-shell, \acs{HF} single reference function and
+construct our correlated wave function as an exponential mapping:
+\begin{equation}
+  \ket{\mathrm{CC}} = \expo{T}\ket{\mathrm{HF}}.
+\end{equation}
+The \emph{cluster} operator appearing in the exponential is given as:
+\begin{equation}
+  T = \sum_{p=1}^\mathcal{M} T_p = \sum_{p=1}^\mathcal{M}\sum_{\mu_p}\tamp{\mu_p}\cluster{\mu_p}
+\end{equation}
+that is, as a truncated sum of excitation operators, $\cluster{\mu_p}$,
+times the corresponding cluster amplitudes, $\tamp{\mu_p}$.
+Here $\mu_p$ is the $\mu$-th excitation at the $p$-th excitation level and
+$\mathcal{M}$ is the truncation level.
+One of the main strengths of the \acs{CC} model is its size extensivity
+which stems directly from the exponential nature of the wave
+operator.~\autocite{Shavitt2009-mr, Helgaker2000-tz}
+
+The \acs{CC} method is a \emph{projective} method. Chosen the truncation
+level $\mathcal{M}$, one projects the nonrelativistic Schrödinger
+equation for this Ansatz on the excitation manifold which comprises the
+reference function and all possible excited determinants, up to the
+chosen truncation order:
+\begin{subequations}
+  \begin{align}
+    \braket{\mathrm{HF} | H_0\expo{T} | \mathrm{HF}} &= E_\mathrm{CC} \\
+    \braket{\mathrm{exc} | H_0\expo{T} | \mathrm{HF}} &=
+    E_\mathrm{CC}\braket{\mathrm{exc} | \expo{T} | \mathrm{HF}}
+  \end{align}
+\end{subequations}
+where $\bra{\mathrm{exc}}$ means any of the excited determinants in Fock
+space compatible with the excitation level truncation.
+This is the \emph{unlinked} form of the \acs{CC}
+equations.~\autocite{Helgaker2000-tz} It is usually more convenient to
+perform a \emph{similarity transformation} of the Hamiltonian operator,
+leading to the \emph{linked} form of the \acs{CC} equations:
+\begin{subequations}
+  \begin{align}
+    \braket{\mathrm{HF} | \expo{-T}H_0\expo{T} | \mathrm{HF}} &= E_\mathrm{CC} \\
+    \braket{\mathrm{exc} | \expo{-T}H_0\expo{T} | \mathrm{HF}} &= 0
+  \end{align}
+\end{subequations}
+The linked and unlinked formulations are completely equivalent. However,
+the former is size extensive term-by-term.
+Use of similarity transformed operators is ubiquitous when dealing with
+\acs{CC} theory and we introduce the following notation for it:
+\begin{equation}
+  \overline{O} =\expo{-T}O\expo{T}
+\end{equation}
+The cluster operator is not Hermitian. Hence, the similarity transformation
+will not, in general, preserve any of the symmetries, such as
+hermiticity, that the bare operators might have possessed.
+Similarity-transformed operators can be expanded in a commutator series,
+also known as the \ac{BCH} expansion:
+\begin{equation}
+  \overline{O} =
+  \sum_{n=0}^{\infty}\frac{1}{n!}[A, B]^n
+\end{equation}
+where $[A,B]^n$ is the $n$-fold nested commutator. It can be shown that
+the \acs{BCH} expansion of the similarity-transformed Hamiltonian
+\emph{exactly} truncates after the four-fold nested commutator, greatly
+simplifying algebraic derivations and manipulations.
+
+Introducing the \acl{MP} partitioning of the Hamiltonian leads to
+further insight:
+\begin{equation}
+ H_0 = F + \Phi = \sum_P\epsilon_P\cons{P}\anni{P} + (g- V_\mathrm{HF})
+\end{equation}
+where $F$ if the Fock matrix and $\Phi$ is the fluctuation potential.
+Given our initial assumption on the reference function, the Fock
+operator is diagonal and expressed in terms of spin-orbital energies and
+number operators.
+Its similarity transformation truncates after the second term and has a
+relatively compact form:
+\begin{equation}
+  \overline{F} = F + \sum_{i}^4\sum_{\mu_i}\tamp{\mu_i}\cluster{\mu_i}\denom{\mu_i}
+\end{equation}
+where $\denom{\mu_i}$ is the difference in orbital energies between the
+virtual and occupied spin-orbitals of excitation $i$.
+For the fluctuation potential the similarity transformation truncates
+after the fifth term, including up to four-fold nested commutator.
+
+As formulated so far, the \acs{CC} method is not variational. It is
+however possible to introduce a fully variational Lagrangian that
+leads to the same amplitudes equations. The Lagrangian is built by
+treating the amplitude equations as constraints in the optimization of
+the \acs{CC} energy:
+\begin{equation}\label{eq:cc-lagrangian}
+  \begin{aligned}
+    \mathcal{L}^\mathcal{M}(\tamp{}, \tbar{})
+    &=
+    \braket{\mathrm{HF} | \overline{H_0} | \mathrm{HF}}
+    + \sum_{i=1}^{\mathcal{M}}\sum_{\mu_i}
+    \tbar{\mu_i}\braket{\mu_i | \overline{H_0} | \mathrm{HF}} \\
+    &=
+  E_0
+  + \sum_{i=1}^{\mathcal{M}}\tbar{\mu_i}\denom{\mu_i}\tamp{\mu_i}
+  + \braket{\mathrm{HF} | \overline{\Phi} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\Phi} | \mathrm{HF}}
+  \end{aligned}
+\end{equation}
+where the shorthand notation for the Lagrangian multiplier state was
+introduced:
+\begin{equation}
+  \bra{\tbar{j}} = \sum_{\mu_j}\tbar{\mu_j}\bra{\mu_j}
+\end{equation}
+and the \acs{MP}-partitioned form of the Hamiltonian was also exploited.
+Differentiation of the Lagrangian with respect to the multipliers
+correctly yields the amplitudes equations, while differentiation with
+respect to the amplitudes leads to the governing equations for the
+multipliers. While these are not needed for the evaluation of the
+\acs{CC} energy, their calculation is mandatory when expectation values
+and molecular properties in general are sought.
+The \acs{CC} expectation values are formed using the left and
+right \acs{CC} states and we introduce the following notation:
+\begin{equation}
+  O(\tamp{}, \tbar{}) = \braket{\mathrm{HF} | \overline{O} | \mathrm{HF} }
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{O} | \mathrm{HF} }
+\end{equation}
+When required by the context, we will explicitly write at which
+truncation level in the cluster operator the expectation was evaluated:
+$O^\mathcal{M}(\tamp{}, \tbar{})$.
+
+Having summarized the essential points of \acs{CC} theory and having
+introduced the relevant notation, we can proceed to analyze its coupling
+with quantum/classical polarizable Hamiltonians.

### 2016-08-28 Chapters/ResponseTheory.tex
Reshuffle some sections and dump some reference around

@@
     \item Something about exact state response theory: quasienergy,
-      response functions, residues of response functions.~\autocite{}
+      response functions, residues of response functions.
     \item  Derivation of the linear response function with
@@
 \end{itemize}
+
+\pagebreak
+
+\section{Exact State Response Theory}\label{sec:exact-response}
+
+General theory~\autocite{Olsen1985-nr, Christiansen1998-pe,
+Norman2011-ad, Helgaker2012-cz, Pawlowski2015-sq}
+Solution strategy for response equations~\autocite{Kauczor2011-rd}
+Open-ended \acs{AO}-based \acs{SCF} response~\autocite{
+Larsen2000-hj, Kjaergaard2008-hy, Thorvaldsen2008-sg,
+Kristensen2008-hv, Ringholm2014-gx}
+Open-ended \acs{SCF} residues~\autocite{Friese2015-kb}
+
+\section{SCF Response Theory for Quantum/Classical Polarizable Hamiltonians}\label{sec:csm-response}
+
+\acs{PCM}~\autocite{Cammi1994-qj, Cammi1996-wf, Cammi1996-vx,
+Cammi1999-rb, Cammi2003-qy, Frediani2005-nc, Ferrighi2010-pm}
+\acs{FQ} and \acs{FQ}/\acs{PCM}~\autocite{Lipparini2012-hx, Lipparini2012-tl, Lipparini2013-ud}
+\acs{PE}~\autocite{Olsen2010-wa}
+\acs{QM}/\acs{MM}/Continuum~\autocite{Stenidal2011-ki, Caprasecca2012-ir}
+Paper V~\autocite{pcm-openrsp}

### 2016-08-28 Chapters/SolvationAndCorrelation.tex
Reshuffle some sections and dump some reference around

@@
 \pagebreak
-
-\section{A Brief Summary of Coupled Cluster Theory}
-
-The analyses and derivations in this Chapter are preeminently based
-on the \acl{CC} wave function Ansatz.
-We will assume a closed-shell, \acs{HF} single reference function and
-construct our correlated wave function as an exponential mapping:
-\begin{equation}
-  \ket{\mathrm{CC}} = \expo{T}\ket{\mathrm{HF}}.
-\end{equation}
-The \emph{cluster} operator appearing in the exponential is given as:
-\begin{equation}
-  T = \sum_{p=1}^\mathcal{M} T_p = \sum_{p=1}^\mathcal{M}\sum_{\mu_p}\tamp{\mu_p}\cluster{\mu_p}
-\end{equation}
-that is, as a truncated sum of excitation operators, $\cluster{\mu_p}$,
-times the corresponding cluster amplitudes, $\tamp{\mu_p}$.
-Here $\mu_p$ is the $\mu$-th excitation at the $p$-th excitation level and
-$\mathcal{M}$ is the truncation level.
-One of the main strengths of the \acs{CC} model is its size extensivity
-which stems directly from the exponential nature of the wave
-operator.~\autocite{Shavitt2009-mr, Helgaker2000-tz}
-
-The \acs{CC} method is a \emph{projective} method. Chosen the truncation
-level $\mathcal{M}$, one projects the nonrelativistic Schrödinger
-equation for this Ansatz on the excitation manifold which comprises the
-reference function and all possible excited determinants, up to the
-chosen truncation order:
-\begin{subequations}
-  \begin{align}
-    \braket{\mathrm{HF} | H_0\expo{T} | \mathrm{HF}} &= E_\mathrm{CC} \\
-    \braket{\mathrm{exc} | H_0\expo{T} | \mathrm{HF}} &=
-    E_\mathrm{CC}\braket{\mathrm{exc} | \expo{T} | \mathrm{HF}}
-  \end{align}
-\end{subequations}
-where $\bra{\mathrm{exc}}$ means any of the excited determinants in Fock
-space compatible with the excitation level truncation.
-This is the \emph{unlinked} form of the \acs{CC}
-equations.~\autocite{Helgaker2000-tz} It is usually more convenient to
-perform a \emph{similarity transformation} of the Hamiltonian operator,
-leading to the \emph{linked} form of the \acs{CC} equations:
-\begin{subequations}
-  \begin{align}
-    \braket{\mathrm{HF} | \expo{-T}H_0\expo{T} | \mathrm{HF}} &= E_\mathrm{CC} \\
-    \braket{\mathrm{exc} | \expo{-T}H_0\expo{T} | \mathrm{HF}} &= 0
-  \end{align}
-\end{subequations}
-The linked and unlinked formulations are completely equivalent. However,
-the former is size extensive term-by-term.
-Use of similarity transformed operators is ubiquitous when dealing with
-\acs{CC} theory and we introduce the following notation for it:
-\begin{equation}
-  \overline{O} =\expo{-T}O\expo{T}
-\end{equation}
-The cluster operator is not Hermitian. Hence, the similarity transformation
-will not, in general, preserve any of the symmetries, such as
-hermiticity, that the bare operators might have possessed.
-Similarity-transformed operators can be expanded in a commutator series,
-also known as the \ac{BCH} expansion:
-\begin{equation}
-  \overline{O} =
-  \sum_{n=0}^{\infty}\frac{1}{n!}[A, B]^n
-\end{equation}
-where $[A,B]^n$ is the $n$-fold nested commutator. It can be shown that
-the \acs{BCH} expansion of the similarity-transformed Hamiltonian
-\emph{exactly} truncates after the four-fold nested commutator, greatly
-simplifying algebraic derivations and manipulations.
-
-Introducing the \acl{MP} partitioning of the Hamiltonian leads to
-further insight:
-\begin{equation}
- H_0 = F + \Phi = \sum_P\epsilon_P\cons{P}\anni{P} + (g- V_\mathrm{HF})
-\end{equation}
-where $F$ if the Fock matrix and $\Phi$ is the fluctuation potential.
-Given our initial assumption on the reference function, the Fock
-operator is diagonal and expressed in terms of spin-orbital energies and
-number operators.
-Its similarity transformation truncates after the second term and has a
-relatively compact form:
-\begin{equation}
-  \overline{F} = F + \sum_{i}^4\sum_{\mu_i}\tamp{\mu_i}\cluster{\mu_i}\denom{\mu_i}
-\end{equation}
-where $\denom{\mu_i}$ is the difference in orbital energies between the
-virtual and occupied spin-orbitals of excitation $i$.
-For the fluctuation potential the similarity transformation truncates
-after the fifth term, including up to four-fold nested commutator.
-
-As formulated so far, the \acs{CC} method is not variational. It is
-however possible to introduce a fully variational Lagrangian that
-leads to the same amplitudes equations. The Lagrangian is built by
-treating the amplitude equations as constraints in the optimization of
-the \acs{CC} energy:
-\begin{equation}\label{eq:cc-lagrangian}
-  \begin{aligned}
-    \mathcal{L}^\mathcal{M}(\tamp{}, \tbar{})
-    &=
-    \braket{\mathrm{HF} | \overline{H_0} | \mathrm{HF}}
-    + \sum_{i=1}^{\mathcal{M}}\sum_{\mu_i}
-    \tbar{\mu_i}\braket{\mu_i | \overline{H_0} | \mathrm{HF}} \\
-    &=
-  E_0
-  + \sum_{i=1}^{\mathcal{M}}\tbar{\mu_i}\denom{\mu_i}\tamp{\mu_i}
-  + \braket{\mathrm{HF} | \overline{\Phi} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\Phi} | \mathrm{HF}}
-  \end{aligned}
-\end{equation}
-where the shorthand notation for the Lagrangian multiplier state was
-introduced:
-\begin{equation}
-  \bra{\tbar{j}} = \sum_{\mu_j}\tbar{\mu_j}\bra{\mu_j}
-\end{equation}
-and the \acs{MP}-partitioned form of the Hamiltonian was also exploited.
-Differentiation of the Lagrangian with respect to the multipliers
-correctly yields the amplitudes equations, while differentiation with
-respect to the amplitudes leads to the governing equations for the
-multipliers. While these are not needed for the evaluation of the
-\acs{CC} energy, their calculation is mandatory when expectation values
-and molecular properties in general are sought.
-The \acs{CC} expectation values are formed using the left and
-right \acs{CC} states and we introduce the following notation:
-\begin{equation}
-  O(\tamp{}, \tbar{}) = \braket{\mathrm{HF} | \overline{O} | \mathrm{HF} }
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{O} | \mathrm{HF} }
-\end{equation}
-When required by the context, we will explicitly write at which
-truncation level in the cluster operator the expectation was evaluated:
-$O^\mathcal{M}(\tamp{}, \tbar{})$.
-
-Having summarized the essential points of \acs{CC} theory and having
-introduced the relevant notation, we can proceed to analyze its coupling
-with quantum/classical polarizable Hamiltonians.
 

### 2016-08-28 FrontBackmatter/Introduction.tex
Reshuffle some sections and dump some reference around

@@
 
-\todo[inline]{CURRENTLY VERBATIM FROM MASTER THESIS}
-
-Continuum models arise as limits, in a proper sense, to the discrete
-models. To stress this point we introduce the \emph{solution}
-Hamiltonian in the Born--Oppenheimer approximation as:
-\begin{equation}\label{eq:solution-ham}
- H(\vect{r}_\mathrm{M}, \vect{r}_\mathrm{S}) =
-  H_\mathrm{M}(\vect{r}_\mathrm{M}) +  H_\mathrm{S}(\vect{r}_\mathrm{S})
-+ H_\mathrm{MS}(\vect{r}_\mathrm{M}, \vect{r}_\mathrm{S})
-\end{equation}
-where the subscript M refers to the solute while S to the solvent. The
-coordinates $(\vect{r}_\mathrm{M}, \vect{r}_\mathrm{S})$ refer to
-\emph{both} nuclei and electrons, the interaction term is given by the
-usual Coulombic electrostatic Hamiltonian.
-
-It is important to remark that nuclear coordinates must be considered
-explicitly. Liquid phases are, by definition, characterized by the high
-mobility of their molecular constituents. Changes in the internal
-geometry of solvent molecules, triggered by intermolecular interactions
-between themselves and with the solute, always play a role in the
-solvation process. Whether this role is prominent or not cannot be
-decided \emph{a priori}.
-
-From a computational point of view, using the Hamiltonian in Eq.~\eqref{eq:solution-ham} is simply not feasible. We would have to
-determine a \acs{PES} of enormous dimensionality and then explore the full
-dynamics of the system. Even assuming that the minimum energy
-conformation on the \acs{PES} can be fully characterized, its knowledge will
-not lead us to a correct description of the physics of solvation. In
-fact, macroscopic observables of systems with a large number of degrees
-of freedom arise as \emph{average} values of the microscopic behaviour
-of every single component, the average being either on the phase space
-trajectory or on the appropriate ensemble.~\autocite{Hill1960-ql}. It is always
-necessary to perform some kind of average upon the energies
-corresponding to all the accessible conformations of the whole
-solute+solvent system.
-
-Of course, this formidable task may be accomplished using statistical
-thermodynamics and the idea of \emph{statistical ensembles} developed by
-Gibbs. To apply the idea of an ensemble average we must resort to a
-molecular dynamics or Monte Carlo simulation of the solution, in order
-to obtain the appropriate partition function, $Z$. This introduces a
-number of points that will be central in the development of continuum
-models:
-\begin{enumerate}
- \item statistical simulations rely on \emph{macroscopic} parameters
-   such as the temperature or the density. These parameters are not
-   present in the Hamiltonian \eqref{eq:solution-ham}, but are necessary
-   for the description of molecular systems in the condensed phase. This
-   is true whether one considers discrete or continuum approaches;
- \item the basic energetic quantity is the thermodynamic potential
-   characterizing the chosen ensemble. If a direct comparison with the
-   experiment is sought, one should run a simulation in the $(NpT)$
-   ensemble whose potential is the Gibbs free energy $G$. This point is
-   of paramount importance in continuum models and we will have the
-   opportunity to stress it in this Chapter as well as in Chapter
-   \ref{chap:relpcmscf};
- \item even starting from a completely atomistic point of view, our
-   final results will always be in the form of an average over solvent
-   conformations. We have explicitly replaced a discrete description
-   with a continuous distribution function. Atomistic details can of
-   course be recovered by further analysis, but the point remains:
-   properties of solutions (the solvation energy among them) have an
-   intrinsic average nature.
-\end{enumerate}
-
-The fact that an averaging procedure is essential in studying solvation
-processes should convince the reader that, at least from a purely
-mathematical point of view, we can replace the full Hamiltonian
-\eqref{eq:solution-ham} with an \emph{effective} Hamiltonian, averaging
-the solvent degrees of freedom in a preliminary
-step.~\autocite{Angyan1992-vo, Tapia1992-pu}
-\todo[inline]{Some pieces missing here, FILL IN}
-Alternatively, an approximate decoupling can be obtained replacing the
-mean-field operator in Eq.~\eqref{eq:ave1} by its classical average over
-a suitable ensemble:
-\begin{equation}\label{eq:ave_classic}
- \Braket{H_\mathrm{MS}}_\mathrm{S} = \frac{\int\diff\Gamma H_\mathrm{MS}\exp
-\left(-\beta H_\mathrm{S}\right)}{\int\diff\Gamma\exp\left(-\beta H_\mathrm{S}\right)}.
-\end{equation}
-This classical average can be written in terms of the $g(\vect{r})$ distribution functions. These functions are defined
-for the electrons and the nuclei of the solvent with respect to the fixed frame of reference of the solute molecule, see \noparcite[ref.][]{hansen}.
-This amounts to considering the solvent as a structureless continuum, hence defining continuum models.
-
-Continuum models are then characterized, by a solute-only
-\emph{effective} Hamiltonian:\footnote{We will use the symbol
-$H_\mathrm{MS}$ for the interaction Hamiltonian even if it is not the
-full interaction term of Eq.~\eqref{eq:solution-ham}.}
-\begin{equation}\label{eq:effective-ham}
- H(\vect{r}_\mathrm{M}) = H_\mathrm{M}(\vect{r}_\mathrm{M})
-+ H_\mathrm{MS}(\vect{r}_\mathrm{M}),
-\end{equation}
-the explicit form of the averaged interaction in
-Eq.~\eqref{eq:effective-ham} is in terms of solvent \emph{response
-functions}, $Q_x(\vect{r},\vect{r}^\prime)$.
-As will be apparent in the following, this term may depend on the solute
-density, making the effective Hamiltonian \emph{nonlinear}.
-
-An hierarchy of approximations for the resolution of the effective
-Schr\"odinger equation can be devised, as was done for the discrete
-models. The distinguishing feature is now the \emph{order} of the
-solvent response functions used in the definition of the interaction
-term.
-
-If we limit ourselves to a linear response of the solvent we obtain the
-class of models commonly called continuum solvation models, the
-Polarizable Continuum Model among them. Computer simulations can be used
-to determine these functions, but almost always one uses bulk properties
-of the pure solvent obtained from experiment.
-
-This, at first sight, crude approximation is rather well justified by
-the following argument: from a physical point of view, our solution is
-extremely dilute. The presence of the solute is an extremely small
-perturbation for the bulk of the solvent and it is safe to assume that
-the macroscopic properties of the solution are not far from those of the
-pure solvent. Furthermore, the use of macroscopic experimental values
-implicitly takes into account the averaging procedure one would carry
-out in atomistic simulations of the solution.
-
-As pictorially represented in Figure \ref{fig:discretetocontinuum} the
-solvent is now a \emph{structureless polarizable} continuum and the
-solute is accommodated inside a cavity formed in this continuum.
-\begin{figure}[!h]
- \centering
-\includegraphics[width=.5\textwidth]{discretetocontinuum}
-\caption{Pictorial representation of the transition from a discrete to a
-continuum model of a solution.}
-\label{fig:discretetocontinuum}
-\end{figure}
-
 \section*{The Road to Reality or Molecular Response Properties}

### 2016-08-28 Chapters/ContinuumSolvationModels.tex
Save work on Chapter 1

--- 2016-08-26 Chapters/ContinuumSolvationModels.tex
@@
 
-\begin{itemize}
-    \item why and how continuum solvation models
-    \item derivation of \acs{IEF}-\acs{PCM}, importance of Green's
-      functions.
-    \item \ac{BEM} and wavelet \acs{BEM} I think it's necessary to
-      mention also ddCOSMO and ddPCM as alternative strategies
-      ~\autocite{Cances2013-jg, Lipparini2013-cy, Lipparini2014-to,
-      Lipparini2014-fq, Stamm2016-fv}
-    \item variational formulation (no derivations, just mention the
-      relevant literature)
-\end{itemize}
+Developments of the \ac{PCM}, a continuum solvation model, are at the
+heart of this thesis.
+This Chapter is a brief exposition of the major points of the \acs{PCM}
+we have worked upon in the thesis.
+Section \ref{sec:IEF} will present some of the mathematical details of the
+formulation of the model. I will introduce the concept of \emph{weak}
+formulations of \acp{PDE} in Section \ref{sec:weak} and eventually show their
+use in the formulation of quantum/classical polarizable Hamiltonians in
+Section \ref{sec:variational}.
+Section \ref{sec:BEM} will be devoted to the numerical strategies used
+in solving the \acs{PCM} equation.
+
+Continuum models have a venerable history in quantum chemistry:
+\citeauthor{Onsager1936-wf}'s model\autocite{Onsager1936-wf} appeared in
+1936 and the first form of the \acs{PCM} entered the stage in
+1981.~\autocite{Miertus1981-mm}
+Section \ref{sec:CSM-why-how} will offer a heuristic "derivation" of
+continuum models.
+While it cannot be thought as formally rigorous, it highlights the major
+physical insights that have informed the creation of continuum models.
+Still nowadays much research activity is expended on the
+\acs{PCM}.~\autocite{Tomasi2011-un, Mennucci2012-dv, Lipparini2016-mo}
+I will point out throughout the Chapter to developments outside our
+group that are particularly interesting.
 
@@
 
+\section{Continuum Solvation Models: Why and How}\label{sec:CSM-why-how}
+
+The basic idea of implicit models is that the environment can
+be replaced by structureless continuum. The continuum will interact
+\emph{classically} with the explicit part of the multiscale model, an
+interaction mediated by some macroscopic property of the bulk material
+the continuum represents in the model.
+Thus, continuum models bypass both problems plaguing explicit models we
+mentioned in the Introduction, namely:
+\begin{enumerate*}[label={\alph*)},font={\color{PMS1797}}]
+\item the statistical averaging of environment configurations and
+\item the \acs{MM} region electrostatics cutoff choice.
+\end{enumerate*}
+
+Following \citeauthor{Tomasi2007-es}, we introduce the complete system
+Hamiltonian in the \acs{BO} approximation:~\autocite{Tomasi2004-dc,
+Tomasi2007-es}
+\begin{equation}\label{eq:solution-ham}
+ H(\vect{r}_\mathrm{S}, \vect{r}_\mathrm{E}) =
+  H_\mathrm{S}(\vect{r}_\mathrm{S}) +  H_\mathrm{E}(\vect{r}_\mathrm{E})
++ H_\mathrm{SE}(\vect{r}_\mathrm{S}, \vect{r}_\mathrm{E})
+\end{equation}
+The Hamiltonian features solute terms, marked by the S subscript,
+environment terms, marked by the E subscript, and interaction terms.
+The coordinates $(\vect{r}_\mathrm{S}, \vect{r}_\mathrm{E})$ refer to
+\emph{both} nuclei and electrons.
+The interaction term is given by the usual Coulomb electrostatic
+Hamiltonian.
+\todo[inline]{What is a nonpolarizable model? What's a polarizable
+model? HAS TO BE MENTIONED HERE!}
+\todo[inline]{What is the nature of the interaction? Mention that we
+only include polarization, but in a mutual manner.}
+
+One can replace the pure environment and interaction terms with their
+classical counterparts and obtain the quantum/classical, possibly
+\emph{polarizable}, Hamiltonian for an explicit \acs{QM}/\acs{MM} model.
+This brings about the first important point: the need for statistical
+averaging.
+Whenever a large number of degrees of freedom is involved, one can
+access macroscopic observables of the system by the appropriate
+averaging of the microscopically detailed motion over phase space
+trajectories or on the appropriate statistical
+ensemble.~\autocite{Hill1960-ql}
+The need for ensemble averages leads us to the following observations:
+\begin{enumerate*}[label={\alph*)},font={\color{PMS1797}}]
+ \item
+   the need for \emph{macroscopic} parameters, absent from the
+   microscopic Hamiltonian Eq.~\eqref{eq:solution-ham}, in carrying out
+   statistical simulations.
+ \item
+   Chosen a thermodynamic ensemble, the basic energetic quantity is
+   accordingly determined. For example, the Gibbs free energy  $G$ is
+   intrinsically related to the $(NpT)$ ensemble.
+ \item
+   Even atomistic simulations will lead to results that are essentially
+   averages. The discrete picture of the system has been replaced by
+   some continuous distribution function.
+\end{enumerate*}
+
+These observations strongly suggest that one can perform the averaging
+step \emph{before} embarking into the solution of the quantum mechanical
+problem and replace the full Hamiltonian with and \emph{effective}
+Hamiltonian in which the environment degrees of freedom are properly
+averaged.~\autocite{Angyan1992-vo, Tapia1992-pu}
+The solute-environment interaction term would then be replaced by
+a term only depending on the solute degrees of freedom, represented by
+the averaged, continuous response functions of the solvent.
+
+\todo[inline]{Introduction of the cavity. PCM as a problem in classical
+electrostatics}
+
 \section{Continuum Electrostatics as a Boundary Integral Problem}\label{sec:IEF}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.65\textwidth]{IEF}
+  \caption{
+  The physical setting of the \ac{PCM}. The molecular solute,
+  represented by its, possibly quantum mechanical, charge density
+  $\rho_\mathrm{i}$ is enclosed in a \emph{cavity} $\Omegai$.
+  The boundary of the cavity $\Gamma\equiv\partial\Omegai$ is a
+  2-manifold and is assumed to be $\mathcal{C}^1.
+  We assume the material properties of the cavity to be those of vacuum,
+  hence characterized by the Green's function
+  $\Gi=\frac{1}{|\vect{r}-\vect{r}^\prime|}$.
+  The cavity is carved out of an infinite, structureless, continuum
+  characterized by its Green's function $\Ge$ and fully covering the
+  external subdomain $\Omegae$.}
+  \label{fig:IEF}
+\end{figure}
 
@@
 \end{alignat}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=.65\textwidth]{IEF}
-  \caption{Vacuum Cavity in dielectric with permittivity $\diel_\mathrm{e}>1$.
-  Source charge density $\rho_\mathrm{i}$ inside the cavity
-  ($\diel_\mathrm{i}=1$). \textcolor{red}{Regular \acs{IEF}}}
-  \label{fig:case0}
-\end{figure}
 
@@
 SWIG~\autocite{Lange2010-jp, Lange2010-qo, Lange2011-eu}
+ddCOSMO and ddPCM~\autocite{Cances2013-jg, Lipparini2013-cy, Lipparini2014-to,
+Lipparini2014-fq, Stamm2016-fv}
 

### 2016-08-28 Chapters/ContinuumSolvationModels.tex
Save the day's work

@@
 
-\begin{figure}[h!]
+The physical setting of the \ac{PCM} is depicted in Figure \ref{fig:IEF}.
+The molecular solute, represented by its charge density
+$\rhoi$ is enclosed in a \emph{cavity} $\Omegai$.
+The boundary of the cavity $\Gamma\equiv\partial\Omegai$ is a
+2-manifold and is assumed to be continuously differentiable, \ie $\mathcal{C}^1$.
+The cavity is carved out of an infinite, structureless, continuum
+characterized by the bulk properties of the solvent.
+The continuum fully spans the external subdomain $\Omegae$.
+The source charge density is not assumed to be continuous: both
+classical point multipoles and quantum mechanical charge densities can
+be treated within the model.
+However, we assume that its support is entirely within the cavity,
+$\Supp(\rhoi) \subseteq \Omegai$. This assumption is, of course, false
+for quantum mechanical charge densities. It can however be proven that
+the errors induced by this \emph{charge penetration} are not so
+severe.~\autocite{Chipman2000-us, Cances2001-qs, Cances2001-qn}
+
+\begin{figure}[tb]
   \centering
-  \includegraphics[width=.65\textwidth]{IEF}
-  \caption{
+  \includegraphics[width=.5\textwidth]{IEF}
+  \caption[The physical setting of the polarizable continuum model.]{
   The physical setting of the \ac{PCM}. The molecular solute,
@@
   The boundary of the cavity $\Gamma\equiv\partial\Omegai$ is a
-  2-manifold and is assumed to be $\mathcal{C}^1.
+  continuously differentiable 2-manifold.
   We assume the material properties of the cavity to be those of vacuum,
@@
   characterized by its Green's function $\Ge$ and fully covering the
-  external subdomain $\Omegae$.}
+  external subdomain $\Omegae$.
+  }
   \label{fig:IEF}
@@
 
-We assume Euclidean space $\mathbb{R}^3$ to be partitioned into two
-subdomains $\Omegai$ and $\Omegae$ sharing a boundary
-$\Gamma\equiv\partial\Omegai$.
-We further assume that $\Omegai$ is a closed domain, entirely contained
-inside $\Omegae$. The boundary $\Gamma$ is thus a closed surface, \ie~a
-closed 2-manifold in $\mathbb{R}^3$.
-The transmission problem is posed as follows:
-\todo[inline]{Some consideration on suitable functional spaces and the metric thereof.}
+This is a well-known problem in electrostatics~\autocite{Jackson1999, Vanderlinde2005-gf}
+and is called a \emph{transmission} problem in the mathematical
+literature on \acp{PDE}.~\autocite{Hackbusch1995-uq, Sauter2011-an}
+The problem is commonly posed as follows. Given the partition of Euclidean
+space $\mathbb{R}^3$ outlined above, find the solution $u(\vect{r})$
+to the following set:
 \begin{subequations}\label{eq:transmission}
   \begin{align}[left={\empheqlbrace}]
-  \Li u &= f_\mathrm{i} \,\, \text{in}\,\, \Omegai \label{eq:internal} \\
-  \Le u &= f_\mathrm{e} \,\, \text{in}\,\, \Omegae \label{eq:external} \\
-  [u] &= \ue - \ui = g_\mathrm{D} \,\, \text{on}\,\, \Gamma
+  \Li u &= \nabla^2 u = -4\pi\rhoi \,\, \text{in}\,\, \Omegai \label{eq:internal} \\
+  \Le u &= 0 \,\, \text{in}\,\, \Omegae \label{eq:external} \\
+  [u] &= \ue - \ui = 0 \,\, \text{on}\,\, \Gamma
   \label{eq:trace-jump} \\
-[\partial_L u] &= \partiale u - \partiali u = g_\mathrm{N} \,\,
-\text{on}\,\, \Gamma \label{eq:conormal-jump} \\
-|u| &\leq C \|x \|^{-1} \,\,\text{for}\,\,\| x \|\rightarrow\infty
+[\partial_L u] &= \partiale u - \partiali u = 0 \,\, \text{on}\,\, \Gamma \label{eq:conormal-jump} \\
+|u(\vect{r})| &\leq C \|\vect{r} \|^{-1} \,\,\text{for}\,\,\| \vect{r} \|\rightarrow\infty
 \label{eq:radiation}
@@
 \end{subequations}
-where the jump conditions (eqs. \eqref{eq:trace-jump} and
-\eqref{eq:conormal-jump}) are given in terms of Dirichlet and Neumann
-data for the solution $u$. The jump conditions are expressed in terms of
-trace operators for the solution $u$ and its conormal derivative. For
-notational simplicity, will use the symbols $\partiale$ and $\partiali$
+Eqs.~\eqref{eq:trace-jump} and \eqref{eq:conormal-jump}
+are the \emph{jump conditions}, expressed in terms of trace operators
+for the solution $u$ and its \emph{conormal} derivative.
+For notational simplicity, will use the symbols $\partiale$ and $\partiali$
 for the latter and only give it in explicit form when needed.
@@
 \citeauthor{Sauter2011-an}.~\autocite{Sauter2011-an}.
-The
-fundamental solutions, or Green's functions, for the elliptic
+The fundamental solutions, or Green's functions, for the elliptic
 differential operators $\Li$ and $\Le$ will be denoted by $\Gi$ and
 $\Ge$, respectively.
+As such the problem is formulated in the so-called \emph{strong} form:
+the sought-after solution $u$ has to be at least two times continuously
+differentiable.
+Moreover, the solution is a function supported over $\mathbb{R}^3$ which
+poses challenges to the numerical solution of the problem.
+We will take a different approach and reformulate it in terms of
+\emph{boundary integral operators}, \ie transform the \acsp{PDE} into a
+\ac{BIE}.
 
-We introduce the relevant boundary integral operators for the
-abovementioned transmission problem. They are three of the four
-components of the Calder\'on projector.
-For $v \in L^2(\Gamma)$, $x, y \in \Gamma$:
-\footnote{We can assume a less restrictive functional space, \ie a
-Sobolev space of fractional order.}
+The first step is to introduce the relevant boundary integral operators for the
+abovementioned transmission problem.
+They are three of the four components of the Calder\'on projector.
+For any function $v$ in the suitable function space and for $\vect{s}, \vect{s}^\prime \in \Gamma$:
 \begin{equation}
 \begin{aligned}
-(\bi{S}_\star v)(x) &= \int_\Gamma G_\star(x, y)v(y)\diff y \\
-(\bi{D}_\star v)(x) &= \int_\Gamma \partial_{L_\star,y}G_\star(x, y)v(y)\diff y \\
-(\bi{D}^\dagger_\star v)(x) &= \int_\Gamma \partial_{L_\star,x}G_\star(x, y)v(y)\diff y \\
+  (\bi{S}_\star v)(\vect{s}) &= \int_\Gamma
+  G_\star(\vect{s}, \vect{s}^\prime)v(\vect{s}^\prime)\diff \vect{s}^\prime \\
+  (\bi{D}_\star v)(\vect{s}) &= \int_\Gamma \partial_{L_\star,\vect{s}}
+  G_\star(\vect{s}, \vect{s}^\prime)v(\vect{s}^\prime)\diff \vect{s}^\prime \\
+  (\bi{D}^\dagger_\star v)(\vect{s}) &= \int_\Gamma
+\partial_{L_\star,\vect{s}^\prime}
+  G_\star(\vect{s}, \vect{s}^\prime)v(\vect{s}^\prime)\diff \vect{s}^\prime \\
 \end{aligned}
@@
 where $\star$ is a placeholder for the $\mathrm{i}$ or $\mathrm{e}$ subscript.
-
-\begin{lemma}[Properties of the boundary integral operators]
-  The integral operators introduced above enjoy the following
-  properties~\autocite{Hsiao2008-xb, Sauter2011-an}:
-  \begin{enumerate}
-      \item on $L^2(\Gamma)$ $\bi{S}_\star$ is self-adjoint,
-        $\bi{D}^\dagger_\star$ is the adjoint operator of
-        $\bi{D}_\star$.
-      \item The commutation relations:
-        \begin{alignat}{2}
-          \bi{D}_\star\bi{S}_\star = \bi{S}_\star\bi{D}^\dagger_\star, \quad&
-          \bi{S}_\star\bi{D}_\star = \bi{D}^\dagger_\star\bi{S}_\star
-        \end{alignat}
-        hold
-      \item The operators:
-        \begin{subequations}
-          \begin{align}
-   \bi{S}_\star &: H^{-\frac{1}{2}}(\Gamma) \rightarrow H^{\frac{1}{2}}(\Gamma) \\
-   \bi{D}_\star &: H^{\frac{1}{2}}(\Gamma) \rightarrow H^{\frac{1}{2}}(\Gamma) \\
-   \bi{D}^\dagger_\star &: H^{-\frac{1}{2}}(\Gamma) \rightarrow H^{-\frac{1}{2}}(\Gamma)
-          \end{align}
-        \end{subequations}
-      \item The operator $\bi{S}_\star$ is coervice and admits a
-        continuous inverse in the aforementioned Sobolev spaces.
-      \item The operators $\lambda - \bi{D}_\star$ and $\lambda -
-        \bi{D}^\dagger_\star$ with $\lambda \in (-2\pi, +\infty)$
-        admit a continuous inverse in the aforementioned
-        Sobolev spaces.
-  \end{enumerate}
-\end{lemma}
-
-\subsection{Lemma of Integral Representation}
-For the transmission problem \ref{eq:transmission} there holds:
-\begin{itemize}
-\item $\forall x \in \Omegai$
+Our next step is to rewrite the $u$ as the sum of two contributions:
 \begin{equation}
-u = \bi{S}_\mathrm{i}(\partiali u)
-- \bi{D}_\mathrm{i}(\ui) + \int_{\Omegai} \Gi f_\mathrm{i}\diff y
+  u(\vect{r}) = \esp(\vect{r}) + \xi(\vect{r}) = \int_{\Omegai}
+  \frac{\rhoi(\vect{r}^\prime)}{|\vect{r}-\vect{r}^\prime|}
+  + \xi(\vect{r})
 \end{equation}
-\item $\forall x \in \Omegae$
-\begin{equation}
-u = -\bi{S}_\mathrm{e}(\partiale u)
-+ \bi{D}_\mathrm{e}(\ue) + \int_{\Omegae}\Ge f_\mathrm{e}\diff y
-\end{equation}
-\item $\forall x \in \Gamma$
-  \begin{equation}\label{eq:stat3}
-    \frac{1}{2}\ui = \bi{S}_\mathrm{i}(\partiali u)
-    - \bi{D}_\mathrm{i}(\ui) + \int_{\Omegai}\Gi f_\mathrm{i}\diff y
-  \end{equation}
-\item $\forall x \in \Gamma$
-  \begin{equation}\label{eq:stat4}
-    \frac{1}{2}\ue = -\bi{S}_\mathrm{e}(\partiale u)
-    + \bi{D}_\mathrm{e}(\ue) + \int_{\Omegae}\Ge f_\mathrm{e}\diff y
-  \end{equation}
-\end{itemize}
-
-Note that the figures refer to cases where the dielectric is supposed to
-be homogeneous and isotropic. The derivation are, however, general and
-do not assume homogeneity and isotropy.
-The source charge densities are not assumed to be continuous, the
-hypothesis that they are entirely supported in either of the internal or
-external subdomains is however made:
-\begin{alignat}{2}
-  \Supp(\rhoi) \subseteq \Omegai, \quad& \Supp(\rhoe) \subseteq \Omegae
-\end{alignat}
-
-This is the regular \acs{IEF} as derived in:~\autocite{Cances1998-og}
-\begin{subequations}
-\begin{align}
-  \nabla^2 u &= \rhoi \,\, \text{in}\,\, \Omegai \\
-  \Le u &= 0 \,\, \text{in}\,\, \Omegae \\
-  [u] &= \ue - \ui = 0 \,\, \text{on}\,\, \Gamma \\
-[\partial_L u] &= \partiale u - \partiali u = 0 \,\,
-\text{on}\,\, \Gamma \\
-|u| &\leq C \|x \|^{-1} \,\,\text{for}\,\,\| x \|\rightarrow\infty
-\end{align}
-\end{subequations}
-
-We introduce the following auxiliary potential:
-\begin{equation}
-  h(x) =
-  \begin{cases}
-    &\int_{\mathbb{R}^3}\Gi\rhoi(y)\diff y  \quad x \in \Omegai \\
-    &\int_{\mathbb{R}^3}\Ge\rhoi(y)\diff y  \quad x \in \Omegae
-  \end{cases}
-\end{equation}
-for which we have:\footnote{This can be checked by applying the
-differential operators to the definition of $h$ and then swapping
-differentiation and integration.}
-\begin{equation}
-  \begin{cases}
-    \nabla^2 h &= \rhoi \,\, \text{in}\,\, \Omegai \\
-    \Le h &= 0 \,\, \text{in}\,\, \Omegae
-  \end{cases}
-\end{equation}
-We then define the \emph{reaction potential} as:
-\begin{equation}
-  \xi = u - h
-\end{equation}
-such that:
-\begin{subequations}
-\begin{align}
-  \nabla^2 \xi &= 0 \,\, \text{in}\,\, \Omegai \\
-  \Le \xi &= 0 \,\, \text{in}\,\, \Omegae \\
-  -[\xi] &= [h] \,\, \text{on}\,\, \Gamma \\
-  -[\partial_L \xi] &= [\partial_L h] \,\, \text{on}\,\, \Gamma
-\end{align}
-\end{subequations}
-inside the cavity, the reaction potential can be represented by a
-\emph{single layer potential}, \ie by an apparent surface charge (ASC):
+the former is the electrostatic potential of $\rhoi$ \emph{in vacuo}
+(Newton potential), while the latter is the \emph{reaction} potential.
+The reaction potential describes the polarization in the medium.
+Moreover, it admits a integral representation as a single-layer
+potential:
 \begin{equation}
@@
 \end{equation}
-
-Applying Eqs. \eqref{eq:stat3} and \eqref{eq:stat4} of the Lemma of Integral
-Representation to the reaction potential yields:
-\begin{align}
-  \frac{1}{2}\xi_\mathrm{i} &= \bi{S}_\mathrm{i}(\partiali \xi) -
-  \bi{D}_\mathrm{i}(\xi_\mathrm{i}) \\
-  -\frac{1}{2}\xi_\mathrm{e} &= \bi{S}_\mathrm{e}(\partiale \xi) -
-  \bi{D}_\mathrm{e}(\xi_\mathrm{e})
-\end{align}
-in both cases the last term, including the integration over the entire
-subdomain volume, drops out ($f_\mathrm{i} = 0 = f_\mathrm{e}$).
-While use of Eq. \eqref{eq:stat4} on the auxiliary potential $h$ yields:
-\begin{equation}
-  -\frac{1}{2}h_\mathrm{e} =
-  \bi{S}_\mathrm{e}(\partiale h) -
-  \bi{D}_\mathrm{e}(h_\mathrm{e})
-\end{equation}
-Together with the jump conditions one obtains:
-\begin{subequations}
-  \begin{align}
-    &\bi{S}_\mathrm{i}(\partiali \xi) - \bi{D}_\mathrm{i}(\xi_\mathrm{i})
-    = \frac{1}{2}\xi_\mathrm{i} \label{eq:first} \\
-    &\bi{S}_\mathrm{e}(\partiale \xi) - \bi{D}_\mathrm{e}(\xi_\mathrm{e})
-    = -\frac{1}{2}\xi_\mathrm{e} \label{eq:second} \\
-  &\bi{S}_\mathrm{e}(\partiale h) - \bi{D}_\mathrm{e}(h_\mathrm{e})
-  = -\frac{1}{2}h_\mathrm{e} \label{eq:third} \\
-  &\xi_\mathrm{i} - \xi_\mathrm{e} = h_\mathrm{e} - h_\mathrm{i}
-  \label{eq:fourth} \\
-  &\partiali \xi - \partiale \xi = \partiale h - \partiali h
-  \label{eq:fifth}
-  \end{align}
-\end{subequations}
-
-\begin{enumerate}
-  \item Add Eqs. \eqref{eq:second} and \eqref{eq:third}:
-    \begin{equation}
-    \bi{S}_\mathrm{e}(\partiale \xi + \partiale h) -
-    \bi{D}_\mathrm{e}(\xi_\mathrm{e} + h_\mathrm{e})
-    = -\frac{1}{2}(\xi_\mathrm{e} + h_\mathrm{e})
-    \end{equation}
-  \item Exploit Eqs. \eqref{eq:fourth} and \eqref{eq:fifth} to
-    eliminate the external traces and conormal derivatives:
-    \begin{equation}
-      \bi{S}_\mathrm{e}(\partiali \xi) + \left(\frac{1}{2} -
-      \bi{D}_\mathrm{e}\right)\xi_\mathrm{i} =
-      - \bi{S}_\mathrm{e}(\partiali h) - \left(\frac{1}{2} -
-      \bi{D}_\mathrm{e}\right)h_\mathrm{i}
-    \end{equation}
-  \item Introduce the integral representation of
-    $\xi_\mathrm{i}$:
-    \begin{equation}
-      \bi{S}_\mathrm{e}(\partiali \xi) + \left(\frac{1}{2} -
-      \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}\sigma =
-      - \bi{S}_\mathrm{e}(\partiali h) - \left(\frac{1}{2} -
-      \bi{D}_\mathrm{e}\right)h_\mathrm{i}
-    \end{equation}
-  \item Reorganize Eq. \eqref{eq:first}, using the commutation
-    relation and the integral representation of $\xi_\mathrm{i}$:
-    \begin{equation}
-    \partiali \xi = \left(\frac{1}{2} + \bi{D}^\dagger_\mathrm{i}\right)\sigma
-    \end{equation}
-    such that:
-    \begin{equation}
-      \left[ \bi{S}_\mathrm{e}\left(\frac{1}{2} + \bi{D}^\dagger_\mathrm{i}\right)
-      +
-      \left(\frac{1}{2} - \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}
-      \right]\sigma =
-      - \bi{S}_\mathrm{e}(\partiali h) - \left(\frac{1}{2} -
-      \bi{D}_\mathrm{e}\right)h_\mathrm{i}
-    \end{equation}
-  \item Employ the \ac{DtN} map to simplify
-    the RHS. The \acs{DtN} map is derived by applying Eq. \eqref{eq:stat3} to
-    the Newton potential:
-    \begin{equation}
-      \phi(x) = (\bi{N}\rhoi)(x) = \int_{\mathbb{R}^3}\Gi\rhoi(y)\diff y
-      = \left.h(x)\right|_{\Omegai}
-    \end{equation}
-    which is equal, in $\Omegai$, to the auxiliary potential $h(x)$.
-    With this definition one has:
-    \begin{equation}
-      \frac{1}{2}\phi_\mathrm{i} =
-      \bi{S}_\mathrm{i}(\partiali \phi) -
-      \bi{D}_\mathrm{i}(\phi_\mathrm{i}) + \int_{\Omegai}\Gi\rhoi(y)\diff y
-      =
-      \bi{S}_\mathrm{i}(\partiali \phi) -
-      \bi{D}_\mathrm{i}(\phi_\mathrm{i}) + \phi_\mathrm{i}
-    \end{equation}
-    eventually leading to the \acs{DtN} map:
-    \begin{equation}\label{eq:regular-DtN}
-      \left(\frac{1}{2} - \bi{D}_\mathrm{i}\right)\phi_\mathrm{i}
-      +\bi{S}_\mathrm{i}(\partiali \phi) = 0
-    \end{equation}
-  \item Having identified $\phi_\mathrm{i} \equiv h_\mathrm{i}$
-    and $\partiali \phi \equiv \partiali h$ and employing the \acs{DtN} map,
-    Eq.\eqref{eq:regular-DtN}, one obtains the \acs{IEF}-\acs{PCM} equation:
+where $\sigma$ is the \ac{ASC}, the key quantity in the \acs{PCM}.
+As shown by \citeauthor{Cances1998-og}, the \acs{ASC} can be computed
+as the \emph{unique} solution to the following integral equation:
 \begin{equation}\label{eq:full-IEF}
-      \left[ \bi{S}_\mathrm{e}\left(\frac{1}{2} + \bi{D}^\dagger_\mathrm{i}\right)
-      +
-      \left(\frac{1}{2} - \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}
-      \right]\sigma =
-      -\left[\left(\frac{1}{2}-\bi{D}_\mathrm{e}\right)
-      -\bi{S}_\mathrm{e}\bi{S}_\mathrm{i}^{-1}
-       \left(\frac{1}{2}-\bi{D}_\mathrm{i}\right)
-      \right]\phi
-    \end{equation}
-    where the subscript $\mathrm{i}$ for the Newton potential $\phi$ has
-    been dropped, given that it is continuous across the boundary
-    $\Gamma$.
-\end{enumerate}
-The above proves the existence of a representation of the reaction
-potential valid \emph{inside} the cavity in terms of an apparent surface
-charge density. We do not prove the uniqueness of this solution and
-refer the interested reader to the original
-paper.~\autocite{Cances1998-og}
-
-In the code, the Green's functions are defined \emph{without} the $4\pi$
-factors in the denominator. Instead of $\frac{1}{2}$ as coefficients to
-the identity operator, one will then have $2\pi$:
-\begin{equation}\label{eq:full-IEF-2pi}
   \left[ \bi{S}_\mathrm{e}\left(2\pi + \bi{D}^\dagger_\mathrm{i}\right)
@@
   \left(2\pi-\bi{D}_\mathrm{i}\right)
-  \right]\phi
+  \right]\esp
 \end{equation}
+commonly called the \ac{IEF} equation.
+The Green's functions for the elliptic differential operators
+assume a central role in the boundary integral formulation and make the
+\acs{IEF} equation valid for a wide range of external
+environments: homogeneous isotropic, ionic and anisotropic
+liquids,~\autocite{Cances1998-og}
+and systems with interfaces.~\autocite{Corni2002-dr, Frediani2004-er,
+Delgado2013-kd, DiRemigio2016-nn}
 
-From Eq. \eqref{eq:full-IEF-2pi} we can derive the isotropic
-\acs{IEF}, the \acs{DPCM} and the \acs{CPCM} (by letting
-$\diel\rightarrow\infty$) equations.
-For a homogeneous, isotropic dielectric with permittivity $\diel$ the
-Green's function outside the cavity is simply given by scaling the one
-inside:
+In the special and very common case of homogeneous and isotropic
+environments, \ie regular dielectric materials characterized by a scalar
+permittivity $\diel$, the Green's functions are:
+\begin{alignat}{2}
+  \Ge = \frac{1}{\diel|\vect{r}-\vect{r}^\prime|} = \frac{1}{\diel} \Gi,
+  \quad&
+  \Gi = \frac{1}{|\vect{r}-\vect{r}^\prime|}
+\end{alignat}
+yielding the isotropic \acs{IEF} equation:
+\begin{equation}\label{eq:isotropic-IEF}
+  \bi{R}_\diel\bi{S}\sigma = - \bi{R}_\infty\esp
+\end{equation}
+where the auxiliary operators:
+\begin{alignat}{2}
+  \bi{R}_\diel = \left[
+  2\pi \left(\frac{\diel+1}{\diel-1}\right) - \bi{D}
+  \right],
+  \quad&
+  \bi{R}_\infty =  \lim_{\diel\rightarrow\infty}\bi{R}_\diel
+  = 2\pi-\bi{D}
+\end{alignat}
+have been introduced.
+By letting the $\diel\rightarrow\infty$, one recovers the limiting case
+where conductor boundary conditions are imposed. This leads to the
+well-known \ac{COSMO} equation:~\autocite{Klamt1993-mj, Cossi2003-xe}
 \begin{equation}
-  \Ge = \frac{1}{\diel|x-y|} = \frac{1}{\diel} \Gi = \frac{1}{|x-y|}
-\end{equation}
-leading to the boundary integral operators:
-\begin{alignat}{2}
-  \bi{S}_\mathrm{i} = \diel\bi{S}_\mathrm{e} = \bi{S}, \quad&
-  \bi{D}_\mathrm{i} = \bi{D}_\mathrm{e} = \bi{D}
-\end{alignat}
-Eventually, one has:
-\begin{equation}\label{eq:isotropic-IEF}
-  \left[ 2\pi \left(\frac{\diel+1}{\diel-1}\right)
-  - \bi{D} \right]\bi{S}\sigma
-  =
-  -\left[2\pi-\bi{D}\right]\phi
+  \bi{S}\sigma = - \esp
 \end{equation}

### 2016-08-28 Chapters/MathematicalResults.tex
Save the day's work

\chapter{Some Mathematical Results}\label{app:mathematical-results}

\begin{lemma}[Properties of the boundary integral operators]
  The integral operators introduced above enjoy the following
  properties~\autocite{Hsiao2008-xb, Sauter2011-an}:
  \begin{enumerate}
      \item on $L^2(\Gamma)$, $\bi{S}_\star$ is self-adjoint,
        $\bi{D}^\dagger_\star$ is the adjoint operator of
        $\bi{D}_\star$.
      \item The commutation relations hold:
        \begin{alignat}{2}
          \bi{D}_\star\bi{S}_\star = \bi{S}_\star\bi{D}^\dagger_\star, \quad&
          \bi{S}_\star\bi{D}_\star = \bi{D}^\dagger_\star\bi{S}_\star
        \end{alignat}
      \item The boundary integral operators are continuous
        mappings between Sobolev spaces of fractional order:
        \begin{subequations}
          \begin{align}
   \bi{S}_\star &: H^{-\frac{1}{2}}(\Gamma) \rightarrow H^{\frac{1}{2}}(\Gamma) \\
   \bi{D}_\star &: H^{\frac{1}{2}}(\Gamma) \rightarrow H^{\frac{1}{2}}(\Gamma) \\
   \bi{D}^\dagger_\star &: H^{-\frac{1}{2}}(\Gamma) \rightarrow H^{-\frac{1}{2}}(\Gamma)
          \end{align}
        \end{subequations}
      \item The operator $\bi{S}_\star$ is coervice and admits a
        continuous inverse in the aforementioned Sobolev spaces.
      \item The operators $\lambda - \bi{D}_\star$ and $\lambda -
        \bi{D}^\dagger_\star$ with $\lambda \in (-2\pi, +\infty)$
        admit a continuous inverse in the aforementioned
        Sobolev spaces.
  \end{enumerate}
\end{lemma}

\subsection{Lemma of Integral Representation}
For the transmission problem \ref{eq:transmission} there holds:
\begin{itemize}
\item $\forall x \in \Omegai$
\begin{equation}
u = \bi{S}_\mathrm{i}(\partiali u)
- \bi{D}_\mathrm{i}(\ui) + \int_{\Omegai} \Gi f_\mathrm{i}\diff y
\end{equation}
\item $\forall x \in \Omegae$
\begin{equation}
u = -\bi{S}_\mathrm{e}(\partiale u)
+ \bi{D}_\mathrm{e}(\ue) + \int_{\Omegae}\Ge f_\mathrm{e}\diff y
\end{equation}
\item $\forall x \in \Gamma$
  \begin{equation}\label{eq:stat3}
    \frac{1}{2}\ui = \bi{S}_\mathrm{i}(\partiali u)
    - \bi{D}_\mathrm{i}(\ui) + \int_{\Omegai}\Gi f_\mathrm{i}\diff y
  \end{equation}
\item $\forall x \in \Gamma$
  \begin{equation}\label{eq:stat4}
    \frac{1}{2}\ue = -\bi{S}_\mathrm{e}(\partiale u)
    + \bi{D}_\mathrm{e}(\ue) + \int_{\Omegae}\Ge f_\mathrm{e}\diff y
  \end{equation}
\end{itemize}

We introduce the following auxiliary potential:
\begin{equation}
  h(x) =
  \begin{cases}
    &\int_{\mathbb{R}^3}\Gi\rhoi(y)\diff y  \quad x \in \Omegai \\
    &\int_{\mathbb{R}^3}\Ge\rhoi(y)\diff y  \quad x \in \Omegae
  \end{cases}
\end{equation}
for which we have:\footnote{This can be checked by applying the
differential operators to the definition of $h$ and then swapping
differentiation and integration.}
\begin{equation}
  \begin{cases}
    \nabla^2 h &= \rhoi \,\, \text{in}\,\, \Omegai \\
    \Le h &= 0 \,\, \text{in}\,\, \Omegae
  \end{cases}
\end{equation}
We then define the \emph{reaction potential} as:
\begin{equation}
  \xi = u - h
\end{equation}
such that:
\begin{subequations}
\begin{align}
  \nabla^2 \xi &= 0 \,\, \text{in}\,\, \Omegai \\
  \Le \xi &= 0 \,\, \text{in}\,\, \Omegae \\
  -[\xi] &= [h] \,\, \text{on}\,\, \Gamma \\
  -[\partial_L \xi] &= [\partial_L h] \,\, \text{on}\,\, \Gamma
\end{align}
\end{subequations}
inside the cavity, the reaction potential can be represented by a
\emph{single layer potential}, \ie by an apparent surface charge (ASC):
\begin{equation}
  \xi_\mathrm{i} = \bi{S}_\mathrm{i}\sigma
\end{equation}

Applying Eqs. \eqref{eq:stat3} and \eqref{eq:stat4} of the Lemma of Integral
Representation to the reaction potential yields:
\begin{align}
  \frac{1}{2}\xi_\mathrm{i} &= \bi{S}_\mathrm{i}(\partiali \xi) -
  \bi{D}_\mathrm{i}(\xi_\mathrm{i}) \\
  -\frac{1}{2}\xi_\mathrm{e} &= \bi{S}_\mathrm{e}(\partiale \xi) -
  \bi{D}_\mathrm{e}(\xi_\mathrm{e})
\end{align}
in both cases the last term, including the integration over the entire
subdomain volume, drops out ($f_\mathrm{i} = 0 = f_\mathrm{e}$).
While use of Eq. \eqref{eq:stat4} on the auxiliary potential $h$ yields:
\begin{equation}
  -\frac{1}{2}h_\mathrm{e} =
  \bi{S}_\mathrm{e}(\partiale h) -
  \bi{D}_\mathrm{e}(h_\mathrm{e})
\end{equation}
Together with the jump conditions one obtains:
\begin{subequations}
  \begin{align}
    &\bi{S}_\mathrm{i}(\partiali \xi) - \bi{D}_\mathrm{i}(\xi_\mathrm{i})
    = \frac{1}{2}\xi_\mathrm{i} \label{eq:first} \\
    &\bi{S}_\mathrm{e}(\partiale \xi) - \bi{D}_\mathrm{e}(\xi_\mathrm{e})
    = -\frac{1}{2}\xi_\mathrm{e} \label{eq:second} \\
  &\bi{S}_\mathrm{e}(\partiale h) - \bi{D}_\mathrm{e}(h_\mathrm{e})
  = -\frac{1}{2}h_\mathrm{e} \label{eq:third} \\
  &\xi_\mathrm{i} - \xi_\mathrm{e} = h_\mathrm{e} - h_\mathrm{i}
  \label{eq:fourth} \\
  &\partiali \xi - \partiale \xi = \partiale h - \partiali h
  \label{eq:fifth}
  \end{align}
\end{subequations}

\begin{enumerate}
  \item Add Eqs. \eqref{eq:second} and \eqref{eq:third}:
    \begin{equation}
    \bi{S}_\mathrm{e}(\partiale \xi + \partiale h) -
    \bi{D}_\mathrm{e}(\xi_\mathrm{e} + h_\mathrm{e})
    = -\frac{1}{2}(\xi_\mathrm{e} + h_\mathrm{e})
    \end{equation}
  \item Exploit Eqs. \eqref{eq:fourth} and \eqref{eq:fifth} to
    eliminate the external traces and conormal derivatives:
    \begin{equation}
      \bi{S}_\mathrm{e}(\partiali \xi) + \left(\frac{1}{2} -
      \bi{D}_\mathrm{e}\right)\xi_\mathrm{i} =
      - \bi{S}_\mathrm{e}(\partiali h) - \left(\frac{1}{2} -
      \bi{D}_\mathrm{e}\right)h_\mathrm{i}
    \end{equation}
  \item Introduce the integral representation of
    $\xi_\mathrm{i}$:
    \begin{equation}
      \bi{S}_\mathrm{e}(\partiali \xi) + \left(\frac{1}{2} -
      \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}\sigma =
      - \bi{S}_\mathrm{e}(\partiali h) - \left(\frac{1}{2} -
      \bi{D}_\mathrm{e}\right)h_\mathrm{i}
    \end{equation}
  \item Reorganize Eq. \eqref{eq:first}, using the commutation
    relation and the integral representation of $\xi_\mathrm{i}$:
    \begin{equation}
    \partiali \xi = \left(\frac{1}{2} + \bi{D}^\dagger_\mathrm{i}\right)\sigma
    \end{equation}
    such that:
    \begin{equation}
      \left[ \bi{S}_\mathrm{e}\left(\frac{1}{2} + \bi{D}^\dagger_\mathrm{i}\right)
      +
      \left(\frac{1}{2} - \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}
      \right]\sigma =
      - \bi{S}_\mathrm{e}(\partiali h) - \left(\frac{1}{2} -
      \bi{D}_\mathrm{e}\right)h_\mathrm{i}
    \end{equation}
  \item Employ the \ac{DtN} map to simplify
    the RHS. The \acs{DtN} map is derived by applying Eq. \eqref{eq:stat3} to
    the Newton potential:
    \begin{equation}
      \phi(x) = (\bi{N}\rhoi)(x) = \int_{\mathbb{R}^3}\Gi\rhoi(y)\diff y
      = \left.h(x)\right|_{\Omegai}
    \end{equation}
    which is equal, in $\Omegai$, to the auxiliary potential $h(x)$.
    With this definition one has:
    \begin{equation}
      \frac{1}{2}\phi_\mathrm{i} =
      \bi{S}_\mathrm{i}(\partiali \phi) -
      \bi{D}_\mathrm{i}(\phi_\mathrm{i}) + \int_{\Omegai}\Gi\rhoi(y)\diff y
      =
      \bi{S}_\mathrm{i}(\partiali \phi) -
      \bi{D}_\mathrm{i}(\phi_\mathrm{i}) + \phi_\mathrm{i}
    \end{equation}
    eventually leading to the \acs{DtN} map:
    \begin{equation}\label{eq:regular-DtN}
      \left(\frac{1}{2} - \bi{D}_\mathrm{i}\right)\phi_\mathrm{i}
      +\bi{S}_\mathrm{i}(\partiali \phi) = 0
    \end{equation}
  \item Having identified $\phi_\mathrm{i} \equiv h_\mathrm{i}$
    and $\partiali \phi \equiv \partiali h$ and employing the \acs{DtN} map,
    Eq.\eqref{eq:regular-DtN}, one obtains the \acs{IEF}-\acs{PCM} equation:
    \begin{equation}\label{eq:full-IEF}
      \left[ \bi{S}_\mathrm{e}\left(\frac{1}{2} + \bi{D}^\dagger_\mathrm{i}\right)
      +
      \left(\frac{1}{2} - \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}
      \right]\sigma =
      -\left[\left(\frac{1}{2}-\bi{D}_\mathrm{e}\right)
      -\bi{S}_\mathrm{e}\bi{S}_\mathrm{i}^{-1}
       \left(\frac{1}{2}-\bi{D}_\mathrm{i}\right)
      \right]\phi
    \end{equation}
    where the subscript $\mathrm{i}$ for the Newton potential $\phi$ has
    been dropped, given that it is continuous across the boundary
    $\Gamma$.
\end{enumerate}
The above proves the existence of a representation of the reaction
potential valid \emph{inside} the cavity in terms of an apparent surface
charge density. We do not prove the uniqueness of this solution and
refer the interested reader to the original
paper.~\autocite{Cances1998-og}

In the code, the Green's functions are defined \emph{without} the $4\pi$
factors in the denominator. Instead of $\frac{1}{2}$ as coefficients to
the identity operator, one will then have $2\pi$:
\begin{equation}\label{eq:full-IEF-2pi}
  \left[ \bi{S}_\mathrm{e}\left(2\pi + \bi{D}^\dagger_\mathrm{i}\right)
  +
  \left(2\pi - \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}
  \right]\sigma =
  -\left[\left(2\pi-\bi{D}_\mathrm{e}\right)
  -\bi{S}_\mathrm{e}\bi{S}_\mathrm{i}^{-1}
  \left(2\pi-\bi{D}_\mathrm{i}\right)
  \right]\phi
\end{equation}

From Eq. \eqref{eq:full-IEF-2pi} we can derive the isotropic
\acs{IEF}, the \acs{DPCM} and the \acs{CPCM} (by letting
$\diel\rightarrow\infty$) equations.
For a homogeneous, isotropic dielectric with permittivity $\diel$ the
Green's function outside the cavity is simply given by scaling the one
inside:
\begin{equation}
  \Ge = \frac{1}{\diel|x-y|} = \frac{1}{\diel} \Gi = \frac{1}{|x-y|}
\end{equation}
leading to the boundary integral operators:
\begin{alignat}{2}
  \bi{S}_\mathrm{i} = \diel\bi{S}_\mathrm{e} = \bi{S}, \quad&
  \bi{D}_\mathrm{i} = \bi{D}_\mathrm{e} = \bi{D}
\end{alignat}
Eventually, one has:
\begin{equation}\label{eq:isotropic-IEF}
  \left[ 2\pi \left(\frac{\diel+1}{\diel-1}\right)
  - \bi{D} \right]\bi{S}\sigma
  =
  -\left[2\pi-\bi{D}\right]\phi
\end{equation}

### 2016-08-28 Chapters/PaperSummary.tex
Save the day's work

@@
 independently of the particular host program in which they will be
-used.~\cite{Dijkstra1968-vy, Parnas1972-im}
+used.~\cite{Dijkstra1968-zp, Parnas1972-im}
 A well-defined \ac{API} clearly delimits the boundaries of the functionality

### 2016-08-28 FrontBackmatter/Introduction.tex
Save work on Chapter 1
Save the day's work

@@
 \scalebox{0.8}{\input{gfx/RQC-axes.tikz}}
-\caption{
+\caption[Pictorial depiction of the concept of \emph{model chemistries}.]{
 Pictorial depiction of the concept of \emph{model chemistries} as the
@@
 \section*{The Problem of Solvation or Taming Complexity with Models}
+
+\todo[inline]{Classification of solute-solvent interactions. The focus
+on electrostatics.}
 

### 2016-08-29 Chapters/PaperSummary.tex
Contributions written

@@
 
-% Very ugly
-\epigraph{\textonehalf\,\,research\,\,\,\,\,\,\,\,\,\,\,\,and \textonehalf\,\,\,\,\,\,\,\,Tεχνη
+\epigraph{\textonehalf\,\,research \kern 20pt and \textonehalf \kern 20pt \textgreek{Tέχνη}
 
-          \textonehalf\,\,observation,\,\,\,\,\,\,\,\,\,\,\,\,\textonehalf\,\,\,\,\,\,\,\,Tεχνη
+          \textonehalf\,\,observation, \kern 22pt \textonehalf \kern 20pt \textgreek{Tέχνη}
 
-          \textonehalf\,\,training,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\textonehalf\,\,\,\,\,\,\,\,Tεχνη}{
+          \textonehalf\,\,training, \kern 39pt \textonehalf \kern 20pt \textgreek{Tέχνη}
+          }{
   --- \textsc{Ezra Pound}, \textit{Canto LXXXV}}
@@
 \pcmsolver by providing the implementation and testing of:
-\begin{itemize}
+\begin{enumerate*}[label={\alph*)},font={\color{PMS1797}}]
   \item \acs{MEP} integrals for four-component wave functions,
-  \item the additional Fock matrix contributions,
+  \item the additional Fock matrix contributions and
   \item the additional terms in the response equations.
-\end{itemize}
+\end{enumerate*}
 I performed all the calculations and large part of the data analysis
@@
 \acs{KS}-\acs{DFT} single-point and linear response calculations.
-I, together with coauthor Krzysztof Mozgawa, performed the benchmark quantum
+Together with coauthor Krzysztof Mozgawa, I performed the benchmark quantum
 chemical calculations presented in the paper.
-Finally, I coordinated the editing of all manuscript drafts. I wrote Sections 2.1 and 2.3,
-while Sections 3 and 4 were co-written with the first author, Monica Bugeanu.
+Finally, I coordinated the editing of all manuscript drafts.
+In particular, I wrote the first draft of Sections 2.1 and 2.3.
+The first draft of Sections 3 and 4 was co-written with the first author, Monica Bugeanu.
 I performed most of the data analysis and produced tables and graphs.
@@
 
-I contributed most of the theoretical work for this paper, based on early drafts from coauthors
+I contributed the theoretical work for this paper, based on earlier drafts from coauthors
 Ville Weijo and Hui Cao. In particular, I derived the separation of
@@
 \refstepcounter{dummy}
-\addcontentsline{toc}{section}{Relativistic calculation of EPR and pNMR parameters in solution}
+\addcontentsline{toc}{section}{Relativistic Calculation of EPR and pNMR Parameters in Solution}
 \section*{Paper IV: Relativistic calculation of EPR and pNMR parameters in solution}
+
+My contributions to this paper include prototyping the interface between the
+\pcmsolver library and the \ReSpect quantum chemistry code.
+The interface is maintained in collaboration with coauthor Michal Repisky, who also
+refined the implementation to achieve better computational performance.
+I tested the interface against one-component and four-component results obtained with
+the \LSDALTON and \DIRAC codes, respectively.
+I helped coauthors Michal Repisky, Stanislav Komorovsky and Peter Hrobarik
+with setting up the \acs{PCM} calculations described in the paper.
+Finally, I provided the first draft for Section 2 of the paper and took part in all editing stages.
+The interface to \ReSpect will be released in the next public version of the software
+package, providing \acs{PCM} and \acs{COSMO} capabilities.
 
@@
 \section*{Paper V: Open-ended self-consistent field response theory in solution}
+
+I developed the theoretical framework for the open-ended \acs{SCF} formulation
+of molecular response properties when a quantum/classical polarizable continuum
+Hamiltonian is used.~\autocite{Thorvaldsen2008-sg, Lipparini2010-be}
+I provided its implementation within the \DALTON code, by interfacing the
+\pcmsolver library and the open-ended \acs{SCF} response code of
+\citeauthor{Ringholm2014-gx}~\autocite{Ringholm2014-gx, Friese2015-kb}
+I performed extensive testing of the code by comparing with previously
+published implementations of the \acs{PCM}-\acs{SCF} response functions within
+\DALTON.~\autocite{Cammi2003-qy, Frediani2005-nc, Ferrighi2010-pm}
+Together with coauthors Maarten T.~P.~Beerepoot and Yann Cornaton, I carried out
+the multiphoton absorption calculations presented in the paper. I contributed
+to data collection and data analysis.
+I drafted the initial versions of Sections 2 and 3 of the manuscript and coordinated all
+editing stages with coauthor Maarten T. P. Beerepoot.

### 2016-08-29 Chapters/PaperSummary.tex
Paper summaries done

@@
 \LSDALTON~\autocite{LSDALTON16, Aidas2013-rp}
-\ReSpect~\autocite{ReSpect citation}
+\ReSpect~\autocite{ReSpect-3.5.0, OtherStuff}
 KOALA~\autocite{Hofener2014-ex, Hofener2016-qz}
@@
 \acs{PCM} code and implementation "checkpoints" across altogether different
-problem domains.
+problem domains, Fig.~\ref{fig:algorithm}.
 
-As a by-product of the 4-component implementation, we were able to obtain and visualize \acs{MEP} maps
-from 4-component \acs{SCF} wave functions. These add yet another interpretive tool to the toolbox available
-in the 4-component relativistic regime.
+As a by-product of the four-component implementation, we were able to obtain and visualize \acs{MEP} maps
+from four-component \acs{SCF} wave functions. These add yet another interpretive tool to the toolbox available
+in the four-component relativistic regime.
 The interface to \DIRAC was later released in the 2014 and later versions,
@@
 generator\autocite{Harbrecht2009-no, Harbrecht2011-dk} and wavelet Galerkin
-\acs{BEM} solver with \pcmsolver.
+\acs{BEM} solver\autocite{Harbrecht2004-uo, Harbrecht2006-ug} with \pcmsolver.
 These were used to interface with the new C++ implementation of the wavelet
@@
 homogeneous environments.
-The general integral equation formulation exposed in Chapter~\ref{ch:CSM} is in
-fact transparent with respect to the definition of the Green's function for the
-space portion exterior to the cavity.
-As already shown by \citeauthor{Cances1998-og}, the same mathematical formalism can easily accommodate
-environments characterized by distance-dependent or anisotropic Green's functions, such as those for
-ionic liquids (in the weakly interacting regime described by the linearized Poisson--Boltzmann equation)
-and liquid crystals, respectively.
+The general integral equation formulation is in fact transparent with respect
+to the definition of the Green's function for the space portion exterior to the
+cavity.
+As outlined in Chapter \ref{ch:CSM}, the boundary integral operators in the
+\acs{IEF} equation:
+\begin{equation}
+  \left[ \bi{S}_\mathrm{e}\left(2\pi + \bi{D}^\dagger_\mathrm{i}\right)
+  +
+  \left(2\pi - \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}
+  \right]\sigma =
+  -\left[\left(2\pi-\bi{D}_\mathrm{e}\right)
+  -\bi{S}_\mathrm{e}\bi{S}_\mathrm{i}^{-1}
+  \left(2\pi-\bi{D}_\mathrm{i}\right)
+  \right]\esp,
+  \tag{\ref{eq:full-IEF} from Chapter \ref{ch:CSM}}
+\end{equation}
+can be set up once the Green's functions $\Gi$ and $Ge$ are known.~\autocite{Cances1998-og}
 \citeauthor{Frediani2004-er} showed that a numerical representation of the
@@
 in spherical symmetry.
-In contrast to previously existing work, our implementation offer a more robust
+In contrast to previously existing work, our implementation offers a more robust
 treatment of thin interfaces, with a rather generic functional form for the
@@
 \addcontentsline{toc}{section}{Relativistic Calculation of EPR and pNMR Parameters in Solution}
-\section*{Paper IV: Relativistic calculation of EPR and pNMR parameters in solution}
+\section*{Paper IV: Relativistic Calculation of EPR and pNMR Parameters in Solution}
+
+Paper IV is a step further in our exploration of the interplay between
+relativistic and solvent effects initiated with Paper I.
+Whereas Paper I presented the essential framework for the coupling of
+four-component \acs{SCF} wave functions with a classical polarizable continuum,
+in this paper we explored the calculation of first-order magnetic properties:
+\ac{EPR} and \ac{pNMR} parameters.~\autocite{Repisky2010-ls, Malkin2011-nm,
+Komorovsky2013-xa, Cherry2016-ij}
+The two works are thus complementary since they explore two different classes
+of properties and present implementations in two algorithmically different
+relativistic quantum chemistry codes.
+In the relativistic framework, spin-orbit interactions are included from the
+outset in the variational optimization of the wave function.
+Hence, \acs{EPR} and \acs{pNMR} parameters are formulated as expectation
+values, by virtue of the Hellmann--Feynman theorem.~\autocite{Konishi2009-zb,
+Helgaker2000-tz}
+Moreover, the \ReSpect code can exploit the \emph{Kramers unrestricted} formalism,
+allowing for spin polarization and thus granting facile access to the computation
+of spin-dependent properties.~\autocite{Dyall2007-tu}
+The same modular programming strategy was adopted in crafting an interface between the
+relativistic four-component code \ReSpect\autocite{ReSpect-3.5.0} and \pcmsolver.
 
@@
 
+In recent years, the availability of strong lasers has allowed to
+design and carry out experiments where the high-order response of
+molecular materials can be routinely probed.
+The more intense the light source, the more complicated the interpretation of
+the measured signal.
+Our group has newly developed an open-ended methodology for the computation of
+\acs{SCF} response functions~\autocite{Thorvaldsen2008-sg, Ringholm2014-gx} and
+their residues.~\autocite{Friese2015-kb}
+These developments offer a route towards a synergistic experimental and
+theoretical approach to high-order absorption spectroscopies.
+Paper V grafts a classical polarizable continuum approach to solvation on top
+of the open-ended methodology of \citeauthor{Thorvaldsen2008-sg}.
+Still nowadays, continuum models represent a cost-effective methodology for the
+approximate inclusion of solvent effects, albeit their known limitations with respect
+to specific solute-solvent interactions.
+
 I developed the theoretical framework for the open-ended \acs{SCF} formulation

### 2016-08-29 Chapters/QuantumMechanics.tex
Paper summaries done

@@
 
-\section{A Brief Summary of Coupled Cluster Theory}
+\section{Molecular Quantum Mechanical Hamiltonians}\label{sec:hamiltonians}
+
+Konishi and Paffuti\autocite{Konishi2009-zb}
+Dyall's book\autocite{Dyall2007-tu}
+Reiher's book\autocite{Reiher2014-cp}
+The Pink Book\autocite{Helgaker2000-tz}
+Shavitt's book\autocite{Shavitt2009-mr}
+
+\section{A Brief Summary of Coupled Cluster Theory}\label{sec:coupled-cluster}
 

### 2016-08-29 Chapters/ContributionsSummary.tex
Change name of one chapter
Fix a typo

--- 2016-08-29 Chapters/PaperSummary.tex
@@
 %************************************************
-\chapter{Summary of the publications}\label{ch:publications-summary}
+\chapter{Summary of Contributions}\label{ch:contributions-summary}
 
@@
 \end{equation}
-can be set up once the Green's functions $\Gi$ and $Ge$ are known.~\autocite{Cances1998-og}
+can be set up once the Green's functions $\Gi$ and $\Ge$ are known.~\autocite{Cances1998-og}
 \citeauthor{Frediani2004-er} showed that a numerical representation of the

### 2016-08-29 Chapters/ContinuumSolvationModels.tex
Contributions written
Change name of one chapter
Correction factor for COSMO added

@@
 % mainfile: ../RobertoDiRemigioPhDThesis.tex
-\chapter{Continuum solvation models}\label{ch:CSM}
+\chapter{Continuum Solvation Models}\label{ch:CSM}
+
+\epigraph{
+\textitalian{
+L'acqua è la forza che ti tempra,
+
+nell'acqua ti ritrovi e ti rinnovi:
+
+noi ti pensiamo come un'alga, un ciottolo,
+
+come un'equorea creatura
+
+che la salsedine non intacca
+
+ma torna al lito più pura.
+}
+}{
+    --- \textsc{E. Montale}, \textit{Falsetto}}
 
@@
 \begin{equation}
-  \bi{S}\sigma = - \esp
+  \bi{S}\sigma = - f(\diel)\esp
 \end{equation}
+where the factor $f(\diel) = \frac{\diel-1}{\diel+x},\, 0\leq x \leq 1$
+has been introduced to account for the error introduced by modelling the
+dielectric as a perfect conductor.
 

### 2016-08-29 Chapters/ContributionsSummary.tex
Save the day's work

@@
 All publications included in this thesis have involved some programming effort.
-I will also describe the principles we have striven to follow in developing our
-\acs{PCM} software library and its management philosophy.
+In Section \ref{sec:software} I will describe the principles we have
+striven to follow in developing our \acs{PCM} software library and its
+management philosophy.
 I will also give some details on the interfaces we have developed to different
 quantum chemistry codes.
-Finally, for each paper, I will give a description of my contributions to the
-work leading to its publication.
-\todo[inline]{Write something along these lines: "My coauthors have all read and approved."}
+A shorter version of the material in Section \ref{sec:software} has been
+accepted for publication in the proceedings of the \emph{Producing High
+Performance and Sustainable Software for Molecular Simulation} workshop
+held at the 2015 Supercomputing Conference.~\autocite{SC15}
+Sections \ref{sec:relapcm}--\ref{sec:pcmopenrsp} contain very short
+summaries of the papers and a list of my contributions.
+My coauthors have all read and approved the description of contributions.
 
@@
 
-\refstepcounter{dummy}
-\addcontentsline{toc}{section}{Software}
-\section*{Software}
-
-\todo[inline]{Have a closer look at this. I think it needs to be updated for the discussion
-on open-source software that went down on JPC Lett. some time ago.
-Also, acronyms should be properly set up.}
-
-A shorter version of this discussion has been accepted for publication in the
-proceedings of the \emph{Producing High Performance and Sustainable Software
-for Molecular Simulation} workshop held at the 2015 Supercomputing Conference.
-
-We have created an open-source, standalone library to provide \acs{PCM} functionality across
-altogether different quantum chemical program packages.
-The library serves as our development platform for the \acs{PCM}.
-Moreover, the interfaces developed to different quantum chemistry codes allow the exploration
-of new methodologies, significantly cutting down development times.
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=.8\textwidth]{notpretty.png}
-\end{figure}
-
-The codes used in the papers are all interfaced to the \pcmsolver library,
-of which I am the main developer and maintainer.
-
-\begin{figure}[!htb]
-\centering
-\scalebox{0.8}{\input{gfx/algorithm.tikz}}
-\caption[Modular approach to programming a \acs{PCM} functionality into an existing \acs{SCF} code.]{
-  Schematic view of the implemented SCF algorithm. Computations/data in
-  \textcolor{Blue}{blue} are on the \pcmsolver side, in
-  \textcolor{Green}{green} on the \DIRAC side.
-  }
-\label{fig:algorithm}
-\end{figure}
+\section{Software}\label{sec:software}
 
@@
 by actively extending the codebase with new functionality.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The codes used in the papers are all interfaced to the \pcmsolver library,
+of which I am the main developer and maintainer.
+
+Boost\autocite{boost}
+Eigen\autocite{eigen}
+libgetkw\autocite{libgetkw}
+libtaylor\autocite{libtaylor}
+
+We have created an open-source, standalone library to provide \acs{PCM}
+functionality across altogether different quantum chemical program
+packages.
+The library serves as our development platform for the \acs{PCM}.
+Moreover, the interfaces developed to different quantum chemistry codes
+allow the exploration of new methodologies, significantly cutting down
+development times.
+
+\begin{figure}[!htb]
+\centering
+\scalebox{0.8}{\input{gfx/algorithm.tikz}}
+\caption[Modular approach to programming a \acs{PCM} functionality into an existing \acs{SCF} code.]{
+  Schematic view of the implemented SCF algorithm. Computations/data in
+  \textcolor{Blue}{blue} are on the \pcmsolver side, in
+  \textcolor{Green}{green} on the \DIRAC side.
+  }
+\label{fig:algorithm}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\begin{figure}[!htb]
+\begin{figure}[tb]
   \centering
-  \includegraphics[width=.5\textwidth]{pcmsolver-scheme}
+  \scalebox{0.8}{\input{gfx/pcmsolver-scheme.tikz}}
+  \caption[Schematic view of the relationship between a host quantum chemistry
+  program and the \pcmsolver library.]{
+  Schematic view of the relationship between a host quantum chemistry
+  program and the \pcmsolver library.
+  }
 \end{figure}
@@
 simulations.~\autocite{Corni2015-pe}
+
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=.8\textwidth]{pcmsolver-internals.png}
+  \caption[Internal structure of \pcmsolver.]
+  {
+  Internal structure of \pcmsolver. In the grey boxes the external
+  projects the library depends on.
+  }
+\end{figure}
 
@@
 
-\refstepcounter{dummy}
-\addcontentsline{toc}{section}{Continuum solvation in the relativistic regime}
-\section*{Paper I. Continuum solvation in the relativistic regime}
+\section{Continuum solvation in the relativistic regime}\label{sec:relapcm}
 
@@
 
-\refstepcounter{dummy}
-\addcontentsline{toc}{section}{The wavelet Galerkin boundary element method for PCM}
-\section*{Paper II. The wavelet Galerkin boundary element method for PCM}
+\section{The Wavelet Galerkin Boundary Element Method for PCM}\label{sec:wemlin}
 
@@
 
-\refstepcounter{dummy}
-\addcontentsline{toc}{section}{Non homogeneous environments}
-\section*{Paper III: Non homogeneous environments}
+\section{Non homogeneous environments}\label{sec:spherical}
 
@@
 
-\refstepcounter{dummy}
-\addcontentsline{toc}{section}{Relativistic Calculation of EPR and pNMR Parameters in Solution}
-\section*{Paper IV: Relativistic Calculation of EPR and pNMR Parameters in Solution}
+\section{Relativistic Calculation of EPR and pNMR Parameters in
+Solution}\label{sec:pcmepr}
 
@@
 
-\refstepcounter{dummy}
-\addcontentsline{toc}{section}{Open-ended self-consistent field response theory in solution}
-\section*{Paper V: Open-ended self-consistent field response theory in solution}
+\section{Open-ended self-consistent field response theory in
+solution}\label{sec:pcmopenrsp}
 

### 2016-08-29 FrontBackmatter/Introduction.tex
Contributions written
Save the day's work

@@
 \epigraph{
-[...] og fordi jeg alltid har hatt en dragning mot det skjulte og hemmelige.
+[...] \textnorsk{og fordi jeg alltid har hatt en dragning mot det skjulte og hemmelige.}
 }{
-    --- \textsc{K. O. Knausgård}, \textit{Om Høsten}}
+--- \textnorsk{\textsc{K. O. Knausgård}, \textit{Om Høsten}}}
 
@@
 \missingfigure{Add picture with multiscale models}
-\caption{From quantum mechanical models to multiscale models.}
+\caption[From quantum mechanical to multiscale models]{
+From quantum mechanical to multiscale models.
+The images used in the scheme are reproduced courtesy of Dr.~Stefano
+Caprasecca (MoLEcoLab, Università di Pisa).}
 \label{fig:qm-to-multiscale}

### 2016-08-29 FrontBackmatter/Publications.tex
Work on the summary of papers
Save the day's work

@@
   \\
+  \textsc{
   Four-Component Relativistic Calculations in Solution with the
@@
   \ce{Po})
+  }
   \\
@@
   \\
+  \textsc{
   Wavelet Formulation of the Polarizable Continuum Model. II. Use of
   Piecewise Bilinear Boundary Elements
+  }
   \\
@@
   \\
+  \textsc{
   A Polarizable Continuum Model for Molecules at Spherical
   Diffuse Interfaces
+  }
   \\
@@
   \\
+  \textsc{
   Four-Component Relativistic Density Functional Theory with the
@@
   and Paramagnetic NMR Shifts
+  }
   \\
-  submitted for publication to \textit{Mol. Phys.}
+  Accepted for publication in \textit{Mol. Phys.}
 \label{pcmepr}
@@
   \\
+  \textsc{
   Open-ended formulation of self-consistent field response theory with
   the polarizable continuum model for solvation
+  }
   \\

### 2016-08-30 Chapters/ContributionsSummary.tex
Switch to B5 paper
Update Makefile

@@
 maintainability and extensibility of the code play an equally important role in
-managing this growing complexity.~\autocite{Hatton1994-us, Hatton1997-ml,
-Hatton1997-bb, Ioannidis2005-aj, Merali2010-uy, Prinz2011-kz, Wilson2014-vh}
+managing software complexity\autocite{Wilson2014-vh} and ensuring
+scientific reproducibility.~\autocite{Hatton1994-us, Hatton1997-ml,
+Hatton1997-bb, Ioannidis2005-aj, Merali2010-uy, Prinz2011-kz}
 
-The use of a \emph{modular programming paradigm} has been recognized as
-beneficial in many other scientific computing contexts. New features are
-isolated into libraries that can be accessed by host programs through a
-well-defined \ac{API}.
+The idea of a \emph{modular programming paradigm} can be traced back to
+the dawn of computer science. \citeauthor{Dijkstra1968-zp}
+successfully used it in the development of the THE operating
+system~\autocite{Dijkstra1968-zp} already in 1968, while
+\citeauthor{Parnas1972-im} advocated it as a best practice already in
+1972.~\autocite{Parnas1972-im}
+Modularity is nothing more than a \emph{divide et impera} strategy
+applied to programming and it has been recognized as beneficial in many
+other scientific computing contexts.
+New features are isolated into libraries that can be accessed by host
+programs through a well-defined \ac{API}.
 Computational tasks are thus implemented into separate, independent and
-interchangeable modules. These are \emph{developed} and \emph{tested}
+interchangeable modules. These are \emph{developed}, \emph{tested},
+\emph{packaged} and \emph{distributed}
 independently of the particular host program in which they will be
-used.~\cite{Dijkstra1968-zp, Parnas1972-im}
+used.
 A well-defined \ac{API} clearly delimits the boundaries of the functionality
@@
 step towards ensuring reproducibility of results from scientific
-simulation software \cite{Gezelter2015-gz, Krylov2015-fs, Jacob2016-oq}.
-Open-source development can fully leverage the benefits
-of widespread, cloud-based, free, code development services, such as
-hosted \acp{DVCS}, \ac{CI}, code
-coverage, static and dynamic code analyses, nightly regression testing,
-public issue tracking, code review and so forth:
-adoption of a modern code development workflow is easily within reach.
+simulation software.~\autocite{Gezelter2015-gz, Krylov2015-fs,
+Jacob2016-oq}
+Open-source development can fully leverage the benefits of widespread,
+cloud-based, free, code development services, such as hosted \acp{DVCS},
+continuous integration (CI), code coverage, static and dynamic code
+analyses, nightly regression testing, public issue tracking, code review
+and so forth: adoption of a modern code development workflow is easily
+within reach.
 
 It is, of course, true that the above mentioned services are not
-exclusive prerogatives of open-source projects.
-However, an open review process of scientific software can often help
-to establish
+exclusive prerogatives of open-source projects. However, an open review
+process of scientific software can often help to establish
 reproducibility, extensibility and sustainability of the software
 ecosystem.
-Open-source software, modular development of new functionalities
-and full-fledged exploitation of \acsp{DVCS} ensure a much larger scientific
-impact. Third-parties can easily contribute to
-the project: by improving the documentation, by reporting bugs or
-by actively extending the codebase with new functionality.
+Open-source software, modular development of new functionalities and
+full-fledged exploitation of \acsp{DVCS} ensure a much larger scientific
+impact. Third-parties can easily contribute to the project: by improving
+the documentation, by reporting bugs or by actively extending the
+codebase with new functionality.
+
+The \acs{PCM} is an ideal candidate for the creation of a solvation
+\acs{API}.
+Consider for example the schematic representation of a
+\acs{PCM}-\acs{SCF} algorithm given in Figure
+Its input from and output to \emph{any} quantum chemistry code is
+limited and well-defined.
+This provides a natural \acs{API} design: the \acs{API} functions can be compared
+\emph{vis-à-vis} with the working equations derived for the different
+quantum chemical methods.
+
+\begin{figure}[!htb]
+\centering
+\scalebox{0.8}{\input{gfx/algorithm.tikz}}
+\caption[Modular approach to programming a \acs{PCM} functionality into an existing \acs{SCF} code.]{
+  Schematic view of the implemented SCF algorithm. Computations/data in
+  \textcolor{PMS2229}{blue} are on the \pcmsolver side, in
+  \textcolor{Green}{green} on the \DIRAC side.
+  }
+\label{fig:algorithm}
+\end{figure}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 We have created an open-source, standalone library to provide \acs{PCM}
@@
 development times.
-
-\begin{figure}[!htb]
-\centering
-\scalebox{0.8}{\input{gfx/algorithm.tikz}}
-\caption[Modular approach to programming a \acs{PCM} functionality into an existing \acs{SCF} code.]{
-  Schematic view of the implemented SCF algorithm. Computations/data in
-  \textcolor{Blue}{blue} are on the \pcmsolver side, in
-  \textcolor{Green}{green} on the \DIRAC side.
-  }
-\label{fig:algorithm}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The description of chemical phenomena in complex environments, such as
-solutions or proteins, is as challenging as it is important
-for an accurate prediction of a wide range of properties.~\autocite{Reichardt2010-le}
-  Quantum chemistry is nowadays a valuable
-tool: its simulations can provide significant insight into complex
-chemical phenomena.~\autocite{Nobel1998, Nobel2013}.
-The major challenge lies in providing a faithful, realistic, yet cost-effective
-description of the environment of the molecule:
-a molecule in solution can be represented by a quantum mechanical model
-comprising a great number of atoms, thousands or more. Such a simulation would,
-however, be impossible to carry out with the current state-of-the-art
-computational resources.
-Models must be devised to overcome the dimensionality ``disease'' and
-the idea of introducing different approximation levels for different
-parts of the system under study is particularly appealing, as it helps
-our chemical understanding. This is the idea at the basis of
-\emph{multiscale} models.
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[width=.45\textwidth]{multiscale.jpg}
-\end{figure}
-
-\emph{Continuum} (or \emph{implicit}) models, deal explicitly
-only with the degrees of freedom associated with the solute, while
-replacing the solvent with a structureless continuum characterized by
-its bulk properties \cite{Onsager1936-wf}. Among these models, the \acl{PCM} (\acs{PCM})
-is the most widespread and successful \cite{Tomasi2005-vm, Mennucci2008-xl}.
-
-The \acs{PCM} is also an ideal candidate for the creation of a solvation \acs{API}.
-Its input from and output to \emph{any} quantum chemistry code is
-limited and well-defined.
-This provides a natural \acs{API} design: the \acs{API} functions can be compared
-\emph{vis-à-vis} with the working equations derived for the different
-quantum chemical methods.
 

### 2016-08-30 Chapters/ContributionsSummary.tex
Contributions chapter done

@@
 maintainability and extensibility of the code play an equally important role in
-managing software complexity\autocite{Wilson2014-vh} and ensuring
+managing software complexity\autocite{Wilson2014-vh, ssi, nsf-s2i2,
+netherlands-escience, coderefinery, molssi} and ensuring
 scientific reproducibility.~\autocite{Hatton1994-us, Hatton1997-ml,
@@
 successfully used it in the development of the THE operating
-system~\autocite{Dijkstra1968-zp} already in 1968, while
+system in 1968,~\autocite{Dijkstra1968-zp} while
 \citeauthor{Parnas1972-im} advocated it as a best practice already in
@@
 programs through a well-defined \ac{API}.
-Computational tasks are thus implemented into separate, independent and
-interchangeable modules. These are \emph{developed}, \emph{tested},
-\emph{packaged} and \emph{distributed}
-independently of the particular host program in which they will be
-used.
+
+In this paradigm, computational tasks are implemented into separate,
+independent and interchangeable modules.
+Modules are to be thought as computational black boxes \emph{developed},
+\emph{tested}, \emph{packaged} and \emph{distributed} independently of
+the particular host program in which they will be used.
 A well-defined \ac{API} clearly delimits the boundaries of the functionality
@@
 Open-source development can fully leverage the benefits of widespread,
-cloud-based, free, code development services, such as hosted \acp{DVCS},
-continuous integration (CI), code coverage, static and dynamic code
-analyses, nightly regression testing, public issue tracking, code review
-and so forth: adoption of a modern code development workflow is easily
-within reach.
+cloud-based, free, code development services, such as hosted
+\acp{DVCS},~\autocite{github, gitlab} continuous
+integration,~\autocite{travis-ci, appveyor-ci} code
+coverage,~\autocite{coveralls} static and dynamic code
+analyses,~\autocite{coverity-scan} nightly regression testing, public
+issue tracking, code review and so forth: adoption of a modern code
+development workflow is easily within reach.
 
@@
 The \acs{PCM} is an ideal candidate for the creation of a solvation
-\acs{API}.
-Consider for example the schematic representation of a
-\acs{PCM}-\acs{SCF} algorithm given in Figure
-Its input from and output to \emph{any} quantum chemistry code is
-limited and well-defined.
-This provides a natural \acs{API} design: the \acs{API} functions can be compared
-\emph{vis-à-vis} with the working equations derived for the different
-quantum chemical methods.
+\acs{API}. Consider for example the schematic representation of a
+\acs{PCM}-\acs{SCF} algorithm given in Figure \ref{fig:algorithm}.
+The input to and from the \acs{PCM} library is limited and well-defined,
+as are the components that host quantum chemistry codes need to
+implement.
+This provides a natural \acs{API} design: the \acs{API} functions can be
+compared \emph{vis-à-vis} with the working equations derived for the
+different quantum chemical methods.
 
-\begin{figure}[!htb]
+\begin{figure}[tb]
 \centering
-\scalebox{0.8}{\input{gfx/algorithm.tikz}}
+\scalebox{0.7}{\input{gfx/algorithm.tikz}}
 \caption[Modular approach to programming a \acs{PCM} functionality into an existing \acs{SCF} code.]{
-  Schematic view of the implemented SCF algorithm. Computations/data in
-  \textcolor{PMS2229}{blue} are on the \pcmsolver side, in
-  \textcolor{Green}{green} on the \DIRAC side.
+Schematic view of a \acs{PCM}-\acs{SCF} algorithm. Computations/data in
+\textcolor{PMS2229}{blue} are implemented by the \acs{PCM} \acs{API}.
+Computations/data in \textcolor{Green}{green} are implemented by the
+host quantum chemistry code.
   }
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The codes used in the papers are all interfaced to the \pcmsolver library,
-of which I am the main developer and maintainer.
-
-Boost\autocite{boost}
-Eigen\autocite{eigen}
-libgetkw\autocite{libgetkw}
-libtaylor\autocite{libtaylor}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-We have created an open-source, standalone library to provide \acs{PCM}
-functionality across altogether different quantum chemical program
-packages.
-The library serves as our development platform for the \acs{PCM}.
-Moreover, the interfaces developed to different quantum chemistry codes
-allow the exploration of new methodologies, significantly cutting down
-development times.
-
-We will elucidate the design principles adopted in the development of our
-\pcmsolver library: an \emph{open-source} \acs{API} for the inclusion of the
-PCM in \emph{any} quantum chemistry code \cite{PCMSolver}.  Use of our
-\acs{API} \emph{significantly} limits coding effort on the side of the host:
-continuum solvation at the \acs{SCF}
-level of theory can be implemented in the host program almost out-of-the-box.  The module is
-agnostic of the host code internals and is tested separately from it.
-
-We will emphasize the role of the open-source licensing model and the
-importance of the modern workflows adopted in our project.  The adoption
-of \git as \acs{DVCS}, \cmake for cross-platform builds and automatic
-documentation deployment on \readthedocs simplify extensibility of the
-module and promote third-party contributions to the code base, while
-continuous integration and nightly testing offer an invaluable level of
-confidence in the code.
-
-\begin{figure}[tb]
-  \centering
-  \scalebox{0.8}{\input{gfx/pcmsolver-scheme.tikz}}
-  \caption[Schematic view of the relationship between a host quantum chemistry
-  program and the \pcmsolver library.]{
-  Schematic view of the relationship between a host quantum chemistry
-  program and the \pcmsolver library.
-  }
-\end{figure}
-
-\pcmsolver currently offers the basic PCM functionality already
-available in many widespread software packages, such as traditional
-collocation solvers for isotropic media.~\autocite{Tomasi2005-vm}
-Some unique functionalities are the wavelet Galerkin solvers on smooth
-molecular surfaces~\autocite{Weijo2010-hy, Bugeanu2015-tp}, the Green’s
-functions for spherical diffuse interfaces~\autocite{DiRemigio2016-nn} and
-the delayed solvers for real-time \acl{TDSCF}
-simulations.~\autocite{Corni2015-pe}
-
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=.8\textwidth]{pcmsolver-internals.png}
-  \caption[Internal structure of \pcmsolver.]
-  {
-  Internal structure of \pcmsolver. In the grey boxes the external
-  projects the library depends on.
-  }
-\end{figure}
+The \pcmsolver library was developed to fulfill these requirements and
+offer the community an easy-to-use implementation of the \acs{PCM}.
+\pcmsolver is written in C++ with Fortran, C and Python components.
+Architectural and conceptual organization of the code is provided by the
+C++ layer. To ensure the largest possible impact, the code is C++03 ISO standard
+compliant, with external dependencies kept to a minimum.
+Tuples, container algorithms and functional programming tools are not
+part of the standard we chose to comply to.
+Parts of the Boost C++ libraries are used to alleviate these
+shortcomings.~\autocite{boost}
+Linear algebra operations are managed through the Eigen C++ template
+library.~\autocite{eigen}
+The Python layer provides input reading and parsing facilities,
+as implemented by the GetKw library .~\autocite{libgetkw}
+Cavity construction and discretization is handled either by the Fortan
+or by the C layer, depending on which computational \emph{backend} is
+needed.
+Evaluation of the Green's functions, needed to set up the representation
+of the boundary integral operators, leverages \ac{AD}. We use the Taylor library, a template C++
+implementation of \acs{AD}.~\autocite{Bartholomew-Biggs2000-db, libtaylor}
+The adoption of \git as \acs{DVCS}, \cmake for cross-platform builds and
+automatic documentation deployment on \readthedocs simplify
+extensibility of the module and promote third-party contributions to the
+code base. Continuous integration and nightly testing offer an
+invaluable level of confidence in the code.
 
@@
 modules, communicating by means of composition at the outermost
-level of the design.
-This allows us to mix languages internally: we currently have Fortran
-and C modules alongside the main C++ structure.
+level of design.
+This is an indispensable feature to achieve frictionless language
+mixing.
 Library internal classes balance the dynamic polymorphism offered by
@@
 
-References for the various codes we have created interfaces to.
-\psicode~\autocite{Turney2012-de}
-\DIRAC~\autocite{DIRAC15}
-\DALTON~\autocite{DALTON16, Aidas2013-rp}
-\LSDALTON~\autocite{LSDALTON16, Aidas2013-rp}
-\ReSpect~\autocite{ReSpect-3.5.0, OtherStuff}
-KOALA~\autocite{Hofener2014-ex, Hofener2016-qz}
+The library is released under the terms of the
+version 3 \ac{LGPL}, a standard open-source license.~\autocite{LGPLv3}
+We strongly believe that open-source software is the key to a larger
+impact for scientific software and a standard license means lower
+thresholds to adoption and third-party contributions.~\autocite{LGPLv3}
+The library serves as our development platform for the \acs{PCM}.
+The interfaces to different quantum chemistry codes allow us to explore
+new methodologies, significantly cutting down development times.
+Use of our \acs{API} \emph{significantly} limits coding effort on the
+side of the host: continuum solvation at the \acs{SCF} level of theory
+can be implemented in the host program almost out-of-the-box.
+
+The library was designed to allow host-\acs{API} communication through
+\emph{pure} functions, \ie{} free from side-effects.
+Such an implementation hinges on the theoretical realization that the
+\acs{PCM} layer is independent of the \acs{AO} or \acs{MO} spaces
+defined in the quantum chemical layer.
+Passing and modifying large data structures, such as the Fock and
+density matrices, can be completely avoided.
+As schematically depicted in Figure \ref{fig:algorithm},
+computation of the \acs{ASC} only depends on
+the electrostatic potential sampled at the cavity boundary:
+\begin{equation}
+ \esp(\vect{s}) =
+ \sum_A \frac{Z_A}{|\vect{R}_A-\vect{s}|} +  \sum_{\kappa\lambda} D_{\lambda\kappa}
+ \left[\int\diff\vect{r}\frac{-\Omega_{\kappa\lambda}(\vect{r})}{|\vect{r}-\vect{s}|}\right],
+\end{equation}
+which is the contraction of charge attraction integrals with a, possibly
+perturbed, density matrix, a task handled by the host program.
+Formation of the solvent contribution to the Fock matrix requires the
+computation of the scalar product
+$\scalprod{\sigma(\vect{s})}{\esp_{\kappa\lambda}(\vect{s})}$
+between the \acs{ASC} and the charge attraction integral matrix, again
+a task easily accomplished by the host program.
+As shown in Figure \ref{fig:pcmsolver-scheme}, data transfer between
+\pcmsolver and the host is limited to the communication of
+$\esp(\vect{s})$ and $\sigma(\vect{s})$ and can be implemented
+\emph{without} storing any quantity to disk, avoiding possibly costly
+I/O operations.
+
+\begin{figure}[tb]
+  \centering
+  \scalebox{0.7}{\input{gfx/pcmsolver-scheme.tikz}}
+  \caption[Schematic view of the relationship between a host quantum chemistry
+  program and the \pcmsolver library.]{
+  Schematic view of the relationship between a host quantum chemistry
+  program and the \pcmsolver library.
+  }
+  \label{fig:pcmsolver-scheme}
+\end{figure}
+
+The \acs{API} is implemented in ISO C99 as a \emph{context-aware} set of
+functions.~\autocite{library-ronacher, context-api-example}
+Upon initialization of the library, a \emph{context} object containing
+the state of the \acs{PCM} \acs{API} is constructed and a handle to it
+(a pointer in C++ parlance) is returned to the host program. Further
+transactions between \pcmsolver and the host program proceed through
+manipulations of the context object.
+\pcmsolver introduces the concept of surface functions:
+labelled instances of any quantity defined on the cavity surface.
+Bookkeeping of a possibly arbitrary number of such quantities is managed
+through a map storing labels and pointers to the contents of the
+functions.
+Listings \ref{lst:C-host} and \ref{lst:Fortran-host} show how calls to
+\pcmsolver \acs{API} functions might look like in an actual C or Fortran
+program, respectively.
+
+\lstinputlisting[language=C,
+                 caption={Relevant calls to \pcmsolver in a C host program.
+                 Ellipsis in the functions arguments list stand for omitted parameters.
+                 For the full, functional example visit
+                 \url{http://pcmsolver.readthedocs.io/en/latest/users/C-example.html}},
+                 label={lst:C-host}
+                ]{snippets/C_host.c}
+
+A context-aware \acs{API} offers a simple strategy for parallelization.
+The host program could spawn contexts on a per-process basis,
+limiting ownership and access to a context just to its parent process.
+Race conditions are trivially avoided in such a scenario.
+
+\lstinputlisting[language=Fortran,
+                 caption={Relevant calls to \pcmsolver in a Fortran host program.
+                 Notice the use of the standard Fortan ISO C bindings.
+                 Ellipsis in the functions arguments list stand for omitted parameters.
+                 For the full, functional example visit
+                 \url{http://pcmsolver.readthedocs.io/en/latest/users/fortran-example.html}},
+                 label={lst:Fortran-host}
+                 ]{snippets/Fortran_host.F90}
+
+\pcmsolver currently offers an implementation of the traditional collocation
+solvers for isotropic media,~\autocite{Tomasi2005-vm}
+together with some unique functionalities: the wavelet Galerkin solvers on smooth
+molecular surfaces,~\autocite{Weijo2010-hy, Bugeanu2015-tp} the Green’s
+functions for spherical diffuse interfaces~\autocite{DiRemigio2016-nn} and
+the delayed solvers for real-time \acl{TDSCF}
+simulations.~\autocite{Corni2015-pe}
+
+Programming effort in this thesis has not only been directed at the
+creation of the \pcmsolver library, but also at interfacing it with a
+number of quantum chemistry codes.
+The first interface was implemented and released within the \DIRAC code
+and is described in Paper I.~\autocite{DIRAC15, DiRemigio2015-ou}.
+Interfaces to the \LSDALTON\autocite{LSDALTON16, Aidas2013-rp} and
+\DALTON\autocite{LSDALTON16, Aidas2013-rp} codes have followed, as
+described in Paper II and Paper V, respectively.~\autocite{Bugeanu2015-tp, pcm-openrsp}
+Details about the interface to \ReSpect~\autocite{ReSpect-3.5.0} can be
+found in Paper IV.~\autocite{pcm-respect}
+Coupling with \psicode~\autocite{Turney2012-de} has allowed us to implement the
+\acs{PCM}-\acs{CC} developments described in this thesis.
+Finally, development of an interface with the KOALA code is underway.~\autocite{Hofener2014-ex, Hofener2016-qz}
 

### 2016-08-30 snippets/C_host.c

@@ -1,0 +1,16 @@
+// Include API functions and data structures definitions
+#include "pcmsolver.h"
+#include "PCMInput.h"
+
+// Initialize API context
+pcmsolver_context_t * pcm_context = pcmsolver_new(...);
+// Register surface function from host program to API context
+pcmsolver_set_surface_function(pcm_context, ...);
+// Compute the ASC from a MEP surface function
+pcmsolver_compute_asc(pcm_context, ...);
+// Retrieve surface function to host from API context
+pcmsolver_get_surface_function(pcm_context, ...);
+// Compute polarization energy from surface functions
+double energy = pcmsolver_compute_polarization_energy(pcm_context, ...);
+// Clean up API context
+pcmsolver_delete(pcm_context);

### 2016-08-30 snippets/Fortran_host.F90

@@ -1,0 +1,20 @@
+! Use standard ISO C bindings
+use, intrinsic :: iso_c_binding
+! Use API functions and data structures
+use pcmsolver
+! Declare handle to API context as C pointer
+type(c_ptr) :: pcm_context
+! Declare polarization energy as a C-interoperable real number
+real(c_double) :: energy
+! Initialize API context
+pcm_context = pcmsolver_new(...)
+! Register surface function from host program to API context
+call pcmsolver_set_surface_function(pcm_context, ...)
+! Compute the ASC from a MEP surface function
+call pcmsolver_compute_asc(pcm_context, ...)
+! Retrieve surface function to host from API context
+call pcmsolver_get_surface_function(pcm_context, ...)
+! Compute polarization energy from surface functions
+energy = pcmsolver_compute_polarization_energy(pcm_context, ...)
+! Clean up API context
+call pcmsolver_delete(pcm_context)

### 2016-08-31 Chapters/ContinuumSolvationModels.tex
Contributions chapter done
Work on CSM chapter

@@
 we have worked upon in the thesis.
-Section \ref{sec:IEF} will present some of the mathematical details of the
-formulation of the model. I will introduce the concept of \emph{weak}
-formulations of \acp{PDE} in Section \ref{sec:weak} and eventually show their
-use in the formulation of quantum/classical polarizable Hamiltonians in
-Section \ref{sec:variational}.
+Appendix \ref{app:mathematical-results} contains much of the
+mathematical details omitted from the exposition in this Chapter.
+Section \ref{sec:IEF} will present the formulation of the model.
+I will show how a variational approach can facilitate the formulation of
+quantum/classical polarizable Hamiltonians in Section
+\ref{sec:variational}.
 Section \ref{sec:BEM} will be devoted to the numerical strategies used
@@
 
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=.5\textwidth]{IEF}
-  \caption[The physical setting of the polarizable continuum model.]{
-  The physical setting of the \ac{PCM}. The molecular solute,
-  represented by its, possibly quantum mechanical, charge density
-  $\rho_\mathrm{i}$ is enclosed in a \emph{cavity} $\Omegai$.
-  The boundary of the cavity $\Gamma\equiv\partial\Omegai$ is a
-  continuously differentiable 2-manifold.
-  We assume the material properties of the cavity to be those of vacuum,
-  hence characterized by the Green's function
-  $\Gi=\frac{1}{|\vect{r}-\vect{r}^\prime|}$.
-  The cavity is carved out of an infinite, structureless, continuum
-  characterized by its Green's function $\Ge$ and fully covering the
-  external subdomain $\Omegae$.
-  }
-  \label{fig:IEF}
-\end{figure}
-
 This is a well-known problem in electrostatics~\autocite{Jackson1999, Vanderlinde2005-gf}
@@
 literature on \acp{PDE}.~\autocite{Hackbusch1995-uq, Sauter2011-an}
-The problem is commonly posed as follows. Given the partition of Euclidean
-space $\mathbb{R}^3$ outlined above, find the solution $u(\vect{r})$
-to the following set:
+Given the partition of Euclidean space $\mathbb{R}^3$ outlined above,
+find the solution $u(\vect{r})$ to the following set:
 \begin{subequations}\label{eq:transmission}
@@
 
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=.5\textwidth]{IEF}
+  \caption[The physical setting of the polarizable continuum model.]{
+  The physical setting of the \ac{PCM}. The molecular solute,
+  represented by its, possibly quantum mechanical, charge density
+  $\rho_\mathrm{i}$ is enclosed in a \emph{cavity} $\Omegai$.
+  The boundary of the cavity $\Gamma\equiv\partial\Omegai$ is a
+  continuously differentiable 2-manifold.
+  We assume the material properties of the cavity to be those of vacuum,
+  hence characterized by the Green's function
+  $\Gi=\frac{1}{|\vect{r}-\vect{r}^\prime|}$.
+  The cavity is carved out of an infinite, structureless, continuum
+  characterized by its Green's function $\Ge$ and fully covering the
+  external subdomain $\Omegae$.
+  }
+  \label{fig:IEF}
+\end{figure}
+
 The first step is to introduce the relevant boundary integral operators for the
-abovementioned transmission problem.
+above mentioned transmission problem.
 They are three of the four components of the Calder\'on projector.
 For any function $v$ in the suitable function space and for $\vect{s}, \vect{s}^\prime \in \Gamma$:
-\begin{equation}
-\begin{aligned}
+\begin{subequations}
+\begin{align}
   (\bi{S}_\star v)(\vect{s}) &= \int_\Gamma
-  G_\star(\vect{s}, \vect{s}^\prime)v(\vect{s}^\prime)\diff \vect{s}^\prime \\
+  G_\star(\vect{s}, \vect{s}^\prime)v(\vect{s}^\prime)\diff
+  \vect{s}^\prime \label{eq:S} \\
   (\bi{D}_\star v)(\vect{s}) &= \int_\Gamma \partial_{L_\star,\vect{s}}
-  G_\star(\vect{s}, \vect{s}^\prime)v(\vect{s}^\prime)\diff \vect{s}^\prime \\
+  G_\star(\vect{s}, \vect{s}^\prime)v(\vect{s}^\prime)\diff
+  \vect{s}^\prime \label{eq:D} \\
   (\bi{D}^\dagger_\star v)(\vect{s}) &= \int_\Gamma
 \partial_{L_\star,\vect{s}^\prime}
-  G_\star(\vect{s}, \vect{s}^\prime)v(\vect{s}^\prime)\diff \vect{s}^\prime \\
-\end{aligned}
-\end{equation}
+  G_\star(\vect{s}, \vect{s}^\prime)v(\vect{s}^\prime)\diff
+  \vect{s}^\prime \label{eq:D-dagger}
+\end{align}
+\end{subequations}
 where $\star$ is a placeholder for the $\mathrm{i}$ or $\mathrm{e}$ subscript.
@@
 The reaction potential describes the polarization in the medium.
-Moreover, it admits a integral representation as a single-layer
+Moreover, it admits an integral representation as a single-layer
 potential:
@@
 
-\section{Detour and Teaser: Weak Formulation of Partial Differential Equations}\label{sec:weak}
-
-From the strong to the weak formulation of \acp{PDE}.\autocite{Ern2004-oo}
-
-\begin{itemize}
-  \item Poisson's equation with conductor boundary conditions
-        % This is COSMO!
-        % This equation makes sense iff \Psi \in \mathcal{C}^2_0(C) (at least)
-        % \Psi = 0 \quad \forall \vect{r}\in\partial C
-        \begin{equation*}
-            \nabla^2\Psi = -4\pi\rho
-            % Dirichlet boundary conditions for a conductor
-          \quad \Psi \in \mathcolor{ForestGreen}{\mathcal{C}^2_0(C)}
-        \end{equation*}
-  \item Project onto the space of test functions
-        % Project onto the space of test functions
-        \begin{equation*}
-          \left\{
-          \begin{aligned}
-            &\text{Seek $\Psi\in\mathcolor{ForestGreen}{V_0}$ such that:}\\
-            &(\zeta, \nabla^2\Psi) =
-            -4\pi(\zeta, \rho) \quad
-            \forall
-            \zeta\in\mathcolor{BurntOrange}{\mathcal{C}^\infty_0(C)}
-          \end{aligned}
-          \right.
-        \end{equation*}
-  \item Use
-    $\eta\nabla^2\gamma = \nabla\cdot(\eta\nabla\gamma) - \nabla\eta\cdot\nabla\gamma$
-        % The integral of the first term on the RHS vanishes.
-        % It can be shown using Gauss' Theorem and the conductor
-        % boundary conditions.
-        \begin{equation*}
-          \left\{
-          \begin{aligned}
-            &\text{Seek $\Psi \in \mathcolor{ForestGreen}{H_0^1(C)}$ such that:}\\
-            &\mathcolor{Red}{(\nabla\zeta, \nabla\Psi)} =
-            \mathcolor{Blue}{-4\pi(\zeta, \rho)} \quad
-            \forall \zeta\in \mathcolor{BurntOrange}{H_0^1(C)}
-          \end{aligned}
-          \right.
-        \end{equation*}
-\end{itemize}
-\begin{equation*}
-  \Psi\in \mathcolor{BurntOrange}{H_0^1(C)} = \lbrace f : C \rightarrow
-  \mathbb{R} | f \in L^2(C), \nabla f \in L^2(C), f = 0 \forall \vect{s}
-  \in\partial C \rbrace
-\end{equation*}
-\begin{equation}
-  \left\{
-  \begin{aligned}
-    &\text{Seek $\Psi \in V$ such that:}\\
-    &\forall \zeta \in V \,\,\,\,\,
-    \mathcolor{Red}{a(\zeta, \Psi)} = \mathcolor{Blue}{b(\zeta)}
-  \end{aligned}
-  \right.
-  \label{eq:weak}
-\end{equation}
-
-% Hadamard well-posedness definition:
-% Problem \eqref{eq:weak} is well-posed if it admits one and only
-% one solution and the solution is bounded by the initial data, by
-% an a priori estimate (see Ern&Guermond, p.82, Def. 2.1)
-\begin{defin}[Well-posedness]\label{def:hadamard}
-  The weak problem Eq. \eqref{eq:weak} is \emph{well-posed} if it
-  admits one and only one solution and the solution is bounded by
-  the initial data by an \emph{a priori} estimate.
-\end{defin}
-\begin{defin}[Continuity]\label{def:continuity}
-  A bilinear form on a normed vector space $V$ is \emph{bounded}, or
-  \emph{continuous}, if there is a constant $C$ such that
-  $\forall u, v \in V$:
-  \[
-  a(u, v) \leq C\lVert u\rVert\lVert v\rVert
-  \]
-\end{defin}
-\begin{defin}[Coercivity]\label{def:coercivity}
-  A bilinear form on a normed vector space $V$ is \emph{coercive},
-  or \emph{elliptic}, if there is a constant $\alpha >0$ such that
-  $\forall u \in V$:
-  \[
-  a(u, u) \geq \alpha\lVert u\rVert^2
-  \]
-\end{defin}
-This implies that no eigenvalue of the linear operator associated
-to the bilinear form can be zero, hence that the operator is
-invertible!!~\autocite{Ern2004-oo}
-
-% Maybe present connection with Steepest Descent and Conjugate
-% Gradient?
-% Introduce a suitable vector space, equipped with a norm
-% derived from a scalar product.
-% The solution belongs to such a vector space, which is
-% bigger that in the strong formulation (less constraints on
-% the regularity of the solution)
-\begin{equation}
-  \left\{
-  \begin{aligned}
-    &\text{Seek $\Psi \in V$ such that:}\\
-    &\forall \zeta \in V \,\,\,\,\,
-    \mathcolor{Red}{a(\zeta, \Psi)} = \mathcolor{Blue}{b(\zeta)}
-  \end{aligned}
-  \right.
-\end{equation}
-\begin{lemma}[Lax--Milgram]
-  If the bilinear form $a$ is \emph{continous} and \emph{coercive}
-  in $V$, then, for any continuous linear form $b$, Problem~\eqref{eq:weak} is
-  \emph{well-posed}.
-\end{lemma}
-\begin{corollary}[Variational property]
-  If the bilinear form is \emph{symmetric} and \emph{positive}
-  the unique solution to Problem \eqref{eq:weak} is
-  the \emph{unique minimum} on $V$ of the functional:
-  \[
-  \mathcal{F}(\zeta) = \frac{1}{2}a(\zeta, \zeta) - b(\zeta)
-  \]
-\end{corollary}
-
-Weak formulation of \acs{BIE}s.\autocite{Hsiao2008-xb}
-
 \section{Numerical Approaches to Boundary Integral Equations}\label{sec:BEM}
 
-Approximability, conformity, consistency and asymptotic consistency
-of the approximation setting, $\inf$-$\sup$ condition, Céa's Lemma
-and generalized $\inf$-$\sup$ condition.~\autocite{Ern2004-oo,
-TesiFilippo}
-Additional cavity generators
-TsLess~\autocite{Pomelli2004-lb}
-DefPol~\autocite{Pomelli1998-ob, Pomelli1999-dc}
-FIXPVA~\autocite{Su2009-em}
-Other discretization schemes
-CSC~\autocite{York1999-xy, Scalmani2010-tw}
-SWIG~\autocite{Lange2010-jp, Lange2010-qo, Lange2011-eu}
-ddCOSMO and ddPCM~\autocite{Cances2013-jg, Lipparini2013-cy, Lipparini2014-to,
-Lipparini2014-fq, Stamm2016-fv}
+The solution of the \acs{IEF}-\acs{PCM} boundary integral equation can
+only be achieved analytically for simple cavity geometries, such as
+single spheres and ellipsoids.
+When more general molecular cavities are employed one has to resort to a
+numerical technique: the \ac{BEM}. This is implemented in four steps:
+\begin{enumerate*}[label={\alph*)},font={\color{PMS1797}}]
+    \item define the molecular cavity
+    \item mesh the molecular surface \label{item:mesh}
+    \item discretize the integral equation \label{item:discretize} and
+    \item solve the linear systems by a suitable numerical technique.
+\end{enumerate*}
+Many different approaches to these four steps have been presented, both
+in the field of the \acs{BEM} and in the context of the \acs{PCM}.
 
-\todo[inline]{CURRENTLY COPIED VERBATIM FROM THE SPHERICAL GREEN'S
-FUNCTION PAPER}
+The molecular cavity is usually defined as a set of atom-centered
+interlocking spheres. The radii are chosen from a suitable set, usually
+van der Waals radii.~\autocite{Bondi1964-dt, Mantina2009-hb}
+If one imagines to model a solvent molecule as a rigid sphere,
+three types of molecular surfaces can be defined:
+\begin{enumerate*}[label={\alph*)},font={\color{PMS1797}}]
+  \item
+    the union of the interlocking spheres surfaces is the \ac{vdWS},
+    Figure \ref{fig:vdWS}
+  \item the locus of points defined by the center of the solvent
+    spherical probe while rolling over the \acs{vdWS} is the
+    \ac{SAS}, Figure \ref{fig:SAS}
+  \item the locus of points defined by the contact point of the solvent
+    spherical probe while rolling over the \acs{vdWS} is the \ac{SES}, Figure \ref{fig:SES}.
+\end{enumerate*}
+\begin{figure}[tb]
+ \centering
+ \begin{subfigure}{.45\textwidth}
+   \centering
+   \includegraphics[width=.85\textwidth]{vdWS.png}
+   \caption{\acl{vdWS} (\acs{vdWS})}
+   \label{fig:vdWS}
+ \end{subfigure}
+ ~
+ \begin{subfigure}{.45\textwidth}
+   \centering
+   \includegraphics[width=.85\textwidth]{SAS.png}
+   \caption{\acl{SAS} (\acs{SAS})}
+   \label{fig:SAS}
+ \end{subfigure}
 
-The numerical solution of boundary integral equations, such as the
-\acs{IEF}-\acs{PCM}
-Eqs.~\eqref{eq:full-IEF-2pi}-\eqref{eq:isotropic-IEF}, can be achieved
-by the application of the \ac{BEM}.
-A discretization of the molecular cavity
-boundary is necessary.~\autocite{Pomelli2007-wq}
-The discretization of the
-molecular surface into finite elements leads to a corresponding
-discretization of both the integral operators and the functions in
-Eqs.~\eqref{eq:full-IEF-2pi}-\eqref{eq:isotropic-IEF}.
+ \begin{subfigure}{.45\textwidth}
+   \centering
+   \includegraphics[width=.85\textwidth]{SES.png}
+   \caption{\acl{SES} (\acs{SES})}
+   \label{fig:SES}
+ \end{subfigure}
+ \caption{Molecular surfaces from interlocking sets of atom-centered spheres.
+ Reproduced from \noparcite[ref.][]{Harbrecht2011-dk}.}
+\end{figure}
+Alternative definitions based on the isodensity surface of the molecular solute
+have been put forth.~\autocite{Foresman1996-wv}
 
-Different strategies are available to perform the
-discretization: the \emph{collocation} or the \emph{Galerkin}
-approach.~\cite{Hackbusch1995-uq, Ern2004-oo}
-The final result is the transformation of
-the original integral equation to a system of linear equations whose
-dimension relates to the number of finite elements generated by
-discretizing the boundary.  For general integral equations and
-discretization schemes, a dense system matrix is obtained. Moreover, the
-discretization scheme might not, in general, preserve the symmetry
-property of the original integral equation.
+Steps \ref{item:mesh} and \ref{item:discretize} are closely intertwined,
+in the technical \acs{BEM} literature they go under the name of
+\emph{finite element discretization}.
+Let us define a mesh of the domain $\Gamma$ as an indexed collection of
+intervals with non-zero measure $\lbrace I_i\rbrace_{i=1}^{N_\mathrm{mesh}}$, where
+$N_\mathrm{mesh}$ is the size of the mesh.
+Then, in a rather informal sense, a finite element is the
+mathematical entity formed by the vector space of functions that are
+piecewise regular on any given interval in the mesh.~\autocite{Ern2004-oo}
+Basically, given an element in the mesh, we "attach" a basis set of
+piecewise polynomials of a prescribed degree and use this basis to
+provide a representation of quantities defined on $\Gamma$ within each
+mesh interval.
+Mesh generation for the molecular surfaces mentioned above is usually
+achieved by modified triangulation procedures. In this thesis we used
+the GePol algorithm~\autocite{Pascual-Ahuir1987-uo, Pascual-Ahuir1990-lp,
+Pomelli1998-qp, Pomelli2001-sj, Frediani2004-ua, Pomelli2007-wq}
+and the mesh generators of Harbrecht and Randrianarivony especially
+tailored for the wavelet \acs{BEM}.~\autocite{Harbrecht2009-no,
+Harbrecht2011-dk}
 
-In this paper we apply the traditional centroid collocation
-method~\autocite{Tomasi2005-vm, Pomelli2007-wq} on the cavity boundary
-discretization obtained \emph{via} the GePol
-algorithm.~\autocite{Pascual-Ahuir1987-uo, Pascual-Ahuir1990-lp,
-Pomelli1998-qp, Pomelli2001-sj, Frediani2004-ua, Pomelli2007-wq}
-In this algorithm, the cavity is made of interlocking spheres, not all
-of them necessarily centered on the atoms. The surface is then
-discretized by means of $N_\mathrm{ts}$ small triangles $\Delta_i$ whose
-centroid defines the \emph{collocation point}. Once the centroid is
-properly defined, functions and operators are discretized by applying a
-one-point quadrature rule in the finite element centroid.
-The \acs{IEF}-\acs{PCM} equation is thus recast as:
+The finite element discretization is usually carried out either by the
+\emph{collocation} or by the \emph{Galerkin} approach.\footnote{
+Indeed, it is possible to show that discretization by collocation
+corresponds to a Galerkin method with a very specific (and asymmetric)
+choice of \emph{trial} and \emph{test}
+spaces.~\autocite{Hackbusch1995-uq, Ern2004-oo}}
+This step transforms the original integral equation into a system of
+linear equations whose dimension relates to the underlying mesh size.
+For general integral equations and finite element discretizations,
+the \emph{stiffness matrix} obtained is dense and proper linear algebra
+techniques need to be employed in order to avoid a detrimental impact on
+computational performance.
+In addition, finite element discretization might destroy intrinsic
+symmetries of the original integral
+operators. These can be recovered by appropriate \emph{a posteriori}
+symmetrization procedures introducing, however, a degree of
+arbitrariness in the \acs{BEM} procedure.~\autocite{Lange2011-eu}
+
+In the GePol procedure, the molecular surface is meshed by means of
+$N_\mathrm{ts}$ spherical polygons $\Pi_i$ whose centroids define the
+collocation points. One-point quadrature rules are then applied at the
+centroid to obtain the matrix equation:
 \begin{equation}
-  \vect{q}
+ \mat{R}^{-1}\mat{T} \vect{q}
   =
-  \mat{T}^{-1}\mat{R}\vect{v}.
+  \vect{v}.
 \end{equation}
@@
 collecting the values of the \acs{ASC} and \ac{MEP}
-at the collocation points.
+at the collocation points. $\mat{R}^{-1}\mat{T}$ is the stiffness
+matrix.
 The matrices representing the boundary integral operators are:
@@
 other matrices are defined as:
-\begin{alignat}{2}
-  &\mat{A} = \mathrm{diag}(\cdots\,\, a_i = \mathrm{area}(\Delta_i)\,\, \cdots) \\
-  &S_{ii, \mathrm{i}} = k\sqrt{\frac{4\pi}{a_i}}, \quad
-  S_{ij, \mathrm{i}} = \frac{1}{|\vect{s}_i-\vect{s}_j|}\quad (i \neq j) \\
-  &D_{ii, \mathrm{i}} = -k\sqrt{\frac{\pi}{a_i}}\frac{1}{R_i}, \quad
-  D_{ij, \mathrm{i}} =
-  \frac{(\vect{s}_i-\vect{s}_j)\cdot\vect{n}_j}{|\vect{s}_i-\vect{s}_j|^3}\quad (i \neq j) \\
-  &S_{ii, \mathrm{e}} = \frac{S_{ii, \mathrm{i}}}{C(\vect{s}_i, \vect{s}_i)} +
-        G_\mathrm{img}(\vect{s}_i, \vect{s}_i), \quad
-  S_{ij, \mathrm{e}} =
-  \frac{1}{C(\vect{s}_i, \vect{s}_j)|\vect{s}_i-\vect{s}_j|} +
-  G_\mathrm{img}(\vect{s}_i, \vect{s}_j) \quad (i \neq j) \\
-  &D_{ii, \mathrm{e}} = \diel(\vect{s}_i)\left[
-  \frac{D_{ii, \mathrm{i}}}{C(\vect{s}_i, \vect{s}_i)}
-  -S_{ii, \mathrm{i}}
-  \frac{\nabla C(\vect{s}_i,
-  \vect{s}_i)\cdot\vect{n}_i}{C^2(\vect{s}_i, \vect{s}_i)}
-  +
-  \nabla G_\mathrm{img}(\vect{s}_i, \vect{s}_i)\cdot\vect{n}_i
-  \right] \\
-  &D_{ij, \mathrm{e}} =
-  \diel(\vect{s}_j)\left[
-  \frac{1}{C(\vect{s}_i, \vect{s}_j)}
-  \frac{(\vect{s}_i-\vect{s}_j)\cdot\vect{n}_j}{|\vect{s}_i-\vect{s}_j|^3}
-  -\frac{\nabla C(\vect{s}_i,
-  \vect{s}_j)\cdot\vect{n}_j}{C^2(\vect{s}_i, \vect{s}_j)}
-  \frac{1}{|\vect{s}_i-\vect{s}_j|}
-  + \nabla G_\mathrm{img}(\vect{s}_i, \vect{s}_j)\cdot\vect{n}_j
-  \right]\quad (i \neq j)
-\end{alignat}
+\begin{subequations}
+  \begin{align}
+    A_{ij} &= a_i\delta_{ij}, \quad a_i = \mathrm{area}(\Pi_i) \\
+    S_{ij, \mathrm{i}} &=
+    k\sqrt{\frac{4\pi}{a_i}}\delta_{ij} +
+    \frac{1}{|\vect{s}_i-\vect{s}_j|}(1-\delta_{ij}) \\
+    D_{ii, \mathrm{i}} &=
+    -k\sqrt{\frac{\pi}{a_i}}\frac{1}{R_i}\delta_{ij}
+    + \frac{(\vect{s}_i-\vect{s}_j)\cdot\vect{n}_j}{|\vect{s}_i-\vect{s}_j|^3}(1-\delta_{ij})
+  \end{align}
+\end{subequations}
 in terms of the centroids $\vect{s}_i, \vect{s}_j$, the finite elements
-areas and their curvatures $R_i$. The factor $k=1.07$ is introduced to achieve a better precision on the diagonal elements~\cite{Tomasi2005}.
+areas and their curvatures $R_i$.
+The factor $k=1.07$ is introduced to achieve a better precision on the
+diagonal elements.~\autocite{Tomasi2005-vm}
+Expressions for $\mat{S}_\mathrm{e}$, $\mat{D}_\mathrm{e}$ and
+$\mat{D}^\dagger_\mathrm{e}$ depend on the specific outer environment
+and can be found in the literature.~\autocite{Corni2002-dr, Frediani2004-er,
+Tomasi2005-vm, Delgado2013-kd, DiRemigio2016-nn}
 
-\todo[inline]{CURRENTLY COPIED VERBATIM FROM THE WAVELET PAPER}
+The wavelet Galerkin \acs{BEM} was also used in this thesis. This method
+preserves the fundamental symmetries of the boundary integral operators
+and achieves sparsity in the stiffness matrix.
+This sparsity can be leveraged to implement efficient algorithms with
+linear space and time complexity. The approach is thus superior
+to the traditional collocation method.
 
-Boundary integral equations, such as the \acs{IEF}-\acs{PCM}
-Eqs.~\eqref{eq:full-IEF-2pi}-\eqref{eq:isotropic-IEF},
-can conveniently be solved numerically by the application of the
-\ac{BEM}.
-Both the integral operators and the functions on which these act are
-defined solely on the boundary of the cavity.
+Given a hierarchical sequence of trial spaces
+$\lbrace V_l\rbrace_{l=1}^J$, one can decompose the $l$-th element in
+the hierarchy into the direct sum $V_l = V_{l-1} \oplus W_l$.
+$W_l$ is called the \emph{wavelet space} and is the, not necessarily
+orthogonal, complement to $V_{l-1}$.
+Recursively applying this splitting to the trial spaces hierarchy
+generates the wavelet decomposition, where all complementary spaces are
+spanned by the wavelet basis.
+The fundamental insight in the wavelet \acs{BEM} is to use the wavelet
+basis for the discretization step and employ the compression technique
+described by \citeauthor{Dahmen2006-pj} to build up the sparse system
+matrix.~\autocite{Harbrecht2004-uo, Harbrecht2004-ug, Dahmen2006-pj}
+Matrix compression is carried out in two rounds, \emph{a priori} and \emph{a
+posteriori}, resulting in a finger-like sparsity pattern for the
+stiffness matrix, Figure \ref{fig:finger-structure}.
+The number of relevant matrix entries scales proportional to $O(N_J)$
+with $N_J$ the number of degrees of freedom for a refinement of the
+geometry up to level $J$.
 
-The application of the boundary element method requires that the boundary of
-the molecular cavity is discretized by a number of suitable finite elements.
-The discretization of the boundary leads to a discretization of the integral
-operators. This discretization can be carried out by using either the
-\emph{collocation} or the \emph{Galerkin}
-approach.~\autocite{Hackbusch1995-uq, Ern2004-oo}
-In both cases, the integral equation is transformed to a system of
-linear equations whose dimension is related to the number of finite
-elements used in the discretization of the boundary.
-The resulting system matrix is, in general, a dense matrix. The
-boundary element method thus suffers from limitations imposed by the
-number of matrix elements to be stored and the memory and time
-requirements of solving the resulting linear system.
+\begin{figure}[tb]
+   \includegraphics[width=\textwidth]{a_2_l_0_5_fs.png}
+   \caption[Sparsity pattern achieved by the wavelet Galerkin \acs{BEM}.]{
+   Sparsity pattern achieved by the wavelet Galerkin \acs{BEM} for a
+   model problem (Laplace equation on 6 patches).
+   Reproduced from \noparcite[ref.][]{Bugeanu2015-tp}.
+   }
+   \label{fig:finger-structure}
+\end{figure}
 
-The use of a wavelet basis in the Galerkin approach has been proven
-beneficial in this respect.~\autocite{Harbrecht2004-uo, Dahmen2006-pj,
-Harbrecht2006-ug}.
-The resulting system matrices are quasi-sparse
-and can be further reduced to a sparse form by discarding
-negligible entries without considerable loss of precision.
+Continuity of meshes with respect to molecular distortions is a
+central issue for stable implementations of molecular gradients and
+recurs frequently in the literature.
+Numerous schemes have been proposed to solve or alleviate this issue.
+We here mention the TsLess~\autocite{Pomelli2004-lb} and
+DefPol~\autocite{Pomelli1998-ob, Pomelli1999-dc} approaches of
+\citeauthor{Pomelli1998-ob}, the FIXPVA
+modification~\autocite{Su2009-em} to the GePol algorithm put forth by
+\citeauthor{Su2009-em}.
+\citeauthor{Scalmani2010-tw} have recently proposed the \ac{CSC} scheme
+which uses a smooth \acs{vdWS} and 3D spherical Gaussian basis
+functions.~\autocite{Scalmani2010-tw, York1999-xy}
+Gaussian blurring was also advocated by \citeauthor{Lange2010-jp} in
+their \ac{SWIG} method.~\autocite{Lange2010-jp, Lange2010-qo}
+Both \acs{CSC} and \acs{SWIG} achieve continuous \acsp{PES} and smooth
+molecular gradients. However, their convergence to the exact solution of
+the underlying integral equation has yet to be proved.
+A completely different approach to the problem has been taken by
+\citeauthor{Cances2013-jg} by employing a domain decomposition
+method, guaranteed to converge to the exact solution.~\autocite{Quarteroni1999-jt}
+Moreover, for \acs{COSMO}, the formulation is embarrassingly
+parallel and has been implemented in a linearly scaling fashion.~\autocite{Cances2013-jg, Lipparini2013-cy, Lipparini2014-to,
+Lipparini2014-fq}
+Very recently the method has also been extended to general dielectric
+environments.~\autocite{Stamm2016-fv}
 
-The wavelet boundary element method starts by defining
-a sequence of hierarchical trial spaces,
-spanned by standard finite element ansatz functions:
+In the following, the \acs{PCM} equations will be written in the
+"complete basis" meaning that we will be working with the exact integral
+equation and not with its discretized counterpart.
+As a consequence, the apparent surface charge $\sigma$ and
+the electrostatic potential $\esp$ will have a \emph{continuous}
+dependence on a cavity surface "index" $\vect{s}$.
+A product involving any of these quantities will have to be
+interpreted as the \emph{surface integral}, \ie the scalar product in
+the suitable, infinite-dimensional vector space on the cavity boundary
+$\Gamma$.
+The following shorthand notations will be adopted:
 \begin{equation}
-  \{0\} = V_{-1} \subset V_0 \subset V_1 \subset \ldots
-  \subset V_J, \qquad V_j = \Span{\phi_{j,k} : k\in \Delta_{j}}.
+  \begin{aligned}
+  \sigma\PCM\sigma &=
+  \int_\Gamma\diff\vect{s}\sigma(\vect{s})\PCM\sigma(\vect{s})
+  =
+  \scalprod{\sigma(\vect{s})}{\PCM\sigma(\vect{s})} \\
+  \sigma\esp &=
+  \int_\Gamma\diff\vect{s}\sigma(\vect{s})\esp(\vect{s})
+  =
+  \scalprod{\sigma(\vect{s})}{\esp(\vect{s})}
+  \end{aligned}
 \end{equation}
-Here, $\Delta_j$ is an index set for the
-single-scale basis of the space $V_j$. In the wavelet method,
-the trial space $V_j$ is split into the direct sum $V_j = V_{j-1}
-\oplus W_j$. The resulting complementary space $W_j$ is called the
-wavelet space and is not necessarily orthogonal to $V_{j-1}$.
-% It suffices the equation $W_j \cap V_{j-1} = \{0\}$,
-Recursive splitting of the trial spaces leads to the wavelet
-decomposition $V_j = \bigoplus_{l = 0}^{j} W_l$.
-
-The complementary space is spanned by the wavelet basis:
-\begin{equation} W_j =
-  \Span{\psi_{j,k}: k \in \Delta_{j+1}
-\setminus \Delta_{j}}.
-\end{equation}
-The choice of this basis turns
-out to be very convenient, since we can exploit the compression
-technique described by \citeauthor{Dahmen2006-pj} to build up the sparse
-system matrix in the wavelet basis directly, avoiding the computation of
-non-relevant matrix entries.~\autocite{Dahmen2006-pj}
-The \emph{a priori} compression is carried out in two steps. In the
-first step, contributions from basis functions sufficiently far away
-from each other are discarded. In the following, elements of the matrix
-where one wavelet is in the smooth part of the other one are ignored.
-The number of relevant matrix entries scales proportional to $O(N_J)$ with
-$N_J$ the number of degrees of freedom for a refinement of the geometry
-up to level $J$.
-
-In order to understand the compression steps, let us introduce some
-notation. Let the \emph{convex hull} of the support of the wavelet $\psi_{j,k}$
-be:
-\begin{equation}
-  \Omega_{j,k} = \conv\hull\left(\supp \psi_{j,k}\right).
-\end{equation}
-Moreover, let the \emph{singular support} of the wavelet $\psi_{j,k}$ be:
-\begin{equation}
-  \Omega'_{j,k} = \sing\supp\psi_{j,k}
-\end{equation}
-It contains all the points in the support where the wavelet is not smooth.
-The compression steps are then governed by the following set of
-equations
-\begin{equation}
-  \left[A_J\right]_{\left(j,k\right)\left(j',k'\right)} =
-  \begin{cases}
-    \quad\quad 0 & \dist(\Omega_{j,k}, \Omega_{j',k'}) > B_{j,j'} \text{ and }j,j' \geq 0, \\
-    \quad\quad 0 & \dist(\Omega_{j,k}, \Omega_{j',k'}) \lesssim 2^{-\min\left\{j,j'\right\}} \,\text{and} \\
-      & \quad \dist(\Omega'_{j,k}, \Omega_{j',k'}) > B'_{j,j'} \text{ if }j' > j \geq 0, \\
-      & \quad \dist(\Omega_{j,k}, \Omega'_{j',k'}) > B'_{j,j'} \text{ if }j > j' \geq 0, \\
-    \langle A\psi_{j',k'},\psi_{j,k}\rangle, & \text{otherwise}
-  \end{cases}
-\label{eq:compression}
-\end{equation}
-where $\dist(\cdot, \cdot)$ denotes the distance, either between
-the bounding boxes of the wavelets or between the singular support and the
-bounding box. The first and second conditions represent the first and second
-compression step, respectively. The parameters $j,j'$ are the
-levels of the wavelets under consideration and the
-level-dependent cut-off parameters $B_{j,j'}$ and $B'_{j,j'}$ are given
-by
-\begin{align*}
-  B_{j,j'} &= a \max
-  \left\{2^{-\min\left\{j,j'\right\}},
-  2^{\frac{J\left(2d^\prime-op\right) -
-  \left(j+j'\right)\left(d^\prime+\tilde{d}\right)}{\left(2\tilde{d}+op\right)}}\right\}
-  \text{ and }\\ B'_{j,j'} &= a \max
-  \left\{2^{-\min\left\{j,j'\right\}},
-  2^{\frac{J\left(2d^\prime-op\right) -
-\left(j+j'\right)d^\prime-\max\left\{j,j'\right\}\tilde{d}}{\left(\tilde{d}+op\right)}}\right\}
-\end{align*}
-with $op$ being the order of the integral operator under consideration.
-For the first kind integral equation $op = -1$, while $op = 0$ for the second
-kind integral equation. The integer $\tilde{d}$ is related to the vanishing
-moments of the wavelet basis:
-\begin{equation}
-  \int \diff\vect{x}\,\,\vect{x}^r \psi_{j,k}(\vect{x}) = 0,
-\quad r = 0, \ldots \tilde{d}-1
-\end{equation}
-In the particular implementation, $\tilde{d}=3$ for the piecewise
-constant wavelet basis,  while $\tilde{d}=4$ for the piecewise bilinear
-parametrization.~\autocite{Harbrecht2013-hb}
-The compression can thus be adjusted by the parameters $a$ and
-$d^\prime$:
-\begin{equation}\label{eq:param}
-  a \geq 1, \quad d < d^\prime < \tilde{d} + op
-\end{equation}
-where $d$ is the approximation order of the trial spaces $V_j$:
-$d=1, 2$ for piecewise constant and
-piecewise bilinear ansatz functions, respectively.
-Note that, if the interaction between two wavelets, $\psi_{j,k}$ and $\psi_{j',k'}$, on
-level $j$ and $j'$ is neglected in the system matrix, all other
-interactions between wavelets resulting from the refinement of
-$\psi_{j,k}$ and $\psi_{j',k'}$ can also be ignored. Thus, the
-compression pattern of the system matrix is calculated
-hierarchically starting from the coarsest level.
-
-Once the compressed system matrix is assembled, we arrive at a sparse
-system matrix in the wavelet basis which can be compressed further by
-leaving out the sufficiently small elements. This post-processing step
-is governed by the rule
-\begin{equation}
-  \left[A_J\right]_{\left(j,k\right)\left(j',k'\right)}
-  \begin{cases}
-    \quad\quad 0 & \text{if } |\left[ A_J \right]_{\left(j,k\right)\left(j',k'\right)}| \leq \epsilon_{j,j'}\\
-    \left[A_J\right]_{\left(j,k\right)\left(j',k'\right)}, & \text{otherwise,}
-  \end{cases}
-  \label{eq:aposteriori}
-\end{equation}
-where the coefficients $\epsilon_{j,j'}$ are given by
-\begin{equation}
-%\epsilon_{j,j'} = b\min \left\{2^{-|j-j'|}, 2^{-\left(2J-\left(j+j'\right)\right)\frac{2d^\prime-op}{2\tilde{d}+op}}\right\} 2^{J\,op}2^{-2d^\prime\left(J-\frac{j+j'}{2}\right)},
-  \epsilon_{j,j'} = b\min \left\{2^{-|j-j'|}, 2^{-\left(2J-\left(j+j'\right)\right)\frac{2d^\prime-op}{2\tilde{d}+op}}\right\} 2^{-2d^\prime\left(J-\frac{j+j'}{2}\right)}
-\end{equation}
-with the \emph{a posteriori} compression parameter $b < 1$.
 
@@
 
+\begin{equation}\label{eq:Y-operator}
+  \PCM = \bi{R}_\infty^{-1}\bi{R}_\diel\bi{S}
+\end{equation}
+
 Variational functional for implicit models, \acs{PCM} in this
@@
 \begin{equation}
- U_\mathrm{PCM} = \frac{1}{2}\scalprod{\sigma}{\PCM\sigma} + \scalprod{\esp}{\sigma}
+ U_\mathrm{PCM} = \frac{1}{2}\sigma\PCM\sigma + \esp\sigma
 \end{equation}
@@
   U_\mathrm{MM} = \frac{1}{2}\kappa\MM\kappa + \kappa\zeta
-\end{equation}
-
-The PCM equations will be written in the ``complete basis'' meaning that
-we will introduce the usual \ac{BEM} discretization
-at the very end of the derivation. In other words, we will be working
-with the exact integral equation and not with its discretized
-counterpart. As a consequence, the apparent surface charge $\sigma$ and
-the electrostatic potential $\esp$ will have a \emph{continuous}
-dependence on a ``cavity surface'' index $\vect{s}$. Whenever a
-charge-potential product is present it is then to be interpreted as the
-\emph{surface integral}, i.e. the scalar product in the suitable,
-infinite-dimensional vector space on the cavity boundary $\Gamma$. The
-following shorthand notation will be adopted:
-\begin{equation}
- \sigma\esp = \scalprod{\sigma}{\esp}
 \end{equation}

### 2016-08-31 Chapters/MathematicalResults.tex
Paper summaries done
Work on CSM chapter

@@
 
+This appendix collects some results in the field of functional analysis and
+boundary integral equations. No proofs or examples are given, the interested
+reader is referred to the monographs by \citeauthor{Ern2004-oo},
+\citeauthor{Hsiao2008-xb} and \citeauthor{Sauter2011-an}\autocite{Ern2004-oo,
+Hsiao2008-xb, Sauter2011-an}
+
+\section{Preliminary Definitions}
+
+\begin{defin}[Sobolev spaces]
+Let $s$ and $p$ be two integers with $s\geq 0$ and $1\leq p \leq +\infty$.
+The so-called \emph{Sobolev space} $W^{s, p}(\Omega)$ is defined as
+\begin{equation}
+W^{s,p}(\Omega) =
+\left\lbrace
+u \in \mathcal{D}^\prime(\Omega) | \partial^\alpha u \in L^p(\Omega), |\alpha|\leq s
+\right\rbrace
+\end{equation}
+where $\mathcal{D}^\prime(\Omega)$ is the space of Schwartz distributions and the
+derivatives $\partial^\alpha u$ are understood in a distributional sense.
+\end{defin}
+
+\begin{lemma}[Hilbert Sobolev spaces]
+Let $s\geq 0$. The space $H^s(\Omega) = W^{s, 2}(\Omega)$ is a Hilbert space
+when equipped with the scalar product
+\begin{equation}
+(u, v)_{s, \Omega} = \sum_{|\alpha|\leq s}\int_\Omega\partial^\alpha u \partial^\alpha v.
+\end{equation}
+The associated norm is denoted by $\| \cdot \|_{s, \Omega}$.
+\end{lemma}
+
+\begin{defin}[Fractional Sobolev spaces]
+For $0 < s < 1$ and $1 \leq p < +\infty$, the so-called \emph{Sobolev space with fractional
+exponent} is defined as
+\begin{equation}
+W^{s, p}(\Omega) =
+\left\lbrace
+u \in L^p(\Omega) |
+\frac{u(\vect{r}) - u(\vect{r}^\prime)}{\| \vect{r} - \vect{r}^\prime\|^{s+\frac{d}{p}}}
+\in L^p(\Omega \times \Omega)
+\right\rbrace.
+\end{equation}
+Furthermore, when $s> 1$ is not integer, letting $\sigma = s - [s]$, where $[s]$ is the integer
+part of $s$, $W^{s,p}(\Omega)$ is defined as
+\begin{equation}
+W^{s, p}(\Omega) =
+\left\lbrace
+u \in W^{[s], p}(\Omega) |
+\partial^\alpha u \in W^{\sigma, p}(\Omega)\, \forall \alpha,\, |\alpha| = [s]
+\right\rbrace.
+\end{equation}
+When $p=2$, we denote $H^s(\Omega) = W^{s,2}(\Omega)$.
+\end{defin}
+
+\begin{defin}[Continuity]\label{def:continuity}
+  A bilinear form on a normed vector space $V$ is \emph{bounded}, or
+  \emph{continuous}, if there is a constant $C$ such that
+  $\forall u, v \in V$:
+  \[
+  a(u, v) \leq C\lVert u\rVert\lVert v\rVert
+  \]
+\end{defin}
+
+\begin{defin}[Coercivity]\label{def:coercivity}
+  A bilinear form on a normed vector space $V$ is \emph{coercive},
+  or \emph{elliptic}, if there is a constant $\alpha >0$ such that
+  $\forall u \in V$:
+  \[
+  a(u, u) \geq \alpha\lVert u\rVert^2
+  \]
+\end{defin}
+Coercivity implies that no eigenvalue of the linear operator associated to the
+bilinear form can be zero, hence its invertibility.~\autocite{Ern2004-oo}
+
+\begin{defin}[Abstract weak problem]\label{def:abstract-weak}
+An abstract weak problem is posed as follows:
+\begin{equation}
+  \left\{
+  \begin{aligned}
+    &\text{Seek $u \in W$ such that:}\\
+    &\forall v \in V \,\,\,\,\,
+    a(u, v) = b(v)
+  \end{aligned}
+  \right.
+  \label{eq:weak}
+\end{equation}
+where:
+\begin{itemize}
+  \item $W$ and $V$ are normed vector spaces. $W$ is the \emph{solution space},
+  $V$ the \emph{test space}.
+  \item $a$ is a \emph{continuous} bilinear form on $W \times V$.
+  \item $b$ is a \emph{continous} linear form on $V$.
+\end{itemize}
+\end{defin}
+
+\begin{defin}[Well-posedness]\label{def:hadamard}
+The abstract weak problem in Definition \ref{def:abstract-weak} is
+\emph{well-posed} if it admits one and only one solution and the solution is
+bounded by the \emph{a priori} estimate:
+\begin{equation}
+\exists c > 0, \forall f \in V^\prime, \|u\|_W \leq c \|f\|_{V^\prime}
+\end{equation}
+where $V^\prime$ is the dual space of $V$.
+\end{defin}
+
+\begin{defin}[Transmission problem]
+We assume Euclidean space $\mathbb{R}^3$ to be partitioned into two subdomains
+$\Omegai$ and $\Omegae$ sharing a boundary $\Gamma$. We further assume that
+$\Omegai$ is a closed domain, entirely contained inside $\Omegae$.
+The transmission problem is posed as follows:
+\begin{subequations}\label{eq:transmission-general}
+\begin{align}[left={\empheqlbrace}]
+  \Li u &= f_\mathrm{i} \,\, \text{in}\,\, \Omegai \\
+  \Le u &= f_\mathrm{e} \,\, \text{in}\,\, \Omegae \\
+  [u] &= \ue - \ui = g_\mathrm{D} \,\, \text{on}\,\, \Gamma
+   \\
+[\partial_L u] &= \partiale u - \partiali u = g_\mathrm{N} \,\,
+\text{on}\,\, \Gamma \\
+|u(\vect{r})| &\leq C \|\vect{r} \|^{-1} \,\,\text{for}\,\,\| \vect{r} \|\rightarrow\infty
+\end{align}
+\end{subequations}
+where the differential operators are assumed elliptic and the jump conditions are
+given in terms of Dirichlet $g_\mathrm{D}$ and Neumann $g_\mathrm{N}$ data.
+\end{defin}
+
+\section{Derivation of the IEF equation}
+I will show a detailed derivation of the \acs{IEF} equation
+for the \acs{PCM} transmission problem Eq.~\eqref{eq:transmission}:
+\begin{subequations}
+  \begin{align}[left={\empheqlbrace}]
+  \Li u &= \nabla^2 u = -4\pi\rhoi \,\, \text{in}\,\, \Omegai \\
+  \Le u &= 0 \,\, \text{in}\,\, \Omegae \\
+  [u] &= \ue - \ui = 0 \,\, \text{on}\,\, \Gamma \\
+[\partial_L u] &= \partiale u - \partiali u = 0 \,\, \text{on}\,\, \Gamma \\
+|u(\vect{r})| &\leq C \|\vect{r} \|^{-1} \,\,\text{for}\,\,\| \vect{r} \|\rightarrow\infty
+\end{align}
+\end{subequations}
+
+We first state two important results in the theory of boundary integral
+equations.
 \begin{lemma}[Properties of the boundary integral operators]
-  The integral operators introduced above enjoy the following
-  properties~\autocite{Hsiao2008-xb, Sauter2011-an}:
+  The integral operators introduced in Eqs.~\eqref{eq:S}-\eqref{eq:D-dagger}
+  enjoy the following properties~\autocite{Hsiao2008-xb, Sauter2011-an}:
   \begin{enumerate}
@@
 
-\subsection{Lemma of Integral Representation}
-For the transmission problem \ref{eq:transmission} there holds:
-\begin{itemize}
-\item $\forall x \in \Omegai$
+\begin{lemma}[Integral Representation]
+For the transmission problem \ref{eq:transmission-general} there holds:
+\begin{enumerate}
+\item $\forall \vect{r} \in \Omegai$
 \begin{equation}
 u = \bi{S}_\mathrm{i}(\partiali u)
-- \bi{D}_\mathrm{i}(\ui) + \int_{\Omegai} \Gi f_\mathrm{i}\diff y
+- \bi{D}_\mathrm{i}(\ui) + \int_{\Omegai} \Gi f_\mathrm{i}\diff \vect{r}^\prime
 \end{equation}
-\item $\forall x \in \Omegae$
+\item $\forall \vect{r} \in \Omegae$
 \begin{equation}
 u = -\bi{S}_\mathrm{e}(\partiale u)
-+ \bi{D}_\mathrm{e}(\ue) + \int_{\Omegae}\Ge f_\mathrm{e}\diff y
++ \bi{D}_\mathrm{e}(\ue) + \int_{\Omegae}\Ge f_\mathrm{e}\diff \vect{r}^\prime
 \end{equation}
-\item $\forall x \in \Gamma$
+\item $\forall \vect{r} \in \Gamma$
   \begin{equation}\label{eq:stat3}
     \frac{1}{2}\ui = \bi{S}_\mathrm{i}(\partiali u)
-    - \bi{D}_\mathrm{i}(\ui) + \int_{\Omegai}\Gi f_\mathrm{i}\diff y
+    - \bi{D}_\mathrm{i}(\ui) + \int_{\Omegai}\Gi f_\mathrm{i}\diff \vect{r}^\prime
   \end{equation}
-\item $\forall x \in \Gamma$
+\item $\forall \vect{r} \in \Gamma$
   \begin{equation}\label{eq:stat4}
     \frac{1}{2}\ue = -\bi{S}_\mathrm{e}(\partiale u)
-    + \bi{D}_\mathrm{e}(\ue) + \int_{\Omegae}\Ge f_\mathrm{e}\diff y
+    + \bi{D}_\mathrm{e}(\ue) + \int_{\Omegae}\Ge f_\mathrm{e}\diff \vect{r}^\prime
   \end{equation}
-\end{itemize}
+\end{enumerate}
+\end{lemma}
 
 We introduce the following auxiliary potential:
-\begin{equation}
-  h(x) =
-  \begin{cases}
-    &\int_{\mathbb{R}^3}\Gi\rhoi(y)\diff y  \quad x \in \Omegai \\
-    &\int_{\mathbb{R}^3}\Ge\rhoi(y)\diff y  \quad x \in \Omegae
-  \end{cases}
-\end{equation}
-for which we have:\footnote{This can be checked by applying the
-differential operators to the definition of $h$ and then swapping
-differentiation and integration.}
-\begin{equation}
-  \begin{cases}
+\begin{empheq}[left={h(\vect{r}) = \empheqlbrace}]{align}
+    &\int_{\mathbb{R}^3}\Gi\rhoi(\vect{r}^\prime)\diff \vect{r}^\prime  \quad \vect{r} \in \Omegai \\
+    &\int_{\mathbb{R}^3}\Ge\rhoi(\vect{r}^\prime)\diff \vect{r}^\prime  \quad \vect{r} \in \Omegae
+\end{empheq}
+for which we have:
+\begin{empheq}[left={\empheqlbrace}]{align}
     \nabla^2 h &= \rhoi \,\, \text{in}\,\, \Omegai \\
     \Le h &= 0 \,\, \text{in}\,\, \Omegae
-  \end{cases}
-\end{equation}
+\end{empheq}
 We then define the \emph{reaction potential} as:
@@
 \begin{subequations}
-\begin{align}
+  \begin{align}[left={\empheqlbrace}]
   \nabla^2 \xi &= 0 \,\, \text{in}\,\, \Omegai \\
@@
 inside the cavity, the reaction potential can be represented by a
-\emph{single layer potential}, \ie by an apparent surface charge (ASC):
+\emph{single layer potential}:
 \begin{equation}
@@
 \end{equation}
+where the function $\sigma$ is the, yet unknown, \acl{ASC}.
 
-Applying Eqs. \eqref{eq:stat3} and \eqref{eq:stat4} of the Lemma of Integral
-Representation to the reaction potential yields:
-\begin{align}
-  \frac{1}{2}\xi_\mathrm{i} &= \bi{S}_\mathrm{i}(\partiali \xi) -
-  \bi{D}_\mathrm{i}(\xi_\mathrm{i}) \\
-  -\frac{1}{2}\xi_\mathrm{e} &= \bi{S}_\mathrm{e}(\partiale \xi) -
-  \bi{D}_\mathrm{e}(\xi_\mathrm{e})
-\end{align}
-in both cases the last term, including the integration over the entire
-subdomain volume, drops out ($f_\mathrm{i} = 0 = f_\mathrm{e}$).
-While use of Eq. \eqref{eq:stat4} on the auxiliary potential $h$ yields:
-\begin{equation}
-  -\frac{1}{2}h_\mathrm{e} =
-  \bi{S}_\mathrm{e}(\partiale h) -
-  \bi{D}_\mathrm{e}(h_\mathrm{e})
-\end{equation}
-Together with the jump conditions one obtains:
-\begin{subequations}
-  \begin{align}
+To derive an equation for $\sigma$, we set up a system of equations containing
+the jump conditions and the integral representations of the reaction and auxiliary potentials:
+\begin{empheq}[left={\empheqlbrace}]{align}
     &\bi{S}_\mathrm{i}(\partiali \xi) - \bi{D}_\mathrm{i}(\xi_\mathrm{i})
@@
   \label{eq:fifth}
-  \end{align}
-\end{subequations}
-
-\begin{enumerate}
-  \item Add Eqs. \eqref{eq:second} and \eqref{eq:third}:
+\end{empheq}
+The final ingredient is the \ac{DtN} map, which can be derived by employing Eq.~\eqref{eq:stat3}
+to the Newton potential:
 \begin{equation}
-    \bi{S}_\mathrm{e}(\partiale \xi + \partiale h) -
-    \bi{D}_\mathrm{e}(\xi_\mathrm{e} + h_\mathrm{e})
-    = -\frac{1}{2}(\xi_\mathrm{e} + h_\mathrm{e})
+  \phi(\vect{r}) = (\bi{N}\rhoi)(\vect{r}) = \int_{\mathbb{R}^3}\Gi\rhoi(\vect{r}^\prime)\diff \vect{r}^\prime
+   = \left.h(\vect{r})\right|_{\Omegai}
 \end{equation}
-  \item Exploit Eqs. \eqref{eq:fourth} and \eqref{eq:fifth} to
-    eliminate the external traces and conormal derivatives:
-    \begin{equation}
-      \bi{S}_\mathrm{e}(\partiali \xi) + \left(\frac{1}{2} -
-      \bi{D}_\mathrm{e}\right)\xi_\mathrm{i} =
-      - \bi{S}_\mathrm{e}(\partiali h) - \left(\frac{1}{2} -
-      \bi{D}_\mathrm{e}\right)h_\mathrm{i}
-    \end{equation}
-  \item Introduce the integral representation of
-    $\xi_\mathrm{i}$:
-    \begin{equation}
-      \bi{S}_\mathrm{e}(\partiali \xi) + \left(\frac{1}{2} -
-      \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}\sigma =
-      - \bi{S}_\mathrm{e}(\partiali h) - \left(\frac{1}{2} -
-      \bi{D}_\mathrm{e}\right)h_\mathrm{i}
-    \end{equation}
-  \item Reorganize Eq. \eqref{eq:first}, using the commutation
-    relation and the integral representation of $\xi_\mathrm{i}$:
-    \begin{equation}
-    \partiali \xi = \left(\frac{1}{2} + \bi{D}^\dagger_\mathrm{i}\right)\sigma
-    \end{equation}
-    such that:
-    \begin{equation}
-      \left[ \bi{S}_\mathrm{e}\left(\frac{1}{2} + \bi{D}^\dagger_\mathrm{i}\right)
-      +
-      \left(\frac{1}{2} - \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}
-      \right]\sigma =
-      - \bi{S}_\mathrm{e}(\partiali h) - \left(\frac{1}{2} -
-      \bi{D}_\mathrm{e}\right)h_\mathrm{i}
-    \end{equation}
-  \item Employ the \ac{DtN} map to simplify
-    the RHS. The \acs{DtN} map is derived by applying Eq. \eqref{eq:stat3} to
-    the Newton potential:
-    \begin{equation}
-      \phi(x) = (\bi{N}\rhoi)(x) = \int_{\mathbb{R}^3}\Gi\rhoi(y)\diff y
-      = \left.h(x)\right|_{\Omegai}
-    \end{equation}
-    which is equal, in $\Omegai$, to the auxiliary potential $h(x)$.
-    With this definition one has:
-    \begin{equation}
-      \frac{1}{2}\phi_\mathrm{i} =
-      \bi{S}_\mathrm{i}(\partiali \phi) -
-      \bi{D}_\mathrm{i}(\phi_\mathrm{i}) + \int_{\Omegai}\Gi\rhoi(y)\diff y
-      =
-      \bi{S}_\mathrm{i}(\partiali \phi) -
-      \bi{D}_\mathrm{i}(\phi_\mathrm{i}) + \phi_\mathrm{i}
-    \end{equation}
-    eventually leading to the \acs{DtN} map:
+which is equal, in $\Omegai$, to the auxiliary potential $h(\vect{r})$.
+The \acs{DtN} map is thus given as:
 \begin{equation}\label{eq:regular-DtN}
@@
 \end{equation}
-  \item Having identified $\phi_\mathrm{i} \equiv h_\mathrm{i}$
-    and $\partiali \phi \equiv \partiali h$ and employing the \acs{DtN} map,
-    Eq.\eqref{eq:regular-DtN}, one obtains the \acs{IEF}-\acs{PCM} equation:
-    \begin{equation}\label{eq:full-IEF}
-      \left[ \bi{S}_\mathrm{e}\left(\frac{1}{2} + \bi{D}^\dagger_\mathrm{i}\right)
-      +
-      \left(\frac{1}{2} - \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}
-      \right]\sigma =
-      -\left[\left(\frac{1}{2}-\bi{D}_\mathrm{e}\right)
-      -\bi{S}_\mathrm{e}\bi{S}_\mathrm{i}^{-1}
-       \left(\frac{1}{2}-\bi{D}_\mathrm{i}\right)
-      \right]\phi
-    \end{equation}
-    where the subscript $\mathrm{i}$ for the Newton potential $\phi$ has
-    been dropped, given that it is continuous across the boundary
-    $\Gamma$.
-\end{enumerate}
-The above proves the existence of a representation of the reaction
-potential valid \emph{inside} the cavity in terms of an apparent surface
-charge density. We do not prove the uniqueness of this solution and
-refer the interested reader to the original
-paper.~\autocite{Cances1998-og}
-
-In the code, the Green's functions are defined \emph{without} the $4\pi$
-factors in the denominator. Instead of $\frac{1}{2}$ as coefficients to
-the identity operator, one will then have $2\pi$:
-\begin{equation}\label{eq:full-IEF-2pi}
+With this last ingredient at hand, algebraic manipulations lead to the \acs{IEF}
+equation:
+\begin{equation}
   \left[ \bi{S}_\mathrm{e}\left(2\pi + \bi{D}^\dagger_\mathrm{i}\right)
@@
   \left(2\pi-\bi{D}_\mathrm{i}\right)
-  \right]\phi
+  \right]\esp,
+  \tag{\ref{eq:full-IEF} from Chapter \ref{ch:CSM}}
 \end{equation}
 
-From Eq. \eqref{eq:full-IEF-2pi} we can derive the isotropic
-\acs{IEF}, the \acs{DPCM} and the \acs{CPCM} (by letting
-$\diel\rightarrow\infty$) equations.
-For a homogeneous, isotropic dielectric with permittivity $\diel$ the
-Green's function outside the cavity is simply given by scaling the one
-inside:
+\section{Weak Formulation of Partial Differential Equations}\label{sec:weak}
+
+Let us now take a step back and formulate a different strategy for the
+solution of the transmission problem.
+We will follow the exposition of \citeauthor{Ern2004-oo} quite closely
+in introducing the \emph{weak} formulation of \acp{PDE}.
+For simplicity, we assume Dirichlet boundary conditions for a conductor,
+\ie the basic assumption behind \acs{COSMO}.
+The \emph{strong} formulation of the electrostatic problem then reads:
+\begin{equation}\label{eq:poisson-cosmo}
+  \nabla^2\esp = -4\pi\rho,
+  \quad \esp \in \mathcolor{ForestGreen}{\mathcal{C}^2_0(\Omegai)},
+\end{equation}
+where $\mathcal{C}^2_0(\Omegai)$ is the vector space of twice
+continuously differentiable functions in $\Omegai$ with null trace on
+$\Gamma$.
+We can relax this regularity requirement on $\Psi$ by introducing the
+\emph{Hilbert Sobolev space} of test functions $H_0^1(\Omegai)$:
 \begin{equation}
-  \Ge = \frac{1}{\diel|x-y|} = \frac{1}{\diel} \Gi = \frac{1}{|x-y|}
+  H_0^1(\Omegai) = \lbrace
+  f : \Omegai \rightarrow \mathbb{R}
+  | f, \nabla f \in L^2(\Omegai), f_\Gamma = 0
+  \rbrace.
 \end{equation}
-leading to the boundary integral operators:
-\begin{alignat}{2}
-  \bi{S}_\mathrm{i} = \diel\bi{S}_\mathrm{e} = \bi{S}, \quad&
-  \bi{D}_\mathrm{i} = \bi{D}_\mathrm{e} = \bi{D}
-\end{alignat}
-Eventually, one has:
-\begin{equation}\label{eq:isotropic-IEF}
-  \left[ 2\pi \left(\frac{\diel+1}{\diel-1}\right)
-  - \bi{D} \right]\bi{S}\sigma
-  =
-  -\left[2\pi-\bi{D}\right]\phi
+Projecting the differential problem onto this space and using the
+$\eta\nabla^2\gamma = \nabla\cdot(\eta\nabla\gamma) - \nabla\eta\cdot\nabla\gamma$ identity
+one obtains the \emph{weak} formulation of the differential problem:
+\begin{equation}\label{eq:poisson-weak}
+  \left\{
+  \begin{aligned}
+    &\text{Seek $\Psi \in \mathcolor{ForestGreen}{H_0^1(\Omegai)}$ such that:}\\
+    &\mathcolor{PMS1797}{(\nabla\zeta, \nabla\Psi)} =
+    \mathcolor{PMS2229}{-4\pi(\zeta, \rho)} \quad
+    \forall \zeta\in \mathcolor{PMS130}{H_0^1(\Omegai)}
+  \end{aligned}
+  \right.
 \end{equation}
+The form $(\nabla\cdot, \nabla\cdot)$ is \emph{bilinear} and continuous
+in $H_0^1(\Omegai)\times H_0^1(\Omegai)$, while $-4\pi(\cdot, \rho)$ is
+\emph{linear} and continuous in $H_0^1(\Omegai)$.
+We rewrite Eq.~\eqref{eq:poisson-weak} in the abstract form:
+\begin{equation}
+  \left\{
+  \begin{aligned}
+    &\text{Seek $u \in V$ such that:}\\
+    &\forall v \in V \,\,\,\,\,
+    a(u, v) = b(v)
+  \end{aligned}
+  \right.
+  \label{eq:weak}
+\end{equation}
+and state the following fundamental results:
+\begin{lemma}[Lax--Milgram]
+  If the bilinear form $a$ is \emph{continous} and \emph{coercive}
+  in $V$, then, for any continuous linear form $b$, Problem~\eqref{eq:weak} is
+  \emph{well-posed}.
+\end{lemma}
+\begin{corollary}[Variational property]
+  If the bilinear form is \emph{symmetric} and \emph{positive}
+  the unique solution to Problem \eqref{eq:weak} is
+  the \emph{unique minimum} on $V$ of the functional:
+  \[
+  \mathcal{F}(u) = \frac{1}{2}a(u, u) - b(u)
+  \]
+\end{corollary}
+
+\todo[inline]{To write or not to write these????}
+
+Weak formulation of \acs{BIE}s.\autocite{Hsiao2008-xb}
+
+Approximability, conformity, consistency and asymptotic consistency
+of the approximation setting, $\inf$-$\sup$ condition, Céa's Lemma
+and generalized $\inf$-$\sup$ condition.~\autocite{Ern2004-oo,
+TesiFilippo}

### 2016-08-31 Chapters/QuantumMechanics.tex
Work on CSM chapter

@@
 
-\section{A Brief Summary of Coupled Cluster Theory}\label{sec:coupled-cluster}
+\section{The Coupled Cluster Ansatz}\label{sec:coupled-cluster}
 
@@
 with quantum/classical polarizable Hamiltonians.
+
+\todo[inline]{Provide a primer here so that the Chapter on
+solvation+correlation is not superlong}
+
+\section{Many-Body Perturbation Theory}\label{sec:MBPT}
+
+\section{Iterative Approximations: CC2 and CC3}\label{sec:ccn}
+
+\section{Noniterative Triples Corrections}\label{sec:ccsd-t}

### 2016-08-31 Chapters/ContinuumSolvationModels.tex
Almost done with CSM chapter

@@
    \caption[Sparsity pattern achieved by the wavelet Galerkin \acs{BEM}.]{
-   Sparsity pattern achieved by the wavelet Galerkin \acs{BEM} for a
-   model problem (Laplace equation on 6 patches).
+   Sparsity pattern achieved by the wavelet Galerkin \acs{BEM} when
+   solving the Laplace equation on 6 patches.
    Reproduced from \noparcite[ref.][]{Bugeanu2015-tp}.
@@
 DefPol~\autocite{Pomelli1998-ob, Pomelli1999-dc} approaches of
-\citeauthor{Pomelli1998-ob}, the FIXPVA
+\citeauthor{Pomelli1998-ob} and the FIXPVA
 modification~\autocite{Su2009-em} to the GePol algorithm put forth by
@@
 
-\begin{equation}\label{eq:Y-operator}
-  \PCM = \bi{R}_\infty^{-1}\bi{R}_\diel\bi{S}
+The attentive reader will have noticed that, despite the focus of this
+thesis on continuum models for the quantum mechanical modelling of
+solvation, no mention has been made so far in this Chapter of any
+quantum aspect.
+The strategy usually adopted is to tailor a specific Hamiltonian, which
+includes classical contributions form the continuum and use it in the
+development of the quantum chemical model.
+Since the \acs{MEP} is a linear functional of the solute charge density,
+the \acs{ASC} itself incurs in a dependency on the solute charge
+density.
+Properly accounting for the mutual solute-solvent polarization induces a
+\emph{nonlinearity} in an otherwise linear problem.
+Manipulating such a \emph{nonlinear} Schr\"{o}dinger equation is
+theoretically cumbersome,\autocite{Sanhueza1979-bp, Heimsoeth1990-ki}
+but has been successfully exploited in the past.\autocite{Tomasi2005-vm}
+
+In this thesis we advocate a different theoretical approach: the
+variational formulation of classical polarizable models.
+Why so? In such a framework the \acs{ASC} becomes an additional,
+independent, variationally optimizable degree of freedom.
+This makes the coupling to, \eg, extended Lagrangian dynamics
+trivial.\autocite{Car1985-jw}
+It is well known that electrodynamics can be derived as the minimization
+of an action functional.\autocite{Jackson1999}
+The principle of least action is however formulated in terms of
+\emph{fields} and not more easily manageable quantities, such as the
+\acs{MEP} or the \acs{ASC}.
+A true energy functional has to fulfill the following
+properties:\autocite{Solis2013-ef}
+\begin{enumerate*}[label={\alph*)},font={\color{PMS1797}}]
+    \item physical extremal points,
+    \item equilibrium values are true energies \label{item:equil-prop} and
+    \item convexity (positive-definiteness)
+\end{enumerate*}
+\citeauthor{Allen2001-fp} and \citeauthor{Attard2003-vr} pioneering
+attempts resulted in functionals of scalar densities violating
+Property \ref{item:equil-prop} and could not thus be used in extended Lagrangian
+dynamics.\autocite{Allen2001-fp, Attard2003-vr, Attard2007-lk}
+
+In a seminal paper, \citeauthor{Lipparini2010-be} showed how the
+\acs{PCM} can be recast in a variational fashion yielding a true energy
+functional of the \acs{ASC}:\autocite{Lipparini2010-be, Lipparini2011-aj, Lipparini2016-mo}
+\begin{equation}\label{eq:pcm-functional}
+ U_\mathrm{PCM} = \frac{1}{2}\sigma\PCM\sigma + \esp\sigma,
+ \quad
+  \PCM \coloneq \bi{R}_\infty^{-1}\bi{R}_\diel\bi{S}.
 \end{equation}
+yielding Eq.~\eqref{eq:isotropic-IEF} upon minimization:
+\begin{equation}\label{eq:pcm-stationarity}
+  \pderiv{U_\mathrm{PCM}}{\sigma} = \PCM\sigma + \varphi = 0.
+\end{equation}
+It is interesting to note how this functional follows quite immediately
+from the integral equation formulation and standard results in the
+theory of boundary integral equations.\autocite{Hsiao2008-xb}
 
-Variational functional for implicit models, \acs{PCM} in this
-case:~\autocite{Lipparini2010-be, Lipparini2011-aj, Lipparini2016-mo}
-\begin{equation}
- U_\mathrm{PCM} = \frac{1}{2}\sigma\PCM\sigma + \esp\sigma
+Explicit classical polarizable models can also be recast in a
+variational fashion allowing for their treatment on par with implicit
+classical polarizable models.
+The functional looks as follows:
+\begin{equation}\label{eq:mm-functional}
+  U_\mathrm{MM} = \frac{1}{2}\kappa\MM\kappa + \kappa\zeta,
 \end{equation}
+here $\kappa$ is the variational degree of freedom representing
+the polarization of the explicit environment. $\zeta$ is the inducing
+field, \emph{vide infra}. The reader is referred to the works of
+\citeauthor{Lipparini2011-rd} for a derivation of this functional.\autocite{Lipparini2011-rd, Lipparini2015-lq, Loco2016-oy}
+We have made no specific reference to the particular explicit
+model used. The variational framework is flexible enough to encompass
+either the MMpol,\autocite{Mennucci2013-go}
+the \ac{PE},\autocite{Olsen2010-wa, Olsen2011-io} or
+the \ac{FQ} model.\autocite{Rick1994-mn, Rick1995-wu, Rick1996-om,
+Lipparini2011-rd}
 
-Variational functional for explicit models,~\autocite{Lipparini} could be
-MMpol~\autocite{Curutchet} or PE~\autocite{Steindal, Kongsted, Olsen} or
-FQ~\autocite{Rick1994-mn, Rick1995-wu, Rick1996-om, Lipparini}:
-\begin{equation}
-  U_\mathrm{MM} = \frac{1}{2}\kappa\MM\kappa + \kappa\zeta
-\end{equation}
-
-Coupled implicit/explicit polarization energy
-functional:~\autocite{Lipparini}
+Coupling the implicit and explicit polarizable models is straightforward
+in the variational framework:\autocite{Steindal2011-ki, Lipparini2011-rd,
+Caprasecca2012-ir, Lipparini2013-ud}
 \begin{equation}\label{eq:pcm-mm-functional}
@@
  + \frac{1}{2}\kappa\MM\kappa + \kappa\zeta
- + \sigma\bi{X}\kappa
+ + \sigma\bi{X}\kappa.
 \end{equation}
-where one has:
+An additional term, as first derived by
+\citeauthor{Steindal2011-ki},\autocite{Steindal2011-ki} was added to
+account for the mutual polarization between the implicit and explicit
+layers:
 \begin{equation}
@@
 \end{equation}
-and $\MMPCM$ is the implicit/explicit interaction kernel. This is
-charge/dipole or charge/charge interaction kernel for the MMpol and
-\ac{PE} models
-or the \ac{FQ} model, respectively.
+$\MMPCM$ is the implicit/explicit interaction kernel, whose form depends
+on the specific explicit model chosen.
+
 The global minimum of the convex functional is found by differentiating
-it with respect to the variational degrees of freedom, \ie{} the
-\acl{ASC} (\acs{ASC}) density $\sigma$ and the classical point
-charges/dipoles $\kappa$. This leads to the coupled equations:
-\begin{align}
-  \PCM\sigma + \esp + \MMPCM\kappa &= 0 \\
-  \MM\kappa  + \zeta + \MMPCM^\dagger\sigma &= 0
-\end{align}
-or in the commonly found matrix form:
-\begin{equation}
+with respect to the variational degrees of freedom:
+\begin{equation}\label{eq:pcm-mm-stationarity}
   \begin{pmatrix}
@@
 \end{equation}
+While $\esp$ is quite clearly the \acs{MEP}, $\zeta$ can either be the
+molecular electric field (MMpol and \acs{PE} models) or again the
+\acs{MEP} (\acs{FQ} model).
+In any case, both will be determined by the quantum mechanical molecular
+charge density and can thus be formulated as expectation values of
+one-electron operators.
+Eventually, this achieves the coupling between the classical -- $\sigma$
+and $\kappa$ -- and the quantum mechanical variational degrees of
+freedom.
+
+The functionals in Eqs.~\eqref{eq:pcm-functional},
+\eqref{eq:mm-functional} and \eqref{eq:pcm-mm-functional} all have a
+clear physical meaning.
+They describe charge distributions interacting (unfavorably) with
+themselves and (favorably) with their respective inducing external
+fields and constitutes the polarization energy of the medium.
+The topic of variational formulations is particularly important in the
+field of \acs{MD} simulations.\autocite{Jadhao2012-gf, Jadhao2013-ry, Jadhao2013-hs,
+Solis2013-ef}
+Our focus however is on the quantum mechanical calculations of molecular
+properties and the variational formulation is advantageous since it
+dispenses us from manipulating cumbersome nonlinear coupling terms in
+the Hamiltonian.
+Moreover, the stationarity condition Eq.~\ref{eq:pcm-mm-stationarity}
+entails the existence of a classical Hellmann--Feynman theorem:
+\begin{equation}\label{eq:classical-HF}
+  \deriv{U_\mathrm{pol}}{\lambda} = \pderiv{U_\mathrm{pol}}{\lambda}
+  + \pderiv{U_\mathrm{pol}}{\sigma}\pderiv{\sigma}{\lambda}
+  + \pderiv{U_\mathrm{pol}}{\kappa}\pderiv{\kappa}{\lambda}
+  = \pderiv{U_\mathrm{pol}}{\lambda},
+\end{equation}
+which is paramount in the formulation of molecular properties, see
+Chapter \ref{ch:molprop}.
+The variational formulation also lends itself to an implementation
+in terms of simultaneous optimization.\autocite{Lipparini2011-aj}
+
 Finally, let us re-express the equations above in a "supermatrix"
@@
 derivations.
-
-\todo[inline]{PUT REFERENCES HERE!!!}
-The variational formulation offers a series of advantages.
-The classical energy functional:
-\begin{equation}
-  U_\mathrm{pol} =
-   \frac{1}{2}\sigma\PCM\sigma + \sigma\esp
- + \frac{1}{2}\kappa\MM\kappa + \kappa\zeta
- + \sigma\bi{X}\kappa
-\end{equation}
-renders itself to a straightforward physical interpretation. In the
-three bilinear terms, \ie{} the ones mediated by an interaction
-operator, we can identify the self-interaction of the classical
-polarization with itself, be it purely implicit:
-$\frac{1}{2}\sigma\PCM\sigma$, purely explicit
-$\frac{1}{2}\kappa\MM\kappa$ or mixed $\sigma\MMPCM\kappa$.
-These terms are positive definite and give rise to an unfavorable
-contribution, which is counterbalanced by the linear terms. These terms
-mediate the interaction between the induced polarization and the
-inducing fields: $\esp$ and $\zeta$. While $\esp$ is quite clearly the
-\acl{MEP} (\acs{MEP}), $\zeta$ can either be the molecular electric
-field (MMpol and \acs{PE} models) or again the \acs{MEP} (\acs{FQ} model).
-In any case, both will be determined by the quantum mechanical molecular
-charge density and can thus be formulated as expectation values of
-one-electron operators. Eventually, this achieves the coupling between
-the classical -- $\sigma$ and $\kappa$ -- and
-the quantum mechanical variational degrees of freedom.
-
-\todo[inline]{Mention the classical equivalent of the Hellmann--Feynman
-theorem and why it is particularly useful in formulating response
-theory.}

### 2016-09-01 Chapters/ContinuumSolvationModels.tex
Switch to glossaries package for acronyms

@@
 
-Developments of the \ac{PCM}, a continuum solvation model, are at the
+Developments of the \gls{PCM}, a continuum solvation model, are at the
 heart of this thesis.
-This Chapter is a brief exposition of the major points of the \acs{PCM}
+This Chapter is a brief exposition of the major points of the \acrshort{PCM}
 we have worked upon in the thesis.
@@
 Section \ref{sec:BEM} will be devoted to the numerical strategies used
-in solving the \acs{PCM} equation.
+in solving the \acrshort{PCM} equation.
 
@@
 \citeauthor{Onsager1936-wf}'s model\autocite{Onsager1936-wf} appeared in
-1936 and the first form of the \acs{PCM} entered the stage in
+1936 and the first form of the \acrshort{PCM} entered the stage in
 1981.~\autocite{Miertus1981-mm}
@@
 Still nowadays much research activity is expended on the
-\acs{PCM}.~\autocite{Tomasi2011-un, Mennucci2012-dv, Lipparini2016-mo}
+\acrshort{PCM}.~\autocite{Tomasi2011-un, Mennucci2012-dv, Lipparini2016-mo}
 I will point out throughout the Chapter to developments outside our
@@
 The basic idea of implicit models is that the environment can
-be replaced by structureless continuum. The continuum will interact
-\emph{classically} with the explicit part of the multiscale model, an
-interaction mediated by some macroscopic property of the bulk material
-the continuum represents in the model.
+be replaced by a structureless continuum.
+The continuum will interact \emph{classically} with the explicit part of
+the multiscale model, an interaction mediated by some macroscopic
+property of the bulk material the continuum represents.
 Thus, continuum models bypass both problems plaguing explicit models we
@@
 \item the statistical averaging of environment configurations and
-\item the \acs{MM} region electrostatics cutoff choice.
+\item the \acrshort{MM} region electrostatics cutoff choice.
 \end{enumerate*}
@@
 Following \citeauthor{Tomasi2007-es}, we introduce the complete system
-Hamiltonian in the \acs{BO} approximation:~\autocite{Tomasi2004-dc,
+Hamiltonian in the \acrshort{BO} approximation:~\autocite{Tomasi2004-dc,
 Tomasi2007-es}
@@
 classical counterparts and obtain the quantum/classical, possibly
-\emph{polarizable}, Hamiltonian for an explicit \acs{QM}/\acs{MM} model.
+\emph{polarizable}, Hamiltonian for an explicit \acrshort{QM}/\acrshort{MM} model.
 This brings about the first important point: the need for statistical
@@
 
-The physical setting of the \ac{PCM} is depicted in Figure \ref{fig:IEF}.
+The physical setting of the \gls{PCM} is depicted in Figure \ref{fig:IEF}.
 The molecular solute, represented by its charge density
@@
 \emph{boundary integral operators}, \ie transform the \acsp{PDE} into a
-\ac{BIE}.
+\gls{BIE}.
 
@@
   \caption[The physical setting of the polarizable continuum model.]{
-  The physical setting of the \ac{PCM}. The molecular solute,
+  The physical setting of the \gls{PCM}. The molecular solute,
   represented by its, possibly quantum mechanical, charge density
@@
 \end{equation}
-where $\sigma$ is the \ac{ASC}, the key quantity in the \acs{PCM}.
-As shown by \citeauthor{Cances1998-og}, the \acs{ASC} can be computed
+where $\sigma$ is the \gls{ASC}, the key quantity in the \acrshort{PCM}.
+As shown by \citeauthor{Cances1998-og}, the \acrshort{ASC} can be computed
 as the \emph{unique} solution to the following integral equation:
@@
 \end{equation}
-commonly called the \ac{IEF} equation.
+commonly called the \gls{IEF} equation.
 The Green's functions for the elliptic differential operators
 assume a central role in the boundary integral formulation and make the
-\acs{IEF} equation valid for a wide range of external
+\acrshort{IEF} equation valid for a wide range of external
 environments: homogeneous isotropic, ionic and anisotropic
@@
 \end{alignat}
-yielding the isotropic \acs{IEF} equation:
+yielding the isotropic \acrshort{IEF} equation:
 \begin{equation}\label{eq:isotropic-IEF}
@@
 where conductor boundary conditions are imposed. This leads to the
-well-known \ac{COSMO} equation:~\autocite{Klamt1993-mj, Cossi2003-xe}
+well-known \gls{COSMO} equation:~\autocite{Klamt1993-mj, Cossi2003-xe}
 \begin{equation}
@@
 
-The solution of the \acs{IEF}-\acs{PCM} boundary integral equation can
+The solution of the \acrshort{IEF}-\acrshort{PCM} boundary integral equation can
 only be achieved analytically for simple cavity geometries, such as
@@
 When more general molecular cavities are employed one has to resort to a
-numerical technique: the \ac{BEM}. This is implemented in four steps:
+numerical technique: the \gls{BEM}. This is implemented in four steps:
 \begin{enumerate*}[label={\alph*)},font={\color{PMS1797}}]
@@
 Many different approaches to these four steps have been presented, both
-in the field of the \acs{BEM} and in the context of the \acs{PCM}.
+in the field of the \acrshort{BEM} and in the context of the \acrshort{PCM}.
 
@@
   \item
-    the union of the interlocking spheres surfaces is the \ac{vdWS},
+    the union of the interlocking spheres surfaces is the \gls{vdWS},
     Figure \ref{fig:vdWS}
   \item the locus of points defined by the center of the solvent
-    spherical probe while rolling over the \acs{vdWS} is the
-    \ac{SAS}, Figure \ref{fig:SAS}
+    spherical probe while rolling over the \acrshort{vdWS} is the
+    \gls{SAS}, Figure \ref{fig:SAS}
   \item the locus of points defined by the contact point of the solvent
-    spherical probe while rolling over the \acs{vdWS} is the \ac{SES}, Figure \ref{fig:SES}.
+    spherical probe while rolling over the \acrshort{vdWS} is the \gls{SES}, Figure \ref{fig:SES}.
 \end{enumerate*}
@@
    \includegraphics[width=.85\textwidth]{vdWS.png}
-   \caption{\acl{vdWS} (\acs{vdWS})}
+   \caption{\acrlong*{vdWS} (\acrshort{vdWS})}
    \label{fig:vdWS}
@@
    \includegraphics[width=.85\textwidth]{SAS.png}
-   \caption{\acl{SAS} (\acs{SAS})}
+   \caption{\acrlong*{SAS} (\acrshort{SAS})}
    \label{fig:SAS}
@@
    \includegraphics[width=.85\textwidth]{SES.png}
-   \caption{\acl{SES} (\acs{SES})}
+   \caption{\acrlong*{SES} (\acrshort{SES})}
    \label{fig:SES}
@@
 Steps \ref{item:mesh} and \ref{item:discretize} are closely intertwined,
-in the technical \acs{BEM} literature they go under the name of
+in the technical \acrshort{BEM} literature they go under the name of
 \emph{finite element discretization}.
@@
 and the mesh generators of Harbrecht and Randrianarivony especially
-tailored for the wavelet \acs{BEM}.~\autocite{Harbrecht2009-no,
+tailored for the wavelet \acrshort{BEM}.~\autocite{Harbrecht2009-no,
 Harbrecht2011-dk}
@@
 symmetrization procedures introducing, however, a degree of
-arbitrariness in the \acs{BEM} procedure.~\autocite{Lange2011-eu}
+arbitrariness in the \acrshort{BEM} procedure.~\autocite{Lange2011-eu}
 
@@
 $\vect{q}$ and $\vect{v}$ are vectors of dimension $N_\mathrm{ts}$
-collecting the values of the \acs{ASC} and \ac{MEP}
+collecting the values of the \acrshort{ASC} and \gls{MEP}
 at the collocation points. $\mat{R}^{-1}\mat{T}$ is the stiffness
@@
 
-The wavelet Galerkin \acs{BEM} was also used in this thesis. This method
+The wavelet Galerkin \acrshort{BEM} was also used in this thesis. This method
 preserves the fundamental symmetries of the boundary integral operators
@@
 spanned by the wavelet basis.
-The fundamental insight in the wavelet \acs{BEM} is to use the wavelet
+The fundamental insight in the wavelet \acrshort{BEM} is to use the wavelet
 basis for the discretization step and employ the compression technique
 described by \citeauthor{Dahmen2006-pj} to build up the sparse system
-matrix.~\autocite{Harbrecht2004-uo, Harbrecht2004-ug, Dahmen2006-pj}
+matrix.~\autocite{Harbrecht2004-uo, Harbrecht2006-ug, Dahmen2006-pj}
 Matrix compression is carried out in two rounds, \emph{a priori} and \emph{a
@@
    \includegraphics[width=\textwidth]{a_2_l_0_5_fs.png}
-   \caption[Sparsity pattern achieved by the wavelet Galerkin \acs{BEM}.]{
-   Sparsity pattern achieved by the wavelet Galerkin \acs{BEM} when
+   \caption[Sparsity pattern achieved by the wavelet Galerkin \acrshort{BEM}.]{
+   Sparsity pattern achieved by the wavelet Galerkin \acrshort{BEM} when
    solving the Laplace equation on 6 patches.
@@
 \citeauthor{Su2009-em}.
-\citeauthor{Scalmani2010-tw} have recently proposed the \ac{CSC} scheme
-which uses a smooth \acs{vdWS} and 3D spherical Gaussian basis
+\citeauthor{Scalmani2010-tw} have recently proposed the \gls{CSC} scheme
+which uses a smooth \acrshort{vdWS} and 3D spherical Gaussian basis
 functions.~\autocite{Scalmani2010-tw, York1999-xy}
 Gaussian blurring was also advocated by \citeauthor{Lange2010-jp} in
-their \ac{SWIG} method.~\autocite{Lange2010-jp, Lange2010-qo}
-Both \acs{CSC} and \acs{SWIG} achieve continuous \acsp{PES} and smooth
+their \gls{SWIG} method.~\autocite{Lange2010-jp, Lange2010-qo}
+Both \acrshort{CSC} and \acrshort{SWIG} achieve continuous \acsp{PES} and smooth
 molecular gradients. However, their convergence to the exact solution of
@@
 method, guaranteed to converge to the exact solution.~\autocite{Quarteroni1999-jt}
-Moreover, for \acs{COSMO}, the formulation is embarrassingly
+Moreover, for \acrshort{COSMO}, the formulation is embarrassingly
 parallel and has been implemented in a linearly scaling fashion.~\autocite{Cances2013-jg, Lipparini2013-cy, Lipparini2014-to,
@@
 
-In the following, the \acs{PCM} equations will be written in the
+In the following, the \acrshort{PCM} equations will be written in the
 "complete basis" meaning that we will be working with the exact integral
@@
   =
-  \scalprod{\sigma(\vect{s})}{\PCM\sigma(\vect{s})} \\
+  \scalprod{\sigma}{\PCM\sigma} \\
   \sigma\esp &=
@@
   =
-  \scalprod{\sigma(\vect{s})}{\esp(\vect{s})}
+  \scalprod{\sigma}{\esp}
   \end{aligned}
@@
 \section[Variational Formulation of Classical Polarizable Models]{
-Above and Beyond: Classical Polarizable Models within a Variational Formulation}
+Variational Formulation of Classical Polarizable Models}
 \label{sec:variational}
@@
 development of the quantum chemical model.
-Since the \acs{MEP} is a linear functional of the solute charge density,
-the \acs{ASC} itself incurs in a dependency on the solute charge
+Since the \acrshort{MEP} is a linear functional of the solute charge density,
+the \acrshort{ASC} itself incurs in a dependency on the solute charge
 density.
@@
 variational formulation of classical polarizable models.
-Why so? In such a framework the \acs{ASC} becomes an additional,
+Why so? In such a framework the \acrshort{ASC} becomes an additional,
 independent, variationally optimizable degree of freedom.
@@
 \emph{fields} and not more easily manageable quantities, such as the
-\acs{MEP} or the \acs{ASC}.
+\acrshort{MEP} or the \acrshort{ASC}.
 A true energy functional has to fulfill the following
@@
 In a seminal paper, \citeauthor{Lipparini2010-be} showed how the
-\acs{PCM} can be recast in a variational fashion yielding a true energy
-functional of the \acs{ASC}:\autocite{Lipparini2010-be, Lipparini2011-aj, Lipparini2016-mo}
+\acrshort{PCM} can be recast in a variational fashion yielding a true energy
+functional of the \acrshort{ASC}:\autocite{Lipparini2010-be, Lipparini2011-aj, Lipparini2016-mo}
 \begin{equation}\label{eq:pcm-functional}
@@
 either the MMpol,\autocite{Mennucci2013-go}
-the \ac{PE},\autocite{Olsen2010-wa, Olsen2011-io} or
-the \ac{FQ} model.\autocite{Rick1994-mn, Rick1995-wu, Rick1996-om,
+the \gls{PE},\autocite{Olsen2010-wa, Olsen2011-io} or
+the \gls{FQ} model.\autocite{Rick1994-mn, Rick1995-wu, Rick1996-om,
 Lipparini2011-rd}
@@
 \end{equation}
-While $\esp$ is quite clearly the \acs{MEP}, $\zeta$ can either be the
-molecular electric field (MMpol and \acs{PE} models) or again the
-\acs{MEP} (\acs{FQ} model).
+While $\esp$ is quite clearly the \acrshort{MEP}, $\zeta$ can either be the
+molecular electric field (MMpol and \acrshort{PE} models) or again the
+\acrshort{MEP} (\acrshort{FQ} model).
 In any case, both will be determined by the quantum mechanical molecular
@@
 The topic of variational formulations is particularly important in the
-field of \acs{MD} simulations.\autocite{Jadhao2012-gf, Jadhao2013-ry, Jadhao2013-hs,
+field of \acrshort{MD} simulations.\autocite{Jadhao2012-gf, Jadhao2013-ry, Jadhao2013-hs,
 Solis2013-ef}
@@
 
-Finally, let us re-express the equations above in a "supermatrix"
+Finally, let us re-express the equations above in a supermatrix
 formalism:
@@
 and the ${}^t\p$ symbol denotes the transposed supervector $\p$.
-The supermatrix form will prove particularly useful in the following
-derivations.
+The scalar products are understood to be in the relevant function
+spaces.
+The supermatrix formalism will be adopted throughout the thesis.
+
+How are quantum/classical polarizable Hamiltonians implemented in this
+framework?
+As shown in Chapter \ref{ch:QM} for the \acrshort{CC} method, one can
+formulate methods in quantum chemistry as the unconstrained minimization
+of a Lagrangian.\autocite{Helgaker1982-xl, Arponen1983-qt,
+Helgaker1988-to, Helgaker1989-wl, Koch1990-vr, Kvaal2013-jr}
+Let us call $\mathcal{E}(\vect{\eta})$ the definition of the energy
+for the given quantum chemical method.
+The \emph{amplitudes} $\vect{\eta}$ are determined by a set of
+stationarity conditions $\mat{\Omega}(\vect{\eta}) = 0$.
+Direct minimization of $\mathcal{E}(\vect{\eta})$ is thus a
+\emph{constrained} optimization.
+Introducing a set of Lagrange \emph{multipliers} $\bar{\vect{\eta}}$
+lets us recast the problem as an unconstrained minimization of the
+Lagrangian:
+\begin{equation}
+  \mathcal{L}(\vect{\eta}, \bar{\vect{\eta}}) =
+  \mathcal{E}(\vect{\eta}) +
+  \scalprod[W]{\bar{\vect{\eta}}}{\mat{\Omega}(\vect{\eta})}
+\end{equation}
+where the governing equations are obtained by differentiation with
+respect to both sets of parameters:
+\begin{subequations}
+  \begin{align}
+    \pderiv{\mathcal{L}(\vect{\eta}, \bar{\vect{\eta}})}{\bar{\vect{\eta}}}
+    &= \mat{\Omega}(\vect{\eta}) = 0 \\
+    \pderiv{\mathcal{L}(\vect{\eta}, \bar{\vect{\eta}})}{\vect{\eta}}
+    &= \pderiv{\mathcal{E}(\vect{\eta})}{\vect{\eta}} +
+    \scalprod[W]{\bar{\vect{\eta}}}{\pderiv{\mat{\Omega(\vect{\eta})}}{\vect{\eta}}}
+    = 0
+  \end{align}
+\end{subequations}
+Due to the  variational property of the classical polarizable functional
+one can devise an \emph{effective} quantum/classical polarizable
+Lagrangian:\autocite{Lipparini2016-mo}
+\begin{equation}
+  \mathcal{L}_\mathrm{eff}(\vect{\eta}, \bar{\vect{\eta}}, \p) =
+  \mathcal{L}(\vect{\eta}, \bar{\vect{\eta}}) +
+  \frac{1}{2}{}^t\p\V\p + {}^t\p\s(\vect{\eta}, \bar{\vect{\eta}})
+\end{equation}
+and derive the governing equations in the usual manner:
+\begin{subequations}
+  \begin{align}
+    \pderiv{\mathcal{L}_\mathrm{eff}(\vect{\eta}, \bar{\vect{\eta}}, \p)}{\bar{\vect{\eta}}}
+    &= \mat{\Omega}(\vect{\eta}) + {}^t\p\pderiv{\s(\vect{\eta}, \bar{\vect{\eta}})}{\bar{\vect{\eta}}} = 0 \\
+    \pderiv{\mathcal{L}_\mathrm{eff}(\vect{\eta}, \bar{\vect{\eta}}, \p)}{\vect{\eta}}
+    &= \pderiv{\mathcal{E}(\vect{\eta})}{\vect{\eta}} +
+    \scalprod[W]{\bar{\vect{\eta}}}{\pderiv{\mat{\Omega(\vect{\eta})}}{\vect{\eta}}}
+    + {}^t\p\pderiv{\s(\vect{\eta}, \bar{\vect{\eta}})}{\vect{\eta}}
+    = 0 \\
+    \pderiv{\mathcal{L}_\mathrm{eff}(\vect{\eta}, \bar{\vect{\eta}}, \p)}{\p} &=
+    \V\p + \s(\vect{\eta}, \bar{\vect{\eta}}) = 0
+  \end{align}
+\end{subequations}
+The quantum mechanical and the classical model are thus fully coupled,
+due to the dependence of the source term $\s$ on the amplitudes and
+multipliers.

### 2016-09-01 Chapters/ContributionsSummary.tex
Switch to glossaries package for acronyms

@@
 In Section \ref{sec:software} I will describe the principles we have
-striven to follow in developing our \acs{PCM} software library and its
+striven to follow in developing our \acrshort{PCM} software library and its
 management philosophy.
@@
 New features are isolated into libraries that can be accessed by host
-programs through a well-defined \ac{API}.
+programs through a well-defined \gls{API}.
 
@@
 the particular host program in which they will be used.
-A well-defined \ac{API} clearly delimits the boundaries of the functionality
+A well-defined \gls{API} clearly delimits the boundaries of the functionality
 offered, thus forcing a programming style and standardisation of the
@@
 
-The \acs{PCM} is an ideal candidate for the creation of a solvation
-\acs{API}. Consider for example the schematic representation of a
-\acs{PCM}-\acs{SCF} algorithm given in Figure \ref{fig:algorithm}.
-The input to and from the \acs{PCM} library is limited and well-defined,
+The \acrshort{PCM} is an ideal candidate for the creation of a solvation
+\acrshort{API}. Consider for example the schematic representation of a
+\acrshort{PCM}-\acrshort{SCF} algorithm given in Figure \ref{fig:algorithm}.
+The input to and from the \acrshort{PCM} library is limited and well-defined,
 as are the components that host quantum chemistry codes need to
 implement.
-This provides a natural \acs{API} design: the \acs{API} functions can be
+This provides a natural \acrshort{API} design: the \acrshort{API} functions can be
 compared \emph{vis-à-vis} with the working equations derived for the
@@
 \scalebox{0.7}{\input{gfx/algorithm.tikz}}
-\caption[Modular approach to programming a \acs{PCM} functionality into an existing \acs{SCF} code.]{
-Schematic view of a \acs{PCM}-\acs{SCF} algorithm. Computations/data in
-\textcolor{PMS2229}{blue} are implemented by the \acs{PCM} \acs{API}.
+\caption[Modular approach to programming a \acrshort{PCM} functionality into an existing \acrshort{SCF} code.]{
+Schematic view of a \acrshort{PCM}-\acrshort{SCF} algorithm. Computations/data in
+\textcolor{PMS2229}{blue} are implemented by the \acrshort{PCM} \acrshort{API}.
 Computations/data in \textcolor{Green}{green} are implemented by the
@@
 The \pcmsolver library was developed to fulfill these requirements and
-offer the community an easy-to-use implementation of the \acs{PCM}.
+offer the community an easy-to-use implementation of the \acrshort{PCM}.
 \pcmsolver is written in C++ with Fortran, C and Python components.
@@
 Evaluation of the Green's functions, needed to set up the representation
-of the boundary integral operators, leverages \ac{AD}. We use the Taylor library, a template C++
-implementation of \acs{AD}.~\autocite{Bartholomew-Biggs2000-db, libtaylor}
-The adoption of \git as \acs{DVCS}, \cmake for cross-platform builds and
+of the boundary integral operators, leverages \gls{AD}. We use the Taylor library, a template C++
+implementation of \acrshort{AD}.~\autocite{Bartholomew-Biggs2000-db, libtaylor}
+The adoption of \git as \acrshort{DVCS}, \cmake for cross-platform builds and
 automatic documentation deployment on \readthedocs simplify
@@
 The library is released under the terms of the
-version 3 \ac{LGPL}, a standard open-source license.~\autocite{LGPLv3}
+version 3 \gls{LGPL}, a standard open-source license.~\autocite{LGPLv3}
 We strongly believe that open-source software is the key to a larger
@@
 thresholds to adoption and third-party contributions.~\autocite{LGPLv3}
-The library serves as our development platform for the \acs{PCM}.
+The library serves as our development platform for the \acrshort{PCM}.
 The interfaces to different quantum chemistry codes allow us to explore
 new methodologies, significantly cutting down development times.
-Use of our \acs{API} \emph{significantly} limits coding effort on the
-side of the host: continuum solvation at the \acs{SCF} level of theory
+Use of our \acrshort{API} \emph{significantly} limits coding effort on the
+side of the host: continuum solvation at the \acrshort{SCF} level of theory
 can be implemented in the host program almost out-of-the-box.
 
-The library was designed to allow host-\acs{API} communication through
+The library was designed to allow host-\acrshort{API} communication through
 \emph{pure} functions, \ie{} free from side-effects.
 Such an implementation hinges on the theoretical realization that the
-\acs{PCM} layer is independent of the \acs{AO} or \acs{MO} spaces
+\acrshort{PCM} layer is independent of the \acrshort{AO} or \acrshort{MO} spaces
 defined in the quantum chemical layer.
@@
 As schematically depicted in Figure \ref{fig:algorithm},
-computation of the \acs{ASC} only depends on
+computation of the \acrshort{ASC} only depends on
 the electrostatic potential sampled at the cavity boundary:
@@
 $\scalprod{\sigma(\vect{s})}{\esp_{\kappa\lambda}(\vect{s})}$
-between the \acs{ASC} and the charge attraction integral matrix, again
+between the \acrshort{ASC} and the charge attraction integral matrix, again
 a task easily accomplished by the host program.
@@
 
-The \acs{API} is implemented in ISO C99 as a \emph{context-aware} set of
+The \acrshort{API} is implemented in ISO C99 as a \emph{context-aware} set of
 functions.~\autocite{library-ronacher, context-api-example}
 Upon initialization of the library, a \emph{context} object containing
-the state of the \acs{PCM} \acs{API} is constructed and a handle to it
+the state of the \acrshort{PCM} \acrshort{API} is constructed and a handle to it
 (a pointer in C++ parlance) is returned to the host program. Further
@@
 Listings \ref{lst:C-host} and \ref{lst:Fortran-host} show how calls to
-\pcmsolver \acs{API} functions might look like in an actual C or Fortran
+\pcmsolver \acrshort{API} functions might look like in an actual C or Fortran
 program, respectively.
@@
 
-A context-aware \acs{API} offers a simple strategy for parallelization.
+A context-aware \acrshort{API} offers a simple strategy for parallelization.
 The host program could spawn contexts on a per-process basis,
@@
 functions for spherical diffuse interfaces~\autocite{DiRemigio2016-nn} and
-the delayed solvers for real-time \acl{TDSCF}
+the delayed solvers for real-time \acrlong*{TDSCF}
 simulations.~\autocite{Corni2015-pe}
@@
 Coupling with \psicode~\autocite{Turney2012-de} has allowed us to implement the
-\acs{PCM}-\acs{CC} developments described in this thesis.
+\acrshort{PCM}-\acrshort{CC} developments described in this thesis.
 Finally, development of an interface with the KOALA code is underway.~\autocite{Hofener2014-ex, Hofener2016-qz}
@@
 approximation of these effects.
-In this paper, we presented the first derivation and implementation of the \acs{PCM}
-coupled to a \acs{SCF} description of the solute in the fully relativistic, four-component
+In this paper, we presented the first derivation and implementation of the \acrshort{PCM}
+coupled to a \acrshort{SCF} description of the solute in the fully relativistic, four-component
 regime.
@@
 programming strategy. We showed that it is entirely possible to adopt the same
-\acs{PCM} code and implementation "checkpoints" across altogether different
+\acrshort{PCM} code and implementation "checkpoints" across altogether different
 problem domains, Fig.~\ref{fig:algorithm}.
 
-As a by-product of the four-component implementation, we were able to obtain and visualize \acs{MEP} maps
-from four-component \acs{SCF} wave functions. These add yet another interpretive tool to the toolbox available
+As a by-product of the four-component implementation, we were able to obtain and visualize \acrshort{MEP} maps
+from four-component \acrshort{SCF} wave functions. These add yet another interpretive tool to the toolbox available
 in the four-component relativistic regime.
 The interface to \DIRAC was later released in the 2014 and later versions,
-providing \acs{PCM} capabilities to the software package.
+providing \acrshort{PCM} capabilities to the software package.
 
 I contributed the theoretical derivation of the quantum/classical polarizable
-terms in a four-component \acs{SCF} framework, for energies and linear response
+terms in a four-component \acrshort{SCF} framework, for energies and linear response
 properties. I devised the coupling of the four-component program \DIRAC with
@@
 \begin{enumerate*}[label={\alph*)},font={\color{PMS1797}}]
-  \item \acs{MEP} integrals for four-component wave functions,
+  \item \acrshort{MEP} integrals for four-component wave functions,
   \item the additional Fock matrix contributions and
@@
 in determining the performance of the models.
-However, the numerical accuracy of the \acs{BEM} procedure used to numerically solve the
-underlying \acs{BIE} is a not-so-often studied aspect of these models.
+However, the numerical accuracy of the \acrshort{BEM} procedure used to numerically solve the
+underlying \acrshort{BIE} is a not-so-often studied aspect of these models.
 Traditionally, collocation methods have been used, but these require parametrization of some of the necessary surface integrals.
@@
 The use of biorthogonal wavelet bases as finite elements achieves sparsity in
-the \acs{BEM} procedure, due to their intrinsically hierarchical structure and
+the \acrshort{BEM} procedure, due to their intrinsically hierarchical structure and
 the existence of \emph{a priori} and \emph{a posteriori} matrix compression
 estimates.
-Thus, wavelet Galerkin \acs{BEM} represents a valid alternative to traditional
+Thus, wavelet Galerkin \acrshort{BEM} represents a valid alternative to traditional
 collocation methods, both to achieve a better computational scaling and to
@@
 
-Already~\citeauthor{Weijo2010-hy} had shown that using \ac{PWC} bases can lead
+Already~\citeauthor{Weijo2010-hy} had shown that using \gls{PWC} bases can lead
 to superior accuracy and convergence in the calculation of quantum mechanical
@@
 In this work we showed that even faster convergence can be achieved when
-\ac{PWL} bases are used instead.
+\gls{PWL} bases are used instead.
 Moreover, the same holds for the calculation of static electric properties.
@@
 even for very large finite element bases. This suggests that, in some cases,
-\acs{BEM} collocation methodologies might slow down or even prevent the
+\acrshort{BEM} collocation methodologies might slow down or even prevent the
 convergence of the quantum mechanical response equations solvers.
@@
 generator\autocite{Harbrecht2009-no, Harbrecht2011-dk} and wavelet Galerkin
-\acs{BEM} solver\autocite{Harbrecht2004-uo, Harbrecht2006-ug} with \pcmsolver.
+\acrshort{BEM} solver\autocite{Harbrecht2004-uo, Harbrecht2006-ug} with \pcmsolver.
 These were used to interface with the new C++ implementation of the wavelet
@@
 I implemented the interface between the \LSDALTON quantum chemistry software
-package and the \pcmsolver software library. The interface allows to run \acs{HF} and
-\acs{KS}-\acs{DFT} single-point and linear response calculations.
+package and the \pcmsolver software library. The interface allows to run \acrshort{HF} and
+\acrshort{KS}-\acrshort{DFT} single-point and linear response calculations.
 Together with coauthor Krzysztof Mozgawa, I performed the benchmark quantum
@@
 Finally, the interface to \LSDALTON was later released in the 2016 version,
-providing \acs{PCM} capabilities to the software package.
+providing \acrshort{PCM} capabilities to the software package.
 
@@
 As outlined in Chapter \ref{ch:CSM}, the boundary integral operators in the
-\acs{IEF} equation:
+\acrshort{IEF} equation:
 \begin{equation}
@@
 Green's function is sufficient to obtain the boundary integral operators in the
-\acs{PCM} integral equation.
+\acrshort{PCM} integral equation.
 The authors introduced a numerical integration procedure to calculate the
@@
 Whereas Paper I presented the essential framework for the coupling of
-four-component \acs{SCF} wave functions with a classical polarizable continuum,
+four-component \acrshort{SCF} wave functions with a classical polarizable continuum,
 in this paper we explored the calculation of first-order magnetic properties:
-\ac{EPR} and \ac{pNMR} parameters.~\autocite{Repisky2010-ls, Malkin2011-nm,
+\gls{EPR} and \gls{pNMR} parameters.~\autocite{Repisky2010-ls, Malkin2011-nm,
 Komorovsky2013-xa, Cherry2016-ij}
@@
 outset in the variational optimization of the wave function.
-Hence, \acs{EPR} and \acs{pNMR} parameters are formulated as expectation
+Hence, \acrshort{EPR} and \acrshort{pNMR} parameters are formulated as expectation
 values, by virtue of the Hellmann--Feynman theorem.~\autocite{Konishi2009-zb,
@@
 I helped coauthors Michal Repisky, Stanislav Komorovsky and Peter Hrobarik
-with setting up the \acs{PCM} calculations described in the paper.
+with setting up the \acrshort{PCM} calculations described in the paper.
 Finally, I provided the first draft for Section 2 of the paper and took part in all editing stages.
 The interface to \ReSpect will be released in the next public version of the software
-package, providing \acs{PCM} and \acs{COSMO} capabilities.
+package, providing \acrshort{PCM} and \acrshort{COSMO} capabilities.
 
@@
 Our group has newly developed an open-ended methodology for the computation of
-\acs{SCF} response functions~\autocite{Thorvaldsen2008-sg, Ringholm2014-gx} and
+\acrshort{SCF} response functions~\autocite{Thorvaldsen2008-sg, Ringholm2014-gx} and
 their residues.~\autocite{Friese2015-kb}
@@
 
-I developed the theoretical framework for the open-ended \acs{SCF} formulation
+I developed the theoretical framework for the open-ended \acrshort{SCF} formulation
 of molecular response properties when a quantum/classical polarizable continuum
@@
 I provided its implementation within the \DALTON code, by interfacing the
-\pcmsolver library and the open-ended \acs{SCF} response code of
+\pcmsolver library and the open-ended \acrshort{SCF} response code of
 \citeauthor{Ringholm2014-gx}~\autocite{Ringholm2014-gx, Friese2015-kb}
 I performed extensive testing of the code by comparing with previously
-published implementations of the \acs{PCM}-\acs{SCF} response functions within
+published implementations of the \acrshort{PCM}-\acrshort{SCF} response functions within
 \DALTON.~\autocite{Cammi2003-qy, Frediani2005-nc, Ferrighi2010-pm}

### 2016-09-01 Chapters/MathematicalResults.tex
Switch to glossaries package for acronyms

@@
 \section{Derivation of the IEF equation}
-I will show a detailed derivation of the \acs{IEF} equation
-for the \acs{PCM} transmission problem Eq.~\eqref{eq:transmission}:
+I will show a detailed derivation of the \acrshort{IEF} equation
+for the \acrshort{PCM} transmission problem Eq.~\eqref{eq:transmission}:
 \begin{subequations}
@@
 \end{equation}
-where the function $\sigma$ is the, yet unknown, \acl{ASC}.
+where the function $\sigma$ is the, yet unknown, \acrlong*{ASC}.
 
@@
 \end{empheq}
-The final ingredient is the \ac{DtN} map, which can be derived by employing Eq.~\eqref{eq:stat3}
+The final ingredient is the \gls{DtN} map, which can be derived by employing Eq.~\eqref{eq:stat3}
 to the Newton potential:
@@
 which is equal, in $\Omegai$, to the auxiliary potential $h(\vect{r})$.
-The \acs{DtN} map is thus given as:
+The \acrshort{DtN} map is thus given as:
 \begin{equation}\label{eq:regular-DtN}
@@
 \end{equation}
-With this last ingredient at hand, algebraic manipulations lead to the \acs{IEF}
+With this last ingredient at hand, algebraic manipulations lead to the \acrshort{IEF}
 equation:
@@
 For simplicity, we assume Dirichlet boundary conditions for a conductor,
-\ie the basic assumption behind \acs{COSMO}.
+\ie the basic assumption behind \acrshort{COSMO}.
 The \emph{strong} formulation of the electrostatic problem then reads:
@@
 
-Weak formulation of \acs{BIE}s.\autocite{Hsiao2008-xb}
+Weak formulation of \acrshort{BIE}s.\autocite{Hsiao2008-xb}
 
+\section{The Galerkin Method}
 Approximability, conformity, consistency and asymptotic consistency

### 2016-09-01 Chapters/QuantumMechanics.tex
Switch to glossaries package for acronyms

@@
     quantum mechanics
-  \item \ac{BO} approximation and the concept of \ac{PES}
+  \item \gls{BO} approximation and the concept of \gls{PES}
   \item Fock space
-  \item single-reference methods: \ac{HF} and \ac{KS} \ac{DFT}
-  \item the \ac{CC}  model
+  \item single-reference methods: \gls{HF} and \gls{KS} \gls{DFT}
+  \item the \gls{CC}  model
 \end{itemize}
@@
 
+\section{Single-Reference Methods}
+
 \section{The Coupled Cluster Ansatz}\label{sec:coupled-cluster}
@@
 The analyses and derivations in this Chapter are preeminently based
-on the \acl{CC} wave function Ansatz.
-We will assume a closed-shell, \acs{HF} single reference function and
+on the \acrlong*{CC} wave function Ansatz.
+We will assume a closed-shell, \acrshort{HF} single reference function and
 construct our correlated wave function as an exponential mapping:
@@
 $\mathcal{M}$ is the truncation level.
-One of the main strengths of the \acs{CC} model is its size extensivity
+One of the main strengths of the \acrshort{CC} model is its size extensivity
 which stems directly from the exponential nature of the wave
@@
 
-The \acs{CC} method is a \emph{projective} method. Chosen the truncation
+The \acrshort{CC} method is a \emph{projective} method. Chosen the truncation
 level $\mathcal{M}$, one projects the nonrelativistic Schrödinger
@@
 space compatible with the excitation level truncation.
-This is the \emph{unlinked} form of the \acs{CC}
+This is the \emph{unlinked} form of the \acrshort{CC}
 equations.~\autocite{Helgaker2000-tz} It is usually more convenient to
 perform a \emph{similarity transformation} of the Hamiltonian operator,
-leading to the \emph{linked} form of the \acs{CC} equations:
+leading to the \emph{linked} form of the \acrshort{CC} equations:
 \begin{subequations}
@@
 Use of similarity transformed operators is ubiquitous when dealing with
-\acs{CC} theory and we introduce the following notation for it:
+\acrshort{CC} theory and we introduce the following notation for it:
 \begin{equation}
@@
 Similarity-transformed operators can be expanded in a commutator series,
-also known as the \ac{BCH} expansion:
+also known as the \gls{BCH} expansion:
 \begin{equation}
@@
 where $[A,B]^n$ is the $n$-fold nested commutator. It can be shown that
-the \acs{BCH} expansion of the similarity-transformed Hamiltonian
+the \acrshort{BCH} expansion of the similarity-transformed Hamiltonian
 \emph{exactly} truncates after the four-fold nested commutator, greatly
@@
 
-Introducing the \acl{MP} partitioning of the Hamiltonian leads to
+Introducing the \acrlong*{MP} partitioning of the Hamiltonian leads to
 further insight:
@@
 
-As formulated so far, the \acs{CC} method is not variational. It is
+As formulated so far, the \acrshort{CC} method is not variational. It is
 however possible to introduce a fully variational Lagrangian that
@@
 treating the amplitude equations as constraints in the optimization of
-the \acs{CC} energy:
+the \acrshort{CC} energy:\autocite{Helgaker1982-xl, Arponen1983-qt,
+Helgaker1988-to, Helgaker1989-wl, Koch1990-vr, Kvaal2013-jr}
 \begin{equation}\label{eq:cc-lagrangian}
@@
 \end{equation}
-and the \acs{MP}-partitioned form of the Hamiltonian was also exploited.
+and the \acrshort{MP}-partitioned form of the Hamiltonian was also exploited.
 Differentiation of the Lagrangian with respect to the multipliers
@@
 multipliers. While these are not needed for the evaluation of the
-\acs{CC} energy, their calculation is mandatory when expectation values
+\acrshort{CC} energy, their calculation is mandatory when expectation values
 and molecular properties in general are sought.
-The \acs{CC} expectation values are formed using the left and
-right \acs{CC} states and we introduce the following notation:
+The \acrshort{CC} expectation values are formed using the left and
+right \acrshort{CC} states and we introduce the following notation:
 \begin{equation}
@@
 
-Having summarized the essential points of \acs{CC} theory and having
+Having summarized the essential points of \acrshort{CC} theory and having
 introduced the relevant notation, we can proceed to analyze its coupling

### 2016-09-01 Chapters/ResponseTheory.tex
Change name of one chapter
Switch to glossaries package for acronyms

--- 2016-08-26 Chapters/ResponseTheory.tex
@@
-\chapter{Response theory and molecular properties}\label{ch:molprop}
+\chapter{Response Theory and Molecular Properties}\label{ch:molprop}
 
@@
 \end{itemize}
+
+\pagebreak
+
+\section{Exact State Response Theory}\label{sec:exact-response}
+
+General theory~\autocite{Olsen1985-nr, Christiansen1998-pe,
+Norman2011-ad, Helgaker2012-cz, Pawlowski2015-sq}
+Solution strategy for response equations~\autocite{Kauczor2011-rd}
+Open-ended \acrshort{AO}-based \acrshort{SCF} response~\autocite{
+Larsen2000-hj, Kjaergaard2008-hy, Thorvaldsen2008-sg,
+Kristensen2008-hv, Ringholm2014-gx}
+Open-ended \acrshort{SCF} residues~\autocite{Friese2015-kb}
+
+\section{SCF Response Theory for Quantum/Classical Polarizable Hamiltonians}\label{sec:csm-response}
+
+\acrshort{PCM}~\autocite{Cammi1994-qj, Cammi1996-wf, Cammi1996-vx,
+Cammi1999-rb, Cammi2003-qy, Frediani2005-nc, Ferrighi2010-pm}
+\acrshort{FQ} and \acrshort{FQ}/\acrshort{PCM}~\autocite{Lipparini2012-hx, Lipparini2012-tl, Lipparini2013-ud}
+\acrshort{PE}~\autocite{Olsen2010-wa}
+\acrshort{QM}/\acrshort{MM}/Continuum~\autocite{Stenidal2011-ki, Caprasecca2012-ir}
+Paper V~\autocite{pcm-openrsp}

### 2016-09-01 Chapters/SolvationAndCorrelation.tex
Contributions written
Switch to glossaries package for acronyms

@@
 \epigraph{Potius sero quam numquam}{--- \textsc{Titus Livius}, \textit{Ab Urbe Condita}}
-\epigraph{Non temerai i terrori della notte non temerai il terrore}{--- \textsc{CCCP}}
+\epigraph{\textitalian{Non temerai i terrori della notte non temerai il terrore}}{--- \textsc{CCCP}}
 
@@
 polarizable models when the quantum part of the system includes a
-many-body description of electron correlation either by a \ac{MBPT},
-\ac{CC} or hybrid approach.
+many-body description of electron correlation either by a \gls{MBPT},
+\gls{CC} or hybrid approach.
 These results leverage the variational formulation of quantum/classical
@@
 will show how a fully variational framework enables transparent
-derivations of \acs{MBPT} and \acs{CC} theory, encompassing previous
+derivations of \acrshort{MBPT} and \acrshort{CC} theory, encompassing previous
 work~\cite{Cammi, Olivares del Valle, Caricato, Sneskov}
@@
 
-After introducing the notation, I will derive the form of the \acs{CC}
+After introducing the notation, I will derive the form of the \acrshort{CC}
 effective Lagrangian and proceed to show how the work of
 \citeauthor{Cammi2009-gu} and \citeauthor{Caricato2011-tx} on
-\acs{PCM}-\acs{CC} and \citeauthor{Sneskov} on \acs{PE}-\acs{CC}
+\acrshort{PCM}-\acrshort{CC} and \citeauthor{Sneskov} on \acrshort{PE}-\acrshort{CC}
 can easily be derived in the variational framework.
@@
 The second part of this Chapter will be devoted to the derivation of
-\acs{MBPT} for quantum/classical polarizable Hamiltonians.
-I will exploit the \acs{CC} Ansatz for the wave function and formulate
-\acs{MBPT} in terms of the effective \acs{CC} Lagrangian, an approach
+\acrshort{MBPT} for quantum/classical polarizable Hamiltonians.
+I will exploit the \acrshort{CC} Ansatz for the wave function and formulate
+\acrshort{MBPT} in terms of the effective \acrshort{CC} Lagrangian, an approach
 guaranteeing term-by-term size extensivity.
 
-Iterative approximations to the \acs{CCSD} and \acs{CCSDT} models will
+Iterative approximations to the \acrshort{CCSD} and \acrshort{CCSDT} models will
 be the subject of Section \ref{sec:cc2-and-cc3} where the equations
-governing the \acs{CC2} and \acs{CC3} models will be derived.
+governing the \acrshort{CC2} and \acrshort{CC3} models will be derived.
 Eventually, I will describe schemes for the noniterative inclusion of
 the effect of connected triples excitations on top of
-the \acs{CCSD} wave function. The symmetric and asymmetric [T] and (T)
+the \acrshort{CCSD} wave function. The symmetric and asymmetric [T] and (T)
 corrections will be derived.
@@
 
-Our purpose in this Section is to derive an \emph{effective} \acs{CC}
+Our purpose in this Section is to derive an \emph{effective} \acrshort{CC}
 Lagrangian that takes into account the coupling of the correlated
@@
 Once again, let us assume that the reference state is a closed-shell
-\acs{HF} determinant. Further, we assume that the reference determinant
+\acrshort{HF} determinant. Further, we assume that the reference determinant
 was optimized in the presence of the classical polarizable environment.
@@
 \end{equation}
-The \acs{BCH} expansions of the source terms have been
+The \acrshort{BCH} expansions of the source terms have been
 rewritten as:
@@
 appears among the terms in the functional but will obviously not enter
-in the optimization of the \acs{CC} wave function.
+in the optimization of the \acrshort{CC} wave function.
 Recasting the polarization functional in the supermatrix formalism (see
@@
 ${}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})$ term to yield the
-\acl{PTE} (\acs{PTE}) Hamiltonian -- $H$ -- and eventually the
-\acl{PTED} (\acs{PTED}) Lagrangian:~\autocite{Olivares del Valle,
+\acrlong*{PTE} (\acrshort{PTE}) Hamiltonian -- $H$ -- and eventually the
+\acrlong*{PTED} (\acrshort{PTED}) Lagrangian:~\autocite{Olivares del Valle,
 Lipparini}
@@
 \end{equation}
-where the \acs{PTE} Lagrangian has been introduced:
+where the \acrshort{PTE} Lagrangian has been introduced:
 \begin{equation}
@@
 
-As noted by \citeauthor{Cammi2009-gu} for the implicit \acs{PCM} model,
-the \acs{PTE} Hamiltonian is computed in \acs{MP}-partitioned form by
+As noted by \citeauthor{Cammi2009-gu} for the implicit \acrshort{PCM} model,
+the \acrshort{PTE} Hamiltonian is computed in \acrshort{MP}-partitioned form by
 employing the ``solvated'' orbitals and Fock matrix:
@@
 fluctuation potential.
-The simple nature of the Fock operator lets us re-express the \acs{PTE}
+The simple nature of the Fock operator lets us re-express the \acrshort{PTE}
 Lagrangian as follows:
@@
 \end{subequations}
-By virtue of Eq.~\eqref{eq:HF-double-commutator} the \acs{HF}
+By virtue of Eq.~\eqref{eq:HF-double-commutator} the \acrshort{HF}
 expectation value of the double commutator can be dropped. Thus, for a
@@
 \end{equation}
-Note that the singles amplitudes assume a unique role in the \acs{CC}
+Note that the singles amplitudes assume a unique role in the \acrshort{CC}
 expression of the source term.
@@
 \commutator{\commutator{\s}{T_1}}{T_3} | \mathrm{HF}} \neq 0$
-More explicitly, for the \ac{CCS} model ($\mathcal{M} = 1$) one has:
+More explicitly, for the \gls{CCS} model ($\mathcal{M} = 1$) one has:
 \begin{equation}
@@
 \end{equation}
-while for the \ac{CCSD} model ($\mathcal{M} = 2$):
+while for the \gls{CCSD} model ($\mathcal{M} = 2$):
 \begin{equation}
@@
 \end{equation}
-Eventually, within the \ac{CCSDT} model ($\mathcal{M} = 3$) four more
+Eventually, within the \gls{CCSDT} model ($\mathcal{M} = 3$) four more
 terms are added:
@@
 with respect to the variational parameters $\tamp{\mu_i}$,
-$\tbar{\mu_i}$ and $\p$ yields the \acs{PTED}-\acs{CC}
+$\tbar{\mu_i}$ and $\p$ yields the \acrshort{PTED}-\acrshort{CC}
 equations:
@@
 scheme for its implementation.~\autocite{Cammi2009-gu, Caricato2010-hx}
-As a consequence, a single-point \acs{PTED}-\acs{CC} will suffer from a
+As a consequence, a single-point \acrshort{PTED}-\acrshort{CC} will suffer from a
 $2\times$ prefactor in its computational cost with respect to an
-\emph{in vacuo} \acs{CC} calculation.
+\emph{in vacuo} \acrshort{CC} calculation.
 
@@
 singles cluster operator and the singles multiplier state.
-Three approximation have currently been proposed: \acs{PTES}, \acs{PTE(S)}
-and \acs{PTE}.
+Three approximation have currently been proposed: \acrshort{PTES}, \acrshort{PTE(S)}
+and \acrshort{PTE}.
 
@@
 
-The first approximate scheme truncates the source term to its \acs{CCS}
+The first approximate scheme truncates the source term to its \acrshort{CCS}
 expectation value, yielding the effective Lagrangian:
@@
 These equations are, however, still coupled due to the presence of the
-multipliers in the \acs{CCS} source term in the polarization equation.
+multipliers in the \acrshort{CCS} source term in the polarization equation.
 To break the coupling, we further approximate $\s_{\mathrm{N}}^1(\tamp{}, \tbar{})$
@@
 \end{equation}
-so that the governing equations for the \acs{PTES} scheme become:
+so that the governing equations for the \acrshort{PTES} scheme become:
 \begin{subequations}
@@
 
-The \acs{PTES} approximation can be brought one step further, by
-truncating the \acs{CCS} expectation value to its multiplier-independent
+The \acrshort{PTES} approximation can be brought one step further, by
+truncating the \acrshort{CCS} expectation value to its multiplier-independent
 term:
@@
 The amplitude equations are formally equivalent to those derived in the
-\acs{PTE} scheme, \emph{vide infra}. The equations for multipliers are also formally
-equivalent to the \acs{PTE} equations, apart from the equation
+\acrshort{PTE} scheme, \emph{vide infra}. The equations for multipliers are also formally
+equivalent to the \acrshort{PTE} equations, apart from the equation
 determining the singles multipliers which features an additional term.
@@
 
-Practical implementation of the \acs{PTE(S)} scheme will solve the
-\acs{PTE}-like amplitude equation, form the source term
+Practical implementation of the \acrshort{PTE(S)} scheme will solve the
+\acrshort{PTE}-like amplitude equation, form the source term
 $\braket{\mathrm{HF} | \check{\s}_\mathrm{N} | \mathrm{HF}}$, solve for
 the polarization degrees of freedom and calculate the polarization
-energy as a correction to the \acs{CC} energy.
+energy as a correction to the \acrshort{CC} energy.
 Only when a first-order molecular property, such as the molecular
 gradient, is needed, will the multipliers equations need to be solved.
-In other words, only the $T_1$-dependent part of the \acs{CC} density is
+In other words, only the $T_1$-dependent part of the \acrshort{CC} density is
 used to define the classical sources.
@@
 drastic approximation to the full quantum/classical polarizable
-effective Lagrangian, the \acs{PTE} scheme.
+effective Lagrangian, the \acrshort{PTE} scheme.
 This is equivalent to differentiating the polarization-independent
-\acs{PTE} effective Lagrangian:
+\acrshort{PTE} effective Lagrangian:
 \begin{subequations}
@@
 \end{subequations}
-The \acs{PTE} scheme naturally preserves the scaling of the underlying
-\acs{CC} method, since no coupling between the \acs{CC} amplitudes and
+The \acrshort{PTE} scheme naturally preserves the scaling of the underlying
+\acrshort{CC} method, since no coupling between the \acrshort{CC} amplitudes and
 multipliers equations is introduced.
@@
 thus readily implemented.
-The \acs{PTE} model offers an efficient approximation to the full
-\acs{PTED} model with a computational cost comparable to that of
-\emph{in vacuo} \acs{CC} theory.
+The \acrshort{PTE} model offers an efficient approximation to the full
+\acrshort{PTED} model with a computational cost comparable to that of
+\emph{in vacuo} \acrshort{CC} theory.
 
@@
 We will now develop a perturbative many-body expansion of the effective
-\acs{CC} Lagrangian when a quantum/classical polarizable Hamiltonian is
+\acrshort{CC} Lagrangian when a quantum/classical polarizable Hamiltonian is
 used.
-The use of a \acs{CC} Ansatz for the wave function ensures term-by-term
+The use of a \acrshort{CC} Ansatz for the wave function ensures term-by-term
 size extensivity of the energy at all orders.~\autocite{Helgaker2000-tz}
@@
 it will be considered of order 1: $\Phi = O(1)$.
-We still assume a closed-shell \acs{HF} reference function and
+We still assume a closed-shell \acrshort{HF} reference function and
 perturbation orders will be counted accordingly.
@@
 
-The \acs{PTED}-\acs{CC} equations are expanded in orders of the
+The \acrshort{PTED}-\acrshort{CC} equations are expanded in orders of the
 perturbation and terms are collected order by order. The use of an
@@
 \end{subequations}
-Thus, despite the closed-shell \acs{HF} reference, the singles
+Thus, despite the closed-shell \acrshort{HF} reference, the singles
 amplitudes will already appear in first order, due to the
@@
 \end{equation}
-Thus, as for the amplitudes, despite the closed-shell \acs{HF} reference, the singles
+Thus, as for the amplitudes, despite the closed-shell \acrshort{HF} reference, the singles
 multipliers will already appear in first order, due to the
@@
 There are no explicit contributions from triples. This means that the
-\acs{CCSD} polarization is correct through fourth order in the
+\acrshort{CCSD} polarization is correct through fourth order in the
 fluctuation potential.
@@
 
-The \acs{PTED}-\acs{CCSD} scheme for the quantum/classical coupling is
+The \acrshort{PTED}-\acrshort{CCSD} scheme for the quantum/classical coupling is
 only includes singles and doubles in its excitation manifold and is thus
@@
 \end{subequations}
-We will denote the \acs{CCSD} converged parameters with a $*$ superscript.
+We will denote the \acrshort{CCSD} converged parameters with a $*$ superscript.
 \todo[inline]{Check the statement above about $\p^{*}$.}
@@
 The [T] noniterative triples corrections are formed by adding the fourth
-order \acs{CC}\acs{PT} terms arising from connected triples on top of
-the \acs{CCSD} energy. The \emph{converged} \acs{CCSD} amplitudes,
+order \acrshort{CC}\acrshort{PT} terms arising from connected triples on top of
+the \acrshort{CCSD} energy. The \emph{converged} \acrshort{CCSD} amplitudes,
 multipliers and polarization are however used instead of the
 perturbative quantities in Eq.~\eqref{eq:triples-fourth-order}.
-The correction term is thus improved, since the \acs{CCSD} quantities
+The correction term is thus improved, since the \acrshort{CCSD} quantities
 are determined to infinite order, within the singles and doubles space.
@@
 \end{equation}
-Replacing the perturbation multipliers by converged \acs{CCSD}
+Replacing the perturbation multipliers by converged \acrshort{CCSD}
 amplitudes leads to the symmetric CCSD[T] model:
@@
 order terms to the fourth order [T] correction given above.
-Also for the (T) corrections, the \emph{converged} \acs{CCSD} amplitudes, multipliers and polarization
+Also for the (T) corrections, the \emph{converged} \acrshort{CCSD} amplitudes, multipliers and polarization
 are used instead of the perturbative quantities.
@@
 
-The correction term is thus improved, since the \acs{CCSD} quantities
+The correction term is thus improved, since the \acrshort{CCSD} quantities
 are determined to infinite order, within the singles and doubles space.
@@
 \end{equation}
-Replacing the perturbation multipliers by converged \acs{CCSD}
+Replacing the perturbation multipliers by converged \acrshort{CCSD}
 amplitudes leads to the symmetric CCSD(T) model:
@@
 
-Triples excitations enter the \acs{MBPT} equations for the singles and
-doubles amplitudes and multipliers at third order. Thus the \acs{CCSD}
+Triples excitations enter the \acrshort{MBPT} equations for the singles and
+doubles amplitudes and multipliers at third order. Thus the \acrshort{CCSD}
 amplitudes and multipliers are correct through second order in the

### 2016-09-01 FrontBackmatter/Introduction.tex
Switch to glossaries package for acronyms

@@
    This is either achieved by a cheaper quantum mechanical
-   method~\autocite{Vreven2006-gx} or by \ac{MM}.~\autocite{Senn2009-sk}
-   In the latter approach, commonly dubbed \acs{QM}/\acs{MM}, the \acs{MM}
+   method~\autocite{Vreven2006-gx} or by \gls{MM}.~\autocite{Senn2009-sk}
+   In the latter approach, commonly dubbed \acrshort{QM}/\acrshort{MM}, the \acrshort{MM}
    region can either be polarizable~\autocite{MMPol} or non
@@
 \end{itemize}
-\acs{QM}/\acs{MM} models can capture, albeit approximately, the effect
+\acrshort{QM}/\acrshort{MM} models can capture, albeit approximately, the effect
 of the atomistic nature of the environment on the active part of the
@@
 to yield results of any significance. Moreover, a large number
-\emph{cutoff} for the \acs{MM} region is usually required to converge
+\emph{cutoff} for the \acrshort{MM} region is usually required to converge
 long-range electrostatic interactions.~\autocite{Something here??}
@@
 recover a satisfactory description of specific effects.
-To partly alleviate these sources of error, the \acs{QM}/\acs{MM} and
-\acs{QM}/Continuum methods can and have been successfully combined to
-yield the three-layer \acs{QM}/\acs{MM}/Continuum method.~\autocite{Curutchet, Steindal Lipparini}
+To partly alleviate these sources of error, the \acrshort{QM}/\acrshort{MM} and
+\acrshort{QM}/Continuum methods can and have been successfully combined to
+yield the three-layer \acrshort{QM}/\acrshort{MM}/Continuum method.~\autocite{Curutchet, Steindal Lipparini}
 Figure \ref{fig:qm-to-multiscale} schematically portrays the transition
-from a full \acs{QM} model of the relevant system to its multiscale
+from a full \acrshort{QM} model of the relevant system to its multiscale
 representations.
@@
 
-Chapter \ref{ch:CSM} will present an overview of the \ac{PCM} for
+Chapter \ref{ch:CSM} will present an overview of the \gls{PCM} for
 solvation.
@@
 Lipparini2015-lq} I will introduce a unifying theoretical formalism for
-\acs{QM}/Continuum, \acs{QM}/\acs{MM} and \acs{QM}/\acs{MM}/Continuum models that will
+\acrshort{QM}/Continuum, \acrshort{QM}/\acrshort{MM} and \acrshort{QM}/\acrshort{MM}/Continuum models that will
 be extensively used throughout the thesis.

### 2016-09-01 FrontBackmatter/Notation.tex
Appearance of notation should be fixed
Switch to glossaries package for acronyms

@@
+%\chapter*{Notations, conventions and units}
+\addcontentsline{toc}{chapter}{\tocEntry{Notations, conventions and units}}
+
 \renewcommand\nomgroup[1]{%
@@
 }
+
 \renewcommand{\nompreamble}{
@@
 \nomenclature[A, 03]{$\delta_{ij}$}{Kronecker symbol: $\delta_{ij}=1i$ if $i=j$ and 0 otherwise}
+\nomenclature[A, 04]{$\scalprod[X]{u}{v}$}{Scalar product of $u$ and $v$ in the Hilbert space $X$}
+\nomenclature[A, 05]{$\norm{u}{X}$}{Norm of $u$ in the normed space $X$}
 % Vectors and Matrices
@@
 % Function Spaces
-\nomenclature[D, 01]{$\norm{u}{X}$}{Norm of $u$ in the normed space $X$}
 \nomenclature[D, 02]{$\mathcal{C}^0(\Omega)$}{Space of continuous functions on $\Omega\subset\mathbb{R}^d$}
@@
 % Molecular Quantum Mechanics
-\nomenclature[E, 01]{$r, s, t$}{General \acl{MO} indices}
-\nomenclature[E, 02]{$i, j, k$}{Occupied \acl{MO} indices}
-\nomenclature[E, 03]{$a, b, c$}{Virtual \acl{MO} indices}
+\nomenclature[E, 01]{$r, s, t$}{General \acrlong*{MO} indices}
+\nomenclature[E, 02]{$i, j, k$}{Occupied \acrlong*{MO} indices}
+\nomenclature[E, 03]{$a, b, c$}{Virtual \acrlong*{MO} indices}
 \nomenclature[E, 04]{$\kappa, \lambda, \mu$}{One-electron basis functions indices}

### 2016-09-01 acronyms.tex
Update notations
Write some of the abstract
Massage the introduction
Save the day's work
Paper summaries done
Contributions chapter done
Work on CSM chapter
Switch to glossaries package for acronyms

--- 2016-07-17 acronyms.tex
@@
-\begin{acronym}
-\acro{CC}{coupled cluster}
-\acro{CCSD}{coupled cluster with single and double substitutions}
-\acro{CI}{configuration interaction}
-\acro{MO}{molecular orbital}
-\acro{LR}{linear response}
-\acro{DFT}{density-functional theory}
-\acro{PCM}{polarizable continuum model}
-\acro{MM}{molecular mechanics}
-\acro{MP}{M{\o}ller--Plesset}
-\acro{HF}{Hartree--Fock}
-\acro{ASC}{apparent surface charge}
-\acro{PTE}{perturbation to the energy}
-\acro{PTED}{perturbation to the energy and the density}
-\acro{QM/MM}{Quantum Mechanics/Molecular Mechanics}
-\acro{DC}{Dielectric Continuum}
-\acro{GEPOL}{GEPOL}
-\acro{IEF}{Integral Equation Formalism}
-\acro{LHS}{left-hand side}
-\acro{MD}{Molecular Dynamics}
-\acro{MM}{Molecular Mechanics}
-\acro{BVP}{boundary value problem}
-\acro{PDE}{partial differential equation}
-\acro{PE}{Polarizable Embedding}
-\acro{QC}{Quantum Chemistry}
-\acro{RHS}{right-hand side}
-\acro{SAS}{Solvent-Accessible Surface}
-\acro{SES}{Solvent-Excluded Surface}
-\acro{MEP}{molecular electrostatic potential}
-\acro{TDSCF}{time-dependent SCF}
-\acro{KS}{Kohn--Sham}
-\end{acronym}
+\newacronym{API}{API}{application programming interface}
+\newacronym{COSMO}{COSMO}{conductor-like screening model}
+\newacronym{BO}{BO}{Born--Oppenheimer}
+\newacronym[plural=PES, firstplural={potential energy surfaces (PESs)}]{PES}{PES}{potential energy surface}
+\newacronym{CI}{CI}{configuration interaction}
+\newacronym{CC}{CC}{coupled cluster}
+\newacronym{CCS}{CCS}{coupled cluster with single substitutions}
+\newacronym{CCSD}{CCSD}{coupled cluster with single and double substitutions}
+\newacronym{CCSDT}{CCSDT}{coupled cluster with single, double and triple substitutions}
+\newacronym{CC2}{CC2}{approximate coupled cluster singles and doubles}
+\newacronym{CC3}{CC3}{approximate coupled cluster singles, doubles and triples}
+\newacronym{PT}{PT}{perturbation theory}
+\newacronym{MBPT}{MBPT}{many-body perturbation theory}
+\newacronym{MO}{MO}{molecular orbital}
+\newacronym{AO}{AO}{atomic orbital}
+\newacronym{LR}{LR}{linear response}
+\newacronym{DFT}{DFT}{density-functional theory}
+\newacronym{PCM}{PCM}{polarizable continuum model}
+\newacronym{MM}{MM}{molecular mechanics}
+\newacronym{MP}{MP}{M{\o}ller--Plesset}
+\newacronym{HF}{HF}{Hartree--Fock}
+\newacronym{ASC}{ASC}{apparent surface charge}
+\newacronym{PTE}{PTE}{perturbation-to-the-energy}
+\newacronym{PTES}{PTES}{perturbation to the energy}
+\newacronym{PTE(S)}{PTE(S)}{perturbation to the energy}
+\newacronym{PTED}{PTED}{perturbation-to-the-energy-and-the-density}
+\newacronym{IEF}{IEF}{integral equation formalism}
+\newacronym{LHS}{LHS}{left-hand side}
+\newacronym{MD}{MD}{molecular dynamics}
+\newacronym{QM}{QM}{quantum mechanics}
+\newacronym{BVP}{BVP}{boundary value problem}
+\newacronym{BEM}{BEM}{boundary element method}
+\newacronym{BIE}{BIE}{boundary integral equation}
+\newacronym[plural=PDE, firstplural={partial differential equations (PDEs)}]{PDE}{PDE}{partial differential equation}
+\newacronym{DtN}{DtN}{Dirichlet-to-Neumann}
+\newacronym{PE}{PE}{polarizable embedding}
+\newacronym{FQ}{FQ}{fluctuating charges}
+\newacronym{QC}{QC}{quantum chemistry}
+\newacronym{RHS}{RHS}{right-hand side}
+\newacronym{vdWS}{vdWS}{van der Waals surface}
+\newacronym{SAS}{SAS}{solvent-accessible surface}
+\newacronym{SES}{SES}{solvent-excluded surface}
+\newacronym{MEP}{MEP}{molecular electrostatic potential}
+\newacronym{SCF}{SCF}{self-consistent field}
+\newacronym{TDSCF}{TDSCF}{time-dependent SCF}
+\newacronym{KS}{KS}{Kohn--Sham}
+\newacronym{PWC}{PWC}{piecewise constant wavelet}
+\newacronym{PWL}{PWL}{piecewise constant linear}
+\newacronym{DC}{DC}{Dirac--Coulomb}
+\newacronym{BCH}{BCH}{Baker--Campbell--Hausdorff}
+\newacronym[plural=DVCS, firstplural={distributed version control systems (DVCSs)}]{DVCS}{DVCS}{distributed version control system}
+\newacronym{EPR}{EPR}{electron paramagnetic resonance}
+\newacronym{NMR}{NMR}{nuclear magnetic resonance}
+\newacronym{pNMR}{pNMR}{paramagnetic nuclear magnetic resonance}
+\newacronym{LGPL}{LGPL}{GNU Lesser General Public License}
+\newacronym{AD}{AD}{automatic differentiation}
+\newacronym{CSC}{CSC}{continuous surface charge}
+\newacronym{SWIG}{SWIG}{switching/Gaussian}

### 2016-09-01 Chapters/SolvationAndCorrelation.tex
Fix some missing references in CC chapter

@@
 \gls{CC} or hybrid approach.
-These results leverage the variational formulation of quantum/classical
-polarizable Hamiltonians described in Section \ref{sec:variational}.
-This is an element of novelty with respect to the existing literature: I
-will show how a fully variational framework enables transparent
-derivations of \acrshort{MBPT} and \acrshort{CC} theory, encompassing previous
-work~\cite{Cammi, Olivares del Valle, Caricato, Sneskov}
-and solving some inconsistencies.~\cite{Angyan1992}
-
 The contents of this Chapter are the result of an ongoing collaboration
@@
 
-After introducing the notation, I will derive the form of the \acrshort{CC}
-effective Lagrangian and proceed to show how the work of
-\citeauthor{Cammi2009-gu} and \citeauthor{Caricato2011-tx} on
-\acrshort{PCM}-\acrshort{CC} and \citeauthor{Sneskov} on \acrshort{PE}-\acrshort{CC}
-can easily be derived in the variational framework.
+I will leverage the variational formulation of quantum/classical
+polarizable Hamiltonians described in Section \ref{sec:variational}.
+This is an element of novelty with respect to the existing literature.
+I will show how a fully variational framework enables transparent
+derivations of \acrshort{MBPT} and \acrshort{CC} theory, encompassing
+previous work~\autocite{Christiansen1999-tj, Cammi2009-gu,
+Caricato2010-hx, Caricato2011-tx, Olivares_del_Valle1991-of,
+Aguilar1991-vq, Olivares_del_Valle1991-tq, Olivares_del_Valle1993-xq,
+Olivares_del_Valle1993-ra, Surjan1983-oc, Angyan1991-mr, Angyan1993-ay,
+Angyan2008-nj, Nielsen2001-yl, Kongsted2003-py, Sneskov2010-dz,
+Sneskov2011-jm, Schwabe2012-cf}
+and solving some inconsistencies in the formulation of \acrshort{MBPT}
+noted by \citeauthor{Angyan1995-co}.\autocite{Angyan1995-co,
+Lipparini2009-io}
+
+Sections \ref{sec:effective-cc-lagrangian} and \ref{sec:source-terms}
+will introduce the effective \acrshort{CC} Lagrangian and the relevant
+notation for the source terms in the classical energy functional,
+respectively.
+In Section \ref{sec:pcm-cc-models} I will derive the governing equation
+for the effective Lagrangian and show how previous work on
+continuum\autocite{Christiansen1999-tj, Cammi2009-gu, Caricato2011-tx}
+and explicit models\autocite{Kongsted2003-py, Sneskov2011-jm} can easily
+be derived in the variational framework.
 
@@
 \section{Effective Coupled Cluster Lagrangian}
+\label{sec:effective-cc-lagrangian}
 
@@
 
-\section{Source terms in the classical energy functional}
+\section{Source Terms in the Classical Energy Functional}
+\label{sec:source-terms}
 

### 2016-09-02 Chapters/QuantumMechanics.tex
Work on MQM chapter

@@
 
-\chapter{A Quick Tour of Quantum Chemistry}\label{ch:QM}
+\chapter{A Quick Tour of Molecular Electronic Structure Theory}\label{ch:QM}
 
-\begin{itemize}
-  \item nonrelativistic and relativistic Hamiltonians used in molecular
-    quantum mechanics
-  \item \gls{BO} approximation and the concept of \gls{PES}
-  \item Fock space
-  \item single-reference methods: \gls{HF} and \gls{KS} \gls{DFT}
-  \item the \gls{CC}  model
-\end{itemize}
+The purpose of quantum chemistry is to provide models based on first
+principles that can help understand and predict molecular properties.
+In quantum chemistry, we apply physical models based on
+quantum many-body methods to molecular systems, employ their
+mathematical realizations and devise computable approximations.
+The central idea is, in fact, to be able to obtain an algorithmic
+implementation of the methods that can be applied to interesting
+chemical systems.
 
-\pagebreak
+This Chapter presents a brief overview of molecular electronic structure
+theory with particular emphasis on the methods used in this thesis.
+Section \ref{sec:mqm} is a general introduction to the methods of
+molecular quantum mechanics, relativistic or nonrelativistic.
+I will provide a brief introduction to the language of second
+quantization, used throughout the thesis.
+Section \ref{sec:mean-field} presents the mean-field approximation to
+the molecular electronic structure problem, the workhorse of quantum
+chemistry.
+Sections \ref{sec:coupled-cluster}--\ref{sec:cc-approximate} are
+concerned with the \gls{CC} and \gls{MBPT} methods for the inclusion of
+electron correlation.
 
-\section{Molecular Quantum Mechanical Hamiltonians}\label{sec:hamiltonians}
+\section{Molecular Quantum Mechanics}\label{sec:mqm}
 
-Konishi and Paffuti\autocite{Konishi2009-zb}
-Dyall's book\autocite{Dyall2007-tu}
-Reiher's book\autocite{Reiher2014-cp}
-The Pink Book\autocite{Helgaker2000-tz}
-Shavitt's book\autocite{Shavitt2009-mr}
+Quantum mechanics is the theory describing the motion and interactions
+of microscopic particles. In quantum theory every observable of the
+system is represented mathematically by an operator $O$ in the
+appropriate Hilbert space of \emph{wave functions}.
+A wave function $\psi$ is the mathematical object describing the state of the
+system. In the usual Copenhagen interpretation of quantum mechanics, the
+modulus square of the wave function $|\psi|^2$ provides the key to
+predicting experimentally measurable quantities.\autocite{Konishi2009-zb}
+The wave function $\psi$ is the solution to the Schr\"{o}dinger equation:
+\begin{equation}\label{eq:schrodinger}
+  H\psi = E\psi
+\end{equation}
+where $H$ is the \emph{Hamiltonian} operator and $E$ is the energy of
+the system. The Schr\"{o}dinger equation is an eigenvalue equation for
+the Hamiltonian operator and $\psi$ is thus an
+eigenfunction.\autocite{Arfken2013-pt}
 
-\section{Single-Reference Methods}
+Molecular quantum mechanics is concerned with the microscopic motion of
+nuclei and electrons in molecules and is thus an intrinsically many-body
+theory. The problem is clearly very complicated to solve, as the number of
+interactions to be considered is large.
+Since the nuclei are much heavier than the electrons, the \gls{BO} or
+clamped-nuclei approximation is typically assumed. The nuclei are fixed
+in a specific configuration (called a \emph{molecular geometry}) and
+treated as static electric sources.\autocite{Born1927-ce}
+This separation of motions results in an electron-only Hamiltonian and
+an electronic wave function that depends \emph{parametrically} on the
+nuclear positions. The eigenvalue of this Hamiltonian, also dependent on
+the nuclear positions, is the \gls{PES} of the system, that is the
+potential function in which the nuclei move.\autocite{Szabo1989-vl,
+McWeeny1992-oj}
+In first quantization, the general expression for the clamped-nuclei,
+$N$-electron, molecular electronic Hamiltonian is:
+\begin{alignat}{2}\label{eq:fq-hamiltonian}
+  H = \sum_{i=1}^N h(\vect{r}_i)
+  + \frac{1}{2}\sum_{i\neq j}^N g(\vect{r}_i, \vect{r}_j)
+  + V_\mathrm{NN},
+  \quad&
+  V_\mathrm{NN} =
+  \frac{1}{2}\sum_{A \neq B}^{N_\mathrm{nuclei}}\frac{Z_A Z_B}{|\vect{R}_A - \vect{R}_B|}
+\end{alignat}
+The last term is the nuclear repulsion energy, a constant shift of the
+energy scale in the \acrshort{BO} approximation.
+As noted by \citeauthor{Saue2011-qg}, the same general expression for
+the clamped-nuclei Hamiltonian Eq.~\eqref{eq:fq-hamiltonian} is valid in
+the nonrelativistic and relativistic no-pair
+regimes.\autocite{Sucher1980-vf, Saue2011-qg}
+In the nonrelativistic domain one defines the one- and two-electron
+operators as:\autocite{Szabo1989-vl, McWeeny1992-oj, Konishi2009-zb}
+\begin{subequations}
+ \begin{align}
+   h(\vect{r}_i) &=
+   -\frac{1}{2}\nabla^2_i + V_\mathrm{Ne}(\vect{r}_i)
+   = -\frac{1}{2}\nabla^2_i - \sum_{A=1}^{N_\mathrm{nuclei}}\frac{Z_A}{|\vect{R}_A -
+   \vect{r}_i|} \label{eq:h} \\
+   g(\vect{r}_i, \vect{r}_j) &= \frac{1}{|\vect{r}_i - \vect{r}_j|},
+   \label{eq:g-coulomb}
+ \end{align}
+\end{subequations}
+while in the relativistic regime with the \gls{DC}
+approximation:\autocite{Dyall2007-tu, Konishi2009-zb, Reiher2014-cp}
+\begin{subequations}
+ \begin{align}
+   h(\vect{r}_i) &=
+   \begin{pmatrix}
+     V_{\mathrm{Ne}}(\vect{r}_i)I_2 & c \vect{\sigma}\cdot\vect{p}_i \\
+     \vect{\sigma}\cdot\vect{p}_i & (V_{\mathrm{Ne}}(\vect{r}_i)-2\si{\electronmass}c^2)I_2
+   \end{pmatrix} \label{eq:h-dirac} \\
+   g(\vect{r}_i, \vect{r}_j) &= \frac{I_4\cdot I_4}{|\vect{r}_i - \vect{r}_j|}
+   \label{eq:g-r-coulomb}
+ \end{align}
+\end{subequations}
+and the operators are implicitly assumed to embed the proper no-pair,
+positive-energy manifold projectors.\autocite{Sucher1980-vf}
+Since these are fermionic Hamiltonians, the spin-statistics theorem
+dictates the corresponding eigenfunctions to be
+antisymmetric.
+Using the expansion theorem,\autocite{Gross1991-hi} we thus seek the wave
+function for the $N$-electron system as a linear combination of
+$N$-electron \emph{Slater determinants}.
+
+It the language of second quantization, we assume that a, not
+necessarily complete, set of orthonormal one-particle states is
+available. These are usually called \glspl{MO}.
+In general, we would like to embed as many of the symmetries of the
+system into the one-particle basis. This is accomplished by requiring
+the one-particle space to be the common eigenbasis for a complete set of
+suitable, commuting one-particle operators.\autocite{Gross1991-hi}
+Any $N$-particle state can be constructed by distributing particles
+into one-particle states and accordingly labelling them by the number of
+particles per one-particle states present.
+This is the so-called \gls{ON} vector representation:
+\begin{equation}
+  \ket{\vect{n}} = \ket{n1,n2,\ldots},
+\end{equation}
+and the span of all \acrshort{ON} vectors is called the \emph{Fock
+space} $F^N$, a space isomorphic to the Hilbert space $H^N$ of
+$N$-particle wave functions.
+To preserve consistency, the \emph{vacuum} state with no particles:
+\begin{equation}
+ \ket{\mathrm{vac}} = \ket{0,0,\ldots}
+\end{equation}
+is included in the construction of Fock space.
+Any $N$-particle state can be generated by application of
+\emph{creation} operators $\cons{i}$. Creation operators increase the
+occupation number in the one-particle state $i$ by $1$, up to a phase.
+Conversely, \emph{annihilation} operators $\anni{i}$, defined as the Hermitian
+conjugates of the creation operators, decrease occupation in state $i$
+by $1$, again up to a phase. Clearly, annihilation of any one-particle
+state in the vacuum results yields $0$.
+Pauli's exclusion principle is enforced by the \emph{canonical
+anticommutation relations}:
+\begin{subequations}
+  \begin{align}
+    \cons{r}\cons{s} + \cons{s}\cons{r} &= 0 \\
+    \anni{r}\anni{s} + \anni{s}\anni{r} &= 0 \\
+    \cons{r}\anni{s} + \anni{s}\cons{r} &= \delta_{rs},
+  \end{align}
+\end{subequations}
+showing how the construction of the Fock space representation
+embeds antisymmetry at the operator level and not in the wave functions.
+Finally, the second-quantized molecular electronic Hamiltonian in the \acrlong{BO}
+approximation is:\autocite{Helgaker2000-tz, Shavitt2009-mr}
+\begin{equation}\label{eq:sq-hamiltonian}
+  H = \sum_{rs}h_{rs}\cons{r}\anni{s}
+  + \frac{1}{2}\sum_{rstu}g_{rstu}\cons{r}\cons{t}\anni{u}\anni{s}
+  + V_\mathrm{NN}
+\end{equation}
+where the matrix elements are given as integrals over the chosen
+one-particle basis:
+\begin{subequations}
+ \begin{align}
+   h_{rs} &= \int \diff\vect{r} \varphi_r^\dagger(\vect{r}) h(\vect{r})\varphi_s(\vect{r}) \label{eq:one-el}\\
+   g_{rstu} &=
+   \int\diff\vect{r}\int\diff\vect{r}^\prime
+   \Omega_{rs}(\vect{r}) g(\vect{r},\vect{r}^\prime) \Omega_{tu}(\vect{r}^\prime) \label{eq:two-el} \\
+  \Omega_{rs}(\vect{r}) &= \varphi_r^\dagger(\vect{r})\varphi_s(\vect{r}). \label{eq:overlap}
+ \end{align}
+\end{subequations}
+The second-quantized Hamiltonian Eq.~\eqref{eq:sq-hamiltonian} is thus a
+\emph{projected} operator, exact only within the subspace determined by
+the chosen one-particle set.~\autocite{Gross1991-hi, Helgaker2000-tz}
+
+In second quantization, the exact molecular electronic wave function can
+be written as a linear combination of \acrshort{ON} vectors:
+\begin{equation}
+  \ket{\psi} = \sum_{\vect{k}}C_\vect{k}\ket{\vect{k}},
+\end{equation}
+also known at the \gls{FCI} expansion.
+\todo[inline]{Variational property and size-extensivity.}
+Coefficients in the expansion corresponding to the molecular ground
+state can be determined by seeking the lowest eigenvalue of the matrix
+representation of the Hamiltonian in the space of \acrshort{ON} vectors.
+The \acrshort{FCI} method is exact, within the chosen one-particle set,
+but scales exponentially with the number of electrons in the system and
+has thus very limited practical applications.
+Approximations can be tailored by appropriate truncations of the
+$N$-electron space.
+In \emph{single-reference} methods, one first determines an
+approximation to the solution by means of a single \acrshort{ON} vector.
+This \acrshort{ON} vector is treated as the \emph{physical} vacuum:
+vectors in Fock space can be classified according to how different their
+occupation is from the reference, \ie~\emph{by excitation
+level}.\autocite{Helgaker2000-tz, Shavitt2009-mr}
+A truncated expansion can then be constructed by including only those
+\acrshort{ON} vectors that differ from the reference up to a certain
+number of excitations.
+Further details on how the expansions are actually tailored and
+algorithmically implemented can be found elsewhere.\autocite{Helgaker2000-tz}
+In the rest of this Chapter we will describe how a single \acrshort{ON}
+vector approximation can be constructed and how we can improve on it by
+means of \acrlong{CC} and \acrlong{MBPT}.
+
+\section{Mean-Field Theory}\label{sec:mean-field}
+
+Basing ourselves on a variational principle, we seek the best
+description of the many-electron wave function as a single Slater
+determinant. This implies that there is no variational freedom in the
+$N$-electron space. Our variational degrees of freedom reside in
+the one-electron space, \ie in the space of \gslpl{MO}
+$\lbrace\varphi_r\rbrace$, used to build the determinant $\Ket{0}$.
+The one-electron space is developed by considering the behaviour of the
+single electron in the mean field of all the other electrons, and while
+this neglects a smaller part of the interaction energy, the electron
+correlation, it provides a suitable starting point for further
+variational or perturbational treatments to recover more of the
+electron--electron interaction.
+
+\todo[inline]{CURRENTLY VERBATIM FROM MASTER THESIS!!!!}
+
+The problem with introducing more electrons in the theory is that
+electron--electron interactions that were previously small now occur to
+the same order as the kinetic energy and the interaction with the
+nuclear potential. The technique for dealing with this problem is
+well-known from nonrelativistic theory of many-electron systems:
+mean-field theory.
+
+Basing ourselves on a variational principle, we seek the best
+description of the many-electron wave function as a single Slater
+determinant. This implies that there is no variational freedom in the
+$N$-electron space (\ie the subset of Fock space describing $N$ particle
+systems, $\mathcal{F}^N$). Our variational degrees of freedom reside in
+the one-electron space, \ie in the space of Molecular Orbitals,
+$\lbrace\varphi_r\rbrace$, used to build the determinant $\Ket{0}$.
+The one-electron space is developed by considering the behaviour of the
+single electron in the mean field of all the other electrons, and while
+this neglects a smaller part of the interaction energy, the electron
+correlation, it provides a suitable starting point for further
+variational or perturbational treatments to recover more of the
+electron--electron interaction.
+
+To derive a set of relativistic one-particle functions, \ie 4-spinors,
+from a mean-field approach we typically start by making a simple guess
+at these functions and then try to refine them iteratively by applying
+suitable rotations in the entire one-particle space available. This
+iterative procedure normally reaches a stage were further rotations do
+not change the spinors, \ie they are \emph{self-consistent}.
+
+As the molecular 4-spinors $\lbrace\varphi_r\rbrace$ are orthonormal,
+the resulting $N$-electron space is made of normalized determinants.
+Consequently, only norm-conserving transformations of the one-electron
+space underlying $\mathcal{F}^N$ must be considered.\footnote{In
+general, any suitable transformation of the one-electron space must
+enforce \emph{all} the symmetries presented by the system under
+consideration.}
+This restricts our attention to the group of unitary transformations:
+\begin{equation}
+ U^\dagger U = id = UU^\dagger.
+\end{equation}
+
+Any unitary transformation may be parametrized as the exponential of an
+anti-Hermitian operator:\autocite{Arfken2013-pt, Helgaker2000-tz}
+\begin{alignat}{2}
+ U = \exp(-\kappa) \quad&\quad \kappa^\dagger = -\kappa
+\end{alignat}
+the anti-Hermitian operator being given as a linear combination of
+single excitation operators:
+\begin{alignat}{2}\label{eq:orb-rot}
+  \kappa = \sum_{rs} \kappa_{rs} \cons{r}\anni{s} -\kappa^*_{rs}\cons{s}\anni{r}
+= \sum_{rs}\kappa_{rs}q^\dagger_{rs} - \kappa^*_{rs}q_{rs}
+\quad&\quad \kappa_{sr}^* = -\kappa_{rs}
+\end{alignat}
+and the \emph{orbital rotation} parameters $\kappa_{rs}$ were introduced.
+It may be easily verified that the matrix representation of $U$ in the
+$\lbrace\varphi_{r}\rbrace$ basis is given by:
+\begin{equation}
+ \mat{U} = \exp(-\vect{\kappa}).
+\end{equation}
+
+Before proceeding with the minimization procedure, let us rewrite the
+orbital rotation operator in a more convenient notation. As already
+pointed out, orbitals will in general be complex. The usual reduction in
+the parametrization of $\kappa$ in the nonrelativistic regime will not
+be observed. Instead of working with the real and imaginary parts of the
+orbital rotation parameters, we will treat the $\vect{\kappa}$ and
+$\vect{\kappa}^*$ matrices as \emph{independent} variables. Hence,
+defining the following vectors:
+\begin{alignat}{2}
+\vect{K} =
+ \begin{pmatrix}
+  \vect{\kappa} \\
+  \vect{\kappa}^*
+ \end{pmatrix}
+\quad&\quad
+\vect{Q} =
+ \begin{pmatrix}
+  -\vect{q} \\
+  \vect{q}^\dagger
+ \end{pmatrix}
+\end{alignat}
+Eq. \eqref{eq:orb-rot} may be rewritten as:
+\begin{equation}\label{eq:orb-rot-new}
+  \kappa = \frac{1}{2}\left[ \vect{K}^\dagger\vect{Q} - \vect{Q}^\dagger\vect{K}\right]
+\end{equation}
+
+Let us note that the parametrization of $\kappa$ is, in general
+\emph{redundant}, in the sense that it contains certain rotations which
+do not change the energy of the state. Redundant rotations are
+inessential in the optimization of the wave function and may lead to
+numerical instabilities in second order algorithms.\autocite{Helgaker2000-tz,
+Shepard1987-sn}
+It is then mandatory to identify and eliminate them. For a single
+determinant this is an easy task as all redundant rotations satisfy the
+condition:
+\begin{equation}
+ \cons{r}\anni{s}\Ket{0} = 0.
+\end{equation}
+
+For a closed shell only those rotations connecting occupied and virtual
+orbitals will be nonredundant. The redundancy-free orbital rotation
+generator is then:
+\begin{equation}\label{eq:orb-rot-nonred}
+  \kappa = \sum_{ai}\kappa_{ai}\cons{a}\anni{i} -\kappa_{ai}^* \cons{i}\anni{a}
+\end{equation}
+The advantage of this parametrization is that the orbital rotation
+operator $\exp(-\kappa)$ ensures orthonormality of the one-particle
+orbitals without the need to introduce Lagrange multipliers:
+unconstrained optimization approaches can be used.
+
+By means of $\exp(-\kappa)$ the energy is established as a function of
+the orbital rotation parameters:
+\begin{equation}\label{eq:param-energy}
+ E(\vect{K}) = \Braket{0(\vect{K})|H_0|0(\vect{K})} = \Braket{0|\exp(\kappa)H\exp(-\kappa)|0}.
+\end{equation}
+The optimal 4-spinors are then found by minimizing $E(\vect{K})$. The
+Hamiltonian in Eq.~\eqref{eq:param-energy} has the second-quantized
+expression of Eq.~\eqref{eq:sq-hamiltonian} and the integrals are given as
+in Eqs.~\eqref{eq:one-el}-\eqref{eq:overlap}.
+
+We perform a Taylor expansion up to second-order of the right-hand-side
+of Eq.~\eqref{eq:param-energy}, all derivatives are evaluated at the
+current expansion point arbitrarily set at
+$\vect{K}=0$:
+\begin{equation}\label{eq:taylor-exp}
+ E(\vect{K}) = E^{[0]} +\vect{K}^\dagger \mat{E}^{[1]} + \frac{1}{2}\vect{K}^\dagger \mat{E}^{[2]}\vect{K} + O(\vect{K}^3)
+\end{equation}
+where the electronic gradient and Hessian are given as:
+\begin{subequations}
+ \begin{align}
+ \mat{E}^{[1]}
+&=
+\begin{pmatrix}
+ \pderiv{E}{{\vect{\kappa}^*}} \\
+ \pderiv{E}{{\vect{\kappa}}}
+\end{pmatrix}
+\\
+\mat{E}^{[2]}
+&=
+\begin{pmatrix}
+ \frac{\partial^2 E}{\partial\boldsymbol{\kappa}^*\partial\boldsymbol{\kappa}} &
+ \frac{\partial^2 E}{\partial\boldsymbol{\kappa}^*\partial\boldsymbol{\kappa}^*} \\[0.3em]
+ \frac{\partial^2 E}{\partial\boldsymbol{\kappa}\partial\boldsymbol{\kappa}} &
+ \frac{\partial^2 E}{\partial\boldsymbol{\kappa}\partial\boldsymbol{\kappa}^*}
+\end{pmatrix}
+=
+\begin{pmatrix}
+                  \mat{A} & \mat{B} \\
+                  \mat{B}^* & \mat{A}^*
+\end{pmatrix}
+ \end{align}
+\end{subequations}
+
+Expansion of Eq. \eqref{eq:param-energy} in a \gls{BCH}
+commutator series yields:\autocite{Arfken2013-pt}
+\begin{equation}
+ E(\vect{K}) = E(0) + \braket{0|\commutator{\kappa}{H_0}|0} +\frac{1}{2}\braket{0|\commutator{\kappa}{\commutator{\kappa}{H_0}}|0}+O(\vect{\kappa}^3)
+\end{equation}
+
+We will now concentrate only on the expression of the electronic energy
+gradient, the Hessian will be analyzed in Chapter
+\ref{ch:molprop}.
+Differentiating the expression above with respect to $\vect{K}$ gives the gradient as:
+\begin{equation}\label{eq:brillouin}
+ E^{[1]}_{ai} = \Braket{0|\commutator{-\cons{i}\anni{a}}{H_0}|0}.
+\end{equation}
+The derivative has been evaluated at $\vect{K}=0$, \ie the current
+expansion point. The stationarity condition Eq.~\eqref{eq:brillouin} is
+also known as \emph{Brillouin's Theorem}.
+The explicit expression of the electronic energy gradient follows from a
+straightforward application of Wick's theorem:\autocite{Wick1950-iy,
+Shavitt2009-mr}
+\begin{equation}\label{eq:stationary}
+  E^{[1]}_{ai} = -h_{ai} - \sum_{j}[g_{aijj} - g_{ajji}]
+\end{equation}
+
+We define the \emph{Fock matrix} $\mat{F}$ with elements:
+\begin{equation}
+  F_{rs} = h_{rs} + \sum_{j} [g_{rsjj} - g_{rjjs}]
+\end{equation}
+apart from the one-electron term, whose expression was given in Eq.~\eqref{eq:one-el},
+the Fock matrix features two types of two-electron
+terms Eq.~\eqref{eq:two-el}.
+The former is due to the Coulomb interaction, while the second is the
+\emph{exchange} contribution arising because of antisymmetry in the
+many-electron wave function, in accordance with Pauli principle.
+
+In terms of the Fock matrix, the stationarity condition Eq.~\eqref{eq:stationary} reads:
+\begin{equation}
+  f_{ai} = h_{ai} + \sum_{j}[g_{aijj} - g_{ajji}] = 0
+\end{equation}
+meaning that the Fock matrix is block-diagonal in the basis of the
+optimal MO 4-spinors, \ie~the occupied-virtual and virtual-occupied
+blocks are zero. Nothing is said about the diagonal blocks (\ie~
+Occupied-Occupied and Virtual-Virtual) of this matrix, as the
+corresponding orbital rotations are redundant. This arbitrariness may be
+used to ease the wave function optimization, \eg~applying level-shifts,
+or to require that the eigenvalues of the Fock matrix have particular
+properties.\autocite{Szabo1989-vl, McWeeny1992-oj, Helgaker2000-tz}
+
+For example, we can fix the diagonal blocks in order to reinterpret the
+eigenvalues of the Fock matrix as ionization potentials and electron
+affinities (Koopmans' Theorem).
+The optimal 4-spinor rotation will bring the Fock matrix to diagonal
+form, what is called its \emph{canonical representation}. Thus the
+optimization of the one-particle space may be recast as a
+diagonalization problem:
+\begin{equation}\label{eq:fock-eq}
+ \mat{F}\vect{\varphi} = \vect{\varphi}\mat{\epsilon}
+\end{equation}
+where the eigenvalues of the Fock matrix are called \emph{orbital
+energies}. The theory exposed above goes under several names that we
+will use interchangeably in the remainder of this thesis. Apart from
+mean-field theory the names \gls{SCF} and
+\gls{HF} theory are also in common usage.
+
+%%%% Introduce DFT and KS-DFT
+An interesting approach to the quantum mechanical description of
+many-electron systems is based on the idea that it should be possible to
+find a quantum theory that solely refers to observable quantities.
+\gls{DFT}, instead of relying on a wave
+function, is based on the electron density.
+The Hohenberg--Kohn theorems\autocite{Hohenberg1964-wo, Eschrig2012-as}
+and their relativistic counterparts\autocite{Rajagopal1973-ns,
+Dreizler2012-ay} establish the energy as a \emph{functional} of the
+density:
+\begin{equation}
+ E = E[\rho]
+\end{equation}
+
+In the \gls{KS} approach to \acrshort{DFT}, the density of the system of interest
+is supposed to be equal to that of a fictitious noninteracting
+system.\autocite{Kohn1965-hg}
+The energy is then a sum of five terms:
+\begin{equation}\label{eq:ks-energy}
+ E[\rho] = T_s[\rho] + V_\mathrm{ext}[\rho] + J[\rho] + E_\mathrm{xc}[\rho] + V_\mathrm{NN}
+\end{equation}
+the first four of which  are functionals of the density. The first term
+$T_s[\rho]$ is the kinetic energy of the fictitious noninteracting
+system represented by a single Slater determinant $\ket{\widetilde{0}}$:
+\begin{equation}
+ T_s[\rho] = \sum_i \braket{\widetilde{0}|h_\mathrm{D}(i)|\widetilde{0}}.
+\end{equation}
+The second and third term in Eq.~\eqref{eq:ks-energy} represent,
+respectively, the classical interaction of the electrons with the
+external potential (including the nuclear attraction potential) and the
+classical Coulomb interaction of the density with itself:
+\begin{subequations}
+ \begin{align}
+   V_\mathrm{ext}[\rho] &= \int\diff\vect{r} V_\mathrm{ext}(\vect{r})\rho(\vect{r}) \\
+   J[\rho] &=
+   \frac{1}{2}\int\diff\vect{r}\int\diff\vect{r}^\prime
+   \rho(\vect{r})g(\vect{r}, \vect{r}^\prime)\rho(\vect{r}^\prime)
+ \end{align}
+\end{subequations}
+The fourth term in Eq.~\eqref{eq:ks-energy} is the
+\gls{XC} functional and is a great unknown. This term
+contains all two-electron interactions except for the classical
+$J[\rho]$ term, \ie~it contains the effects of electron exchange and
+electron correlation. It also corrects for the self-interaction present
+in the classical Coulomb term and for the error in the evaluation of the
+kinetic energy.
+A number of approximations exist for this functional.
+In the \gls{LDA}, the functional is expressed in terms of an
+\emph{energy density} $e_\mathrm{xc}$ which is a local function of the
+density:
+\begin{equation}
+ E^\mathrm{LDA}_\mathrm{xc}[\rho] = \int e_\mathrm{xc}(\rho(\vect{r}))\diff\vect{r}.
+\end{equation}
+In the \gls{GGA}, the energy
+density is a local function of both the density and the norm of its
+gradient $\zeta(\vect{r})=\vect{\nabla}\rho\cdot \vect{\nabla}\rho$:
+\begin{equation}
+ E^\mathrm{GGA}_\mathrm{xc}[\rho] = \int e_\mathrm{xc}(\rho(\vect{r}),\zeta(\vect{r}))\diff\vect{r}.
+\end{equation}
+Finally, in \emph{hybrid} functionals, some proportion of the
+\acrshort{HF} exchange is included:
+\begin{equation}
+ E^\mathrm{hybrid}_\mathrm{xc}[\rho] = E^\mathrm{GGA}_\mathrm{xc}[\rho] + \gamma E^\mathrm{HF}_\mathrm{x}[\rho]
+\end{equation}
+The last term of Eq.~\eqref{eq:ks-energy} is the nuclear--nuclear repulsion energy.
+
+In complete analogy with the mean-field theory exposed above,
+\citeauthor{Saue2002-cu} have proposed a unitary parametrization of the
+\acrshort{KS} state in terms of orbital rotations:\autocite{Saue2002-cu}
+\begin{equation}
+ \ket{\widetilde{0}} = \exp(-\kappa)\ket{0}.
+\end{equation}
+The density matrix is then a function of the orbital rotation parameters:
+\begin{equation}
+ \widetilde{D}_{rs}(\vect{\kappa}) = \braket{0|\exp(\kappa)\cons{r}\anni{s}\exp(-\kappa)|0}
+\end{equation}
+thus establishing the density $\rho$ and the norm of its gradient
+$\zeta$ as functions of the orbital rotation parameters.
+Using the overlap distributions defined in Eq.~\eqref{eq:overlap}:
+\begin{subequations}
+ \begin{align}
+   \rho(\vect{r},\vect{\kappa}) &= \sum_{rs}\widetilde{D}_{rs}(\vect{\kappa})\Omega_{rs}(\vect{r}) \\
+   \vect{\nabla} \rho(\vect{r},\vect{\kappa}) &= \sum_{rs}\widetilde{D}_{rs}(\vect{\kappa})\vect{\nabla}\Omega_{rs}(\vect{r}).
+ \end{align}
+\end{subequations}
+The electronic energy gradient leads to the definition of the Fock-like
+\emph{KS} matrix as:
+\begin{equation}\label{eq:ksmatrix}
+  f_{rs} = h_{rs} + \sum_{j}[g_{rsjj} - \gamma g_{rjjs}] + v_{\mathrm{xc};rs}
+\end{equation}
+the actual form of $v_{\mathrm{xc};rs}$ can be found in
+\noparcite[refs.][]{Saue2002-cu}
+and \noparcite[][]{Salek2005-at}
+
+In the \acrshort{KS} approach to \acrshort{DFT}, the one-particle space
+is then optimized in complete analogy with \acrshort{HF} theory, \ie~by
+diagonalization of the \acrshort{KS} matrix:
+\begin{equation}
+ \mat{F}\vect{\varphi} = \vect{\varphi}\mat{\epsilon}
+\end{equation}
+unlike \acrshort{HF} theory, the \acrshort{KS} approach enables us to
+recover some electron correlation, albeit through the use of \emph{ad
+hoc} \acrshort{XC} functionals.\autocite{Koch2015-wq}
 
@@
 $\mathcal{M}$ is the truncation level.
-One of the main strengths of the \acrshort{CC} model is its size extensivity
+One of the main strengths of the \acrshort{CC} model is its size-extensivity
 which stems directly from the exponential nature of the wave
@@
 The linked and unlinked formulations are completely equivalent. However,
-the former is size extensive term-by-term.
+the former is size-extensive term-by-term.
 Use of similarity transformed operators is ubiquitous when dealing with
@@
 hermiticity, that the bare operators might have possessed.
-Similarity-transformed operators can be expanded in a commutator series,
-also known as the \gls{BCH} expansion:
+Similarity-transformed operators can be expanded in a \acrshort{BCH} commutator series:
 \begin{equation}
@@
 
-Having summarized the essential points of \acrshort{CC} theory and having
-introduced the relevant notation, we can proceed to analyze its coupling
-with quantum/classical polarizable Hamiltonians.
-
-\todo[inline]{Provide a primer here so that the Chapter on
-solvation+correlation is not superlong}
-
-\section{Many-Body Perturbation Theory}\label{sec:MBPT}
-
-\section{Iterative Approximations: CC2 and CC3}\label{sec:ccn}
-
-\section{Noniterative Triples Corrections}\label{sec:ccsd-t}
+\section{Many-Body Perturbation Theory}\label{sec:mbpt}
+\section{Approximate Coupled Cluster Methods}\label{sec:cc-approximate}

### 2016-09-02 Chapters/SolvationAndCorrelation.tex
Work on MQM chapter

@@
 \acrshort{MBPT} in terms of the effective \acrshort{CC} Lagrangian, an approach
-guaranteeing term-by-term size extensivity.
+guaranteeing term-by-term size-extensivity.
 
@@
 corrections will be derived.
-
-\pagebreak
 
@@
 The use of a \acrshort{CC} Ansatz for the wave function ensures term-by-term
-size extensivity of the energy at all orders.~\autocite{Helgaker2000-tz}
+size-extensivity of the energy at all orders.~\autocite{Helgaker2000-tz}
 The fluctuation potential $\Phi$ is assumed as the perturbation, hence

### 2016-09-03 Chapters/ContinuumSolvationModels.tex
Work on MQM chapter
Save the day's work. Chapter 1 almost done

@@
 group that are particularly interesting.
-
-\pagebreak
 
@@
 formulate methods in quantum chemistry as the unconstrained minimization
-of a Lagrangian.\autocite{Helgaker1982-xl, Arponen1983-qt,
-Helgaker1988-to, Helgaker1989-wl, Koch1990-vr, Kvaal2013-jr}
-Let us call $\mathcal{E}(\vect{\eta})$ the definition of the energy
-for the given quantum chemical method.
-The \emph{amplitudes} $\vect{\eta}$ are determined by a set of
-stationarity conditions $\mat{\Omega}(\vect{\eta}) = 0$.
-Direct minimization of $\mathcal{E}(\vect{\eta})$ is thus a
-\emph{constrained} optimization.
-Introducing a set of Lagrange \emph{multipliers} $\bar{\vect{\eta}}$
-lets us recast the problem as an unconstrained minimization of the
-Lagrangian:
-\begin{equation}
-  \mathcal{L}(\vect{\eta}, \bar{\vect{\eta}}) =
-  \mathcal{E}(\vect{\eta}) +
-  \scalprod[W]{\bar{\vect{\eta}}}{\mat{\Omega}(\vect{\eta})}
-\end{equation}
-where the governing equations are obtained by differentiation with
-respect to both sets of parameters:
-\begin{subequations}
-  \begin{align}
-    \pderiv{\mathcal{L}(\vect{\eta}, \bar{\vect{\eta}})}{\bar{\vect{\eta}}}
-    &= \mat{\Omega}(\vect{\eta}) = 0 \\
-    \pderiv{\mathcal{L}(\vect{\eta}, \bar{\vect{\eta}})}{\vect{\eta}}
-    &= \pderiv{\mathcal{E}(\vect{\eta})}{\vect{\eta}} +
-    \scalprod[W]{\bar{\vect{\eta}}}{\pderiv{\mat{\Omega(\vect{\eta})}}{\vect{\eta}}}
-    = 0
-  \end{align}
-\end{subequations}
+of a Lagrangian.
 Due to the  variational property of the classical polarizable functional
@@
 multipliers.
+
+\todo[inline]{Provide an example with single-reference theories.}
+As an example, let us consider the single-reference \acrshort{HF} and
+\acrshort{KS}-\acrshort{DFT} methods.

### 2016-09-03 Chapters/ContributionsSummary.tex
Save the day's work. Chapter 1 almost done

@@
                  For the full, functional example visit
-                 \url{http://pcmsolver.readthedocs.io/en/latest/users/C-example.html}},
+                 \url{http://pcmsolver.readthedocs.io/en/v1.1.5/users/C-example.html}},
                  label={lst:C-host}
@@
                  For the full, functional example visit
-                 \url{http://pcmsolver.readthedocs.io/en/latest/users/fortran-example.html}},
+                 \url{http://pcmsolver.readthedocs.io/en/v1.1.5/users/fortran-example.html}},
                  label={lst:Fortran-host}

### 2016-09-03 Chapters/QuantumMechanics.tex
Save the day's work. Chapter 1 almost done

--- 2016-09-01 Chapters/QuantumMechanics.tex
@@
 
-\chapter{A Quick Tour of Quantum Chemistry}\label{ch:QM}
+\chapter{A Quick Tour of Molecular Electronic Structure Theory}\label{ch:QM}
 
-\begin{itemize}
-  \item nonrelativistic and relativistic Hamiltonians used in molecular
-    quantum mechanics
-  \item \gls{BO} approximation and the concept of \gls{PES}
-  \item Fock space
-  \item single-reference methods: \gls{HF} and \gls{KS} \gls{DFT}
-  \item the \gls{CC}  model
-\end{itemize}
+The purpose of quantum chemistry is to provide models based on first
+principles that can help understand and predict molecular properties.
+In quantum chemistry, we apply physical models based on
+quantum many-body methods to molecular systems, employ their
+mathematical realizations and devise computable approximations.
+The central idea is, in fact, to be able to obtain an algorithmic
+implementation of the methods that can be applied to interesting
+chemical systems.
 
-\pagebreak
+This Chapter presents a brief overview of molecular electronic structure
+theory with particular emphasis on the methods used in this thesis.
+Section \ref{sec:mqm} is a general introduction to the methods of
+molecular quantum mechanics, relativistic or nonrelativistic.
+I will provide a brief introduction to the language of second
+quantization, used throughout the thesis.
+Section \ref{sec:mean-field} presents the mean-field approximation to
+the molecular electronic structure problem, the workhorse of quantum
+chemistry.
+Sections \ref{sec:coupled-cluster}--\ref{sec:cc-approximate} are
+concerned with the \gls{CC} and \gls{MBPT} methods for the inclusion of
+electron correlation.
 
-\section{Molecular Quantum Mechanical Hamiltonians}\label{sec:hamiltonians}
+\section{Molecular Quantum Mechanics}\label{sec:mqm}
 
-Konishi and Paffuti\autocite{Konishi2009-zb}
-Dyall's book\autocite{Dyall2007-tu}
-Reiher's book\autocite{Reiher2014-cp}
-The Pink Book\autocite{Helgaker2000-tz}
-Shavitt's book\autocite{Shavitt2009-mr}
+Quantum mechanics is the theory describing the motion and interactions
+of microscopic particles. In quantum theory every observable of the
+system is represented mathematically by an operator $O$ in the
+appropriate Hilbert space of \emph{wave functions}.
+A wave function $\psi$ is the mathematical object describing the state of the
+system. In the usual Copenhagen interpretation of quantum mechanics, the
+modulus square of the wave function $|\psi|^2$ provides the key to
+predicting experimentally measurable quantities.\autocite{Konishi2009-zb}
+The wave function $\psi$ is the solution to the Schr\"{o}dinger equation:
+\begin{equation}\label{eq:schrodinger}
+  H\psi = E\psi
+\end{equation}
+where $H$ is the \emph{Hamiltonian} operator and $E$ is the energy of
+the system. The Schr\"{o}dinger equation is an eigenvalue equation for
+the Hamiltonian operator and $\psi$ is thus an
+eigenfunction.\autocite{Arfken2013-pt}
 
-\section{Single-Reference Methods}
+Molecular quantum mechanics is concerned with the microscopic motion of
+nuclei and electrons in molecules and is thus an intrinsically many-body
+theory. The problem is clearly very complicated to solve, as the number of
+interactions to be considered is large.
+Since the nuclei are much heavier than the electrons, the \gls{BO} or
+clamped-nuclei approximation is typically assumed. The nuclei are fixed
+in a specific configuration (called a \emph{molecular geometry}) and
+treated as static electric sources.\autocite{Born1927-ce}
+This separation of motions results in an electron-only Hamiltonian and
+an electronic wave function that depends \emph{parametrically} on the
+nuclear positions. The eigenvalue of this Hamiltonian, also dependent on
+the nuclear positions, is the \gls{PES} of the system, that is the
+potential function in which the nuclei move.\autocite{Szabo1989-vl,
+McWeeny1992-oj}
+In first quantization, the general expression for the clamped-nuclei,
+$N$-electron, molecular electronic Hamiltonian is:
+\begin{alignat}{2}\label{eq:fq-hamiltonian}
+  H = \sum_{i=1}^N h(\vect{r}_i)
+  + \frac{1}{2}\sum_{i\neq j}^N g(\vect{r}_i, \vect{r}_j)
+  + V_\mathrm{NN},
+  \quad&
+  V_\mathrm{NN} =
+  \frac{1}{2}\sum_{A \neq B}^{N_\mathrm{nuclei}}\frac{Z_A Z_B}{|\vect{R}_A - \vect{R}_B|}
+\end{alignat}
+The last term is the nuclear repulsion energy, a constant shift of the
+energy scale in the \acrshort{BO} approximation.
+As noted by \citeauthor{Saue2011-qg}, the same general expression for
+the clamped-nuclei Hamiltonian Eq.~\eqref{eq:fq-hamiltonian} is valid in
+the nonrelativistic and relativistic no-pair
+regimes.\autocite{Sucher1980-vf, Saue2011-qg}
+In the nonrelativistic domain one defines the one- and two-electron
+operators as:\autocite{Szabo1989-vl, McWeeny1992-oj, Konishi2009-zb}
+\begin{subequations}
+ \begin{align}
+   h(\vect{r}_i) &=
+   -\frac{1}{2}\nabla^2_i + V_\mathrm{Ne}(\vect{r}_i)
+   = -\frac{1}{2}\nabla^2_i - \sum_{A=1}^{N_\mathrm{nuclei}}\frac{Z_A}{|\vect{R}_A -
+   \vect{r}_i|} \label{eq:h} \\
+   g(\vect{r}_i, \vect{r}_j) &= \frac{1}{|\vect{r}_i - \vect{r}_j|},
+   \label{eq:g-coulomb}
+ \end{align}
+\end{subequations}
+while in the relativistic regime with the \gls{DC}
+approximation:\autocite{Dyall2007-tu, Konishi2009-zb, Reiher2014-cp}
+\begin{subequations}
+ \begin{align}
+   h(\vect{r}_i) &=
+   \begin{pmatrix}
+     V_{\mathrm{Ne}}(\vect{r}_i)I_2 & c \vect{\sigma}\cdot\vect{p}_i \\
+     \vect{\sigma}\cdot\vect{p}_i & (V_{\mathrm{Ne}}(\vect{r}_i)-2\si{\electronmass}c^2)I_2
+   \end{pmatrix} \label{eq:h-dirac} \\
+   g(\vect{r}_i, \vect{r}_j) &= \frac{I_4\cdot I_4}{|\vect{r}_i - \vect{r}_j|}
+   \label{eq:g-r-coulomb}
+ \end{align}
+\end{subequations}
+and the operators are implicitly assumed to embed the proper no-pair,
+positive-energy manifold projectors.\autocite{Sucher1980-vf}
+Since these are fermionic Hamiltonians, the spin-statistics theorem
+dictates the corresponding eigenfunctions to be
+antisymmetric.
+Using the expansion theorem,\autocite{Gross1991-hi} we thus seek the wave
+function for the $N$-electron system as a linear combination of
+$N$-electron \emph{Slater determinants}.
+
+It the language of second quantization, we assume that a, not
+necessarily complete, set of orthonormal one-particle states is
+available. These are usually called \glspl{MO}.
+In general, we would like to embed as many of the symmetries of the
+system into the one-particle basis. This is accomplished by requiring
+the one-particle space to be the common eigenbasis for a complete set of
+suitable, commuting one-particle operators.\autocite{Gross1991-hi}
+Any $N$-particle state can be constructed by distributing particles
+into one-particle states and accordingly labelling them by the number of
+particles per one-particle states present.
+This is the so-called \gls{ON} vector representation:
+\begin{equation}
+  \ket{\vect{n}} = \ket{n1,n2,\ldots},
+\end{equation}
+and the span of all \acrshort{ON} vectors is called the \emph{Fock
+space} $F^N$, a space isomorphic to the Hilbert space $H^N$ of
+$N$-particle wave functions.
+To preserve consistency, the \emph{vacuum} state with no particles:
+\begin{equation}
+ \ket{\mathrm{vac}} = \ket{0,0,\ldots}
+\end{equation}
+is included in the construction of Fock space.
+Any $N$-particle state can be generated by application of
+\emph{creation} operators $\cons{i}$. Creation operators increase the
+occupation number in the one-particle state $i$ by $1$, up to a phase.
+Conversely, \emph{annihilation} operators $\anni{i}$, defined as the Hermitian
+conjugates of the creation operators, decrease occupation in state $i$
+by $1$, again up to a phase. Clearly, annihilation of any one-particle
+state in the vacuum results yields $0$.
+Pauli's exclusion principle is enforced by the \emph{canonical
+anticommutation relations}:
+\begin{subequations}
+  \begin{align}
+    \cons{r}\cons{s} + \cons{s}\cons{r} &= 0 \\
+    \anni{r}\anni{s} + \anni{s}\anni{r} &= 0 \\
+    \cons{r}\anni{s} + \anni{s}\cons{r} &= \delta_{rs},
+  \end{align}
+\end{subequations}
+showing how the construction of the Fock space representation
+embeds antisymmetry at the operator level and not in the wave functions.
+Finally, the second-quantized molecular electronic Hamiltonian in the \acrlong{BO}
+approximation is:\autocite{Helgaker2000-tz, Shavitt2009-mr}
+\begin{equation}\label{eq:sq-hamiltonian}
+  H = \sum_{rs}h_{rs}\cons{r}\anni{s}
+  + \frac{1}{2}\sum_{rstu}g_{rstu}\cons{r}\cons{t}\anni{u}\anni{s}
+  + V_\mathrm{NN}
+\end{equation}
+where the matrix elements are given as integrals over the chosen
+one-particle basis:
+\begin{subequations}
+ \begin{align}
+   h_{rs} &= \int \diff\vect{r} \varphi_r^\dagger(\vect{r}) h(\vect{r})\varphi_s(\vect{r}) \label{eq:one-el}\\
+   g_{rstu} &=
+   \int\diff\vect{r}\int\diff\vect{r}^\prime
+   \Omega_{rs}(\vect{r}) g(\vect{r},\vect{r}^\prime) \Omega_{tu}(\vect{r}^\prime) \label{eq:two-el} \\
+  \Omega_{rs}(\vect{r}) &= \varphi_r^\dagger(\vect{r})\varphi_s(\vect{r}). \label{eq:overlap}
+ \end{align}
+\end{subequations}
+The second-quantized Hamiltonian Eq.~\eqref{eq:sq-hamiltonian} is thus a
+\emph{projected} operator, exact only within the subspace determined by
+the chosen one-particle set.~\autocite{Gross1991-hi, Helgaker2000-tz}
+
+In second quantization, the exact molecular electronic wave function can
+be written as a linear combination of \acrshort{ON} vectors:
+\begin{equation}\label{eq:fci}
+  \ket{\psi} = \sum_{\vect{k}}C_\vect{k}\ket{\vect{k}},
+\end{equation}
+also known at the \gls{FCI} expansion.
+Among the many properties the exact wave function enjoys, the
+\emph{variational} property and \emph{size-extensivity} are the ones
+that are almost always embedded into approximate wave function ans\"{a}tze.\autocite{Helgaker2000-tz}
+According to the former, the energy is stable with respect to orthogonal
+variations of the wave function. In practice this guarantees that
+minimization with respect to the parameters of a well-crafted ansatz
+ensures convergence from above to the exact energy.\autocite{Szabo1989-vl, McWeeny1992-oj, Helgaker2000-tz}
+A wave function ansatz enjoying the latter property guarantees that the
+energy of a system made of noninteracting subsystems is equal to the sum of
+the energies of the isolated subsystems.\autocite{Helgaker2000-tz, Nooijen2005-ux}
+
+Coefficients in the \acrshort{FCI} expansion in Eq. \eqref{eq:fci} corresponding to the molecular ground
+state can be determined by seeking the lowest eigenvalue of the matrix
+representation of the Hamiltonian in the space of \acrshort{ON} vectors.
+The \acrshort{FCI} method is exact, within the chosen one-particle set,
+but scales exponentially with the number of electrons in the system and
+has thus very limited practical applications.
+Approximations can be tailored by appropriate truncations of the
+$N$-electron space.
+In \emph{single-reference} methods, one first determines an
+approximation to the solution by means of a single \acrshort{ON} vector.
+This \acrshort{ON} vector is treated as the \emph{physical} vacuum:
+vectors in Fock space can be classified according to how different their
+occupation is from the reference, \ie~\emph{by excitation
+level}.\autocite{Helgaker2000-tz, Shavitt2009-mr}
+A truncated expansion can then be constructed by including only those
+\acrshort{ON} vectors that differ from the reference up to a certain
+number of excitations.
+Further details on how the expansions are actually tailored and
+algorithmically implemented can be found elsewhere.\autocite{Helgaker2000-tz}
+In the rest of this Chapter we will describe how a single \acrshort{ON}
+vector approximation can be constructed, how we can improve on it by
+means of \acrlong{CC} and \acrlong{MBPT} and how the properties of the
+exact wave function are embedded into these ans\"{a}tze.
+
+\section{Mean-Field Theory}\label{sec:mean-field}
+
+The idea behind mean-field theory is simple: seek the best description
+of the many-electron wave function using just one Slater determinant.
+How to direct this search towards the "best"? We will use the
+variational principle and optimize the degrees of freedom in our
+single-determinantal trial wave function accordingly.
+It is evident that there is no variational freedom in the
+$N$-electron space. All variational degrees of freedom reside in
+the one-electron space, \ie in the space of \glspl{MO}
+$\lbrace\varphi_r\rbrace$, used to build the determinant $\Ket{0}$.
+The one-electron \glspl{MO} are thus optimized considering each single
+electron in the mean field generated by the motion of all other
+electrons, hence the name of the method. Clearly this neglects
+parts of the electron--electron interaction energy, the \emph{electron
+correlation}. However, in most cases, it provides a suitable starting
+approximation for more sophisticated many-body treatments that recover
+larger portions of the correlation energy.
+
+Starting from an initial guess for the one-particle states, we
+iteratively refine them by applying suitable rotations.
+The iterations are carried out until a suitable convergence criterion
+has been met, \ie the orbitals are \emph{self-consistent}.
+The unitary rotation operator can be expressed as the exponential of a
+one-electron, anti-Hermitian operator:
+\begin{alignat}{3}
+  U = \expo{-\kappa},
+  \quad&
+  \kappa = \sum_{rs}
+  [\kappa_{rs} \cons{r}\anni{s} - \kappa^*_{rs}\cons{s}\anni{r}],
+  \quad&
+  \kappa^\dagger = -\kappa.
+\end{alignat}
+The energy of the system is now a function of the rotation parameters
+$\kappa_{rs}, \kappa^*_{rs}$ and setting its gradient $\mat{E}^{[1]}$ to
+zero will locate a stationary point. The gradient of the electronic
+energy can be identified by comparing the Taylor and \gls{BCH}
+commutator expansions of the Hamiltonian expectation value.
+For a closed-shell this yields \emph{Brillouin's theorem}:
+\begin{equation}\label{eq:brillouin}
+  E^{[1]}_{ai} = -f_{ai}
+ = h_{ai} + \sum_{j}[g_{aijj} - g_{ajji}] = 0,
+\end{equation}
+with $\mat{F}$ being the \emph{Fock matrix}.
+The two-electron terms appearing in the Fock matrix are called the
+Coulomb and exchange integrals, respectively.
+The stationarity condition implies that the Fock matrix is
+block-diagonal in the basis of the optimal \glspl{MO}.
+In the \emph{canonical representation} one seeks the set of orbitals
+that make the Fock matrix diagonal. Hence the optimization problem is
+equivalent to diagonalization of an effective one-electron operator:
+\begin{equation}\label{eq:fock-eq}
+ \mat{F}\vect{\varphi} = \vect{\varphi}\mat{\epsilon}
+\end{equation}
+where the eigenvalues of the Fock matrix are called \emph{orbital
+energies}.
+
+A different approach is taken in \gls{DFT}. The electron density is
+taken as the central quantity in the theory, instead of the wave
+function.
+The Hohenberg--Kohn theorems\autocite{Hohenberg1964-wo, Eschrig2012-as}
+and their relativistic counterparts\autocite{Rajagopal1973-ns,
+Dreizler2012-ay} establish the energy as a \emph{functional} of the
+density.
+The \gls{KS} approach to \acrshort{DFT} takes us a step further by
+assuming the density of the system to be equal to that of a fictitious,
+noninteracting system once again described by a single Slater
+determinant.\autocite{Kohn1965-hg}
+We rewrite the energy as a sum of five terms:
+\begin{equation}\label{eq:ks-energy}
+ E[\rho] = T_s[\rho] + V_\mathrm{ext}[\rho] + J[\rho] + E_\mathrm{xc}[\rho] + V_\mathrm{NN}
+\end{equation}
+the first four of which are functionals of the density. The first term
+$T_s[\rho]$ is the kinetic energy of the fictitious noninteracting
+system.
+The second and third terms in Eq. \eqref{eq:ks-energy} represent,
+respectively, the classical interaction of the electrons with the
+external potential (including the nuclear attraction potential) and the
+classical Coulomb interaction of the density with itself:
+\begin{subequations}
+ \begin{align}
+   V_\mathrm{ext}[\rho] &= \int\diff\vect{r} V_\mathrm{ext}(\vect{r})\rho(\vect{r}) \\
+   J[\rho] &=
+   \frac{1}{2}\int\diff\vect{r}\int\diff\vect{r}^\prime
+   \rho(\vect{r})g(\vect{r}, \vect{r}^\prime)\rho(\vect{r}^\prime)
+ \end{align}
+\end{subequations}
+The fourth term in Eq. \eqref{eq:ks-energy} is the \gls{XC} functional.
+This term accounts for the nonclassical part of the two-electron
+interaction: electron exchange and electron correlation.
+Moreover, it also corrects for the self-interaction present in $J[\rho]$
+and the error introduced in the evaluation of the kinetic energy.
+The exact form of this functional is not known and a number of
+parametrized approximations exist.\autocite{Koch2015-wq}
+In the \gls{LDA}, the functional is expressed in terms of an
+\emph{energy density} $e_\mathrm{xc}$ which is a local function of the
+density:
+\begin{equation}
+ E^\mathrm{LDA}_\mathrm{xc}[\rho] = \int e_\mathrm{xc}(\rho(\vect{r}))\diff\vect{r}.
+\end{equation}
+In the \gls{GGA}, the energy density is a local function of both the
+density and the norm of its gradient
+$\zeta(\vect{r})=\vect{\nabla}\rho\cdot \vect{\nabla}\rho$:
+\begin{equation}
+ E^\mathrm{GGA}_\mathrm{xc}[\rho] = \int e_\mathrm{xc}(\rho(\vect{r}),\zeta(\vect{r}))\diff\vect{r}.
+\end{equation}
+Finally, in \emph{hybrid} functionals, some proportion of the
+\acrshort{HF} exchange is included:
+\begin{equation}
+ E^\mathrm{hybrid}_\mathrm{xc}[\rho] = E^\mathrm{GGA}_\mathrm{xc}[\rho] + \gamma E^\mathrm{HF}_\mathrm{x}[\rho]
+\end{equation}
+The \acrshort{KS}-\acrshort{DFT} state is optimized by the same
+iterative process described above.
+Repeated diagonalizations of the \acrshort{KS} matrix are carried out
+until self-consistency is reached.
+The \acrshort{KS} matrix contains the one-electron \acrshort{XC}
+potential in addition to the other terms contained in the Fock matrix:
+\begin{equation}\label{eq:ksmatrix}
+  f_{rs} = h_{rs} + \sum_{j}[g_{rsjj} - \gamma g_{rjjs}] + v_{\mathrm{xc};rs}
+\end{equation}
 
@@
 
-\todo[inline]{MOVED FROM CHAPTER 4 HERE: READJUST!}
-
-The analyses and derivations in this Chapter are preeminently based
-on the \acrlong*{CC} wave function Ansatz.
-We will assume a closed-shell, \acrshort{HF} single reference function and
-construct our correlated wave function as an exponential mapping:
+In the \acrlong*{CC} method we construct our correlated wave function as
+an exponential mapping on top of a single reference
+function:\autocite{Crawford2000-ey, Helgaker2000-tz, Bartlett2007-mz,
+Shavitt2009-mr}
 \begin{equation}
-  \ket{\mathrm{CC}} = \expo{T}\ket{\mathrm{HF}}.
+  \ket{\mathrm{CC}} = \expo{T}\ket{\mathrm{HF}},
 \end{equation}
+out of convenience we assume, here and in the following, that the
+reference function is a closed-shell \acrshort{HF} wave function.
 The \emph{cluster} operator appearing in the exponential is given as:
@@
 
-The \acrshort{CC} method is a \emph{projective} method. Chosen the truncation
-level $\mathcal{M}$, one projects the nonrelativistic Schrödinger
-equation for this Ansatz on the excitation manifold which comprises the
-reference function and all possible excited determinants, up to the
-chosen truncation order:
+The \acrshort{CC} method is not a variational but a \emph{projective} method.
+Chosen the truncation level $\mathcal{M}$, one projects the
+nonrelativistic Schrödinger equation for this ansatz on the excitation
+manifold which comprises the reference function and all possible excited
+determinants, up to the chosen truncation order:
 \begin{subequations}
   \begin{align}
-    \braket{\mathrm{HF} | H_0\expo{T} | \mathrm{HF}} &= E_\mathrm{CC} \\
-    \braket{\mathrm{exc} | H_0\expo{T} | \mathrm{HF}} &=
-    E_\mathrm{CC}\braket{\mathrm{exc} | \expo{T} | \mathrm{HF}}
+    \braket{\mathrm{HF} | \expo{-T}H\expo{T} | \mathrm{HF}} &= E_\mathrm{CC} \\
+    \braket{\mathrm{exc} | \expo{-T}H\expo{T} | \mathrm{HF}} &= 0
   \end{align}
@@
 space compatible with the excitation level truncation.
-This is the \emph{unlinked} form of the \acrshort{CC}
-equations.~\autocite{Helgaker2000-tz} It is usually more convenient to
-perform a \emph{similarity transformation} of the Hamiltonian operator,
-leading to the \emph{linked} form of the \acrshort{CC} equations:
-\begin{subequations}
-  \begin{align}
-    \braket{\mathrm{HF} | \expo{-T}H_0\expo{T} | \mathrm{HF}} &= E_\mathrm{CC} \\
-    \braket{\mathrm{exc} | \expo{-T}H_0\expo{T} | \mathrm{HF}} &= 0
-  \end{align}
-\end{subequations}
-The linked and unlinked formulations are completely equivalent. However,
-the former is size extensive term-by-term.
+Notice that we have also performed a \emph{similarity transformation} of
+the Hamiltonian operator. This leads to the \emph{linked} form of the
+\acrshort{CC} equations, which is manifestly size-extensive term-by-term.
 Use of similarity transformed operators is ubiquitous when dealing with
@@
 hermiticity, that the bare operators might have possessed.
-Similarity-transformed operators can be expanded in a commutator series,
-also known as the \gls{BCH} expansion:
-\begin{equation}
-  \overline{O} =
-  \sum_{n=0}^{\infty}\frac{1}{n!}[A, B]^n
-\end{equation}
-where $[A,B]^n$ is the $n$-fold nested commutator. It can be shown that
-the \acrshort{BCH} expansion of the similarity-transformed Hamiltonian
-\emph{exactly} truncates after the four-fold nested commutator, greatly
-simplifying algebraic derivations and manipulations.
+Similarity-transformed operators can be expanded in a \acrshort{BCH}
+commutator series.
+It can be shown that the \acrshort{BCH} expansion of the
+similarity-transformed Hamiltonian \emph{exactly} truncates after the
+four-fold nested commutator, greatly simplifying algebraic derivations
+and manipulations.
 
-Introducing the \acrlong*{MP} partitioning of the Hamiltonian leads to
+Introducing the \gls{MP} partitioning of the Hamiltonian leads to
 further insight:
-\begin{equation}
- H_0 = F + \Phi = \sum_P\epsilon_P\cons{P}\anni{P} + (g- V_\mathrm{HF})
-\end{equation}
+\begin{alignat}{2}
+ H = F + \Phi = \sum_r\epsilon_r\cons{r}\anni{r} + (g- V_\mathrm{HF}),
+ \quad&
+ V_\mathrm{HF} = \sum_{i}[g_{rsii} - g_{riis}]\cons{r}\anni{s}
+\end{alignat}
 where $F$ if the Fock matrix and $\Phi$ is the fluctuation potential.
@@
 For the fluctuation potential the similarity transformation truncates
-after the fifth term, including up to four-fold nested commutator.
+after the fifth term, including up to four-fold nested commutators.
 
-As formulated so far, the \acrshort{CC} method is not variational. It is
-however possible to introduce a fully variational Lagrangian that
-leads to the same amplitudes equations. The Lagrangian is built by
-treating the amplitude equations as constraints in the optimization of
-the \acrshort{CC} energy:\autocite{Helgaker1982-xl, Arponen1983-qt,
+As formulated so far, the \acrshort{CC} method is not variational.
+It is however possible to introduce a fully variational Lagrangian that
+leads to the same amplitudes equations.\autocite{Helgaker1982-xl, Arponen1983-qt,
 Helgaker1988-to, Helgaker1989-wl, Koch1990-vr, Kvaal2013-jr}
+Let us call $\mathcal{E}(\vect{\eta})$ the definition of the energy
+for the given quantum chemical method.
+The \emph{amplitudes} $\vect{\eta}$ are determined by a set of
+stationarity conditions $\mat{\Omega}(\vect{\eta}) = 0$.
+Direct minimization of $\mathcal{E}(\vect{\eta})$ is thus a
+\emph{constrained} optimization.
+Introducing a set of Lagrange \emph{multipliers} $\bar{\vect{\eta}}$
+lets us recast the problem as an unconstrained minimization of the
+Lagrangian:
+\begin{equation}
+  \mathcal{L}(\vect{\eta}, \bar{\vect{\eta}}) =
+  \mathcal{E}(\vect{\eta}) +
+  \scalprod[W]{\bar{\vect{\eta}}}{\mat{\Omega}(\vect{\eta})}
+\end{equation}
+where the governing equations are obtained by differentiation with
+respect to both sets of parameters:
+\begin{subequations}
+  \begin{align}
+    \pderiv{\mathcal{L}(\vect{\eta}, \bar{\vect{\eta}})}{\bar{\vect{\eta}}}
+    &= \mat{\Omega}(\vect{\eta}) = 0 \\
+    \pderiv{\mathcal{L}(\vect{\eta}, \bar{\vect{\eta}})}{\vect{\eta}}
+    &= \pderiv{\mathcal{E}(\vect{\eta})}{\vect{\eta}} +
+    \scalprod[W]{\bar{\vect{\eta}}}{\pderiv{\mat{\Omega(\vect{\eta})}}{\vect{\eta}}}
+    = 0
+  \end{align}
+\end{subequations}
+The \acrshort*{CC} Lagrangian can thus be expressed as:
 \begin{equation}\label{eq:cc-lagrangian}
@@
     &=
-    \braket{\mathrm{HF} | \overline{H_0} | \mathrm{HF}}
+    \braket{\mathrm{HF} | \overline{H} | \mathrm{HF}}
     + \sum_{i=1}^{\mathcal{M}}\sum_{\mu_i}
-    \tbar{\mu_i}\braket{\mu_i | \overline{H_0} | \mathrm{HF}} \\
+    \tbar{\mu_i}\braket{\mu_i | \overline{H} | \mathrm{HF}} \\
     &=
@@
 \end{equation}
-where the shorthand notation for the Lagrangian multiplier state was
-introduced:
+where $E_0 = \sum_{i}\epsilon_i$ is the sum of the one-electron orbital
+energies and the shorthand notation for the Lagrangian multiplier state (left
+\acrshort*{CC} state) was introduced:
 \begin{equation}
-  \bra{\tbar{j}} = \sum_{\mu_j}\tbar{\mu_j}\bra{\mu_j}
+  \bra{\tbar{j}} = \sum_{\mu_j}\tbar{\mu_j}\bra{\mu_j}.
 \end{equation}
-and the \acrshort{MP}-partitioned form of the Hamiltonian was also exploited.
 Differentiation of the Lagrangian with respect to the multipliers
@@
 respect to the amplitudes leads to the governing equations for the
-multipliers. While these are not needed for the evaluation of the
+multipliers:
+\begin{subequations}
+  \begin{align}
+   \tampEq{\mu_i}(\tamp{}, \tbar{}, \p)  &=
+   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
+             = 0 \label{eq:cc-amplitudes}\\
+   \tbarEq{\mu_i}(\tamp{}, \tbar{}, \p)
+    &=
+    \denom{\mu_i}\tbar{\mu_i} +
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
+    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} |
+    \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}}
+             = 0 \label{eq:cc-multipliers}
+  \end{align}
+\end{subequations}
+While the multipliers are not needed for the evaluation of the
 \acrshort{CC} energy, their calculation is mandatory when expectation values
@@
 
-Having summarized the essential points of \acrshort{CC} theory and having
-introduced the relevant notation, we can proceed to analyze its coupling
-with quantum/classical polarizable Hamiltonians.
+\section{Many-Body Perturbation Theory}\label{sec:mbpt}
 
-\todo[inline]{Provide a primer here so that the Chapter on
-solvation+correlation is not superlong}
+An alternative method to recover the correlation energy missing in the
+mean-field approximation is to employ perturbation
+theory.\autocite{Konishi2009-zb}
+The single determinant optimized in the mean-field approximation is used
+as zeroth-order wave function and the perturbation series is developed
+considering the fluctuation potential in the \acrlong*{MP} partitioning
+of the Hamiltonian as the perturbation.\autocite{McWeeny1992-oj, Helgaker2000-tz}
+Notice that the partitioning of the Hamiltonian is not unique and
+other choices are available.\autocite{Shavitt2009-mr}
 
-\section{Many-Body Perturbation Theory}\label{sec:MBPT}
+As shown by \citeauthor{Koch1997-nm}, the \acrshort*{MBPT} series can be
+derived from a \acrshort*{CC} ansatz for the wave function.
+This ensures term-by-term size-extensivity of the energy at all
+orders.\autocite{Helgaker2000-tz}
+The governing equations
+\eqref{eq:cc-amplitudes}--\eqref{eq:cc-multipliers} are expanded in
+terms of the fluctuation potential. Energy corrections are identified
+from the corresponding order expansion of the \acrshort*{CC} Lagrangian
+and retaining only terms complying to the $2n+1$ and $2n+2$ rules.
+A superscript index in square brackets - $[i]$ -  will denote an $i$-th
+order contribution.
 
-\section{Iterative Approximations: CC2 and CC3}\label{sec:ccn}
+It is clear from the stationarity conditions that $\tamp{\mu_i}^{[0]} = 0, \,\forall i$ and
+$\tbar{\mu_i}^{[0]} = 0, \, \forall i$.
+Thus cluster operators can be of order 1 and higher. We introduce the
+following notation for the $i$-th order cluster operators and
+multipliers states:
+\begin{alignat}{2}
+  T_p^{[i]} = \sum_{\mu_p}\tamp{\mu_p}^{[i]}\cluster{\mu_p},
+  \quad&
+  \bra{\tbar{p}^{[i]}} = \sum_{\mu_p}\tbar{\mu_p}^{[i]}\bra{\mu_p}
+\end{alignat}
 
-\section{Noniterative Triples Corrections}\label{sec:ccsd-t}
+To first order, one has:
+\begin{equation}\label{eq:1st-order}
+  \begin{aligned}
+  \denom{\mu_2}\tamp{\mu_2}^{[1]}
+  &= - \braket{\mu_2 | \Phi | \mathrm{HF}}
+  = - \braket{\mathrm{HF} | \commutator{\Phi}{\cluster{\mu_2}} |
+  \mathrm{HF}} \\
+  &= \denom{\mu_2}\tbar{\mu_2}^{[1]}
+  \end{aligned}
+\end{equation}
+which shows that singles, due to Brillouin's condition, and triples
+amplitudes are of second and higher order.
+To second order, one obtains:
+\begin{equation}\label{eq:2nd-order}
+  \begin{aligned}
+  \denom{\mu_i}\tamp{\mu_i}^{[2]}
+  &= - \braket{\mu_i | \commutator{\Phi}{T_2^{[1]}} | \mathrm{HF}}
+  = - \braket{\tbar{2}^{[1]} | \commutator{\Phi}{\cluster{\mu_i}} |
+  \mathrm{HF}} \\
+  &= \denom{\mu_i}\tbar{\mu_i}^{[2]},\quad i = 1, 2, 3.
+  \end{aligned}
+\end{equation}
+According to the $2n+1$ and $2n+2$ rules, energy corrections up to fifth
+order can be obtained from amplitudes and multipliers up to second
+order.
+\todo[inline]{CHECK THE LOGIC OF THIS STATEMENT}
+Since the triples first appear to second order the \gls{CCSD}
+energy is correct to third order in the fluctuation potential.
+The triples contributions to the fourth and fifth order corrections to
+the energy are given as:
+\begin{subequations}
+  \begin{align}
+    E_T^{[4]} &=
+ \braket{\tbar{2}^{[1]}| \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+ \label{eq:4th-order-triples} \\
+    E_T^{[5]} &=
+    \braket{\tbar{1}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+  + \braket{\tbar{2}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+  \nonumber \\
+  &+ \braket{\tbar{3}^{[2]} | \commutator{\Phi}{T_2^{[2]}} | \mathrm{HF}}
+  + \braket{\tbar{3}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+  \nonumber \\
+    &+
+    \frac{1}{2}\braket{\tbar{3}^{[2]} |
+    \commutator{\commutator{\Phi}{T_2^{[1]}}}{T_2^{[1]}}
+    | \mathrm{HF}}.
+ \label{eq:5th-order-triples}
+  \end{align}
+\end{subequations}
+These expressions form the basis for the development of reduced-scaling
+\acrshort*{CC} schemes were the double and triple excitation are treated
+approximately. These are the subject of the next Section.
+
+\section{Approximate Coupled Cluster Methods}\label{sec:cc-approximate}
+
+The \acrshort*{CCSD} scheme, which scales as $N^6$, is really successful
+in recovering a large portion of the correlation energy, but as hinted
+at in the previous Section is only correct to third order in
+perturbation theory.
+Triple excitations are needed to get higher accuracy, but the full
+\gls{CCSDT} model, with its steep $N^8$ scaling, is unfeasible for all
+but the smallest systems.
+One way around this problem is to exploit the perturbation theory
+expression for the fourth and fifth order energy corrections of Eqs.
+\eqref{eq:4th-order-triples}--\eqref{eq:5th-order-triples} to guide the
+construction of approximate schemes for the inclusion of higher
+excitations. These schemes can either be \emph{iterative} or
+\emph{noniterative}.
+
+In the iterative schemes, one approximates the amplitude equations for
+the higher excitation to be correct up to a given order in perturbation
+theory.
+For the development of iterative schemes, we assume that $T_1 = O(0)$
+and use $T_1$-transformed operators:
+\begin{equation}
+  \check{O} = \expo{-T_1}O\expo{T_1}.
+\end{equation}
+Since $T_1$ is a one-electron operator, the $T_1$-transformation does
+not affect the particle rank of the electronic
+Hamiltonian.\autocite{Koch1994-vr, Helgaker2000-tz}
+The assumption that $T_1$ is of zeroth order is contrary to the usual
+\acrshort*{MP} approach but highlights the role of the singles
+amplitudes as orbital optimization parameters.
+In the \gls{CC2} method, the amplitudes equations are:\autocite{Christiansen1995-vy, Helgaker2000-tz}
+\begin{subequations}
+  \begin{align}
+    \denom{\mu_1}\tamp{\mu_1} &+ \braket{\mu_1 | \check{\Phi} +
+    \commutator{\check{\Phi}}{Q_2} | \mathrm{HF}} = 0 \\
+    \denom{\mu_2}\tamp{\mu_2} &+ \braket{\mu_2 | \check{\Phi} | \mathrm{HF}} = 0
+  \end{align}
+\end{subequations}
+and are obtained from the \acrshort{CCSD} equations by retaining only
+first order terms in the doubles equations.
+The approximate doubles cluster operator was introduced:
+\begin{equation}
+  Q_2 = - \sum_{\mu_2}
+  \denom{\mu_2}^{-1}\braket{\mu_2 | \check{\Phi} | \mathrm{HF}}\cluster{\mu_2}
+\end{equation}
+and we shall refer to it as quasi-$T_2$.
+The \acrshort*{CC2} Lagrangian is constructed based on the approximate
+equations obtained for the singles and
+doubles:\autocite{Christiansen1995-vy, Helgaker2000-tz}
+\begin{equation}\label{eq:cc2-lagrangian}
+  \begin{aligned}
+  \mathcal{L}_\mathrm{CC2}(\tamp{}, \tbar{})
+  &=
+  E_0
+  + \sum_{i=1}^{\mathcal{2}}\tbar{\mu_i}\denom{\mu_i}\tamp{\mu_i}
+  + \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{Q_2} | \mathrm{HF}} \\
+  &+ \sum_{\mu_1}\tbar{\mu_1}\braket{\mu_1 |
+  \check{\Phi} + \commutator{\check{\Phi}}{Q_2}
+  | \mathrm{HF}} \\
+  &+ \sum_{\mu_2}\tbar{\mu_2}\braket{\mu_2 | \check{\Phi} | \mathrm{HF}}
+  \end{aligned}
+\end{equation}
+\todo[inline]{How does CC2 scale in the iterative step?}
+
+The \gls{CC3} model is constructed analogously, the amplitudes
+equations:\autocite{Christiansen1995-pv, Koch1997-nm, Helgaker2000-tz,
+Smith2005-an}
+\begin{subequations}
+  \begin{align}
+    \denom{\mu_1}\tamp{\mu_1} &+ \braket{\mu_1 |
+    \check{\Phi} + \commutator{\check{\Phi}}{T_2}
+    + \commutator{\check{\Phi}}{Q_3}
+    | \mathrm{HF}} = 0 \\
+    \denom{\mu_2}\tamp{\mu_2} &+ \braket{\mu_2 |
+    \check{\Phi} + \commutator{\check{\Phi}}{T_2}
+    + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2}
+    + \commutator{\check{\Phi}}{Q_3}
+    | \mathrm{HF}} = 0 \\
+    \denom{\mu_3}\tamp{\mu_3} &+ \braket{\mu_3 |
+    \commutator{\check{\Phi}}{T_2}
+    | \mathrm{HF}} = 0
+  \end{align}
+\end{subequations}
+are obtained from the \acrshort*{CCSDT} equations by retaining only
+second order terms in the triples equation.
+The approximate triples cluster operator was introduced:
+\begin{equation}
+  Q_3 = - \sum_{\mu_3}
+  \denom{\mu_3}^{-1}\braket{\mu_3 | \commutator{\check{\Phi}}{T_2} | \mathrm{HF}}\cluster{\mu_3}
+\end{equation}
+and we shall refer to it as quasi-$T_3$.
+The \acrshort*{CC3} Lagrangian is constructed based on the approximate
+equations obtained for the singles, doubles and
+triples:\autocite{Christiansen1995-pv, Koch1997-nm, Helgaker2000-tz, Smith2005-an}
+\begin{equation}\label{eq:cc3-lagrangian}
+  \begin{aligned}
+  \mathcal{L}_\mathrm{CC3}(\tamp{}, \tbar{})
+  &=
+  E_0
+  + \sum_{i=1}^{\mathcal{3}}\tbar{\mu_i}\denom{\mu_i}\tamp{\mu_i}
+  + \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
+  &+ \sum_{\mu_1}\tbar{\mu_1}\braket{\mu_1 |
+  \check{\Phi} + \commutator{\check{\Phi}}{T_2} + \commutator{\check{\Phi}}{Q_3}
+  | \mathrm{HF}} \\
+  &+ \sum_{\mu_2}\tbar{\mu_2}\braket{\mu_2 |
+  \check{\Phi} + \commutator{\check{\Phi}}{T_2}
+  + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{Q_3}
+  | \mathrm{HF}} \\
+  &+ \sum_{\mu_3}\tbar{\mu_3}\braket{\mu_3 | \commutator{\check{\Phi}}{T_2} | \mathrm{HF}}
+  \end{aligned}
+\end{equation}
+\todo[inline]{How does CC3 scale in the iterative step?}
+
+\todo[inline]{REFERENCES MISSING HERE!!!!!}
+Finally, within noniterative methods the fourth and fifth order triples
+energy corrections are added on top of the \acrshort*{CCSD} energy.
+The \acrshort*{CCSD} converged singles and doubles amplitudes are used
+instead of their perturbational counterparts. A $*$ superscript will be
+used to denote the converged parameters.
+Since the left and right \acrshort*{CC} states differ, symmetric and
+asymmetric variants of the noniterative corrections exist.
+The $\Lambda$\acrshort*{CCSD}[T] and \acrshort*{CCSD}[T] methods only
+include the fourth order correction Eq. \eqref{eq:4th-order-triples}.
+The former uses both the left and right state, while the latter uses
+only the right state and can be obtained from the former by replacing
+the multipliers with the amplitudes:\autocite{Koch1997-nm,
+Eriksen2015-il}
+\begin{subequations}
+  \begin{align}
+    E_{\Lambda\text{[T]}} &= {}^* E^{[4]}_{\Lambda\text{[T]}} =
+    \braket{\tbar{2}^{*}| \commutator{\Phi}{T_3^{*}} | \mathrm{HF}} \\
+    E_{\text{[T]}} &= {}^* E^{[4]}_{\text{[T]}} = \braket{\tamp{2}^{*}|
+    \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}.
+  \end{align}
+\end{subequations}
+From Eqs. \eqref{eq:1st-order}--\eqref{eq:2nd-order}, one can derive
+which \acrshort*{MBPT} terms are included in the corrections:
+\begin{equation}
+  \begin{aligned}
+  {}^* E^{[4]}_{\Lambda\text{[T]}}
+  &=
+    \braket{\tbar{2}^{[1]}| \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+  + \braket{\tbar{2}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}} \\
+  &+ \braket{\tbar{3}^{[2]} | \commutator{\Phi}{T_2^{[2]}} | \mathrm{HF}}
+  + O(6)
+  \end{aligned}
+\end{equation}
+and similarly for ${}^* E^{[4]}_{\text{[T]}}$.
+
+The $\Lambda$\acrshort*{CCSD}(T) and \acrshort*{CCSD}(T) methods also
+include the first of the fifth order triples corrections in Eq.
+\eqref{eq:5th-order-triples}:
+\begin{subequations}
+ \begin{align}
+  E_{\Lambda(\text{T})} = {}^* E^{[4]}_{\Lambda\text{[T]}} + {}^* E^{[5]}_{\Lambda\text{(T)}}
+  &=
+    \braket{\tbar{1}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
+    +
+    \braket{\tbar{2}^{*}| \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
+    \\
+  E_{(\text{T})} = {}^* E^{[4]}_{\text{[T]}} + {}^* E^{[5]}_{\text{(T)}}
+  &=
+    \braket{\tamp{1}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
+    +
+    \braket{\tamp{2}^{*}| \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
+ \end{align}
+\end{subequations}
+once again, the symmetric method can be obtained from the asymmetric
+variant by replacing multipliers with amplitudes in the left state.
+The \acrshort*{MBPT} terms included in
+${}^* E^{[5]}_{\Lambda\text{(T)}}$ can be seen to be:
+\begin{equation}
+  {}^* E^{[5]}_{\Lambda\text{(T)}}
+  =
+  \braket{\tbar{1}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+  + O(6)
+\end{equation}
+and similarly for ${}^* E^{[5]}_{\text{(T)}}$.
+
+\todo[inline]{COST IN THE NONITERATIVE STEP. CHECK REFERENCES!!!!!!!!}

### 2016-09-03 Chapters/SolvationAndCorrelation.tex
Save the day's work. Chapter 1 almost done

@@
 \acrshort{MBPT} for quantum/classical polarizable Hamiltonians.
-I will exploit the \acrshort{CC} Ansatz for the wave function and formulate
+I will exploit the \acrshort{CC} ansatz for the wave function and formulate
 \acrshort{MBPT} in terms of the effective \acrshort{CC} Lagrangian, an approach
@@
 expansions truncate at the third term. Moreover, normal-ordering of the
-operators has been introduced.~\autocite{Crawford2000-ey, Shavitt}
+operators has been introduced.~\autocite{Crawford2000-ey, Shavitt2009-mr}
 This is equivalent to applying a shift to the correlation part of the
@@
 used.
-The use of a \acrshort{CC} Ansatz for the wave function ensures term-by-term
+The use of a \acrshort{CC} ansatz for the wave function ensures term-by-term
 size-extensivity of the energy at all orders.~\autocite{Helgaker2000-tz}

### 2016-09-04 Chapters/QuantumMechanics.tex
Minor adjustments to text

@@
  \begin{align}
-   h(\vect{r}_i) &=
-   \begin{pmatrix}
-     V_{\mathrm{Ne}}(\vect{r}_i)I_2 & c \vect{\sigma}\cdot\vect{p}_i \\
-     \vect{\sigma}\cdot\vect{p}_i & (V_{\mathrm{Ne}}(\vect{r}_i)-2\si{\electronmass}c^2)I_2
-   \end{pmatrix} \label{eq:h-dirac} \\
+   h(\vect{r}_i) &= (\mat{\beta} - I_4)\si{\electronmass}c^2 +
+   c(\mat{\alpha}_i\cdot\vect{p}_i) + V_{\mathrm{Ne}}(\vect{r}_i)I_4 \\
    g(\vect{r}_i, \vect{r}_j) &= \frac{I_4\cdot I_4}{|\vect{r}_i - \vect{r}_j|}
@@
 \end{subequations}
+The Dirac matrices, expressed in terms of the vector $\mat{\sigma}$
+collecting the three Pauli spin matrices, were introduced:
+\begin{alignat}{2}
+  \mat{\alpha}_i =
+  \begin{pmatrix}
+    0_2 & \mat{\sigma}_i \\
+    \mat{\sigma}_i & 0_2
+  \end{pmatrix},
+  \quad&
+  \mat{\beta} =
+  \begin{pmatrix}
+    I_2 & 0_2 \\
+    0_2 & -I_2
+  \end{pmatrix},
+\end{alignat}
 and the operators are implicitly assumed to embed the proper no-pair,
@@
 \end{equation}
-\todo[inline]{How does CC2 scale in the iterative step?}
+The \acrshort*{CC2} methods brings the scaling of the \acrshort*{CCSD}
+method down to $N^5$ in the iterative step.
 
@@
 \end{equation}
-\todo[inline]{How does CC3 scale in the iterative step?}
+The \acrshort*{CC3} methods brings the scaling of the \acrshort*{CCSDT}
+method down to $N^7$ in the iterative step.
 
-\todo[inline]{REFERENCES MISSING HERE!!!!!}
-Finally, within noniterative methods the fourth and fifth order triples
-energy corrections are added on top of the \acrshort*{CCSD} energy.
-The \acrshort*{CCSD} converged singles and doubles amplitudes are used
-instead of their perturbational counterparts. A $*$ superscript will be
-used to denote the converged parameters.
+Finally, one can construct \emph{noniterative} methods, where the
+expensive $N^7$ step for the inclusion of connected triples is
+calculated as a one-off correction on top of the converged
+\acrshort*{CCSD} energy.
+The expression for the correction is based on the fourth and fifth order
+triples energy corrections Eqs.
+\eqref{eq:4th-order-triples}--\eqref{eq:5th-order-triples}, but is
+computed using the converged \acrshort*{CCSD} singles and doubles
+amplitudes,\autocite{Urban1985-xp, Raghavachari1989-bn, Bartlett1990-lk,
+Watts1993-ff, Kucharski1998-qq, Kucharski1998-oi, Crawford1998-vj,
+Taube2008-xv, Taube2008-oq} here and in the following denoted by a $*$
+superscript.
 Since the left and right \acrshort*{CC} states differ, symmetric and
 asymmetric variants of the noniterative corrections exist.
-The $\Lambda$\acrshort*{CCSD}[T] and \acrshort*{CCSD}[T] methods only
-include the fourth order correction Eq. \eqref{eq:4th-order-triples}.
+
+The $\Lambda\text{CCSD}[\text{T}]$\autocite{Eriksen2015-il} and
+$\text{CCSD}[\text{T}]$\autocite{Urban1985-xp} methods only include the
+fourth order correction Eq. \eqref{eq:4th-order-triples}.
 The former uses both the left and right state, while the latter uses
@@
 \end{subequations}
-From Eqs. \eqref{eq:1st-order}--\eqref{eq:2nd-order}, one can derive
-which \acrshort*{MBPT} terms are included in the corrections:
-\begin{equation}
-  \begin{aligned}
-  {}^* E^{[4]}_{\Lambda\text{[T]}}
-  &=
-    \braket{\tbar{2}^{[1]}| \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
-  + \braket{\tbar{2}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}} \\
-  &+ \braket{\tbar{3}^{[2]} | \commutator{\Phi}{T_2^{[2]}} | \mathrm{HF}}
-  + O(6)
-  \end{aligned}
-\end{equation}
-and similarly for ${}^* E^{[4]}_{\text{[T]}}$.
-
-The $\Lambda$\acrshort*{CCSD}(T) and \acrshort*{CCSD}(T) methods also
+The $\Lambda\text{CCSD}(\text{T})$\autocite{Kucharski1998-qq,
+Kucharski1998-oi, Crawford1998-vj} and
+$\text{CCSD}(\text{T})$\autocite{Raghavachari1989-bn} methods also
 include the first of the fifth order triples corrections in Eq.
@@
 variant by replacing multipliers with amplitudes in the left state.
-The \acrshort*{MBPT} terms included in
-${}^* E^{[5]}_{\Lambda\text{(T)}}$ can be seen to be:
+
+Results from noniterative methods can be rationalized within the
+framework of \acrshort*{MBPT}.\autocite{Koch1997-nm, Eriksen2014-gy,
+Eriksen2015-il, Kristensen2016-od}
+From Eqs. \eqref{eq:1st-order}--\eqref{eq:2nd-order}, one can derive
+which \acrshort*{MBPT} terms are included in the corrections:
+\begin{equation}
+  \begin{aligned}
+  {}^* E^{[4]}_{\Lambda\text{[T]}}
+  &=
+    \braket{\tbar{2}^{[1]}| \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+  + \braket{\tbar{2}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}} \\
+  &+ \braket{\tbar{3}^{[2]} | \commutator{\Phi}{T_2^{[2]}} | \mathrm{HF}}
+  + O(6)
+  \end{aligned}
+\end{equation}
+and similarly for ${}^* E^{[4]}_{\text{[T]}}$.
+The \acrshort*{MBPT} terms included in ${}^*
+E^{[5]}_{\Lambda\text{(T)}}$ can be seen to be:
 \begin{equation}
@@
 and similarly for ${}^* E^{[5]}_{\text{(T)}}$.
-
-\todo[inline]{COST IN THE NONITERATIVE STEP. CHECK REFERENCES!!!!!!!!}

### 2016-09-04 Chapters/SolvationAndCorrelation.tex
Minor adjustments to text

@@
 
-\epigraph{Potius sero quam numquam}{--- \textsc{Titus Livius}, \textit{Ab Urbe Condita}}
-\epigraph{\textitalian{Non temerai i terrori della notte non temerai il terrore}}{--- \textsc{CCCP}}
+\begin{epigraphs}
+\qitem{Potius sero quam numquam}{--- \textsc{Titus Livius}, \textit{Ab Urbe Condita}}
+\qitem{\textitalian{Non temerai i terrori della notte non temerai il terrore}}{--- \textsc{CCCP}}
+\end{epigraphs}
 

### 2016-09-06 Chapters/ContinuumSolvationModels.tex
Chaps. 1, 2, and 5 are complete

@@
 }{
-    --- \textsc{E. Montale}, \textit{Falsetto}}
+    --- \textsc{Eugenio Montale}, \textit{Falsetto}}
 
@@
 Hamiltonian.
-\todo[inline]{What is a nonpolarizable model? What's a polarizable
-model? HAS TO BE MENTIONED HERE!}
-\todo[inline]{What is the nature of the interaction? Mention that we
-only include polarization, but in a mutual manner.}
 
@@
 the averaged, continuous response functions of the solvent.
+This is the physical setting of the \gls{PCM} and is schematically
+depicted in Figure \ref{fig:IEF}.
+The solute molecule is enclose in a \emph{cavity} inside the
+structureless continuum representing the solvent and the basic energetic
+quantity is a \emph{free energy}:\autocite{Tomasi1994-gv,
+Amovilli1998-pv, Tomasi2005-vm}
+\begin{equation}\label{eq:free-energy}
+ G = G_\mathrm{cav} + G_\mathrm{el} + G_\mathrm{dis} + G_\mathrm{rep} + G_\mathrm{Mm}(T).
+\end{equation}
+The decomposition of the solute free energy takes into account the
+different physical origins of the different solute-solvent interactions.
+$G_\mathrm{cav}$ is the cavity formation free energy, a constant as long
+as the solute geometry does not change.\autocite{Tomasi1994-gv,
+Floris1997-tu}
+The second term $G_\mathrm{el}$ is the most important one in the
+framework of continuum models. It accounts for the mutual electrostatic
+polarization of solute and solvent, see Section \ref{sec:IEF}.
+The third $G_\mathrm{dis}$ and fourth $G_\mathrm{rep}$ terms account for
+dispersion and repulsion interactions,
+respectively.\autocite{Amovilli1997-fo, Weijo2010-jy}
+Finally, $G_\mathrm{Mm}(T)$ accounts for the collective molecular
+motions of the solute-solvent system. It explicitly depends on the
+temperature $T$ and can be calculated from the solution partition
+function, see \noparcite[ref.][]{Amovilli1998-pv} for further details.
 
-\todo[inline]{Introduction of the cavity. PCM as a problem in classical
-electrostatics}
+In this thesis, we focused exclusively on the purely electrostatic
+component of the interaction energy, completely neglecting the other
+terms.
+While a balanced description of solvation chemistry certainly demands
+a proper treatment of all solute-solvent interactions, especially when
+non homogeneous environments are studied,\autocite{Mozgawa2014-ad,
+Mozgawa2016-wf} it is also true that the electrostatic component is the
+most numerically relevant and the most sensitive to changes in the
+system studied and in the level of quantum mechanical description.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=.5\textwidth]{IEF}
+  \caption[The physical setting of the polarizable continuum model.]{
+  The physical setting of the \gls{PCM}. The molecular solute,
+  represented by its, possibly quantum mechanical, charge density
+  $\rho_\mathrm{i}$ is enclosed in a \emph{cavity} $\Omegai$.
+  The boundary of the cavity $\Gamma\equiv\partial\Omegai$ is a
+  continuously differentiable 2-manifold.
+  We assume the material properties of the cavity to be those of vacuum,
+  hence characterized by the Green's function
+  $\Gi=\frac{1}{|\vect{r}-\vect{r}^\prime|}$.
+  The cavity is carved out of an infinite, structureless, continuum
+  characterized by its Green's function $\Ge$ and fully covering the
+  external subdomain $\Omegae$.
+  }
+  \label{fig:IEF}
+\end{figure}
 
@@
 
-The physical setting of the \gls{PCM} is depicted in Figure \ref{fig:IEF}.
 The molecular solute, represented by its charge density
@@
 characterized by the bulk properties of the solvent.
-The continuum fully spans the external subdomain $\Omegae$.
+The continuum fully spans the external subdomain $\Omegae$, see Fig. \ref{fig:IEF}.
 The source charge density is not assumed to be continuous: both
@@
 
-This is a well-known problem in electrostatics~\autocite{Jackson1999, Vanderlinde2005-gf}
-and is called a \emph{transmission} problem in the mathematical
-literature on \acp{PDE}.~\autocite{Hackbusch1995-uq, Sauter2011-an}
+Formulated as such, the \gls*{PCM} is a well-known problem in classical
+electrostatics: find the electrostatic potential in space generated by
+a charge distribution confined in a cavity in a \emph{polarizable}
+continuous medium.\autocite{Jackson1999, Vanderlinde2005-gf}
+In the mathematical literature on \acp{PDE}, it is usually referred to as a
+\emph{transmission} problem.\autocite{Hackbusch1995-uq, Sauter2011-an}
+
 Given the partition of Euclidean space $\mathbb{R}^3$ outlined above,
@@
 \gls{BIE}.
-
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=.5\textwidth]{IEF}
-  \caption[The physical setting of the polarizable continuum model.]{
-  The physical setting of the \gls{PCM}. The molecular solute,
-  represented by its, possibly quantum mechanical, charge density
-  $\rho_\mathrm{i}$ is enclosed in a \emph{cavity} $\Omegai$.
-  The boundary of the cavity $\Gamma\equiv\partial\Omegai$ is a
-  continuously differentiable 2-manifold.
-  We assume the material properties of the cavity to be those of vacuum,
-  hence characterized by the Green's function
-  $\Gi=\frac{1}{|\vect{r}-\vect{r}^\prime|}$.
-  The cavity is carved out of an infinite, structureless, continuum
-  characterized by its Green's function $\Ge$ and fully covering the
-  external subdomain $\Omegae$.
-  }
-  \label{fig:IEF}
-\end{figure}
 
@@
 The finite element discretization is usually carried out either by the
-\emph{collocation} or by the \emph{Galerkin} approach.\footnote{
+\emph{collocation} or by the \emph{Galerkin} approach.
 Indeed, it is possible to show that discretization by collocation
@@
 choice of \emph{trial} and \emph{test}
-spaces.~\autocite{Hackbusch1995-uq, Ern2004-oo}}
+spaces.~\autocite{Hackbusch1995-uq, Ern2004-oo}
 This step transforms the original integral equation into a system of
@@
 
-\todo[inline]{Provide an example with single-reference theories.}
 As an example, let us consider the single-reference \acrshort{HF} and
-\acrshort{KS}-\acrshort{DFT} methods.
+\acrshort{KS}-\acrshort{DFT} methods introduced in Section
+\ref{sec:mean-field}.
+First of all, we need the appropriate \acrshort{SCF} energy functional
+\emph{in vacuo}.
+The \acrshort{AO} basis set expansion of the \glspl{MO} lets us define
+the one-electron \acrshort{AO} density matrix as:
+\begin{equation}
+  D_{\mu\nu} = \sum_{i=1}^{N_O}C_{\mu i}C^\dagger_{\nu i}
+\end{equation}
+where $C_{\alpha r}$ are the \acrshort*{MO} expansion coefficients and
+$N_O$ is the total number of occupied orbitals.
+The \acrshort*{SCF} energy functional \emph{in vacuo} can now be
+expressed as:
+\begin{equation}\label{eq:scf-functional}
+  \mathcal{E}(\mat{C}) = \Tr \mat{h}\mat{D} + \frac{1}{2}\Tr
+  \mat{G}^\gamma(\mat{D})\mat{D} + E_\mathrm{xc}[\rho(\vect{r})] +
+  V_\mathrm{NN},
+\end{equation}
+and has to be optimized under the \acrshort*{MO} orthonormality
+constraint, easily introduced by means of Lagrangian multipliers:
+\begin{equation}\label{eq:scf-lagrangian}
+  \mathcal{L}(\mat{C}, \mat{\epsilon}) = \mathcal{E}(\mat{D}) -
+  \sum_{ij}\epsilon_{ji}(\braket{\phi_i|\phi_j} - \delta_{ij})
+\end{equation}
+The various contributions are expressed in terms of \acrshort*{AO} basis
+integral and density matrices:
+\begin{subequations}
+  \begin{align}
+  h_{\mu\nu} = &= \int \diff\vect{r}
+  \chi_\mu^\dagger(\vect{r})h(\vect{r})\chi_\nu(\vect{r}) \label{eq:ao-one-el} \\
+  G_{\mu\nu}^\gamma &=
+  \sum_{\kappa\lambda}D_{\kappa\lambda}(g_{\mu\nu\kappa\lambda} -
+  \gamma g_{\mu\lambda\kappa\nu}) \label{eq:two-el-matrix} \\
+  g_{\mu\nu\kappa\lambda} &=
+  \int\diff\vect{r}\int\diff\vect{r}^\prime
+  \Omega_{\mu\nu}(\vect{r}) g(\vect{r},\vect{r}^\prime)
+  \Omega_{\kappa\lambda}(\vect{r}^\prime) \label{eq:ao-two-el} \\
+  E_\mathrm{xc}[\rho(\vect{r})] &=
+  \int\diff\vect{r}
+  v_\mathrm{xc}(\vect{r})\Tr\mat{\Omega}(\vect{r})\mat{D} \label{eq:ao-xc} \\
+  \Omega_{\mu\nu}(\vect{r}) &= \chi_\mu^\dagger(\vect{r})\chi_\nu(\vect{r}).
+  \label{eq:ao-overlap}
+  \end{align}
+\end{subequations}
+The quantum/classical polarizable \acrshort*{SCF} functional is obtained by adding the
+polarization functional to Eq. \eqref{eq:scf-functional}:
+\begin{equation}\label{eq:polarizable-scf-functional}
+  \mathcal{L}_\mathrm{eff}(\mat{C}, \mat{\epsilon}, \s) =
+  \mathcal{L}(\mat{C}, \mat{\epsilon}) +
+  \frac{1}{2}{}^t\p\V\p + {}^t\p\Tr\mathbfup{s}\mat{D}.
+\end{equation}
+Differentiation with respect to the variational degrees of freedom
+yields the \acrshort*{KS} and polarization equations:
+\begin{subequations}
+  \begin{align}
+    \mat{F}\mat{C} &= \mat{S}\mat{C}\mat{\epsilon} \\
+    \V\p &+ \Tr\mathbfup{s}\mat{D} = 0
+  \end{align}
+\end{subequations}
+The \acrshort*{KS} matrix now includes environment contributions:
+\begin{equation}
+  f_{\mu\nu} = h_{\mu\nu}  + G^\gamma_{\mu\nu}(\mat{D}) + f_{\mathrm{xc};\mu\nu}
+  + {}^t\p\s_{\mu\nu}.
+\end{equation}
+With respect to the coupling described in abstract terms above, the
+\acrshort*{SCF} case is formally simpler. In fact, whereas in the
+general case the source term depends on \emph{both} primal (amplitudes)
+and dual (multipliers) variational degrees of freedom, here only the
+dependence on the primal parameters (\acrshort*{MO} coefficients) needs
+to be considered.\autocite{Lipparini2011-aj}

### 2016-09-06 Chapters/ContributionsSummary.tex
Chaps. 1, 2, and 5 are complete

@@
 My coauthors have all read and approved the description of contributions.
-
-\pagebreak
 
@@
 A well-defined \gls{API} clearly delimits the boundaries of the functionality
-offered, thus forcing a programming style and standardisation of the
+offered, effectively enforcing a programming style and standardisation of the
 functionality, eventually.~\autocite{Reddy2011-sd}
 
-When coupled with \emph{open-source} licensing, modularity is a proper
-step towards ensuring reproducibility of results from scientific
-simulation software.~\autocite{Gezelter2015-gz, Krylov2015-fs,
+Recently, a debate over \emph{open-source}
+licensing\autocite{St_Laurent2004-fa} and \emph{open data} practices has
+sparked in the quantum chemistry community.\autocite{Gezelter2015-gz, Krylov2015-fs,
 Jacob2016-oq}
+Open-source licensing enables an open code review process that is, in
+our opinion, essential in guaranteeing scientific
+reproducibility.\autocite{Gezelter2015-gz}
+The points raised \emph{against} wholesale adoption of open-source
+licensing models,\autocite{Krylov2015-fs} especially the concerns about
+code quality and maintenance, do partially resonate with us.
+However, we still think that a modular approach to programming can only
+have significant impact, in terms of improved reproducibility and
+overall code quality, when an open-source licensing model is adopted.
 Open-source development can fully leverage the benefits of widespread,
@@
 \acp{DVCS},~\autocite{github, gitlab} continuous
-integration,~\autocite{travis-ci, appveyor-ci} code
-coverage,~\autocite{coveralls} static and dynamic code
+integration,~\autocite{travis-ci, magnum-ci, appveyor-ci} code
+coverage analysis,~\autocite{coveralls} static and dynamic code
 analyses,~\autocite{coverity-scan} nightly regression testing, public
@@
 The Python layer provides input reading and parsing facilities,
-as implemented by the GetKw library .~\autocite{libgetkw}
+as implemented by the GetKw library.~\autocite{libgetkw}
 Cavity construction and discretization is handled either by the Fortan
@@
 Upon initialization of the library, a \emph{context} object containing
-the state of the \acrshort{PCM} \acrshort{API} is constructed and a handle to it
-(a pointer in C++ parlance) is returned to the host program. Further
-transactions between \pcmsolver and the host program proceed through
-manipulations of the context object.
+the state of the \acrshort{PCM} \acrshort{API} is constructed and an
+opaque handle to it (a pointer in C++ parlance) is returned to the host
+program. Further transactions between \pcmsolver and the host program
+proceed through manipulations of the context object.
 \pcmsolver introduces the concept of surface functions:
@@
 The first interface was implemented and released within the \DIRAC code
-and is described in Paper I.~\autocite{DIRAC15, DiRemigio2015-ou}.
+and is described in \paper{I}.~\autocite{DIRAC15, DiRemigio2015-ou}.
 Interfaces to the \LSDALTON\autocite{LSDALTON16, Aidas2013-rp} and
 \DALTON\autocite{LSDALTON16, Aidas2013-rp} codes have followed, as
-described in Paper II and Paper V, respectively.~\autocite{Bugeanu2015-tp, pcm-openrsp}
+described in \paper{II} and \paper{V}, respectively.~\autocite{Bugeanu2015-tp, pcm-openrsp}
 Details about the interface to \ReSpect~\autocite{ReSpect-3.5.0} can be
-found in Paper IV.~\autocite{pcm-respect}
+found in \paper{IV}.~\autocite{pcm-respect}
 Coupling with \psicode~\autocite{Turney2012-de} has allowed us to implement the
@@
 
-Paper IV is a step further in our exploration of the interplay between
-relativistic and solvent effects initiated with Paper I.
-Whereas Paper I presented the essential framework for the coupling of
+\paper{IV} is a step further in our exploration of the interplay between
+relativistic and solvent effects initiated with \paper{I}.
+Whereas \paper{I} presented the essential framework for the coupling of
 four-component \acrshort{SCF} wave functions with a classical polarizable continuum,
@@
 theoretical approach to high-order absorption spectroscopies.
-Paper V grafts a classical polarizable continuum approach to solvation on top
+\paper{V} grafts a classical polarizable continuum approach to solvation on top
 of the open-ended methodology of \citeauthor{Thorvaldsen2008-sg}.

### 2016-09-06 Chapters/QuantumMechanics.tex
Chaps. 1, 2, and 5 are complete

@@
  \begin{align}
-   h_{rs} &= \int \diff\vect{r} \varphi_r^\dagger(\vect{r}) h(\vect{r})\varphi_s(\vect{r}) \label{eq:one-el}\\
+   h_{rs} &= \int \diff\vect{r} \phi_r^\dagger(\vect{r}) h(\vect{r})\phi_s(\vect{r}) \label{eq:one-el}\\
    g_{rstu} &=
@@
    \Omega_{rs}(\vect{r}) g(\vect{r},\vect{r}^\prime) \Omega_{tu}(\vect{r}^\prime) \label{eq:two-el} \\
-  \Omega_{rs}(\vect{r}) &= \varphi_r^\dagger(\vect{r})\varphi_s(\vect{r}). \label{eq:overlap}
+  \Omega_{rs}(\vect{r}) &= \phi_r^\dagger(\vect{r})\phi_s(\vect{r}). \label{eq:overlap}
  \end{align}
@@
 the one-electron space, \ie in the space of \glspl{MO}
-$\lbrace\varphi_r\rbrace$, used to build the determinant $\Ket{0}$.
+$\lbrace\phi_r\rbrace$, used to build the determinant $\Ket{0}$.
 The one-electron \glspl{MO} are thus optimized considering each single
@@
 \begin{equation}\label{eq:fock-eq}
- \mat{F}\vect{\varphi} = \vect{\varphi}\mat{\epsilon}
+ \mat{F}\vect{\phi} = \vect{\phi}\mat{\epsilon}
 \end{equation}
@@
 
-A different approach is taken in \gls{DFT}. The electron density is
-taken as the central quantity in the theory, instead of the wave
-function.
+A different approach is taken in \gls{DFT}, where the electron density
+is the central quantity in the theory, instead of the wave function.
 The Hohenberg--Kohn theorems\autocite{Hohenberg1964-wo, Eschrig2012-as}
@@
 until self-consistency is reached.
-The \acrshort{KS} matrix contains the one-electron \acrshort{XC}
-potential in addition to the other terms contained in the Fock matrix:
+The \acrshort{KS} matrix:
 \begin{equation}\label{eq:ksmatrix}
-  f_{rs} = h_{rs} + \sum_{j}[g_{rsjj} - \gamma g_{rjjs}] + v_{\mathrm{xc};rs}
+  f_{rs} = h_{rs} + \sum_{j}[g_{rsjj} - \gamma g_{rjjs}] + f_{\mathrm{xc};rs}.
 \end{equation}
+contains an \acrshort{XC} one-electron contribution due to the
+in addition to the other terms contained in the Fock matrix.
+The functional derivative of the \acrshort*{XC} functional enters in the
+definition of this additional term:
+\begin{alignat}{2}
+  f_{\mathrm{xc}; rs}
+  =
+  \int\diff\vect{r}\Omega_{rs}(\vect{r})v_\mathrm{xc}(\vect{r}),
+  \quad&
+  v_\mathrm{xc}(\vect{r})
+  \coloneq
+  \frac{\delta E_\mathrm{xc}[\rho(\vect{r})]}{\delta \rho(\vect{r})}
+\end{alignat}
+notice moreover the exact exchange admixing factor $\gamma$ in the
+two-electron part of the \acrshort*{KS} matrix.
+
+How are \glspl*{MO} constructed? The usual approach is to expand them in
+a basis of one-particle, atom-centered basis functions of known
+functional form:
+\begin{equation}
+  \phi_r(\vect{r})
+  =
+  \sum_{\alpha=1}^{N_\mathrm{basis}}
+  C_{\alpha r}\chi_{\alpha}(\vect{r})
+\end{equation}
+and optimize the expansion coefficients in the \acrshort*{SCF} procedure.
+Choosing the one-particle \acrshort*{MO} space is then equivalent to
+choosing the one-particle \gls{AO} space.
+The functional form of the \acrshort*{AO} basis is chosen to strike a
+balance between a physically accurate description of the short-range
+Coulomb interaction and algorithmic robustness and
+efficiency.\autocite{Helgaker2000-tz}
+The usual choice, also adopted in this thesis, is to use bases of
+Gaussian functions.\autocite{Helgaker1995-gi, Reine2012-tq}
 
@@
 and retaining only terms complying to the $2n+1$ and $2n+2$ rules.
+\autocite{Helgaker1988-to, Helgaker1989-wl, Helgaker1992-ph}
 A superscript index in square brackets - $[i]$ -  will denote an $i$-th
@@
 order can be obtained from amplitudes and multipliers up to second
-order.
-\todo[inline]{CHECK THE LOGIC OF THIS STATEMENT}
-Since the triples first appear to second order the \gls{CCSD}
+order. Since the triples first appear to second order, the \gls{CCSD}
 energy is correct to third order in the fluctuation potential.

### 2016-09-06 Chapters/ResponseTheory.tex
Work on MQM chapter
Minor adjustments to text
Chaps. 1, 2, and 5 are complete

@@
 
-\epigraph{
+\begin{epigraphs}
+\qitem{
 White light goin' messin' up my brain
@@
 }
-\epigraph{What's the frequency, Kenneth? [...]
+\qitem{What's the frequency, Kenneth? [...]
 
@@
           --- \textsc{R.E.M.}}
+\end{epigraphs}
 
-\begin{itemize}
-    \item Something about exact state response theory: quasienergy,
-      response functions, residues of response functions.
-    \item  Derivation of the linear response function with
-      quantum/classical polarizable Hamiltonian and then refer to the
-      OpenRSP paper for the open-ended treatment.~\autocite{pcm-openrsp}
-\end{itemize}
+The scope of quantum chemistry is to understand and predict molecular
+properties, but so far we have made no mention of properties other than
+the energy.
+The experimentalists' view of molecular systems is built mainly around the use
+of spectroscopic techniques that explore the interaction of light and matter.
+When a system is exposed to an external perturbing electromagnetic field, it will
+respond with a detectable change in its properties.\autocite{Pedersen2012-il, Jaszunski2012-wy}
+By "external" we mean an electromagnetic field which is "weak" when
+compared to the electron--electron and electron--nuclei interactions in
+the molecule.
+\emph{Response theory} is the missing link between theory and
+experiment. It allows to describe perturbation-induced changes in
+molecular properties in terms of \emph{response functions}, built solely
+by means of \emph{unperturbed} molecular states and energies.
+Moreover, excitation energies and transition moments can be inferred
+from response functions without recurring to an explicit modelling of
+excited states.
+Section \ref{sec:exact-response} will give a brief introduction to exact
+state response theory for isolated molecules. The concepts of
+quasienergy,\autocite{Christiansen1998-pe} variational perturbation
+theory\autocite{Helgaker1992-ph} and pole-and-residue
+analysis of response functions\autocite{Olsen1985-nr} will be introduced.
+I will explicitly derive the \acrshort{SCF} response function for a
+quantum/classical polarizable Hamiltonian in Section
+\ref{sec:csm-response}. The derivation will employ the
+density matrix-based, \acrshort{AO} formalism introduced by
+\citeauthor{Thorvaldsen2008-sg} in \noparcite[ref.][]{Thorvaldsen2008-sg}.
+This formalism is suitable for the derivation of arbitrary-order
+response functions and their residues, also when perturbation-dependent
+basis sets are considered.\autocite{Friese2015-kb, Friese2015-bu}
+Such a general derivation for the \acrshort{PCM} is presented in
+\paper{V}.
 
-\pagebreak
+The response treatment of molecular properties has its roots in
+time-dependent perturbation theory\autocite{Konishi2009-zb} and has been
+continually developed in quantum chemistry for the past 30 years:
+the exposition will be necessarily brief and incomplete, the reader is
+referred to the cited works in the literature for a more detailed
+presentation of the topic.\autocite{Olsen1985-nr, Helgaker1992-ph, Olsen1995-pf,
+Christiansen1998-pe, Norman2011-ad, Helgaker2012-cz, Pawlowski2015-sq}
 
@@
 
-General theory~\autocite{Olsen1985-nr, Christiansen1998-pe,
-Norman2011-ad, Helgaker2012-cz, Pawlowski2015-sq}
-Solution strategy for response equations~\autocite{Kauczor2011-rd}
+General theory~\autocite{Olsen1985-nr, Helgaker1992-ph, Olsen1995-pf,
+Christiansen1998-pe, Norman2011-ad, Helgaker2012-cz, Pawlowski2015-sq}
+Solution strategy for response equations~\autocite{Saad2003-oa,
+Saad2011-gm, Kauczor2011-rd, Malmqvist2013-vw}
 Open-ended \acrshort{AO}-based \acrshort{SCF} response~\autocite{
@@
 \acrshort{PE}~\autocite{Olsen2010-wa}
-\acrshort{QM}/\acrshort{MM}/Continuum~\autocite{Stenidal2011-ki, Caprasecca2012-ir}
+\acrshort{QM}/\acrshort{MM}/Continuum~\autocite{Steindal2011-ki, Caprasecca2012-ir}
 Paper V~\autocite{pcm-openrsp}

### 2016-09-06 FrontBackmatter/Introduction.tex
Save the day's work. Chapter 1 almost done
Minor adjustments to text
Chaps. 1, 2, and 5 are complete

@@
 
-\epigraph{
+\begin{epigraphs}
+\qitem{
 pity this busy monster, manunkind,
@@
     --- \textsc{E. E. Cummings}}
-
-\epigraph{
+\qitem{
 [...] \textnorsk{og fordi jeg alltid har hatt en dragning mot det skjulte og hemmelige.}
@@
 --- \textnorsk{\textsc{K. O. Knausgård}, \textit{Om Høsten}}}
+\end{epigraphs}
 
@@
 
-\begin{figure}[!h]
+\begin{figure}[tb]
 \centering
-\scalebox{0.8}{\input{gfx/RQC-axes.tikz}}
+\scalebox{0.7}{\input{gfx/RQC-axes.tikz}}
 \caption[Pictorial depiction of the concept of \emph{model chemistries}.]{
@@
 \section*{The Problem of Solvation or Taming Complexity with Models}
-
-\todo[inline]{Classification of solute-solvent interactions. The focus
-on electrostatics.}
 
@@
    In the latter approach, commonly dubbed \acrshort{QM}/\acrshort{MM}, the \acrshort{MM}
-   region can either be polarizable~\autocite{MMPol} or non
-   polarizable.~\autocite{MM}
+   region can either be polarizable~\autocite{Mennucci2013-go,
+   Olsen2010-wa, Lipparini2011-rd} or non
+   polarizable. While the former method allows for mutual polarization
+   between the \acrshort*{QM} and \acrshort*{MM} subsystems, the latter
+   treats the \acrshort*{MM} region as fixed.
  \item
@@
 \emph{cutoff} for the \acrshort{MM} region is usually required to converge
-long-range electrostatic interactions.~\autocite{Something here??}
+long-range electrostatic interactions.\autocite{Steindal2011-ki}
 Continuum models avoid both problems at once. Statistical averaging is
@@
 recover a satisfactory description of specific effects.
-To partly alleviate these sources of error, the \acrshort{QM}/\acrshort{MM} and
-\acrshort{QM}/Continuum methods can and have been successfully combined to
-yield the three-layer \acrshort{QM}/\acrshort{MM}/Continuum method.~\autocite{Curutchet, Steindal Lipparini}
+To partly alleviate these sources of error, the
+\acrshort{QM}/\acrshort{MM} and \acrshort{QM}/Continuum methods can and
+have been successfully combined to yield the three-layer
+\acrshort{QM}/\acrshort{MM}/Continuum method.\autocite{Steindal2011-ki,
+Lipparini2011-rd, Caprasecca2012-ir, Lipparini2013-ud}
 Figure \ref{fig:qm-to-multiscale} schematically portrays the transition
@@
 Notice that we have deliberately ruled out so-called \emph{cluster}
-models from the above discussion.~\autocite{ClusterModels}
+models from the above discussion.
 These approaches replace the actual physical setting with a suitable

### 2016-09-06 Chapters/QuantumMechanics.tex
Write another section in the introduction

@@
 \chapter{A Quick Tour of Molecular Electronic Structure Theory}\label{ch:QM}
-
-The purpose of quantum chemistry is to provide models based on first
-principles that can help understand and predict molecular properties.
-In quantum chemistry, we apply physical models based on
-quantum many-body methods to molecular systems, employ their
-mathematical realizations and devise computable approximations.
-The central idea is, in fact, to be able to obtain an algorithmic
-implementation of the methods that can be applied to interesting
-chemical systems.
 

### 2016-09-06 FrontBackmatter/Introduction.tex
Write another section in the introduction

@@
 \chapter*{Introduction}
-\addtocontents{toc}{\protect\vspace{\beforebibskip}} % to have the bib a bit from the rest in the toc
-\addcontentsline{toc}{chapter}{\tocEntry{Introduction}}
+\pdfbookmark[1]{Introduction}{introduction}
 
@@
 \begin{epigraphs}
+\setlength\epigraphwidth{7.5cm}
 \qitem{
-pity this busy monster, manunkind,
+plays with the bigness of his littleness
+
+--- electrons deify one razorblade
+
+into a mountainrange; lenses extend
+
+unwish through curving wherewhen till unwish
+
+returns on its unself.
     }{
     --- \textsc{E. E. Cummings}}
+\setlength\epigraphwidth{7cm}
 \qitem{
@@
 
-\todo[inline]{Quantum chemistry as they study of
-molecular properties by means of \emph{ab initio} physico-mathematical
-theories and computer programs.
-Introduce the concept of model chemistries, but then mention that we are
-interested in different aspects and will look at different "axes".}
-
-This question was posed by R.~S.~Mulliken over a half-century ago, as
-lecture title for the acceptance of the Gilbert N.~Lewis award in 1960.
+This question was posed by R.~S.~Mulliken over a half-century
+ago\footnote{This was the title Mulliken chose for his Gilbert N.~Lewis
+award acceptance speech in 1960.}
+and can be considered the fundamental research question behind the
+development of quantum chemistry.
+The purpose of quantum chemistry is to provide models based on first
+principles that can help \emph{understand} and \emph{predict} molecular properties.
+As stated by \citeauthor{Dirac1929-gn} in his \citetitle{Dirac1929-gn}
+paper:\autocite{Dirac1929-gn, Kutzelnigg2000-fl}
+\blockquote{The underlying physical laws necessary for the mathematical
+theory of a large part of physics and the whole of chemistry are thus
+completely known, and the difficulty is only that the exact application of
+these laws leads to equations much too complicated to be soluble. It
+therefore becomes desirable that approximate practical methods of
+applying quantum mechanics should be developed, which can lead to an
+explanation of the main features of complex atomic systems without too
+much computation.}
+Following Dirac's \emph{dictum}, in quantum chemistry we apply physical
+models based on quantum many-body methods to molecular systems, employ
+their mathematical realizations and devise computable approximations.
+The central idea is, in fact, to be able to obtain an algorithmic
+implementation of the methods that can be applied to interesting
+chemical systems: model building and software implementations are two
+closely intertwined aspects in the practice of quantum chemistry.
 
@@
 
-This will be the subject of Chapter \ref{ch:QM}
+Quantum chemical methods have become invaluable tools in the modern
+development of chemistry,\autocite{Lee1995-pw, Helgaker2004-oz,
+Tajti2004-ye} as witnessed by the Nobel prizes awarded in 1998\autocite{Nobel1998}
+and 2013.\autocite{Nobel2013}
+The concept of \emph{model chemistries} is at the heart of these
+successes. Introduced by \citeauthor{Pople1999-gt}, theoretical model
+chemistries are specific combinations of approximations in the basis set
+and molecular electronic structure methodology.\autocite{Pople1999-gt}
+Model chemistries are \emph{systematically improvable} so that it is
+possible to achieve the heaven of chemical accuracy by relaxing
+approximations in the model, albeit at an increased computational cost.
+Figure \ref{fig:RQC-axis} presents the usual depiction of this concept
+as a set of orthogonal axes, where chemical accuracy can be achieved by
+moving away from the origin.
+Relativity can be considered as the third axis of quantum
+chemistry: accuracy can be reached improving on the Hamiltonian, method
+and basis axes.\autocite{Saue2011-qg}
+
+The description of the ideas and methods at the basis of quantum
+chemistry will be the subject of Chapter \ref{ch:QM}.
+I will put emphasis on the methods that have been relevant in the work
+presented in this thesis. Section \ref{sec:mqm} discusses some of the
+ideas of quantum many-body theory relevant to quantum chemistry. Section
+\ref{sec:mean-field} introduces the workhorses of quantum chemistry
+\acrlong*{HF} and \acrlong*{KS} \acrlong*{DFT}.
+The \acrlong*{MBPT} and \acrlong*{CC} rungs on the methodology axis will
+be discussed in Sections \ref{sec:mbpt} and \ref{sec:coupled-cluster},
+respectively.
 
@@
 Environment effects are usually classified as:
-\begin{itemize}
-\item \emph{direct}.
+\begin{description}[font=\normalfont\scshape]
+\item[Direct.]
   These effect stem straightforwardly from the modification underwent by
   the solute electronic density when interacting with the environment.
-\item \emph{Indirect}.
+\item[Indirect.]
   It is common for solutes to exhibit different minimum-energy
@@
   labelled as indirect.
-\item \emph{Local field}.
+\item[Local field.]
   Light-matter interactions are also affected by the environment. Local
@@
   responses.~\autocite{Cammi1998-jp, Pipolo2014-sd}
-\item \emph{Dynamic}.
+\item[Dynamic.]
   Probing the nature of molecular excited states has enormous
@@
   become important.
-\item \emph{Specific}. This catch-all category includes all effects
+\item[Specific.] This catch-all category includes all effects
   stemming from the peculiar solute-solvent pair interactions that
@@
   detail.
-\end{itemize}
+\end{description}
 

### 2016-09-06 Chapters/QuantumMechanics.tex
Adjust TOC bookmarks

@@
 antisymmetric.
-Using the expansion theorem,\autocite{Gross1991-hi} we thus seek the wave
+Using the expansion theorem,\autocite{Gross1991-hi} we seek the wave
 function for the $N$-electron system as a linear combination of
@@
 but scales exponentially with the number of electrons in the system and
-has thus very limited practical applications.
+thus has very limited practical applications.
 Approximations can be tailored by appropriate truncations of the
@@
 of the many-electron wave function using just one Slater determinant.
-How to direct this search towards the "best"? We will use the
+How does one direct this search towards the "best"? We will use the
 variational principle and optimize the degrees of freedom in our
@@
 $N$-electron space. All variational degrees of freedom reside in
-the one-electron space, \ie in the space of \glspl{MO}
+the one-electron space, \ie~in the space of \glspl{MO}
 $\lbrace\phi_r\rbrace$, used to build the determinant $\Ket{0}$.
@@
 \end{equation}
-with $\mat{F}$ being the \emph{Fock matrix}.
+where $f_{ai}$ are elements of the \emph{Fock matrix} $\mat{F}$.
 The two-electron terms appearing in the Fock matrix are called the
@@
 In the \gls{LDA}, the functional is expressed in terms of an
-\emph{energy density} $e_\mathrm{xc}$ which is a local function of the
+\emph{energy density}, $e_\mathrm{xc}$, which is a local function of the
 density:
@@
 In the \gls{GGA}, the energy density is a local function of both the
-density and the norm of its gradient
-$\zeta(\vect{r})=\vect{\nabla}\rho\cdot \vect{\nabla}\rho$:
-\begin{equation}
- E^\mathrm{GGA}_\mathrm{xc}[\rho] = \int e_\mathrm{xc}(\rho(\vect{r}),\zeta(\vect{r}))\diff\vect{r}.
-\end{equation}
+density and the norm of its gradient:
+\begin{alignat}{2}
+ E^\mathrm{GGA}_\mathrm{xc}[\rho] = \int
+ e_\mathrm{xc}(\rho(\vect{r}),\zeta(\vect{r}))\diff\vect{r},
+ \quad&
+ \zeta(\vect{r}) \coloneq \vect{\nabla}\rho\cdot \vect{\nabla}\rho.
+\end{alignat}
 Finally, in \emph{hybrid} functionals, some proportion of the
@@
 Results from noniterative methods can be rationalized within the
-framework of \acrshort*{MBPT}.\autocite{Koch1997-nm, Eriksen2014-gy,
-Eriksen2015-il, Kristensen2016-od}
+framework of \acrshort*{MBPT}.\autocite{Stanton1997-lx, Koch1997-nm,
+Eriksen2014-gy, Eriksen2015-il, Kristensen2016-od}
 From Eqs. \eqref{eq:1st-order}--\eqref{eq:2nd-order}, one can derive

### 2016-09-06 FrontBackmatter/Bibliography.tex
Adjust TOC bookmarks

%********************************************************************
% Bibliography
%*******************************************************
% work-around to have small caps also here in the headline
\manualmark
\markboth{\spacedlowsmallcaps{\bibname}}{\spacedlowsmallcaps{\bibname}} % work-around to have small caps also
\phantomsection
\refstepcounter{dummy}
\addtocontents{toc}{\protect\vspace{\beforebibskip}} % to have the bib a bit from the rest in the toc
\addcontentsline{toc}{chapter}{\tocEntry{\bibname}}
\label{app:bibliography}
\printbibliography

### 2016-09-06 FrontBackmatter/Introduction.tex
Adjust TOC bookmarks

@@
 \chapter*{Introduction}
-\pdfbookmark[1]{Introduction}{introduction}
+\refstepcounter{dummy}
+\phantomsection
+\addcontentsline{toc}{chapter}{\tocEntry{Introduction}}
 
@@
 We hereby define a "solution", or more generally an "environment", as
-a system were the number of solvent molecules exceeds by far the number
+a system where the number of solvent molecules exceeds by far the number
 of solute molecules.~\autocite{Tomasi2004-dc, Tomasi2007-es}
@@
 understanding reality and predicting new and exciting
-phenomena.~\autocite{Anderson1972-ai, Winsberg2010-sy, Kovac2011-ew}
+phenomena and properties.~\autocite{Anderson1972-ai, Winsberg2010-sy, Kovac2011-ew}
 One of the earlier attempts at tackling the problem of solvation is due
@@
 \item[Direct.]
-  These effect stem straightforwardly from the modification underwent by
+  These effects stem straightforwardly from the modification underwent by
   the solute electronic density when interacting with the environment.
@@
 parts of the system with different methods and couple these methods by
-bridging "scales" at the boundary, \emph{vide infra}.
+bridging "scales" at the boundary.
+Figure \ref{fig:qm-to-multiscale} schematically portrays the transition
+from a full \acrshort{QM} model of the relevant system to its multiscale
+representations.
+
+\begin{figure}[tb]
+\missingfigure{Add picture with multiscale models}
+\caption[From quantum mechanical to multiscale models]{
+From quantum mechanical to multiscale models.
+The images used in the scheme are reproduced courtesy of Dr.~Stefano
+Caprasecca (MoLEcoLab, Università di Pisa).}
+\label{fig:qm-to-multiscale}
+\end{figure}
+
 While both models treat the molecular degrees of freedom at the quantum
@@
    \emph{Continuum} (or \emph{implicit}) completely remove the degrees
-   of freedom of the environment from the picture, replacing it with a
+   of freedom of the environment from the model, replacing them with a
    structureless continuum.
@@
 \end{itemize}
+
 \acrshort{QM}/\acrshort{MM} models can capture, albeit approximately, the effect
@@
 However, they demand a statistical average of environment configurations
-to yield results of any significance. Moreover, a large number
-\emph{cutoff} for the \acrshort{MM} region is usually required to converge
+to yield results of any significance. Moreover, a rather large
+\emph{cutoff radius} for the \acrshort{MM} region is usually required to converge
 long-range electrostatic interactions.\autocite{Steindal2011-ki}
@@
 Lipparini2011-rd, Caprasecca2012-ir, Lipparini2013-ud}
-Figure \ref{fig:qm-to-multiscale} schematically portrays the transition
-from a full \acrshort{QM} model of the relevant system to its multiscale
-representations.
-
-\begin{figure}[tb]
-\missingfigure{Add picture with multiscale models}
-\caption[From quantum mechanical to multiscale models]{
-From quantum mechanical to multiscale models.
-The images used in the scheme are reproduced courtesy of Dr.~Stefano
-Caprasecca (MoLEcoLab, Università di Pisa).}
-\label{fig:qm-to-multiscale}
-\end{figure}
 

### 2016-09-07 Chapters/ResponseTheory.tex
Finish introduction. I'll have to add some stuff later..:

@@
 
-The scope of quantum chemistry is to understand and predict molecular
-properties, but so far we have made no mention of properties other than
-the energy.
-The experimentalists' view of molecular systems is built mainly around the use
-of spectroscopic techniques that explore the interaction of light and matter.
-When a system is exposed to an external perturbing electromagnetic field, it will
-respond with a detectable change in its properties.\autocite{Pedersen2012-il, Jaszunski2012-wy}
-By "external" we mean an electromagnetic field which is "weak" when
-compared to the electron--electron and electron--nuclei interactions in
-the molecule.
-\emph{Response theory} is the missing link between theory and
-experiment. It allows to describe perturbation-induced changes in
-molecular properties in terms of \emph{response functions}, built solely
-by means of \emph{unperturbed} molecular states and energies.
-Moreover, excitation energies and transition moments can be inferred
-from response functions without recurring to an explicit modelling of
-excited states.
+Here and in the following, "external" will be used to mean an
+electromagnetic field which is "weak" when compared to the
+electron--electron and electron--nuclei interactions in the molecule.
+
 Section \ref{sec:exact-response} will give a brief introduction to exact
@@
 basis sets are considered.\autocite{Friese2015-kb, Friese2015-bu}
-Such a general derivation for the \acrshort{PCM} is presented in
-\paper{V}.
+\paper{V} presents the arbitrary-order derivation of the \acrshort{PCM}.
 

### 2016-09-07 FrontBackmatter/Introduction.tex
Finish introduction. I'll have to add some stuff later..:

@@
 
-Quantum chemical methods have become invaluable tools in the modern
-development of chemistry,\autocite{Lee1995-pw, Helgaker2004-oz,
-Tajti2004-ye} as witnessed by the Nobel prizes awarded in 1998\autocite{Nobel1998}
-and 2013.\autocite{Nobel2013}
+It is easy to manipulate chemicals in a virtual laboratory. Nowadays
+quantum chemical methods often complement traditional experimental
+approaches. They have become invaluable tools in the modern development
+of chemistry,\autocite{Lee1995-pw, Helgaker2004-oz, Tajti2004-ye} as
+witnessed by the Nobel prizes awarded in 1998\autocite{Nobel1998} and
+2013.\autocite{Nobel2013}
 The concept of \emph{model chemistries} is at the heart of these
@@
 
-Characterizing, explaining and predicting the effect of the chemical
-environment on a large number of measurable properties requires,
-however, a synergistic experimental and theoretical approach.
+The scope of quantum chemistry is to understand and predict molecular
+properties, but so far we have made no mention of properties other than
+the energy.
+The experimentalists' view of molecular systems is built mainly around
+the use of spectroscopic techniques that explore the interaction of
+light and matter. When a system is exposed to an external perturbing
+electromagnetic field, it will respond with a detectable change in its
+properties.\autocite{Pedersen2012-il, Jaszunski2012-wy}
+Characterizing, explaining and predicting a large number of measurable
+properties requires a synergistic experimental and theoretical approach.
+\emph{Response theory} is the missing link between theory and
+experiment, making quantum chemistry a full-fledged virtual laboratory.
+Response theory allows the description and computation of
+perturbation-induced changes in observable molecular properties.
+Electric and magnetic properties, excitation energies and transition
+moments can easily be calculated in the framework of response theory.
+\emph{Response functions} are the central concept in response theory.
+These are built solely by means of \emph{unperturbed} molecular states and
+energies: no explicit modelling of excited states is needed.
 
-This will be the subject of Chapter \ref{ch:molprop}
+Response theory will be the subject of Chapter \ref{ch:molprop},
+where the basic ideas in the computation of response functions will be
+presented. We will discuss the formulation of the linear response
+function for quantum/classical polarizable Hamiltonians and show how the
+variational framework is a powerful theoretical tool.
 
@@
 
-This will be the subject of Chapter \ref{ch:solvation-correlation}
+The concepts of systematic improvability and theoretical model
+chemistries are the foundations for the successful practice of quantum
+chemistry.
+A balanced description of electron correlation is often necessary
+to achieve accurate enough results in the computation of molecular
+properties.\autocite{Lee1995-pw, Helgaker2004-oz, Tajti2004-ye}
+The cost-effective treatment of electron correlation is a challenging
+problem and a very active line of research in the field.
+\Acrlong*{DFT} approaches are cheap and widespread, but their general
+accuracy is hard to assess. \Acrlong*{MBPT} and \acrlong*{CC} approaches
+are more robust in this respect, albeit at an increased computational
+cost.
+
+In this thesis, we are interested in the inclusion of
+environment effects in \emph{ab initio} models of interesting chemical
+systems. As already noted, this is a challenging problem, the more so
+when including electron correlation is necessary to obtain better
+accuracy.
+We will describe our approach to this problem in Chapter
+\ref{ch:solvation-correlation}. Treating solvation and electron
+correlation has been a recurring subject of research in the literature
+since the inception of continuum models. Once again, we will leverage
+the variational formulation of classical polarizable models presented
+in Chapter \ref{ch:CSM}.

### 2016-09-08 Chapters/ResponseTheory.tex
Finished section on exact state response

@@
 
-Here and in the following, "external" will be used to mean an
-electromagnetic field which is "weak" when compared to the
-electron--electron and electron--nuclei interactions in the molecule.
+Response theory provides an \emph{ab initio} framework for the
+formulation and computation of molecular properties and can be
+considered the connection between quantum chemistry and experimental
+chemistry.
+The response treatment of molecular properties has its roots in
+time-dependent perturbation theory\autocite{Konishi2009-zb} and has been
+continually developed in quantum chemistry for the past 30
+years.\autocite{Olsen1985-nr, Helgaker1992-ph, Olsen1995-pf,
+Christiansen1998-pe, Norman2011-ad, Helgaker2012-cz, Pawlowski2015-sq}
 
@@
 analysis of response functions\autocite{Olsen1985-nr} will be introduced.
+The exposition closely follows the one by \citeauthor{Saue2002-ns}
+in \noparcite[ref.][]{Saue2002-ns}.
 I will explicitly derive the \acrshort{SCF} response function for a
@@
 
-The response treatment of molecular properties has its roots in
-time-dependent perturbation theory\autocite{Konishi2009-zb} and has been
-continually developed in quantum chemistry for the past 30 years:
-the exposition will be necessarily brief and incomplete, the reader is
-referred to the cited works in the literature for a more detailed
-presentation of the topic.\autocite{Olsen1985-nr, Helgaker1992-ph, Olsen1995-pf,
-Christiansen1998-pe, Norman2011-ad, Helgaker2012-cz, Pawlowski2015-sq}
-
 \section{Exact State Response Theory}\label{sec:exact-response}
 
-General theory~\autocite{Olsen1985-nr, Helgaker1992-ph, Olsen1995-pf,
-Christiansen1998-pe, Norman2011-ad, Helgaker2012-cz, Pawlowski2015-sq}
-Solution strategy for response equations~\autocite{Saad2003-oa,
-Saad2011-gm, Kauczor2011-rd, Malmqvist2013-vw}
-Open-ended \acrshort{AO}-based \acrshort{SCF} response~\autocite{
-Larsen2000-hj, Kjaergaard2008-hy, Thorvaldsen2008-sg,
-Kristensen2008-hv, Ringholm2014-gx}
-Open-ended \acrshort{SCF} residues~\autocite{Friese2015-kb}
+Let us consider the experimentally relevant situation where a molecular
+system is subject to an external source of electromagnetic radiation.
+Here and in the following, "external" will be used to mean an
+electromagnetic field which is "weak" when compared to the
+electron--electron and electron--nuclei interactions in the molecule.
+
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=.5\textwidth]{Optical-rotation.png}
+  \caption{
+  When a molecular system is subject to an external electromagnetic
+  radiation of frequency $\omega$, a macroscopic, experimentally
+  measurable response can be detected.
+  Intensity and position of the signal can be related to the microscopic
+  detail of the molecular system, as described by quantum mechanics.
+  The picture depicts an optical rotation experiment where
+  the linear polarization of the incident electromagnetic field
+  is rotated by an angle $\alpha$ upon traversing a chiral material.
+  Reproduced, with modifications, from \href{https://commons.wikimedia.org/wiki/File:Optical-rotation.svg}{Wikipedia}.
+  }
+  \label{fig:OR}
+\end{figure}
+
+The system can then be described by means of a combined matter--field
+Hamiltonian:
+\begin{equation}
+  H = H_\mathrm{matter} + H_\mathrm{field} + H_\mathrm{int},
+\end{equation}
+where the three terms describe the molecular sample, the electromagnetic
+field and their interaction, respectively.
+Since our focus is on the molecular system, one can neglect the
+quantized description of the field and treat the interaction term in a
+semiclassical fashion.\autocite{Craig2012-zp}
+The time-dependent Schr\"{o}dinger equation is now the appropriate
+equation of motion:
+\begin{equation}\label{eq:td-schrodinger}
+  H\psi(t) = \mathrm{i}\pderiv{\psi(t)}{t},
+\end{equation}
+where the time-dependent, semiclassical matter--field Hamiltonian is
+used:
+\begin{equation}
+  H = H_0 + V(t).
+\end{equation}
+The perturbation is a one-electron operator. We will further assume that
+it is periodic in time $V(t) = V(t+T)$ and admits a discrete Fourier
+decomposition:
+\begin{equation}\label{eq:fourier}
+  \begin{aligned}
+ V(t) &=
+ \sum_{k=-N}^N \expo{-\mathrm{i}\omega_k t}V(\omega_k)
+ =
+ \sum_{k=-N}^N\expo{-\mathrm{i}\omega_k t} \sum_X \epsilon_X(\omega_k)V_X \\
+ &=
+ \sum_{x}\expo{-\mathrm{i}\omega_x t}\epsilon_x V_X
+  \end{aligned}
+\end{equation}
+where all the frequencies are integer multiples, $\omega_k=n_k \omega$
+of the fundamental frequency $\omega = \frac{2\pi}{T}$.\autocite{Olsen1985-nr}
+Frequencies and amplitudes in the Fourier decomposition obey the
+following relations:
+\begin{alignat}{2}
+ \omega_{-k} = \omega_{k},
+ \quad&
+ \epsilon_x = \epsilon_X^*(\omega_k)
+ =
+ \epsilon_X(\omega_{-k}) = \epsilon^*_{-x},
+\end{alignat}
+since the $V(t)$ is required to be Hermitian.
+For notational convenience, the perturbation operator ($X$) and
+frequency ($k$) indices have been collapsed into the common index $x$.
+
+Under the assumption that the external field is weak with respect to the
+molecular field, time-dependent perturbation theory is an appropriate
+tool to approximately solve Eq. \eqref{eq:td-schrodinger}.
+The response theory route to molecular properties introduces a
+phase-separated ansatz for the time-dependent wave function:
+\begin{equation}\label{eq:phase-separated}
+  \psi(t) = \expo{-\mathrm{i}F(t)}\bar{\psi}(t),
+\end{equation}
+where the time-derivative of the phase factor defines the time-dependent
+\emph{quasienergy}:\autocite{Christiansen1998-pe, Pawlowski2015-sq}
+\begin{equation}
+  Q(t) = \dot{F}(t) = \Braket{\bar{\psi}(t) |
+  H_0 + V(t) - \mathrm{i}\pderiv{}{t}
+  | \bar{\psi}(t)}.
+\end{equation}
+The time-dependent quasienergy is variational, but the corresponding
+time-dependent Hellmann--Feynman theorem cannot straightforwardly be
+applied to the calculation of molecular properties. Fortunately, we can exploit the
+periodicity of the perturbation operator to introduce the time-averaged
+quasienergy:
+\begin{equation}
+  \aveQ =
+ \frac{1}{T} \int_{-\frac{T}{2}}^{\frac{T}{2}} \diff t Q(t) =
+ Q_0 + \sum_{x}\epsilon_x E_x
+\end{equation}
+where:
+\begin{subequations}
+ \begin{align}
+  Q_0 &= \left\lbrace \Braket{\bar{\psi}(t) | H_0 | \bar{\psi}(t)} \right\rbrace_T
+- \left\lbrace \Braket{\bar{\psi}(t) | \mathrm{i} \pderiv{}{t} | \bar{\psi}(t)}\right\rbrace_T = E_0(0) - S \\
+  E_x &= \left\lbrace \Braket{\bar{\psi}(t) | V_X | \bar{\psi}(t)}\exp(-\mathrm{i}\omega_k t) \right\rbrace_T
+ \end{align}
+\end{subequations}
+Differentiating the time-averaged quasienergy yields the time-averaged
+Hellmann--Feynman theorem:
+\begin{equation}\label{eq:tave-hellfeyn}
+  \deriv{\aveQ}{\epsilon_X(\omega_k)}
+  =
+  \left\lbrace\Braket{\bar{\psi}(t) | \pderiv{H}{\epsilon_x} | \bar{\psi}(t)}\right\rbrace_T
+  = E_x,
+\end{equation}
+which can be used to connect response functions to molecular properties.
+Letting $\ket{0}$ represent the unperturbed reference state, we can
+develop the Kubo expansion at zero perturbation strength of the
+expectation value of the observable $H_X$:\autocite{Kubo1957-ay}
+\begin{equation}
+\begin{aligned}
+ \Braket{\bar{\psi}(t) | V_X | \bar{\psi}(t)} &\simeq
+ \Braket{0|V_X|0} +\sum_{k=-N}^N
+ \response{V_X}{V(\omega_k)}{\omega_k}\expo{-\mathrm{i}\omega_k t} \\
+ &+\frac{1}{2}\sum_{k,l=-N}^N
+ \response{V_X}{V(\omega_k),V(\omega_l)}{\omega_k,\omega_l}\expo{-\mathrm{i}(\omega_k+\omega_l)t}
+\end{aligned}
+\end{equation}
+response functions may be identified as the Fourier coefficients in the
+expansion. These are none other but higher order derivatives of the
+time-averaged quasienergy, as a comparison with the same expansion for
+the right-hand side of Eq. \eqref{eq:tave-hellfeyn} will confirm:
+\begin{equation}\label{eq:kubo}
+\begin{aligned}
+  \deriv{\aveQ}{\epsilon_x}
+  &\simeq \Braket{0|V_X|0}\delta_{\omega_x} \\
+  &+\sum_{y}\response{V_X}{V_Y}{\omega_y}\epsilon_y\delta_{\omega_x+\omega_y} \\
+  &+\frac{1}{2}\sum_{y,z}\response{V_X}{V_Y,V_Z}{\omega_y,\omega_z}\epsilon_y\epsilon_z\delta_{\omega_x+\omega_y+\omega_z}
+\end{aligned}
+\end{equation}
+First and higher order molecular properties can now be identified from
+the response functions appearing as Fourier coefficients in the Kubo
+expansion by taking the appropriate perturbation-strength derivatives at
+zero field strength:
+\begin{equation}\label{eq:nth-order-derivative}
+  \left.\frac{\diff^n \aveQ}{\diff\epsilon_{x_1}\diff\epsilon_{x_2}\cdots\diff\epsilon_{x_n}} \right|_{\vect{\epsilon}=0}
+  =
+  \response{V_{X_1}}{V_{X_2}, \ldots, V_{X_n}}{\omega_{x_2},\ldots,\omega_{x_n}}\delta_{\omega_{x_1}+\cdots+\omega_{x_n}},
+\end{equation}
+the $\delta_{\omega_{x_1}+\cdots+\omega_{x_n}}$ notation is used to
+enforce the sum rule on the probing and response frequencies involved in
+the expansion:
+\begin{equation}\label{eq:sum-rule}
+ \sum_{i=1}^{n} \omega_{k_i} = 0
+\end{equation}
+%\begin{subequations}
+%  \begin{align}
+%  \left.\deriv{\aveQ}{\epsilon_a}\right|_{\vect{\epsilon}=0} &= \Braket{0|V_A|0}\delta_{\omega_a} \\
+%  \left.\frac{\diff^2 \aveQ}{\diff\epsilon_a\diff\epsilon_b}\right|_{\vect{\epsilon}=0}
+%         &=
+%         \response{V_A}{V_B}{\omega_b}\delta_{\omega_a+\omega_b} \\
+%  \left.\frac{\diff^3\aveQ}{\diff\epsilon_a\diff\epsilon_b\diff\epsilon_c}\right|_{\vect{\epsilon}=0}
+%  &=
+%  \response{V_A}{V_B, V_C}{\omega_b,
+%  \omega_c}\delta_{\omega_a+\omega_b+\omega_c}.
+%  \end{align}
+%\end{subequations}
+Response functions quantify how, to a certain order in the perturbing
+field tuples, the expectation value of a given observable is modified.
+The semicolon appearing in the response functions separates the operator
+$V_X$, whose response is measured, from the external perturbations
+$V_Y$, $V_Z$\dots{} causing it.
+From the perturbation expansion Eq. \eqref{eq:kubo}, it is clear that
+permutation symmetry exists between the external perturbations provided
+that the corresponding frequencies are permuted concomitantly.
+
+Having introduced response functions as a useful tool to quantify the
+effect of external perturbations on molecular properties, it is now time
+to consider how to actually compute them.
+In this Section, we are concerned with exact wave functions that can
+thus be parametrized in a variational fashion by a set of parameters
+$\vect{\lambda}$: $\ket{0}=\ket{0(\vect{\lambda})}$.
+We recall that the Lagrangian method outlines in Section
+\ref{sec:coupled-cluster} can be employed also in this
+case.\autocite{Christiansen1998-pe, Helgaker2012-cz, Pawlowski2015-sq}
+The optimal parameter set is determined upon minimization of the
+time-averaged quasienergy:
+\begin{equation}\label{eq:var-cond}
+ \deriv{\aveQ}{\vect{\lambda}} = 0.
+\end{equation}
+In variational perturbation theory,\autocite{Helgaker1992-ph,
+Saue2002-ns} we assume the variational parameters to be functions of the
+perturbation strength parameters: $\vect{\lambda} =
+\vect{\lambda}(\vect{\epsilon})$.
+Furthermore, we require the stationarity condition to hold at \emph{all}
+perturbation strengths, with $\lambda(0) = 0$, for convenience.
+
+The perturbation-strength total derivative at zero perturbation strength
+of the time-averaged quasi energy is:
+\begin{equation}
+\begin{aligned}
+ \left[\prod_{i=1}^n \deriv{}{\epsilon_{x_i}}\right]\aveQ
+ &= Q^{[n]}_{x_1 \ldots x_n}
+ =
+ Q^{[n]}_{0; x_1 \ldots x_n}
+ + \sum_{j=1}^{n}E^{[n-1]}_{x_j;x_1\ldots x_{j-1}x_{j+1}\ldots x_n} \\
+ &=
+ \left[\prod_{i=1}^n \deriv{}{\epsilon_{x_i}}\right]Q_0
+ + \sum_{j=1}^n
+ \left[\prod_{i\neq j}^{n-1} \deriv{}{\epsilon_{x_i}}\right]E_{x_j}
+\end{aligned}
+\end{equation}
+Notice that this is just an alternative and more verbose notation for
+Eq. \eqref{eq:nth-order-derivative}, which allows to express response
+functions in terms of wave function perturbed parameters.
+%for first and second-order properties one has:
+%\begin{subequations}
+% \begin{align}
+%   Q^{[1]}_{a} &= \left[Q^{[1]}_{0; a} + E_{a}\right]\delta_{\omega_a} =
+%   \braket{0| V_A | 0}\delta_{\omega_a} \\
+%  Q^{[2]}_{ab} &= \left[Q^{[2]}_{0; ab} + E^{[1]}_{a;b} + E^{[1]}_{b;a}\right]\delta_{\omega_a+\omega_b}
+%  = \response{V_A}{V_B}{\omega_b}\delta_{\omega_a+\omega_b}
+% \end{align}
+%\end{subequations}
+Application of the chain rule to the total derivatives in fact yields:
+\begin{subequations}
+ \begin{align}
+   Q^{[1]}_{a} &=
+   \sum_{\sigma}\left[
+   Q^{[1]}_{0; \sigma}\lambda^{[1]}_{\sigma;a} + E_{a}
+   \right]\delta_{\omega_a} \\
+   Q^{[2]}_{ab} &=
+   \sum_{\sigma\tau}\left[
+   \lambda^{[1]}_{\sigma;a}Q^{[2]}_{0;\sigma\tau}\lambda^{[1]}_{\tau;b}
+   + Q^{[1]}_{0; \sigma}\lambda^{[2]}_{\sigma;a}
+   + E^{[1]}_{a;\sigma}\lambda^{[1]}_{\sigma;b}
+   + E^{[1]}_{b;\sigma}\lambda^{[1]}_{\sigma;a}
+   \right]\delta_{\omega_a+\omega_b} \label{eq:linres}
+ \end{align}
+\end{subequations}
+where the following tensors were introduced:
+\begin{subequations}
+ \begin{align}
+ Q^{[n]}_{0; \sigma_1\cdots \sigma_n} &=
+ \frac{\partial^n Q_0}{\partial\lambda_{\sigma_1}\cdots\partial\lambda_{\sigma_n}} \\
+ E^{[n]}_{x;\sigma_1\cdots \sigma_n} &=
+ \frac{\partial^n E_x}{\partial\lambda_{\sigma_1}\cdots\partial\lambda_{\sigma_n}} \\
+\lambda^{[n]}_{\sigma;x_1\dots x_n} &=
+\frac{\partial^n \lambda_\sigma}{\partial \epsilon_{x_1}\cdots\partial \epsilon_{x_n}}
+ \end{align}
+\end{subequations}
+
+The response equations, determining the $\vect{\lambda}^{[n]}$ tensor of
+perturbed wave function parameters, are obtained by differentiating the
+variational condition on the time-averaged quasienergy Eq.
+\eqref{eq:var-cond}:\autocite{Olsen1985-nr, Helgaker1992-ph, Christiansen1998-pe}
+\begin{equation}
+ R_{\sigma;x_1\cdots x_n}^{[n]} = \left[\prod_{i=1}^n \deriv{}{\epsilon_{x_i}}\right]
+\left. \deriv{\aveQ}{\lambda_\sigma}\right|_{\vect{\epsilon}=0} = 0.
+\end{equation}
+The determination of second-order properties requires the knowledge of
+the first-order response of the wave function:
+\begin{subequations}
+ \begin{align}
+ R_{\sigma}^{[0]} &= Q_{0;\sigma}^{[1]} = 0 \\
+ R_{\sigma;a}^{[1]} &= \sum_{\tau}Q_{0;\sigma\tau}^{[2]}\lambda_{\tau;a}^{[1]} + E_{a;\sigma}^{[1]} = 0.
+ \label{eq:1st-order-response}
+ \end{align}
+\end{subequations}
+The time-averaged quasienergy formalism allows to formulate response
+properties to static and dynamic fields on an equal footing as
+perturbation-strength derivatives.
+Moreover, variational perturbation theory achieves a transparent
+derivation of the necessary response equations.\autocite{Helgaker1992-ph}
+
+Let us consider in more detail the form of the linear response function
+Eq. \eqref{eq:linres}. The formal solution to the response equation Eq.
+\eqref{eq:1st-order-response} can be obtained by inverting the
+\emph{Hessian}. Eq. \eqref{eq:linres} can then be rewritten as:
+\begin{equation}
+  \response{V_A}{V_B}{\omega_b}
+  =
+   -\sum_{\sigma\tau}
+   E^{[1]}_{a;\sigma}
+   \left[\mat{Q}^{[2]}_{0}\right]^{-1}_{\sigma\tau}
+   E^{[1]}_{b;\tau}.
+\end{equation}
+Expansion into the exact eigenbasis of $H_0$ yields the
+\gls{SOS} expression for the linear response function:
+\begin{equation}
+  \response{V_A}{V_B}{\omega_b}
+  =
+  -\sum_{n>0}\left[
+  \frac{\braket{0|V_A|n}\braket{n|V_B|0}}{\omega_{n0}-\omega_b}
+  +
+  \frac{\braket{0|V_B|n}\braket{n|V_A|0}}{\omega_{n0}-\omega_a}
+  \right]\delta_{\omega_a+\omega_b}.
+\end{equation}
+The excitation energies of the system appear in the denominators
+$\omega_{n0} = E_n - E_0$ and can thus be identified as the \emph{poles}
+of the response function. Moreover, the \emph{residues} at
+the poles are the transition moments determining the intensity of
+spectroscopic transitions:
+\begin{equation}
+  \lim_{\omega_b\rightarrow\omega_{n0}}
+  \response{V_A}{V_B}{\omega_b}
+  = \braket{0|V_A|n}\braket{n|V_B|0}.
+\end{equation}
+A similar analysis holds also for higher order response functions. The
+pole structure of the higher order quantities also conveys information
+about the excited states \emph{via} single and double
+residues.\autocite{Olsen1985-nr, Christiansen1998-pe, Helgaker2012-cz}
 
@@
 \acrshort{QM}/\acrshort{MM}/Continuum~\autocite{Steindal2011-ki, Caprasecca2012-ir}
-Paper V~\autocite{pcm-openrsp}
+Open-ended \acrshort{AO}-based \acrshort{SCF} response~\autocite{
+Larsen2000-hj, Kjaergaard2008-hy, Thorvaldsen2008-sg,
+Kristensen2008-hv, Ringholm2014-gx}
+Open-ended \acrshort{SCF} residues~\autocite{Friese2015-kb}
+\paper{V}\autocite{pcm-openrsp}
+Solution strategy for response equations~\autocite{Saad2003-oa,
+Saad2011-gm, Kauczor2011-rd, Malmqvist2013-vw}

### 2016-09-08 Chapters/ResponseTheory.tex
Finishing up response theory chapter

@@
 
-Section \ref{sec:exact-response} will give a brief introduction to exact
-state response theory for isolated molecules. The concepts of
+Section \ref{sec:exact-response} will give a brief introduction to
+response theory for isolated molecules. The concepts of
 quasienergy,\autocite{Christiansen1998-pe} variational perturbation
@@
 
-\section{Exact State Response Theory}\label{sec:exact-response}
+\section{Response Theory in a Nutshell}\label{sec:exact-response}
 
 Let us consider the experimentally relevant situation where a molecular
-system is subject to an external source of electromagnetic radiation.
+system is subject to an external source of electromagnetic radiation,
+Figure \ref{fig:OR}.
 Here and in the following, "external" will be used to mean an
@@
 \end{equation}
-%\begin{subequations}
-%  \begin{align}
-%  \left.\deriv{\aveQ}{\epsilon_a}\right|_{\vect{\epsilon}=0} &= \Braket{0|V_A|0}\delta_{\omega_a} \\
-%  \left.\frac{\diff^2 \aveQ}{\diff\epsilon_a\diff\epsilon_b}\right|_{\vect{\epsilon}=0}
-%         &=
-%         \response{V_A}{V_B}{\omega_b}\delta_{\omega_a+\omega_b} \\
-%  \left.\frac{\diff^3\aveQ}{\diff\epsilon_a\diff\epsilon_b\diff\epsilon_c}\right|_{\vect{\epsilon}=0}
-%  &=
-%  \response{V_A}{V_B, V_C}{\omega_b,
-%  \omega_c}\delta_{\omega_a+\omega_b+\omega_c}.
-%  \end{align}
-%\end{subequations}
 Response functions quantify how, to a certain order in the perturbing
@@
 functions in terms of wave function perturbed parameters.
-%for first and second-order properties one has:
-%\begin{subequations}
-% \begin{align}
-%   Q^{[1]}_{a} &= \left[Q^{[1]}_{0; a} + E_{a}\right]\delta_{\omega_a} =
-%   \braket{0| V_A | 0}\delta_{\omega_a} \\
-%  Q^{[2]}_{ab} &= \left[Q^{[2]}_{0; ab} + E^{[1]}_{a;b} + E^{[1]}_{b;a}\right]\delta_{\omega_a+\omega_b}
-%  = \response{V_A}{V_B}{\omega_b}\delta_{\omega_a+\omega_b}
-% \end{align}
-%\end{subequations}
 Application of the chain rule to the total derivatives in fact yields:
@@
 
+The derivation above holds for an exact state, a situation that is never
+realized in common practice. The power of response theory, however, is
+in its applicability to exact and approximate state theories alike.
+As noted by \citeauthor{Norman2011-ad}, a response theory treatment of
+molecular properties can be summarized into the following four
+steps:\autocite{Norman2011-ad}
+\begin{enumerate}
+    \item Single out a quantity of interest that can be connected to
+      molecular properties. In our case, the time-averaged
+      quasienergy;
+    \item Introduce a suitable parametrization of the time-dependent,
+      phase-separated wave function;
+    \item Devise the appropriate equation of motion, based on the
+      time-dependent Schr\"{o}dinger equation, for the time-dependent
+      wave function parameters;
+    \item Identify response functions and response equations.
+\end{enumerate}
+Thus, one can formulate and compute molecular properties based on
+the approximate state ans\"{a}tze of quantum chemistry.
+An interesting formulation of \acrshort{SCF} response theory is the
+open-ended \acrshort{AO} density matrix-based approach presented by
+\citeauthor{Thorvaldsen2008-sg}.\autocite{Thorvaldsen2008-sg}
+By formulating the perturbation-strenght derivative of the time-averaged
+quasienergy Lagrangian in terms of the \acrshort*{AO} density matrix,
+the authors were able to formulate arbitrary order response functions
+in the \acrshort*{AO} basis.
+The density matrix is the variational degree of freedom and response
+equations can be obtained by differentiation of the \gls{TDSCF}
+equation, the stationarity condition on the density matrix.
+The approach lends itself to an efficient \acrshort*{AO}-based
+implementation,\autocite{Larsen2000-hj, Kjaergaard2008-hy} can fully
+leverage different response parameters elimination
+schemes,\autocite{Thorvaldsen2008-sg, Kristensen2008-hv} and is amenable
+to an open-ended, recursive implementation.\autocite{Ringholm2014-gx,
+Friese2015-kb, Friese2015-bu}
+We will base our derivation of \acrshort*{SCF} response theory for
+quantum/classical polarizable Hamiltonians of this framework.
+
 \section{SCF Response Theory for Quantum/Classical Polarizable Hamiltonians}\label{sec:csm-response}
 
-\acrshort{PCM}~\autocite{Cammi1994-qj, Cammi1996-wf, Cammi1996-vx,
+The development of response theory for quantum/classical polarizable
+Hamiltonians has moved almost at the same pace as the polarizable
+models themselves.
+An ample literature exists in the context of the
+\acrshort{PCM},\autocite{Cammi1994-qj, Cammi1996-wf, Cammi1996-vx,
 Cammi1999-rb, Cammi2003-qy, Frediani2005-nc, Ferrighi2010-pm}
-\acrshort{FQ} and \acrshort{FQ}/\acrshort{PCM}~\autocite{Lipparini2012-hx, Lipparini2012-tl, Lipparini2013-ud}
-\acrshort{PE}~\autocite{Olsen2010-wa}
-\acrshort{QM}/\acrshort{MM}/Continuum~\autocite{Steindal2011-ki, Caprasecca2012-ir}
-Open-ended \acrshort{AO}-based \acrshort{SCF} response~\autocite{
-Larsen2000-hj, Kjaergaard2008-hy, Thorvaldsen2008-sg,
-Kristensen2008-hv, Ringholm2014-gx}
-Open-ended \acrshort{SCF} residues~\autocite{Friese2015-kb}
-\paper{V}\autocite{pcm-openrsp}
+polarizable \acrshort{MM} models,\autocite{Curutchet2009-bt, Olsen2010-wa,
+Lipparini2012-hx, Lipparini2012-tl}
+and general \acrshort*{QM}/\acrshort*{MM}/Continuum models.\autocite{Steindal2011-ki, Caprasecca2012-ir, Lipparini2013-ud}
+
+Solvent relaxation is an important point when formulating response
+properties. Solvation dynamics occurs on timescales and through
+relaxation modes that are typical of the solvent molecular structure and of the
+peculiar solute-solvent interactions.
+This is clearly difficult to model with continuum models, since the
+entire frequency spectrum of the permittivity $\diel(\omega)$ would be
+required.\autocite{Ingrosso2003-ev, Caricato2005-lo,
+Mennucci2005-vi, Caricato2006-ba, Corni2015-pe}
+However, if one consider the initial steps of an electronic excitation
+process, a Franck--Condon-like response of the solvent can be assumed.
+Nuclear motions inside and among solvent molecules will not be able to
+follow the fast rearrangement of the solute electronic density, the
+corresponding part of the response will remain frozen in the state
+immediately previous the change.
+The polarization vector will then be (approximately) a sum of two
+components:
+\begin{equation}
+ \vect{P} \simeq \vect{P}_\mathrm{fast} + \vect{P}_\mathrm{slow},
+\end{equation}
+this is equivalent to consider only the asymptotic limits of the solvent
+dielectric spectrum, \ie the static $\diels$ and optical
+dielectric constants $\dield$.
+
+The variational setting introduced in Section \ref{sec:variational},
+offers a series of advantages also in the formulation of response
+theory. As already noted, a classical counterpart to the
+Hellmann--Feynman theorem holds:
+\begin{equation}\label{eq:classical-HF}
+  \deriv{U_\mathrm{pol}}{F} = \pderiv{U_\mathrm{pol}}{F}
+  + \pderiv{U_\mathrm{pol}}{\s}\pderiv{\s}{F}
+  = \pderiv{U_\mathrm{pol}}{F},
+\end{equation}
+which forms the basis for our formal development in
+\paper{V}.\autocite{pcm-openrsp}
+There we show how the variational polarization functional can
+successfully achieve a transparent derivation of arbitrary order
+response functions, when the fixed-cavity approximation for the
+\acrshort*{PCM} is assumed.
+Here we report a summary of our findings in \paper{V} employing the
+classical functional in supermatrix form in Eq. \eqref{eq:supermatrix-functional}.
+In the notation of \citeauthor{Thorvaldsen2008-sg}, we introduce the
+generalized free energy:
+\begin{equation}\label{eq:generalized-KS-energy}
+  \begin{aligned}
+ \tilde{\mathcal{G}}
+ &=
+ \tilde{\mathcal{E}}
+ + \frac{1}{2}{}^t\tilde{\p}\V\tilde{\p} +
+ {}^t\tilde{\p}\Tr(\tilde{\mathbfup{s}}\pertmat{D}) \\
+ &\eqtr
+ \left[\pertmat{h} + \pertmat{V}^{t} +
+ \frac{1}{2}\pertmat{G}^{\gamma}(\pertmat{D}) -
+ \frac{\mathrm{i}}{2}\pertmat{T}\right]\pertmat{D}
+ + \tilde{E}_\mathrm{xc}[\tilde{\rho}(\pertmat{D})] + h_\mathrm{nuc}
+ + \frac{1}{2}{}^t\tilde{\p}\V\tilde{\p}
+ + {}^t\tilde{\p}\tilde{\mathbfup{s}}\pertmat{D},
+ \end{aligned}
+\end{equation}
+whose perturbation-strength derivatives at zero perturbation strength
+will yield the desired response functions.
+
+Due to the implicit time-dependence of $\pertmat{D}$ and
+$\tilde{\sigma}$, higher-order
+derivatives of the KS generalized energy will require application of the
+chain rule. The ${mn, abc\ldots}$ superscript describes how and to which
+extent the chain rule was applied for a given term, \ie the number of explicit
+differentiations with respect to the variational densities, so that
+\begin{equation}\label{eq:differential}
+  \mathcal{G}^{mn, abc}
+  =
+  \frac{\partial^{m+n+3}\mathcal{G}}{\partial(\mat{D}^T)^m\partial\p^n\partial\epsilon_a\partial\epsilon_b\partial\epsilon_c}
+  =
+    \mathcal{E}^{m, abc} +
+  \frac{\partial^{m+n+3}U_\mathrm{pol}}{\partial(\mat{D}^T)^m\partial\p^n\partial\epsilon_a\partial\epsilon_b\partial\epsilon_c} \text{.}
+\end{equation}
+In this notation, the index $m$ denotes the order of differentiation
+with respect to the density matrix $\mat{D}$, while the index $n$ symbolizes
+the order of differentiation with respect to the \acs{ASC} density $\sigma$.
+Differentiation with respect to the density matrix will result in a
+$2m$-rank tensor, while differentiation with respect to the
+\acs{ASC} density will result in a function of the continuous cavity
+index $\vect{s}$. For higher-order properties, mixed terms involving both density
+matrix and \acs{ASC} density differentiation may generally occur.
+In the \emph{fixed-cavity approximation}, the cavity is kept frozen at a
+given molecular geometry.\autocite{Cammi1994-qj}
+Under this simplifying assumption, only the linear interaction term in
+the polarization functional eq.~\ref{eq:PCM-functional} will be affected
+by the movements of the nuclei \emph{via} the dependence of basis functions on
+the molecular geometry.
+Its perturbation-strength derivative will then be
+\begin{equation}\label{eq:PCM-derivative}
+ \deriv{}{\epsilon_a}\left\lbrace
+ U_\mathrm{pol}
+ \right\rbrace_T
+ \eqavetr
+ {}^t\tilde{\p}\tilde{\mathbfup{s}}^a\pertmat{D} \text{,}
+\end{equation}
+where the second term only involves derivatives of the electrostatic potential integrals.
+We remark that, under the fixed-cavity approximation, both the density
+matrix -- $m$ -- and \acs{ASC} density -- $n$ -- differentiation indices
+in eq.~\ref{eq:differential} can only assume the values 0 or 1 for the
+$\frac{\partial^{m+n+3}U_\mathrm{pol}}{\partial(\mat{D}^T)^m\partial\p^n\partial\epsilon_a\partial\epsilon_b\partial\epsilon_c}$
+term.
+By construction, the density matrix dependence in the
+polarization functional is at most linear, while,
+by virtue of the classical Hellmann--Feynman theorem,
+eq.~\ref{eq:classical-HF}, the \acs{ASC} variational density will also appear
+at most \emph{linearly} in $\tilde{\mathcal{G}}^{00,a}$.
+The free energy term perturbation strength derivative is given as
+\begin{equation}\label{eq:tilde-G00-a}
+  \begin{aligned}
+    \tilde{\mathcal{G}}^{00,a}(\tilde{\mat{D}}, \tilde{\sigma}, t)
+    &= \tilde{\mathcal{E}}^{0,a} +
+    {}^t\tilde{\p}\Tr({\tilde{\mathbfup{s}}^a\pertmat{D}}) \\
+  &\eqtr
+  (\pertmat{h}^a + \pertmat{V}^{t,a} + \frac{1}{2}\pertmat{G}^{\gamma, a}(\pertmat{D}) + \pertmat{F}^{\Omega_a}_\mathrm{xc}
+  -\frac{\mathrm{i}}{2}\pertmat{T}^a)\pertmat{D}
+  + h_\mathrm{nuc}^a + {}^t\tilde{\p}\tilde{\mathbfup{s}}^a\pertmat{D}
+  \end{aligned}
+\end{equation}
+where the matrix $\pertmat{F}^{\Omega_a}_\mathrm{xc}$ denotes the
+the functional derivative matrix defined in terms of the perturbed
+overlap distributions $\tilde{\Omega}^a$.
+
+Response functions can then be obtained by straightforward
+differentiation with respect to further perturbations and subsequent
+evaluation at zero perturbation strength, so that
+\begin{subequations}
+  \begin{align}
+    L^a &\eqavetr
+    \mathcal{G}^{00,a} - \mat{S}^a\mat{W} \label{eq:first-order} \\
+    L^{ab} &\eqavetr \mathcal{G}^{00,ab}
+    + \mathcal{G}^{10,a}\mat{D}^{b}
+    + \mathcal{G}^{01,a}\p^{b}
+    - \mat{S}^{ab}\mat{W}
+    - \mat{S}^{a}\mat{W}^{b} \label{eq:second-order} \\
+    L^{abc} &\eqavetr
+      \mathcal{G}^{00,abc}
+    + \mathcal{G}^{10,ac}\mat{D}^{b}
+    + \mathcal{G}^{10,ab}\mat{D}^{c}
+    + \mathcal{G}^{20,a}\mat{D}^{b}\mat{D}^{c} \nonumber \\
+    &+ \mathcal{G}^{10,a}\mat{D}^{bc}
+    + \mathcal{G}^{11,a}\mat{D}^{b}\p^c \label{eq:third-order} \\
+    &+ \mathcal{G}^{01,ac}\p^{b}
+    + \mathcal{G}^{01,ab}\p^{c}
+    + \mathcal{G}^{01,a}\p^{bc}
+    + \mathcal{G}^{11,a}\p^{b}\mat{D}^c \nonumber \\
+    &- \mat{S}^{abc}\mat{W}
+    - \mat{S}^{ab}\mat{W}^{c}
+    - \mat{S}^{ac}\mat{W}^{b}
+    - \mat{S}^{a}\mat{W}^{bc} \nonumber
+  \end{align}
+\end{subequations}
+and similarly for higher-order response functions.
+
+The governing equations for the perturbed \acs{ASC} densities are
+obtained in complete analogy with the handling of pertubed density
+matrices outlined above.
+We introduce a decomposition of the \acs{ASC} density into frequency
+components into the perturbation-strength derivative of
+Eq. \eqref{eq:polarization-stat}, so that
+\begin{equation}\label{eq:PCM-differentiated}
+  \V\p_\omega^{b_N} + \Tr \mathbfup{s}\mat{D}_\omega^{b_N} =
+  \mathbb{S}^{(n-1)}_\omega
+  \eqtr - \left(\mat{\varphi}\mat{D}\right)_{\omega, n-1}^{b_N} \text{.}
+\end{equation}
+$\mathbb{S}^{(n-1)}_\omega$ contains all terms that depend on lower order density matrices and
+differentiated source integrals.
+The term $\mathbb{S}^{(n-1)}_\omega$ always contains at least a first derivative of the
+electrostatic potential integrals and is thus zero if the basis set is
+independent of the perturbation tuple under consideration.
+We now introduce a partition of the frequency components of the
+polarization $\p_\omega^{b_N}$ into homogeneous (H) and particular (P)
+terms:
+\begin{equation}\label{eq:polarization-partition}
+  \p_\omega^{b_N} = \p_\mathrm{P}^{b_N} + \p_\mathrm{H}^{b_N} \text{,}
+\end{equation}
+a similar partition exists for the density matrix and we thus get
+the following system of equations:
+\begin{subequations}
+  \begin{align}
+    \V\p_\mathrm{H}^{b_N} &+ \Tr \mathbfup{s}\Dh{b_N} = 0
+    \label{eq:homo-polarization-eq}\\
+    \V\p_\mathrm{P}^{b_N} &+ \Tr \mathbfup{s}\Dp{b_N}
+    = \mathbb{S}^{(n-1)}_\omega \text{.}
+ \label{eq:part-polarization-eq}
+  \end{align}
+\end{subequations}
+We note that the particular term of the polarization is nonzero if and
+only if the basis set depends on the external perturbation.
+
+We finally turn our attention to the \acs{TDSCF} equation.
+The perturbation-strength differentiated generalized \acs{KS} matrix is
+first separated into its frequency components $\mathbfcal{F}^{b_N}_\omega$.
+The H/P partition introduced for the variational densities will induce a
+similar partition into these frequency components:
+\begin{equation}\label{eq:generalized-Fock-partition}
+  \mathbfcal{F}_\omega^{b_N} =
+  \mat{G}^{\mathrm{KS}}(\Dh{b_N}) + \ASCh{b_N}\mat{\esp}
+  + \breve{\mathbfcal{F}}_\omega^{b_N} \text{.}
+\end{equation}
+The two-electron and \acs{XC} contributions depending on the homogeneous
+perturbed density matrix have been collected into the
+$\mat{G}^\mathrm{KS}(\Dh{b_N})$ matrix, while all other contributions
+are collected into $\breve{\mathbfcal{F}}_\omega^{b_N}$, and we
+refer to \noparcite[refs.][]{Thorvaldsen2008-sg} and
+\noparcite[][]{Ringholm2014-gx} for a more detailed treatment of this.
+The parametrization of the homogeneous part of the perturbed density
+matrix can be exploited to conveniently reformulate the perturbed
+\acs{TDSCF} equation, so that
+\begin{equation}\label{eq:response-equation}
+ \left[\genHessian - \omega_{b_N}\genMetric \right]\rspParam{b_N} = \RHS{b_N}
+\end{equation}
+where the generalized Hessian $\genHessian$ and metric $\genMetric$
+matrices were introduced and are defined by their transformations on the
+response parameters $\rspParam{b_N}$:\autocite{Larsen2000-hj,
+Kjaergaard2008-hy}
+\begin{align}
+  \genHessian\rspParam{b_N} &=
+  \mat{G}^\mathrm{KS}([\rspParam{b_N}, \mat{D}]_{\mat{S}})\mat{D}\mat{S}
+  -\mat{S}\mat{D}\mat{G}^\mathrm{KS}([\rspParam{b_N}, \mat{D}]_{\mat{S}})
++\mat{F} [\rspParam{b_N}, \mat{D}]_{\mat{S}} \mat{S}
+-\mat{S} [\rspParam{b_N}, \mat{D}]_{\mat{S}} \mat{F}
++ \ASCh{b_N}\mat{\esp}\mat{D}\mat{S}
+-\mat{S}\mat{D}\mat{\esp}\ASCh{b_N}
+\label{eq:E2-lintra}
+\\
+  \genMetric\rspParam{b_N} &= \mat{S} [\rspParam{b_N}, \mat{D}]_{\mat{S}} \mat{S} \text{.}
+  \label{eq:S2-lintra}
+\end{align}
+The generalized Hessian matrix $\genHessian$ includes two types of
+solvent term. \emph{Implicit} terms are those included in the
+zeroth-order Fock matrix, $\mat{F}$.
+The last two terms in Eq. \ref{eq:E2-lintra} are \emph{explicit} terms,
+as they explicitly involve the $N$-th order homogenous \acs{ASC}
+variational density.
+The theoretical treatment of frequency-dependent properties in solution
+within the \acs{PCM} requires adoption of a nonequilibrium response
+framework.\autocite{Cammi1999-rb, Tomasi2005-vm}
+The explicit \acs{PCM} terms in Eq. \ref{eq:E2-lintra} are then
+evaluated using the optical permittivity $\dield$ of the solvent
+instead of the static permittivity $\diels$ used to compute the
+implicit contributions.
+In contrast to $\genHessian$, the generalized metric matrix --
+$\genMetric$ -- is unchanged.
+The \ac{RHS} in the response equation only includes terms that
+depend on particular contributions up to the desired order or lower-order perturbed density
+matrices:
+\begin{equation}\label{eq:RHS}
+  \RHS{b_N} =
+  \left[\left( \tilde{mathbfcal{F}}
+-
+\frac{\mathrm{i}}{2}\pertmat{S}\deriv{}{t}\right)\pertmat{D}\pertmat{S}\right]^{\ominus,
+b_N}_\mathrm{P}
+\end{equation}
+
 Solution strategy for response equations~\autocite{Saad2003-oa,

### 2016-09-09 Chapters/MathematicalResults.tex
Fix some missing references in CC chapter
Finished response theory chapter. One section in appendix missing

@@
 We introduce the following auxiliary potential:
+\begin{subequations}
 \begin{empheq}[left={h(\vect{r}) = \empheqlbrace}]{align}
@@
 \end{empheq}
+\end{subequations}
 for which we have:
+\begin{subequations}
 \begin{empheq}[left={\empheqlbrace}]{align}
@@
 \end{empheq}
+\end{subequations}
 We then define the \emph{reaction potential} as:
@@
 the jump conditions and the integral representations of the reaction and auxiliary potentials:
+\begin{subequations}
 \begin{empheq}[left={\empheqlbrace}]{align}
@@
 \end{empheq}
+\end{subequations}
 The final ingredient is the \gls{DtN} map, which can be derived by employing Eq.~\eqref{eq:stat3}
@@
 
-Let us now take a step back and formulate a different strategy for the
-solution of the transmission problem.
+In this Section, I will show how one can reformulate the transmission problem
+into a variational formulation.
+Such a formulation allows for a larger vector space, with weaker regularity
+conditions, to be explored as solution space for the problem.
 We will follow the exposition of \citeauthor{Ern2004-oo} quite closely
-in introducing the \emph{weak} formulation of \acp{PDE}.
+in introducing the \emph{weak} formulation of \glspl{PDE}.
 For simplicity, we assume Dirichlet boundary conditions for a conductor,
@@
   \nabla^2\esp = -4\pi\rho,
-  \quad \esp \in \mathcolor{ForestGreen}{\mathcal{C}^2_0(\Omegai)},
+  \quad \esp \in \mathcal{C}^2_0(\Omegai),
 \end{equation}
@@
   \begin{aligned}
-    &\text{Seek $\Psi \in \mathcolor{ForestGreen}{H_0^1(\Omegai)}$ such that:}\\
-    &\mathcolor{PMS1797}{(\nabla\zeta, \nabla\Psi)} =
-    \mathcolor{PMS2229}{-4\pi(\zeta, \rho)} \quad
-    \forall \zeta\in \mathcolor{PMS130}{H_0^1(\Omegai)}
+    &\text{Seek $\Psi \in H_0^1(\Omegai)$ such that:}\\
+    &(\nabla\zeta, \nabla\Psi) =
+    -4\pi(\zeta, \rho) \quad
+    \forall \zeta\in H_0^1(\Omegai)
   \end{aligned}
@@
   \right.
-  \label{eq:weak}
+  \tag{\ref{eq:weak}}
 \end{equation}
@@
 
-\todo[inline]{To write or not to write these????}
-
-Weak formulation of \acrshort{BIE}s.\autocite{Hsiao2008-xb}
+It is important to note how \glspl{BIE} can also be reformulated in a
+variational framework, as shown in \noparcite[ref.][]{Hsiao2008-xb}
 
 \section{The Galerkin Method}
+
+\todo[inline]{Write this section}
 Approximability, conformity, consistency and asymptotic consistency
 of the approximation setting, $\inf$-$\sup$ condition, Céa's Lemma
-and generalized $\inf$-$\sup$ condition.~\autocite{Ern2004-oo,
-TesiFilippo}
+and generalized $\inf$-$\sup$ condition.~\autocite{Ern2004-oo}

### 2016-09-09 Chapters/ResponseTheory.tex
Finished response theory chapter. One section in appendix missing

@@
 Here we report a summary of our findings in \paper{V} employing the
-classical functional in supermatrix form in Eq. \eqref{eq:supermatrix-functional}.
+supermatrix form of the classical functional as given in Eq.
+\eqref{eq:supermatrix-functional}.
 In the notation of \citeauthor{Thorvaldsen2008-sg}, we introduce the
@@
 
-Due to the implicit time-dependence of $\pertmat{D}$ and
-$\tilde{\sigma}$, higher-order
-derivatives of the KS generalized energy will require application of the
-chain rule. The ${mn, abc\ldots}$ superscript describes how and to which
-extent the chain rule was applied for a given term, \ie the number of explicit
-differentiations with respect to the variational densities, so that
+To keep track of the number of chain rule applications,
+we will denote higher-order derivatives of the \acrshort*{KS}
+generalized free energy with the composite ${mn, abc\ldots}$
+superscript:
 \begin{equation}\label{eq:differential}
@@
 \end{equation}
-In this notation, the index $m$ denotes the order of differentiation
-with respect to the density matrix $\mat{D}$, while the index $n$ symbolizes
-the order of differentiation with respect to the \acs{ASC} density $\sigma$.
-Differentiation with respect to the density matrix will result in a
-$2m$-rank tensor, while differentiation with respect to the
-\acs{ASC} density will result in a function of the continuous cavity
-index $\vect{s}$. For higher-order properties, mixed terms involving both density
-matrix and \acs{ASC} density differentiation may generally occur.
-In the \emph{fixed-cavity approximation}, the cavity is kept frozen at a
-given molecular geometry.\autocite{Cammi1994-qj}
-Under this simplifying assumption, only the linear interaction term in
-the polarization functional eq.~\ref{eq:PCM-functional} will be affected
-by the movements of the nuclei \emph{via} the dependence of basis functions on
-the molecular geometry.
-Its perturbation-strength derivative will then be
-\begin{equation}\label{eq:PCM-derivative}
+In this notation, the $abc\ldots$ indices denote the perturbations with
+respect to which differentiation occurs. The index $m$ denotes the order
+of differentiation with respect to the density matrix. Eventually, $n$
+eventually denotes the order of differentiation with respect to the
+polarization.
+The classical interaction kernel $\V$ might, in the most general
+scenario, depend on the molecular geometry.
+In \paper{V} and in the following we ignore this possibility and work
+under the simplifying assumption that $\V$ is independent of the nuclear
+positions.
+This goes under the name of \emph{fixed-cavity approximation} in the
+context of the \acrshort*{PCM}.\autocite{Cammi1994-qj}
+Under this simplifying assumption, the perturbation-strength derivative
+of the classical functional will only include terms stemming from the
+linear interaction term:
+\begin{equation}\label{eq:polarization-derivative}
  \deriv{}{\epsilon_a}\left\lbrace
@@
 \end{equation}
-where the second term only involves derivatives of the electrostatic potential integrals.
-We remark that, under the fixed-cavity approximation, both the density
-matrix -- $m$ -- and \acs{ASC} density -- $n$ -- differentiation indices
-in eq.~\ref{eq:differential} can only assume the values 0 or 1 for the
-$\frac{\partial^{m+n+3}U_\mathrm{pol}}{\partial(\mat{D}^T)^m\partial\p^n\partial\epsilon_a\partial\epsilon_b\partial\epsilon_c}$
-term.
-By construction, the density matrix dependence in the
-polarization functional is at most linear, while,
-by virtue of the classical Hellmann--Feynman theorem,
-eq.~\ref{eq:classical-HF}, the \acs{ASC} variational density will also appear
-at most \emph{linearly} in $\tilde{\mathcal{G}}^{00,a}$.
-The free energy term perturbation strength derivative is given as
+where the second term only involves derivatives of the source integrals.
+Moreover, in the general expression for the perturbation-strength
+derivative of the polarization functional in Eq. \eqref{eq:differential},
+the $m$ and $n$ indices can only assume the values 0 or 1.
+In fact, the density matrix dependence is at most linear in the
+polarization functional, while the classical Hellmann--Feynman theorem
+ensures that the polarization appears at most linearly in $\tilde{\mathcal{G}}^{00,a}$:
 \begin{equation}\label{eq:tilde-G00-a}
@@
   -\frac{\mathrm{i}}{2}\pertmat{T}^a)\pertmat{D}
-  + h_\mathrm{nuc}^a + {}^t\tilde{\p}\tilde{\mathbfup{s}}^a\pertmat{D}
+  + h_\mathrm{nuc}^a + {}^t\tilde{\p}\tilde{\mathbfup{s}}^a\pertmat{D}.
   \end{aligned}
 \end{equation}
-where the matrix $\pertmat{F}^{\Omega_a}_\mathrm{xc}$ denotes the
-the functional derivative matrix defined in terms of the perturbed
-overlap distributions $\tilde{\Omega}^a$.
-
-Response functions can then be obtained by straightforward
-differentiation with respect to further perturbations and subsequent
-evaluation at zero perturbation strength, so that
+$\pertmat{F}^{\Omega_a}_\mathrm{xc}$ is the functional derivative matrix
+defined in terms of the perturbed overlap distributions
+$\tilde{\Omega}^a$.
+Further perturbation-strength differentiations and evaluation at zero
+perturbation strength will yield the desired response functions:
 \begin{subequations}
@@
 \end{subequations}
-and similarly for higher-order response functions.
+and similarly for higher orders.
 
-The governing equations for the perturbed \acs{ASC} densities are
+The governing equations for the perturbed polarization densities are
 obtained in complete analogy with the handling of pertubed density
-matrices outlined above.
-We introduce a decomposition of the \acs{ASC} density into frequency
-components into the perturbation-strength derivative of
-Eq. \eqref{eq:polarization-stat}, so that
-\begin{equation}\label{eq:PCM-differentiated}
-  \V\p_\omega^{b_N} + \Tr \mathbfup{s}\mat{D}_\omega^{b_N} =
-  \mathbb{S}^{(n-1)}_\omega
-  \eqtr - \left(\mat{\varphi}\mat{D}\right)_{\omega, n-1}^{b_N} \text{.}
-\end{equation}
-$\mathbb{S}^{(n-1)}_\omega$ contains all terms that depend on lower order density matrices and
-differentiated source integrals.
-The term $\mathbb{S}^{(n-1)}_\omega$ always contains at least a first derivative of the
-electrostatic potential integrals and is thus zero if the basis set is
-independent of the perturbation tuple under consideration.
-We now introduce a partition of the frequency components of the
+matrices described in \noparcite[ref.][]{Thorvaldsen2008-sg}
+Denoting by the superscript $b_N$ a tuple of simulatenous perturbations,
+the polarization can be decomposed into frequency components.
+Similarly to the treatment of the perturbed density matrices,
+we introduce a partition of the frequency components of the
 polarization $\p_\omega^{b_N}$ into homogeneous (H) and particular (P)
@@
 \end{equation}
-a similar partition exists for the density matrix and we thus get
-the following system of equations:
+and we thus get the following system of equations:
 \begin{subequations}
-  \begin{align}
+\begin{empheq}[left={\empheqlbrace}]{align}
     \V\p_\mathrm{H}^{b_N} &+ \Tr \mathbfup{s}\Dh{b_N} = 0
@@
  \label{eq:part-polarization-eq}
-  \end{align}
+\end{empheq}
 \end{subequations}
-We note that the particular term of the polarization is nonzero if and
+$\mathbb{S}^{(n-1)}_\omega$ contains all terms that depend on lower
+order density matrices and differentiated source integrals.
+The term $\mathbb{S}^{(n-1)}_\omega$ always contains at least a first
+derivative of the source integrals and is thus zero if the basis set is
+independent of the perturbation tuple under consideration.
+Consequently, the particular term of the polarization is nonzero if and
 only if the basis set depends on the external perturbation.
 
-We finally turn our attention to the \acs{TDSCF} equation.
-The perturbation-strength differentiated generalized \acs{KS} matrix is
-first separated into its frequency components $\mathbfcal{F}^{b_N}_\omega$.
-The H/P partition introduced for the variational densities will induce a
-similar partition into these frequency components:
-\begin{equation}\label{eq:generalized-Fock-partition}
-  \mathbfcal{F}_\omega^{b_N} =
-  \mat{G}^{\mathrm{KS}}(\Dh{b_N}) + \ASCh{b_N}\mat{\esp}
-  + \breve{\mathbfcal{F}}_\omega^{b_N} \text{.}
-\end{equation}
-The two-electron and \acs{XC} contributions depending on the homogeneous
-perturbed density matrix have been collected into the
-$\mat{G}^\mathrm{KS}(\Dh{b_N})$ matrix, while all other contributions
-are collected into $\breve{\mathbfcal{F}}_\omega^{b_N}$, and we
-refer to \noparcite[refs.][]{Thorvaldsen2008-sg} and
-\noparcite[][]{Ringholm2014-gx} for a more detailed treatment of this.
-The parametrization of the homogeneous part of the perturbed density
-matrix can be exploited to conveniently reformulate the perturbed
-\acs{TDSCF} equation, so that
+Response equations can be obtained by differentiating the
+\acrshort{TDSCF} equation with respect to the perturbation tuple
+$b_N$:\autocite{Thorvaldsen2008-sg, Ringholm2014-gx}
 \begin{equation}\label{eq:response-equation}
@@
 Kjaergaard2008-hy}
+\begin{subequations}
 \begin{align}
@@
   \mat{G}^\mathrm{KS}([\rspParam{b_N}, \mat{D}]_{\mat{S}})\mat{D}\mat{S}
-  -\mat{S}\mat{D}\mat{G}^\mathrm{KS}([\rspParam{b_N}, \mat{D}]_{\mat{S}})
-+\mat{F} [\rspParam{b_N}, \mat{D}]_{\mat{S}} \mat{S}
+  -\mat{S}\mat{D}\mat{G}^\mathrm{KS}([\rspParam{b_N},
+  \mat{D}]_{\mat{S}}) \nonumber \\
+&+\mat{F} [\rspParam{b_N}, \mat{D}]_{\mat{S}} \mat{S}
 -\mat{S} [\rspParam{b_N}, \mat{D}]_{\mat{S}} \mat{F}
-+ \ASCh{b_N}\mat{\esp}\mat{D}\mat{S}
--\mat{S}\mat{D}\mat{\esp}\ASCh{b_N}
-\label{eq:E2-lintra}
+\label{eq:E2-lintra} \\
+&+ \ASCh{b_N}\mat{\esp}\mat{D}\mat{S}
+-\mat{S}\mat{D}\mat{\esp}\ASCh{b_N} \nonumber
 \\
@@
 \end{align}
+\end{subequations}
 The generalized Hessian matrix $\genHessian$ includes two types of
-solvent term. \emph{Implicit} terms are those included in the
+classical terms. \emph{Implicit} terms are those included in the
 zeroth-order Fock matrix, $\mat{F}$.
 The last two terms in Eq. \ref{eq:E2-lintra} are \emph{explicit} terms,
-as they explicitly involve the $N$-th order homogenous \acs{ASC}
+as they explicitly involve the $N$-th order homogenous polarization
 variational density.
-The theoretical treatment of frequency-dependent properties in solution
-within the \acs{PCM} requires adoption of a nonequilibrium response
-framework.\autocite{Cammi1999-rb, Tomasi2005-vm}
-The explicit \acs{PCM} terms in Eq. \ref{eq:E2-lintra} are then
-evaluated using the optical permittivity $\dield$ of the solvent
-instead of the static permittivity $\diels$ used to compute the
-implicit contributions.
-In contrast to $\genHessian$, the generalized metric matrix --
-$\genMetric$ -- is unchanged.
-The \ac{RHS} in the response equation only includes terms that
-depend on particular contributions up to the desired order or lower-order perturbed density
-matrices:
+The nonequilibrium response framework needed in continuum models when
+dealing with frequency-dependent properties,\autocite{Cammi1999-rb, Tomasi2005-vm}
+will affect the explicit terms in Eq. \ref{eq:E2-lintra} which are
+evaluated by means of the optical permittivity $\dield$.
+The \acrlong*{RHS} in the response equation only includes terms that
+depend on particular contributions up to the desired order or
+lower-order perturbed variational densities:\autocite{Ringholm2014-gx}
 \begin{equation}\label{eq:RHS}
   \RHS{b_N} =
-  \left[\left( \tilde{mathbfcal{F}}
+  \left[\left( \tilde{\mathbfcal{F}}
 -
 \frac{\mathrm{i}}{2}\pertmat{S}\deriv{}{t}\right)\pertmat{D}\pertmat{S}\right]^{\ominus,
-b_N}_\mathrm{P}
+b_N}_\mathrm{P},
 \end{equation}
+where $\mathbfcal{F}$ is the generalized \acrshort*{KS} matrix, see
+\noparcite[ref.][]{Thorvaldsen2008-sg} for details.
 
-Solution strategy for response equations~\autocite{Saad2003-oa,
+A direct solution of the response equation \eqref{eq:response-equation} by inversion
+is not feasible due to the size of the $\left[\genHessian -
+\omega_{b_N}\genMetric \right]$ matrix. \emph{Subspace iteration}
+methods are usually employed, the solution is achieved iteratively
+by enlarging a projection subspace of trial vectors.\autocite{Saad2003-oa,
 Saad2011-gm, Kauczor2011-rd, Malmqvist2013-vw}
+These methods can be implemented efficiently and in matrix-free form
+since the linear transformations $\genHessian\rspParam{b_N}$ and
+$\genMetric\rspParam{b_N}$ have known forms only involving integral and
+(perturbed) density matrices.
+Moreover, Jacobi precondinioting of the iterations is usually sufficiently
+robust and efficient, due to the diagonal dominance of the generalized
+Hessian.

### 2016-09-09 Chapters/ContributionsSummary.tex
Finish introduction. I'll have to add some stuff later..:
Almost done

@@
 The \pcmsolver library was developed to fulfill these requirements and
-offer the community an easy-to-use implementation of the \acrshort{PCM}.
+offer the community an easy-to-use implementation of the
+\acrshort{PCM}.\autocite{PCMSolver}
 \pcmsolver is written in C++ with Fortran, C and Python components.
@@
 
-\section{Continuum solvation in the relativistic regime}\label{sec:relapcm}
+\section{Continuum Solvation in the Relativistic Regime}\label{sec:relapcm}
 
@@
 
-\section{Non homogeneous environments}\label{sec:spherical}
+\section{Non Homogeneous Environments}\label{sec:spherical}
 
@@
 
-\section{Open-ended self-consistent field response theory in
-solution}\label{sec:pcmopenrsp}
+\section{Open-Ended Self-Consistent Field Response Theory in
+Solution}\label{sec:pcmopenrsp}
 

### 2016-09-09 FrontBackmatter/Abstract.tex
Almost done

--- 2016-07-27 FrontBackmatter/Abstract.tex
@@
 
-\vfill
+Synergistic theoretical and experimental approaches to challenging
+chemical problems have become more and more widespread, due to the
+availability of efficient and accurate \emph{ab initio} quantum chemical
+models.
+Limitations to such an approach do, however, still exist.
+The vast majority of chemical phenomena happens in complex environments,
+where the molecule of interest can interact with a large number of other
+moieties, solvent molecules or residues in a protein.
+These systems represent an ongoing challenge to our modelling
+capabilities, especially when high accuracy is required for the
+prediction of exotic and novel molecular properties.
+How to achieve the insight needed to understand and predict the physics
+and chemistry of such complex systems is still an open question.
+
+I will present our efforts in answering this question based on the
+development of the polarizable continuum model for solvation.
+While the solute is described by a quantum mechanical method, the
+surrounding environment is replaced by a structureless continuum
+dielectric.
+The mutual polarization of the solute-environment system is described by
+classical electrostatics.
+Despite its inherent simplifications, the model contains the basic
+mathematical features of more refined explicit quantum/classical
+polarizable models.
+Leveraging this fundamental similarity, we show how the
+inclusion of environment effects for relativistic and nonrelativistic
+quantum mechanical Hamiltonians, arbitrary order response properties and
+high-level electron correlation methods can be transparently derived and
+implemented.
+We implemented an application programming interface that can provide the
+continuum solvation functionality to any quantum chemistry software.
+As examples of the flexibility of our implementation approach, we
+present results for the continuum modelling of non homogeneous environments and
+how wavelet-based numerical methods greatly outperform the accuracy of
+traditional methods usually employed in continuum solvation models.
 
@@
 
-\vfill
+\begin{flushright}
+    \begin{tabular}{m{5cm}}
+      \textit{\myLocation, \myTime} \\
+      \includegraphics[width=0.4\textwidth]{gfx/signature.png} \\
+        \hline
+        \centering\myName \\
+    \end{tabular}
+\end{flushright}

### 2016-09-09 FrontBackmatter/Notation.tex
Work on MQM chapter
Write another section in the introduction
Adjust TOC bookmarks
Almost done

@@
-%\chapter*{Notations, conventions and units}
-\addcontentsline{toc}{chapter}{\tocEntry{Notations, conventions and units}}
-
 \renewcommand\nomgroup[1]{%
@@
 
-\renewcommand{\nomname}{
-Notations, Conventions and Units
-}
+\renewcommand{\nomname}{Notations, Conventions and Units}
 
@@
 \end{equation*}
+Complex conjugation will always be shown using a dagger ($\dagger$)
+instead of a star ($*$).
 }
+
+\refstepcounter{dummy}
+\phantomsection
+\addcontentsline{toc}{chapter}{\tocEntry{\nomname}}
 
@@
 \nomenclature[A, 05]{$\norm{u}{X}$}{Norm of $u$ in the normed space $X$}
+\nomenclature[A, 06]{$O(N)$}{A quantity of order $N$ or higher}
 % Vectors and Matrices
@@
 \nomenclature[B, 04]{$\Tr\mat{A}$}{Trace of $\mat{A}$. For $\mat{A}\in\mathbb{R}^{n, n}$, $\Tr\mat{A}=\sum_{i=1}^n A_{ii}$}
-\nomenclature[B, 05]{$0_N, I_N$}{The zero and the identity in an $N$-dimensional vector space}
+\nomenclature[B, 05]{$\eqtr$}{The expression following is to be interpreted as a trace. $\mathcal{E} \eqtr \mat{h}\mat{D} = \Tr\mat{h}\mat{D}$}
+\nomenclature[B, 06]{$\eqavetr$}{The expression following is to be interpreted as a trace followed by time-averaging over a period $T$.
+$\mathcal{E} \eqavetr \mat{h}\mat{D} = \frac{1}{T}\int_{0}^{T}\diff t\Tr\mat{h}\mat{D}$}
+\nomenclature[B, 07]{$0_N, I_N$}{The zero and the identity in an $N$-dimensional vector space}
 % Operators
@@
 \nomenclature[D, 08]{$H_0^s(\Omega)$}{Space of functions in $H^2(\Omega)$ with null trace on the boundary of $\Omega$}
-\nomenclature[D, 09]{$|u|_{s, \Omega}$}{Seminorm in $H^s(\Omega)$: $|u|_{s, \Omega} \coloneq \sum_{|\alpha|=s}\norm{\partial^\alpha u}{\Omega}$}
-\nomenclature[D, 10]{$\norm{u}{s, \Omega}$}{Norm in $H^s(\Omega)$: $\norm{u}{s, \Omega} \coloneq \sum_{l \leq s} |u|_{l, \Omega}$}
+\nomenclature[D, 09]{$|u|_{s, \Omega}$}{Seminorm in $H^s(\Omega)$: $|u|_{s, \Omega} = \sum_{|\alpha|=s}\norm{\partial^\alpha u}{\Omega}$}
+\nomenclature[D, 10]{$\norm{u}{s, \Omega}$}{Norm in $H^s(\Omega)$: $\norm{u}{s, \Omega} = \sum_{l \leq s} |u|_{l, \Omega}$}
 % Molecular Quantum Mechanics

### 2016-09-09 Chapters/MathematicalResults.tex
Almost done
Cleaning up solvation-correlation chapter

@@
 In this Section, I will show how one can reformulate the transmission problem
-into a variational formulation.
+in a variational fashion.
 Such a formulation allows for a larger vector space, with weaker regularity
@@
 and generalized $\inf$-$\sup$ condition.~\autocite{Ern2004-oo}
+
+\begin{comment}
+\section{Mixed Blabbering}
+
+\begin{equation}
+  \begin{pmatrix}
+   \sigma^\mathrm{tot} \\
+   \kappa^\mathrm{tot}
+  \end{pmatrix}
+  =
+  \begin{pmatrix}
+   \sigma_\mathrm{HF} \\
+   \kappa_\mathrm{HF}
+  \end{pmatrix}
+  +
+  \begin{pmatrix}
+   \sigma \\
+   \kappa
+  \end{pmatrix}
+\end{equation}
+and similarly for the sources:
+\begin{equation}
+\begin{aligned}
+  \begin{pmatrix}
+   \esp(\tamp{}, \tbar{}) \\
+   \zeta(\tamp{}, \tbar{})
+  \end{pmatrix}
+  &=
+  \begin{pmatrix}
+   \esp_\mathrm{HF} \\
+   \zeta_\mathrm{HF}
+  \end{pmatrix}
+  +
+  \begin{pmatrix}
+    \esp_{\mathrm{N}}(\tamp{}, \tbar{}) \\
+    \zeta_{\mathrm{N}}(\tamp{}, \tbar{})
+  \end{pmatrix} \\
+  &=
+  \begin{pmatrix}
+  \braket{\mathrm{HF} | \esp  | \mathrm{HF}} \\
+  \braket{\mathrm{HF} | \zeta | \mathrm{HF}}
+  \end{pmatrix}
+  +
+  \begin{pmatrix}
+   \braket{\mathrm{HF} | \tilde{\esp} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} | \mathrm{HF}} \\
+   \braket{\mathrm{HF} | \tilde{\zeta} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} | \mathrm{HF}}
+  \end{pmatrix}
+\end{aligned}
+\end{equation}
+The \acrshort{BCH} expansions of the source terms have been
+rewritten as:
+\begin{subequations}
+  \begin{align}
+  \overline{\esp} &= \esp + \tilde{\esp} = \esp + \commutator{\esp}{T}
+  + \frac{1}{2}\commutator{\commutator{\esp}{T}}{T} \\
+  \overline{\kappa} &= \kappa + \tilde{\kappa} = \kappa + \commutator{\kappa}{T}
+  + \frac{1}{2}\commutator{\commutator{\kappa}{T}}{T}
+  \end{align}
+\end{subequations}
+taking advantage of the fact that both $\esp$ and $\kappa$ are
+nondiagonal, one-electron operators and hence their commutator
+expansions truncate at the third term. Moreover, normal-ordering of the
+operators has been introduced.~\autocite{Crawford2000-ey, Shavitt2009-mr}
+This is equivalent to applying a shift to the correlation part of the
+source term one-electron operators to remove the reference source terms:
+\begin{equation}
+  \begin{aligned}
+  \begin{pmatrix}
+    \esp_{\mathrm{N}}(\tamp{}, \tbar{}) \\
+    \zeta_{\mathrm{N}}(\tamp{}, \tbar{})
+  \end{pmatrix}
+  &=
+  \begin{pmatrix}
+   \braket{\mathrm{HF} | \tilde{\esp} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} | \mathrm{HF}} \\
+   \braket{\mathrm{HF} | \tilde{\zeta} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} | \mathrm{HF}}
+  \end{pmatrix}
+  \\
+  &=
+  \begin{pmatrix}
+    \braket{\mathrm{HF} | \overline{\esp} - \esp_\mathrm{HF} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} - \esp_\mathrm{HF} | \mathrm{HF}} \\
+  \braket{\mathrm{HF} | \overline{\zeta} - \zeta_\mathrm{HF} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} -
+  \zeta_\mathrm{HF}| \mathrm{HF}}
+  \end{pmatrix}
+\end{aligned}
+\end{equation}
+
+Eventually, the polarization functional can be rewritten as:
+\begin{equation}
+  \begin{aligned}
+  U_\mathrm{pol} &=
+  \frac{1}{2}(\sigma_\mathrm{HF}+\sigma)\PCM(\sigma_\mathrm{HF}+\sigma)
++ (\sigma_\mathrm{HF}+\sigma)(\esp_\mathrm{HF}+\esp_{\mathrm{N}}(\tamp{},\tbar{})) \\
+&+ \frac{1}{2}(\kappa_\mathrm{HF}+\kappa)\MM(\kappa_\mathrm{HF}+\kappa)
++ (\kappa_\mathrm{HF}+\kappa)(\zeta_\mathrm{HF}+\zeta_{\mathrm{N}}(\tamp{},\tbar{}))
++ (\sigma_\mathrm{HF}+\sigma)\bi{X}(\kappa_\mathrm{HF}+\kappa) \\
+  %%%% Steps
+&= \left/ \text{Expand and collect} \right/ \\
+&=
+    \frac{1}{2}\sigma_\mathrm{HF}\PCM\sigma_\mathrm{HF}
+  + \sigma_\mathrm{HF}\esp_\mathrm{HF}
+  + \frac{1}{2}\kappa_\mathrm{HF}\MM\kappa_\mathrm{HF}
+  + \kappa_\mathrm{HF}\zeta_\mathrm{HF}
+  + \sigma_\mathrm{HF}\bi{X}\kappa_\mathrm{HF} \\
+&+  \frac{1}{2}\sigma\PCM\sigma
+  + \sigma\esp_{\mathrm{N}}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa\MM\kappa
+  + \kappa\zeta_{\mathrm{N}}(\tamp{},\tbar{})
+  + \sigma\bi{X}\kappa \\
+&+  \frac{1}{2}\sigma_\mathrm{HF}\PCM\sigma
+  + \sigma_\mathrm{HF}\esp_{\mathrm{N}}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa_\mathrm{HF}\MM\kappa
+  + \kappa_\mathrm{HF}\zeta_{\mathrm{N}}(\tamp{},\tbar{})
+  + \sigma_\mathrm{HF}\bi{X}\kappa \\
+&+  \frac{1}{2}\sigma\PCM\sigma_\mathrm{HF}
+  + \sigma\esp_\mathrm{HF}
+  + \frac{1}{2}\kappa\MM\kappa_\mathrm{HF}
+  + \kappa\zeta_\mathrm{HF}
+  + \sigma\bi{X}\kappa_\mathrm{HF} \\
+  %%%% Steps
+&= \left/
+  \frac{1}{2}\sigma_\mathrm{HF}\PCM\sigma_\mathrm{HF}
+  +
+  \frac{1}{2}\kappa_\mathrm{HF}\MM\kappa_\mathrm{HF}
+  +
+  \sigma_\mathrm{HF}\esp_\mathrm{HF}
+  +
+  \kappa_\mathrm{HF}\zeta_\mathrm{HF}
+  +
+  \sigma_\mathrm{HF}\bi{X}\kappa_\mathrm{HF}
+  =
+  \frac{1}{2}\sigma_\mathrm{HF}\esp_\mathrm{HF}
+  +
+  \frac{1}{2}\kappa_\mathrm{HF}\zeta_\mathrm{HF}
+  \right/ \\
+&=
+    \frac{1}{2}\sigma_\mathrm{HF}\esp_\mathrm{HF}
+  + \frac{1}{2}\kappa_\mathrm{HF}\zeta_\mathrm{HF} \\
+&+  \frac{1}{2}\sigma\PCM\sigma
+  + \sigma\esp_{\mathrm{N}}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa\MM\kappa
+  + \kappa\zeta_{\mathrm{N}}(\tamp{},\tbar{})
+  + \sigma\bi{X}\kappa \\
+&+  \frac{1}{2}\sigma_\mathrm{HF}\PCM\sigma
+  + \sigma_\mathrm{HF}\esp_{\mathrm{N}}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa_\mathrm{HF}\MM\kappa
+  + \kappa_\mathrm{HF}\zeta_{\mathrm{N}}(\tamp{},\tbar{})
+  + \sigma_\mathrm{HF}\bi{X}\kappa \\
+&+  \frac{1}{2}\sigma\PCM\sigma_\mathrm{HF}
+  + \sigma\esp_\mathrm{HF}
+  + \frac{1}{2}\kappa\MM\kappa_\mathrm{HF}
+  + \kappa\zeta_\mathrm{HF}
+  + \sigma\bi{X}\kappa_\mathrm{HF} \\
+  %%%% Steps
+&= \left/
+  \frac{1}{2}\sigma\PCM\sigma_\mathrm{HF}
+  =
+  \frac{1}{2}\sigma_\mathrm{HF}\PCM\sigma;\quad
+  \frac{1}{2}\kappa\MM\kappa_\mathrm{HF}
+  =
+  \frac{1}{2}\kappa_\mathrm{HF}\MM\kappa; \quad
+  \sigma_\mathrm{HF}\bi{X}\kappa
+  =
+  \kappa\bi{X}^\dagger\sigma_\mathrm{HF}
+  \right/ \\
+&=
+    \frac{1}{2}\sigma_\mathrm{HF}\esp_\mathrm{HF}
+  + \frac{1}{2}\kappa_\mathrm{HF}\zeta_\mathrm{HF} \\
+&+  \frac{1}{2}\sigma\PCM\sigma
+  + \sigma\esp_{\mathrm{N}}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa\MM\kappa
+  + \kappa\zeta_{\mathrm{N}}(\tamp{},\tbar{})
+  + \sigma\bi{X}\kappa \\
+  &+ \sigma(\cancelto{0}{\bi{Y}\sigma_\mathrm{HF} + \bi{X}\kappa_\mathrm{HF} + \esp_\mathrm{HF}})
+  + \kappa(\cancelto{0}{\bi{W}\kappa_\mathrm{HF} + \bi{X}^\dagger\sigma_\mathrm{HF} + \zeta_\mathrm{HF}}) \\
+&+ \sigma_\mathrm{HF}\esp_{\mathrm{N}}(\tamp{},\tbar{}) + \kappa_\mathrm{HF}\zeta_{\mathrm{N}}(\tamp{},\tbar{}) \\
+  %%%% Steps
+&= \left/
+    \frac{1}{2}\sigma_\mathrm{HF}\esp_\mathrm{HF}
+  + \frac{1}{2}\kappa_\mathrm{HF}\zeta_\mathrm{HF}
+  = U_\mathrm{pol}^\mathrm{ref}
+  \right/ \\
+&=
+  U_\mathrm{pol}^\mathrm{ref}
+  +  \frac{1}{2}\sigma\PCM\sigma
+  + \sigma\esp_{\mathrm{N}}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa\MM\kappa
+  + \kappa\zeta_{\mathrm{N}}(\tamp{},\tbar{})
+  + \sigma\bi{X}\kappa \\
+&+ \sigma_\mathrm{HF}\esp_{\mathrm{N}}(\tamp{},\tbar{}) + \kappa_\mathrm{HF}\zeta_{\mathrm{N}}(\tamp{},\tbar{})
+  \end{aligned}
+\end{equation}
+
+The reference polarization energy $U_\mathrm{pol}^\mathrm{ref}$ still
+appears among the terms in the functional but will obviously not enter
+in the optimization of the \acrshort{CC} wave function.
+Recasting the polarization functional in the supermatrix formalism (see
+Section \ref{sec:variational}) will simplify subsequent algebraic
+manipulations:
+\begin{equation}
+  \begin{aligned}
+  U_\mathrm{pol} &=
+    \frac{1}{2}\sigma\PCM\sigma
+    + \sigma\esp_{\mathrm{N}}(\tamp{},\tbar{})
+  + \frac{1}{2}\kappa\MM\kappa
+  + \kappa\zeta_{\mathrm{N}}(\tamp{},\tbar{})
+  + \sigma\bi{X}\kappa
+  + \sigma_\mathrm{HF}\esp_{\mathrm{N}}(\tamp{},\tbar{}) +
+  \kappa_\mathrm{HF}\zeta_{\mathrm{N}}(\tamp{},\tbar{})
+  + U_\mathrm{pol}^\mathrm{ref}
+  \\
+  &=
+  \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})
+  + {}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})
+  + U_\mathrm{pol}^\mathrm{ref}
+\end{aligned}
+\end{equation}
+
+\end{comment}

### 2016-09-09 Chapters/QuantumMechanics.tex
Almost done
Cleaning up solvation-correlation chapter

@@
  \quad&
- \zeta(\vect{r}) \coloneq \vect{\nabla}\rho\cdot \vect{\nabla}\rho.
+ \zeta(\vect{r}) = \vect{\nabla}\rho\cdot \vect{\nabla}\rho.
 \end{alignat}
@@
   v_\mathrm{xc}(\vect{r})
-  \coloneq
+  =
   \frac{\delta E_\mathrm{xc}[\rho(\vect{r})]}{\delta \rho(\vect{r})}
@@
 truncation level in the cluster operator the expectation was evaluated:
-$O^\mathcal{M}(\tamp{}, \tbar{})$.
+$O(\tamp{}, \tbar{})_{\mathrm{CC}\mathcal{M}}$.
 

### 2016-09-09 Chapters/SolvationAndCorrelation.tex
Almost done
Cleaning up solvation-correlation chapter

@@
 The contents of this Chapter are the result of an ongoing collaboration
-with prof.~T.~Daniel Crawford (Virginia Tech) and Dr.~Andrew
-C.~Simmonett (NIH)
+with prof. T. Daniel Crawford (Virginia Tech) and Dr. Andrew
+C. Simmonett (NIH)
 
@@
 derivations of \acrshort{MBPT} and \acrshort{CC} theory, encompassing
-previous work~\autocite{Christiansen1999-tj, Cammi2009-gu,
+previous work\autocite{Christiansen1999-tj, Cammi2009-gu,
 Caricato2010-hx, Caricato2011-tx, Olivares_del_Valle1991-of,
@@
 
-Iterative approximations to the \acrshort{CCSD} and \acrshort{CCSDT} models will
-be the subject of Section \ref{sec:cc2-and-cc3} where the equations
+Iterative and noniterative approximations to the \acrshort{CCSD} and \acrshort{CCSDT} models will
+be the subject of Section \ref{sec:approximate-cc-quantum-classical} where the equations
 governing the \acrshort{CC2} and \acrshort{CC3} models will be derived.
@@
 was optimized in the presence of the classical polarizable environment.
-Then the variational degrees of freedom corresponding to the environment
-satisfy:
-\begin{equation}
-  \begin{pmatrix}
-    \PCM & \MMPCM \\
-    \MMPCM^\dagger & \MM
-  \end{pmatrix}
-  \begin{pmatrix}
-   \sigma_\mathrm{HF} \\
-   \kappa_\mathrm{HF}
-  \end{pmatrix}
-  =
-  -
-  \begin{pmatrix}
-   \esp_\mathrm{HF} \\
-   \zeta_\mathrm{HF}
-  \end{pmatrix}
-\end{equation}
-and correspond to the \emph{reference reaction field}.
+Then the polarization degree of freedom corresponding to the environment
+satisfies:
+\begin{equation}
+  \V\p_\mathrm{HF} + \s_\mathrm{HF} = 0
+\end{equation}
+and corresponds to the \emph{reference reaction field}.
 In the subsequent correlated treatment one can separate the reaction
@@
 \begin{equation}
-  \begin{pmatrix}
-   \sigma^\mathrm{tot} \\
-   \kappa^\mathrm{tot}
-  \end{pmatrix}
-  =
-  \begin{pmatrix}
-   \sigma_\mathrm{HF} \\
-   \kappa_\mathrm{HF}
-  \end{pmatrix}
-  +
-  \begin{pmatrix}
-   \sigma \\
-   \kappa
-  \end{pmatrix}
-\end{equation}
-and similarly for the sources:
-\begin{equation}
-\begin{aligned}
-  \begin{pmatrix}
-   \esp(\tamp{}, \tbar{}) \\
-   \zeta(\tamp{}, \tbar{})
-  \end{pmatrix}
-  &=
-  \begin{pmatrix}
-   \esp_\mathrm{HF} \\
-   \zeta_\mathrm{HF}
-  \end{pmatrix}
-  +
-  \begin{pmatrix}
-    \esp_{\mathrm{N}}(\tamp{}, \tbar{}) \\
-    \zeta_{\mathrm{N}}(\tamp{}, \tbar{})
-  \end{pmatrix} \\
-  &=
-  \begin{pmatrix}
-  \braket{\mathrm{HF} | \esp  | \mathrm{HF}} \\
-  \braket{\mathrm{HF} | \zeta | \mathrm{HF}}
-  \end{pmatrix}
-  +
-  \begin{pmatrix}
-   \braket{\mathrm{HF} | \tilde{\esp} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} | \mathrm{HF}} \\
-   \braket{\mathrm{HF} | \tilde{\zeta} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} | \mathrm{HF}}
-  \end{pmatrix}
-\end{aligned}
-\end{equation}
-The \acrshort{BCH} expansions of the source terms have been
-rewritten as:
-\begin{subequations}
-  \begin{align}
-  \overline{\esp} &= \esp + \tilde{\esp} = \esp + \commutator{\esp}{T}
-  + \frac{1}{2}\commutator{\commutator{\esp}{T}}{T} \\
-  \overline{\kappa} &= \kappa + \tilde{\kappa} = \kappa + \commutator{\kappa}{T}
-  + \frac{1}{2}\commutator{\commutator{\kappa}{T}}{T}
-  \end{align}
-\end{subequations}
-taking advantage of the fact that both $\esp$ and $\kappa$ are
-nondiagonal, one-electron operators and hence their commutator
-expansions truncate at the third term. Moreover, normal-ordering of the
+  \p^\mathrm{tot} = \p_\mathrm{HF} + \p
+\end{equation}
+and similarly for the source term:
+\begin{equation}
+  \begin{aligned}
+\s(\tamp{}, \tbar{}) &= \s_\mathrm{HF} + \s_{\mathrm{N}}(\tamp{}, \tbar{}) \\
+                     &=
+    \braket{\mathrm{HF} | \s | \mathrm{HF}}
+  + \braket{\mathrm{HF} | \tilde{\s} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\s} | \mathrm{HF}}
+  \end{aligned}
+\end{equation}
+The \acrshort{BCH} expansions of the source term has been rewritten as:
+\begin{equation}
+  \overline{\s} = \s + \tilde{\s} = \s + \commutator{\s}{T}
+  + \frac{1}{2}\commutator{\commutator{\s}{T}}{T}
+\end{equation}
+taking advantage of the fact that $\s$ is a
+nondiagonal, one-electron operators and hence its commutator
+expansion truncates at the third term. Moreover, normal-ordering of the
 operators has been introduced.~\autocite{Crawford2000-ey, Shavitt2009-mr}
@@
   \begin{aligned}
-  \begin{pmatrix}
-    \esp_{\mathrm{N}}(\tamp{}, \tbar{}) \\
-    \zeta_{\mathrm{N}}(\tamp{}, \tbar{})
-  \end{pmatrix}
-  &=
-  \begin{pmatrix}
-   \braket{\mathrm{HF} | \tilde{\esp} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} | \mathrm{HF}} \\
-   \braket{\mathrm{HF} | \tilde{\zeta} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} | \mathrm{HF}}
-  \end{pmatrix}
-  \\
-  &=
-  \begin{pmatrix}
-    \braket{\mathrm{HF} | \overline{\esp} - \esp_\mathrm{HF} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\esp} - \esp_\mathrm{HF} | \mathrm{HF}} \\
-  \braket{\mathrm{HF} | \overline{\zeta} - \zeta_\mathrm{HF} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\zeta} -
-  \zeta_\mathrm{HF}| \mathrm{HF}}
-  \end{pmatrix}
+    \s_{\mathrm{N}}(\tamp{}, \tbar{})
+  &=
+   \braket{\mathrm{HF} | \tilde{\s} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\s} | \mathrm{HF}} \\
+  &=
+    \braket{\mathrm{HF} | \overline{\s} - \s_\mathrm{HF} | \mathrm{HF}}
+  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\s} - \s_\mathrm{HF} | \mathrm{HF}}
   \end{aligned}
@@
   \begin{aligned}
-  U_\mathrm{pol} &=
-  \frac{1}{2}(\sigma_\mathrm{HF}+\sigma)\PCM(\sigma_\mathrm{HF}+\sigma)
-+ (\sigma_\mathrm{HF}+\sigma)(\esp_\mathrm{HF}+\esp_{\mathrm{N}}(\tamp{},\tbar{})) \\
-&+ \frac{1}{2}(\kappa_\mathrm{HF}+\kappa)\MM(\kappa_\mathrm{HF}+\kappa)
-+ (\kappa_\mathrm{HF}+\kappa)(\zeta_\mathrm{HF}+\zeta_{\mathrm{N}}(\tamp{},\tbar{}))
-+ (\sigma_\mathrm{HF}+\sigma)\bi{X}(\kappa_\mathrm{HF}+\kappa) \\
-  %%%% Steps
-&= \left/ \text{Expand and collect} \right/ \\
-&=
-    \frac{1}{2}\sigma_\mathrm{HF}\PCM\sigma_\mathrm{HF}
-  + \sigma_\mathrm{HF}\esp_\mathrm{HF}
-  + \frac{1}{2}\kappa_\mathrm{HF}\MM\kappa_\mathrm{HF}
-  + \kappa_\mathrm{HF}\zeta_\mathrm{HF}
-  + \sigma_\mathrm{HF}\bi{X}\kappa_\mathrm{HF} \\
-&+  \frac{1}{2}\sigma\PCM\sigma
-  + \sigma\esp_{\mathrm{N}}(\tamp{},\tbar{})
-  + \frac{1}{2}\kappa\MM\kappa
-  + \kappa\zeta_{\mathrm{N}}(\tamp{},\tbar{})
-  + \sigma\bi{X}\kappa \\
-&+  \frac{1}{2}\sigma_\mathrm{HF}\PCM\sigma
-  + \sigma_\mathrm{HF}\esp_{\mathrm{N}}(\tamp{},\tbar{})
-  + \frac{1}{2}\kappa_\mathrm{HF}\MM\kappa
-  + \kappa_\mathrm{HF}\zeta_{\mathrm{N}}(\tamp{},\tbar{})
-  + \sigma_\mathrm{HF}\bi{X}\kappa \\
-&+  \frac{1}{2}\sigma\PCM\sigma_\mathrm{HF}
-  + \sigma\esp_\mathrm{HF}
-  + \frac{1}{2}\kappa\MM\kappa_\mathrm{HF}
-  + \kappa\zeta_\mathrm{HF}
-  + \sigma\bi{X}\kappa_\mathrm{HF} \\
-  %%%% Steps
-&= \left/
-  \frac{1}{2}\sigma_\mathrm{HF}\PCM\sigma_\mathrm{HF}
-  +
-  \frac{1}{2}\kappa_\mathrm{HF}\MM\kappa_\mathrm{HF}
-  +
-  \sigma_\mathrm{HF}\esp_\mathrm{HF}
-  +
-  \kappa_\mathrm{HF}\zeta_\mathrm{HF}
-  +
-  \sigma_\mathrm{HF}\bi{X}\kappa_\mathrm{HF}
-  =
-  \frac{1}{2}\sigma_\mathrm{HF}\esp_\mathrm{HF}
-  +
-  \frac{1}{2}\kappa_\mathrm{HF}\zeta_\mathrm{HF}
-  \right/ \\
-&=
-    \frac{1}{2}\sigma_\mathrm{HF}\esp_\mathrm{HF}
-  + \frac{1}{2}\kappa_\mathrm{HF}\zeta_\mathrm{HF} \\
-&+  \frac{1}{2}\sigma\PCM\sigma
-  + \sigma\esp_{\mathrm{N}}(\tamp{},\tbar{})
-  + \frac{1}{2}\kappa\MM\kappa
-  + \kappa\zeta_{\mathrm{N}}(\tamp{},\tbar{})
-  + \sigma\bi{X}\kappa \\
-&+  \frac{1}{2}\sigma_\mathrm{HF}\PCM\sigma
-  + \sigma_\mathrm{HF}\esp_{\mathrm{N}}(\tamp{},\tbar{})
-  + \frac{1}{2}\kappa_\mathrm{HF}\MM\kappa
-  + \kappa_\mathrm{HF}\zeta_{\mathrm{N}}(\tamp{},\tbar{})
-  + \sigma_\mathrm{HF}\bi{X}\kappa \\
-&+  \frac{1}{2}\sigma\PCM\sigma_\mathrm{HF}
-  + \sigma\esp_\mathrm{HF}
-  + \frac{1}{2}\kappa\MM\kappa_\mathrm{HF}
-  + \kappa\zeta_\mathrm{HF}
-  + \sigma\bi{X}\kappa_\mathrm{HF} \\
-  %%%% Steps
-&= \left/
-  \frac{1}{2}\sigma\PCM\sigma_\mathrm{HF}
-  =
-  \frac{1}{2}\sigma_\mathrm{HF}\PCM\sigma;\quad
-  \frac{1}{2}\kappa\MM\kappa_\mathrm{HF}
-  =
-  \frac{1}{2}\kappa_\mathrm{HF}\MM\kappa; \quad
-  \sigma_\mathrm{HF}\bi{X}\kappa
-  =
-  \kappa\bi{X}^\dagger\sigma_\mathrm{HF}
-  \right/ \\
-&=
-    \frac{1}{2}\sigma_\mathrm{HF}\esp_\mathrm{HF}
-  + \frac{1}{2}\kappa_\mathrm{HF}\zeta_\mathrm{HF} \\
-&+  \frac{1}{2}\sigma\PCM\sigma
-  + \sigma\esp_{\mathrm{N}}(\tamp{},\tbar{})
-  + \frac{1}{2}\kappa\MM\kappa
-  + \kappa\zeta_{\mathrm{N}}(\tamp{},\tbar{})
-  + \sigma\bi{X}\kappa \\
-  &+ \sigma(\cancelto{0}{\bi{Y}\sigma_\mathrm{HF} + \bi{X}\kappa_\mathrm{HF} + \esp_\mathrm{HF}})
-  + \kappa(\cancelto{0}{\bi{W}\kappa_\mathrm{HF} + \bi{X}^\dagger\sigma_\mathrm{HF} + \zeta_\mathrm{HF}}) \\
-&+ \sigma_\mathrm{HF}\esp_{\mathrm{N}}(\tamp{},\tbar{}) + \kappa_\mathrm{HF}\zeta_{\mathrm{N}}(\tamp{},\tbar{}) \\
-  %%%% Steps
-&= \left/
-    \frac{1}{2}\sigma_\mathrm{HF}\esp_\mathrm{HF}
-  + \frac{1}{2}\kappa_\mathrm{HF}\zeta_\mathrm{HF}
-  \coloneq U_\mathrm{pol}^\mathrm{ref}
-  \right/ \\
+  U_\mathrm{pol}
+  &=
+  \frac{1}{2}({}^t\p_\mathrm{HF}+{}^t\p)\V(\p_\mathrm{HF}+\p)
+  + ({}^t\p_\mathrm{HF}+{}^t\p)(\s_\mathrm{HF}+\s_{\mathrm{N}}(\tamp{},\tbar{})) \\
+  &=
+    \frac{1}{2}{}^t\p_\mathrm{HF}\V\p_\mathrm{HF} + {}^t\p_\mathrm{HF}\s_\mathrm{HF}
+  + \frac{1}{2}{}^t\p_\mathrm{HF}\V\p + \frac{1}{2}{}^t\p\V\p_\mathrm{HF} \\
+  &+ {}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})
+  + {}^t\p\s_\mathrm{HF}
+  + \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{}) \\
   &=
     U_\mathrm{pol}^\mathrm{ref}
-  +  \frac{1}{2}\sigma\PCM\sigma
-  + \sigma\esp_{\mathrm{N}}(\tamp{},\tbar{})
-  + \frac{1}{2}\kappa\MM\kappa
-  + \kappa\zeta_{\mathrm{N}}(\tamp{},\tbar{})
-  + \sigma\bi{X}\kappa \\
-&+ \sigma_\mathrm{HF}\esp_{\mathrm{N}}(\tamp{},\tbar{}) + \kappa_\mathrm{HF}\zeta_{\mathrm{N}}(\tamp{},\tbar{})
-  \end{aligned}
-\end{equation}
-
+  + \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})
+  + {}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})
+  \end{aligned}
+\end{equation}
 The reference polarization energy $U_\mathrm{pol}^\mathrm{ref}$ still
@@
 in the optimization of the \acrshort{CC} wave function.
-Recasting the polarization functional in the supermatrix formalism (see
-Section \ref{sec:variational}) will simplify subsequent algebraic
-manipulations:
-\begin{equation}
-  \begin{aligned}
-  U_\mathrm{pol} &=
-    \frac{1}{2}\sigma\PCM\sigma
-    + \sigma\esp_{\mathrm{N}}(\tamp{},\tbar{})
-  + \frac{1}{2}\kappa\MM\kappa
-  + \kappa\zeta_{\mathrm{N}}(\tamp{},\tbar{})
-  + \sigma\bi{X}\kappa
-  + \sigma_\mathrm{HF}\esp_{\mathrm{N}}(\tamp{},\tbar{}) +
-  \kappa_\mathrm{HF}\zeta_{\mathrm{N}}(\tamp{},\tbar{})
-  + U_\mathrm{pol}^\mathrm{ref}
-  \\
-  &=
-  \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})
-  + {}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})
-  + U_\mathrm{pol}^\mathrm{ref}
-\end{aligned}
-\end{equation}
-
 One can then write the effective Lagrangian as:
@@
 \acrlong*{PTE} (\acrshort{PTE}) Hamiltonian -- $H$ -- and eventually the
-\acrlong*{PTED} (\acrshort{PTED}) Lagrangian:~\autocite{Olivares del Valle,
-Lipparini}
+\acrlong*{PTED} (\acrshort{PTED}) Lagrangian:\autocite{Olivares_del_Valle1991-of,
+Aguilar1991-vq, Olivares_del_Valle1991-tq, Olivares_del_Valle1993-xq,
+Olivares_del_Valle1993-ra, Lipparini2009-io}
 \begin{equation}\label{eq:pted-cc}
@@
 \begin{equation}
-  \s_{\mathrm{N}}(\tamp{}, \tbar{}) =
-  \braket{\mathrm{HF} | \BCHfirst{\s}{T} + \BCHsecond{\s}{T} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \s + \BCHfirst{\s}{T} +
+  \begin{aligned}
+  \s_{\mathrm{N}}(\tamp{}, \tbar{}) &=
+  \braket{\mathrm{HF} | \BCHfirst{\s}{T} + \BCHsecond{\s}{T} | \mathrm{HF}} \\
+  &+ \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \s + \BCHfirst{\s}{T} +
   \BCHsecond{\s}{T} | \mathrm{HF}},
+  \end{aligned}
 \end{equation}
@@
   \BCHfirst{\s}{T_t}
-  | \mathrm{HF}}
-  &\quad\quad\text{nonzero contributions} \quad 0 \leq t \leq 1
-  \Longleftrightarrow t = 1 \\
+  | \mathrm{HF}} &\neq 0 \quad \text{iff} \quad 0 \leq t \leq 1 \Longleftrightarrow t = 1 \\
   \braket{\mathrm{HF} |
   \commutator{\commutator{\s}{T_t}}{T_u}
-  | \mathrm{HF}}
-  &\quad\quad\text{nonzero contributions} \quad t + u = 1
+  | \mathrm{HF}} &\neq 0 \quad \text{iff} \quad t + u = 1
   \label{eq:HF-double-commutator} \\
@@
   \s
-  | \mathrm{HF} }
-  &\quad\quad\text{nonzero contributions} \quad i = 1 \\
+  | \mathrm{HF} } &\neq 0 \quad \text{iff} \quad i = 1 \\
   \braket{\tbar{i} |
   \BCHfirst{\s}{T_t}
-  | \mathrm{HF} }
-  &\quad\quad\text{nonzero contributions} \quad i \leq t \leq i + 1 \\
+  | \mathrm{HF} } &\neq 0 \quad \text{iff} \quad i \leq t \leq i + 1 \\
   \braket{\tbar{i} |
   \commutator{\commutator{\s}{T_t}}{T_u}
-  | \mathrm{HF} }
-  &\quad\quad\text{nonzero contributions} \quad t + u = i + 1
+  | \mathrm{HF} } &\neq 0 \quad \text{iff} \quad t + u = i + 1
 \end{align}
@@
 operators can only lead to nonzero terms when $T_1$ is also involved and
-by projection onto the highest order multiplier state, \eg
+by projection onto the highest order multiplier state, \eg~
 $\braket{\tbar{2} | \commutator{\commutator{\s}{T_1}}{T_2} |
@@
   \begin{aligned}
-  \s_{\mathrm{N}}^1(\tamp{}, \tbar{}) &=
+    \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCS} &=
   \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
@@
   \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
-  + \braket{\tbar{1} | \s | \mathrm{HF}}
-  + \braket{\tbar{1} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
+  + \braket{\tbar{1} | \s | \mathrm{HF}} \\
+  &+ \braket{\tbar{1} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
   + \frac{1}{2}\braket{\tbar{1} | \commutator{\commutator{\s}{T_1}}{T_1} | \mathrm{HF}}
@@
 \begin{equation}
-  \s_{\mathrm{N}}^2(\tamp{}, \tbar{}) =
-  \s_{\mathrm{N}}^1(\tamp{}, \tbar{})
-  + \braket{\tbar{1} | \BCHfirst{\s}{T_2} | \mathrm{HF}}
-  + \braket{\tbar{2} | \BCHfirst{\s}{T_2} | \mathrm{HF}}
+  \begin{aligned}
+  \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCSD} &=
+  \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCS}
+  + \braket{\tbar{1} | \BCHfirst{\s}{T_2} | \mathrm{HF}} \\
+  &+ \braket{\tbar{2} | \BCHfirst{\s}{T_2} | \mathrm{HF}}
   + \braket{\tbar{2} | \commutator{\commutator{\s}{T_1}}{T_2} | \mathrm{HF}}
+  \end{aligned}
 \end{equation}
@@
   \begin{aligned}
-  \s_{\mathrm{N}}^3(\tamp{}, \tbar{}) &=
-  \s_{\mathrm{N}}^2(\tamp{}, \tbar{})
-  + \braket{\tbar{2} | \BCHfirst{\s}{T_3} | \mathrm{HF}} \\
-  &+ \braket{\tbar{3} | \BCHfirst{\s}{T_3} | \mathrm{HF}}
-  + \frac{1}{2}
+    \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCSDT} &=
+  \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCSD}
+  + \braket{\tbar{2} | \BCHfirst{\s}{T_3} | \mathrm{HF}}
+  + \braket{\tbar{3} | \BCHfirst{\s}{T_3} | \mathrm{HF}} \\
+  &+ \frac{1}{2}
   \braket{\tbar{3} | \commutator{\commutator{\s}{T_2}}{T_2} |
@@
 \begin{equation}\label{eq:ptes-cc}
+  \begin{aligned}
   {}^{\mathrm{PTES}}\lag{M}(\tamp{}, \tbar{}, \p) =
   \braket{\mathrm{HF} | \overline{H} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}}
-  + \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}^1(\tamp{}, \tbar{})
+  &+ \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}} \\
+  &+ \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},
+  \tbar{})_\mathrm{CCS}
+  \end{aligned}
 \end{equation}
@@
     &=
-    \V\p + \s_{\mathrm{N}}^1(\tamp{}, \tbar{}) = 0
+    \V\p + \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCS} = 0
   \end{align}
@@
 multipliers in the \acrshort{CCS} source term in the polarization equation.
-To break the coupling, we further approximate $\s_{\mathrm{N}}^1(\tamp{}, \tbar{})$
+To break the coupling, we further approximate $\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathrm{CCS}$
 with its multiplier-independent part:
 \begin{equation}
-  \s_{\mathrm{N}}^1(\tamp{}, \tbar{})
+  \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCS}
   \simeq
@@
 \begin{equation}
-  \s_{\mathrm{N}}^1(\tamp{}, \tbar{})
+  \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCS}
   \simeq
@@
 \begin{equation}\label{eq:pte-s-cc}
+  \begin{aligned}
   {}^{\mathrm{PTE(S)}}\lag{M}(\tamp{}, \tbar{}, \p) =
   \braket{\mathrm{HF} | \overline{H} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}}
-  + \frac{1}{2}{}^t\p\V\p + {}^t\p\braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}}
+  &+ \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}} \\
+  &+ \frac{1}{2}{}^t\p\V\p + {}^t\p\braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}}
+  \end{aligned}
 \end{equation}
@@
 
-\section[Coupled Cluster Many-Body Perturbation Theory]{
-Coupled Cluster Many-Body Perturbation Theory}\label{sec:mbpt-polarizable}
+\section[MBPT for Quantum/Classical Polarizable Hamiltonians]{
+MBPT for Quantum/Classical Polarizable Hamiltonians}\label{sec:mbpt-polarizable}
 
@@
 
-\section[Iterative Approximations: CC2 and CC3]{Iterative Approximations to CCSD and CCSDT: the CC2 and CC3 Models}\label{sec:cc2-and-cc3}
-
-\section[Noniterative Triples Corrections to PCM-CCSD]{
-Noniterative Perturbative Triples Corrections to PCM-CCSD}\label{sec:ccsd-t-noniterative}
+\section[Approximate Coupled Cluster Methods]{
+Approximate Coupled Cluster Methods with Quantum/Classical Polarizable Hamiltonians
+}\label{sec:cc-approximate-quantum-classical}
 
@@
   = \sum_{\mu_3}\tilde{t}_{\mu_3}\cluster{\mu_3}
-  \coloneq \sum_{\mu_3}
+  = \sum_{\mu_3}
   \left(-\denom{\mu_3}^{-1}\braket{\mu_3 |

### 2016-09-11 Chapters/SolvationAndCorrelation.tex
First draft sent to L.F. and B.M.

@@
 Iterative and noniterative approximations to the \acrshort{CCSD} and \acrshort{CCSDT} models will
-be the subject of Section \ref{sec:approximate-cc-quantum-classical} where the equations
+be the subject of Section \ref{sec:cc-approximate-quantum-classical} where the equations
 governing the \acrshort{CC2} and \acrshort{CC3} models will be derived.
@@
 More explicitly, for the \gls{CCS} model ($\mathcal{M} = 1$) one has:
-\begin{equation}
+\begin{equation}\label{eq:source-ccs}
   \begin{aligned}
@@
 while for the \gls{CCSD} model ($\mathcal{M} = 2$):
-\begin{equation}
+\begin{equation}\label{eq:source-ccsd}
   \begin{aligned}
@@
 terms are added:
-\begin{equation}
+\begin{equation}\label{eq:source-ccsdt}
   \begin{aligned}
@@
 equations:
-\begin{subequations}
+\begin{subequations}\label{eq:pted}
   \begin{align}
@@
 singles cluster operator and the singles multiplier state.
-Three approximation have currently been proposed: \acrshort{PTES}, \acrshort{PTE(S)}
-and \acrshort{PTE}.
+Three approximation have currently been proposed: \gls{PTES}, \gls{PTE(S)}
+and \gls{PTE}.
 
@@
 
-\todo[inline]{This is my claim. I am waiting on an answer from MC to
-clarify this section.}
-
-The first approximate scheme truncates the source term to its \acrshort{CCS}
-expectation value, yielding the effective Lagrangian:
+The first approximate scheme truncates the source term to its
+\acrshort{CCS} expectation value, yielding the effective
+Lagrangian:\autocite{Caricato2011-tx, Krause2016-ee}
 \begin{equation}\label{eq:ptes-cc}
@@
 and the corresponding governing equations:
-\begin{subequations}
+\begin{subequations}\label{eq:ptes}
   \begin{align}
@@
     \V\p + \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCS} = 0
+    \label{eq:ptes-cc-polarization-amp}
   \end{align}
@@
 multipliers in the \acrshort{CCS} source term in the polarization equation.
-To break the coupling, we further approximate $\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathrm{CCS}$
-with its multiplier-independent part:
+To break the coupling, we reformulate the split the system of equations
+\eqref{eq:ptes} into \emph{two} systems of equations, one for the
+amplitudes:
+\begin{subequations}\label{eq:ptes-amp}
+  \begin{align}
+   {}^{\mathrm{PTES}}\tampEq{\mu_i}(\tamp{}, \tbar{}, \p)  &=
+   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
+   + {}^t\p\braket{\mu_i | \check{\s} | \mathrm{HF}}
+             = 0 \label{eq:ptes-cc-amplitudes} \\
+    {}^{\mathrm{PTES}}\Omega_\p(\tamp{}, \p)
+    &=
+    \V\p +
+    \braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}} = 0
+  \end{align}
+\end{subequations}
+where we further approximate the source term
+$\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathrm{CCS}$ with its
+multiplier-independent part:
 \begin{equation}
@@
   =
-  \braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}}
-\end{equation}
-so that the governing equations for the \acrshort{PTES} scheme become:
-\begin{subequations}
+  \braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}},
+\end{equation}
+and one for the multipliers:
+\begin{subequations}\label{eq:ptes-mult}
   \begin{align}
-   {}^{\mathrm{PTES}}\tampEq{\mu_i}(\tamp{}, \tbar{}, \p)  &=
-   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
-   + {}^t\p\braket{\mu_i | \check{\s} | \mathrm{HF}}
-             = 0 \label{eq:ptes-cc-amplitudes}\\
    {}^{\mathrm{PTES}}\tbarEq{\mu_i}(\tamp{}, \tbar{}, \p)
@@
     &=
-    \V\p + \braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}} = 0
-    \label{eq:ptes-cc-polarization}
+    \V\p + \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCS} = 0
+    \label{eq:ptes-cc-polarization-mult}
   \end{align}
 \end{subequations}
-
-\todo[inline]{This thing looks very inconsistent with electrostatics...
-Moreover, it doesn't add up with Caricato's derivation. Maybe there are
-\emph{two different} polarization equations that are used? One when
-solving for the amplitudes and one when solving for the multipliers?}
+where no truncation for the source term is introduced.
+Both the amplitudes and multipliers equations are formally identical to
+the \acrshort*{PTED} equations \eqref{eq:pted} but are coupled to
+approximate polarization equations.
+For the amplitudes equation \eqref{eq:ptes-cc-amplitudes}, one computes
+the polarization from a modified \acrshort*{CCS} expectation value of
+the source term. This adds a negligible $O(N^5)$ term to the iterative
+solution of the equations.
+The amplitudes equations are instead coupled to the polarization
+obtained using the full \acrshort*{CCS} expectation value of the source
+term. However, this adds another $O(N^5)$ step to the iterative
+procedure, which is, overall, negligible.
+Let us notice that this scheme is not consistent with electrostatics, as
+the amplitudes and multipliers are optimized in the presence of two
+different polarizations.
 
@@
 truncating the \acrshort{CCS} expectation value to its multiplier-independent
-term:
-\begin{equation}
-  \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCS}
-  \simeq
-  \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
-  =
-  \braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}}
-\end{equation}
-the coupling between the amplitudes and multipliers is broken, while
+term already in the effective Lagrangian.
+The coupling between the amplitudes and multipliers is broken, while
 still improving upon the reference reaction field with correlated
-contributions.~\autocite{Caricato2011-tx}
+contributions.\autocite{Caricato2011-tx}
 The effective Lagrangian would then read as:
@@
 We further assume that real wave functions are used.
-\todo[inline]{Is this a necessary assumption??}
 
@@
 \end{subequations}
-Thus, despite the closed-shell \acrshort{HF} reference, the singles
-amplitudes will already appear in first order, due to the
-quantum/classical coupling.
+Having chosen a closed-shell \acrshort{HF} as reference,
+the term $\braket{\mu_1 | \Phi | \mathrm{HF}}$ is zero in the
+first-order singles equation. However, due to the quantum/classical
+coupling, there will still be nonzero first-order singles amplitudes.
 
@@
 \paragraph*{Polarization}
-To first order, the polarization degree of freedom will be given as:
 \begin{equation}%\label{eq:1st-order-pol}
@@
 first order amplitudes and multipliers yields:
+\begin{equation}
+  \braket{\mathrm{HF} | \commutator{\s}{T_1^{[1]}}| \mathrm{HF}}
+  = \braket{\tbar{1}^{[1]} | \s | \mathrm{HF}}
+\end{equation}
+so that the first order polarization equation becomes:
+\begin{equation}\label{eq:1st-order-pol}
+  \V\p^{[1]} + 2\braket{\mathrm{HF} | \commutator{\s}{T_1^{[1]}}| \mathrm{HF}} = 0
+\end{equation}
+
+\begin{comment}
 \begin{equation}
@@
 \end{equation}
-so that the first order polarization equation becomes:
-\begin{equation}\label{eq:1st-order-pol}
-  \V\p^{[1]} + 2\braket{\mathrm{HF} | \commutator{\s}{T_1^{[1]}}| \mathrm{HF}} = 0
-\end{equation}
+\end{comment}
 
@@
 \begin{equation}\label{eq:2nd-order-amp}
-  \Omega_{\mu_i}^{[2]} = \denom{\mu_i}\tamp{\mu_i}^{[2]}
+  \begin{aligned}
+  \Omega_{\mu_i}^{[2]} &= \denom{\mu_i}\tamp{\mu_i}^{[2]}
   + \braket{\mu_i | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
-  + {}^t\p^{[2]}\braket{\mu_i | \s | \mathrm{HF}}\delta_{\mu_i\mu_1}
-  + {}^t\p^{[1]}\braket{\mu_i | \commutator{\s}{T^{[1]}} |
+  + {}^t\p^{[2]}\braket{\mu_i | \s | \mathrm{HF}}\delta_{\mu_i\mu_1} \\
+  &+ {}^t\p^{[1]}\braket{\mu_i | \commutator{\s}{T^{[1]}} |
   \mathrm{HF}}
   = 0
-\end{equation}
-Considering the singles, doubles and triples excitation manifolds:
+  \end{aligned}
+\end{equation}
+As already noted elsewhere,\autocite{Koch1997-nm, Helgaker2000-tz}
+the second term in Eq. \eqref{eq:2nd-order-amp} can involve no higher
+than triple excitations.
+Moreover, the triples first appear to second order and are not
+\emph{directly} affected by the quantum/classical coupling, as can be
+seen by explicitly considering the singles, doubles and triples
+excitation manifolds:
 \begin{subequations}
@@
   + \braket{\mu_1 | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
-  + {}^t\p^{[2]}\braket{\mu_1 | \s | \mathrm{HF}}
-  + {}^t\p^{[1]}\braket{\mu_1 | \commutator{\s}{T^{[1]}} |
+  + {}^t\p^{[2]}\braket{\mu_1 | \s | \mathrm{HF}} \nonumber \\
+  &+ {}^t\p^{[1]}\braket{\mu_1 | \commutator{\s}{T^{[1]}} |
   \mathrm{HF}}
@@
 \end{subequations}
-As already noted elsewhere~\autocite{Koch1997-nm, Helgaker2000-tz},
-the second term in Eq.~\eqref{eq:2nd-order-amp} can involve no higher
-than triple excitations.
-Moreover, the triples first appear to second order and are not \emph{directly}
-affected by the quantum/classical coupling.
 
 \paragraph*{Multipliers}
-\begin{equation}
+\begin{equation}\label{eq:2nd-order-mult}
   \begin{aligned}
@@
 \end{equation}
-The double commutator term is null thus:~\autocite{Koch1997-nm, Helgaker2000-tz}
-\todo[inline]{Check this statement... I am not 100\% sure and I have
-found no derivation...}
-\begin{equation}\label{eq:2nd-order-mult}
-  \begin{aligned}
-  \bar{\Omega}_{\mu_i}^{[2]} &=
-    \denom{\mu_i}\tbar{\mu_i}^{[2]}
-    + \sum_j\braket{\tbar{j}^{[1]} |
-    \commutator{\Phi}{\cluster{\mu_i}} | \mathrm{HF}} \\
-    &+ {}^t\p^{[2]}\braket{\mathrm{HF} |
-    \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}\delta_{\mu_i\mu_1}
-    + {}^t\p^{[1]}
-    \sum_j\braket{\tbar{j}^{[1]} |
-    \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}
-    =0
-  \end{aligned}
-\end{equation}
-\citeauthor{Koch1997-nm} showed that the \emph{in vacuo} second order
-amplitudes and multipliers equations are the complex conjugates of each
-other for the singles, doubles and tripled manifold. Under the further
-assumption of real wave functions thus $\tamp{\mu_i}^{[2]} =
-\tbar{\mu_i}^{[2]},\,\,\forall i = 1 - 3$.
-A similar conclusion holds also for the quantum/classical polarizable
-Hamiltonians here considered.
-First of all, the terms involving the second order polarization degree of
-freedom are identical:
-\begin{equation}
-    {}^t\p^{[2]}\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}\delta_{\mu_i\mu_1}
-    =
-    {}^t\p^{[2]}\braket{\mathrm{HF} | \s|\mu_i}\delta_{\mu_i\mu_1}
-    =
-    {}^t\p^{[2]}\braket{\mu_i| \s|\mathrm{HF}}\delta_{\mu_i\mu_1}
-\end{equation}
-since they are complex conjugates and real wave functions are assumed.
-Under the same set of assumptions, the first order terms are also identical:
-\begin{equation}
-  \begin{aligned}
-  {}^t\p^{[1]} \sum_j\braket{\tbar{j}^{[1]} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}
-  %%%% Steps
-  &= \left/ \sum_j\bra{\tbar{j}^{[1]}} = \sum_j\sum_{\mu_j}\tbar{\mu_j}\bra{\mu_j}  \right/ \\
-  &=
-  {}^t\p^{[1]} \sum_j\sum_{\mu_j}\tbar{\mu_j}^{[1]} \braket{\mu_j| \s | \mu_i}
-  %%%% Steps
-  = \left/ \text{Complex conjugation and real wave functions} \right/ \\
-  &=
-  {}^t\p^{[1]} \sum_j\sum_{\mu_j} \braket{\mu_i| \s | \mu_j} \tamp{\mu_j}^{[1]}
-  %%%% Steps
-  = \left/ \sum_j\sum_{\mu_j}\tamp{\mu_j}\ket{\mu_j}  = T^{[1]}\right/ \\
-  &=
-  {}^t\p^{[1]}\braket{\mu_i | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
- \end{aligned}
-\end{equation}
-
-Considering the singles, doubles and triples excitation manifolds:
+Since first-order singles amplitudes are now nonzero, the double
+commutator term will contribute to the singles multipliers equation.
+This is markedly in contrast with the derivation \emph{in vacuo}, see
+Section \ref{sec:mbpt}.\autocite{Koch1997-nm, Helgaker2000-tz}
+For the singles, doubles and triples manifolds the multipliers equations
+to second order are given as:
 \begin{subequations}
@@
       \denom{\mu_1}\tbar{\mu_1}^{[2]}
+      + \braket{\mathrm{HF} |
+      \commutator{\commutator{\Phi}{T_1^{[1]}}}{\cluster{\mu_1}} | \mathrm{HF}}
       + \sum_j\braket{\tbar{j}^{[1]} |
@@
       + {}^t\p^{[1]}
-      \braket{\tbar{1}^{[1]} + \tbar{2}^{[1]} |
+      \sum_j\braket{\tbar{j}^{[1]} |
       \commutator{\s}{\cluster{\mu_2}} | \mathrm{HF}}
@@
 \end{subequations}
-The triples first appear to second order and are not \emph{directly}
-affected by the quantum/classical coupling.
+\citeauthor{Koch1997-nm} showed that, assuming real wave functions, the
+\emph{in vacuo} second order amplitudes and multipliers equations are
+the complex conjugates of each other for the singles, doubles and
+triples manifolds.\autocite{Koch1997-nm, Helgaker2000-tz,
+Shavitt2009-mr}
+This symmetry is a consequence of choosing a closed-shell \acrshort*{HF}
+determinant as reference function\autocite{Shavitt2009-mr} and is
+\emph{broken} by the quantum/classical coupling.
+As an example, consider the triples multipliers equation. \emph{In vacuo}
+one has:
+\begin{equation}\tag{\ref{eq:2nd-order} from Chapter \ref{ch:QM}}
+  \denom{\mu_3}\tamp{\mu_3}^{[2]}
+  = - \braket{\mu_i | \commutator{\Phi}{T_2^{[1]}} | \mathrm{HF}}
+  = - \braket{\tbar{2}^{[1]} | \commutator{\Phi}{\cluster{\mu_3}} |
+  \mathrm{HF}}
+  = \denom{\mu_3}\tbar{\mu_3}^{[2]},
+\end{equation}
+whereas now the first-order singles multiplier state also contributes to
+the second-order triples multipliers.
 
 \paragraph*{Polarization}
-\begin{equation}
-  \V\p^{[2]} +
+\begin{equation}\label{eq:2nd-order-pol}
+  \begin{aligned}
+  \V\p^{[2]} &+
   \braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}}| \mathrm{HF}}
-  + \braket{\tbar{1}^{[2]} | \s | \mathrm{HF} }
-  + \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
+  + \braket{\tbar{1}^{[2]} | \s | \mathrm{HF} } \\
+  &+ \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
   + \braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[1]}} | \mathrm{HF}}
   = 0
-\end{equation}
-There are no explicit contributions from triples. This means that the
-\acrshort{CCSD} polarization is correct through fourth order in the
+  \end{aligned}
+\end{equation}
+There are no explicit contributions from the triples and thus the
+\acrshort*{CCSD} polarization is correct through fourth order in the
 fluctuation potential.
-Moreover, since the second order amplitudes and multipliers are
-identical up to the triples one has $\braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}}| \mathrm{HF}} = \braket{\tbar{1}^{[2]} | \s | \mathrm{HF} }$
-and thus:
-\begin{equation}\label{eq:2nd-order-pol}
-  \V\p^{[2]} +
-  2\braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}}| \mathrm{HF}}
-  + \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
-  + \braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[1]}} | \mathrm{HF}}
-  = 0
-\end{equation}
 
@@
 including $p$-th order and by multipliers up to and including $q$-th order.
-Green boxes will be put around terms that are formally equivalent to the
-vacuum expression for the energy corrections. Red boxes will appear
-around quantum/classical coupling terms that involve the triples
-manifold.
+Red boxes will appear around quantum/classical coupling terms that
+involve the triples manifold.
 
@@
   G^{[2]} &=
-  \highlight{green}{
   \braket{\mathrm{HF} | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
-  }
   + \frac{1}{2}{}^t\p^{[1]}\V\p^{[1]} +
   {}^t\p^{[1]}\s_{\mathrm{N}}^{[1]}(\tamp{}^{[1]}, \tbar{}^{[0]}) \\
-  &= \highlight{green}{E^{[2]}}
+  &= E^{[2]}
   +
@@
   +
-  {}^t\p^{[1]}\braket{\mathrm{HF} | \commutator{\s}{T^{[1]}_1} | \mathrm{HF}} \\
-  &= \left/
-  \V\p^{[1]} + 2\braket{\mathrm{HF} | \commutator{\s}{T_1^{[1]}}| \mathrm{HF}} = 0
-  \right/ \\
-  &= \highlight{green}{E^{[2]}}
-  \end{aligned}
-\end{equation}
-\todo[inline]{I feel like there is some important lesson/conclusion to
-draw from this.}
+  {}^t\p^{[1]}\braket{\mathrm{HF} | \commutator{\s}{T^{[1]}_1} | \mathrm{HF}}
+  = E^{[2]}
+  \end{aligned}
+\end{equation}
+The polarization terms were eliminated by virtue of Eq.
+\eqref{eq:1st-order-pol}. Formally, $G^{[2]}$ has the same expression as
+\emph{in vacuo}. However, the first-order singles amplitudes are nonzero
+and are determined by the quantum/classical polarizable coupling term.
 
@@
     G^{[3]} &=
-    \highlight{green}{\sum_j\braket{\tbar{j}^{[1]} | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}}
+    \sum_j\braket{\tbar{j}^{[1]} | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
     + {}^t\p^{[1]}\s_{\mathrm{N}}^{[2]}(\tamp{}^{[1]}, \tbar{}^{[1]}) \\
     &=
-    \highlight{green}{E^{[3]}}
+    E^{[3]}
     + {}^t\p^{[1]}
-    \lbrace
+    \left\lbrace
       \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
     + \braket{\tbar{2}^{[1]} | \commutator{\s}{T^{[1]}_2} | \mathrm{HF}}
-    \rbrace
+    \right\rbrace
   \end{aligned}
@@
     G^{[4]} &=
-    \highlight{green}
-    {
     \sum_j\braket{\tbar{j}^{[1]} | \commutator{\Phi}{T^{[2]}} | \mathrm{HF}}
@@
     \frac{1}{2}\commutator{\commutator{\Phi}{T^{[1]}}}{T^{[1]}} | \mathrm{HF}}
-    }
     \\
@@
     + {}^t\p^{[1]}\s_{\mathrm{N}}^{[3]}(\tamp{}^{[2]}, \tbar{}^{[1]}) \\
-    &= \highlight{green}{E^{[4]}} + \frac{1}{2}{}^t\p^{[2]}\V\p^{[2]}
-    +{}^t\p^{[2]}
-    \lbrace
+    &= E^{[4]}
+    + \frac{1}{2}{}^t\p^{[2]}\V\p^{[2]} \\
+    &+{}^t\p^{[2]}
+    \left\lbrace
     \braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}} | \mathrm{HF}}
     + \sum_i\braket{\tbar{i}^{[1]}| \commutator{\s}{T^{[1]}} | \mathrm{HF} }
-    \rbrace \\
+    \right\rbrace \\
     &+{}^t\p^{[1]}
-    \lbrace
+    \left\lbrace
     \sum_i\braket{\tbar{i}^{[1]} | \commutator{\s}{T^{[2]}} | \mathrm{HF}}
     +
-    \sum_i\frac{1}{2}\braket{\tbar{i}^{[1]} |
+    \sum_i\braket{\tbar{i}^{[1]} |\frac{1}{2}
     \commutator{\commutator{\s}{T^{[1]}}}{T^{[1]}}
@@
     \mathrm{HF}}
-    \rbrace \\
-  %%%% Steps
-  &= \left/ \text{Remove null terms} \right/ \\
-  &= \highlight{green}{E^{[4]}} \\
-    &+ \frac{1}{2}{}^t\p^{[2]}\V\p^{[2]}
-    +{}^t\p^{[2]}
-    \lbrace
-    \braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}} | \mathrm{HF}}
-    + \braket{\tbar{1}^{[1]}| \commutator{\s}{T^{[1]}} | \mathrm{HF} }
-    + \braket{\tbar{2}^{[1]}| \commutator{\s}{T_2^{[1]}} | \mathrm{HF} }
-    \rbrace \\
+    \right\rbrace \\
+  &= E^{[4]}
+    + \frac{1}{2}{}^t\p^{[2]}\V\p^{[2]}
+    + {}^t\p^{[2]}
+      \braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}} | \mathrm{HF}} \\
+    &+ {}^t\p^{[2]}
+      \braket{\tbar{1}^{[1]}| \commutator{\s}{T^{[1]}} | \mathrm{HF} }
+    + {}^t\p^{[2]}
+       \braket{\tbar{2}^{[1]}| \commutator{\s}{T_2^{[1]}} | \mathrm{HF} } \\
     &+
     {}^t\p^{[1]}
-    \lbrace
     \braket{\tbar{1}^{[1]} | \commutator{\s}{T_1^{[2]} + T_2^{[2]}} | \mathrm{HF}}
     +
+    {}^t\p^{[1]}
     \braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[2]} + T_3^{[2]}} | \mathrm{HF}}
-    \rbrace \\
+    \\
     &+
     {}^t\p^{[1]}
-    \lbrace
-    \frac{1}{2}\braket{\tbar{1}^{[1]} |
-    \commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[1]}}
+    \braket{\tbar{1}^{[1]} |
+    \frac{1}{2}\commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[1]}}
     |
@@
     +
+    {}^t\p^{[1]}
     \braket{\tbar{2}^{[1]} |
@@
     \mathrm{HF}}
-    \rbrace \\
-  %%%% Steps
-  &= \left/
-  \V\p^{[2]} +
-  2\braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}}| \mathrm{HF}}
-  + \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
-  + \braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[1]}} | \mathrm{HF}}
-  = 0
-  \right/ \\
-  &= \highlight{green}{E^{[4]}}
-    + \frac{1}{2}{}^t\p^{[2]}\braket{\tbar{1}^{[1]}| \commutator{\s}{T^{[1]}} | \mathrm{HF} }
+  \end{aligned}
+\end{equation}
+Employing Eq. \eqref{eq:2nd-order-pol}, the terms involving the
+second order polarization can be simplified further:
+\begin{equation}
+  \begin{aligned}
+  G^{[4]}
+  &= E^{[4]}
+    + \frac{1}{2}{}^t\p^{[2]}
+      \braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}} | \mathrm{HF}}
+    - \frac{1}{2}{}^t\p^{[2]}
+       \braket{\tbar{1}^{[2]}| \s | \mathrm{HF}} \\
+    &+ \frac{1}{2}{}^t\p^{[2]}
+      \braket{\tbar{1}^{[1]}| \commutator{\s}{T^{[1]}} | \mathrm{HF} }
+    + \frac{1}{2}{}^t\p^{[2]}
+       \braket{\tbar{2}^{[1]}| \commutator{\s}{T_2^{[1]}} | \mathrm{HF} } \\
+    &+ \frac{1}{2}{}^t\p^{[2]}\braket{\tbar{1}^{[1]}| \commutator{\s}{T^{[1]}} | \mathrm{HF} }
     + \frac{1}{2}{}^t\p^{[2]}\braket{\tbar{2}^{[1]}| \commutator{\s}{T_2^{[1]}} | \mathrm{HF} } \\
@@
     +
-    \highlight{red}{
+    \highlight{PMS1797}{
     {}^t\p^{[1]}
@@
     {}^t\p^{[1]}
-    \lbrace
-    \frac{1}{2}\braket{\tbar{1}^{[1]} |
+    \braket{\tbar{1}^{[1]} |
+    \frac{1}{2}
     \commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[1]}}
@@
     \mathrm{HF}}
-    \rbrace \\
-  \end{aligned}
-\end{equation}
-\todo[inline]{There is one term involving second order triples and first
-order polarization}
+  \end{aligned}
+\end{equation}
+Notice that the term $\frac{1}{2}{}^t\p^{[2]}\braket{\tbar{1}^{[2]}| \s | \mathrm{HF}}$
+has not been included, since it would violate the $2n+2$ rule for the
+multipliers. Moreover, there is only one term that involves both the
+triples amplitudes and the polarization.
 
@@
     G^{[5]} &=
-    \highlight{green}
-    {
     \frac{1}{2}\braket{\mathrm{HF} |
@@
     | \mathrm{HF}}
-    }
     \\
-    &+ \highlight{green}{\sum_j \braket{\tbar{j}^{[2]} |
+    &+ \sum_j \braket{\tbar{j}^{[2]} |
       \commutator{\Phi}{T^{[2]}}
@@
     \commutator{\commutator{\Phi}{T^{[1]}}}{T^{[1]}}
-    | \mathrm{HF}}
-    }
-    + {}^t\p^{[2]}\s_{\mathrm{N}}^{[3]}(\tamp{}^{[2]}, \tbar{}^{[2]})
+    | \mathrm{HF}} \\
+    &+ {}^t\p^{[2]}\s_{\mathrm{N}}^{[3]}(\tamp{}^{[2]}, \tbar{}^{[2]})
     + {}^t\p^{[1]}\s_{\mathrm{N}}^{[4]}(\tamp{}^{[2]}, \tbar{}^{[2]}) \\
-    &= \highlight{green}{E^{[5]}}
+    &= E^{[5]}
     + {}^t\p^{[2]}
-    \lbrace
-    \sum_i\braket{\tbar{i}^{[1]} | \commutator{\s}{T^{[2]}} | \mathrm{HF}}
-    +
+    \sum_i\braket{\tbar{i}^{[1]} | \commutator{\s}{T^{[2]}} | \mathrm{HF}} \\
+    &+ {}^t\p^{[2]}
     \sum_i\braket{\tbar{i}^{[1]} |
@@
     | \mathrm{HF}}
-    +
+    + {}^t\p^{[2]}
     \sum_i\braket{\tbar{i}^{[2]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
-    \rbrace \\
+    \\
     &+ {}^t\p^{[1]}
-    \lbrace
     \sum_i\braket{\tbar{i}^{[2]} | \commutator{\s}{T^{[2]}} | \mathrm{HF}}
-    +
+    + {}^t\p^{[1]}
     \sum_i\braket{\tbar{i}^{[2]} |
     \frac{1}{2}\commutator{\commutator{\s}{T^{[1]}}}{T^{[1]}}
-    | \mathrm{HF}}
-    +
+    | \mathrm{HF}} \\
+    &+ {}^t\p^{[1]}
     \sum_i\braket{\tbar{i}^{[1]} |
@@
     | \mathrm{HF}}
-    \rbrace \\
-  %%%% Steps
-  &= \left/ \text{Remove null terms} \right/ \\
-  &= \highlight{green}{E^{[5]}}
-    + {}^t\p^{[2]}\braket{\tbar{1}^{[1]} | \commutator{\s}{T_1^{[2]} + T_2^{[2]}} | \mathrm{HF}}
-    + \highlight{red}{
+    \\
+  &= E^{[5]}
+    + {}^t\p^{[2]}\braket{\tbar{1}^{[1]} | \commutator{\s}{T_1^{[2]} +
+    T_2^{[2]}} | \mathrm{HF}} \\
+    &+ \highlight{PMS1797}{
     {}^t\p^{[2]}\braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[2]} + T_3^{[2]}} | \mathrm{HF}}
@@
     \frac{1}{2}\commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[1]}}
-    | \mathrm{HF}}
-    + {}^t\p^{[2]}
+    | \mathrm{HF}} \\
+    &+ {}^t\p^{[2]}
     \braket{\tbar{2}^{[1]} |
@@
     | \mathrm{HF}}
-    \\
-    &+
+    +
     {}^t\p^{[1]}
-    \braket{\tbar{1}^{[2]} | \commutator{\s}{T_1^{[2]} + T_2^{[2]}} | \mathrm{HF}}
-    +
-    \highlight{red}{
+    \braket{\tbar{1}^{[2]} | \commutator{\s}{T_1^{[2]} + T_2^{[2]}} |
+    \mathrm{HF}} \\
+    &+
+    \highlight{PMS1797}{
     {}^t\p^{[1]}
@@
     &+
-    \highlight{red}{
+    \highlight{PMS1797}{
     {}^t\p^{[1]}
@@
 \end{equation}
-
-\section[Approximate Coupled Cluster Methods]{
-Approximate Coupled Cluster Methods with Quantum/Classical Polarizable Hamiltonians
-}\label{sec:cc-approximate-quantum-classical}
-
-The \acrshort{PTED}-\acrshort{CCSD} scheme for the quantum/classical coupling is
+Similarly to the exposition in Chapter \ref{ch:QM}, these expressions
+will form the basis to the development of methods for the approximate
+beyond \acrshort*{CCSD} including connected triples.
+
+The \acrshort{PTED}-\acrshort{CCSD} scheme for the quantum/classical coupling
 only includes singles and doubles in its excitation manifold and is thus
@@
     \label{eq:tbar-star} \\
-    \p^{*} &= \p^{[1]} + \p^{[2]} + \p^{[3]} + O(4)
+    \p^{*} &= \p^{[1]} + \p^{[2]} + O(3)
     \label{eq:p-star}
@@
 \end{subequations}
-We will denote the \acrshort{CCSD} converged parameters with a $*$ superscript.
-\todo[inline]{Check the statement above about $\p^{*}$.}
+Here and in the following, the \acrshort{CCSD} converged parameters will
+be denoted with a $*$ superscript.
 
@@
 model going beyond third order must then take their effect into account.
-We first select the fourth and fifth order contributions to the energy
-involving connected triples.
+The fourth and fifth order contributions to the energy due to connected
+triples given in Eqs.
+\eqref{eq:4th-order-triples} and \eqref{eq:5th-order-triples} will be
+augmented by additional terms due to the nonzero first order singles and
+the quantum/classical polarizable coupling.
+To fourth order these are:
 \begin{equation}\label{eq:triples-fourth-order}
-  G_T^{[4]} =
-  \highlight{red}{\braket{\tbar{1}^{[1]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}}
-  + \braket{\tbar{2}^{[1]}| \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
-  + {}^t\p^{[1]}\braket{\tbar{2}^{[1]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
-\end{equation}
+  \begin{aligned}
+  G_T^{[4]} &=
+  \highlight{PMS130}{\braket{\tbar{1}^{[1]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}}
+  + \braket{\tbar{2}^{[1]}| \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}} \\
+  &+ {}^t\p^{[1]}\braket{\tbar{2}^{[1]} | \commutator{\s}{T_3^{[2]}} |
+  \mathrm{HF}},
+  \end{aligned}
+\end{equation}
+while the fifth order contributions are:
 \begin{equation}\label{eq:triples-fifth-order}
@@
   G_T^{[5]} &=
-    \highlight{red}{
+    \highlight{PMS130}{
     \braket{\tbar{2}^{[1]} |
@@
     | \mathrm{HF}}
-    } \\
-    &+
-    \braket{\tbar{1}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
-  + \braket{\tbar{2}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
-  + \braket{\tbar{3}^{[2]} | \commutator{\Phi}{T_2^{[2]}} | \mathrm{HF}}
+    }
+    +
+    \braket{\tbar{1}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}} \\
+  &+ \braket{\tbar{2}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}} \\
+  &+ \braket{\tbar{3}^{[2]} | \commutator{\Phi}{T_2^{[2]}} | \mathrm{HF}}
   + \braket{\tbar{3}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
@@
   +
-  \highlight{red}{
+  \highlight{PMS130}{
     \braket{\tbar{3}^{[2]} |
@@
 \end{equation}
-Terms in red boxes are vacuum-like terms that appear due to nonzero
-first order singles amplitudes and multipliers.
-
-\todo[inline]{WARNING. I am not 100\% sure that the multipliers
-equations are complex conjugate to the amplitudes equations also in
-second order!}
-
-\subsection{The \texorpdfstring{$\Lambda\text{CCSD}[\text{T}]$}{CCSD[aT]} and \texorpdfstring{$\text{CCSD}[\text{T}]$}{CCSD[T]} models}
-
-\todo[inline]{References for the a-CCSD[T] and CCSD[T] models!!!}
-
-The [T] noniterative triples corrections are formed by adding the fourth
-order \acrshort{CC}\acrshort{PT} terms arising from connected triples on top of
-the \acrshort{CCSD} energy. The \emph{converged} \acrshort{CCSD} amplitudes,
-multipliers and polarization are however used instead of the
-perturbative quantities in Eq.~\eqref{eq:triples-fourth-order}.
-The correction term is thus improved, since the \acrshort{CCSD} quantities
-are determined to infinite order, within the singles and doubles space.
-The asymmetric $\Lambda\text{CCSD}[\text{T}]$ correction is then
-obtained as:
-\begin{equation}\label{eq:4-lambda-bracket-t}
-  G_{\Lambda[\text{T}]} =
-  {}^*G_{\Lambda[\text{T}]}^{[4]} =
-  \highlight{red}{\braket{\tbar{1}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}}
-  + \braket{\tbar{2}^{*}| \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
-  + {}^t\p^{*}\braket{\tbar{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
-\end{equation}
-Replacing the perturbation multipliers by converged \acrshort{CCSD}
-amplitudes leads to the symmetric CCSD[T] model:
-\begin{equation}\label{eq:4-bracket-t}
-  G_{[\text{T}]} =
-  {}^*G_{[\text{T}]}^{[4]} =
-  \highlight{red}{\braket{\tamp{1}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}}
-  + \braket{\tamp{2}^{*}| \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
+We have highlighted the vacuum-like terms that appear due to the nonzero
+first order singles amplitudes and multipliers in yellow.
+
+\section[Approximate Coupled Cluster Methods]{
+Approximate Coupled Cluster Methods with Quantum/Classical Polarizable Hamiltonians
+}\label{sec:cc-approximate-quantum-classical}
+
+\todo[inline]{Finish up on CC2 and CC3.}
+In complete analogy to the discussion in Section
+\ref{sec:cc-approximate}, we can derive iterative and noniterative
+schemes for the approximate inclusion of higher order excitations in the
+\acrshort*{CC} hierarchy.
+Starting from the \acrshort{CC2} and \acrshort{CC3} Lagrangians, Eqs.
+\eqref{eq:cc2-lagrangian} and \eqref{eq:cc3-lagrangian} respectively,
+we can write down the corresponding effective Lagrangians. The correct
+source term $\s_\mathrm{N}(\tamp{},\tbar{})$ must however be identified.
+For the \acrshort*{CC2} model, we start from its \acrshort*{CCSD}
+expression, Eq. \eqref{eq:source-ccsd}, and retain only terms up to first order:
+\begin{equation}
+  \begin{aligned}
+  \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCSD} &=
+  \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCS}
+  + \braket{\tbar{1} | \BCHfirst{\s}{T_2} | \mathrm{HF}} \\
+  &+ \braket{\tbar{2} | \BCHfirst{\s}{T_2} | \mathrm{HF}}
+  + \braket{\tbar{2} | \commutator{\commutator{\s}{T_1}}{T_2} | \mathrm{HF}}
+  \end{aligned}
+\end{equation}
+where the $T_1$-transformed operator was used.
+
+We can construct the \emph{noniterative}
+$\Lambda\text{CCSD}(\text{T})$\autocite{Kucharski1998-qq,
+Kucharski1998-oi, Crawford1998-vj} and
+$\text{CCSD}(\text{T})$\autocite{Raghavachari1989-bn} methods as shown
+in Chapter \ref{ch:QM}.
+We propose to include all the fourth order terms in Eq.
+\eqref{eq:triples-fourth-order} and the following two terms from the
+fifth order connected triples corrections in Eq.
+\eqref{eq:triples-fifth-order}:
+\begin{alignat}{2}
+ \braket{\tbar{1}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}},
+    \quad&
+ {}^t\p^{[1]}\braket{\tbar{2}^{[2]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
+\end{alignat}
+Inserting the converged \acrshort{CCSD} parameters yields the following
+expressions:
+\begin{subequations}\label{eq:g-parentheses-t}
+ \begin{align}
+  G_{\Lambda(\text{T})}
+  &=
+    \sum_{j}\braket{\tbar{j}^{*}| \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
+    + {}^t\p^{*}\braket{\tbar{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}} \\
+  G_{(\text{T})}
+ &=
+    \sum_{j}\braket{\tamp{j}^{*}|\commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
     + {}^t\p^{*}\braket{\tamp{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
-\end{equation}
-
-\subsection{The
-\texorpdfstring{$\Lambda\text{CCSD}(\text{T})$}{CCSD(aT)} and CCSD(T) models}
-
-\todo[inline]{References for the a-CCSD(T) and CCSD(T) models!!!}
-
-The (T) noniterative triples corrections are formed by adding some fifth
-order terms to the fourth order [T] correction given above.
-Also for the (T) corrections, the \emph{converged} \acrshort{CCSD} amplitudes, multipliers and polarization
-are used instead of the perturbative quantities.
-The fifth order terms involving projection of
-$\commutator{\Phi}{T_3^{*}}$ on the singles and doubles manifold are
-chosen to form the additional perturbative correction.
-
-The correction term is thus improved, since the \acrshort{CCSD} quantities
-are determined to infinite order, within the singles and doubles space.
-The asymmetric $\Lambda\text{CCSD}(\text{T})$ correction is then
-obtained as:
-\begin{equation}\label{eq:5-lambda-parens-t}
- \begin{aligned}
-  G_{\Lambda(\text{T})} =
-  {}^*G_{\Lambda[\text{T}]}^{[4]} +
-  {}^*G_{\Lambda(\text{T})}^{[5]} =
-  \sum_{j=1}^2
-    \braket{\tbar{j}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
-  + {}^t\p^{*}\braket{\tbar{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
- \end{aligned}
-\end{equation}
-Replacing the perturbation multipliers by converged \acrshort{CCSD}
-amplitudes leads to the symmetric CCSD(T) model:
-\begin{equation}\label{eq:5-parens-t}
- \begin{aligned}
-  G_{(\text{T})} =
-  {}^*G_{[\text{T}]}^{[4]} +
-  {}^*G_{(\text{T})}^{[5]} =
-  \sum_{j=1}^2
-    \braket{\tamp{j}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
-  + {}^t\p^{*}\braket{\tamp{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
- \end{aligned}
-\end{equation}
-
-\subsection{MBPT Terms in the
-\texorpdfstring{$\Lambda\text{CCSD}(\text{T})$}{CCSD(aT)} and CCSD(T) models}
-
-Triples excitations enter the \acrshort{MBPT} equations for the singles and
-doubles amplitudes and multipliers at third order. Thus the \acrshort{CCSD}
-amplitudes and multipliers are correct through second order in the
-fluctuation potential, while the polarization is correct through third
-order. Moreover, amplitudes and multipliers are equal to one another.
-
-We can rework the energy correction given above so that the perturbative
-counterparts clearly appear.
-First of all, we notice that $T_3^{*} = T_3^{[2]} + \tilde{T}_3$ contains higher than second order
-contributions where $\tilde{T}_3$ contains third and higher order contributions to the
-connected triples. In particular, the following holds:
+ \end{align}
+\end{subequations}
+We recall that both the left and right state are used in the
+$\Lambda\text{CCSD}(\text{T})$ method, while $\text{CCSD}(\text{T})$
+uses the right state only, replacing the multipliers with the
+amplitudes.
+Further insight into the proposed correction can be gained by
+inserting the \acrshort{MBPT} expansion of the converged \acrshort{CCSD}
+parameters:\autocite{Stanton1997-lx, Koch1997-nm,
+Eriksen2014-gy, Eriksen2015-il, Kristensen2016-od}
+First of all, we notice that $T_3^{*} = T_3^{[2]} + \tilde{T}_3 + O(3)$
+contains higher than second order contributions. In particular, $\tilde{T}_3$
+is a third order contributions to the connected triples arising from
+second order connected doubles:
 \begin{equation}
@@
 \end{equation}
-such that:
+such that the third order $\tilde{T}_3$ can re rewritten as:
 \begin{equation}
   \tilde{T}_3
-  = \sum_{\mu_3}\tilde{t}_{\mu_3}\cluster{\mu_3}
   = \sum_{\mu_3}
   \left(-\denom{\mu_3}^{-1}\braket{\mu_3 |
-  \commutator{\Phi}{T_3^{[2]}}
+  \commutator{\Phi}{T_2^{[2]}}
   | \mathrm{HF}}\right)
@@
 
-Using Eqs.~\eqref{eq:tamp-star}--\eqref{eq:p-star} one can expand the
+Using Eqs. \eqref{eq:tamp-star}--\eqref{eq:p-star} one can expand the
 energy correction as follows:
@@
   G_{\Lambda(\text{T})} &=
-  \sum_{j=1}^2 \braket{\tbar{j}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
-  + {}^t\p^{*}\braket{\tbar{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}} \\
-  &= \left/ \text{Insert perturbative expansions of parameters} \right/ \\
-  &=
-  \sum_{j=1}^2 \braket{\tbar{j}^{[1]}+\tbar{j}^{[2]} |
-  \commutator{\Phi}{T_3^{[2]} + \tilde{T}_3} | \mathrm{HF}} \\
-  &+ ({}^t\p^{[1]} + {}^t\p^{[2]})
-  \braket{\tbar{2}^{[1]} + \tbar{2}^{[2]} | \commutator{\s}{T_3^{[2]} + \tilde{T}_3} | \mathrm{HF}} \\
-  &= \left/ \text{Expand} \right/ \\
-  &\highlight{green}{
   \sum_{j=1}^2 \braket{\tbar{j}^{[1]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
-  }
-  + \highlight{blue}{
-  \sum_{j=1}^2 \braket{\tbar{j}^{[1]} | \commutator{\Phi}{\tilde{T}_3} | \mathrm{HF}}
-  }
-  + \highlight{blue}{
+  +
+  \braket{\tbar{3}^{[2]} | \commutator{\Phi}{T_2^{[2]}} | \mathrm{HF}}
+  +
   \sum_{j=1}^2 \braket{\tbar{j}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
-  } \\
-  &+
-  \highlight{green}{
+  \\
+  &+
   {}^t\p^{[1]} \braket{\tbar{2}^{[1]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
-  }
-  + \highlight{blue}{
+  +
   {}^t\p^{[1]} \braket{\tbar{2}^{[2]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
-  } \\
-  &+ \highlight{blue}{
+   \\
+  &+
   {}^t\p^{[1]} \braket{\tbar{2}^{[1]} | \commutator{\s}{\tilde{T}_3} | \mathrm{HF}}
-  }
-  + \highlight{blue}{
+  +
   {}^t\p^{[2]} \braket{\tbar{2}^{[1]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
-  }
-  + O(6) \\
-  &= \left/ \text{Use definition of $\tilde{T}_3$ and triples multipliers equation} \right/ \\
-  &\highlight{green}{
-  \sum_{j=1}^2 \braket{\tbar{j}^{[1]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
-  }
-  + \highlight{blue}{
-  \braket{\tbar{3}^{[2]} | \commutator{\Phi}{T_2^{[2]}} | \mathrm{HF}}
-  }
-  + \highlight{blue}{
-  \sum_{j=1}^2 \braket{\tbar{j}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
-  } \\
-  &+
-  \highlight{green}{
-  {}^t\p^{[1]} \braket{\tbar{2}^{[1]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
-  }
-  + \highlight{blue}{
-  {}^t\p^{[1]} \braket{\tbar{2}^{[2]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
-  } \\
-  &+ \highlight{blue}{
-  {}^t\p^{[1]} \braket{\tbar{2}^{[1]} | \commutator{\s}{\tilde{T}_3} | \mathrm{HF}}
-  }
-  + \highlight{blue}{
-  {}^t\p^{[2]} \braket{\tbar{2}^{[1]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
-  }
-  + O(6) \\
-  \end{aligned}
-\end{equation}
-
-\todo[inline]{Don't really know what to do with the
-${}^t\p^{[1]} \braket{\tbar{2}^{[1]} | \commutator{\s}{\tilde{T}_3} | \mathrm{HF}}$ term.
-I think it is correctly identified as $O(5)$, but that it's among the
-discarded terms.}
+  + O(6)
+  \end{aligned}
+\end{equation}

### 2016-09-13 Chapters/QuantumMechanics.tex
First draft sent to L.F. and B.M.
Multipliers equations for CC2 and CC3 added

@@
 \chapter{A Quick Tour of Molecular Electronic Structure Theory}\label{ch:QM}
+
+\begin{epigraphs}
+\setlength\epigraphwidth{7.5cm}
+\qitem{
+plays with the bigness of his littleness
+
+--- electrons deify one razorblade
+
+into a mountainrange; lenses extend
+
+unwish through curving wherewhen till unwish
+
+returns on its unself.
+    }{
+    --- \textsc{E. E. Cummings}}
+\setlength\epigraphwidth{7cm}
+\end{epigraphs}
 
@@
 \begin{equation}
-  \overline{F} = F + \sum_{i}^4\sum_{\mu_i}\tamp{\mu_i}\cluster{\mu_i}\denom{\mu_i}
+  \overline{F} = F + \sum_{u=1}^\mathcal{M}\sum_{\mu_u}\denom{\mu_u}\tamp{\mu_u}\cluster{\mu_u}
 \end{equation}
-where $\denom{\mu_i}$ is the difference in orbital energies between the
-virtual and occupied spin-orbitals of excitation $i$.
+where $\denom{\mu_u}$ is the difference in orbital energies between the
+virtual and occupied spin-orbitals of excitation $u$.
 For the fluctuation potential the similarity transformation truncates
@@
   \begin{aligned}
-    \mathcal{L}^\mathcal{M}(\tamp{}, \tbar{})
+    \mathcal{L}(\tamp{}, \tbar{})_\mathcal{M}
     &=
     \braket{\mathrm{HF} | \overline{H} | \mathrm{HF}}
-    + \sum_{i=1}^{\mathcal{M}}\sum_{\mu_i}
-    \tbar{\mu_i}\braket{\mu_i | \overline{H} | \mathrm{HF}} \\
+    + \sum_{u=1}^{\mathcal{M}}\sum_{\mu_u}
+    \tbar{\mu_u}\braket{\mu_u | \overline{H} | \mathrm{HF}} \\
     &=
   E_0
-  + \sum_{i=1}^{\mathcal{M}}\tbar{\mu_i}\denom{\mu_i}\tamp{\mu_i}
+  + \sum_{u=1}^{\mathcal{M}}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
   + \braket{\mathrm{HF} | \overline{\Phi} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\Phi} | \mathrm{HF}}
+  + \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \overline{\Phi} | \mathrm{HF}}
   \end{aligned}
@@
 \begin{equation}
-  \bra{\tbar{j}} = \sum_{\mu_j}\tbar{\mu_j}\bra{\mu_j}.
+  \bra{\tbar{u}} = \sum_{\mu_u}\tbar{\mu_u}\bra{\mu_u}.
 \end{equation}
@@
   \begin{align}
-   \tampEq{\mu_i}(\tamp{}, \tbar{}, \p)  &=
-   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
+   \tampEq{\mu_q}(\tamp{}, \tbar{}, \p)  &=
+   \denom{\mu_q}\tamp{\mu_q} + \braket{\mu_q | \overline{\Phi} | \mathrm{HF}}
              = 0 \label{eq:cc-amplitudes}\\
-   \tbarEq{\mu_i}(\tamp{}, \tbar{}, \p)
+   \tbarEq{\mu_q}(\tamp{}, \tbar{}, \p)
     &=
-    \denom{\mu_i}\tbar{\mu_i} +
-    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
-    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} |
-    \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}}
+    \denom{\mu_q}\tbar{\mu_q} +
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}} +
+    \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
+    \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}}
              = 0 \label{eq:cc-multipliers}
@@
 \begin{equation}
-  O(\tamp{}, \tbar{}) = \braket{\mathrm{HF} | \overline{O} | \mathrm{HF} }
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{O} | \mathrm{HF} }
+  O(\tamp{}, \tbar{})_\mathcal{M} = \braket{\mathrm{HF} | \overline{O} | \mathrm{HF} }
+  + \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \overline{O} | \mathrm{HF} }
 \end{equation}
-When required by the context, we will explicitly write at which
-truncation level in the cluster operator the expectation was evaluated:
-$O(\tamp{}, \tbar{})_{\mathrm{CC}\mathcal{M}}$.
 
@@
 
-It is clear from the stationarity conditions that $\tamp{\mu_i}^{[0]} = 0, \,\forall i$ and
-$\tbar{\mu_i}^{[0]} = 0, \, \forall i$.
+It is clear from the stationarity conditions that $\tamp{\mu_u}^{[0]} = 0, \,\forall u$ and
+$\tbar{\mu_u}^{[0]} = 0, \, \forall u$.
 Thus cluster operators can be of order 1 and higher. We introduce the
@@
 \begin{alignat}{2}
-  T_p^{[i]} = \sum_{\mu_p}\tamp{\mu_p}^{[i]}\cluster{\mu_p},
+  T_u^{[i]} = \sum_{\mu_u}\tamp{\mu_u}^{[i]}\cluster{\mu_u},
   \quad&
-  \bra{\tbar{p}^{[i]}} = \sum_{\mu_p}\tbar{\mu_p}^{[i]}\bra{\mu_p}
+  \bra{\tbar{u}^{[i]}} = \sum_{\mu_u}\tbar{\mu_u}^{[i]}\bra{\mu_u}
 \end{alignat}
@@
 \begin{equation}\label{eq:1st-order}
-  \begin{aligned}
   \denom{\mu_2}\tamp{\mu_2}^{[1]}
-  &= - \braket{\mu_2 | \Phi | \mathrm{HF}}
+  = - \braket{\mu_2 | \Phi | \mathrm{HF}}
   = - \braket{\mathrm{HF} | \commutator{\Phi}{\cluster{\mu_2}} |
-  \mathrm{HF}} \\
-  &= \denom{\mu_2}\tbar{\mu_2}^{[1]}
-  \end{aligned}
+  \mathrm{HF}}
+  = \denom{\mu_2}\tbar{\mu_2}^{[1]}
 \end{equation}
@@
 amplitudes are of second and higher order.
-To second order, one obtains:
+For second order singles, doubles and triples one obtains:
 \begin{equation}\label{eq:2nd-order}
-  \begin{aligned}
-  \denom{\mu_i}\tamp{\mu_i}^{[2]}
-  &= - \braket{\mu_i | \commutator{\Phi}{T_2^{[1]}} | \mathrm{HF}}
-  = - \braket{\tbar{2}^{[1]} | \commutator{\Phi}{\cluster{\mu_i}} |
-  \mathrm{HF}} \\
-  &= \denom{\mu_i}\tbar{\mu_i}^{[2]},\quad i = 1, 2, 3.
-  \end{aligned}
+  \denom{\mu_q}\tamp{\mu_q}^{[2]}
+  = - \braket{\mu_q | \commutator{\Phi}{T_2^{[1]}} | \mathrm{HF}}
+  = - \braket{\tbar{2}^{[1]} | \commutator{\Phi}{\cluster{\mu_q}} |
+  \mathrm{HF}}
+  = \denom{\mu_q}\tbar{\mu_q}^{[2]}
 \end{equation}
@@
 In the \gls{CC2} method, the amplitudes equations are:\autocite{Christiansen1995-vy, Helgaker2000-tz}
-\begin{subequations}
+\begin{subequations}\label{eq:cc2-amplitudes}
   \begin{align}
@@
   \begin{aligned}
-  \mathcal{L}_\mathrm{CC2}(\tamp{}, \tbar{})
+  \mathcal{L}(\tamp{}, \tbar{})_\mathrm{CC2}
   &=
   E_0
-  + \sum_{i=1}^{\mathcal{2}}\tbar{\mu_i}\denom{\mu_i}\tamp{\mu_i}
+  + \sum_{u=1}^{\mathcal{2}}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
   + \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{Q_2} | \mathrm{HF}} \\
-  &+ \sum_{\mu_1}\tbar{\mu_1}\braket{\mu_1 |
+  &+ \braket{\tbar{1} |
   \check{\Phi} + \commutator{\check{\Phi}}{Q_2}
-  | \mathrm{HF}} \\
-  &+ \sum_{\mu_2}\tbar{\mu_2}\braket{\mu_2 | \check{\Phi} | \mathrm{HF}}
+  | \mathrm{HF}}
+  + \braket{\tbar{2} | \check{\Phi} | \mathrm{HF}}
   \end{aligned}
@@
 method down to $N^5$ in the iterative step.
+From the approximate \acrshort*{CC2} Lagrangian we can derive the
+multipliers equations:
+\begin{subequations}\label{eq:cc2-multipliers}
+  \begin{align}
+    \denom{\mu_1}\tbar{\mu_1} &+
+     \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}}
+   + \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_1}}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{Q_2}
+     | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{2} | \commutator{\check{\Phi}}{\cluster{\mu_1}} |
+   \mathrm{HF}} = 0 \\
+    \denom{\mu_2}\tbar{\mu_2} &+
+    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}}
+   + \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_2}}
+     | \mathrm{HF}}
+    = 0
+  \end{align}
+\end{subequations}
 
-The \gls{CC3} model is constructed analogously, the amplitudes
-equations:\autocite{Christiansen1995-pv, Koch1997-nm, Helgaker2000-tz,
+The \gls{CC3} model is constructed analogously. Starting from the
+\acrshort*{CCSDT} amplitudes equations in the $T_1$-transformed
+representation, we retain only second order terms in the triples
+equation:\autocite{Christiansen1995-pv, Koch1997-nm, Helgaker2000-tz,
 Smith2005-an}
-\begin{subequations}
+\begin{subequations}\label{eq:cc3-amplitudes}
   \begin{align}
@@
 \end{subequations}
-are obtained from the \acrshort*{CCSDT} equations by retaining only
-second order terms in the triples equation.
-The approximate triples cluster operator was introduced:
+In complete analogy with \acrshort*{CC2}, the quasi-$T_3$ operator was introduced:
 \begin{equation}
   Q_3 = - \sum_{\mu_3}
-  \denom{\mu_3}^{-1}\braket{\mu_3 | \commutator{\check{\Phi}}{T_2} | \mathrm{HF}}\cluster{\mu_3}
+  \denom{\mu_3}^{-1}\braket{\mu_3 | \commutator{\check{\Phi}}{T_2} | \mathrm{HF}}\cluster{\mu_3}.
 \end{equation}
-and we shall refer to it as quasi-$T_3$.
 The \acrshort*{CC3} Lagrangian is constructed based on the approximate
-equations obtained for the singles, doubles and
-triples:\autocite{Christiansen1995-pv, Koch1997-nm, Helgaker2000-tz, Smith2005-an}
+amplitude equations
+\eqref{eq:cc3-amplitudes}:\autocite{Christiansen1995-pv, Koch1997-nm,
+Helgaker2000-tz, Smith2005-an}
 \begin{equation}\label{eq:cc3-lagrangian}
   \begin{aligned}
-  \mathcal{L}_\mathrm{CC3}(\tamp{}, \tbar{})
+  \mathcal{L}(\tamp{}, \tbar{})_\mathrm{CC3}
   &=
   E_0
-  + \sum_{i=1}^{\mathcal{3}}\tbar{\mu_i}\denom{\mu_i}\tamp{\mu_i}
+  + \sum_{u=1}^{\mathcal{3}}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
   + \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
-  &+ \sum_{\mu_1}\tbar{\mu_1}\braket{\mu_1 |
+  &+ \braket{\tbar{1} |
   \check{\Phi} + \commutator{\check{\Phi}}{T_2} + \commutator{\check{\Phi}}{Q_3}
   | \mathrm{HF}} \\
-  &+ \sum_{\mu_2}\tbar{\mu_2}\braket{\mu_2 |
+  &+ \braket{\tbar{2} |
   \check{\Phi} + \commutator{\check{\Phi}}{T_2}
@@
   | \mathrm{HF}} \\
-  &+ \sum_{\mu_3}\tbar{\mu_3}\braket{\mu_3 | \commutator{\check{\Phi}}{T_2} | \mathrm{HF}}
+  &+ \braket{\tbar{3} | \commutator{\check{\Phi}}{T_2} | \mathrm{HF}}
   \end{aligned}
@@
 method down to $N^7$ in the iterative step.
+We can now derive the \acrshort*{CC3} multipliers equations:
+\begin{subequations}\label{eq:cc3-multipliers}
+  \begin{align}
+    \denom{\mu_1}\tbar{\mu_1} &+
+     \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}}
+   + \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_1}}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+     | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{2} |
+       \commutator{\check{\Phi}}{\cluster{\mu_1}}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{Q_3}
+   | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{3} |
+      \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+     | \mathrm{HF}} = 0 \\
+    \denom{\mu_2}\tbar{\mu_2} &+
+    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}}
+   + \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_2}}
+     | \mathrm{HF}} \nonumber \\
+  &+ \braket{\tbar{2} |
+       \commutator{\check{\Phi}}{\cluster{\mu_2}}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_2}}}{T_2}
+     | \mathrm{HF}} \nonumber \\
+  &+ \braket{\tbar{3} |
+       \commutator{\check{\Phi}}{\cluster{\mu_2}}
+     | \mathrm{HF}}
+    = 0 \\
+    \denom{\mu_3}\tbar{\mu_3} &+
+    \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_3}}
+     | \mathrm{HF}}
+  + \braket{\tbar{2} |
+       \commutator{\check{\Phi}}{\cluster{\mu_3}}
+     | \mathrm{HF}}
+    = 0
+  \end{align}
+\end{subequations}
 
@@
 Eriksen2015-il}
-\begin{subequations}
+\begin{subequations}\label{eq:brackets-t}
   \begin{align}
@@
 \eqref{eq:5th-order-triples}:
-\begin{subequations}
+\begin{subequations}\label{eq:parentheses-t}
  \begin{align}

### 2016-09-14 Chapters/MathematicalResults.tex
Save the day's work

@@
 
-This appendix collects some results in the field of functional analysis and
-boundary integral equations. No proofs or examples are given, the interested
-reader is referred to the monographs by \citeauthor{Ern2004-oo},
-\citeauthor{Hsiao2008-xb} and \citeauthor{Sauter2011-an}\autocite{Ern2004-oo,
-Hsiao2008-xb, Sauter2011-an}
+For brevity's sake, some results and derivations have been omitted from the main
+body of the thesis. I collect the ones I judge most relevant in this Appendix.
+Section \ref{sec:T1-cc} presents the $T_1$-transformed form of the
+\acrshort{CCS}, \acrshort{CCSD} and \acrshort{CCSDT} Lagrangians, the corresponding amplitudes
+and multipliers equations and the one-electron operators expectation values.
 
-\section{Preliminary Definitions}
+Basic results in the fields of functional analysis and boundary integral
+equations that were omitted from Chapter \ref{ch:CSM} are presented here.
+No proofs or examples are given, the interested reader is referred to the
+monographs by \citeauthor{Ern2004-oo}, \citeauthor{Hsiao2008-xb} and
+\citeauthor{Sauter2011-an}\autocite{Ern2004-oo, Hsiao2008-xb, Sauter2011-an}
+
+\section{The \texorpdfstring{$T_1$}{T1}-Transformation}\label{sec:T1-cc}
+
+$T_1$ is a one-electron operator, carrying out a similarity transformation of
+an operator $O$ by means of its exponential mapping will preserve the particle
+rank of $O$.\autocite{Koch1994-vr}
+Starting from the arbitrary order Lagrangian \eqref{eq:cc-lagrangian},
+we want to derive their $T_1$-transformed
+expressions.
+For the \acrshort{CCS} model, this is straightforward.
+The Lagrangian is:
+\begin{equation}\label{eq:ccs-lagrangian}
+ \mathcal{L}(\tamp{}, \tbar{})_\mathrm{CCS}
+ =
+  E_0
+  + \tbar{\mu_1}\denom{\mu_1}\tamp{\mu_1}
+  + \braket{\mathrm{HF} | \check{\Phi} | \mathrm{HF}}
+  + \braket{\tbar{1} | \check{\Phi} | \mathrm{HF}}
+\end{equation}
+the governing equations:
+\begin{subequations}\label{eq:ccs-equations}
+  \begin{align}
+   \denom{\mu_1}\tamp{\mu_1} &+ \braket{\mu_1 | \check{\Phi} | \mathrm{HF}}
+             = 0 \label{eq:ccs-amplitudes}\\
+    \denom{\mu_1}\tbar{\mu_1} &+
+    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}} +
+    \braket{\tbar{1} |
+    \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}}
+             = 0 \label{eq:ccs-multipliers}
+  \end{align}
+\end{subequations}
+
+When including higher order excitations we will seek simplifications in the commutator expansions
+by employing the following result:\autocite{Helgaker2000-tz}
+\begin{lemma}[Excitation ranks manifold]\label{lemma:exc-rank}
+The $k$-fold nested commutator of a \emph{particle rank} $m_O$ operator $O$ with
+cluster operators $T_{n_i}$ of rank $n_i$ acting on the reference determinant:
+\begin{equation}
+  \commutator{\commutator{\commutator{O}{T_{n_1}}}{\ldots}}{T_{n_k}} \ket{\mathrm{HF}}
+\end{equation}
+generates a linear combination of determinants with excitation ranks $v$ in the
+range:
+\begin{equation}
+  \sum_{i=1}^k n_i - m_O \leq v \leq \sum_{i=1}^k n_i + m_O -k
+\end{equation}
+\end{lemma}
+\begin{corollary}[Excited states overlaps]
+The overlap of a determinant $\mu_n$ with excitation rank $n$ onto the linear
+combination of determinants generated by the $k$-fold nested commutator of an
+operator $O$ with particle rank $m_O$ with cluster operators $T_{n_i}$ of rank $n_i$ is
+nonzero if and only if the sum of cluster operator ranks satisfies:
+\begin{equation}
+  n - m_O + k \leq \sum_{i=1}^k n_i \leq n + m_O
+\end{equation}
+In compact form:
+\begin{equation}
+  \braket{\mu_n |
+  \commutator{\commutator{\commutator{O}{T_{n_1}}}{\ldots}}{T_{n_k}} | \mathrm{HF}} \neq 0
+  \Longleftrightarrow
+  n - m_O + k \leq \sum_{i=1}^k n_i \leq n + m_O
+\end{equation}
+\end{corollary}
+
+The \acrshort*{CCSD} Lagrangian is then:
+\begin{equation}\label{eq:ccsd-lagrangian}
+ \begin{aligned}
+ \mathcal{L}(\tamp{}, \tbar{})_\mathrm{CCSD}
+&=
+ E_0
+ + \sum_{u=1}^{2}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
+ + \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
+ &+ \braket{\tbar{1} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
+ &+ \braket{\tbar{2} | \check{\Phi} + \commutator{\check{\Phi}}{T_2}
+   +\frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} | \mathrm{HF}}
+  \end{aligned}
+\end{equation}
+with the governing equations:
+\begin{subequations}\label{eq:ccsd-equations}
+  \begin{align}
+   \denom{\mu_1}\tamp{\mu_1} &+ \braket{\mu_1 | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}}
+             = 0 \label{eq:ccsd-singles-amplitudes}\\
+   \denom{\mu_2}\tamp{\mu_2} &+
+   \braket{\mu_2 | \check{\Phi} + \commutator{\check{\Phi}}{T_2} +\frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} | \mathrm{HF}}
+             = 0 \label{eq:ccsd-doubles-amplitudes}\\
+    \denom{\mu_1}\tbar{\mu_1} &+
+    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}} +
+    \braket{\tbar{1} |
+      \commutator{\check{\Phi}}{\cluster{\mu_1}}
+    + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+    | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{2} |
+      \commutator{\check{\Phi}}{\cluster{\mu_1}}
+    + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+    | \mathrm{HF}}
+             = 0 \label{eq:ccsd-singles-multipliers} \\
+    \denom{\mu_2}\tbar{\mu_2} &+
+    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}} +
+    \braket{\tbar{1} |
+      \commutator{\check{\Phi}}{\cluster{\mu_2}}
+    | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{2} |
+      \commutator{\check{\Phi}}{\cluster{\mu_2}}
+    + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_2}}}{T_2}
+    | \mathrm{HF}}
+             = 0 \label{eq:ccsd-doubles-multipliers}
+  \end{align}
+\end{subequations}
+Equations \eqref{eq:ccsd-doubles-amplitudes},
+\eqref{eq:ccsd-singles-multipliers} and \eqref{eq:ccsd-doubles-multipliers} can
+be compared with the corresponding \acrshort{CC2} equations
+\eqref{eq:cc2-doubles-amplitudes}, \eqref{eq:cc2-singles-multipliers} and
+\eqref{eq:cc2-doubles-multipliers}, respectively, to highlight which terms were neglected in
+constructing the iterative method for the approximate inclusion of double excitations.
+
+The \acrshort{CCSDT} Lagrangian is built in the same exact manner:
+\begin{equation}\label{eq:ccsdt-lagrangian}
+ \begin{aligned}
+ \mathcal{L}(\tamp{}, \tbar{})_\mathrm{CCSDT}
+&=
+ E_0
+ + \sum_{u=1}^{3}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
+ + \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
+ &+ \braket{\tbar{1} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} + \commutator{\check{\Phi}}{T_3} | \mathrm{HF}} \\
+ &+ \braket{\tbar{2} | \check{\Phi} + \commutator{\check{\Phi}}{T_2}
+   +\frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{T_3} | \mathrm{HF}} \\
+ &+ \braket{\tbar{3} | \commutator{\check{\Phi}}{T_2}
+   + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{T_3}
+   + \commutator{\commutator{\check{\Phi}}{T_2}}{T_3}
+   | \mathrm{HF}}
+  \end{aligned}
+\end{equation}
+Notice that the reference expectation value of the similarity transformed
+fluctuation potential $\overline{\Phi}$ is unchanged with respect to the
+expression in the \acrshort{CCSD} Lagrangian.
+This is a consequence of the well-known fact that the \acrshort*{CC} energy can
+be expressed purely in terms of $T_1$ and $T_2$:
+\begin{equation}
+E_\mathrm{CC} = \braket{\mathrm{HF} | \overline{H} | \mathrm{HF}}
+=
+\braket{\mathrm{HF} | H(1 + T_1 + T_2 + T_1^2) | \mathrm{HF}}
+\end{equation}
+The \acrshort{CCSDT} amplitude equations are:
+\begin{subequations}\label{eq:ccsdt-amplitudes}
+  \begin{align}
+   \denom{\mu_1}\tamp{\mu_1} &+ \braket{\mu_1 |
+   \check{\Phi} + \commutator{\check{\Phi}}{T_2} + \commutator{\check{\Phi}}{T_3}
+   | \mathrm{HF}}
+             = 0 \label{eq:ccsdt-singles-amplitudes}\\
+   \denom{\mu_2}\tamp{\mu_2} &+ \braket{\mu_2 |
+   \check{\Phi} + \commutator{\check{\Phi}}{T_2}
+   + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{T_3}
+   | \mathrm{HF}}
+             = 0 \label{eq:ccsdt-doubles-amplitudes} \\
+   \denom{\mu_3}\tamp{\mu_3} &+ \braket{\mu_3 |
+   \commutator{\check{\Phi}}{T_2}
+   + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{T_3}
+   + \commutator{\commutator{\check{\Phi}}{T_2}}{T_3}
+   | \mathrm{HF}}
+             = 0 \label{eq:ccsdt-triples-amplitudes}
+  \end{align}
+\end{subequations}
+while the multipliers are determined by solving the following:
+\begin{subequations}\label{eq:ccsdt-multipliers}
+  \begin{align}
+    \denom{\mu_1}\tbar{\mu_1} &+
+    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}} +
+    \braket{\tbar{1} |
+      \commutator{\check{\Phi}}{\cluster{\mu_1}}
+    + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+    | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{2} |
+      \commutator{\check{\Phi}}{\cluster{\mu_1}}
+    + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+    + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_3}
+    | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{3} |
+      \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+    + \frac{1}{2}\commutator{\commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}}{T_2}
+    + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_3}
+    | \mathrm{HF}}
+             = 0 \label{eq:ccsdt-singles-multipliers} \\
+    \denom{\mu_2}\tbar{\mu_2} &+
+    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}} +
+    \braket{\tbar{1} |
+      \commutator{\check{\Phi}}{\cluster{\mu_2}}
+    | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{2} |
+      \commutator{\check{\Phi}}{\cluster{\mu_2}}
+    + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_2}}}{T_2}
+    | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{3} |
+      \commutator{\check{\Phi}}{\cluster{\mu_2}}
+    +  \commutator{\commutator{\check{\Phi}}{\cluster{\mu_2}}}{T_2}
+    +  \commutator{\commutator{\check{\Phi}}{\cluster{\mu_2}}}{T_3}
+    | \mathrm{HF}}
+             = 0 \label{eq:ccsdt-doubles-multipliers} \\
+    \denom{\mu_3}\tbar{\mu_3} &+
+    \braket{\tbar{1} |
+      \commutator{\check{\Phi}}{\cluster{\mu_3}}
+    | \mathrm{HF}}
+    + \braket{\tbar{2} |
+      \commutator{\check{\Phi}}{\cluster{\mu_3}}
+    | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{3} |
+      \commutator{\check{\Phi}}{\cluster{\mu_3}}
+    +  \commutator{\commutator{\check{\Phi}}{\cluster{\mu_3}}}{T_2}
+    | \mathrm{HF}}
+             = 0 \label{eq:ccsdt-triples-multipliers}
+  \end{align}
+\end{subequations}
+A comparison of equations \eqref{eq:ccsdt-triples-amplitudes},
+\eqref{eq:ccsdt-singles-multipliers}, \eqref{eq:ccsdt-doubles-multipliers} and
+\eqref{eq:ccsdt-triples-multipliers}
+with equations \eqref{eq:cc3-triples-amplitudes},
+\eqref{eq:cc3-singles-multipliers}, \eqref{eq:cc3-doubles-multipliers} and
+\eqref{eq:cc3-triples-multipliers}, respectively, serves to highlight which terms were
+neglected in constructing the \acrshort{CC3} method.
+
+\section{Coupled Cluster Expectation Values}\label{sec:CC-expval}
+
+A similar analysis in terms of $T_1$-transformed operators can be given for the
+expectation value of one-electron operators, Eq.
+\eqref{eq:cc-expectation-value}.
+First of all, by virtue of the cluster commutation
+condition,\autocite{Helgaker2000-tz} the commutator expansion of the similarity
+transformation of a general one-electron operator truncates after the twofold
+nested commutator:
+\begin{equation}
+\overline{O} = O
+  + \commutator{O}{T}
+  + \frac{1}{2}\commutator{\commutator{O}{T}}{T}
+\end{equation}
+where $T$ is the complete cluster operator. However, due to Lemma
+\ref{lemma:exc-rank}, some of the cluster operators will not contribute to the
+expectation value.
+
+For a general truncation level $\mathcal{M}$ one has:
+\begin{equation}
+  \begin{aligned}
+  O(\tamp{}, \tbar{}) &=
+  \braket{\mathrm{HF} | O + \BCHfirst{O}{T} + \BCHsecond{O}{T} | \mathrm{HF}} \\
+  &+ \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | O + \BCHfirst{O}{T} +
+  \BCHsecond{O}{T} | \mathrm{HF}},
+  \end{aligned}
+\end{equation}
+and by virtue of Lemma \ref{lemma:exc-rank} and its Corollary:
+\begin{equation}
+\begin{aligned}
+  O(\tamp{}, \tbar{})
+  &=
+  \braket{\mathrm{HF} | O + \BCHfirst{O}{T_1} | \mathrm{HF}}
+  + \braket{\tbar{1} | O | \mathrm{HF}} \\
+  &+ \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \BCHfirst{O}{T} + \BCHsecond{O}{T} | \mathrm{HF}}
+\end{aligned}
+\end{equation}
+Note that the singles amplitudes assume a unique role in the \acrshort{CC}
+expectation value.
+In the following, we give explicit expressions for the \acrshort{CCS}, \acrshort{CCSD}
+and \acrshort{CCSDT} models. These results are at the basis of our developments
+in Chapter \ref{ch:solvation-correlation}.
+We will refer to the $T_1$-transformed expressions as \emph{dressed}, in contrast to the
+\emph{bare} expressions, where the operator $O$ appears untransformed.
+
+For the \acrshort{CCS} model, the expectation value of a one-electron operator
+is simply:
+\begin{equation}\label{eq:ccs-expectation-value}
+\begin{aligned}
+  O(\tamp{}, \tbar{})_\mathrm{CCS} &=
+  \braket{\mathrm{HF} | \check{O} | \mathrm{HF} }
+  + \braket{\tbar{1} | \check{O} | \mathrm{HF} } \\
+  &=
+  \braket{\mathrm{HF} | O + \commutator{O}{T_1} | \mathrm{HF} }
+  + \braket{\tbar{1} | O + \commutator{O}{T_1}
+  + \frac{1}{2}\commutator{\commutator{O}{T_1}}{T_1} | \mathrm{HF} }
+\end{aligned}
+\end{equation}
+Adding double excitation to the manifold yields:
+\begin{equation}\label{eq:ccsd-expectation-value}
+  \begin{aligned}
+  O(\tamp{}, \tbar{})_\mathrm{CCSD}
+  &=
+    O(\tamp{}, \tbar{})_\mathrm{CCS}
+  + \braket{\tbar{1} | \commutator{\check{O}}{T_2} | \mathrm{HF} }
+  + \braket{\tbar{2} | \commutator{\check{O}}{T_2} | \mathrm{HF} } \\
+  &=
+    O(\tamp{}, \tbar{})_\mathrm{CCS}
+  + \braket{\tbar{1} | \commutator{O}{T_2} | \mathrm{HF} }
+  + \braket{\tbar{2} | \commutator{O}{T_2}
+      + \commutator{\commutator{O}{T_1}}{T_2}
+  | \mathrm{HF} }
+  \end{aligned}
+\end{equation}
+a number of terms was dropped thanks to Lemma \ref{lemma:exc-rank} and its
+Corollary.
+Eventually, including triples one obtains:
+\begin{equation}\label{eq:ccsdt-expectation-value}
+\begin{aligned}
+  O(\tamp{}, \tbar{})_\mathrm{CCSDT}
+  &=
+    O(\tamp{}, \tbar{})_\mathrm{CCSD}
+  + \braket{\tbar{2} | \commutator{\check{O}}{T_3} | \mathrm{HF} } \\
+  &+ \braket{\tbar{3} | \commutator{\check{O}}{T_3}
+    + \frac{1}{2}\commutator{\commutator{\check{O}}{T_2}}{T_2}
+  | \mathrm{HF} } \\
+  &=
+    O(\tamp{}, \tbar{})_\mathrm{CCSD}
+  + \braket{\tbar{2} | \commutator{O}{T_3} | \mathrm{HF} } \\
+  &+ \braket{\tbar{3} | \commutator{O}{T_3}
+    + \commutator{\commutator{O}{T_1}}{T_3}
+    + \frac{1}{2}\commutator{\commutator{O}{T_2}}{T_2}
+  | \mathrm{HF} }
+\end{aligned}
+\end{equation}
+where Lemma \ref{lemma:exc-rank} and its corollary were again extensively employed.
+
+The use of normal-ordered operators is common in \acrlong*{CC} theory. Both for
+the dressed and bare representations, the expectation values of such operators
+can be formed by simply replacing operator with its normal-ordered counterpart,
+$\check{O}_\mathrm{N}$ or $O$, respectively.
+This achieves elimination of the reference expectation value from the first
+term in equation \eqref{eq:ccs-expectation-value}.
+
+\section{Selected Results in Functional Analysis}
 
@@
 
+\begin{comment}
+
 \section{The Galerkin Method}
@@
 
-\begin{comment}
 \section{Mixed Blabbering}

### 2016-09-14 Chapters/QuantumMechanics.tex
Save the day's work

@@
   \begin{align}
-   \tampEq{\mu_q}(\tamp{}, \tbar{}, \p)  &=
+   \tampEq{\mu_q}(\tamp{}, \tbar{})  &=
    \denom{\mu_q}\tamp{\mu_q} + \braket{\mu_q | \overline{\Phi} | \mathrm{HF}}
              = 0 \label{eq:cc-amplitudes}\\
-   \tbarEq{\mu_q}(\tamp{}, \tbar{}, \p)
+   \tbarEq{\mu_q}(\tamp{}, \tbar{})
     &=
@@
 right \acrshort{CC} states and we introduce the following notation:
-\begin{equation}
+\begin{equation}\label{eq:cc-expectation-value}
   O(\tamp{}, \tbar{})_\mathcal{M} = \braket{\mathrm{HF} | \overline{O} | \mathrm{HF} }
@@
 not affect the particle rank of the electronic
-Hamiltonian.\autocite{Koch1994-vr, Helgaker2000-tz}
+Hamiltonian,\autocite{Koch1994-vr, Helgaker2000-tz}
+see Appendix \ref{app:mathematical-results} for details.
 The assumption that $T_1$ is of zeroth order is contrary to the usual
@@
     \denom{\mu_1}\tamp{\mu_1} &+ \braket{\mu_1 | \check{\Phi} +
-    \commutator{\check{\Phi}}{Q_2} | \mathrm{HF}} = 0 \\
+    \commutator{\check{\Phi}}{Q_2} | \mathrm{HF}} = 0
+    \label{eq:cc2-singles-amplitudes} \\
     \denom{\mu_2}\tamp{\mu_2} &+ \braket{\mu_2 | \check{\Phi} | \mathrm{HF}} = 0
+    \label{eq:cc2-doubles-amplitudes}
   \end{align}
@@
    &+ \braket{\tbar{2} | \commutator{\check{\Phi}}{\cluster{\mu_1}} |
-   \mathrm{HF}} = 0 \\
+   \mathrm{HF}} = 0 \label{eq:cc2-singles-multipliers} \\
     \denom{\mu_2}\tbar{\mu_2} &+
@@
      | \mathrm{HF}}
-    = 0
+     = 0 \label{eq:cc2-doubles-multipliers}
   \end{align}
 \end{subequations}
+Formally, \acrshort*{CC2} expectation values can be computed using the expression for
+the \acrshort*{CCSD} and the approximate amplitudes and multipliers:
+\begin{equation}\label{eq:cc2-expectation-value}
+  \begin{aligned}
+  O(\tamp{}, \tbar{})_\mathrm{CC2}
+  &=
+  \braket{\mathrm{HF} | \check{O} | \mathrm{HF} }
+  + \braket{\tbar{1} | \check{O} | \mathrm{HF} }
+  + \braket{\tbar{1} | \commutator{\check{O}}{Q_2} | \mathrm{HF} }
+  + \braket{\tbar{2} | \commutator{\check{O}}{Q_2} | \mathrm{HF} } \\
+  &=
+  O(\tamp{}, \tbar{})_\mathrm{CCS}
+  + \braket{\tbar{1} | \commutator{\check{O}}{Q_2} | \mathrm{HF} }
+  + \braket{\tbar{2} | \commutator{\check{O}}{Q_2} | \mathrm{HF} }
+  \end{aligned}
+\end{equation}
 
@@
     + \commutator{\check{\Phi}}{Q_3}
-    | \mathrm{HF}} = 0 \\
+    | \mathrm{HF}} = 0 \label{eq:cc3-singles-amplitudes} \\
     \denom{\mu_2}\tamp{\mu_2} &+ \braket{\mu_2 |
@@
     + \commutator{\check{\Phi}}{Q_3}
-    | \mathrm{HF}} = 0 \\
+    | \mathrm{HF}} = 0 \label{eq:cc3-doubles-amplitudes} \\
     \denom{\mu_3}\tamp{\mu_3} &+ \braket{\mu_3 |
     \commutator{\check{\Phi}}{T_2}
-    | \mathrm{HF}} = 0
+    | \mathrm{HF}} = 0 \label{eq:cc3-triples-amplitudes}
   \end{align}
@@
       \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
-     | \mathrm{HF}} = 0 \\
+     | \mathrm{HF}} = 0 \label{eq:cc3-singles-multipliers} \\
     \denom{\mu_2}\tbar{\mu_2} &+
@@
      | \mathrm{HF}}
-    = 0 \\
+    = 0 \label{eq:cc3-doubles-multipliers} \\
     \denom{\mu_3}\tbar{\mu_3} &+
@@
      | \mathrm{HF}}
-    = 0
+    = 0 \label{eq:cc3-triples-multipliers}
   \end{align}
 \end{subequations}
+\acrshort*{CC3} expectation values can be computed as follows:
+\begin{equation}\label{eq:cc3-expectation-value}
+  \begin{aligned}
+  O(\tamp{}, \tbar{})_\mathrm{CC3}
+  &=
+  O(\tamp{}, \tbar{})_\mathrm{CCS}
+  + \braket{\tbar{1} | \commutator{\check{O}}{T_2} | \mathrm{HF} }
+  + \braket{\tbar{2} | \commutator{\check{O}}{T_2} | \mathrm{HF} } \\
+  &+
+  \braket{\tbar{2} | \commutator{\check{O}}{Q_3} | \mathrm{HF} }
+  + \braket{\tbar{3} |
+     \commutator{\check{O}}{Q_3}
+   + \frac{1}{2}\commutator{\commutator{\check{O}}{T_2}}{T_2}
+                     | \mathrm{HF} } \\
+  &=
+  O(\tamp{}, \tbar{})_\mathrm{CCSD}
+  +
+  \braket{\tbar{2} | \commutator{\check{O}}{Q_3} | \mathrm{HF} }
+  + \braket{\tbar{3} |
+     \commutator{\check{O}}{Q_3}
+                     | \mathrm{HF} } \\
+  &+ \braket{\tbar{3} |
+     \frac{1}{2}\commutator{\commutator{\check{O}}{T_2}}{T_2}
+                     | \mathrm{HF} }
+  \end{aligned}
+\end{equation}
 

### 2016-09-14 Chapters/SolvationAndCorrelation.tex
Save the day's work

@@
   \begin{aligned}
-\s(\tamp{}, \tbar{}) &= \s_\mathrm{HF} + \s_{\mathrm{N}}(\tamp{}, \tbar{}) \\
+    \s(\tamp{}, \tbar{})_\mathcal{M} &= \s_\mathrm{HF} + \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathcal{M} \\
                      &=
@@
   + \braket{\mathrm{HF} | \tilde{\s} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\s} | \mathrm{HF}}
+  + \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \overline{\s} | \mathrm{HF}}
   \end{aligned}
@@
   \begin{aligned}
-    \s_{\mathrm{N}}(\tamp{}, \tbar{})
+    \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathcal{M}
   &=
    \braket{\mathrm{HF} | \tilde{\s} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\s} | \mathrm{HF}} \\
+  + \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \overline{\s} | \mathrm{HF}} \\
   &=
     \braket{\mathrm{HF} | \overline{\s} - \s_\mathrm{HF} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\s} - \s_\mathrm{HF} | \mathrm{HF}}
+  + \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \overline{\s} - \s_\mathrm{HF} | \mathrm{HF}}
   \end{aligned}
@@
   \frac{1}{2}({}^t\p_\mathrm{HF}+{}^t\p)\V(\p_\mathrm{HF}+\p)
-  + ({}^t\p_\mathrm{HF}+{}^t\p)(\s_\mathrm{HF}+\s_{\mathrm{N}}(\tamp{},\tbar{})) \\
+  + ({}^t\p_\mathrm{HF}+{}^t\p)(\s_\mathrm{HF}+\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}) \\
   &=
@@
   + \frac{1}{2}{}^t\p_\mathrm{HF}\V\p + \frac{1}{2}{}^t\p\V\p_\mathrm{HF} \\
-  &+ {}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})
+  &+ {}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}
   + {}^t\p\s_\mathrm{HF}
-  + \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{}) \\
+  + \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M} \\
   &=
     U_\mathrm{pol}^\mathrm{ref}
-  + \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})
-  + {}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})
+  + \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}
+  + {}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}
   \end{aligned}
@@
   \begin{aligned}
-  \lag{M}(\tamp{}, \tbar{}, \p) &=
+    \lag(\tamp{}, \tbar{}, \p)_\mathcal{M} &=
   \braket{\mathrm{HF} | \overline{H_0} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tamp{i} | \overline{H_0} | \mathrm{HF}} \\
-  &+
-  \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})
-  + {}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})
+  + \sum_{u=1}^{\mathcal{M}}\braket{\tamp{u} | \overline{H_0} | \mathrm{HF}} \\
+  &+
+  \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}
+  + {}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}
   + U_\mathrm{pol}^\mathrm{ref}
@@
 \begin{equation}\label{eq:pted-cc}
-  \lag{M}(\tamp{}, \tbar{}, \p) =
-  {}^\mathrm{PTE}\lag{M}
-  + \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})
+  \lag(\tamp{}, \tbar{}, \p)_\mathcal{M} =
+  {}^\mathrm{PTE}\lag(\tamp{}, \tbar{})_\mathcal{M}
+  + \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}
 \end{equation}
@@
 \begin{equation}
-{}^\mathrm{PTE}\lag{M}
+  {}^\mathrm{PTE}\lag(\tamp{}, \tbar{})_\mathcal{M}
   =
   \braket{\mathrm{HF} | \overline{H} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}}
+  + \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \overline{H} | \mathrm{HF}}
   + U_\mathrm{pol}^\mathrm{ref}
@@
 \begin{equation}
-{}^\mathrm{PTE}\lag{M}
+  {}^\mathrm{PTE}\lag(\tamp{}, \tbar{})_\mathcal{M}
   =
   G_0
-  + \sum_{i=1}^{\mathcal{M}}\tbar{\mu_i}\denom{\mu_i}\tamp{\mu_i}
+  + \sum_{u=1}^{\mathcal{M}}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
   + \braket{\mathrm{HF} | \overline{\Phi} | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{\Phi} | \mathrm{HF}}
+  + \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \overline{\Phi} | \mathrm{HF}}
 \end{equation}
@@
 involved.
-We develop this analysis in terms of excitation ranks of the operators
-involved, based on the following:
-\begin{equation}
-  \braket{\mu_n |
-  \commutator{\commutator{\commutator{O}{T_{n_1}}}{\ldots}}{T_{n_k}} | \mathrm{HF}} \neq 0
-  \Longleftrightarrow
-  n - m_O + k \leq \sum_{i=1}^k n_i \leq n + m_O
-\end{equation}
-where $n$ is the excitation rank of the $\ket{\mu_n}$ determinant, $k$
-is the total number of cluster operators appearing in the nested
-commutator, $n_i$ are their respective excitations ranks and $m_O$
-is the \emph{particle rank} of the operator $O$.\autocite{Helgaker2000-tz}
-This follows from the realization that:
-\begin{equation}
-  \commutator{\commutator{\commutator{O}{T_{n_1}}}{\ldots}}{T_{n_k}} \ket{\mathrm{HF}}
-\end{equation}
-is a linear combination of determinants with excitation ranks $v$ in the
-range:
-\begin{equation}
-  \sum_{i=1}^k n_i - m_O \leq v \leq \sum_{i=1}^k n_i + m_O -k
-\end{equation}
-
-We first expand the source terms entering the Lagrangian as follows:
-\begin{equation}
-  \begin{aligned}
-  \s_{\mathrm{N}}(\tamp{}, \tbar{}) &=
-  \braket{\mathrm{HF} | \BCHfirst{\s}{T} + \BCHsecond{\s}{T} | \mathrm{HF}} \\
-  &+ \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \s + \BCHfirst{\s}{T} +
-  \BCHsecond{\s}{T} | \mathrm{HF}},
-  \end{aligned}
-\end{equation}
-and since $\s$ has particle rank $m_\s = 1$ we can conclude that the
-following relationships hold:
-\begin{subequations}
-\begin{align}
-  \braket{\mathrm{HF} |
-  \BCHfirst{\s}{T_t}
-  | \mathrm{HF}} &\neq 0 \quad \text{iff} \quad 0 \leq t \leq 1 \Longleftrightarrow t = 1 \\
-  \braket{\mathrm{HF} |
-  \commutator{\commutator{\s}{T_t}}{T_u}
-  | \mathrm{HF}} &\neq 0 \quad \text{iff} \quad t + u = 1
-  \label{eq:HF-double-commutator} \\
-  \braket{\tbar{i} |
-  \s
-  | \mathrm{HF} } &\neq 0 \quad \text{iff} \quad i = 1 \\
-  \braket{\tbar{i} |
-  \BCHfirst{\s}{T_t}
-  | \mathrm{HF} } &\neq 0 \quad \text{iff} \quad i \leq t \leq i + 1 \\
-  \braket{\tbar{i} |
-  \commutator{\commutator{\s}{T_t}}{T_u}
-  | \mathrm{HF} } &\neq 0 \quad \text{iff} \quad t + u = i + 1
-\end{align}
-\end{subequations}
-By virtue of Eq.~\eqref{eq:HF-double-commutator} the \acrshort{HF}
-expectation value of the double commutator can be dropped. Thus, for a
-general truncation level we obtain:
-\begin{equation}\label{eq:source-term}
-  \s_{\mathrm{N}}(\tamp{}, \tbar{}) =
-  \braket{\mathrm{HF} | \BCHfirst{\s}{T_1} | \mathrm{HF}}
-  + \braket{\tbar{1} | \s | \mathrm{HF}}
-  + \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \BCHfirst{\s}{T} + \BCHsecond{\s}{T} | \mathrm{HF}}
-\end{equation}
-Note that the singles amplitudes assume a unique role in the \acrshort{CC}
-expression of the source term.
-In addition, double commutators involving the highest order cluster
-operators can only lead to nonzero terms when $T_1$ is also involved and
-by projection onto the highest order multiplier state, \eg~
-$\braket{\tbar{2} | \commutator{\commutator{\s}{T_1}}{T_2} |
-\mathrm{HF}} \neq 0$ and $\braket{\tbar{3} |
-\commutator{\commutator{\s}{T_1}}{T_3} | \mathrm{HF}} \neq 0$
-More explicitly, for the \gls{CCS} model ($\mathcal{M} = 1$) one has:
+This analysis is carried out in detail in Appendix \ref{sec:CC-expval}
+and we report here the final results for the \gls{CCS}, \gls{CCSD} and
+\gls{CCSDT} models.
+
+For the \gls{CCS} model ($\mathcal{M} = 1$) one has:
 \begin{equation}\label{eq:source-ccs}
@@
   \begin{align}
-   \tampEq{\mu_i}(\tamp{}, \tbar{}, \p)  &=
-   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
-   + {}^t\p\braket{\mu_i | \overline{\s} | \mathrm{HF}}
+   \tampEq{\mu_q}(\tamp{}, \tbar{}, \p)  &=
+   \denom{\mu_q}\tamp{\mu_q} + \braket{\mu_q | \overline{\Phi} | \mathrm{HF}}
+   + {}^t\p\braket{\mu_q | \overline{\s} | \mathrm{HF}}
              = 0 \label{eq:pted-cc-amplitudes}\\
-   \tbarEq{\mu_i}(\tamp{}, \tbar{}, \p)
-    &=
-    \denom{\mu_i}\tbar{\mu_i} +
-    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
-    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} |
-    \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}}
+   \tbarEq{\mu_q}(\tamp{}, \tbar{}, \p)
+    &=
+    \denom{\mu_q}\tbar{\mu_q} +
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}} +
+    \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
+    \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}}
     \nonumber \\
     &+
-    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}\delta_{\mu_i\mu_1} \nonumber \\
-    &+
-    {}^t\p\sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}
-    +
-    {}^t\p\sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\commutator{\s}{T}}{\cluster{\mu_i}} | \mathrm{HF}}
+    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1} \nonumber \\
+    &+
+    {}^t\p\sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}
+    +
+    {}^t\p\sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \commutator{\commutator{\s}{T}}{\cluster{\mu_q}} | \mathrm{HF}}
              = 0 \label{eq:pted-cc-multipliers}\\
@@
     &=
-    \V\p + \s_{\mathrm{N}}(\tamp{}, \tbar{}) = 0
+    \V\p + \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathcal{M} = 0
     \label{eq:pted-cc-polarization}
@@
 A coupling of the amplitudes and multipliers equations is introduced by
-the presence of the $\s_{\mathrm{N}}(\tamp{}, \tbar{})$ expectation value
+the presence of the $\s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathcal{M}$ expectation value
 in the polarization equation.
@@
 Moreover, the leading terms in the correlated source operator
-expectation value, Eq.~\eqref{eq:source-term}, prominently involve the
+expectation value prominently involve the
 singles cluster operator and the singles multiplier state.
@@
 \begin{equation}\label{eq:ptes-cc}
-  \begin{aligned}
-  {}^{\mathrm{PTES}}\lag{M}(\tamp{}, \tbar{}, \p) =
-  \braket{\mathrm{HF} | \overline{H} | \mathrm{HF}}
-  &+ \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}} \\
-  &+ \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},
+  {}^{\mathrm{PTES}}\lag(\tamp{}, \tbar{}, \p)_\mathcal{M} =
+  {}^\mathrm{PTE}\lag(\tamp{}, \tbar{})_\mathcal{M}
+  + \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},
   \tbar{})_\mathrm{CCS}
-  \end{aligned}
 \end{equation}
@@
   \begin{align}
-   \tampEq{\mu_i}(\tamp{}, \tbar{}, \p)  &=
-   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
-   + {}^t\p\braket{\mu_i | \check{\s} | \mathrm{HF}}
+   \tampEq{\mu_q}(\tamp{}, \tbar{}, \p)  &=
+   \denom{\mu_q}\tamp{\mu_q} + \braket{\mu_q | \overline{\Phi} | \mathrm{HF}}
+   + {}^t\p\braket{\mu_q | \check{\s} | \mathrm{HF}}
              = 0 \\
-   \tbarEq{\mu_i}(\tamp{}, \tbar{}, \p)
-    &=
-    \denom{\mu_i}\tbar{\mu_i} +
-    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
-    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} |
-    \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}}
+   \tbarEq{\mu_q}(\tamp{}, \tbar{}, \p)
+    &=
+    \denom{\mu_q}\tbar{\mu_q} +
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}} +
+    \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
+    \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}}
     \nonumber \\
     &+
-    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}\delta_{\mu_i\mu_1} \nonumber \\
-    &+
-    {}^t\p\sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}
-    +
-    {}^t\p\sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\commutator{\s}{T}}{\cluster{\mu_i}} | \mathrm{HF}}
+    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1} \nonumber \\
+    &+
+    {}^t\p\sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
+    \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}
+    +
+    {}^t\p\sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
+    \commutator{\commutator{\s}{T}}{\cluster{\mu_q}} | \mathrm{HF}}
              = 0 \\
@@
   \begin{align}
-   {}^{\mathrm{PTES}}\tampEq{\mu_i}(\tamp{}, \tbar{}, \p)  &=
-   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
-   + {}^t\p\braket{\mu_i | \check{\s} | \mathrm{HF}}
+   {}^{\mathrm{PTES}}\tampEq{\mu_q}(\tamp{}, \tbar{}, \p)  &=
+   \denom{\mu_i}\tamp{\mu_q} + \braket{\mu_q | \overline{\Phi} | \mathrm{HF}}
+   + {}^t\p\braket{\mu_q | \check{\s} | \mathrm{HF}}
              = 0 \label{eq:ptes-cc-amplitudes} \\
@@
   \begin{align}
-   {}^{\mathrm{PTES}}\tbarEq{\mu_i}(\tamp{}, \tbar{}, \p)
-    &=
-    \denom{\mu_i}\tbar{\mu_i} +
-    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
-    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} |
-    \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}}
+   {}^{\mathrm{PTES}}\tbarEq{\mu_q}(\tamp{}, \tbar{}, \p)
+    &=
+    \denom{\mu_q}\tbar{\mu_q} +
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}} +
+    \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
+    \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}}
     \nonumber \\
     &+
-    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}\delta_{\mu_i\mu_1} \nonumber \\
-    &+
-    {}^t\p\sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}
-    +
-    {}^t\p\sum_{j}^{\mathcal{M}}\braket{\tbar{j} | \commutator{\commutator{\s}{T}}{\cluster{\mu_i}} | \mathrm{HF}}
+    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_q}} |
+    \mathrm{HF}}\delta_{\mu_q\mu_1} \nonumber \\
+    &+
+    {}^t\p\sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
+    \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}
+    +
+    {}^t\p\sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
+    \commutator{\commutator{\s}{T}}{\cluster{\mu_q}} | \mathrm{HF}}
              = 0 \label{eq:ptes-cc-multipliers}\\
@@
 \begin{equation}\label{eq:pte-s-cc}
-  \begin{aligned}
-  {}^{\mathrm{PTE(S)}}\lag{M}(\tamp{}, \tbar{}, \p) =
-  \braket{\mathrm{HF} | \overline{H} | \mathrm{HF}}
-  &+ \sum_{i=1}^{\mathcal{M}}\braket{\tbar{i} | \overline{H} | \mathrm{HF}} \\
-  &+ \frac{1}{2}{}^t\p\V\p + {}^t\p\braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}}
-  \end{aligned}
+    {}^{\mathrm{PTE(S)}}\lag(\tamp{}, \tbar{}, \p)_\mathcal{M} =
+  {}^\mathrm{PTE}\lag(\tamp{}, \tbar{})_\mathcal{M}
+  + \frac{1}{2}{}^t\p\V\p + {}^t\p\braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}}
 \end{equation}
@@
   \begin{align}
-  {}^{\mathrm{PTE(S)}}\tampEq{\mu_i}(\tamp{}, \tbar{})  &=
-   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
+  {}^{\mathrm{PTE(S)}}\tampEq{\mu_q}(\tamp{}, \tbar{})  &=
+   \denom{\mu_q}\tamp{\mu_q} + \braket{\mu_q | \overline{\Phi} | \mathrm{HF}}
    = 0 \\
-   {}^{\mathrm{PTE(S)}}\tbarEq{\mu_i}(\tamp{}, \tbar{}, \p) &=
-    \denom{\mu_i}\tbar{\mu_i} +
-    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
-    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} |
-    \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}}
+   {}^{\mathrm{PTE(S)}}\tbarEq{\mu_q}(\tamp{}, \tbar{}, \p) &=
+    \denom{\mu_q}\tbar{\mu_q} +
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}} +
+    \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
+    \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}}
     \nonumber \\
     &+
-    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}\delta_{\mu_i\mu_1}
+    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
     = 0 \\
@@
   \begin{align}
- {}^{\mathrm{PTE}}\tampEq{\mu_i}(\tamp{}, \tbar{}, \p)  &=
-   \denom{\mu_i}\tamp{\mu_i} + \braket{\mu_i | \overline{\Phi} | \mathrm{HF}}
+ {}^{\mathrm{PTE}}\tampEq{\mu_q}(\tamp{}, \tbar{}, \p)  &=
+   \denom{\mu_q}\tamp{\mu_q} + \braket{\mu_q | \overline{\Phi} | \mathrm{HF}}
     = 0 \\
-   {}^{\mathrm{PTE}}\tbarEq{\mu_i}(\tamp{}, \tbar{}, \p)
-    &=
-    \denom{\mu_i}\tbar{\mu_i} +
-    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}} +
-    \sum_{j}^{\mathcal{M}}\braket{\tbar{j} |
-    \commutator{\overline{\Phi}}{\cluster{\mu_i}} | \mathrm{HF}}
+   {}^{\mathrm{PTE}}\tbarEq{\mu_q}(\tamp{}, \tbar{}, \p)
+    &=
+    \denom{\mu_q}\tbar{\mu_q} +
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}} +
+    \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
+    \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}}
     = 0
@@
 Eqs.~\eqref{eq:pted-cc-amplitudes} and \eqref{eq:pted-cc-multipliers},
-already shows that $\tamp{\mu_i}^{[0]} = 0, \,\forall i$ and
-$\tbar{\mu_i}^{[0]} = 0, \, \forall i$.
+already shows that $\tamp{\mu_u}^{[0]} = 0, \,\forall u$ and
+$\tbar{\mu_u}^{[0]} = 0, \, \forall u$.
 Thus cluster operators can be of order 1 and higher. We introduce the
@@
 effective, variational Lagrangian implies the validity of the $2n+1$
-rule for the amplitudes $\tamp{\mu_i}$ and polarization $\p$ and of the
-$2n+2$ rule for the multipliers $\tbar{\mu_i}$.
+rule for the amplitudes $\tamp{\mu_u}$ and polarization $\p$ and of the
+$2n+2$ rule for the multipliers $\tbar{\mu_u}$.
 We can thus derive energy corrections up to fifth order by means of the
@@
 The polarization equation couples to the multipliers and amplitudes
-equations \emph{via} the source term, see Eq.~\eqref{eq:source-term}.
+equations \emph{via} the source term.
 Its perturbative expansion will be given as:
 \begin{equation}\label{eq:source-term-pt}
-  \s_{\mathrm{N}}(\tamp{},\tbar{}) =
+  \s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M} =
   \s_{\mathrm{N}}^{[0]}
@@
 the perturbation.
-The order of the contributions to $\s_{\mathrm{N}}(\tamp{},\tbar{})$ will be
-solely determined by the amplitudes and multipliers.
-From the structure of \eqref{eq:pted-cc-polarization}, it is already
-clear that $\s_{\mathrm{N}}(\tamp{},\tbar{})$ is at least first order in the
-fluctuation potential:
-\begin{equation}
-  \s_{\mathrm{N}}(\tamp{},\tbar{}) =
+The order of the contributions to
+$\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}$ will be solely determined
+by the amplitudes and multipliers. From the structure of
+\eqref{eq:pted-cc-polarization}, it is already clear that
+$\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}$ is at least first order
+in the fluctuation potential:
+\begin{equation}
+  \s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M} =
   \s_{\mathrm{N}}^{[1]}
@@
 \begin{equation}\label{eq:1st-order-amp}
-  \Omega_{\mu_i}^{[1]} = \denom{\mu_i}\tamp{\mu_i}^{[1]}
-  + \braket{\mu_i | \Phi | \mathrm{HF}}
-  + {}^t\p^{[1]}\braket{\mu_i | \s | \mathrm{HF}}\delta_{\mu_i\mu_1}
+  \Omega_{\mu_q}^{[1]} = \denom{\mu_q}\tamp{\mu_q}^{[1]}
+  + \braket{\mu_q | \Phi | \mathrm{HF}}
+  + {}^t\p^{[1]}\braket{\mu_q | \s | \mathrm{HF}}\delta_{\mu_q\mu_1}
   = 0
@@
 \begin{equation}%\label{eq:1st-order-mult}
-  \bar{\Omega}_{\mu_i}^{[1]} =
-    \denom{\mu_i}\tbar{\mu_i}^{[1]}
-    + \braket{\mathrm{HF} | \commutator{\Phi}{\cluster{\mu_i}} | \mathrm{HF}}
+  \bar{\Omega}_{\mu_q}^{[1]} =
+    \denom{\mu_q}\tbar{\mu_q}^{[1]}
+    + \braket{\mathrm{HF} | \commutator{\Phi}{\cluster{\mu_q}} | \mathrm{HF}}
     + {}^t\p^{[1]}\braket{\mathrm{HF} |
-    \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}\delta_{\mu_i\mu_1} =0
+    \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1} =0
 \end{equation}
@@
 \begin{equation}\label{eq:1st-order-mult}
-  \bar{\Omega}_{\mu_i}^{[1]} =
-    \denom{\mu_i}\tbar{\mu_i}^{[1]}
-    + \braket{\mathrm{HF} | \Phi | \mu_i}
-    + {}^t\p^{[1]}\braket{\mathrm{HF} | \s | \mu_i}\delta_{\mu_i\mu_1} =0
+  \bar{\Omega}_{\mu_q}^{[1]} =
+    \denom{\mu_q}\tbar{\mu_q}^{[1]}
+    + \braket{\mathrm{HF} | \Phi | \mu_q}
+    + {}^t\p^{[1]}\braket{\mathrm{HF} | \s | \mu_q}\delta_{\mu_q\mu_1} =0
 \end{equation}
@@
 \begin{equation}
-  \tamp{\mu_i}^{[1]} = \tbar{\mu_i}^{[1]},\,\,\forall i
+  \tamp{\mu_u}^{[1]} = \tbar{\mu_u}^{[1]},\,\,\forall u
 \end{equation}
@@
   \begin{aligned}
-  \Omega_{\mu_i}^{[2]} &= \denom{\mu_i}\tamp{\mu_i}^{[2]}
-  + \braket{\mu_i | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
-  + {}^t\p^{[2]}\braket{\mu_i | \s | \mathrm{HF}}\delta_{\mu_i\mu_1} \\
-  &+ {}^t\p^{[1]}\braket{\mu_i | \commutator{\s}{T^{[1]}} |
+  \Omega_{\mu_q}^{[2]} &= \denom{\mu_q}\tamp{\mu_q}^{[2]}
+  + \braket{\mu_q | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
+  + {}^t\p^{[2]}\braket{\mu_q | \s | \mathrm{HF}}\delta_{\mu_q\mu_1} \\
+  &+ {}^t\p^{[1]}\braket{\mu_q | \commutator{\s}{T^{[1]}} |
   \mathrm{HF}}
@@
   \begin{aligned}
-  \bar{\Omega}_{\mu_i}^{[2]} &=
-    \denom{\mu_i}\tbar{\mu_i}^{[2]}
-    + \braket{\mathrm{HF} | \commutator{\commutator{\Phi}{T^{[1]}}}{\cluster{\mu_i}} | \mathrm{HF}}
-    + \sum_j\braket{\tbar{j}^{[1]} |
-    \commutator{\Phi}{\cluster{\mu_i}} | \mathrm{HF}} \\
+  \bar{\Omega}_{\mu_q}^{[2]} &=
+    \denom{\mu_q}\tbar{\mu_q}^{[2]}
+    + \braket{\mathrm{HF} | \commutator{\commutator{\Phi}{T^{[1]}}}{\cluster{\mu_q}} | \mathrm{HF}}
+    + \sum_{u=1}^\mathcal{M}\braket{\tbar{u}^{[1]} |
+    \commutator{\Phi}{\cluster{\mu_q}} | \mathrm{HF}} \\
     &+ {}^t\p^{[2]}\braket{\mathrm{HF} |
-    \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}\delta_{\mu_i\mu_1}
+    \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
     + {}^t\p^{[1]}
-    \sum_j\braket{\tbar{j}^{[1]} |
-    \commutator{\s}{\cluster{\mu_i}} | \mathrm{HF}}
+    \sum_{u=1}^\mathcal{M}\braket{\tbar{u}^{[1]} |
+    \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}
     =0
@@
       \commutator{\commutator{\Phi}{T_1^{[1]}}}{\cluster{\mu_1}} | \mathrm{HF}}
-      + \sum_j\braket{\tbar{j}^{[1]} |
+      + \sum_{u=1}^\mathcal{M}\braket{\tbar{u}^{[1]} |
       \commutator{\Phi}{\cluster{\mu_1}} | \mathrm{HF}} \nonumber \\
@@
       \denom{\mu_2}\tbar{\mu_2}^{[2]}
-      + \sum_j\braket{\tbar{j}^{[1]} |
+      + \sum_{u=1}^\mathcal{M}\braket{\tbar{u}^{[1]} |
       \commutator{\Phi}{\cluster{\mu_2}} | \mathrm{HF}}
       + {}^t\p^{[1]}
-      \sum_j\braket{\tbar{j}^{[1]} |
+      \sum_{u=1}^\mathcal{M}\braket{\tbar{u}^{[1]} |
       \commutator{\s}{\cluster{\mu_2}} | \mathrm{HF}}
@@
       \denom{\mu_3}\tbar{\mu_3}^{[2]}
-      + \sum_j\braket{\tbar{j}^{[1]} |
+      + \sum_{u=1}^\mathcal{M}\braket{\tbar{u}^{[1]} |
       \commutator{\Phi}{\cluster{\mu_3}} | \mathrm{HF}}
@@
     G^{[3]} &=
-    \sum_j\braket{\tbar{j}^{[1]} | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
+    \sum_{u=1}^2\braket{\tbar{u}^{[1]} | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
     + {}^t\p^{[1]}\s_{\mathrm{N}}^{[2]}(\tamp{}^{[1]}, \tbar{}^{[1]}) \\
@@
     G^{[4]} &=
-    \sum_j\braket{\tbar{j}^{[1]} | \commutator{\Phi}{T^{[2]}} | \mathrm{HF}}
-    +
-    \sum_j\braket{\tbar{j}^{[1]} |
+    \sum_{u=1}^2\braket{\tbar{u}^{[1]} | \commutator{\Phi}{T^{[2]}} | \mathrm{HF}}
+    +
+    \sum_{u=1}^2\braket{\tbar{u}^{[1]} |
     \frac{1}{2}\commutator{\commutator{\Phi}{T^{[1]}}}{T^{[1]}} | \mathrm{HF}}
@@
     \braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}} | \mathrm{HF}}
-    + \sum_i\braket{\tbar{i}^{[1]}| \commutator{\s}{T^{[1]}} | \mathrm{HF} }
+    +
+    \sum_{u=1}^2\braket{\tbar{u}^{[1]}|
+    \commutator{\s}{T^{[1]}} | \mathrm{HF} }
     \right\rbrace \\
@@
     \left\lbrace
-    \sum_i\braket{\tbar{i}^{[1]} | \commutator{\s}{T^{[2]}} | \mathrm{HF}}
-    +
-    \sum_i\braket{\tbar{i}^{[1]} |\frac{1}{2}
+    \sum_{u=1}^2\braket{\tbar{u}^{[1]}|
+    \commutator{\s}{T^{[2]}} | \mathrm{HF}}
+    +
+    \sum_{u=1}^2\braket{\tbar{u}^{[1]}|
+    \frac{1}{2}
     \commutator{\commutator{\s}{T^{[1]}}}{T^{[1]}}
@@
     G^{[5]} &=
-    \frac{1}{2}\braket{\mathrm{HF} |
+    \braket{\mathrm{HF} | \frac{1}{2}
     \commutator{\commutator{\Phi}{T^{[2]}}}{T^{[2]}}
@@
     +
-    \sum_j \braket{\tbar{j}^{[1]} |
+    \sum_{u=1}^2\braket{\tbar{u}^{[1]}|
     \commutator{\commutator{\Phi}{T^{[2]}}}{T^{[1]}}
@@
     \\
-    &+ \sum_j \braket{\tbar{j}^{[2]} |
+    &+
+    \sum_{u=1}^3\braket{\tbar{u}^{[2]}|
       \commutator{\Phi}{T^{[2]}}
@@
     +
-    \frac{1}{2}\sum_j\braket{\tbar{j}^{[2]} |
+    \sum_{u=1}^3\braket{\tbar{u}^{[2]}|
+    \frac{1}{2}
     \commutator{\commutator{\Phi}{T^{[1]}}}{T^{[1]}}
@@
     + {}^t\p^{[2]}
-    \sum_i\braket{\tbar{i}^{[1]} | \commutator{\s}{T^{[2]}} | \mathrm{HF}} \\
+    \sum_{u=1}^2\braket{\tbar{u}^{[1]} |
+    \commutator{\s}{T^{[2]}} | \mathrm{HF}} \\
     &+ {}^t\p^{[2]}
-    \sum_i\braket{\tbar{i}^{[1]} |
+    \sum_{u=1}^2\braket{\tbar{u}^{[1]} |
     \frac{1}{2}\commutator{\commutator{\s}{T^{[1]}}}{T^{[1]}}
@@
     + {}^t\p^{[2]}
-    \sum_i\braket{\tbar{i}^{[2]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
+    \sum_{u=1}^3\braket{ \tbar{u}^{[2]}
+                                 | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
     \\
     &+ {}^t\p^{[1]}
-    \sum_i\braket{\tbar{i}^{[2]} | \commutator{\s}{T^{[2]}} | \mathrm{HF}}
+    \sum_{u=1}^3\braket{ \tbar{u}^{[2]}
+    \commutator{\s}{T^{[2]}} | \mathrm{HF}}
     + {}^t\p^{[1]}
-    \sum_i\braket{\tbar{i}^{[2]} |
+    \sum_{u=1}^3\braket{ \tbar{u}^{[2]}
     \frac{1}{2}\commutator{\commutator{\s}{T^{[1]}}}{T^{[1]}}
@@
     &+ {}^t\p^{[1]}
-    \sum_i\braket{\tbar{i}^{[1]} |
+    \sum_{u=1}^2\braket{\tbar{u}^{[1]} |
     \commutator{\commutator{\s}{T^{[1]}}}{T^{[2]}}
@@
   &=
-    \sum_{j}\braket{\tbar{j}^{*}| \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
+    \sum_{u=1}^2\braket{\tbar{u}^{*}| \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
     + {}^t\p^{*}\braket{\tbar{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}} \\
@@
  &=
-    \sum_{j}\braket{\tamp{j}^{*}|\commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
+    \sum_{u=1}^2\braket{\tamp{u}^{*}|\commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
     + {}^t\p^{*}\braket{\tamp{2}^{*} | \commutator{\s}{T_3^{*}} | \mathrm{HF}}
@@
   G_{\Lambda(\text{T})} &=
-  \sum_{j=1}^2 \braket{\tbar{j}^{[1]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+  \sum_{u=1}^2 \braket{\tbar{u}^{[1]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
   +
@@
   +
-  \sum_{j=1}^2 \braket{\tbar{j}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
+  \sum_{u=1}^2 \braket{\tbar{u}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
   \\

### 2016-09-14 Chapters/QuantumMechanics.tex
CC2 and CC3 written

@@
 \end{equation}
+Explicit expressions for truncated \acrshort{CC} models can be found in
+Appendix \ref{sec:CC-expval}
 
@@
   E_0
-  + \sum_{u=1}^{\mathcal{2}}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
+  + \sum_{u=1}^{2}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
   + \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{Q_2} | \mathrm{HF}} \\
@@
 \end{subequations}
-Formally, \acrshort*{CC2} expectation values can be computed using the expression for
-the \acrshort*{CCSD} and the approximate amplitudes and multipliers:
+Formally, \acrshort*{CC2} expectation values can be computed using the
+expression for the \acrshort*{CCSD} (see Appendix \ref{sec:CC-expval})
+and the approximate amplitudes and multipliers:
 \begin{equation}\label{eq:cc2-expectation-value}
@@
   E_0
-  + \sum_{u=1}^{\mathcal{3}}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
+  + \sum_{u=1}^{3}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
   + \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\

### 2016-09-14 Chapters/SolvationAndCorrelation.tex
CC2 and CC3 written

@@
 We will now develop a perturbative many-body expansion of the effective
-\acrshort{CC} Lagrangian when a quantum/classical polarizable Hamiltonian is
-used.
+\acrshort{PTED}-\acrshort{CC} Lagrangian \eqref{eq:pted-cc}
 The use of a \acrshort{CC} ansatz for the wave function ensures term-by-term
@@
 The structure of the stationarity conditions,
-Eqs.~\eqref{eq:pted-cc-amplitudes} and \eqref{eq:pted-cc-multipliers},
+Eqs. \eqref{eq:pted-cc-amplitudes} and \eqref{eq:pted-cc-multipliers},
 already shows that $\tamp{\mu_u}^{[0]} = 0, \,\forall u$ and
@@
 
+\subsection{Approximations to PTED}
+
+Perturbative expansions can also be developed starting from the
+approximations to \acrshort{PTED} presented in Section
+\ref{sec:pcm-cc-models}.
+For the \acrshort{PTE} scheme, the only change with respect to the
+\emph{in vacuo} theory as presented in Section \ref{sec:mbpt} is the use
+of the solvated Fock matrix.\autocite{Olivares_del_Valle1991-of,
+Angyan1995-co, Lipparini2009-io}
+
+In the \acrshort{PTE(S)} scheme
+
 \section[Approximate Coupled Cluster Methods]{
@@
 
-\todo[inline]{Finish up on CC2 and CC3.}
 In complete analogy to the discussion in Section
@@
 \acrshort*{CC} hierarchy.
-Starting from the \acrshort{CC2} and \acrshort{CC3} Lagrangians, Eqs.
-\eqref{eq:cc2-lagrangian} and \eqref{eq:cc3-lagrangian} respectively,
-we can write down the corresponding effective Lagrangians. The correct
-source term $\s_\mathrm{N}(\tamp{},\tbar{})$ must however be identified.
-For the \acrshort*{CC2} model, we start from its \acrshort*{CCSD}
-expression, Eq. \eqref{eq:source-ccsd}, and retain only terms up to first order:
-\begin{equation}
-  \begin{aligned}
-  \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCSD} &=
-  \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCS}
-  + \braket{\tbar{1} | \BCHfirst{\s}{T_2} | \mathrm{HF}} \\
-  &+ \braket{\tbar{2} | \BCHfirst{\s}{T_2} | \mathrm{HF}}
-  + \braket{\tbar{2} | \commutator{\commutator{\s}{T_1}}{T_2} | \mathrm{HF}}
-  \end{aligned}
-\end{equation}
-where the $T_1$-transformed operator was used.
+Starting from the \acrshort{CC2} Lagrangian, Eq.
+\eqref{eq:cc2-lagrangian}, we can write down the corresponding
+\acrshort{PTED} effective Lagrangian:\autocite{Sneskov2011-jm,
+Schwabe2012-cf}
+\begin{equation}\label{eq:pted-cc2-lagrangian}
+  \begin{aligned}
+  \lag(\tamp{}, \tbar{}, \p)_\mathrm{CC2}
+  &=
+  E_0
+  + \sum_{u=1}^{2}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
+  + \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{Q_2} | \mathrm{HF}} \\
+  &+ \braket{\tbar{1} |
+  \check{\Phi} + \commutator{\check{\Phi}}{Q_2}
+  | \mathrm{HF}}
+  + \braket{\tbar{2} | \check{\Phi} | \mathrm{HF}}
+  + \frac{1}{2}{}^t\p\V\p +
+  {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathrm{CC2}
+  \end{aligned}
+\end{equation}
+where the source term is formed according to Eq.
+\eqref{eq:cc2-expectation-value}
+The multipliers equations now become:
+\begin{subequations}\label{eq:pted-cc2-amplitudes}
+  \begin{align}
+    \denom{\mu_1}\tamp{\mu_1} &+
+    \braket{\mu_1 | \check{\Phi} +
+    \commutator{\check{\Phi}}{Q_2} | \mathrm{HF}}
+    + {}^t\p\braket{\mu_1 | \check{\s} +
+    \commutator{\check{\s}}{Q_2} | \mathrm{HF}}
+    = 0
+    \label{eq:pted-cc2-singles-amplitudes} \\
+    \denom{\mu_2}\tamp{\mu_2} &+ \braket{\mu_2 | \check{\Phi} | \mathrm{HF}}
+    + {}^t\p\braket{\mu_2 | \commutator{\check{\s}}{Q_2} | \mathrm{HF}}
+    = 0
+    \label{eq:pted-cc2-doubles-amplitudes}
+  \end{align}
+\end{subequations}
+while for the multipliers one has:
+\begin{subequations}\label{eq:pted-cc2-multipliers}
+  \begin{align}
+    \denom{\mu_1}\tbar{\mu_1} &+
+     \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}}
+   + \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_1}}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{Q_2}
+     | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{2} | \commutator{\check{\Phi}}{\cluster{\mu_1}} |
+   \mathrm{HF}}
+   + {}^t\p\braket{\mathrm{HF} |
+   \commutator{\check{\s}_\mathrm{N}}{\cluster{\mu_1}} | \mathrm{HF}}  \nonumber \\
+   &+ {}^t\p\braket{\tbar{1} | \commutator{\check{\s}}{\cluster{\mu_1}} | \mathrm{HF}}
+   + {}^t\p\braket{\tbar{2} | \commutator{\commutator{\check{\s}}{\cluster{\mu_1}}}{Q_2} | \mathrm{HF}}
+   = 0 \label{eq:pted-cc2-singles-multipliers} \\
+    \denom{\mu_2}\tbar{\mu_2} &+
+    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}}
+   + \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_2}}
+     | \mathrm{HF}} \nonumber \\
+   &+ {}^t\p\braket{\tbar{1} | \commutator{\check{\s}}{\cluster{\mu_2}} | \mathrm{HF}}
+   + {}^t\p\braket{\tbar{2} | \commutator{\check{\s}}{\cluster{\mu_2}} | \mathrm{HF}}
+     = 0 \label{eq:pted-cc2-doubles-multipliers}
+  \end{align}
+\end{subequations}
+Finally, the polarization is obtained by solving:
+\begin{equation}\label{eq:pted-cc2-polarization}
+  \V\p + \s_{\mathrm{N}}(\tamp{},\tbar{})_\mathrm{CC2} = 0
+\end{equation}
+
+The \acrshort{PTED}-\acrshort{CC3} model is obtained similarly. The
+source term is formed according to Eq. \eqref{eq:cc3-expectation-value}.
+Thus, adding the polarization functional to the Lagrangian
+\eqref{eq:cc3-lagrangian}, we obtain the effective Lagrangian:
+\begin{equation}\label{eq:pted-cc3-lagrangian}
+  \begin{aligned}
+  \lag(\tamp{}, \tbar{}, \p)_\mathrm{CC3}
+  &=
+  E_0
+  + \sum_{u=1}^{3}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
+  + \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
+  &+ \braket{\tbar{1} |
+  \check{\Phi} + \commutator{\check{\Phi}}{T_2} + \commutator{\check{\Phi}}{Q_3}
+  | \mathrm{HF}} \\
+  &+ \braket{\tbar{2} |
+  \check{\Phi} + \commutator{\check{\Phi}}{T_2}
+  + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{Q_3}
+  | \mathrm{HF}} \\
+  &+ \braket{\tbar{3} | \commutator{\check{\Phi}}{T_2} | \mathrm{HF}}
+  + \frac{1}{2}{}^t\p\V\p +
+  {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathrm{CC3}
+  \end{aligned}
+\end{equation}
+The polarization equation is straightforwardly:
+\begin{equation}\label{eq:pted-cc3-polarization}
+  \V\p + \s_{\mathrm{N}}(\tamp{},\tbar{})_\mathrm{CC3} = 0
+\end{equation}
+The amplitudes are determined by:
+\begin{subequations}\label{eq:pted-cc3-amplitudes}
+  \begin{align}
+    \denom{\mu_1}\tamp{\mu_1} &+ \braket{\mu_1 |
+    \check{\Phi} + \commutator{\check{\Phi}}{T_2}
+    + \commutator{\check{\Phi}}{Q_3}
+    | \mathrm{HF}} \nonumber \\
+    &+ {}^t\p\braket{\mu_1 | \check{\s} +
+    \commutator{\check{\s}}{T_2} | \mathrm{HF}}
+    = 0 \label{eq:pted-cc3-singles-amplitudes} \\
+    \denom{\mu_2}\tamp{\mu_2} &+ \braket{\mu_2 |
+    \check{\Phi} + \commutator{\check{\Phi}}{T_2}
+    + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2}
+    + \commutator{\check{\Phi}}{Q_3}
+    | \mathrm{HF}} \nonumber  \\
+    &+
+    {}^t\p\braket{\mu_2 | \commutator{\check{\s}}{T_2} +
+    \commutator{\check{\s}}{Q_3} | \mathrm{HF}}
+    = 0 \label{eq:pted-cc3-doubles-amplitudes} \\
+    \denom{\mu_3}\tamp{\mu_3} &+ \braket{\mu_3 |
+    \commutator{\check{\Phi}}{T_2}
+    | \mathrm{HF}}
+    + {}^t\p\braket{\mu_3 | \commutator{\check{\s}}{Q_3}
+     + \frac{1}{2}\commutator{\commutator{\check{\s}}{T_2}}{T_2}
+    | \mathrm{HF}}
+    = 0 \label{eq:pted-cc3-triples-amplitudes}
+    \end{align}
+\end{subequations}
+and finally the multipliers obey:
+\begin{subequations}\label{eq:pted-cc3-multipliers}
+  \begin{align}
+    \denom{\mu_1}\tbar{\mu_1} &+
+     \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}}
+   + \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_1}}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+     | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{2} |
+       \commutator{\check{\Phi}}{\cluster{\mu_1}}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{Q_3}
+   | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{3} |
+      \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+     | \mathrm{HF}}
+   + {}^t\p\braket{\mathrm{HF} |
+   \commutator{\check{\s}_\mathrm{N}}{\cluster{\mu_1}} | \mathrm{HF}}  \nonumber \\
+   &+ {}^t\p\braket{\tbar{1} | \commutator{\check{\s}}{\cluster{\mu_1}} | \mathrm{HF}}
+   + {}^t\p\braket{\tbar{2} |
+   \commutator{\commutator{\check{\s}}{\cluster{\mu_1}}}{Q_2} |
+   \mathrm{HF}} \nonumber \\
+   &+ {}^t\p\braket{\tbar{3} | \commutator{\commutator{\check{\s}}{\cluster{\mu_1}}}{Q_3} | \mathrm{HF}}
+     = 0 \label{eq:pted-cc3-singles-multipliers} \\
+    \denom{\mu_2}\tbar{\mu_2} &+
+    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}}
+   + \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_2}}
+     | \mathrm{HF}} \nonumber \\
+  &+ \braket{\tbar{2} |
+       \commutator{\check{\Phi}}{\cluster{\mu_2}}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_2}}}{T_2}
+     | \mathrm{HF}}
+  + \braket{\tbar{3} |
+       \commutator{\check{\Phi}}{\cluster{\mu_2}}
+     | \mathrm{HF}} \nonumber \\
+   &+ {}^t\p\braket{\tbar{1} | \commutator{\check{\s}}{\cluster{\mu_2}} | \mathrm{HF}}
+   + {}^t\p\braket{\tbar{2} | \commutator{\check{\s}}{\cluster{\mu_2}} | \mathrm{HF}}
+   \nonumber \\
+   &+ {}^t\p\braket{\tbar{3} |
+   \commutator{\commutator{\check{\s}}{\cluster{\mu_2}}}{T_2} | \mathrm{HF}}
+    = 0 \label{eq:pted-cc3-doubles-multipliers} \\
+    \denom{\mu_3}\tbar{\mu_3} &+
+    \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_3}}
+     | \mathrm{HF}}
+  + \braket{\tbar{2} |
+       \commutator{\check{\Phi}}{\cluster{\mu_3}}
+     | \mathrm{HF}} \nonumber \\
+   &+ {}^t\p\braket{\tbar{2} | \commutator{\check{\s}}{\cluster{\mu_3}} | \mathrm{HF}}
+   + {}^t\p\braket{\tbar{3} | \commutator{\check{\s}}{\cluster{\mu_3}} | \mathrm{HF}}
+    = 0 \label{eq:pted-cc3-triples-multipliers}
+  \end{align}
+\end{subequations}
+The \acrshort{PTES}, \acrshort{PTE(S)} and \acrshort{PTE} approximate
+schemes can be applied, \emph{mutatis mutandis}, also within the
+\acrshort{CC2} and \acrshort{CC3} models.\autocite{Caricato2011-tx, Schwabe2012-cf, Krause2016-ee}
+In particular, we note that the \acrshort{PTES} approximation is crucial
+in the implementation of \acrshort{PE}-\acrshort{CC2} presented by
+\citeauthor{Schwabe2012-cf} in \noparcite[ref.][]{Schwabe2012-cf}
+The resolution of the identity can in fact still be applied when
+assuming the \acrshort{PTES} approximation.
+Moreover, the approximate \acrshort{PE}-\acrshort{CC} Lagrangian of
+\citeauthor{Krause2016-ee} is essentially the \acrshort{PTE(S)}
+approximation for \acrshort{PE}-\acrshort{CC}.\autocite{Krause2016-ee}
 

### 2016-09-14 FrontBackmatter/Introduction.tex
First draft sent to L.F. and B.M.
Polish

@@
+\phantomsection
+\refstepcounter{dummy}
+\addcontentsline{toc}{chapter}{\tocEntry{Introduction}}
+
 \chapter*{Introduction}
-\refstepcounter{dummy}
-\phantomsection
-\addcontentsline{toc}{chapter}{\tocEntry{Introduction}}
 
@@
 \begin{epigraphs}
-\setlength\epigraphwidth{7.5cm}
-\qitem{
-plays with the bigness of his littleness
-
---- electrons deify one razorblade
-
-into a mountainrange; lenses extend
-
-unwish through curving wherewhen till unwish
-
-returns on its unself.
-    }{
-    --- \textsc{E. E. Cummings}}
-\setlength\epigraphwidth{7cm}
 \qitem{
@@
 \centering
-\scalebox{0.7}{\input{gfx/RQC-axes.tikz}}
+\includegraphics[width=.6\textwidth]{qc-axes.png}
 \caption[Pictorial depiction of the concept of \emph{model chemistries}.]{
@@
 three dimensions of quantum chemistry.~\autocite{Pople1999-gt, Saue2011-qg}
-The computational scaling of a model chemistry is
-$S=\textcolor{PMS1797}{y}(\textcolor{PMS2229}{x})f(\textcolor{PMS2229}{x})\textcolor{PMS130}{M}^{\textcolor{PMS2229}{x}}$.
+Reproduced from \noparcite[ref.][]{Norman2011-ad} with permission from the PCCP Owner Societies.
 }

### 2016-09-15 Chapters/ContinuumSolvationModels.tex
Finishing up response theory chapter
Almost done
Save the day's work
Polish

@@
 1936 and the first form of the \acrshort{PCM} entered the stage in
-1981.~\autocite{Miertus1981-mm}
+1981.\autocite{Miertus1981-mm}
 Section \ref{sec:CSM-why-how} will offer a heuristic "derivation" of
@@
 Still nowadays much research activity is expended on the
-\acrshort{PCM}.~\autocite{Tomasi2011-un, Mennucci2012-dv, Lipparini2016-mo}
+\acrshort{PCM}.\autocite{Tomasi2011-un, Mennucci2012-dv, Lipparini2016-mo}
 I will point out throughout the Chapter to developments outside our
@@
 Following \citeauthor{Tomasi2007-es}, we introduce the complete system
-Hamiltonian in the \acrshort{BO} approximation:~\autocite{Tomasi2004-dc,
+Hamiltonian in the \acrshort{BO} approximation:\autocite{Tomasi2004-dc,
 Tomasi2007-es}
@@
 trajectories or on the appropriate statistical
-ensemble.~\autocite{Hill1960-ql}
+ensemble.\autocite{Hill1960-ql}
 The need for ensemble averages leads us to the following observations:
@@
    the need for \emph{macroscopic} parameters, absent from the
-   microscopic Hamiltonian Eq.~\eqref{eq:solution-ham}, in carrying out
+   microscopic Hamiltonian Eq. \eqref{eq:solution-ham}, in carrying out
    statistical simulations.
@@
 Hamiltonian in which the environment degrees of freedom are properly
-averaged.~\autocite{Angyan1992-vo, Tapia1992-pu}
+averaged.\autocite{Angyan1992-vo, Tapia1992-pu}
 The solute-environment interaction term would then be replaced by
@@
 the errors induced by this \emph{charge penetration} are not so
-severe.~\autocite{Chipman2000-us, Cances2001-qs, Cances2001-qn}
+severe.\autocite{Chipman2000-us, Cances2001-qs, Cances2001-qn}
 
@@
 \end{subequations}
-Eqs.~\eqref{eq:trace-jump} and \eqref{eq:conormal-jump}
+Eqs. \eqref{eq:trace-jump} and \eqref{eq:conormal-jump}
 are the \emph{jump conditions}, expressed in terms of trace operators
@@
 conormal derivatives can be found in the excellent book by
-\citeauthor{Sauter2011-an}.~\autocite{Sauter2011-an}.
+\citeauthor{Sauter2011-an}.\autocite{Sauter2011-an}.
 The fundamental solutions, or Green's functions, for the elliptic
@@
 environments: homogeneous isotropic, ionic and anisotropic
-liquids,~\autocite{Cances1998-og}
-and systems with interfaces.~\autocite{Corni2002-dr, Frediani2004-er,
+liquids,\autocite{Cances1998-og}
+and systems with interfaces.\autocite{Corni2002-dr, Frediani2004-er,
 Delgado2013-kd, DiRemigio2016-nn}
@@
 where conductor boundary conditions are imposed. This leads to the
-well-known \gls{COSMO} equation:~\autocite{Klamt1993-mj, Cossi2003-xe}
+well-known \gls{COSMO} equation:\autocite{Klamt1993-mj, Cossi2003-xe}
 \begin{equation}
@@
 interlocking spheres. The radii are chosen from a suitable set, usually
-van der Waals radii.~\autocite{Bondi1964-dt, Mantina2009-hb}
+van der Waals radii.\autocite{Bondi1964-dt, Mantina2009-hb}
 If one imagines to model a solvent molecule as a rigid sphere,
@@
 Alternative definitions based on the isodensity surface of the molecular solute
-have been put forth.~\autocite{Foresman1996-wv}
+have been put forth.\autocite{Foresman1996-wv}
 
@@
 mathematical entity formed by the vector space of functions that are
-piecewise regular on any given interval in the mesh.~\autocite{Ern2004-oo}
+piecewise regular on any given interval in the mesh.\autocite{Ern2004-oo}
 Basically, given an element in the mesh, we "attach" a basis set of
@@
 achieved by modified triangulation procedures. In this thesis we used
-the GePol algorithm~\autocite{Pascual-Ahuir1987-uo, Pascual-Ahuir1990-lp,
+the GePol algorithm\autocite{Pascual-Ahuir1987-uo, Pascual-Ahuir1990-lp,
 Pomelli1998-qp, Pomelli2001-sj, Frediani2004-ua, Pomelli2007-wq}
 and the mesh generators of Harbrecht and Randrianarivony especially
-tailored for the wavelet \acrshort{BEM}.~\autocite{Harbrecht2009-no,
+tailored for the wavelet \acrshort{BEM}.\autocite{Harbrecht2009-no,
 Harbrecht2011-dk}
@@
 choice of \emph{trial} and \emph{test}
-spaces.~\autocite{Hackbusch1995-uq, Ern2004-oo}
+spaces.\autocite{Hackbusch1995-uq, Ern2004-oo}
 This step transforms the original integral equation into a system of
@@
 symmetrization procedures introducing, however, a degree of
-arbitrariness in the \acrshort{BEM} procedure.~\autocite{Lange2011-eu}
+arbitrariness in the \acrshort{BEM} procedure.\autocite{Lange2011-eu}
 
@@
 The factor $k=1.07$ is introduced to achieve a better precision on the
-diagonal elements.~\autocite{Tomasi2005-vm}
+diagonal elements.\autocite{Tomasi2005-vm}
 Expressions for $\mat{S}_\mathrm{e}$, $\mat{D}_\mathrm{e}$ and
 $\mat{D}^\dagger_\mathrm{e}$ depend on the specific outer environment
-and can be found in the literature.~\autocite{Corni2002-dr, Frediani2004-er,
+and can be found in the literature.\autocite{Corni2002-dr, Frediani2004-er,
 Tomasi2005-vm, Delgado2013-kd, DiRemigio2016-nn}
@@
 described by \citeauthor{Dahmen2006-pj} to build up the sparse system
-matrix.~\autocite{Harbrecht2004-uo, Harbrecht2006-ug, Dahmen2006-pj}
+matrix.\autocite{Harbrecht2004-uo, Harbrecht2006-ug, Dahmen2006-pj}
 Matrix compression is carried out in two rounds, \emph{a priori} and \emph{a
@@
 Numerous schemes have been proposed to solve or alleviate this issue.
-We here mention the TsLess~\autocite{Pomelli2004-lb} and
-DefPol~\autocite{Pomelli1998-ob, Pomelli1999-dc} approaches of
+We here mention the TsLess\autocite{Pomelli2004-lb} and
+DefPol\autocite{Pomelli1998-ob, Pomelli1999-dc} approaches of
 \citeauthor{Pomelli1998-ob} and the FIXPVA
-modification~\autocite{Su2009-em} to the GePol algorithm put forth by
+modification\autocite{Su2009-em} to the GePol algorithm put forth by
 \citeauthor{Su2009-em}.
@@
 which uses a smooth \acrshort{vdWS} and 3D spherical Gaussian basis
-functions.~\autocite{Scalmani2010-tw, York1999-xy}
+functions.\autocite{Scalmani2010-tw, York1999-xy}
 Gaussian blurring was also advocated by \citeauthor{Lange2010-jp} in
-their \gls{SWIG} method.~\autocite{Lange2010-jp, Lange2010-qo}
+their \gls{SWIG} method.\autocite{Lange2010-jp, Lange2010-qo}
 Both \acrshort{CSC} and \acrshort{SWIG} achieve continuous \acsp{PES} and smooth
@@
 \citeauthor{Cances2013-jg} by employing a domain decomposition
-method, guaranteed to converge to the exact solution.~\autocite{Quarteroni1999-jt}
+method, guaranteed to converge to the exact solution.\autocite{Quarteroni1999-jt}
 Moreover, for \acrshort{COSMO}, the formulation is embarrassingly
-parallel and has been implemented in a linearly scaling fashion.~\autocite{Cances2013-jg, Lipparini2013-cy, Lipparini2014-to,
+parallel and has been implemented in a linearly scaling fashion.\autocite{Cances2013-jg, Lipparini2013-cy, Lipparini2014-to,
 Lipparini2014-fq}
 Very recently the method has also been extended to general dielectric
-environments.~\autocite{Stamm2016-fv}
+environments.\autocite{Stamm2016-fv}
 
@@
  \quad
-  \PCM \coloneq \bi{R}_\infty^{-1}\bi{R}_\diel\bi{S}.
+  \PCM = \bi{R}_\infty^{-1}\bi{R}_\diel\bi{S}.
 \end{equation}
-yielding Eq.~\eqref{eq:isotropic-IEF} upon minimization:
+yielding Eq. \eqref{eq:isotropic-IEF} upon minimization:
 \begin{equation}\label{eq:pcm-stationarity}
@@
 
-The functionals in Eqs.~\eqref{eq:pcm-functional},
+The functionals in Eqs. \eqref{eq:pcm-functional},
 \eqref{eq:mm-functional} and \eqref{eq:pcm-mm-functional} all have a
@@
 \begin{equation}\label{eq:classical-HF}
-  \deriv{U_\mathrm{pol}}{\lambda} = \pderiv{U_\mathrm{pol}}{\lambda}
-  + \pderiv{U_\mathrm{pol}}{\sigma}\pderiv{\sigma}{\lambda}
-  + \pderiv{U_\mathrm{pol}}{\kappa}\pderiv{\kappa}{\lambda}
-  = \pderiv{U_\mathrm{pol}}{\lambda},
+  \deriv{U_\mathrm{pol}}{F} = \pderiv{U_\mathrm{pol}}{F}
+  + \pderiv{U_\mathrm{pol}}{\sigma}\pderiv{\sigma}{F}
+  + \pderiv{U_\mathrm{pol}}{\kappa}\pderiv{\kappa}{F}
+  = \pderiv{U_\mathrm{pol}}{F},
 \end{equation}
@@
 formalism:
-\begin{equation}
+\begin{equation}\label{eq:supermatrix-functional}
   U_\mathrm{pol} =
@@
 \begin{equation}\label{eq:scf-functional}
-  \mathcal{E}(\mat{C}) = \Tr \mat{h}\mat{D} + \frac{1}{2}\Tr
+  \mathcal{E}(\mat{C}) \eqtr \mat{h}\mat{D} + \frac{1}{2}
   \mat{G}^\gamma(\mat{D})\mat{D} + E_\mathrm{xc}[\rho(\vect{r})] +
@@
 \begin{equation}\label{eq:scf-lagrangian}
-  \mathcal{L}(\mat{C}, \mat{\epsilon}) = \mathcal{E}(\mat{D}) -
+  \mathcal{L}(\mat{C}, \mat{\epsilon}) = \mathcal{E}(\mat{C}) -
   \sum_{ij}\epsilon_{ji}(\braket{\phi_i|\phi_j} - \delta_{ij})
@@
     \mat{F}\mat{C} &= \mat{S}\mat{C}\mat{\epsilon} \\
-    \V\p &+ \Tr\mathbfup{s}\mat{D} = 0
+    \V\p &+ \Tr\mathbfup{s}\mat{D} = 0 \label{eq:polarization-stat}
   \end{align}

### 2016-09-15 Chapters/ContributionsSummary.tex
Polish

@@
 Performance and Sustainable Software for Molecular Simulation} workshop
-held at the 2015 Supercomputing Conference.~\autocite{SC15}
+held at the 2015 Supercomputing Conference.\autocite{SC15}
 Sections \ref{sec:relapcm}--\ref{sec:pcmopenrsp} contain very short
@@
 netherlands-escience, coderefinery, molssi} and ensuring
-scientific reproducibility.~\autocite{Hatton1994-us, Hatton1997-ml,
+scientific reproducibility.\autocite{Hatton1994-us, Hatton1997-ml,
 Hatton1997-bb, Ioannidis2005-aj, Merali2010-uy, Prinz2011-kz}
@@
 successfully used it in the development of the THE operating
-system in 1968,~\autocite{Dijkstra1968-zp} while
+system in 1968,\autocite{Dijkstra1968-zp} while
 \citeauthor{Parnas1972-im} advocated it as a best practice already in
-1972.~\autocite{Parnas1972-im}
+1972.\autocite{Parnas1972-im}
 Modularity is nothing more than a \emph{divide et impera} strategy
@@
 offered, effectively enforcing a programming style and standardisation of the
-functionality, eventually.~\autocite{Reddy2011-sd}
+functionality, eventually.\autocite{Reddy2011-sd}
 
@@
 cloud-based, free, code development services, such as hosted
-\acp{DVCS},~\autocite{github, gitlab} continuous
-integration,~\autocite{travis-ci, magnum-ci, appveyor-ci} code
-coverage analysis,~\autocite{coveralls} static and dynamic code
-analyses,~\autocite{coverity-scan} nightly regression testing, public
+\acp{DVCS},\autocite{github, gitlab} continuous
+integration,\autocite{travis-ci, magnum-ci, appveyor-ci} code
+coverage analysis,\autocite{coveralls} static and dynamic code
+analyses,\autocite{coverity-scan} nightly regression testing, public
 issue tracking, code review and so forth: adoption of a modern code
@@
 Parts of the Boost C++ libraries are used to alleviate these
-shortcomings.~\autocite{boost}
+shortcomings.\autocite{boost}
 Linear algebra operations are managed through the Eigen C++ template
-library.~\autocite{eigen}
+library.\autocite{eigen}
 The Python layer provides input reading and parsing facilities,
-as implemented by the GetKw library.~\autocite{libgetkw}
+as implemented by the GetKw library.\autocite{libgetkw}
 Cavity construction and discretization is handled either by the Fortan
@@
 of the boundary integral operators, leverages \gls{AD}. We use the Taylor library, a template C++
-implementation of \acrshort{AD}.~\autocite{Bartholomew-Biggs2000-db, libtaylor}
+implementation of \acrshort{AD}.\autocite{Bartholomew-Biggs2000-db, libtaylor}
 The adoption of \git as \acrshort{DVCS}, \cmake for cross-platform builds and
@@
 inheritance and the static polymorphism offered by template programming.
-~\autocite{Alexandrescu2001-bp, Sutter2004-nt, Langr2012-js}
+\autocite{Alexandrescu2001-bp, Sutter2004-nt, Langr2012-js}
 Preeminent use of composition over inheritance keeps the coupling
-between different modules as low as possible.~\autocite{Gamma1995-fd}
+between different modules as low as possible.\autocite{Gamma1995-fd}
 
 The library is released under the terms of the
-version 3 \gls{LGPL}, a standard open-source license.~\autocite{LGPLv3}
+version 3 \gls{LGPL}, a standard open-source license.\autocite{LGPLv3}
 We strongly believe that open-source software is the key to a larger
 impact for scientific software and a standard license means lower
-thresholds to adoption and third-party contributions.~\autocite{LGPLv3}
+thresholds to adoption and third-party contributions.\autocite{LGPLv3}
 The library serves as our development platform for the \acrshort{PCM}.
@@
 The \acrshort{API} is implemented in ISO C99 as a \emph{context-aware} set of
-functions.~\autocite{library-ronacher, context-api-example}
+functions.\autocite{library-ronacher, context-api-example}
 Upon initialization of the library, a \emph{context} object containing
@@
 \pcmsolver currently offers an implementation of the traditional collocation
-solvers for isotropic media,~\autocite{Tomasi2005-vm}
+solvers for isotropic media,\autocite{Tomasi2005-vm}
 together with some unique functionalities: the wavelet Galerkin solvers on smooth
-molecular surfaces,~\autocite{Weijo2010-hy, Bugeanu2015-tp} the Green’s
-functions for spherical diffuse interfaces~\autocite{DiRemigio2016-nn} and
+molecular surfaces,\autocite{Weijo2010-hy, Bugeanu2015-tp} the Green’s
+functions for spherical diffuse interfaces\autocite{DiRemigio2016-nn} and
 the delayed solvers for real-time \acrlong*{TDSCF}
-simulations.~\autocite{Corni2015-pe}
+simulations.\autocite{Corni2015-pe}
 
@@
 The first interface was implemented and released within the \DIRAC code
-and is described in \paper{I}.~\autocite{DIRAC15, DiRemigio2015-ou}.
+and is described in \paper{I}.\autocite{DIRAC15, DiRemigio2015-ou}.
 Interfaces to the \LSDALTON\autocite{LSDALTON16, Aidas2013-rp} and
 \DALTON\autocite{LSDALTON16, Aidas2013-rp} codes have followed, as
-described in \paper{II} and \paper{V}, respectively.~\autocite{Bugeanu2015-tp, pcm-openrsp}
-Details about the interface to \ReSpect~\autocite{ReSpect-3.5.0} can be
-found in \paper{IV}.~\autocite{pcm-respect}
-Coupling with \psicode~\autocite{Turney2012-de} has allowed us to implement the
+described in \paper{II} and \paper{V}, respectively.\autocite{Bugeanu2015-tp, pcm-openrsp}
+Details about the interface to \ReSpect\autocite{ReSpect-3.5.0} can be
+found in \paper{IV}.\autocite{pcm-respect}
+Coupling with \psicode\autocite{Turney2012-de} has allowed us to implement the
 \acrshort{PCM}-\acrshort{CC} developments described in this thesis.
-Finally, development of an interface with the KOALA code is underway.~\autocite{Hofener2014-ex, Hofener2016-qz}
+Finally, development of an interface with the KOALA code is underway.\autocite{Hofener2014-ex, Hofener2016-qz}
 
@@
 \end{equation}
-can be set up once the Green's functions $\Gi$ and $\Ge$ are known.~\autocite{Cances1998-og}
+can be set up once the Green's functions $\Gi$ and $\Ge$ are known.\autocite{Cances1998-og}
 \citeauthor{Frediani2004-er} showed that a numerical representation of the
@@
 they have been proved to be crucial for non homogeneous
-environments.~\autocite{Mozgawa2014-ad}
+environments.\autocite{Mozgawa2014-ad}
 Nevertheless our implementation represents a first significant step in the continuum treatment
@@
 in this paper we explored the calculation of first-order magnetic properties:
-\gls{EPR} and \gls{pNMR} parameters.~\autocite{Repisky2010-ls, Malkin2011-nm,
+\gls{EPR} and \gls{pNMR} parameters.\autocite{Repisky2010-ls, Malkin2011-nm,
 Komorovsky2013-xa, Cherry2016-ij}
@@
 Hence, \acrshort{EPR} and \acrshort{pNMR} parameters are formulated as expectation
-values, by virtue of the Hellmann--Feynman theorem.~\autocite{Konishi2009-zb,
+values, by virtue of the Hellmann--Feynman theorem.\autocite{Konishi2009-zb,
 Helgaker2000-tz}
@@
 allowing for spin polarization and thus granting facile access to the computation
-of spin-dependent properties.~\autocite{Dyall2007-tu}
+of spin-dependent properties.\autocite{Dyall2007-tu}
 The same modular programming strategy was adopted in crafting an interface between the
@@
 Our group has newly developed an open-ended methodology for the computation of
-\acrshort{SCF} response functions~\autocite{Thorvaldsen2008-sg, Ringholm2014-gx} and
-their residues.~\autocite{Friese2015-kb}
+\acrshort{SCF} response functions\autocite{Thorvaldsen2008-sg, Ringholm2014-gx} and
+their residues.\autocite{Friese2015-kb}
 These developments offer a route towards a synergistic experimental and
@@
 of molecular response properties when a quantum/classical polarizable continuum
-Hamiltonian is used.~\autocite{Thorvaldsen2008-sg, Lipparini2010-be}
+Hamiltonian is used.\autocite{Thorvaldsen2008-sg, Lipparini2010-be}
 I provided its implementation within the \DALTON code, by interfacing the
 \pcmsolver library and the open-ended \acrshort{SCF} response code of
-\citeauthor{Ringholm2014-gx}~\autocite{Ringholm2014-gx, Friese2015-kb}
+\citeauthor{Ringholm2014-gx}\autocite{Ringholm2014-gx, Friese2015-kb}
 I performed extensive testing of the code by comparing with previously
 published implementations of the \acrshort{PCM}-\acrshort{SCF} response functions within
-\DALTON.~\autocite{Cammi2003-qy, Frediani2005-nc, Ferrighi2010-pm}
+\DALTON.\autocite{Cammi2003-qy, Frediani2005-nc, Ferrighi2010-pm}
 Together with coauthors Maarten T.~P.~Beerepoot and Yann Cornaton, I carried out

### 2016-09-15 Chapters/MathematicalResults.tex
Polish

@@
 \chapter{Some Mathematical Results}\label{app:mathematical-results}
+
+\begin{epigraphs}
+\qitem{\textitalian{Non temerai i terrori della notte non temerai il terrore}}{--- \textsc{CCCP}}
+\end{epigraphs}
 
@@
 Coercivity implies that no eigenvalue of the linear operator associated to the
-bilinear form can be zero, hence its invertibility.~\autocite{Ern2004-oo}
+bilinear form can be zero, hence its invertibility.\autocite{Ern2004-oo}
 
@@
 I will show a detailed derivation of the \acrshort{IEF} equation
-for the \acrshort{PCM} transmission problem Eq.~\eqref{eq:transmission}:
+for the \acrshort{PCM} transmission problem Eq. \eqref{eq:transmission}:
 \begin{subequations}
@@
 \begin{lemma}[Properties of the boundary integral operators]
-  The integral operators introduced in Eqs.~\eqref{eq:S}-\eqref{eq:D-dagger}
-  enjoy the following properties~\autocite{Hsiao2008-xb, Sauter2011-an}:
+  The integral operators introduced in Eqs. \eqref{eq:S}--\eqref{eq:D-dagger}
+  enjoy the following properties\autocite{Hsiao2008-xb, Sauter2011-an}:
   \begin{enumerate}
@@
 \end{subequations}
-The final ingredient is the \gls{DtN} map, which can be derived by employing Eq.~\eqref{eq:stat3}
+The final ingredient is the \gls{DtN} map, which can be derived by employing Eq. \eqref{eq:stat3}
 to the Newton potential:
@@
 \emph{linear} and continuous in $H_0^1(\Omegai)$.
-We rewrite Eq.~\eqref{eq:poisson-weak} in the abstract form:
+We rewrite Eq. \eqref{eq:poisson-weak} in the abstract form:
 \begin{equation}
@@
 of the approximation setting, $\inf$-$\sup$ condition, Céa's Lemma
-and generalized $\inf$-$\sup$ condition.~\autocite{Ern2004-oo}
+and generalized $\inf$-$\sup$ condition.\autocite{Ern2004-oo}
 
@@
 expansions truncate at the third term. Moreover, normal-ordering of the
-operators has been introduced.~\autocite{Crawford2000-ey, Shavitt2009-mr}
+operators has been introduced.\autocite{Crawford2000-ey, Shavitt2009-mr}
 This is equivalent to applying a shift to the correlation part of the

### 2016-09-15 Chapters/QuantumMechanics.tex
Polish

@@
 As noted by \citeauthor{Saue2011-qg}, the same general expression for
-the clamped-nuclei Hamiltonian Eq.~\eqref{eq:fq-hamiltonian} is valid in
+the clamped-nuclei Hamiltonian Eq. \eqref{eq:fq-hamiltonian} is valid in
 the nonrelativistic and relativistic no-pair
@@
 \end{subequations}
-The second-quantized Hamiltonian Eq.~\eqref{eq:sq-hamiltonian} is thus a
+The second-quantized Hamiltonian Eq. \eqref{eq:sq-hamiltonian} is thus a
 \emph{projected} operator, exact only within the subspace determined by
-the chosen one-particle set.~\autocite{Gross1991-hi, Helgaker2000-tz}
+the chosen one-particle set.\autocite{Gross1991-hi, Helgaker2000-tz}
 
@@
 \begin{equation}
-  T = \sum_{p=1}^\mathcal{M} T_p = \sum_{p=1}^\mathcal{M}\sum_{\mu_p}\tamp{\mu_p}\cluster{\mu_p}
+  T = \sum_{u=1}^\mathcal{M} T_u = \sum_{u=1}^\mathcal{M}\sum_{\mu_u}\tamp{\mu_u}\cluster{\mu_u}
 \end{equation}
@@
 which stems directly from the exponential nature of the wave
-operator.~\autocite{Shavitt2009-mr, Helgaker2000-tz}
+operator.\autocite{Shavitt2009-mr, Helgaker2000-tz}
 

### 2016-09-15 Chapters/SolvationAndCorrelation.tex
Polish

@@
 \qitem{Potius sero quam numquam}{--- \textsc{Titus Livius}, \textit{Ab Urbe Condita}}
-\qitem{\textitalian{Non temerai i terrori della notte non temerai il terrore}}{--- \textsc{CCCP}}
 \end{epigraphs}
@@
 
-Iterative and noniterative approximations to the \acrshort{CCSD} and \acrshort{CCSDT} models will
-be the subject of Section \ref{sec:cc-approximate-quantum-classical} where the equations
-governing the \acrshort{CC2} and \acrshort{CC3} models will be derived.
+Iterative and noniterative approximations to the \acrshort{CCSD} and
+\acrshort{CCSDT} models will be the subject of Section
+\ref{sec:cc-approximate-quantum-classical} where the equations governing
+the \acrshort{CC2} and \acrshort{CC3} models will be derived.
 Eventually, I will describe schemes for the noniterative inclusion of
-the effect of connected triples excitations on top of
-the \acrshort{CCSD} wave function. The symmetric and asymmetric [T] and (T)
-corrections will be derived.
+the effect of connected triples excitations on top of the
+\acrshort{CCSD} wave function.
+The symmetric and asymmetric (T) corrections will be derived.
 
@@
 environment.
-This can be readily obtained~\autocite{Lipparini2016-mo} by coupling the
-usual Lagrangian in Eq.~\eqref{eq:cc-lagrangian} with the classical
-polarization functional in Eq.~\eqref{eq:pcm-mm-functional}.
+This can be readily obtained\autocite{Lipparini2016-mo} by coupling the
+usual Lagrangian in Eq. \eqref{eq:cc-lagrangian} with the classical
+polarization functional in Eq. \eqref{eq:pcm-mm-functional}.
 Care must however be taken in the definition of the classical
@@
 expansion truncates at the third term. Moreover, normal-ordering of the
-operators has been introduced.~\autocite{Crawford2000-ey, Shavitt2009-mr}
+operators has been introduced.\autocite{Crawford2000-ey, Shavitt2009-mr}
 This is equivalent to applying a shift to the correlation part of the
@@
 
-Differentiation of the effective Lagrangian in Eq.~\eqref{eq:pted-cc}
+Differentiation of the effective Lagrangian in Eq. \eqref{eq:pted-cc}
 with respect to the variational parameters $\tamp{\mu_i}$,
@@
 This requires a proper macroiteration/microiteration self-consistency
-scheme for its implementation.~\autocite{Cammi2009-gu, Caricato2010-hx}
+scheme for its implementation.\autocite{Cammi2009-gu, Caricato2010-hx}
 As a consequence, a single-point \acrshort{PTED}-\acrshort{CC} will suffer from a
@@
 Lagrangian and the governing equations by breaking or weakening the coupling.
-While Eq.~\eqref{eq:pted-cc-polarization} is \emph{directly} coupled to
+While Eq. \eqref{eq:pted-cc-polarization} is \emph{directly} coupled to
 the multipliers equation, since the source terms directly depends on
@@
 The use of a \acrshort{CC} ansatz for the wave function ensures term-by-term
-size-extensivity of the energy at all orders.~\autocite{Helgaker2000-tz}
+size-extensivity of the energy at all orders.\autocite{Helgaker2000-tz}
 The fluctuation potential $\Phi$ is assumed as the perturbation, hence
@@
 \end{equation}
-This clearly shows that Eq.~\eqref{eq:1st-order-amp}
-and Eq.~\eqref{eq:1st-order-mult} are complex conjugates. Under the
+This clearly shows that Eq. \eqref{eq:1st-order-amp}
+and Eq. \eqref{eq:1st-order-mult} are complex conjugates. Under the
 assumption of real wave functions we can also conclude that:
@@
 Similarly to the exposition in Chapter \ref{ch:QM}, these expressions
-will form the basis to the development of methods for the approximate
+will form the basis for the development of methods for the approximate
 beyond \acrshort*{CCSD} including connected triples.
@@
 only includes singles and doubles in its excitation manifold and is thus
-correct up to third order in perturbation theory:
-\begin{subequations}
-  \begin{align}
-    \tamp{\mu_i}^{*} &= \tamp{\mu_i}^{[1]} + \tamp{\mu_i}^{[2]} + O(3)
-    \label{eq:tamp-star} \\
-    \tbar{\mu_i}^{*} &= \tbar{\mu_i}^{[1]} + \tbar{\mu_i}^{[2]} + O(3)
-    \label{eq:tbar-star} \\
-    \p^{*} &= \p^{[1]} + \p^{[2]} + O(3)
-    \label{eq:p-star}
-  \end{align}
-\end{subequations}
-Here and in the following, the \acrshort{CCSD} converged parameters will
-be denoted with a $*$ superscript.
-
+correct up to third order in perturbation theory.
 Already in fourth order connected triples make their appearance and any
@@
 \end{subequations}
-The \acrshort{PTES}, \acrshort{PTE(S)} and \acrshort{PTE} approximate
-schemes can be applied, \emph{mutatis mutandis}, also within the
-\acrshort{CC2} and \acrshort{CC3} models.\autocite{Caricato2011-tx, Schwabe2012-cf, Krause2016-ee}
-In particular, we note that the \acrshort{PTES} approximation is crucial
-in the implementation of \acrshort{PE}-\acrshort{CC2} presented by
-\citeauthor{Schwabe2012-cf} in \noparcite[ref.][]{Schwabe2012-cf}
-The resolution of the identity can in fact still be applied when
-assuming the \acrshort{PTES} approximation.
-Moreover, the approximate \acrshort{PE}-\acrshort{CC} Lagrangian of
-\citeauthor{Krause2016-ee} is essentially the \acrshort{PTE(S)}
-approximation for \acrshort{PE}-\acrshort{CC}.\autocite{Krause2016-ee}
 
@@
 \end{alignat}
-Inserting the converged \acrshort{CCSD} parameters yields the following
-expressions:
+Inserting the converged \acrshort{CCSD} parameters, denoted with a $*$
+superscript, yields the following expressions:
 \begin{subequations}\label{eq:g-parentheses-t}
@@
 
-Using Eqs. \eqref{eq:tamp-star}--\eqref{eq:p-star} one can expand the
-energy correction as follows:
+The converged \acrshort{PTED}-\acrshort{PTED} parameters are correct up
+to second order in the fluctuation potential:
+\begin{subequations}
+  \begin{align}
+    \tamp{\mu_i}^{*} &= \tamp{\mu_i}^{[1]} + \tamp{\mu_i}^{[2]} + O(3)
+    \label{eq:tamp-star} \\
+    \tbar{\mu_i}^{*} &= \tbar{\mu_i}^{[1]} + \tbar{\mu_i}^{[2]} + O(3)
+    \label{eq:tbar-star} \\
+    \p^{*} &= \p^{[1]} + \p^{[2]} + O(3)
+    \label{eq:p-star}
+  \end{align}
+\end{subequations}
+Using the equations above one can expand the energy correction as
+follows:
 \begin{equation}
@@
 \end{equation}
+
+\subsection{Approximate Coupled Cluster Methods with Approximate
+Quantum/Classical Coupling}\label{sec:approximate-everything}
+
+The \acrshort{PTES}, \acrshort{PTE(S)} and \acrshort{PTE} approximate
+schemes can be applied, \emph{mutatis mutandis}, also within the
+\acrshort{CC2} and \acrshort{CC3} models.\autocite{Caricato2011-tx, Schwabe2012-cf, Krause2016-ee}
+In particular, we note that the \acrshort{PTES} approximation is crucial
+in the implementation of \acrshort{PE}-\acrshort{CC2} presented by
+\citeauthor{Schwabe2012-cf} in \noparcite[ref.][]{Schwabe2012-cf}
+The resolution of the identity can in fact still be applied when
+assuming the \acrshort{PTES} approximation.
+Moreover, the approximate \acrshort{PE}-\acrshort{CC} Lagrangian of
+\citeauthor{Krause2016-ee} is essentially the \acrshort{PTE(S)}
+approximation for \acrshort{PE}-\acrshort{CC}.\autocite{Krause2016-ee}

### 2016-09-15 FrontBackmatter/Introduction.tex
Polish

@@
 Pictorial depiction of the concept of \emph{model chemistries} as the
-three dimensions of quantum chemistry.~\autocite{Pople1999-gt, Saue2011-qg}
+three dimensions of quantum chemistry.\autocite{Pople1999-gt, Saue2011-qg}
 Reproduced from \noparcite[ref.][]{Norman2011-ad} with permission from the PCCP Owner Societies.
@@
 Chemistry can be largely considered a \emph{wet} science: almost always
-chemical phenomena happen in a liquid environment.~\autocite{Reichardt2010-le}
+chemical phenomena happen in a liquid environment.\autocite{Reichardt2010-le}
 We hereby define a "solution", or more generally an "environment", as
 a system where the number of solvent molecules exceeds by far the number
-of solute molecules.~\autocite{Tomasi2004-dc, Tomasi2007-es}
+of solute molecules.\autocite{Tomasi2004-dc, Tomasi2007-es}
 It is then clear that theoretical and computational approaches to such a
@@
 As is well known from statistical mechanics, microscopic detail cannot
-account for the macroscopic behaviour.~\autocite{Hill1960-ql,
+account for the macroscopic behaviour.\autocite{Hill1960-ql,
 Hansen2013-io}
@@
 understanding reality and predicting new and exciting
-phenomena and properties.~\autocite{Anderson1972-ai, Winsberg2010-sy, Kovac2011-ew}
+phenomena and properties.\autocite{Anderson1972-ai, Winsberg2010-sy, Kovac2011-ew}
 One of the earlier attempts at tackling the problem of solvation is due
@@
 that has had a lasting impact and informs much of the developments that
-will be presented in this thesis.~\autocite{Onsager1936-wf}
+will be presented in this thesis.\autocite{Onsager1936-wf}
 
@@
   modifications of externally applied fields subtly influence molecular
-  responses.~\autocite{Cammi1998-jp, Pipolo2014-sd}
+  responses.\autocite{Cammi1998-jp, Pipolo2014-sd}
 \item[Dynamic.]
@@
 Both can be classified as \emph{multiscale} (or \emph{focused})
-models~\autocite{Nobel2013} and hinge on the same idea: treat different
+models\autocite{Nobel2013} and hinge on the same idea: treat different
 parts of the system with different methods and couple these methods by
@@
    This is either achieved by a cheaper quantum mechanical
-   method~\autocite{Vreven2006-gx} or by \gls{MM}.~\autocite{Senn2009-sk}
+   method\autocite{Vreven2006-gx} or by \gls{MM}.\autocite{Senn2009-sk}
    In the latter approach, commonly dubbed \acrshort{QM}/\acrshort{MM}, the \acrshort{MM}
-   region can either be polarizable~\autocite{Mennucci2013-go,
+   region can either be polarizable\autocite{Mennucci2013-go,
    Olsen2010-wa, Lipparini2011-rd} or non
@@
    The effect of this continuum is described, classically, \emph{via}
-   its bulk properties.~\autocite{Onsager1936-wf, Miertus1981-mm}
+   its bulk properties.\autocite{Onsager1936-wf, Miertus1981-mm}
 \end{itemize}

### 2016-09-15 Chapters/SolvationAndCorrelation.tex
More polish + add a couple of TODOs

@@
 and are determined by the quantum/classical polarizable coupling term.
+\todo[inline]{This expression is consistent with electrostatics and
+perturbation theory. Comment how this wouldn't be the case if one uses
+the nonlinear Hamiltonian formulation for the quantum/classical coupling.}
 
@@
 \subsection{Approximations to PTED}
+
+\todo[inline]{Develop MBPT for the approximate Lagrangians}
 
@@
 
+\todo[inline]{Give the CC2, CC3 and the (T) corrections (symmetric and
+asymmetric) equations.}
+
 The \acrshort{PTES}, \acrshort{PTE(S)} and \acrshort{PTE} approximate

### 2016-09-16 Chapters/QuantumMechanics.tex
Only acknowledgements missing

@@
 order can be obtained from amplitudes and multipliers up to second
-order. Since the triples first appear to second order, the \gls{CCSD}
+order:
+\begin{subequations}\label{eq:mbpt-energies}
+  \begin{align}
+    E^{[1]}
+    &= \braket{\mathrm{HF} | \Phi | \mathrm{HF}} \label{eq:mbpt-1st} \\
+    E^{[2]}
+    &= \braket{\mathrm{HF} | \commutator{\Phi}{T_2^{[1]}} | \mathrm{HF}} \label{eq:mbpt-2nd} \\
+    E^{[3]}
+    &= \braket{\tbar{2}^{[1]} | \commutator{\Phi}{T_2^{[1]}} | \mathrm{HF}} \label{eq:mbpt-3rd} \\
+    E^{[4]} &= \braket{\tbar{2}^{[1]} | \commutator{\Phi}{T^{[2]}} |
+    \mathrm{HF} }
+    +
+    \braket{\tbar{2}^{[1]} | \BCHsecond{\Phi}{T_2^{[1]}}|
+    \mathrm{HF} } \label{eq:mbpt-4th} \\
+    E^{[5]} &=
+    \braket{\mathrm{HF} | \BCHsecond{\Phi}{T^{[2]}}|\mathrm{HF} }
+    + \braket{\tbar{2}^{[1]} |
+    \commutator{\commutator{\Phi}{T^{[2]}}}{T_2^{[1]}}
+    | \mathrm{HF} } \nonumber \\
+    &+ \sum_{u=1}^\mathcal{M}
+    \braket{\tbar{u}^{[2]} | \BCHfirst{\Phi}{T^{[2]}}|
+    \mathrm{HF} }
+    +\sum_{u=1}^\mathcal{M}
+    \braket{\tbar{u}^{[2]} | \BCHsecond{\Phi}{T_2^{[1]}}|
+    \mathrm{HF} } \label{eq:mbpt-5th}
+  \end{align}
+\end{subequations}
+
+Since the triples first appear to second order, the \gls{CCSD}
 energy is correct to third order in the fluctuation potential.

### 2016-09-16 Chapters/SolvationAndCorrelation.tex
Only acknowledgements missing

--- 2016-09-15 Chapters/SolvationAndCorrelation.tex
@@
    \denom{\mu_q}\tamp{\mu_q} + \braket{\mu_q | \overline{\Phi} | \mathrm{HF}}
-   + {}^t\p\braket{\mu_q | \check{\s} | \mathrm{HF}}
+   + {}^t\p\braket{\mu_q | \check{\s} | \mathrm{HF}}\delta_{\mu_q\mu_1}
              = 0 \\
@@
     &+
-    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1} \nonumber \\
-    &+
-    {}^t\p\sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
-    \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}
-    +
-    {}^t\p\sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
-    \commutator{\commutator{\s}{T}}{\cluster{\mu_q}} | \mathrm{HF}}
+    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
+    +
+    {}^t\p\braket{\tbar{1} |
+    \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
+     \nonumber \\
+    &+
+    {}^t\p\braket{\tbar{1} |
+    \commutator{\commutator{\s}{T_1}}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
              = 0 \\
@@
    {}^{\mathrm{PTES}}\tampEq{\mu_q}(\tamp{}, \tbar{}, \p)  &=
-   \denom{\mu_i}\tamp{\mu_q} + \braket{\mu_q | \overline{\Phi} | \mathrm{HF}}
-   + {}^t\p\braket{\mu_q | \check{\s} | \mathrm{HF}}
+   \denom{\mu_q}\tamp{\mu_q} + \braket{\mu_q | \overline{\Phi} | \mathrm{HF}}
+   + {}^t\p\braket{\mu_q | \check{\s} | \mathrm{HF}}\delta_{\mu_q\mu_1}
              = 0 \label{eq:ptes-cc-amplitudes} \\
@@
 multiplier-independent part:
-\begin{equation}
+\begin{equation}\label{eq:pte-s-source}
   \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCS}
@@
     &+
-    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_q}} |
-    \mathrm{HF}}\delta_{\mu_q\mu_1} \nonumber \\
-    &+
-    {}^t\p\sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
-    \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}
-    +
-    {}^t\p\sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
-    \commutator{\commutator{\s}{T}}{\cluster{\mu_q}} | \mathrm{HF}}
-             = 0 \label{eq:ptes-cc-multipliers}\\
+    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
+    +
+    {}^t\p\braket{\tbar{1} |
+    \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
+     \nonumber \\
+    &+
+    {}^t\p\braket{\tbar{1} |
+    \commutator{\commutator{\s}{T_1}}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
+             = 0 \label{eq:ptes-cc-multipliers} \\
     {}^{\mathrm{PTES}}\Omega_\p(\tamp{}, \tbar{}, \p)
@@
     \V\p + \braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}} = 0
+    \label{eq:pte-s-polarization}
   \end{align}
@@
 \acrshort{PTED} model with a computational cost comparable to that of
-\emph{in vacuo} \acrshort{CC} theory.
+\emph{in vacuo} \acrshort{CC}
+theory.\autocite{Olivares_del_Valle1991-of, Begue2005-qn,
+Hasegawa2007-jt}
 
@@
   + {}^t\p^{[1]}\braket{\mu_1 | \s | \mathrm{HF}}
-  = 0 \\
+  = 0 \label{eq:1st-order-amp-singles} \\
   \Omega_{\mu_2}^{[1]} &= \denom{\mu_2}\tamp{\mu_2}^{[1]}
@@
 thus have:
-\begin{equation}\label{eq:first-order-G}
+\begin{equation}\label{eq:pted-mbpt-1st}
   G^{[1]} = \braket{\mathrm{HF} | \Phi | \mathrm{HF}}
@@
 \begin{equation}
-  \s_{\mathrm{N}}^{[m]}(\tamp{}^{[p]}, \tbar{}^{[q]})
+  \s_{\mathrm{N}}^{[m]}(\tamp{}^{[p]}, \tbar{}^{[q]}),
 \end{equation}
@@
 amplitudes and polarization, while using zeroth order multipliers:
-\begin{equation}\label{eq:second-order-G}
+\begin{equation}\label{eq:pted-mbpt-2nd}
   \begin{aligned}
@@
 The polarization terms were eliminated by virtue of Eq.
-\eqref{eq:1st-order-pol}. Formally, $G^{[2]}$ has the same expression as
-\emph{in vacuo}. However, the first-order singles amplitudes are nonzero
-and are determined by the quantum/classical polarizable coupling term.
+\eqref{eq:1st-order-pol}. Thus, at least formally, $G^{[2]}$ has the same expression as
+\emph{in vacuo} and is consistent with \emph{both} electrostatics and
+perturbation theory.
+We remark that this would not be the case if one had based the
+development of \acrshort{MBPT} on the use of a nonlinear
+Hamiltonian.\autocite{Angyan1995-co}
 
@@
 amplitudes, multipliers and polarization:
-\begin{equation}\label{eq:third-order-G}
+\begin{equation}\label{eq:pted-mbpt-3rd}
   \begin{aligned}
@@
 order amplitudes and polarization, while using first order multipliers:
-\begin{equation}\label{eq:fourth-order-G}
+\begin{equation}\label{eq:pted-mbpt-4th}
   \begin{aligned}
@@
 amplitudes, multipliers and polarization:
-\begin{equation}\label{eq:fifth-order-G}
+\begin{equation}\label{eq:pted-mbpt-5th}
   \begin{aligned}
@@
 
-In the \acrshort{PTE(S)} scheme
+In the \acrshort{PTE(S)} scheme, the amplitudes equations are uncoupled
+from the multipliers and polarization equations. \acrshort{MBPT} for
+this approximate model can be developed similarly to that in the
+\acrshort{PTE} scheme.
+The first and second order parameters are determined by:
+\begin{subequations}\label{eq:pte-s-mbpt}
+  \begin{align}
+  \denom{\mu_2}\tamp{\mu_2}^{[1]} &=
+  - \braket{\mu_2 | \Phi | \mathrm{HF}}
+  =
+  - \braket{\mathrm{HF} | \commutator{\Phi}{\cluster{\mu_2}} | \mathrm{HF}}
+  =
+  \denom{\mu_2}\tbar{\mu_2}^{[1]}
+  \\
+    \denom{\mu_q}\tamp{\mu_q}^{[2]}
+    &= -
+      \braket{\mu_q | \commutator{\Phi}{T_2^{[1]}} | \mathrm{HF}}
+  \\
+    \denom{\mu_q}\tbar{\mu_q}^{[2]}
+    &=
+    - \left(
+      \braket{\tbar{2}^{[1]} | \commutator{\Phi}{\cluster{\mu_q}} | \mathrm{HF}}
+      + {}^t\p^{[2]}\braket{\mathrm{HF} |
+      \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
+    \right)
+    = 0
+    \label{eq:pte-s-2nd-order-mult}
+  \\
+  \V\p^{[2]} &+ \braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}}| \mathrm{HF}} = 0
+  \label{eq:pte-s-2nd-order-pol}
+  \end{align}
+\end{subequations}
+Notice that the first order singles amplitudes and multipliers are now
+zero, as is the first order polarization. The second order singles
+multipliers are coupled to the second order polarization. Notice however
+that $\p^{[2]}$ only depends on $T_1^{[2]}$ and can thus be computed
+on-the-fly, without iterating between Eqs.
+\eqref{eq:pte-s-2nd-order-mult} and \eqref{eq:pte-s-2nd-order-pol}
+Starting from Eqs. \eqref{eq:mbpt-1st}--\eqref{eq:mbpt-5th}, the free energy
+corrections in the \acrshort{PTE(S)} scheme are:
+\begin{subequations}
+  \begin{align}
+    G^{[1]} &= E^{[1]} \label{eq:pte-s-mbpt-1st} \\
+    G^{[2]} &= E^{[2]} \label{eq:pte-s-mbpt-2nd} \\
+    G^{[3]} &= E^{[3]} \label{eq:pte-s-mbpt-3rd} \\
+    G^{[4]} &= E^{[4]} + \frac{1}{2}
+    {}^t\p^{[2]}\braket{\mathrm{HF} | \BCHfirst{\s}{T_1^{[2]}}
+    |\mathrm{HF}} \label{eq:pte-s-mbpt-4th} \\
+    G^{[5]} &= E^{[5]} \label{eq:pte-s-mbpt-5th}
+  \end{align}
+\end{subequations}
+
+In the \acrshort{PTES} scheme, two different polarizations are
+introduced to achieve an approximate decoupling of the governing
+equations. \acrshort{MBPT} can be developed by considering the
+Lagrangian \eqref{eq:ptes-cc} to form the energy corrections, Eq.
+\eqref{eq:ptes-amp} for the order analysis of the amplitudes, Eq.
+\eqref{eq:ptes-mult} for the order analysis of the multipliers.
+The order analysis of the polarization will be developed based on
+Eq. \eqref{eq:ptes-cc-polarization-mult}, \ie~the polarization coupled
+to the multipliers.
+To first order:
+\begin{subequations}\label{eq:ptes-1st-mbpt}
+  \begin{align}
+  \denom{\mu_2}\tamp{\mu_2}^{[1]} &=
+  - \braket{\mu_2 | \Phi | \mathrm{HF}}
+  =
+  - \braket{\mathrm{HF} | \commutator{\Phi}{\cluster{\mu_2}} | \mathrm{HF}}
+  =
+  \denom{\mu_2}\tbar{\mu_2}^{[1]}
+  \\
+  \denom{\mu_1}\tbar{\mu_1}^{[1]} &=
+  - {}^t\p^{[1]}\braket{\mathrm{HF} |
+      \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
+  \\
+  \V\p^{[1]} &+ \braket{\tbar{1}^{[1]} | \s | \mathrm{HF}} = 0
+  \label{eq:ptes-1st-order-pol}
+  \end{align}
+\end{subequations}
+while to second order:
+\begin{subequations}\label{eq:ptes-2nd-mbpt}
+  \begin{align}
+    \denom{\mu_q}\tamp{\mu_q}^{[2]}
+    &= -
+      \braket{\mu_q | \commutator{\Phi}{T_2^{[1]}} | \mathrm{HF}}
+  \\
+    \denom{\mu_q}\tbar{\mu_q}^{[2]}
+    &=
+    - \braket{\tbar{2}^{[1]} | \commutator{\Phi}{\cluster{\mu_q}} | \mathrm{HF}}
+    - {}^t\p^{[2]}\braket{\mathrm{HF} |
+      \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
+      \nonumber \\
+    &- {}^t\p^{[1]}\braket{\tbar{1}^{[1]} | \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}
+    = 0
+    \label{eq:ptes-2nd-order-mult}
+  \\
+  \V\p^{[2]} &+ \braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}}| \mathrm{HF}}
+  + \braket{\tbar{1}^{[2]} | \s | \mathrm{HF}} = 0
+  \label{eq:ptes-2nd-order-pol}
+  \end{align}
+\end{subequations}
+We can now develop the free energy corrections in the \acrshort{PTES}
+scheme based on Eqs. \eqref{eq:mbpt-1st}--\eqref{eq:mbpt-5th}
+\begin{subequations}
+  \begin{align}
+    G^{[1]} &= E^{[1]} \label{eq:ptes-mbpt-1st} \\
+    G^{[2]} &= E^{[2]} \label{eq:ptes-mbpt-2nd} \\
+    G^{[3]} &= E^{[3]} \label{eq:ptes-mbpt-3rd} \\
+    G^{[4]} &= E^{[4]} + \frac{1}{2}
+    {}^t\p^{[2]}\braket{\mathrm{HF} | \BCHfirst{\s}{T_1^{[2]}}
+    |\mathrm{HF}} \label{eq:ptes-mbpt-4th} \\
+    G^{[5]} &= E^{[5]}
+    + {}^t\p^{[2]}\braket{\tbar{1}^{[1]} | \BCHfirst{\s}{T_1^{[2]}}
+    \label{eq:ptes-mbpt-5th}
+    |\mathrm{HF}}
+  \end{align}
+\end{subequations}
 
@@
 
-\subsection{Approximate Coupled Cluster Methods with Approximate
-Quantum/Classical Coupling}\label{sec:approximate-everything}
-
-The \acrshort{PTES}, \acrshort{PTE(S)} and \acrshort{PTE} approximate
-schemes can be applied, \emph{mutatis mutandis}, also within the
-\acrshort{CC2} and \acrshort{CC3} models.\autocite{Caricato2011-tx, Schwabe2012-cf, Krause2016-ee}
-In particular, we note that the \acrshort{PTES} approximation is crucial
+\section[Approximate CC and Approximate Couplings]{Approximate Coupled Cluster Methods and Approximate
+Quantum/Classical Coupling Schemes}\label{sec:approximate-everything}
+
+The iterative \acrshort{CC2} and \acrshort{CC3} and the noniterative
+$\Lambda\text{CCSD}(\text{T})$ and $\text{CCSD}(\text{T})$ methods can
+be formulated, \emph{mutatis mutandis}, also within the approximate
+quantum/classical coupling schemes mentioned previously.\autocite{Caricato2011-tx, Schwabe2012-cf, Krause2016-ee}
+
+Let us first consider the noniterative models for the inclusion of
+connected triples.
+The \acrshort{PTE} scheme is particularly simple: the equations are
+unchanged with respect to vacuum theory, the only change being in the
+definition of the \glspl{MO} and the Fock matrix.\autocite{Cammi2009-gu,
+Begue2005-qn, Hasegawa2007-jt}
+Comparing the fourth and fifth order \acrshort{MBPT} energy corrections
+in the \acrshort{PTE(S)}, Eqs. \eqref{eq:pte-s-mbpt-1st}--\eqref{eq:pte-s-mbpt-5th}, and
+\acrshort{PTES}, Eqs. \eqref{eq:ptes-mbpt-1st}--\eqref{eq:ptes-mbpt-5th}, schemes to
+those obtained for the \acrshort{PTED} scheme, Eqs.
+\eqref{eq:pted-mbpt-1st}--\eqref{eq:pted-mbpt-5th},
+we can see that there are no contributions from the triples to the
+quantum/classical terms involving the polarization.
+Thus, also for \acrshort{PTE(S)} and \acrshort{PTES} the asymmetric and
+symmetric triples corrections are formally identical to the vacuum
+expression given in equations \eqref{eq:parentheses-t}
+The converged \acrshort{CCSD} parameters are optimized in the presence of an
+approximate polarization and are used to form the triples free energy
+corrections.
+Thus, despite the absence of explicit polarization terms in the
+\acrshort{PTE(S)} and \acrshort{PTES} triples free energy corrections,
+the effect of the polarization on the approximate triples amplitudes is
+implicitly included.
+
+The theory and implementation for iterative \acrshort{CC} methods with
+quantum/classical polarizable Hamiltonians has already been presented in
+the context of the multipolar continuum and \acrlong*{PE}
+models.\autocite{Kongsted2002-zm, Osted2003-qc, Sneskov2011-jm,
+Schwabe2012-cf}
+Once again, the \acrshort{PTE} scheme can be trivially derived from the
+\acrshort{PTED} effective Lagrangians \eqref{eq:pted-cc2-lagrangian} and
+\eqref{eq:pted-cc3-lagrangian} by neglecting the polarization components.
+In the \acrshort{PTE(S)} scheme, we replace the expectation value of the
+source with its \acrshort{CCS} counterpart and further neglect the
+multiplier-dependent part, Eq. \eqref{eq:pte-s-source}.
+For \acrshort{CC2}, the amplitudes equations are the same as \emph{in
+vacuo}, Eqs. \eqref{eq:cc2-amplitudes} The multipliers are instead
+determined by:
+\begin{subequations}\label{eq:pte-s-cc2-multipliers}
+  \begin{align}
+    \denom{\mu_1}\tbar{\mu_1} &+
+     \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}}
+   + \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_1}}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{Q_2}
+     | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{2} | \commutator{\check{\Phi}}{\cluster{\mu_1}} |
+   \mathrm{HF}}
+   + {}^t\p\braket{\mathrm{HF} |
+   \commutator{\check{\s}_\mathrm{N}}{\cluster{\mu_1}} | \mathrm{HF}}
+   = 0 \label{eq:pte-s-cc2-singles-multipliers} \\
+    \denom{\mu_2}\tbar{\mu_2} &+
+    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}}
+   + \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_2}}
+     | \mathrm{HF}}
+     = 0 \label{eq:pte-s-cc2-doubles-multipliers}
+  \end{align}
+\end{subequations}
+The \acrshort{CC3} amplitudes are once again the same as \emph{in
+vacuo}, Eqs. \eqref{eq:cc3-amplitudes}
+The multipliers equations are:
+\begin{subequations}\label{eq:pte-s-cc3-multipliers}
+  \begin{align}
+    \denom{\mu_1}\tbar{\mu_1} &+
+     \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}}
+   + \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_1}}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+     | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{2} |
+       \commutator{\check{\Phi}}{\cluster{\mu_1}}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{Q_3}
+   | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{3} |
+      \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+     | \mathrm{HF}}
+   + {}^t\p\braket{\mathrm{HF} |
+   \commutator{\check{\s}_\mathrm{N}}{\cluster{\mu_1}} | \mathrm{HF}}
+     = 0 \label{eq:pte-s-cc3-singles-multipliers} \\
+    \denom{\mu_2}\tbar{\mu_2} &+
+    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}}
+   + \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_2}}
+     | \mathrm{HF}} \nonumber \\
+  &+ \braket{\tbar{2} |
+       \commutator{\check{\Phi}}{\cluster{\mu_2}}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_2}}}{T_2}
+     | \mathrm{HF}}
+  + \braket{\tbar{3} |
+       \commutator{\check{\Phi}}{\cluster{\mu_2}}
+     | \mathrm{HF}}
+    = 0 \label{eq:pte-s-cc3-doubles-multipliers} \\
+    \denom{\mu_3}\tbar{\mu_3} &+
+    \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_3}}
+     | \mathrm{HF}}
+  + \braket{\tbar{2} |
+       \commutator{\check{\Phi}}{\cluster{\mu_3}}
+    | \mathrm{HF}}
+    = 0 \label{eq:pte-s-cc3-triples-multipliers}
+  \end{align}
+\end{subequations}
+In both \acrlong*{CC} models, the polarization is still determined by
+Eq. \eqref{eq:pte-s-polarization} and only the singles multipliers
+equation is modified.
+
+In the \acrshort{PTES} scheme, we replace the expectation value of the
+source with its \acrshort{CCS} counterpart. The singles amplitudes equation
+thus contains a quantum/classical coupling term. However, to avoid
+introducing a coupling with the multipliers equations, an additional
+approximation is introduced in the polarization equation, where the
+source term is truncated to its multiplier-independent part.
+As already noted, this leads to the use of \emph{two} different
+polarization equations when solving for the amplitudes and the
+multipliers.
+Within the \acrshort{CC2} model we have the following amplitudes
+equations:
+\begin{subequations}\label{eq:ptes-cc2-amp}
+  \begin{align}
+   \denom{\mu_1}\tamp{\mu_1} &+ \braket{\mu_1 | \check{\Phi} +
+   \BCHfirst{\check{\Phi}}{Q_2} | \mathrm{HF}}
+   + {}^t\p\braket{\mu_1 | \check{\s} | \mathrm{HF}}
+             = 0 \label{eq:ptes-cc2-singles-amplitudes} \\
+   \denom{\mu_2}\tamp{\mu_2} &+ \braket{\mu_2 | \check{\Phi} | \mathrm{HF}}
+             = 0 \label{eq:ptes-cc2-doubles-amplitudes} \\
+    \V\p &+
+    \braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}} = 0
+  \end{align}
+\end{subequations}
+For the \acrshort{CC2} multipliers one has:
+\begin{subequations}\label{eq:ptes-cc2-mult}
+  \begin{align}
+    \denom{\mu_1}\tbar{\mu_1} &+
+    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}} +
+    \braket{\tbar{1} |
+    \commutator{\check{\Phi}}{\cluster{\mu_1}}
+    + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{Q_2}
+    | \mathrm{HF}}
+    \nonumber \\
+    &+ \braket{\tbar{2} |
+    \commutator{\check{\Phi}}{\cluster{\mu_1}}
+    | \mathrm{HF}}
+    \nonumber \\
+    &+
+    {}^t\p\braket{\mathrm{HF} | \commutator{\check{\s}_\mathrm{N}}{\cluster{\mu_1}} | \mathrm{HF}}
+    +
+    {}^t\p\braket{\tbar{1} |
+    \commutator{\check{\s}}{\cluster{\mu_1}} | \mathrm{HF}}
+             = 0 \label{eq:ptes-cc2-singles-multipliers} \\
+    \denom{\mu_2}\tbar{\mu_2} &+
+    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}} +
+    \braket{\tbar{1} |
+    \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}}
+             = 0 \label{eq:ptes-cc2-doubles-multipliers} \\
+    \V\p &+ \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCS} = 0
+    \label{eq:ptes-cc2-polarization-mult}
+  \end{align}
+\end{subequations}
+We note that the \acrshort{PTES} approximation is crucial
 in the implementation of \acrshort{PE}-\acrshort{CC2} presented by
@@
 assuming the \acrshort{PTES} approximation.
-Moreover, the approximate \acrshort{PE}-\acrshort{CC} Lagrangian of
-\citeauthor{Krause2016-ee} is essentially the \acrshort{PTE(S)}
-approximation for \acrshort{PE}-\acrshort{CC}.\autocite{Krause2016-ee}
+Moreover, the simplified \acrshort{PE}-\acrshort{CC}
+(s\acrshort*{PE}\acrshort*{CC}) Lagrangian of \citeauthor{Krause2016-ee}
+is essentially the \acrshort{PTE(S)} approximation for
+\acrshort{PE}-\acrshort{CC}.\autocite{Krause2016-ee}
+
+Finally, for the \acrshort{CC3} model the amplitudes are determined by:
+\begin{subequations}\label{eq:ptes-cc3-amplitudes}
+  \begin{align}
+    \denom{\mu_1}\tamp{\mu_1} &+ \braket{\mu_1 |
+    \check{\Phi} + \commutator{\check{\Phi}}{T_2}
+    + \commutator{\check{\Phi}}{Q_3}
+    | \mathrm{HF}}
+    + {}^t\p\braket{\mu_1 | \check{\s} | \mathrm{HF}}
+    = 0 \label{eq:ptes-cc3-singles-amplitudes} \\
+    \denom{\mu_2}\tamp{\mu_2} &+ \braket{\mu_2 |
+    \check{\Phi} + \commutator{\check{\Phi}}{T_2}
+    + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2}
+    + \commutator{\check{\Phi}}{Q_3}
+    | \mathrm{HF}}
+    = 0 \label{eq:ptes-cc3-doubles-amplitudes} \\
+    \denom{\mu_3}\tamp{\mu_3} &+ \braket{\mu_3 |
+    \commutator{\check{\Phi}}{T_2}
+    | \mathrm{HF}}
+    = 0 \label{eq:ptes-cc3-triples-amplitudes} \\
+    \V\p &+
+    \braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}} = 0
+    \end{align}
+\end{subequations}
+and the multipliers by:
+\begin{subequations}\label{eq:ptes-cc3-multipliers}
+  \begin{align}
+    \denom{\mu_1}\tbar{\mu_1} &+
+     \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}}
+   + \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_1}}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+     | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{2} |
+       \commutator{\check{\Phi}}{\cluster{\mu_1}}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{Q_3}
+   | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{3} |
+      \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
+     | \mathrm{HF}}
+   + {}^t\p\braket{\mathrm{HF} |
+   \commutator{\check{\s}_\mathrm{N}}{\cluster{\mu_1}} | \mathrm{HF}}  \nonumber \\
+   &+ {}^t\p\braket{\tbar{1} | \commutator{\check{\s}}{\cluster{\mu_1}} | \mathrm{HF}}
+     = 0 \label{eq:ptes-cc3-singles-multipliers} \\
+    \denom{\mu_2}\tbar{\mu_2} &+
+    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}}
+   + \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_2}}
+     | \mathrm{HF}} \nonumber \\
+  &+ \braket{\tbar{2} |
+       \commutator{\check{\Phi}}{\cluster{\mu_2}}
+     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_2}}}{T_2}
+     | \mathrm{HF}}
+  + \braket{\tbar{3} |
+       \commutator{\check{\Phi}}{\cluster{\mu_2}}
+     | \mathrm{HF}}
+    = 0 \label{eq:ptes-cc3-doubles-multipliers} \\
+    \denom{\mu_3}\tbar{\mu_3} &+
+    \braket{\tbar{1} |
+       \commutator{\check{\Phi}}{\cluster{\mu_3}}
+     | \mathrm{HF}}
+  + \braket{\tbar{2} |
+       \commutator{\check{\Phi}}{\cluster{\mu_3}}
+     | \mathrm{HF}}
+    = 0 \label{eq:ptes-cc3-triples-multipliers} \\
+    \V\p &+ \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCS} = 0
+    \label{eq:ptes-cc3-polarization-mult}
+  \end{align}
+\end{subequations}

### 2016-09-19 Chapters/ContributionsSummary.tex
Add reference to SC15 workshop proceedings

@@
 Performance and Sustainable Software for Molecular Simulation} workshop
-held at the 2015 Supercomputing Conference.\autocite{SC15}
+held at the 2015 Supercomputing Conference.\autocite{SC15, Trott2016-pf}
 Sections \ref{sec:relapcm}--\ref{sec:pcmopenrsp} contain very short

### 2016-09-27 Chapters/ContributionsSummary.tex
Update a picture

--- 2016-09-15 Chapters/ContributionsSummary.tex
@@
 Performance and Sustainable Software for Molecular Simulation} workshop
-held at the 2015 Supercomputing Conference.\autocite{SC15}
+held at the 2015 Supercomputing Conference.\autocite{%SC15, Trott2016-pf,
+DiRemigio2016-es}
 Sections \ref{sec:relapcm}--\ref{sec:pcmopenrsp} contain very short

### 2016-09-27 Chapters/ContinuumSolvationModels.tex
Update a picture
Luca's comments and corrections are in for the main body of the thesis

@@
 The need for ensemble averages leads us to the following observations:
-\begin{enumerate*}[label={\alph*)},font={\color{PMS1797}}]
+\begin{enumerate}%[label={\alph*)},font={\color{PMS1797}}]
  \item
@@
    some continuous distribution function.
-\end{enumerate*}
+\end{enumerate}
 
@@
 depicted in Figure \ref{fig:IEF}.
-The solute molecule is enclose in a \emph{cavity} inside the
+The solute molecule is enclosed in a \emph{cavity} inside the
 structureless continuum representing the solvent and the basic energetic
@@
 framework of continuum models. It accounts for the mutual electrostatic
-polarization of solute and solvent, see Section \ref{sec:IEF}.
-The third $G_\mathrm{dis}$ and fourth $G_\mathrm{rep}$ terms account for
-dispersion and repulsion interactions,
+polarization of solute and solvent (Section \ref{sec:IEF})
+The third and fourth terms account for
+dispersion ($G_\mathrm{dis}$) and repulsion ($G_\mathrm{rep}$) interactions,
 respectively.\autocite{Amovilli1997-fo, Weijo2010-jy}
@@
 Mozgawa2016-wf} it is also true that the electrostatic component is the
-most numerically relevant and the most sensitive to changes in the
-system studied and in the level of quantum mechanical description.
+most numerically relevant, especially when considering energy
+differences between different electronic states.
+%and the most sensitive to changes in the
+%system studied and in the level of quantum mechanical description.
 
@@
   \centering
-  \includegraphics[width=.5\textwidth]{IEF}
+  \includegraphics[width=.5\textwidth]{IEF.png}
   \caption[The physical setting of the polarizable continuum model.]{
@@
 for quantum mechanical charge densities. It can however be proven that
-the errors induced by this \emph{charge penetration} are not so
+the errors induced by this \emph{charge penetration} are not
 severe.\autocite{Chipman2000-us, Cances2001-qs, Cances2001-qn}
@@
 for the solution $u$ and its \emph{conormal} derivative.
-For notational simplicity, will use the symbols $\partiale$ and $\partiali$
+For notational simplicity, we will use the symbols $\partiale$ and $\partiali$
 for the latter and only give it in explicit form when needed.
@@
 numerical technique: the \gls{BEM}. This is implemented in four steps:
-\begin{enumerate*}[label={\alph*)},font={\color{PMS1797}}]
+\begin{enumerate}%[label={\alph*)},font={\color{PMS1797}}]
     \item define the molecular cavity
@@
     \item solve the linear systems by a suitable numerical technique.
-\end{enumerate*}
+\end{enumerate}
 Many different approaches to these four steps have been presented, both
@@
 three types of molecular surfaces can be defined:
-\begin{enumerate*}[label={\alph*)},font={\color{PMS1797}}]
+\begin{enumerate}%[label={\alph*)},font={\color{PMS1797}}]
   \item
@@
     spherical probe while rolling over the \acrshort{vdWS} is the \gls{SES}, Figure \ref{fig:SES}.
-\end{enumerate*}
+\end{enumerate}
 \begin{figure}[tb]
@@
 Alternative definitions based on the isodensity surface of the molecular solute
-have been put forth.\autocite{Foresman1996-wv}
+have also been put forth.\autocite{Foresman1996-wv}
 
@@
 variational formulation of classical polarizable models.
-Why so? In such a framework the \acrshort{ASC} becomes an additional,
+In such a framework the \acrshort{ASC} becomes an additional,
 independent, variationally optimizable degree of freedom.
@@
 properties:\autocite{Solis2013-ef}
-\begin{enumerate*}[label={\alph*)},font={\color{PMS1797}}]
+\begin{enumerate}%[label={\alph*)},font={\color{PMS1797}}]
     \item physical extremal points,
@@
     \item convexity (positive-definiteness)
-\end{enumerate*}
+\end{enumerate}
 \citeauthor{Allen2001-fp} and \citeauthor{Attard2003-vr} pioneering
@@
 Solis2013-ef}
-Our focus however is on the quantum mechanical calculations of molecular
+Our focus is however on the quantum mechanical calculations of molecular
 properties and the variational formulation is advantageous since it
@@
 The variational formulation also lends itself to an implementation
-in terms of simultaneous optimization.\autocite{Lipparini2011-aj}
+as a simultaneous optimization of the electronic structure and
+polarization degrees of freedom.\autocite{Lipparini2011-aj}
 

### 2016-09-27 Chapters/ContributionsSummary.tex
Luca's comments and corrections are in for the main body of the thesis

@@
 
-The \pcmsolver library was developed to fulfill these requirements and
+The \pcmsolver library has been developed to fulfill these requirements and
 offer the community an easy-to-use implementation of the
@@
 part of the standard we chose to comply to.
-Parts of the Boost C++ libraries are used to alleviate these
+Parts of the \Boost C++ libraries are used to alleviate these
 shortcomings.\autocite{boost}
-Linear algebra operations are managed through the Eigen C++ template
+Linear algebra operations are managed through the \Eigen C++ template
 library.\autocite{eigen}
 The Python layer provides input reading and parsing facilities,
-as implemented by the GetKw library.\autocite{libgetkw}
+as implemented by the \GetKw library.\autocite{libgetkw}
 Cavity construction and discretization is handled either by the Fortan
@@
 Evaluation of the Green's functions, needed to set up the representation
-of the boundary integral operators, leverages \gls{AD}. We use the Taylor library, a template C++
+of the boundary integral operators, leverages \gls{AD}. We use the \Taylor library, a template C++
 implementation of \acrshort{AD}.\autocite{Bartholomew-Biggs2000-db, libtaylor}
@@
 Preeminent use of composition over inheritance keeps the coupling
-between different modules as low as possible.\autocite{Gamma1995-fd}
+between different submodules as low as possible.\autocite{Gamma1995-fd}
 
@@
                  caption={Relevant calls to \pcmsolver in a C host program.
-                 Ellipsis in the functions arguments list stand for omitted parameters.
+                 Ellipses in the functions' argument list stand for omitted parameters.
                  For the full, functional example visit
@@
                  Notice the use of the standard Fortan ISO C bindings.
-                 Ellipsis in the functions arguments list stand for omitted parameters.
+                 Ellipses in the functions' argument list stand for omitted parameters.
                  For the full, functional example visit
@@
 
+\begin{tcolorbox}
+\begin{flushright}
+  {\small
+  \textbf{R. Di Remigio}, R. Bast, L. Frediani, and T. Saue
+  \\
+  \textsc{
+  Four-Component Relativistic Calculations in Solution with the
+  Polarizable Continuum Model of Solvation: Theory,
+  Implementation, and Application to the Group 16 Dihydrides
+  \ce{H2X} (\ce{X} = \ce{O}, \ce{S}, \ce{Se}, \ce{Te},
+  \ce{Po})
+  }
+  \\
+  \textit{J. Phys. Chem. A}, \textrm{2015}, \textbf{119}, 5061--5077
+  \\
+  DOI: \url{10.1063/1.4943782}
+  }
+\end{flushright}
+\end{tcolorbox}
+
 Systems containing heavy elements are notoriously challenging for quantum chemistry.
-In addition to their significantly large sizes, one has to properly include the
-effect of special relativity in the quantum chemical description in order to
-achieve at least qualitative agreement with experiment.
+One has to properly include the effect of special relativity in the
+quantum chemical description in order to achieve at least qualitative
+agreement with experiment.
 Additional complications arise if one needs to also include environment effects.
@@
 approximation of these effects.
-In this paper, we presented the first derivation and implementation of the \acrshort{PCM}
+In \paper{I} we presented the first derivation and implementation of the \acrshort{PCM}
 coupled to a \acrshort{SCF} description of the solute in the fully relativistic, four-component
@@
 elements are involved.
-The main point of the paper was, however, the adoption of a fully modular
+The main point of \paper{I} was, however, the adoption of a fully modular
 programming strategy. We showed that it is entirely possible to adopt the same
@@
 in the four-component relativistic regime.
-The interface to \DIRAC was later released in the 2014 and later versions,
+The interface to \DIRAC was first released in the 2014 version,
 providing \acrshort{PCM} capabilities to the software package.
@@
 \pcmsolver by providing the implementation and testing of:
-\begin{enumerate*}[label={\alph*)},font={\color{PMS1797}}]
+\begin{enumerate}%[label={\alph*)},font={\color{PMS1797}}]
   \item \acrshort{MEP} integrals for four-component wave functions,
@@
   \item the additional terms in the response equations.
-\end{enumerate*}
+\end{enumerate}
 I performed all the calculations and large part of the data analysis
@@
 \section{The Wavelet Galerkin Boundary Element Method for PCM}\label{sec:wemlin}
+
+\begin{tcolorbox}
+\begin{flushright}
+  {\small
+  M. Bugeanu, \textbf{R. Di Remigio}, K. Mozgawa, S. S. Reine, H.
+  Harbrecht,  and L. Frediani
+  \\
+  \textsc{
+  Wavelet Formulation of the Polarizable Continuum Model. II. Use of
+  Piecewise Bilinear Boundary Elements
+  }
+  \\
+  \textit{Phys. Chem. Chem. Phys.}, \textrm{2015}, \textbf{17},
+  31566--31581
+  \\
+  DOI: \url{10.1039/C5CP03410H}
+  }
+\end{flushright}
+\end{tcolorbox}
 
@@
 The use of biorthogonal wavelet bases as finite elements achieves sparsity in
-the \acrshort{BEM} procedure, due to their intrinsically hierarchical structure and
+the \acrshort{BEM} procedure, due to the intrinsical hierarchy and
 the existence of \emph{a priori} and \emph{a posteriori} matrix compression
@@
 
-For this paper, I provided template interface and test sets for the cavity
+For \paper{II}, I provided template interface and test sets for the cavity
 generator\autocite{Harbrecht2009-no, Harbrecht2011-dk} and wavelet Galerkin
@@
 \section{Non Homogeneous Environments}\label{sec:spherical}
+
+\begin{tcolorbox}
+\begin{flushright}
+  {\small
+  \textbf{R. Di Remigio}, K. Mozgawa, H. Cao, V. Weijo, and L.
+  Frediani
+  \\
+  \textsc{
+  A Polarizable Continuum Model for Molecules at Spherical
+  Diffuse Interfaces
+  }
+  \\
+  \textit{J. Chem. Phys.}, \textrm{2016}, \textbf{144}, 124103
+  \\
+  DOI: \url{10.1063/1.4943782}
+  }
+\end{flushright}
+\end{tcolorbox}
 
@@
 
-I contributed the theoretical work for this paper, based on earlier drafts from coauthors
+I contributed the theoretical work for \paper{III}, based on earlier drafts from coauthors
 Ville Weijo and Hui Cao. In particular, I derived the separation of
@@
 Moreover, I contributed the implementation and testing of the Green's function code.
-The interface to the \LSDALTON program package, developed within paper II, was
+The interface to the \LSDALTON program package, developed within
+\paper{II}, was
 also used for this paper.
@@
 Solution}\label{sec:pcmepr}
+
+\begin{tcolorbox}
+\begin{flushright}
+  {\small
+  \textbf{R. Di Remigio}, M. Repisky, S. Komorovsky, P. Hrobarik, L.
+  Frediani, and K. Ruud
+  \\
+  \textsc{
+  Four-Component Relativistic Density Functional Theory with the
+  Polarizable Continuum Model: Application to EPR Parameters
+  and Paramagnetic NMR Shifts
+  }
+  \\
+  Accepted for publication in \textit{Mol. Phys.}
+  \\
+  DOI: \url{10.1080/00268976.2016.1239846}
+  }
+\end{flushright}
+\end{tcolorbox}
 
@@
 
+\begin{tcolorbox}
+\begin{flushright}
+  {\small
+  \textbf{R. Di Remigio}, M. T. P. Beerepoot, Y. Cornaton, M. Ringholm,
+  A. H. S. Steindal, K. Ruud, and L. Frediani
+  \\
+  \textsc{
+  Open-Ended Formulation of Self-Consistent Field Response Theory with
+  the Polarizable Continuum Model for Solvation
+  }
+  \\
+  In preparation for submission to \textit{Phys. Chem. Chem. Phys.}
+  }
+\end{flushright}
+\end{tcolorbox}
+
 In recent years, the availability of strong lasers has allowed to

### 2016-09-27 Chapters/QuantumMechanics.tex
Luca's comments and corrections are in for the main body of the thesis

@@
 while in the relativistic regime with the \gls{DC}
-approximation:\autocite{Dyall2007-tu, Konishi2009-zb, Reiher2014-cp}
+approximation they become:\autocite{Dyall2007-tu, Konishi2009-zb, Reiher2014-cp}
 \begin{subequations}
@@
 minimization with respect to the parameters of a well-crafted ansatz
-ensures convergence from above to the exact energy.\autocite{Szabo1989-vl, McWeeny1992-oj, Helgaker2000-tz}
+will converge from above to the exact energy.\autocite{Szabo1989-vl, McWeeny1992-oj, Helgaker2000-tz}
 A wave function ansatz enjoying the latter property guarantees that the
@@
   =
-  \frac{\delta E_\mathrm{xc}[\rho(\vect{r})]}{\delta \rho(\vect{r})}
+  \frac{\delta E_\mathrm{xc}[\rho(\vect{r})]}{\delta \rho(\vect{r})}.
 \end{alignat}
-notice moreover the exact exchange admixing factor $\gamma$ in the
+Notice moreover the exact exchange admixing factor $\gamma$ in the
 two-electron part of the \acrshort*{KS} matrix.
@@
 \begin{equation}
-  \ket{\mathrm{CC}} = \expo{T}\ket{\mathrm{HF}},
+  \label{eq:cluster}
+  \ket{\mathrm{CC}} = \expo{T}\ket{\mathrm{HF}}.
 \end{equation}
-out of convenience we assume, here and in the following, that the
+Out of convenience we assume, here and in the following, that the
 reference function is a closed-shell \acrshort{HF} wave function.
@@
 \acrshort{CC} theory and we introduce the following notation for it:
-\begin{equation}
+\begin{equation}\label{eq:similarity-transformation}
   \overline{O} =\expo{-T}O\expo{T}
@@
 Similarity-transformed operators can be expanded in a \acrshort{BCH}
-commutator series.
+commutator series:
+\begin{equation}\label{eq:bch}
+  \overline{O}
+  =
+  O + \BCHfirst{O}{T} + \BCHsecond{O}{T} + \ldots
+  =
+  O + \tilde{O}.
+\end{equation}
 It can be shown that the \acrshort{BCH} expansion of the
@@
 four-fold nested commutator, greatly simplifying algebraic derivations
-and manipulations.
+and manipulations.\autocite{Helgaker2000-tz}
 
@@
 where $\denom{\mu_u}$ is the difference in orbital energies between the
-virtual and occupied spin-orbitals of excitation $u$.
+occupied and virtual spin-orbitals of excitation $u$.
 For the fluctuation potential the similarity transformation truncates
@@
 \begin{equation}
+  \label{eq:left-state}
   \bra{\tbar{u}} = \sum_{\mu_u}\tbar{\mu_u}\bra{\mu_u}.
@@
 \end{equation}
-which shows that singles, due to Brillouin's condition, and triples
-amplitudes are of second and higher order.
+which shows that singles and triples amplitudes are of second and higher
+order.
+For the former this is a consequence of Brillouin's condition, valid for
+the reference closed-shell determinant.
 For second order singles, doubles and triples one obtains:
@@
 These expressions form the basis for the development of reduced-scaling
-\acrshort*{CC} schemes were the double and triple excitation are treated
+\acrshort*{CC} schemes were the double and triple excitations are treated
 approximately. These are the subject of the next Section.
@@
 in recovering a large portion of the correlation energy, but as hinted
-at in the previous Section is only correct to third order in
+at in the previous Section, is only correct to third order in
 perturbation theory.
@@
 \begin{equation}
+  \label{eq:t1-transformation}
   \check{O} = \expo{-T_1}O\expo{T_1}.
@@
 \end{equation}
-The \acrshort*{CC2} methods brings the scaling of the \acrshort*{CCSD}
+The \acrshort*{CC2} method brings the scaling of the \acrshort*{CCSD}
 method down to $N^5$ in the iterative step.

### 2016-09-27 Chapters/ResponseTheory.tex
First draft sent to L.F. and B.M.
Luca's comments and corrections are in for the main body of the thesis

@@
 Section \ref{sec:exact-response} will give a brief introduction to
-response theory for isolated molecules. The concepts of
-quasienergy,\autocite{Christiansen1998-pe} variational perturbation
-theory\autocite{Helgaker1992-ph} and pole-and-residue
-analysis of response functions\autocite{Olsen1985-nr} will be introduced.
-The exposition closely follows the one by \citeauthor{Saue2002-ns}
-in \noparcite[ref.][]{Saue2002-ns}.
+response theory for isolated molecules.
+
+\noindent The concepts of quasienergy,\autocite{Christiansen1998-pe} variational
+perturbation theory\autocite{Helgaker1992-ph} and pole-and-residue
+analysis of response functions\autocite{Olsen1985-nr} will be
+introduced. The exposition closely follows the one by
+\citeauthor{Saue2002-ns} in \noparcite[ref.][]{Saue2002-ns}.
 I will explicitly derive the \acrshort{SCF} response function for a
@@
 where the time-derivative of the phase factor defines the time-dependent
-\emph{quasienergy}:\autocite{Christiansen1998-pe, Pawlowski2015-sq}
+\emph{quasienergy}:
 \begin{equation}
@@
 time-dependent Hellmann--Feynman theorem cannot straightforwardly be
-applied to the calculation of molecular properties. Fortunately, we can exploit the
-periodicity of the perturbation operator to introduce the time-averaged
-quasienergy:
+applied to the calculation of molecular properties.\autocite{Christiansen1998-pe, Pawlowski2015-sq}
+Fortunately, we can exploit the periodicity of the perturbation operator
+to introduce the time-averaged quasienergy:
 \begin{equation}
@@
 effect of external perturbations on molecular properties, it is now time
-to consider how to actually compute them.
+to consider how to compute them.
 In this Section, we are concerned with exact wave functions that can
-thus be parametrized in a variational fashion by a set of parameters
+be variationally parametrized by a set of parameters
 $\vect{\lambda}$: $\ket{0}=\ket{0(\vect{\lambda})}$.
-We recall that the Lagrangian method outlines in Section
+We recall that the Lagrangian method outlined in Section
 \ref{sec:coupled-cluster} can be employed also in this
@@
 
-The response equations, determining the $\vect{\lambda}^{[n]}$ tensor of
-perturbed wave function parameters, are obtained by differentiating the
+The response equations, determining the tensor of
+perturbed wave function parameters $\vect{\lambda}^{[n]}$, are obtained by differentiating the
 variational condition on the time-averaged quasienergy Eq.
@@
 the poles are the transition moments determining the intensity of
-spectroscopic transitions:
+the spectroscopic transitions:
 \begin{equation}
@@
 The derivation above holds for an exact state, a situation that is never
-realized in common practice. The power of response theory, however, is
-in its applicability to exact and approximate state theories alike.
+realized in common practice. However, response theory is
+also applicable to approximate states.
 As noted by \citeauthor{Norman2011-ad}, a response theory treatment of
@@
 in the \acrshort*{AO} basis.
-The density matrix is the variational degree of freedom and response
+The density matrix is the variational degree of freedom and the response
 equations can be obtained by differentiation of the \gls{TDSCF}
@@
 Mennucci2005-vi, Caricato2006-ba, Corni2015-pe}
-However, if one consider the initial steps of an electronic excitation
+However, if one considers the initial steps of an electronic excitation
 process, a Franck--Condon-like response of the solvent can be assumed.
-Nuclear motions inside and among solvent molecules will not be able to
-follow the fast rearrangement of the solute electronic density, the
+Nuclear motions of the solute and solvent molecules alike will not be able to
+follow the fast rearrangement of the electronic density, therefore the
 corresponding part of the response will remain frozen in the state
-immediately previous the change.
+immediately preceding the change.
+Such an assumption, albeit approximate, leads to the following partition
+of the polarization vector:
 The polarization vector will then be (approximately) a sum of two
@@
 \begin{equation}
- \vect{P} \simeq \vect{P}_\mathrm{fast} + \vect{P}_\mathrm{slow},
+ \vect{P} \simeq \vect{P}_\mathrm{fast} + \vect{P}_\mathrm{slow}.
 \end{equation}
-this is equivalent to consider only the asymptotic limits of the solvent
+This is equivalent to considering only the asymptotic limits of the solvent
 dielectric spectrum, \ie the static $\diels$ and optical
@@
 Hellmann--Feynman theorem holds:
-\begin{equation}\label{eq:classical-HF}
+\begin{equation}\tag{\ref{eq:classical-HF} from Chapter \ref{ch:CSM}}
   \deriv{U_\mathrm{pol}}{F} = \pderiv{U_\mathrm{pol}}{F}
@@
 respect to which differentiation occurs. The index $m$ denotes the order
-of differentiation with respect to the density matrix. Eventually, $n$
-eventually denotes the order of differentiation with respect to the
+of differentiation with respect to the density matrix. $n$
+denotes the order of differentiation with respect to the
 polarization.
@@
 The generalized Hessian matrix $\genHessian$ includes two types of
-classical terms. \emph{Implicit} terms are those included in the
-zeroth-order Fock matrix, $\mat{F}$.
+classical terms: implicit and explicit. \emph{Implicit} terms are those
+included in the zeroth-order Fock matrix, $\mat{F}$.
 The last two terms in Eq. \ref{eq:E2-lintra} are \emph{explicit} terms,
@@
 will affect the explicit terms in Eq. \ref{eq:E2-lintra} which are
-evaluated by means of the optical permittivity $\dield$.
+evaluated by making use of the optical permittivity $\dield$ in the
+polarization equations \eqref{eq:part-polarization-eq}.
 The \acrlong*{RHS} in the response equation only includes terms that
@@
 \omega_{b_N}\genMetric \right]$ matrix. \emph{Subspace iteration}
-methods are usually employed, the solution is achieved iteratively
+methods are usually employed and the solution is achieved iteratively
 by enlarging a projection subspace of trial vectors.\autocite{Saad2003-oa,

### 2016-09-27 Chapters/SolvationAndCorrelation.tex
Luca's comments and corrections are in for the main body of the thesis

@@
 
+\subsection*{Notation}
+
+For ease of reference, I will repeat some of the notational conventions
+introduced in Chapter \ref{ch:QM} here.
+The cluster operator at truncation level $\mathcal{M}$ is:
+\begin{equation}
+  T = \sum_{u=1}^\mathcal{M} T_u =
+  \sum_{u=1}^\mathcal{M}\sum_{\mu_u}\tamp{\mu_u}\cluster{\mu_u},
+  \tag{\ref{eq:cluster} from Chapter \ref{ch:QM}}
+\end{equation}
+where $\tamp{\mu_u}$ and $\cluster{\mu_u}$ are the cluster amplitudes
+and cluster operators for the $\mu_u$-th excited determinant at the
+$u$-th excitation level, respectively.
+The similarity transformed counterpart of the operator $O$ is denoted
+by $\overline{O}$:
+\begin{equation}
+  \overline{O} =\expo{-T}O\expo{T}.
+  \tag{\ref{eq:similarity-transformation} from Chapter \ref{ch:QM}}
+\end{equation}
+We recall that hermiticity of $O$ is not preserved by the similarity
+transformation.
+Manipulation of similarity transformed operators involves the use of
+their \acrlong*{BCH} expansions:
+\begin{equation}
+  \overline{O}
+  =
+  O + \BCHfirst{O}{T} + \BCHsecond{O}{T} + \ldots
+  =
+  O + \tilde{O},
+  \tag{\ref{eq:bch} from Chapter \ref{ch:QM}}
+\end{equation}
+since we are concerned with one- and two-electron operators only,
+such expansions will truncate after the third and fifth terms,
+respectively.\autocite{Helgaker2000-tz}
+The similarity transformation by means of the singles cluster operator
+only, \ie the $T_1$-transformation, will be denoted by a "check" on top
+of the operator:
+\begin{equation}
+  \check{O} = \expo{-T_1}O\expo{T_1}.
+  \tag{\ref{eq:t1-transformation} from Chapter \ref{ch:QM}}
+\end{equation}
+The $T_1$-transformation preserves the particle rank of the operator
+$O$, see Appendix \ref{app:mathematical-results} for details.
+Finally, the $u$-th excited component of the \acrlong*{CC} left state
+will be denoted as:
+\begin{equation}
+  \bra{\tbar{u}} = \sum_{\mu_u}\tbar{\mu_u}\bra{\mu_u},
+  \tag{\ref{eq:left-state} from Chapter \ref{ch:QM}}
+\end{equation}
+where $\tbar{\mu_u}$ are Lagrange multipliers.
+
 \section{Effective Coupled Cluster Lagrangian}
@@
 singles cluster operator and the singles multiplier state.
-Three approximation have currently been proposed: \gls{PTES}, \gls{PTE(S)}
+Three approximations have currently been proposed: \gls{PTES}, \gls{PTE(S)}
 and \gls{PTE}.
@@
 multipliers in the \acrshort{CCS} source term in the polarization equation.
-To break the coupling, we reformulate the split the system of equations
+To break the coupling, we split the system of equations
 \eqref{eq:ptes} into \emph{two} systems of equations, one for the
@@
 obtained using the full \acrshort*{CCS} expectation value of the source
-term. However, this adds another $O(N^5)$ step to the iterative
+term. This adds another $O(N^5)$ step to the iterative
 procedure, which is, overall, negligible.
@@
 The amplitude equations are formally equivalent to those derived in the
-\acrshort{PTE} scheme, \emph{vide infra}. The equations for multipliers are also formally
-equivalent to the \acrshort{PTE} equations, apart from the equation
-determining the singles multipliers which features an additional term.
+\acrshort{PTE} scheme, \emph{vide infra}.
+The equations for the multipliers are also formally equivalent to the
+\acrshort{PTE} equations, apart from the equation determining the
+singles multipliers which features an additional term involving the
+polarization:
+${}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}$.
 This term is not, however, coupled to the other equations given that
@@
 Similarly to the exposition in Chapter \ref{ch:QM}, these expressions
-will form the basis for the development of methods for the approximate
+will form the basis for the development of approximations
 beyond \acrshort*{CCSD} including connected triples.
@@
 thus contains a quantum/classical coupling term. However, to avoid
-introducing a coupling with the multipliers equations, an additional
+the coupling with the multipliers equations, an additional
 approximation is introduced in the polarization equation, where the

### 2016-09-27 FrontBackmatter/Introduction.tex
Add Stefano's picture
Luca's comments and corrections are in for the main body of the thesis

@@
 Relativity can be considered as the third axis of quantum
-chemistry: accuracy can be reached improving on the Hamiltonian, method
+chemistry: accuracy can be improved on the Hamiltonian, method
 and basis axes.\autocite{Saue2011-qg}
@@
 The number of degrees of freedom to be taken into account is, in
-principle, so large, that even the most advanced computing machines
+principle, so large, that even the most powerful computers
 would have a hard time computing the desired observables.
@@
 To tame this complexity and cure the disease, one must devise
-\emph{models} that simplify physical reality, while offering tools for
+\emph{models} that simplify physical picture, while offering tools for
 understanding reality and predicting new and exciting
@@
 \item[Dynamic.]
-  Probing the nature of molecular excited states has enormous
-  technological impact. The presence of the environment radically
-  influences excited states, since relaxation processes in the medium
-  become important.
+  The presence of the environment radically influences excited states,
+  since relaxation processes in the medium become important.
 \item[Specific.] This catch-all category includes all effects
@@
 \begin{figure}[tb]
-\missingfigure{Add picture with multiscale models}
+\centering
+\includegraphics[width=.7\textwidth]{multiscale_models.png}
 \caption[From quantum mechanical to multiscale models]{
-From quantum mechanical to multiscale models.
-The images used in the scheme are reproduced courtesy of Dr.~Stefano
-Caprasecca (MoLEcoLab, Università di Pisa).}
+From quantum mechanical to multiscale models. Picture reproduced
+courtesy of Dr.~Stefano Caprasecca (MoLEcoLab, Università di Pisa).}
 \label{fig:qm-to-multiscale}
@@
  \item
-   \emph{Continuum} (or \emph{implicit}) completely remove the degrees
+   \emph{Continuum} (or \emph{implicit}) models completely remove the degrees
    of freedom of the environment from the model, replacing them with a
    structureless continuum.
-   The effect of this continuum is described, classically, \emph{via}
-   its bulk properties.\autocite{Onsager1936-wf, Miertus1981-mm}
+   Its effect is described, classically, \emph{via} its bulk
+   properties.\autocite{Onsager1936-wf, Miertus1981-mm}
 \end{itemize}

### 2016-09-28 FrontBackmatter/Colophon.tex
Contributions written
Fix some missing references in CC chapter
Adjust colophon

@@
 \section*{Colophon}
-This document was typeset using the typographical look-and-feel
-\texttt{classicthesis} developed by Andr\'e Miede. The style was
-inspired by Robert Bringhurst's seminal book on typography ``\emph{The
-Elements of Typographic Style}''. \texttt{classicthesis} is available
-for both \LaTeX\ and \mLyX:
-\begin{center}
-\url{https://bitbucket.org/amiede/classicthesis/}
-\end{center}
+This document was typeset using \XeLaTeX, with the
+\texttt{classicthesis} template developed by André Miede.\footnote{\url{https://bitbucket.org/amiede/classicthesis/}}
+The style was inspired by Robert Bringhurst's seminal book on typography
+\emph{The Elements of Typographic Style}.\autocite{Bringhurst2004-wn}
+The bibliography was processed using
+Biblatex\footnote{\url{https://www.ctan.org/pkg/biblatex}} and
+Biber.\footnote{\url{http://biblatex-biber.sourceforge.net/}}
+David J. Perry's Cardo provides both the text and display typeface.
+Open Sans is the sans-serif typeface, while Bitstream Vera Mono is used
+for monospaced text.
+Finally, XITS acts at the math typeface.
 
 \bigskip
-
-\noindent\finalVersionString
-
-\emph{Garamond} and \emph{XITS} type faces are used.
+{\small
+\noindent\myTitle
+\\
+\noindent\textcopyright\,\finalVersion, \myName
+}
 

### 2016-09-28 Chapters/MathematicalResults.tex
Switch to Minion Pro

@@
 
-\section{The \texorpdfstring{$T_1$}{T1}-Transformation}\label{sec:T1-cc}
+\section{The \texorpdfstring{$\T1$}{T1}-Transformation}\label{sec:T1-cc}
 
-$T_1$ is a one-electron operator, carrying out a similarity transformation of
-an operator $O$ by means of its exponential mapping will preserve the particle
-rank of $O$.\autocite{Koch1994-vr}
+Carrying out a similarity transformation of
+an operator $O$ by means of the $T_1$ cluster operator will preserve the particle
+rank of $O$, since $T_1$ is a one-electron operator.\autocite{Koch1994-vr}
 Starting from the arbitrary order Lagrangian \eqref{eq:cc-lagrangian},
@@
 \end{equation}
-Adding double excitation to the manifold yields:
+Adding double excitations to the manifold yields:
 \begin{equation}\label{eq:ccsd-expectation-value}
@@
       + \commutator{\commutator{O}{T_1}}{T_2}
-  | \mathrm{HF} }
+  | \mathrm{HF} },
   \end{aligned}
 \end{equation}
-a number of terms was dropped thanks to Lemma \ref{lemma:exc-rank} and its
+where a number of terms was dropped thanks to Lemma \ref{lemma:exc-rank} and its
 Corollary.
@@
 the dressed and bare representations, the expectation values of such operators
-can be formed by simply replacing operator with its normal-ordered counterpart,
+can be formed by simply replacing the operator with its normal-ordered counterpart,
 $\check{O}_\mathrm{N}$ or $O$, respectively.
@@
 
-In this Section, I will show how one can reformulate the transmission problem
-in a variational fashion.
+The transmission problem can be reformulated in a variational fashion.
 Such a formulation allows for a larger vector space, with weaker regularity

### 2016-09-28 FrontBackmatter/Colophon.tex
Switch to Minion Pro

@@
 
-\pdfbookmark[0]{Colophon}{colophon}
+\pdfbookmark[1]{Colophon}{colophon}
 \section*{Colophon}
@@
 Biber.\footnote{\url{http://biblatex-biber.sourceforge.net/}}
-David J. Perry's Cardo provides both the text and display typeface.
-Open Sans is the sans-serif typeface, while Bitstream Vera Mono is used
-for monospaced text.
+Roberto Slimbach's Minion Pro provides both the text and display typeface.
+\textsf{Open Sans} is the sans-serif typeface, while Jim Lyles's
+\texttt{Bitstream Vera Mono} is used for monospaced text.
 Finally, XITS acts at the math typeface.
@@
 {\small
-\noindent\myTitle
+\noindent\textit{\myTitle}
 \\

### 2016-09-28 Chapters/ContributionsSummary.tex
Fix bibliography

@@
                  For the full, functional example visit
-                 \url{http://pcmsolver.readthedocs.io/en/v1.1.5/users/C-example.html}},
+                 \url{http://pcmsolver.readthedocs.io/en/v1.1.6/users/C-example.html}},
                  label={lst:C-host}
@@
                  For the full, functional example visit
-                 \url{http://pcmsolver.readthedocs.io/en/v1.1.5/users/fortran-example.html}},
+                 \url{http://pcmsolver.readthedocs.io/en/v1.1.6/users/fortran-example.html}},
                  label={lst:Fortran-host}
@@
 \begin{tcolorbox}
-\begin{flushright}
   {\small
@@
   \\
-  \textsc{
+  \textsf{
   Four-Component Relativistic Calculations in Solution with the
@@
   }
-\end{flushright}
 \end{tcolorbox}
@@
   \\
-  \textsc{
+  \textsf{
   Wavelet Formulation of the Polarizable Continuum Model. II. Use of
@@
   \\
-  \textsc{
+  \textsf{
   A Polarizable Continuum Model for Molecules at Spherical
@@
   \\
-  \textsc{
+  \textsf{
   Four-Component Relativistic Density Functional Theory with the
@@
   \\
-  \textsc{
+  \textsf{
   Open-Ended Formulation of Self-Consistent Field Response Theory with

### 2016-09-28 FrontBackmatter/Publications.tex
Adjust TOC bookmarks
First draft sent to L.F. and B.M.
Fix bibliography

@@
 \thispagestyle{empty}
+\phantomsection
 \refstepcounter{dummy}
-\pdfbookmark[1]{Publications}{Publications}
+\addcontentsline{toc}{chapter}{\tocEntry{Publications}}
+
 \chapter*{Publications}
@@
 
-\begin{enumerate}[label=\textbf{\Roman{*} }, ref=\Roman{*}]
-
+\begin{enumerate}[label=\paper{\Roman{*}}, ref=\Roman{*}]
 \item
@@
 \textit{J. Phys. Chem. A}, \textrm{2015}, \textbf{119}, 5061--5077
+  \\
+  DOI: \url{10.1063/1.4943782}
 \label{relapcm}
@@
   31566--31581
+  \\
+  DOI: \url{10.1039/C5CP03410H}
 \label{wemlin}
@@
   \textit{J. Chem. Phys.}, \textrm{2016}, \textbf{144}, 124103
+  \\
+  DOI: \url{10.1063/1.4943782}
 \label{spherical}
@@
   Accepted for publication in \textit{Mol. Phys.}
+  \\
+  DOI: \url{10.1080/00268976.2016.1239846}
 \label{pcmepr}
@@
   \\
-  \textsc{
-  Open-ended formulation of self-consistent field response theory with
-  the polarizable continuum model for solvation
+  \textsf{
+  Open-Ended Formulation of Self-Consistent Field Response Theory with
+  the Polarizable Continuum Model for Solvation
   }

### 2016-09-29 FrontBackmatter/Abstract.tex
Update abstract

@@
 implemented.
-We implemented an application programming interface that can provide the
-continuum solvation functionality to any quantum chemistry software.
+
+The computer implementation of the polarizable continuum model is
+central to the work presented in this thesis.
+The quantum chemistry software ecosystem suffers from a growing
+complexity. Modular programming offers an extensible, flexible and
+sustainable paradigm to implement new features with reduced effort.
+\pcmsolver, our open-source application programming interface, can
+provide continuum solvation functionality to any quantum chemistry
+software: continuum solvation goes \emph{viral}.
+Our strategy affords simpler programming workflows, more thorough
+testing and lower overall code complexity.
 As examples of the flexibility of our implementation approach, we
-present results for the continuum modelling of non homogeneous environments and
-how wavelet-based numerical methods greatly outperform the accuracy of
-traditional methods usually employed in continuum solvation models.
+present results for the continuum modelling of non homogeneous
+environments and how wavelet-based numerical methods greatly outperform
+the accuracy of traditional methods usually employed in continuum
+solvation models.
 

### 2016-09-30 FrontBackmatter/Colophon.tex
Fine-tuning

@@
 The bibliography was processed using
-Biblatex\footnote{\url{https://www.ctan.org/pkg/biblatex}} and
+Bib\LaTeX\footnote{\url{https://www.ctan.org/pkg/biblatex}} and
 Biber.\footnote{\url{http://biblatex-biber.sourceforge.net/}}
-Roberto Slimbach's Minion Pro provides both the text and display typeface.
+Robert Slimbach's Minion Pro provides both the text and display typeface.
 \textsf{Open Sans} is the sans-serif typeface, while Jim Lyles's

### 2016-09-30 FrontBackmatter/Colophon.tex
Use TeX Gyre Termes and TeX Gyre Adventor fonts

@@
 Biber.\footnote{\url{http://biblatex-biber.sourceforge.net/}}
-Robert Slimbach's Minion Pro provides both the text and display typeface.
-\textsf{Open Sans} is the sans-serif typeface, while Jim Lyles's
+TeX Gyre Termes provides both the text and display typeface.
+\textsf{TeX Gyre Adventor} is the sans-serif typeface, while Jim Lyles's
 \texttt{Bitstream Vera Mono} is used for monospaced text.
-Finally, XITS acts at the math typeface.
+Finally, XITS acts as the math typeface.
 

### 2016-09-30 Chapters/ContributionsSummary.tex
Fine-tuning
Minor polish

@@
 
-          \textonehalf\,\,observation, \kern 22pt \textonehalf \kern 20pt \textgreek{Tέχνη}
+          \textonehalf\,\,observation, \kern 21pt \textonehalf \kern 20pt \textgreek{Tέχνη}
 
-          \textonehalf\,\,training, \kern 39pt \textonehalf \kern 20pt \textgreek{Tέχνη}
+          \textonehalf\,\,training, \kern 35pt \textonehalf \kern 20pt \textgreek{Tέχνη}
           }{
@@
   {\small
-  \textbf{R. Di Remigio}, R. Bast, L. Frediani, and T. Saue
-  \\
-  \textsf{
-  Four-Component Relativistic Calculations in Solution with the
+  \textsf{Four-Component Relativistic Calculations in Solution with the
   Polarizable Continuum Model of Solvation: Theory,
@@
   }
+  \\
+  \textbf{R. Di Remigio}, R. Bast, L. Frediani, and T. Saue
   \\
@@
   {\small
+  \textsf{Wavelet Formulation of the Polarizable Continuum Model. II. Use of
+  Piecewise Bilinear Boundary Elements
+  }
+  \\
   M. Bugeanu, \textbf{R. Di Remigio}, K. Mozgawa, S. S. Reine, H.
   Harbrecht,  and L. Frediani
-  \\
-  \textsf{
-  Wavelet Formulation of the Polarizable Continuum Model. II. Use of
-  Piecewise Bilinear Boundary Elements
-  }
   \\
@@
   {\small
+  \textsf{A Polarizable Continuum Model for Molecules at Spherical
+  Diffuse Interfaces
+  }
+  \\
   \textbf{R. Di Remigio}, K. Mozgawa, H. Cao, V. Weijo, and L.
   Frediani
-  \\
-  \textsf{
-  A Polarizable Continuum Model for Molecules at Spherical
-  Diffuse Interfaces
-  }
   \\
@@
   {\small
-  \textbf{R. Di Remigio}, M. Repisky, S. Komorovsky, P. Hrobarik, L.
-  Frediani, and K. Ruud
-  \\
-  \textsf{
-  Four-Component Relativistic Density Functional Theory with the
+  \textsf{Four-Component Relativistic Density Functional Theory with the
   Polarizable Continuum Model: Application to EPR Parameters
@@
   }
+  \\
+  \textbf{R. Di Remigio}, M. Repisky, S. Komorovsky, P. Hrobarik, L.
+  Frediani, and K. Ruud
   \\
@@
   {\small
+  \textsf{Open-Ended Formulation of Self-Consistent Field Response Theory with
+  the Polarizable Continuum Model for Solvation
+  }
+  \\
   \textbf{R. Di Remigio}, M. T. P. Beerepoot, Y. Cornaton, M. Ringholm,
   A. H. S. Steindal, K. Ruud, and L. Frediani
-  \\
-  \textsf{
-  Open-Ended Formulation of Self-Consistent Field Response Theory with
-  the Polarizable Continuum Model for Solvation
-  }
   \\

### 2016-09-30 FrontBackmatter/Publications.tex
Fine-tuning
Minor polish

@@
 
-\begin{enumerate}[label=\paper{\Roman{*}}, ref=\Roman{*}]
-\item
-  \textbf{R. Di Remigio}, R. Bast, L. Frediani, and T. Saue
-  \\
-  \textsf{
+\renewcommand{\descriptionlabel}[1]{\hspace*{\labelsep}#1}
+\begin{description}[leftmargin=2mm, style=nextline]
+\item[\paper{I}]
+  {\small\textsf{
   Four-Component Relativistic Calculations in Solution with the
@@
   \\
+  \textbf{R. Di Remigio}, R. Bast, L. Frediani, and T. Saue
+  \\
 \textit{J. Phys. Chem. A}, \textrm{2015}, \textbf{119}, 5061--5077
@@
   DOI: \url{10.1063/1.4943782}
+  }
 \label{relapcm}
 
-\item
-  M. Bugeanu, \textbf{R. Di Remigio}, K. Mozgawa, S. S. Reine, H.
-  Harbrecht,  and L. Frediani
-  \\
-  \textsf{
+\item[\paper{II}]
+  {\small\textsf{
   Wavelet Formulation of the Polarizable Continuum Model. II. Use of
@@
   }
+  \\
+  M. Bugeanu, \textbf{R. Di Remigio}, K. Mozgawa, S. S. Reine, H.
+  Harbrecht,  and L. Frediani
   \\
@@
   DOI: \url{10.1039/C5CP03410H}
+  }
 \label{wemlin}
 
-\item
-  \textbf{R. Di Remigio}, K. Mozgawa, H. Cao, V. Weijo, and L.
-  Frediani
-  \\
-  \textsf{
+\item[\paper{III}]
+  \small{\textsf{
   A Polarizable Continuum Model for Molecules at Spherical
@@
   \\
+  \textbf{R. Di Remigio}, K. Mozgawa, H. Cao, V. Weijo, and L.
+  Frediani
+  \\
   \textit{J. Chem. Phys.}, \textrm{2016}, \textbf{144}, 124103
@@
   DOI: \url{10.1063/1.4943782}
+  }
 \label{spherical}
 
-\item
-  \textbf{R. Di Remigio}, M. Repisky, S. Komorovsky, P. Hrobarik, L.
-  Frediani, and K. Ruud
-  \\
-  \textsf{
+\item[\paper{IV}]
+  \small{\textsf{
   Four-Component Relativistic Density Functional Theory with the
@@
   \\
+  \textbf{R. Di Remigio}, M. Repisky, S. Komorovsky, P. Hrobarik, L.
+  Frediani, and K. Ruud
+  \\
   Accepted for publication in \textit{Mol. Phys.}
@@
   DOI: \url{10.1080/00268976.2016.1239846}
+  }
 \label{pcmepr}
 
-\item
-  \textbf{R. Di Remigio}, M. T. P. Beerepoot, Y. Cornaton, M. Ringholm,
-  A. H. S. Steindal, K. Ruud, and L. Frediani
-  \\
-  \textsf{
+\item[\paper{V}]
+  {\small\textsf{
   Open-Ended Formulation of Self-Consistent Field Response Theory with
@@
   \\
+  \textbf{R. Di Remigio}, M. T. P. Beerepoot, Y. Cornaton, M. Ringholm,
+  A. H. S. Steindal, K. Ruud, and L. Frediani
+  \\
   In preparation for submission to \textit{Phys. Chem. Chem. Phys.}
+  }
 \label{pcmopenrsp}
-\end{enumerate}
+\end{description}
+\renewcommand{\descriptionlabel}[1]{\hspace*{\labelsep}\spacedlowsmallcaps{#1}}

### 2016-10-01 Chapters/ResponseTheory.tex
Change colorscheme and extract only neede stuff from classicthesis

@@
 dielectric spectrum, \ie the static $\diels$ and optical
-dielectric constants $\dield$.
+dielectric constants $\dield$.\autocite{Aguilar1993-pu, Cammi1995-ah,
+Cammi2005-tb, Corni2005-nd}
 

### 2016-10-01 FrontBackmatter/Colophon.tex
Titlepage done, microtype is in (thanks LuaLaTeX), text area similar to Tufte
Change colorscheme and extract only neede stuff from classicthesis

@@
 \section*{Colophon}
-This document was typeset using \XeLaTeX, with the
-\texttt{classicthesis} template developed by André Miede.\footnote{\url{https://bitbucket.org/amiede/classicthesis/}}
+This document was typeset using \LuaLaTeX, with a modified version of
+the \texttt{classicthesis} template developed by André
+Miede.\footnote{\url{https://bitbucket.org/amiede/classicthesis/}}
 The style was inspired by Robert Bringhurst's seminal book on typography
-\emph{The Elements of Typographic Style}.\autocite{Bringhurst2004-wn}
+\emph{The Elements of Typographic Style}\autocite{Bringhurst2004-wn}
+and Edward Tufte's work.\autocite{Tufte1983-pr, Tufte1994-wf,
+Tufte1997-fx, Tufte2006-ui}
 The bibliography was processed using
@@
 \texttt{Bitstream Vera Mono} is used for monospaced text.
-Finally, XITS acts as the math typeface.
+Finally, XITS is the typeface for mathematical formulas.
 

### 2016-10-02 Chapters/ContinuumSolvationModels.tex
Change colorscheme and extract only neede stuff from classicthesis
Small font change

@@
 Developments of the \gls{PCM}, a continuum solvation model, are at the
-heart of this thesis.
+heart of this dissertation.
 This Chapter is a brief exposition of the major points of the \acrshort{PCM}
@@
 mentioned in the Introduction, namely:
-\begin{enumerate*}[label={\alph*)},font={\color{PMS1797}}]
+\begin{enumerate}
 \item the statistical averaging of environment configurations and
 \item the \acrshort{MM} region electrostatics cutoff choice.
-\end{enumerate*}
+\end{enumerate}
 
@@
 The need for ensemble averages leads us to the following observations:
-\begin{enumerate}%[label={\alph*)},font={\color{PMS1797}}]
+\begin{enumerate}
  \item
@@
 
-In this thesis, we focused exclusively on the purely electrostatic
+In this dissertation, we focused exclusively on the purely electrostatic
 component of the interaction energy, completely neglecting the other
@@
 numerical technique: the \gls{BEM}. This is implemented in four steps:
-\begin{enumerate}%[label={\alph*)},font={\color{PMS1797}}]
+\begin{enumerate}
     \item define the molecular cavity
@@
 three types of molecular surfaces can be defined:
-\begin{enumerate}%[label={\alph*)},font={\color{PMS1797}}]
+\begin{enumerate}
   \item
@@
 Mesh generation for the molecular surfaces mentioned above is usually
-achieved by modified triangulation procedures. In this thesis we used
+achieved by modified triangulation procedures. In this dissertation we used
 the GePol algorithm\autocite{Pascual-Ahuir1987-uo, Pascual-Ahuir1990-lp,
@@
 
-The wavelet Galerkin \acrshort{BEM} was also used in this thesis. This method
+The wavelet Galerkin \acrshort{BEM} was also used in this dissertation. This method
 preserves the fundamental symmetries of the boundary integral operators
@@
 
-In this thesis we advocate a different theoretical approach: the
+In this dissertation we advocate a different theoretical approach: the
 variational formulation of classical polarizable models.
@@
 properties:\autocite{Solis2013-ef}
-\begin{enumerate}%[label={\alph*)},font={\color{PMS1797}}]
+\begin{enumerate}
     \item physical extremal points,

### 2016-10-02 Chapters/ContributionsSummary.tex
Change colorscheme and extract only neede stuff from classicthesis
Small font change

@@
 This final Chapter provides a brief overview of the motivations, results and
-conclusions of the papers this thesis is based on.
-All publications included in this thesis have involved some programming effort.
+conclusions of the papers this dissertation is based on.
+All publications included in this dissertation have involved some programming effort.
 In Section \ref{sec:software} I will describe the principles we have
@@
 Schematic view of a \acrshort{PCM}-\acrshort{SCF} algorithm. Computations/data in
-\textcolor{PMS2229}{blue} are implemented by the \acrshort{PCM} \acrshort{API}.
+\textcolor{brewerBlue}{blue} are implemented by the \acrshort{PCM} \acrshort{API}.
 Computations/data in \textcolor{Green}{green} are implemented by the
@@
 
-Programming effort in this thesis has not only been directed at the
+Programming effort in this dissertation has not only been directed at the
 creation of the \pcmsolver library, but also at interfacing it with a
@@
 Coupling with \psicode\autocite{Turney2012-de} has allowed us to implement the
-\acrshort{PCM}-\acrshort{CC} developments described in this thesis.
+\acrshort{PCM}-\acrshort{CC} developments described in this dissertation.
 Finally, development of an interface with the KOALA code is underway.\autocite{Hofener2014-ex, Hofener2016-qz}
@@
 \pcmsolver by providing the implementation and testing of:
-\begin{enumerate}%[label={\alph*)},font={\color{PMS1797}}]
+\begin{enumerate}
   \item \acrshort{MEP} integrals for four-component wave functions,

### 2016-10-02 Chapters/QuantumMechanics.tex
Small font change

@@
 This Chapter presents a brief overview of molecular electronic structure
-theory with particular emphasis on the methods used in this thesis.
+theory with particular emphasis on the methods used in this dissertation.
 Section \ref{sec:mqm} is a general introduction to the methods of
@@
 efficiency.\autocite{Helgaker2000-tz}
-The usual choice, also adopted in this thesis, is to use bases of
+The usual choice, also adopted in this dissertation, is to use bases of
 Gaussian functions.\autocite{Helgaker1995-gi, Reine2012-tq}

### 2016-10-02 FrontBackmatter/Introduction.tex
Fine-tuning
Small font change

@@
 I will put emphasis on the methods that have been relevant in the work
-presented in this thesis. Section \ref{sec:mqm} discusses some of the
+presented in this dissertation. Section \ref{sec:mqm} discusses some of the
 ideas of quantum many-body theory relevant to quantum chemistry. Section
@@
 that has had a lasting impact and informs much of the developments that
-will be presented in this thesis.\autocite{Onsager1936-wf}
+will be presented in this dissertation.\autocite{Onsager1936-wf}
 
@@
 Environment effects are usually classified as:
-\begin{description}[font=\normalfont\scshape]
+\begin{description}[leftmargin=2mm, font=\normalfont\scshape]
 \item[Direct.]
@@
 \centering
-\includegraphics[width=.7\textwidth]{multiscale_models.png}
+\includegraphics[width=\textwidth]{multiscale_models-landscape.png}
 \caption[From quantum mechanical to multiscale models]{
@@
 Cluster models can be used to benchmark more approximate multiscale
-models, but their description is outside the scope of this thesis.
+models, but their description is outside the scope of this dissertation.
 
@@
 
-In this thesis, we are interested in the inclusion of
+In this dissertation, we are interested in the inclusion of
 environment effects in \emph{ab initio} models of interesting chemical

### 2016-10-02 FrontBackmatter/Introduction.tex
Add a couple of slick macros

@@
 
-This question was posed by R.~S.~Mulliken over a half-century
-ago\footnote{This was the title Mulliken chose for his Gilbert N.~Lewis
+\beginningthought{This question was posed} by R.~S.~Mulliken over a half-century
+ago\footnote{Mulliken chose it as the title for his Gilbert N.~Lewis
 award acceptance speech in 1960.}
@@
 The purpose of quantum chemistry is to provide models based on first
-principles that can help \emph{understand} and \emph{predict} molecular properties.
+principles that can help \emph{understand} and \emph{predict}
+macroscopic, observable chemical phenomena, such as reaction mechanisms and
+spectroscopic experiments.
 As stated by \citeauthor{Dirac1929-gn} in his \citetitle{Dirac1929-gn}
 paper:\autocite{Dirac1929-gn, Kutzelnigg2000-fl}
-\blockquote{The underlying physical laws necessary for the mathematical
+\blockquote{\small\textit{The underlying physical laws necessary for the mathematical
 theory of a large part of physics and the whole of chemistry are thus
@@
 explanation of the main features of complex atomic systems without too
-much computation.}
+much computation.}}
 Following Dirac's \emph{dictum}, in quantum chemistry we apply physical
@@
 
-It is easy to manipulate chemicals in a virtual laboratory. Nowadays
+\newthought{It is easy to manipulate chemicals in a virtual laboratory.} Nowadays
 quantum chemical methods often complement traditional experimental

### 2016-10-02 FrontBackmatter/Introduction.tex
Going throgh the introduction

@@
 
-\beginningthought{This question was posed} by R.~S.~Mulliken over a half-century
+\begthought{This question was posed} by R.~S.~Mulliken over a half-century
 ago\footnote{Mulliken chose it as the title for his Gilbert N.~Lewis
@@
 macroscopic, observable chemical phenomena, such as reaction mechanisms and
-spectroscopic experiments.
+spectroscopic experiments.\autocite{Kovac2011-ew}
 As stated by \citeauthor{Dirac1929-gn} in his \citetitle{Dirac1929-gn}
@@
 chemical systems: model building and software implementations are two
-closely intertwined aspects in the practice of quantum chemistry.
+closely intertwined aspects in the practice of quantum
+chemistry.\autocite{Winsberg2010-sy}
 
@@
 \centering
-\includegraphics[width=.6\textwidth]{qc-axes.png}
+\includegraphics[width=.8\textwidth]{qc-axes.png}
 \caption[Pictorial depiction of the concept of \emph{model chemistries}.]{
@@
 
-\newthought{It is easy to manipulate chemicals in a virtual laboratory.} Nowadays
+\arrowthought{It is easy to manipulate chemicals in a virtual laboratory.} Nowadays
 quantum chemical methods often complement traditional experimental
@@
 
-Chemistry can be largely considered a \emph{wet} science: almost always
+\begarrowthought{Chemistry can be largely considered a \emph{wet} science}: almost always
 chemical phenomena happen in a liquid environment.\autocite{Reichardt2010-le}
-We hereby define a "solution", or more generally an "environment", as
+We hereby define a \enquote{solution}, or more generally an
+\enquote{environment}, as
 a system where the number of solvent molecules exceeds by far the number
@@
 To tame this complexity and cure the disease, one must devise
-\emph{models} that simplify physical picture, while offering tools for
+\emph{models} that simplify the physical picture, while offering tools for
 understanding reality and predicting new and exciting
@@
 
-Before introducing our model of choice, let us consider how an
+\newthought{Before introducing our model of choice}, let us consider how an
 environment might affect molecular observables of interest.
 Environment effects are usually classified as:
-\begin{description}[leftmargin=2mm, font=\normalfont\scshape]
+\begin{description}[leftmargin=2mm, font=\normalfont]
 \item[Direct.]
@@
   The presence of the environment radically influences excited states,
-  since relaxation processes in the medium become important.
+  since relaxation processes in the medium become
+  important.\autocite{Aguilar1993-pu, Cammi1995-ah}
 \item[Specific.] This catch-all category includes all effects
@@
 parts of the system with different methods and couple these methods by
-bridging "scales" at the boundary.
+bridging \enquote{scales} at the boundary.
 Figure \ref{fig:qm-to-multiscale} schematically portrays the transition
@@
 \caption[From quantum mechanical to multiscale models]{
-From quantum mechanical to multiscale models. Picture reproduced
-courtesy of Dr.~Stefano Caprasecca (MoLEcoLab, Università di Pisa).}
+Pictorial depiction of the transition from quantum mechanical to
+multiscale models for the aqueous solvation of \emph{para}-nitroaniline.
+Leftmost panel: a fully quantum mechanical cluster model.
+Upper central panel: a quantum/classical discrete (explicit) model.
+Lower central panel: a quantum/classical continuum (implicit) model.
+Rightmost panel: a quantum/classical mixed explicit/implicit model.
+Picture reproduced courtesy of Dr. Stefano Caprasecca (MoLEcoLab, Università di Pisa).}
 \label{fig:qm-to-multiscale}
@@
 system.
-However, they demand a statistical average of environment configurations
+However, they demand statistical averaging of environment configurations
 to yield results of any significance. Moreover, a rather large
@@
 \acrshort{QM}/Continuum, \acrshort{QM}/\acrshort{MM} and \acrshort{QM}/\acrshort{MM}/Continuum models that will
-be extensively used throughout the thesis.
+be extensively used throughout.
 
@@
 
-The scope of quantum chemistry is to understand and predict molecular
-properties, but so far we have made no mention of properties other than
+\begarrowthought{The scope of quantum chemistry is to understand and predict molecular
+properties}, but so far we have made no mention of properties other than
 the energy.
@@
 
-The concepts of systematic improvability and theoretical model
-chemistries are the foundations for the successful practice of quantum
+\begarrowthought{The concepts of systematic improvability and theoretical model
+chemistries} are the foundations for the successful practice of quantum
 chemistry.

### 2016-10-03 FrontBackmatter/Notation.tex
First draft sent to L.F. and B.M.
Add cover and modify Makefile to generate the complete pdf
Polish
Going throgh the introduction
Finally get openright, right

@@
 
+\phantomsection
+\refstepcounter{dummy}
+\addcontentsline{toc}{chapter}{\tocEntry{\nomname}}
+
 \renewcommand{\nompreamble}{
@@
 order to maintain consistency throughout. We summarize them here.
-Hartree atomic units are used throughout:~\autocite{Whiffen1978-xx, Mohr2012-zc}
+Hartree atomic units are used throughout:\autocite{Whiffen1978-xx, Mohr2012-zc}
 \begin{equation*}
@@
 instead of a star ($*$).
+
+This document makes use of hyperlinks within and without to ease reading
+and navigation. A numbered citation style was chosen.
+Citations are color-coded in \textcolor{brewerCitation}{green}, appear
+(most often) as superscripts and are hyperlinked to the bibliography.
+Acronyms are color-coded in \textcolor{brewerGray}{gray} and
+hyperlinked to the list of acronyms.
+Finally, URLs and placeholders for URLs are color-coded in
+\textcolor{brewerURL}{orange}.
+They are hyperlinked to the respective online resource.
 }
-
-\refstepcounter{dummy}
-\phantomsection
-\addcontentsline{toc}{chapter}{\tocEntry{\nomname}}
 
@@
 \nomenclature[A, 02]{$\Span{\vect{v}_1, \ldots, \vect{v}_n}$}{Vector space spanned by the vectors $\vect{v}_1, \ldots, \vect{v}_n$}
-\nomenclature[A, 03]{$\delta_{ij}$}{Kronecker symbol: $\delta_{ij}=1i$ if $i=j$ and 0 otherwise}
+\nomenclature[A, 03]{$\delta_{ij}$}{Kronecker symbol: $\delta_{ij}=1$ if $i=j$ and $0$ otherwise}
 \nomenclature[A, 04]{$\scalprod[X]{u}{v}$}{Scalar product of $u$ and $v$ in the Hilbert space $X$}
@@
 % Function Spaces
-\nomenclature[D, 02]{$\mathcal{C}^0(\Omega)$}{Space of continuous functions on $\Omega\subset\mathbb{R}^d$}
-\nomenclature[D, 03]{$\mathcal{C}^k(\Omega)$}{Space of $k$ times continuously differentiable functions on $\Omega$}
-\nomenclature[D, 04]{$L^p(\Omega)$}{Space of functions whose $p$-th power is Lebesgue integrable on $\Omega$}
-\nomenclature[D, 05]{$\norm{u}{0, p, \Omega}$}{Norm in $L^p(\Omega)$: $\norm{u}{0, p, \Omega} = (\int_\Omega\diff\vect{r} |u(\vect{r})|^p)^\frac{1}{p}$}
-\nomenclature[D, 06]{$\norm{u}{\Omega}$}{Norm in $L^2(\Omega)$: $\norm{u}{\Omega} = \norm{u}{0, 2,\Omega} = \sqrt{\int_\Omega\diff\vect{r} |u(\vect{r})|^2}$}
-\nomenclature[D, 07]{$H^s(\Omega)$}{Sobolev space of order $s$ of functions whose derivative up to order $s$ are in $L^2(\Omega)$}
-\nomenclature[D, 08]{$H_0^s(\Omega)$}{Space of functions in $H^2(\Omega)$ with null trace on the boundary of $\Omega$}
-\nomenclature[D, 09]{$|u|_{s, \Omega}$}{Seminorm in $H^s(\Omega)$: $|u|_{s, \Omega} = \sum_{|\alpha|=s}\norm{\partial^\alpha u}{\Omega}$}
-\nomenclature[D, 10]{$\norm{u}{s, \Omega}$}{Norm in $H^s(\Omega)$: $\norm{u}{s, \Omega} = \sum_{l \leq s} |u|_{l, \Omega}$}
+%\nomenclature[D, 02]{$\mathcal{C}^0(\Omega)$}{Space of continuous functions on $\Omega\subset\mathbb{R}^d$}
+%\nomenclature[D, 03]{$\mathcal{C}^k(\Omega)$}{Space of $k$ times continuously differentiable functions on $\Omega$}
+%\nomenclature[D, 04]{$L^p(\Omega)$}{Space of functions whose $p$-th power is Lebesgue integrable on $\Omega$}
+%\nomenclature[D, 05]{$\norm{u}{0, p, \Omega}$}{Norm in $L^p(\Omega)$: $\norm{u}{0, p, \Omega} = (\int_\Omega\diff\vect{r} |u(\vect{r})|^p)^\frac{1}{p}$}
+%\nomenclature[D, 06]{$\norm{u}{\Omega}$}{Norm in $L^2(\Omega)$: $\norm{u}{\Omega} = \norm{u}{0, 2,\Omega} = \sqrt{\int_\Omega\diff\vect{r} |u(\vect{r})|^2}$}
+%\nomenclature[D, 07]{$H^s(\Omega)$}{Sobolev space of order $s$ of functions whose derivative up to order $s$ are in $L^2(\Omega)$}
+%\nomenclature[D, 08]{$H_0^s(\Omega)$}{Space of functions in $H^2(\Omega)$ with null trace on the boundary of $\Omega$}
+%\nomenclature[D, 09]{$|u|_{s, \Omega}$}{Seminorm in $H^s(\Omega)$: $|u|_{s, \Omega} = \sum_{|\alpha|=s}\norm{\partial^\alpha u}{\Omega}$}
+%\nomenclature[D, 10]{$\norm{u}{s, \Omega}$}{Norm in $H^s(\Omega)$: $\norm{u}{s, \Omega} = \sum_{l \leq s} |u|_{l, \Omega}$}
 % Molecular Quantum Mechanics

### 2016-10-03 Chapters/ContinuumSolvationModels.tex
Always link acronyms back to list of acronyms

@@
    \includegraphics[width=.85\textwidth]{vdWS.png}
-   \caption{\acrlong*{vdWS} (\acrshort{vdWS})}
+   \caption{\acrlong{vdWS} (\acrshort{vdWS})}
    \label{fig:vdWS}
@@
    \includegraphics[width=.85\textwidth]{SAS.png}
-   \caption{\acrlong*{SAS} (\acrshort{SAS})}
+   \caption{\acrlong{SAS} (\acrshort{SAS})}
    \label{fig:SAS}
@@
    \includegraphics[width=.85\textwidth]{SES.png}
-   \caption{\acrlong*{SES} (\acrshort{SES})}
+   \caption{\acrlong{SES} (\acrshort{SES})}
    \label{fig:SES}
@@
 \end{equation}
-where $C_{\alpha r}$ are the \acrshort*{MO} expansion coefficients and
+where $C_{\alpha r}$ are the \acrshort{MO} expansion coefficients and
 $N_O$ is the total number of occupied orbitals.
-The \acrshort*{SCF} energy functional \emph{in vacuo} can now be
+The \acrshort{SCF} energy functional \emph{in vacuo} can now be
 expressed as:
@@
 \end{equation}
-and has to be optimized under the \acrshort*{MO} orthonormality
+and has to be optimized under the \acrshort{MO} orthonormality
 constraint, easily introduced by means of Lagrangian multipliers:
@@
 \end{equation}
-The various contributions are expressed in terms of \acrshort*{AO} basis
+The various contributions are expressed in terms of \acrshort{AO} basis
 integral and density matrices:
@@
 \end{subequations}
-The quantum/classical polarizable \acrshort*{SCF} functional is obtained by adding the
+The quantum/classical polarizable \acrshort{SCF} functional is obtained by adding the
 polarization functional to Eq. \eqref{eq:scf-functional}:
@@
 Differentiation with respect to the variational degrees of freedom
-yields the \acrshort*{KS} and polarization equations:
+yields the \acrshort{KS} and polarization equations:
 \begin{subequations}
@@
 \end{subequations}
-The \acrshort*{KS} matrix now includes environment contributions:
+The \acrshort{KS} matrix now includes environment contributions:
 \begin{equation}
@@
 With respect to the coupling described in abstract terms above, the
-\acrshort*{SCF} case is formally simpler. In fact, whereas in the
+\acrshort{SCF} case is formally simpler. In fact, whereas in the
 general case the source term depends on \emph{both} primal (amplitudes)
 and dual (multipliers) variational degrees of freedom, here only the
-dependence on the primal parameters (\acrshort*{MO} coefficients) needs
+dependence on the primal parameters (\acrshort{MO} coefficients) needs
 to be considered.\autocite{Lipparini2011-aj}

### 2016-10-03 Chapters/ContributionsSummary.tex
Always link acronyms back to list of acronyms

@@
 functions for spherical diffuse interfaces\autocite{DiRemigio2016-nn} and
-the delayed solvers for real-time \acrlong*{TDSCF}
+the delayed solvers for real-time \acrlong{TDSCF}
 simulations.\autocite{Corni2015-pe}

### 2016-10-03 Chapters/MathematicalResults.tex
Always link acronyms back to list of acronyms

@@
 
-The \acrshort*{CCSD} Lagrangian is then:
+The \acrshort{CCSD} Lagrangian is then:
 \begin{equation}\label{eq:ccsd-lagrangian}
@@
 expression in the \acrshort{CCSD} Lagrangian.
-This is a consequence of the well-known fact that the \acrshort*{CC} energy can
+This is a consequence of the well-known fact that the \acrshort{CC} energy can
 be expressed purely in terms of $T_1$ and $T_2$:
@@
 
-The use of normal-ordered operators is common in \acrlong*{CC} theory. Both for
+The use of normal-ordered operators is common in \acrlong{CC} theory. Both for
 the dressed and bare representations, the expectation values of such operators
@@
 \end{equation}
-where the function $\sigma$ is the, yet unknown, \acrlong*{ASC}.
+where the function $\sigma$ is the, yet unknown, \acrlong{ASC}.
 

### 2016-10-03 Chapters/QuantumMechanics.tex
Finally get openright, right
Always link acronyms back to list of acronyms

@@
 
-\chapter{A Quick Tour of Molecular Electronic Structure Theory}\label{ch:QM}
+\chapter{A Tour of Molecular Electronic Structure Theory}\label{ch:QM}
 
@@
 in addition to the other terms contained in the Fock matrix.
-The functional derivative of the \acrshort*{XC} functional enters in the
+The functional derivative of the \acrshort{XC} functional enters in the
 definition of this additional term:
@@
 Notice moreover the exact exchange admixing factor $\gamma$ in the
-two-electron part of the \acrshort*{KS} matrix.
+two-electron part of the \acrshort{KS} matrix.
 
@@
 \end{equation}
-and optimize the expansion coefficients in the \acrshort*{SCF} procedure.
-Choosing the one-particle \acrshort*{MO} space is then equivalent to
+and optimize the expansion coefficients in the \acrshort{SCF} procedure.
+Choosing the one-particle \acrshort{MO} space is then equivalent to
 choosing the one-particle \gls{AO} space.
-The functional form of the \acrshort*{AO} basis is chosen to strike a
+The functional form of the \acrshort{AO} basis is chosen to strike a
 balance between a physically accurate description of the short-range
@@
 
-In the \acrlong*{CC} method we construct our correlated wave function as
+In the \acrlong{CC} method we construct our correlated wave function as
 an exponential mapping on top of a single reference
@@
 \end{subequations}
-The \acrshort*{CC} Lagrangian can thus be expressed as:
+The \acrshort{CC} Lagrangian can thus be expressed as:
 \begin{equation}\label{eq:cc-lagrangian}
@@
 energies and the shorthand notation for the Lagrangian multiplier state (left
-\acrshort*{CC} state) was introduced:
+\acrshort{CC} state) was introduced:
 \begin{equation}
@@
 as zeroth-order wave function and the perturbation series is developed
-considering the fluctuation potential in the \acrlong*{MP} partitioning
+considering the fluctuation potential in the \acrlong{MP} partitioning
 of the Hamiltonian as the perturbation.\autocite{McWeeny1992-oj, Helgaker2000-tz}
@@
 
-As shown by \citeauthor{Koch1997-nm}, the \acrshort*{MBPT} series can be
-derived from a \acrshort*{CC} ansatz for the wave function.
+As shown by \citeauthor{Koch1997-nm}, the \acrshort{MBPT} series can be
+derived from a \acrshort{CC} ansatz for the wave function.
 This ensures term-by-term size-extensivity of the energy at all
@@
 terms of the fluctuation potential. Energy corrections are identified
-from the corresponding order expansion of the \acrshort*{CC} Lagrangian
+from the corresponding order expansion of the \acrshort{CC} Lagrangian
 and retaining only terms complying to the $2n+1$ and $2n+2$ rules.
@@
 These expressions form the basis for the development of reduced-scaling
-\acrshort*{CC} schemes were the double and triple excitations are treated
+\acrshort{CC} schemes were the double and triple excitations are treated
 approximately. These are the subject of the next Section.
@@
 
-The \acrshort*{CCSD} scheme, which scales as $N^6$, is really successful
+The \acrshort{CCSD} scheme, which scales as $N^6$, is really successful
 in recovering a large portion of the correlation energy, but as hinted
@@
 The assumption that $T_1$ is of zeroth order is contrary to the usual
-\acrshort*{MP} approach but highlights the role of the singles
+\acrshort{MP} approach but highlights the role of the singles
 amplitudes as orbital optimization parameters.
@@
 and we shall refer to it as quasi-$T_2$.
-The \acrshort*{CC2} Lagrangian is constructed based on the approximate
+The \acrshort{CC2} Lagrangian is constructed based on the approximate
 equations obtained for the singles and
@@
 \end{equation}
-The \acrshort*{CC2} method brings the scaling of the \acrshort*{CCSD}
+The \acrshort{CC2} method brings the scaling of the \acrshort{CCSD}
 method down to $N^5$ in the iterative step.
-From the approximate \acrshort*{CC2} Lagrangian we can derive the
+From the approximate \acrshort{CC2} Lagrangian we can derive the
 multipliers equations:
@@
 \end{subequations}
-Formally, \acrshort*{CC2} expectation values can be computed using the
-expression for the \acrshort*{CCSD} (see Appendix \ref{sec:CC-expval})
+Formally, \acrshort{CC2} expectation values can be computed using the
+expression for the \acrshort{CCSD} (see Appendix \ref{sec:CC-expval})
 and the approximate amplitudes and multipliers:
@@
 The \gls{CC3} model is constructed analogously. Starting from the
-\acrshort*{CCSDT} amplitudes equations in the $T_1$-transformed
+\acrshort{CCSDT} amplitudes equations in the $T_1$-transformed
 representation, we retain only second order terms in the triples
@@
 \end{subequations}
-In complete analogy with \acrshort*{CC2}, the quasi-$T_3$ operator was introduced:
+In complete analogy with \acrshort{CC2}, the quasi-$T_3$ operator was introduced:
 \begin{equation}
@@
 \end{equation}
-The \acrshort*{CC3} Lagrangian is constructed based on the approximate
+The \acrshort{CC3} Lagrangian is constructed based on the approximate
 amplitude equations
@@
 \end{equation}
-The \acrshort*{CC3} methods brings the scaling of the \acrshort*{CCSDT}
+The \acrshort{CC3} methods brings the scaling of the \acrshort{CCSDT}
 method down to $N^7$ in the iterative step.
-We can now derive the \acrshort*{CC3} multipliers equations:
+We can now derive the \acrshort{CC3} multipliers equations:
 \begin{subequations}\label{eq:cc3-multipliers}
@@
 \end{subequations}
-\acrshort*{CC3} expectation values can be computed as follows:
+\acrshort{CC3} expectation values can be computed as follows:
 \begin{equation}\label{eq:cc3-expectation-value}
@@
 calculated as a one-off correction on top of the converged
-\acrshort*{CCSD} energy.
+\acrshort{CCSD} energy.
 The expression for the correction is based on the fourth and fifth order
@@
 \eqref{eq:4th-order-triples}--\eqref{eq:5th-order-triples}, but is
-computed using the converged \acrshort*{CCSD} singles and doubles
+computed using the converged \acrshort{CCSD} singles and doubles
 amplitudes,\autocite{Urban1985-xp, Raghavachari1989-bn, Bartlett1990-lk,
@@
 superscript.
-Since the left and right \acrshort*{CC} states differ, symmetric and
+Since the left and right \acrshort{CC} states differ, symmetric and
 asymmetric variants of the noniterative corrections exist.
@@
 Results from noniterative methods can be rationalized within the
-framework of \acrshort*{MBPT}.\autocite{Stanton1997-lx, Koch1997-nm,
+framework of \acrshort{MBPT}.\autocite{Stanton1997-lx, Koch1997-nm,
 Eriksen2014-gy, Eriksen2015-il, Kristensen2016-od}
 From Eqs. \eqref{eq:1st-order}--\eqref{eq:2nd-order}, one can derive
-which \acrshort*{MBPT} terms are included in the corrections:
+which \acrshort{MBPT} terms are included in the corrections:
 \begin{equation}
@@
 and similarly for ${}^* E^{[4]}_{\text{[T]}}$.
-The \acrshort*{MBPT} terms included in ${}^*
+The \acrshort{MBPT} terms included in ${}^*
 E^{[5]}_{\Lambda\text{(T)}}$ can be seen to be:

### 2016-10-03 Chapters/ResponseTheory.tex
Always link acronyms back to list of acronyms

@@
 By formulating the perturbation-strenght derivative of the time-averaged
-quasienergy Lagrangian in terms of the \acrshort*{AO} density matrix,
+quasienergy Lagrangian in terms of the \acrshort{AO} density matrix,
 the authors were able to formulate arbitrary order response functions
-in the \acrshort*{AO} basis.
+in the \acrshort{AO} basis.
 The density matrix is the variational degree of freedom and the response
@@
 equation, the stationarity condition on the density matrix.
-The approach lends itself to an efficient \acrshort*{AO}-based
+The approach lends itself to an efficient \acrshort{AO}-based
 implementation,\autocite{Larsen2000-hj, Kjaergaard2008-hy} can fully
@@
 Friese2015-kb, Friese2015-bu}
-We will base our derivation of \acrshort*{SCF} response theory for
+We will base our derivation of \acrshort{SCF} response theory for
 quantum/classical polarizable Hamiltonians of this framework.
@@
 Lipparini2012-hx, Lipparini2012-tl}
-and general \acrshort*{QM}/\acrshort*{MM}/Continuum models.\autocite{Steindal2011-ki, Caprasecca2012-ir, Lipparini2013-ud}
+and general \acrshort{QM}/\acrshort{MM}/Continuum models.\autocite{Steindal2011-ki, Caprasecca2012-ir, Lipparini2013-ud}
 
@@
 response functions, when the fixed-cavity approximation for the
-\acrshort*{PCM} is assumed.
+\acrshort{PCM} is assumed.
 Here we report a summary of our findings in \paper{V} employing the
@@
 To keep track of the number of chain rule applications,
-we will denote higher-order derivatives of the \acrshort*{KS}
+we will denote higher-order derivatives of the \acrshort{KS}
 generalized free energy with the composite ${mn, abc\ldots}$
@@
 This goes under the name of \emph{fixed-cavity approximation} in the
-context of the \acrshort*{PCM}.\autocite{Cammi1994-qj}
+context of the \acrshort{PCM}.\autocite{Cammi1994-qj}
 Under this simplifying assumption, the perturbation-strength derivative
@@
 polarization equations \eqref{eq:part-polarization-eq}.
-The \acrlong*{RHS} in the response equation only includes terms that
+The \acrlong{RHS} in the response equation only includes terms that
 depend on particular contributions up to the desired order or
@@
 \end{equation}
-where $\mathbfcal{F}$ is the generalized \acrshort*{KS} matrix, see
+where $\mathbfcal{F}$ is the generalized \acrshort{KS} matrix, see
 \noparcite[ref.][]{Thorvaldsen2008-sg} for details.

### 2016-10-03 Chapters/SolvationAndCorrelation.tex
Change colorscheme and extract only neede stuff from classicthesis
Always link acronyms back to list of acronyms

@@
 Manipulation of similarity transformed operators involves the use of
-their \acrlong*{BCH} expansions:
+their \acrlong{BCH} expansions:
 \begin{equation}
@@
 $O$, see Appendix \ref{app:mathematical-results} for details.
-Finally, the $u$-th excited component of the \acrlong*{CC} left state
+Finally, the $u$-th excited component of the \acrlong{CC} left state
 will be denoted as:
@@
 ${}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})$ term to yield the
-\acrlong*{PTE} (\acrshort{PTE}) Hamiltonian -- $H$ -- and eventually the
-\acrlong*{PTED} (\acrshort{PTED}) Lagrangian:\autocite{Olivares_del_Valle1991-of,
+\acrlong{PTE} (\acrshort{PTE}) Hamiltonian -- $H$ -- and eventually the
+\acrlong{PTED} (\acrshort{PTED}) Lagrangian:\autocite{Olivares_del_Valle1991-of,
 Aguilar1991-vq, Olivares_del_Valle1991-tq, Olivares_del_Valle1993-xq,
@@
 Both the amplitudes and multipliers equations are formally identical to
-the \acrshort*{PTED} equations \eqref{eq:pted} but are coupled to
+the \acrshort{PTED} equations \eqref{eq:pted} but are coupled to
 approximate polarization equations.
 For the amplitudes equation \eqref{eq:ptes-cc-amplitudes}, one computes
-the polarization from a modified \acrshort*{CCS} expectation value of
+the polarization from a modified \acrshort{CCS} expectation value of
 the source term. This adds a negligible $O(N^5)$ term to the iterative
@@
 The amplitudes equations are instead coupled to the polarization
-obtained using the full \acrshort*{CCS} expectation value of the source
+obtained using the full \acrshort{CCS} expectation value of the source
 term. This adds another $O(N^5)$ step to the iterative
@@
 Shavitt2009-mr}
-This symmetry is a consequence of choosing a closed-shell \acrshort*{HF}
+This symmetry is a consequence of choosing a closed-shell \acrshort{HF}
 determinant as reference function\autocite{Shavitt2009-mr} and is
@@
 There are no explicit contributions from the triples and thus the
-\acrshort*{CCSD} polarization is correct through fourth order in the
+\acrshort{CCSD} polarization is correct through fourth order in the
 fluctuation potential.
@@
     +
-    \highlight{PMS1797}{
+    \highlight{brewerRed}{
     {}^t\p^{[1]}
@@
     T_2^{[2]}} | \mathrm{HF}} \\
-    &+ \highlight{PMS1797}{
+    &+ \highlight{brewerRed}{
     {}^t\p^{[2]}\braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[2]} + T_3^{[2]}} | \mathrm{HF}}
@@
     &+
-    \highlight{PMS1797}{
+    \highlight{brewerRed}{
     {}^t\p^{[1]}
@@
 will form the basis for the development of approximations
-beyond \acrshort*{CCSD} including connected triples.
+beyond \acrshort{CCSD} including connected triples.
 
@@
   G_T^{[4]} &=
-  \highlight{PMS130}{\braket{\tbar{1}^{[1]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}}
+  \highlight{brewerYellow}{\braket{\tbar{1}^{[1]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}}
   + \braket{\tbar{2}^{[1]}| \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}} \\
@@
   G_T^{[5]} &=
-    \highlight{PMS130}{
+    \highlight{brewerYellow}{
     \braket{\tbar{2}^{[1]} |
@@
   +
-  \highlight{PMS130}{
+  \highlight{brewerYellow}{
     \braket{\tbar{3}^{[2]} |
@@
 schemes for the approximate inclusion of higher order excitations in the
-\acrshort*{CC} hierarchy.
+\acrshort{CC} hierarchy.
 Starting from the \acrshort{CC2} Lagrangian, Eq.
@@
 quantum/classical polarizable Hamiltonians has already been presented in
-the context of the multipolar continuum and \acrlong*{PE}
+the context of the multipolar continuum and \acrlong{PE}
 models.\autocite{Kongsted2002-zm, Osted2003-qc, Sneskov2011-jm,
@@
 \end{subequations}
-In both \acrlong*{CC} models, the polarization is still determined by
+In both \acrlong{CC} models, the polarization is still determined by
 Eq. \eqref{eq:pte-s-polarization} and only the singles multipliers
@@
 Moreover, the simplified \acrshort{PE}-\acrshort{CC}
-(s\acrshort*{PE}\acrshort*{CC}) Lagrangian of \citeauthor{Krause2016-ee}
+(s\acrshort{PE}\acrshort{CC}) Lagrangian of \citeauthor{Krause2016-ee}
 is essentially the \acrshort{PTE(S)} approximation for

### 2016-10-03 FrontBackmatter/Introduction.tex
Simplify configuration and add higher resolution images in introduction
Some more repo clean up
Finally get openright, right
Always link acronyms back to list of acronyms

@@
 \centering
-\includegraphics[width=.8\textwidth]{qc-axes.png}
+\includegraphics[width=.8\textwidth]{model-chem.pdf}
 \caption[Pictorial depiction of the concept of \emph{model chemistries}.]{
@@
 \ref{sec:mean-field} introduces the workhorses of quantum chemistry
-\acrlong*{HF} and \acrlong*{KS} \acrlong*{DFT}.
-The \acrlong*{MBPT} and \acrlong*{CC} rungs on the methodology axis will
+\acrlong{HF} and \acrlong{KS} \acrlong{DFT}.
+The \acrlong{MBPT} and \acrlong{CC} rungs on the methodology axis will
 be discussed in Sections \ref{sec:mbpt} and \ref{sec:coupled-cluster},
@@
 
-\section*{The Problem of Solvation or Taming Complexity with Models}
+\section*{The Problem of Solvation}
 
@@
 \centering
-\includegraphics[width=\textwidth]{multiscale_models-landscape.png}
+\includegraphics[width=\textwidth]{multiscale_models-landscape.pdf}
 \caption[From quantum mechanical to multiscale models]{
@@
 Rightmost panel: a quantum/classical mixed explicit/implicit model.
-Picture reproduced courtesy of Dr. Stefano Caprasecca (MoLEcoLab, Università di Pisa).}
+Picture reproduced courtesy of Dr. Stefano Caprasecca
+(MoLEcoLab, Università di Pisa).}
 \label{fig:qm-to-multiscale}
@@
 
-While both models treat the molecular degrees of freedom at the quantum
+While both models treat the molecular degrees of freedom at the
+quantum
 mechanical level, their approach to the microscopic description of the
@@
    polarizable. While the former method allows for mutual polarization
-   between the \acrshort*{QM} and \acrshort*{MM} subsystems, the latter
-   treats the \acrshort*{MM} region as fixed.
+   between the \acrshort{QM} and \acrshort{MM} subsystems, the latter
+   treats the \acrshort{MM} region as fixed.
  \item
@@
 \Acrlong*{DFT} approaches are cheap and widespread, but their general
-accuracy is hard to assess. \Acrlong*{MBPT} and \acrlong*{CC} approaches
+accuracy is hard to assess. \Acrlong*{MBPT} and \acrlong{CC} approaches
 are more robust in this respect, albeit at an increased computational

### 2016-10-03 FrontBackmatter/Acknowledgments.tex
Luca's comments and corrections are in for the main body of the thesis
Going through typos and typography

@@
 
-\bigskip
-
-\begingroup
-\let\clearpage\relax
-\let\cleardoublepage\relax
-\let\cleardoublepage\relax
 \chapter*{Acknowledgments}
 
-\epigraph{
-Just a perfect day
+It has not been one day, it's been four years.\footnotetext{
+Just a perfect day \\
+You made me forget myself \\
+I thought I was \\
+Someone else, someone good \\
+\hspace*{15pt}
+--- \textsc{Lou Reed}}
+And it's not always been perfect.
+It is however true that the people I have met and interacted with during
+this adventure have made me feel \emph{someone good}.
+I don't feel it's an exaggeration to say that if I have made it this far
+is also because of \emph{you}.
+Thus, these few, poorly written lines are
+a \emph{dedication} of this dissertation to you
+and an \emph{acknowledgment}
+of the patience and love you have shown towards me.
+Thanks for not giving up on me.
 
-You made me forget myself
+First and foremost, \textitalian{ai miei genitori,
+Adriano e Vilma. Ai miei fratelli, Francesca e Simone. A mia zia, Marina
+ed ai miei nonni, Cherubino e Irma, Maria e Piero. Questo è dedicato a
+voi, per tutto quello che mi avete insegnato, per il sostegno continuo e
+la pazienza infinita che dimostrate ogni giorno.}
 
-I thought I was
-
-Someone else, someone good}{
---- \textsc{Lou Reed}}
-
-\endgroup
+in no particular order:\footnotetext{
+And all of my friends were there, \\
+and no one cared \\
+\hspace*{15pt}
+--- \textsc{The Kinks}}
+\begin{itemize}
+  \item To \emph{Radovan Bast}, \emph{il miglior fabbro},
+  \item To \emph{Davide Michetti}, the inventor of nicknames,
+  \item To \emph{The Little Lebowski Urban Achievers}, Erlend, Ana and
+    Torgeir,
+  \item To the \emph{Schrodinger killed my cat!}
+\end{itemize}

### 2016-10-03 FrontBackmatter/Introduction.tex
Going through typos and typography

@@
 
-\begin{epigraphs}
-\qitem{
+\section*{What Are the Electrons Really Doing in Molecules?}\footnotetext{
 [...] \textnorsk{og fordi jeg alltid har hatt en dragning mot det skjulte og hemmelige.}
-}{
---- \textnorsk{\textsc{K. O. Knausgård}, \textit{Om Høsten}}}
-\end{epigraphs}
-
-\section*{What Are the Electrons Really Doing in Molecules?}
+\\
+\hspace*{15pt}
+--- \textnorsk{\textsc{K. O. Knausgård}, \textit{Om Høsten}}
+}
 
@@
 Model chemistries are \emph{systematically improvable} so that it is
-possible to achieve the heaven of chemical accuracy by relaxing
+possible to achieve the heaven of chemical
+accuracy\footnote{A theoretical result is said to be \emph{chemically
+accurate} if it is within \SI{1}{\kilo\calorie\per\mol} of the corresponding
+experimental measurement.} by relaxing
 approximations in the model, albeit at an increased computational cost.
@@
 I will put emphasis on the methods that have been relevant in the work
-presented in this dissertation. Section \ref{sec:mqm} discusses some of the
-ideas of quantum many-body theory relevant to quantum chemistry. Section
-\ref{sec:mean-field} introduces the workhorses of quantum chemistry
-\acrlong{HF} and \acrlong{KS} \acrlong{DFT}.
-The \acrlong{MBPT} and \acrlong{CC} rungs on the methodology axis will
-be discussed in Sections \ref{sec:mbpt} and \ref{sec:coupled-cluster},
-respectively.
+presented in this dissertation.
 
@@
 Environment effects are usually classified as:
-\begin{description}[leftmargin=2mm, font=\normalfont]
+\begin{description}[leftmargin=2mm, font=\itshape]
 \item[Direct.]
@@
 
-\section*{The Road to Reality or Molecular Response Properties}
+\section*{A Road to Reality or Molecular Response Properties}
 
-\begarrowthought{The scope of quantum chemistry is to understand and predict molecular
-properties}, but so far we have made no mention of properties other than
-the energy.
 The experimentalists' view of molecular systems is built mainly around
@@
 \emph{Response theory} is the missing link between theory and
-experiment, making quantum chemistry a full-fledged virtual laboratory.
+these vast classes of
+experiments, making quantum chemistry a full-fledged virtual laboratory.
 Response theory allows the description and computation of
@@
 A balanced description of electron correlation is often necessary
-to achieve accurate enough results in the computation of molecular
-properties.\autocite{Lee1995-pw, Helgaker2004-oz, Tajti2004-ye}
+to achieve results that are accurate enough for meaningful comparisons
+with experiments.\autocite{Lee1995-pw, Helgaker2004-oz, Tajti2004-ye}
 The cost-effective treatment of electron correlation is a challenging
@@
 environment effects in \emph{ab initio} models of interesting chemical
-systems. As already noted, this is a challenging problem, the more so
+systems. As already noted, this is a difficult problem, the more so
 when including electron correlation is necessary to obtain better

### 2016-10-04 Chapters/ContinuumSolvationModels.tex
Adjust equations

@@
 \chapter{Continuum Solvation Models}\label{ch:CSM}
+\footnotetext{\textitalian{L'acqua è la forza che ti tempra, \\
+nell'acqua ti ritrovi e ti rinnovi: \\
+\hspace*{15pt} --- \textsc{Eugenio Montale}, \\ \hspace*{15pt}
+\textit{Falsetto}}}
 
-\epigraph{
-\textitalian{
-L'acqua è la forza che ti tempra,
-
-nell'acqua ti ritrovi e ti rinnovi:
-
-noi ti pensiamo come un'alga, un ciottolo,
-
-come un'equorea creatura
-
-che la salsedine non intacca
-
-ma torna al lito più pura.
-}
-}{
-    --- \textsc{Eugenio Montale}, \textit{Falsetto}}
-
-Developments of the \gls{PCM}, a continuum solvation model, are at the
+Developments of the \glsfirst{PCM}, a continuum solvation model, are at the
 heart of this dissertation.
@@
 we have worked upon in the thesis.
-Appendix \ref{app:mathematical-results} contains much of the
-mathematical details omitted from the exposition in this Chapter.
-Section \ref{sec:IEF} will present the formulation of the model.
-I will show how a variational approach can facilitate the formulation of
-quantum/classical polarizable Hamiltonians in Section
-\ref{sec:variational}.
-Section \ref{sec:BEM} will be devoted to the numerical strategies used
-in solving the \acrshort{PCM} equation.
-
 Continuum models have a venerable history in quantum chemistry:
@@
 1981.\autocite{Miertus1981-mm}
-Section \ref{sec:CSM-why-how} will offer a heuristic "derivation" of
+Section \ref{sec:CSM-why-how} will offer a heuristic
+\enquote{derivation} of
 continuum models.
@@
 physical insights that have informed the creation of continuum models.
+Section \ref{sec:IEF} will present the formulation of the model.
+Section \ref{sec:BEM} will be devoted to the numerical strategies used
+in solving the \acrshort{PCM} equation.
+I will show how a variational approach can facilitate the formulation of
+quantum/classical polarizable Hamiltonians in Section
+\ref{sec:variational}.
+Appendix \ref{app:mathematical-results} contains much of the
+mathematical details omitted from the exposition in this Chapter.
+
 Still nowadays much research activity is expended on the
@@
 
-The basic idea of implicit models is that the environment can
-be replaced by a structureless continuum.
+\begarrowthought{The basic idea of implicit models is to replace the environment
+with a structureless continuum.}
 The continuum will interact \emph{classically} with the explicit part of
@@
 \begin{enumerate}
-\item the statistical averaging of environment configurations and
-\item the \acrshort{MM} region electrostatics cutoff choice.
+\item the statistical averaging of environment configurations, and
+\item the \acrshort{MM} region electrostatics cutoff radius choice.
 \end{enumerate}
@@
 step \emph{before} embarking into the solution of the quantum mechanical
-problem and replace the full Hamiltonian with and \emph{effective}
+problem and replace the full Hamiltonian with an \emph{effective}
 Hamiltonian in which the environment degrees of freedom are properly
@@
 depicted in Figure \ref{fig:IEF}.
-The solute molecule is enclosed in a \emph{cavity} inside the
-structureless continuum representing the solvent and the basic energetic
-quantity is a \emph{free energy}:\autocite{Tomasi1994-gv,
-Amovilli1998-pv, Tomasi2005-vm}
-\begin{equation}\label{eq:free-energy}
- G = G_\mathrm{cav} + G_\mathrm{el} + G_\mathrm{dis} + G_\mathrm{rep} + G_\mathrm{Mm}(T).
-\end{equation}
-The decomposition of the solute free energy takes into account the
-different physical origins of the different solute-solvent interactions.
-$G_\mathrm{cav}$ is the cavity formation free energy, a constant as long
-as the solute geometry does not change.\autocite{Tomasi1994-gv,
-Floris1997-tu}
-The second term $G_\mathrm{el}$ is the most important one in the
-framework of continuum models. It accounts for the mutual electrostatic
-polarization of solute and solvent (Section \ref{sec:IEF})
-The third and fourth terms account for
-dispersion ($G_\mathrm{dis}$) and repulsion ($G_\mathrm{rep}$) interactions,
-respectively.\autocite{Amovilli1997-fo, Weijo2010-jy}
-Finally, $G_\mathrm{Mm}(T)$ accounts for the collective molecular
-motions of the solute-solvent system. It explicitly depends on the
-temperature $T$ and can be calculated from the solution partition
-function, see \noparcite[ref.][]{Amovilli1998-pv} for further details.
-
-In this dissertation, we focused exclusively on the purely electrostatic
-component of the interaction energy, completely neglecting the other
-terms.
-While a balanced description of solvation chemistry certainly demands
-a proper treatment of all solute-solvent interactions, especially when
-non homogeneous environments are studied,\autocite{Mozgawa2014-ad,
-Mozgawa2016-wf} it is also true that the electrostatic component is the
-most numerically relevant, especially when considering energy
-differences between different electronic states.
-%and the most sensitive to changes in the
-%system studied and in the level of quantum mechanical description.
 
@@
 
+The solute molecule is enclosed in a \emph{cavity} inside the
+structureless continuum representing the solvent and the basic energetic
+quantity is a \emph{free energy}:\autocite{Tomasi1994-gv,
+Amovilli1998-pv, Tomasi2005-vm}
+\begin{equation}\label{eq:free-energy}
+ G = G_\mathrm{cav} + G_\mathrm{el} + G_\mathrm{dis} + G_\mathrm{rep} + G_\mathrm{Mm}(T).
+\end{equation}
+The decomposition of the solute free energy takes into account the
+different physical origins of the different solute-solvent interactions.
+$G_\mathrm{cav}$ is the cavity formation free energy, a constant as long
+as the solute geometry does not change.\autocite{Tomasi1994-gv,
+Floris1997-tu}
+The second term $G_\mathrm{el}$ is the most important one in the
+framework of continuum models. It accounts for the mutual electrostatic
+polarization of solute and solvent, see Section \ref{sec:IEF}.
+The third and fourth terms account for
+dispersion ($G_\mathrm{dis}$) and repulsion ($G_\mathrm{rep}$) interactions,
+respectively.\autocite{Amovilli1997-fo, Weijo2010-jy}
+Finally, $G_\mathrm{Mm}(T)$ accounts for the collective molecular
+motions of the solute-solvent system. It explicitly depends on the
+temperature $T$ and can be calculated from the solution partition
+function, see \noparcite[ref.][]{Amovilli1998-pv} for further details.
+
+In this dissertation, we focused exclusively on the purely electrostatic
+component of the interaction energy, completely neglecting the other
+terms.
+While a balanced description of solvation chemistry certainly demands
+a proper treatment of all solute-solvent interactions, especially when
+non homogeneous environments are studied,\autocite{Mozgawa2014-ad,
+Mozgawa2016-wf} it is also true that the electrostatic component is the
+most numerically relevant, especially when considering energy
+differences between different electronic states.
+
 \section{Continuum Electrostatics as a Boundary Integral Problem}\label{sec:IEF}
@@
 The boundary of the cavity $\Gamma\equiv\partial\Omegai$ is a
-2-manifold and is assumed to be continuously differentiable, \ie $\mathcal{C}^1$.
+2-manifold and is assumed to be continuously differentiable.
 The cavity is carved out of an infinite, structureless, continuum
@@
 
-Given the partition of Euclidean space $\mathbb{R}^3$ outlined above,
+Given the partition of Euclidean space $\mathbb{R}^3$ outlined above, we
+want to
 find the solution $u(\vect{r})$ to the following set:
@@
 conormal derivatives can be found in the excellent book by
-\citeauthor{Sauter2011-an}.\autocite{Sauter2011-an}.
+\citeauthor{Sauter2011-an}\autocite{Sauter2011-an}
 The fundamental solutions, or Green's functions, for the elliptic
@@
 
-The first step is to introduce the relevant boundary integral operators for the
+\arrowthought{The first step is to introduce the relevant boundary
+integral operators} for the
 above mentioned transmission problem.
@@
 \begin{align}
-  (\bi{S}_\star v)(\vect{s}) &= \int_\Gamma
-  G_\star(\vect{s}, \vect{s}^\prime)v(\vect{s}^\prime)\diff
-  \vect{s}^\prime \label{eq:S} \\
-  (\bi{D}_\star v)(\vect{s}) &= \int_\Gamma \partial_{L_\star,\vect{s}}
-  G_\star(\vect{s}, \vect{s}^\prime)v(\vect{s}^\prime)\diff
-  \vect{s}^\prime \label{eq:D} \\
-  (\bi{D}^\dagger_\star v)(\vect{s}) &= \int_\Gamma
+  (\bi{S}_\star v)(\vect{s}) &= \int_\Gamma\diff\vect{s}^\prime
+  G_\star(\vect{s}, \vect{s}^\prime)v(\vect{s}^\prime)
+  \label{eq:S} \\
+  (\bi{D}_\star v)(\vect{s}) &= \int_\Gamma\diff\vect{s}^\prime  \partial_{L_\star,\vect{s}}
+  G_\star(\vect{s}, \vect{s}^\prime)v(\vect{s}^\prime)
+  \label{eq:D} \\
+  (\bi{D}^\dagger_\star v)(\vect{s}) &= \int_\Gamma\diff\vect{s}^\prime
 \partial_{L_\star,\vect{s}^\prime}
-  G_\star(\vect{s}, \vect{s}^\prime)v(\vect{s}^\prime)\diff
-  \vect{s}^\prime \label{eq:D-dagger}
+  G_\star(\vect{s}, \vect{s}^\prime)v(\vect{s}^\prime)
+  \label{eq:D-dagger}
 \end{align}
@@
   u(\vect{r}) = \esp(\vect{r}) + \xi(\vect{r}) = \int_{\Omegai}
+  \diff\vect{r}^\prime
   \frac{\rhoi(\vect{r}^\prime)}{|\vect{r}-\vect{r}^\prime|}
@@
 (Newton potential), while the latter is the \emph{reaction} potential.
-The reaction potential describes the polarization in the medium.
-Moreover, it admits an integral representation as a single-layer
+The reaction potential describes the polarization in the medium
+and it admits an integral representation as a single-layer
 potential:
 \begin{equation}
-  \xi_\mathrm{i} = \bi{S}_\mathrm{i}\sigma
+  \xi_\mathrm{i} = \bi{S}_\mathrm{i}\sigma,
 \end{equation}
@@
 \begin{equation}\label{eq:full-IEF}
-  \left[ \bi{S}_\mathrm{e}\left(2\pi + \bi{D}^\dagger_\mathrm{i}\right)
-  +
-  \left(2\pi - \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}
-  \right]\sigma =
-  -\left[\left(2\pi-\bi{D}_\mathrm{e}\right)
+  \begin{aligned}
+  \big[ \bi{S}_\mathrm{e}\big(2\pi + \bi{D}^\dagger_\mathrm{i}\big)
+  &+
+  \big(2\pi - \bi{D}_\mathrm{e}\big)\bi{S}_\mathrm{i}
+  \big]\sigma = \\
+  &-\big[\big(2\pi-\bi{D}_\mathrm{e}\big)
   -\bi{S}_\mathrm{e}\bi{S}_\mathrm{i}^{-1}
-  \left(2\pi-\bi{D}_\mathrm{i}\right)
-  \right]\esp
+  \big(2\pi-\bi{D}_\mathrm{i}\big)
+  \big]\esp
+  \end{aligned}
 \end{equation}
@@
 have been introduced.
-By letting the $\diel\rightarrow\infty$, one recovers the limiting case
+By letting $\diel\rightarrow\infty$, one recovers the limiting case
 where conductor boundary conditions are imposed. This leads to the
@@
 
-The solution of the \acrshort{IEF}-\acrshort{PCM} boundary integral equation can
+\begarrowthought{The solution of the \acrshort{IEF}-\acrshort{PCM}
+boundary integral equation} can
 only be achieved analytically for simple cavity geometries, such as
@@
 \begin{enumerate}
-    \item define the molecular cavity
-    \item mesh the molecular surface \label{item:mesh}
-    \item discretize the integral equation \label{item:discretize} and
+    \item define the molecular cavity,
+    \item mesh the molecular surface, \label{item:mesh}
+    \item discretize the integral equation, \label{item:discretize} and
     \item solve the linear systems by a suitable numerical technique.
@@
 van der Waals radii.\autocite{Bondi1964-dt, Mantina2009-hb}
-If one imagines to model a solvent molecule as a rigid sphere,
-three types of molecular surfaces can be defined:
-\begin{enumerate}
-  \item
-    the union of the interlocking spheres surfaces is the \gls{vdWS},
-    Figure \ref{fig:vdWS}
-  \item the locus of points defined by the center of the solvent
-    spherical probe while rolling over the \acrshort{vdWS} is the
-    \gls{SAS}, Figure \ref{fig:SAS}
-  \item the locus of points defined by the contact point of the solvent
-    spherical probe while rolling over the \acrshort{vdWS} is the \gls{SES}, Figure \ref{fig:SES}.
-\end{enumerate}
+
 \begin{figure}[tb]
@@
 \end{figure}
+
+If one imagines to model a solvent molecule as a rigid sphere,
+three types of molecular surfaces can be defined:
+\begin{enumerate}
+  \item
+    the union of the interlocking spheres surfaces is the \gls{vdWS},
+    Figure \ref{fig:vdWS}
+  \item the locus of points defined by the center of the solvent
+    spherical probe while rolling over the \acrshort{vdWS} is the
+    \gls{SAS}, Figure \ref{fig:SAS}
+  \item the locus of points defined by the contact point of the solvent
+    spherical probe while rolling over the \acrshort{vdWS} is the \gls{SES}, Figure \ref{fig:SES}.
+\end{enumerate}
 Alternative definitions based on the isodensity surface of the molecular solute
@@
 piecewise regular on any given interval in the mesh.\autocite{Ern2004-oo}
-Basically, given an element in the mesh, we "attach" a basis set of
+Basically, given an element in the mesh, we \enquote{attach} a basis set of
 piecewise polynomials of a prescribed degree and use this basis to
@@
 The finite element discretization is usually carried out either by the
-\emph{collocation} or by the \emph{Galerkin} approach.
-Indeed, it is possible to show that discretization by collocation
-corresponds to a Galerkin method with a very specific (and asymmetric)
-choice of \emph{trial} and \emph{test}
-spaces.\autocite{Hackbusch1995-uq, Ern2004-oo}
+\emph{collocation} or by the \emph{Galerkin}
+approach.\autocite{Hackbusch1995-uq, Ern2004-oo}\footnote{ Indeed, it is
+possible to show that discretization by collocation corresponds to a
+Galerkin method with a very specific (and asymmetric) choice of
+\emph{trial} and \emph{test} spaces.}
 This step transforms the original integral equation into a system of
@@
 
-In the GePol procedure, the molecular surface is meshed by means of
+\arrowthought{In the GePol procedure,} the molecular surface is meshed by means of
 $N_\mathrm{ts}$ spherical polygons $\Pi_i$ whose centroids define the
@@
 areas and their curvatures $R_i$.
-The factor $k=1.07$ is introduced to achieve a better precision on the
-diagonal elements.\autocite{Tomasi2005-vm}
+The factor $k=1.07$ is introduced to achieve a better precision in the
+integration of the diagonal elements.\autocite{Tomasi2005-vm}
 Expressions for $\mat{S}_\mathrm{e}$, $\mat{D}_\mathrm{e}$ and
@@
 
-The wavelet Galerkin \acrshort{BEM} was also used in this dissertation. This method
+\arrowthought{The wavelet Galerkin \acrshort{BEM}}
 preserves the fundamental symmetries of the boundary integral operators
 and achieves sparsity in the stiffness matrix.
-This sparsity can be leveraged to implement efficient algorithms with
+This sparsity was leveraged in this dissertation to implement efficient algorithms with
 linear space and time complexity. The approach is thus superior
@@
 stiffness matrix, Figure \ref{fig:finger-structure}.
-The number of relevant matrix entries scales proportional to $O(N_J)$
+The number of relevant matrix entries scales linearly
 with $N_J$ the number of degrees of freedom for a refinement of the
@@
 
-Continuity of meshes with respect to molecular distortions is a
+\newthought{Continuity of meshes with respect to molecular distortions} is a
 central issue for stable implementations of molecular gradients and
@@
 modification\autocite{Su2009-em} to the GePol algorithm put forth by
-\citeauthor{Su2009-em}.
+\citeauthor{Su2009-em}
 \citeauthor{Scalmani2010-tw} have recently proposed the \gls{CSC} scheme
@@
 molecular gradients. However, their convergence to the exact solution of
-the underlying integral equation has yet to be proved.
-A completely different approach to the problem has been taken by
-\citeauthor{Cances2013-jg} by employing a domain decomposition
+the underlying integral equation has yet to be
+proved.\autocite{Ern2004-oo}
+\citeauthor{Cances2013-jg} took
+a completely different approach to the problem
+by employing a domain decomposition
 method, guaranteed to converge to the exact solution.\autocite{Quarteroni1999-jt}
@@
 In the following, the \acrshort{PCM} equations will be written in the
-"complete basis" meaning that we will be working with the exact integral
+\enquote{complete basis} meaning that we will be working with the exact integral
 equation and not with its discretized counterpart.
@@
 the electrostatic potential $\esp$ will have a \emph{continuous}
-dependence on a cavity surface "index" $\vect{s}$.
+dependence on a cavity surface \enquote{index} $\vect{s}$.
 A product involving any of these quantities will have to be
@@
 The strategy usually adopted is to tailor a specific Hamiltonian, which
-includes classical contributions form the continuum and use it in the
+includes classical contributions from the continuum and use it in the
 development of the quantum chemical model.
@@
 theoretically cumbersome,\autocite{Sanhueza1979-bp, Heimsoeth1990-ki}
-but has been successfully exploited in the past.\autocite{Tomasi2005-vm}
+but has been successfully done in the past.\autocite{Tomasi2005-vm}
 
-In this dissertation we advocate a different theoretical approach: the
+\arrowthought{In this dissertation, we advocate a different theoretical
+approach:} the
 variational formulation of classical polarizable models.
@@
     \item physical extremal points,
-    \item equilibrium values are true energies \label{item:equil-prop} and
+    \item equilibrium values that are true energies, \label{item:equil-prop} and
     \item convexity (positive-definiteness)
@@
 the \gls{PE},\autocite{Olsen2010-wa, Olsen2011-io} or
-the \gls{FQ} model.\autocite{Rick1994-mn, Rick1995-wu, Rick1996-om,
+the \gls{FQ} models.\autocite{Rick1994-mn, Rick1995-wu, Rick1996-om,
 Lipparini2011-rd}
@@
 \begin{equation}
-  \sigma\MMPCM\kappa = \kappa\MMPCM^\dagger\sigma
+  \sigma\MMPCM\kappa = \kappa\MMPCM^\dagger\sigma.
 \end{equation}
@@
    \zeta
-  \end{pmatrix}
+  \end{pmatrix}.
 \end{equation}
@@
 
-How are quantum/classical polarizable Hamiltonians implemented in this
-framework?
+\arrowthought{How are quantum/classical polarizable Hamiltonians implemented in this
+framework?}
 As shown in Chapter \ref{ch:QM} for the \acrshort{CC} method, one can
@@
     \scalprod[W]{\bar{\vect{\eta}}}{\pderiv{\mat{\Omega(\vect{\eta})}}{\vect{\eta}}}
-    + {}^t\p\pderiv{\s(\vect{\eta}, \bar{\vect{\eta}})}{\vect{\eta}}
+    \nonumber \\
+    &+ {}^t\p\pderiv{\s(\vect{\eta}, \bar{\vect{\eta}})}{\vect{\eta}}
     = 0 \\
@@
   \int\diff\vect{r}
-  v_\mathrm{xc}(\vect{r})\Tr\mat{\Omega}(\vect{r})\mat{D} \label{eq:ao-xc} \\
+  v_\mathrm{xc}(\vect{r})\left(\Tr\mat{\Omega}(\vect{r})\mat{D}\right) \label{eq:ao-xc} \\
   \Omega_{\mu\nu}(\vect{r}) &= \chi_\mu^\dagger(\vect{r})\chi_\nu(\vect{r}).
@@
   \mathcal{L}(\mat{C}, \mat{\epsilon}) +
-  \frac{1}{2}{}^t\p\V\p + {}^t\p\Tr\mathbfup{s}\mat{D}.
+  \frac{1}{2}{}^t\p\V\p + {}^t\p(\Tr\mathbfup{s}\mat{D}).
 \end{equation}

### 2016-10-04 Chapters/QuantumMechanics.tex
Adjust equations

@@
 \chapter{A Tour of Molecular Electronic Structure Theory}\label{ch:QM}
-
-\begin{epigraphs}
-\setlength\epigraphwidth{7.5cm}
-\qitem{
-plays with the bigness of his littleness
-
---- electrons deify one razorblade
-
-into a mountainrange; lenses extend
-
-unwish through curving wherewhen till unwish
-
-returns on its unself.
-    }{
-    --- \textsc{E. E. Cummings}}
-\setlength\epigraphwidth{7cm}
-\end{epigraphs}
+\footnotetext{--- electrons deify one razorblade \\
+into a mountainrange; [...] \\
+\hspace*{15pt} --- \textsc{E. E. Cummings}}
 
@@
 
-Quantum mechanics is the theory describing the motion and interactions
-of microscopic particles. In quantum theory every observable of the
+\begarrowthought{Quantum mechanics is the theory describing the motion and interactions
+of microscopic particles.} In quantum theory every observable of the
 system is represented mathematically by an operator $O$ in the
@@
 
-Molecular quantum mechanics is concerned with the microscopic motion of
-nuclei and electrons in molecules and is thus an intrinsically many-body
+\begarrowthought{Molecular quantum mechanics is concerned with the microscopic motion of
+nuclei and electrons in molecules} and is thus an intrinsically many-body
 theory. The problem is clearly very complicated to solve, as the number of
@@
 $N$-electron, molecular electronic Hamiltonian is:
-\begin{alignat}{2}\label{eq:fq-hamiltonian}
+\begin{equation}\label{eq:fq-hamiltonian}
   H = \sum_{i=1}^N h(\vect{r}_i)
   + \frac{1}{2}\sum_{i\neq j}^N g(\vect{r}_i, \vect{r}_j)
-  + V_\mathrm{NN},
-  \quad&
+  + V_\mathrm{NN}.
+\end{equation}
+The last term is the nuclear repulsion energy:
+\begin{equation}
   V_\mathrm{NN} =
-  \frac{1}{2}\sum_{A \neq B}^{N_\mathrm{nuclei}}\frac{Z_A Z_B}{|\vect{R}_A - \vect{R}_B|}
-\end{alignat}
-The last term is the nuclear repulsion energy, a constant shift of the
+  \frac{1}{2}\sum_{A \neq B}^{N_\mathrm{nuclei}}\frac{Z_A
+  Z_B}{|\vect{R}_A - \vect{R}_B|},
+\end{equation}
+a constant shift of the
 energy scale in the \acrshort{BO} approximation.
@@
 
-It the language of second quantization, we assume that a, not
-necessarily complete, set of orthonormal one-particle states is
-available. These are usually called \glspl{MO}.
+\begarrowthought{In the language of second quantization}, we assume that a
+complete and orthonormal
+set of one-particle states is
+available.\footnote{Neither completeness
+nor orthonormality of the one-particle set are necessary assumptions.
+Relaxing both assumptions would however make the formalism slightly more
+involved.}
+These are usually called \glspl{MO}.
 In general, we would like to embed as many of the symmetries of the
@@
 into one-particle states and accordingly labelling them by the number of
-particles per one-particle states present.
+particles per one-particle state present.
 This is the so-called \gls{ON} vector representation:
 \begin{equation}
-  \ket{\vect{n}} = \ket{n1,n2,\ldots},
+  \ket{\vect{n}} = \ket{n_1,n_2,\ldots},
 \end{equation}
@@
 \emph{creation} operators $\cons{i}$. Creation operators increase the
-occupation number in the one-particle state $i$ by $1$, up to a phase.
-Conversely, \emph{annihilation} operators $\anni{i}$, defined as the Hermitian
+occupation number in the one-particle state $i$ by $1$, up to a
+phase. Conversely, \emph{annihilation} operators $\anni{i}$, defined as the Hermitian
 conjugates of the creation operators, decrease occupation in state $i$
 by $1$, again up to a phase. Clearly, annihilation of any one-particle
-state in the vacuum results yields $0$.
+state in the vacuum yields $0$.
 Pauli's exclusion principle is enforced by the \emph{canonical
@@
 showing how the construction of the Fock space representation
-embeds antisymmetry at the operator level and not in the wave functions.
+embeds antisymmetry at the operator level and not in the \gls{ON} vectors.
 Finally, the second-quantized molecular electronic Hamiltonian in the \acrlong{BO}
@@
 \end{equation}
-also known at the \gls{FCI} expansion.
+also known as the \gls{FCI} expansion.
 Among the many properties the exact wave function enjoys, the
@@
 vectors in Fock space can be classified according to how different their
-occupation is from the reference, \ie~\emph{by excitation
+occupation is from the reference, \ie by \emph{excitation
 level}.\autocite{Helgaker2000-tz, Shavitt2009-mr}
@@
 number of excitations.
-Further details on how the expansions are actually tailored and
+Further details on how these expansions are actually tailored and
 algorithmically implemented can be found elsewhere.\autocite{Helgaker2000-tz}
@@
 
-The idea behind mean-field theory is simple: seek the best description
+  \begarrowthought{The idea behind mean-field theory is simple:} seek the best description
 of the many-electron wave function using just one Slater determinant.
-How does one direct this search towards the "best"? We will use the
+How does one direct this search towards the \enquote{best}? We will use the
 variational principle and optimize the degrees of freedom in our
@@
 $N$-electron space. All variational degrees of freedom reside in
-the one-electron space, \ie~in the space of \glspl{MO}
+the one-electron space, \ie in the space of \glspl{MO}
 $\lbrace\phi_r\rbrace$, used to build the determinant $\Ket{0}$.
-The one-electron \glspl{MO} are thus optimized considering each single
-electron in the mean field generated by the motion of all other
-electrons, hence the name of the method. Clearly this neglects
-parts of the electron--electron interaction energy, the \emph{electron
-correlation}. However, in most cases, it provides a suitable starting
+The one-electron \glspl{MO}
+are thus optimized considering each single electron in the mean field generated
+by the motion of all other electrons, hence the name of the
+method. Clearly this neglects parts of the electron-- electron
+interaction energy: \emph{electron correlation}.
+According to \citeauthor{Lowdin1955-yt}, this can be defined as the
+energy difference between the exact nonrelativistic and the
+complete-basis set mean-field results for the system under
+consideration.\autocite{Lowdin1955-yt}
+However, in most cases, it provides a suitable starting
 approximation for more sophisticated many-body treatments that recover
@@
 $\kappa_{rs}, \kappa^*_{rs}$ and setting its gradient $\mat{E}^{[1]}$ to
-zero will locate a stationary point. The gradient of the electronic
+zero will identify its stationary points. The gradient of the electronic
 energy can be identified by comparing the Taylor and \gls{BCH}
@@
 
-A different approach is taken in \gls{DFT}, where the electron density
+\newthought{A different approach is taken in \gls{DFT}}, where the electron density
 is the central quantity in the theory, instead of the wave function.
@@
 \begin{equation}
- E^\mathrm{LDA}_\mathrm{xc}[\rho] = \int e_\mathrm{xc}(\rho(\vect{r}))\diff\vect{r}.
+ E^\mathrm{LDA}_\mathrm{xc}[\rho] = \int \diff\vect{r}e_\mathrm{xc}(\rho(\vect{r})).
 \end{equation}
@@
 \begin{alignat}{2}
- E^\mathrm{GGA}_\mathrm{xc}[\rho] = \int
- e_\mathrm{xc}(\rho(\vect{r}),\zeta(\vect{r}))\diff\vect{r},
+ E^\mathrm{GGA}_\mathrm{xc}[\rho] = \int\diff\vect{r}
+ e_\mathrm{xc}(\rho(\vect{r}),\zeta(\vect{r})),
  \quad&
@@
 
-How are \glspl*{MO} constructed? The usual approach is to expand them in
+\newthought{How are \glspl{MO} constructed?} The usual approach is to expand them in
 a basis of one-particle, atom-centered basis functions of known
@@
 \end{equation}
-that is, as a truncated sum of excitation operators, $\cluster{\mu_p}$,
-times the corresponding cluster amplitudes, $\tamp{\mu_p}$.
-Here $\mu_p$ is the $\mu$-th excitation at the $p$-th excitation level and
+that is, as a truncated sum of excitation operators, $\cluster{\mu_u}$,
+times the corresponding cluster amplitudes, $\tamp{\mu_u}$.
+Here $\mu_u$ is the $\mu$-th excitation at the $u$-th excitation level and
 $\mathcal{M}$ is the truncation level.
@@
   =
+  O + \tilde{O}
+  =
   O + \BCHfirst{O}{T} + \BCHsecond{O}{T} + \ldots
-  =
-  O + \tilde{O}.
 \end{equation}
@@
 further insight:
-\begin{alignat}{2}
+\begin{equation}
  H = F + \Phi = \sum_r\epsilon_r\cons{r}\anni{r} + (g- V_\mathrm{HF}),
- \quad&
- V_\mathrm{HF} = \sum_{i}[g_{rsii} - g_{riis}]\cons{r}\anni{s}
-\end{alignat}
-where $F$ if the Fock matrix and $\Phi$ is the fluctuation potential.
+\end{equation}
+where $F$ if the Fock matrix and $\Phi$ is the fluctuation potential,
+expressed as the difference between the full two-electron interaction
+and its \gls{HF} mean-field approximation:
+\begin{equation}
+ V_\mathrm{HF} = \sum_{i}[g_{rsii} - g_{riis}]\cons{r}\anni{s}.
+\end{equation}
 Given our initial assumption on the reference function, the Fock
@@
 
-As formulated so far, the \acrshort{CC} method is not variational.
+\newthought{We have already remarked that the \acrshort{CC} method is
+not variational.}
 It is however possible to introduce a fully variational Lagrangian that
@@
   \mathcal{E}(\vect{\eta}) +
-  \scalprod[W]{\bar{\vect{\eta}}}{\mat{\Omega}(\vect{\eta})}
+  \scalprod[W]{\bar{\vect{\eta}}}{\mat{\Omega}(\vect{\eta})},
 \end{equation}
-where the governing equations are obtained by differentiation with
+where $\scalprod[W]{\cdot}{\cdot}$ denotes the scalar product in the
+vector space $W$.
+The governing equations are then obtained by differentiation with
 respect to both sets of parameters:
@@
   + \sum_{u=1}^{\mathcal{M}}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
-  + \braket{\mathrm{HF} | \overline{\Phi} | \mathrm{HF}}
-  + \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \overline{\Phi} | \mathrm{HF}}
+  + \braket{\mathrm{HF} | \overline{\Phi} | \mathrm{HF}} \\
+  &+ \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \overline{\Phi} | \mathrm{HF}}
   \end{aligned}
@@
     \denom{\mu_q}\tbar{\mu_q} +
-    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}} +
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_q}}
+    | \mathrm{HF}}  \nonumber \\
+    &+
     \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
@@
 
-An alternative method to recover the correlation energy missing in the
-mean-field approximation is to employ perturbation
-theory.\autocite{Konishi2009-zb}
+\begarrowthought{Perturbation theory offers an alternative method to
+recover the correlation energy} missing in the
+mean-field approximation.\autocite{Konishi2009-zb}
 The single determinant optimized in the mean-field approximation is used
@@
 from the corresponding order expansion of the \acrshort{CC} Lagrangian
-and retaining only terms complying to the $2n+1$ and $2n+2$ rules.
-\autocite{Helgaker1988-to, Helgaker1989-wl, Helgaker1992-ph}
+and retaining only terms complying to the $2n+1$ and $2n+2$ rules,
+\emph{vide infra}.\autocite{Helgaker1988-to, Helgaker1989-wl, Helgaker1992-ph}
 A superscript index in square brackets - $[i]$ -  will denote an $i$-th
@@
 \begin{equation}\label{eq:2nd-order}
+  \begin{aligned}
   \denom{\mu_q}\tamp{\mu_q}^{[2]}
-  = - \braket{\mu_q | \commutator{\Phi}{T_2^{[1]}} | \mathrm{HF}}
-  = - \braket{\tbar{2}^{[1]} | \commutator{\Phi}{\cluster{\mu_q}} |
+  &= - \braket{\mu_q | \commutator{\Phi}{T_2^{[1]}} | \mathrm{HF}} \\
+  &= - \braket{\tbar{2}^{[1]} | \commutator{\Phi}{\cluster{\mu_q}} |
   \mathrm{HF}}
   = \denom{\mu_q}\tbar{\mu_q}^{[2]}
+  \end{aligned}
 \end{equation}
@@
     \braket{\tbar{u}^{[2]} | \BCHfirst{\Phi}{T^{[2]}}|
-    \mathrm{HF} }
-    +\sum_{u=1}^\mathcal{M}
+    \mathrm{HF} } \nonumber \\
+    &+\sum_{u=1}^\mathcal{M}
     \braket{\tbar{u}^{[2]} | \BCHsecond{\Phi}{T_2^{[1]}}|
@@
 
-In the iterative schemes, one approximates the amplitude equations for
+\arrowthought{In the iterative schemes,} one approximates the amplitude equations for
 the higher excitation to be correct up to a given order in perturbation
@@
   E_0
-  + \sum_{u=1}^{2}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
-  + \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{Q_2} | \mathrm{HF}} \\
+  + \sum_{u=1}^{2}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u} \\
+  &+ \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{Q_2} | \mathrm{HF}} \\
   &+ \braket{\tbar{1} |
@@
   \braket{\mathrm{HF} | \check{O} | \mathrm{HF} }
-  + \braket{\tbar{1} | \check{O} | \mathrm{HF} }
-  + \braket{\tbar{1} | \commutator{\check{O}}{Q_2} | \mathrm{HF} }
+  + \braket{\tbar{1} | \check{O} | \mathrm{HF} } \\
+  &+ \braket{\tbar{1} | \commutator{\check{O}}{Q_2} | \mathrm{HF} }
   + \braket{\tbar{2} | \commutator{\check{O}}{Q_2} | \mathrm{HF} } \\
@@
   O(\tamp{}, \tbar{})_\mathrm{CCS}
-  + \braket{\tbar{1} | \commutator{\check{O}}{Q_2} | \mathrm{HF} }
-  + \braket{\tbar{2} | \commutator{\check{O}}{Q_2} | \mathrm{HF} }
+  + \braket{\tbar{1} | \commutator{\check{O}}{Q_2} | \mathrm{HF} } \\
+  & + \braket {
+    \tbar{2} | \commutator{\check{O}} {Q_2} | \mathrm { HF }
+  }
   \end{aligned}
@@
     \denom{\mu_2}\tamp{\mu_2} &+ \braket{\mu_2 |
-    \check{\Phi} + \commutator{\check{\Phi}}{T_2}
-    + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2}
+    \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}}
+    \nonumber \\
+    &+ \braket{\mu_2 | \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2}
     + \commutator{\check{\Phi}}{Q_3}
@@
   &+ \braket{\tbar{2} |
-  \check{\Phi} + \commutator{\check{\Phi}}{T_2}
-  + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{Q_3}
+  \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
+  &+ \braket{\tbar{2} |
+  \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{Q_3}
   | \mathrm{HF}} \\
@@
 \end{equation}
-The \acrshort{CC3} methods brings the scaling of the \acrshort{CCSDT}
+The \acrshort{CC3} method brings the scaling of the \acrshort{CCSDT}
 method down to $N^7$ in the iterative step.
@@
   O(\tamp{}, \tbar{})_\mathrm{CCS}
-  + \braket{\tbar{1} | \commutator{\check{O}}{T_2} | \mathrm{HF} }
-  + \braket{\tbar{2} | \commutator{\check{O}}{T_2} | \mathrm{HF} } \\
+  + \braket{\tbar{1} | \commutator{\check{O}}{T_2} | \mathrm{HF} } \\
+  &+ \braket{\tbar{2} | \commutator{\check{O}}{T_2} | \mathrm{HF} } \\
   &+
-  \braket{\tbar{2} | \commutator{\check{O}}{Q_3} | \mathrm{HF} }
-  + \braket{\tbar{3} |
+  \braket{\tbar{2} | \commutator{\check{O}}{Q_3} | \mathrm{HF} } \\
+  &+ \braket{\tbar{3} |
      \commutator{\check{O}}{Q_3}
@@
   +
-  \braket{\tbar{2} | \commutator{\check{O}}{Q_3} | \mathrm{HF} }
-  + \braket{\tbar{3} |
+  \braket{\tbar{2} | \commutator{\check{O}}{Q_3} | \mathrm{HF} } \\
+  &+ \braket{\tbar{3} |
      \commutator{\check{O}}{Q_3}
@@
 
-Finally, one can construct \emph{noniterative} methods, where the
+\arrowthought{Finally, one can construct \emph{noniterative} methods,} where the
 expensive $N^7$ step for the inclusion of connected triples is
@@
   &=
-    \braket{\tbar{1}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
-    +
+    \braket{\tbar{1}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}} \\
+    &+
     \braket{\tbar{2}^{*}| \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
@@
   &=
-    \braket{\tamp{1}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}
-    +
+    \braket{\tamp{1}^{*} | \commutator{\Phi}{T_3^{*}} | \mathrm{HF}} \\
+    &+
     \braket{\tamp{2}^{*}| \commutator{\Phi}{T_3^{*}} | \mathrm{HF}}

### 2016-10-04 Chapters/ResponseTheory.tex
Better quality figure!
Adjust equations

@@
 \chapter{Response Theory and Molecular Properties}\label{ch:molprop}
-
-\begin{epigraphs}
-\qitem{
-White light goin' messin' up my brain
-
-White light it's gonna drive me insane
-}{
---- \textsc{The Velvet Underground}
+\footnotetext{White light goin' messin' up my brain \\
+White light it's gonna drive me insane \\
+\hspace*{15pt} --- \textsc{The Velvet Underground}
 }
-\qitem{What's the frequency, Kenneth? [...]
-
-          I never understood the frequency, uh-huh}{
-          --- \textsc{R.E.M.}}
-\end{epigraphs}
 
 Response theory provides an \emph{ab initio} framework for the
-formulation and computation of molecular properties and can be
-considered the connection between quantum chemistry and experimental
-chemistry.
+formulation and computation of molecular properties and thus a
+connection
+between \emph{ab initio} and experimental methods based on spectroscopy.
 The response treatment of molecular properties has its roots in
@@
 Christiansen1998-pe, Norman2011-ad, Helgaker2012-cz, Pawlowski2015-sq}
+\footnotetext{What's the frequency, Kenneth? \\
+I never understood the frequency \\
+\hspace*{15pt} --- \textsc{R.E.M.}}
 
@@
 response theory for isolated molecules.
-
-\noindent The concepts of quasienergy,\autocite{Christiansen1998-pe} variational
+The concepts of quasienergy,\autocite{Christiansen1998-pe} variational
 perturbation theory\autocite{Helgaker1992-ph} and pole-and-residue
 analysis of response functions\autocite{Olsen1985-nr} will be
-introduced. The exposition closely follows the one by
+introduced. The exposition closely follows that given by
 \citeauthor{Saue2002-ns} in \noparcite[ref.][]{Saue2002-ns}.
@@
 \citeauthor{Thorvaldsen2008-sg} in \noparcite[ref.][]{Thorvaldsen2008-sg}.
-This formalism is suitable for the derivation of arbitrary-order
-response functions and their residues, also when perturbation-dependent
-basis sets are considered.\autocite{Friese2015-kb, Friese2015-bu}
-\paper{V} presents the arbitrary-order derivation of the \acrshort{PCM}.
 
@@
 
-Let us consider the experimentally relevant situation where a molecular
+\begarrowthought{Let us consider the experimentally relevant
+situation} where a molecular
 system is subject to an external source of electromagnetic radiation,
@@
 
-\begin{figure}[tb]
+\begin{figure}[!htb]
   \centering
-  \includegraphics[width=.5\textwidth]{Optical-rotation.png}
+  \includegraphics[width=.7\textwidth]{Optical-rotation.pdf}
   \caption{
@@
  &=
- \sum_{x}\expo{-\mathrm{i}\omega_x t}\epsilon_x V_X
+ \sum_{x}\expo{-\mathrm{i}\omega_x t}\epsilon_x V_X,
   \end{aligned}
 \end{equation}
-where all the frequencies are integer multiples, $\omega_k=n_k \omega$
-of the fundamental frequency $\omega = \frac{2\pi}{T}$.\autocite{Olsen1985-nr}
+where all the frequencies are integer multiples, $\omega_k=n_k \omega$,
+of a fundamental frequency $\omega = \frac{2\pi}{T}$.\autocite{Olsen1985-nr}
 Frequencies and amplitudes in the Fourier decomposition obey the
@@
 
-Under the assumption that the external field is weak with respect to the
+\arrowthought{Under the assumption that the external field is weak} with respect to the
 molecular field, time-dependent perturbation theory is an appropriate
-tool to approximately solve Eq. \eqref{eq:td-schrodinger}.
+tool to approximately solve Eq.
+\eqref{eq:td-schrodinger}.\footnote{\emph{Real-time} methods, where the
+time-dependent Schr\"{o}dinger equation is integrated directly in a
+set time interval, offer an alternative approach to perturbation
+theory.\autocite{Castro2004-tn, Tannor2007-xq, Konishi2009-zb}}
 The response theory route to molecular properties introduces a
@@
   \aveQ =
- \frac{1}{T} \int_{-\frac{T}{2}}^{\frac{T}{2}} \diff t Q(t) =
+  \frac{1}{T} \int_{0}^{T} \diff t Q(t) =
  Q_0 + \sum_{x}\epsilon_x E_x
@@
   Q_0 &= \left\lbrace \Braket{\bar{\psi}(t) | H_0 | \bar{\psi}(t)} \right\rbrace_T
-- \left\lbrace \Braket{\bar{\psi}(t) | \mathrm{i} \pderiv{}{t} | \bar{\psi}(t)}\right\rbrace_T = E_0(0) - S \\
+- \left\lbrace \Braket{\bar{\psi}(t) | \mathrm{i} \pderiv{}{t} |
+\bar{\psi}(t)}\right\rbrace_T \nonumber \\
+&= E_0(0) - S \\
   E_x &= \left\lbrace \Braket{\bar{\psi}(t) | V_X | \bar{\psi}(t)}\exp(-\mathrm{i}\omega_k t) \right\rbrace_T
@@
 \begin{aligned}
- \Braket{\bar{\psi}(t) | V_X | \bar{\psi}(t)} &\simeq
- \Braket{0|V_X|0} +\sum_{k=-N}^N
+  \langle \bar{\psi}(t) | V_X &| \bar{\psi}(t) \rangle
+  \simeq
+ \Braket{0|V_X|0} \\
+ &+\sum_{k=-N}^N
  \response{V_X}{V(\omega_k)}{\omega_k}\expo{-\mathrm{i}\omega_k t} \\
@@
   &+\sum_{y}\response{V_X}{V_Y}{\omega_y}\epsilon_y\delta_{\omega_x+\omega_y} \\
-  &+\frac{1}{2}\sum_{y,z}\response{V_X}{V_Y,V_Z}{\omega_y,\omega_z}\epsilon_y\epsilon_z\delta_{\omega_x+\omega_y+\omega_z}
+  &+\frac{1}{2}\sum_{y,z}\response{V_X}{V_Y,V_Z}{\omega_y,\omega_z}\epsilon_y\epsilon_z\delta_{\omega_x+\omega_y+\omega_z}.
 \end{aligned}
@@
 \begin{equation}\label{eq:nth-order-derivative}
+  \begin{aligned}
   \left.\frac{\diff^n \aveQ}{\diff\epsilon_{x_1}\diff\epsilon_{x_2}\cdots\diff\epsilon_{x_n}} \right|_{\vect{\epsilon}=0}
-  =
-  \response{V_{X_1}}{V_{X_2}, \ldots, V_{X_n}}{\omega_{x_2},\ldots,\omega_{x_n}}\delta_{\omega_{x_1}+\cdots+\omega_{x_n}},
+  &= \\
+  \langle\langle
+  V_{X_1}&; V_{X_2}, \ldots, V_{X_n}
+  \rangle\rangle_{\omega_{x_2},\ldots,\omega_{x_n}}
+  \delta_{\omega_{x_1}+\cdots+\omega_{x_n}},
+  \end{aligned}
 \end{equation}
@@
 \begin{equation}\label{eq:sum-rule}
- \sum_{i=1}^{n} \omega_{k_i} = 0
+ \sum_{i=1}^{n} \omega_{x_i} = 0
 \end{equation}
@@
 $V_X$, whose response is measured, from the external perturbations
-$V_Y$, $V_Z$\dots{} causing it.
+$V_Y$, $V_Z$, \dots{} causing it.
 From the perturbation expansion Eq. \eqref{eq:kubo}, it is clear that
@@
 
-Having introduced response functions as a useful tool to quantify the
+\newthought{Having introduced response functions as a useful tool} to quantify the
 effect of external perturbations on molecular properties, it is now time
@@
 be variationally parametrized by a set of parameters
-$\vect{\lambda}$: $\ket{0}=\ket{0(\vect{\lambda})}$.
+$\vect{\eta}$: $\ket{0}=\ket{0(\vect{\eta})}$.
 We recall that the Lagrangian method outlined in Section
@@
 \begin{equation}\label{eq:var-cond}
- \deriv{\aveQ}{\vect{\lambda}} = 0.
+ \deriv{\aveQ}{\vect{\eta}} = 0.
 \end{equation}
@@
 Saue2002-ns} we assume the variational parameters to be functions of the
-perturbation strength parameters: $\vect{\lambda} =
-\vect{\lambda}(\vect{\epsilon})$.
+perturbation strength parameters: $\vect{\eta} =
+\vect{\eta}(\vect{\epsilon})$.
 Furthermore, we require the stationarity condition to hold at \emph{all}
-perturbation strengths, with $\lambda(0) = 0$, for convenience.
+perturbation strengths, with $\eta(0) = 0$, for convenience.
 
@@
 \begin{aligned}
- \left[\prod_{i=1}^n \deriv{}{\epsilon_{x_i}}\right]\aveQ
- &= Q^{[n]}_{x_1 \ldots x_n}
- =
+ \left[\prod_{i=1}^n \deriv{}{\epsilon_{x_i}}\right]&\aveQ
+ = Q^{[n]}_{x_1 \ldots x_n} \\
+ &=
  Q^{[n]}_{0; x_1 \ldots x_n}
@@
    \sum_{\sigma}\left[
-   Q^{[1]}_{0; \sigma}\lambda^{[1]}_{\sigma;a} + E_{a}
+   Q^{[1]}_{0; \sigma}\eta^{[1]}_{\sigma;a} + E_{a}
    \right]\delta_{\omega_a} \\
    Q^{[2]}_{ab} &=
-   \sum_{\sigma\tau}\left[
-   \lambda^{[1]}_{\sigma;a}Q^{[2]}_{0;\sigma\tau}\lambda^{[1]}_{\tau;b}
-   + Q^{[1]}_{0; \sigma}\lambda^{[2]}_{\sigma;a}
-   + E^{[1]}_{a;\sigma}\lambda^{[1]}_{\sigma;b}
-   + E^{[1]}_{b;\sigma}\lambda^{[1]}_{\sigma;a}
-   \right]\delta_{\omega_a+\omega_b} \label{eq:linres}
+   \sum_{\sigma\tau}\big[
+   \eta^{[1]}_{\sigma;a}Q^{[2]}_{0;\sigma\tau}\eta^{[1]}_{\tau;b}
+   + Q^{[1]}_{0; \sigma}\eta^{[2]}_{\sigma;a} \nonumber \\
+ \hspace{10pt} &+ E^{[1]}_{a;\sigma}\eta^{[1]}_{\sigma;b}
+   + E^{[1]}_{b;\sigma}\eta^{[1]}_{\sigma;a}
+   \big]\delta_{\omega_a+\omega_b} \label{eq:linres}
  \end{align}
@@
  Q^{[n]}_{0; \sigma_1\cdots \sigma_n} &=
- \frac{\partial^n Q_0}{\partial\lambda_{\sigma_1}\cdots\partial\lambda_{\sigma_n}} \\
+ \frac{\partial^n Q_0}{\partial\eta_{\sigma_1}\cdots\partial\eta_{\sigma_n}} \\
  E^{[n]}_{x;\sigma_1\cdots \sigma_n} &=
- \frac{\partial^n E_x}{\partial\lambda_{\sigma_1}\cdots\partial\lambda_{\sigma_n}} \\
-\lambda^{[n]}_{\sigma;x_1\dots x_n} &=
-\frac{\partial^n \lambda_\sigma}{\partial \epsilon_{x_1}\cdots\partial \epsilon_{x_n}}
+ \frac{\partial^n E_x}{\partial\eta_{\sigma_1}\cdots\partial\eta_{\sigma_n}} \\
+\eta^{[n]}_{\sigma;x_1\dots x_n} &=
+\frac{\partial^n \eta_\sigma}{\partial \epsilon_{x_1}\cdots\partial \epsilon_{x_n}}
  \end{align}
@@
 The response equations, determining the tensor of
-perturbed wave function parameters $\vect{\lambda}^{[n]}$, are obtained by differentiating the
-variational condition on the time-averaged quasienergy Eq.
-\eqref{eq:var-cond}:\autocite{Olsen1985-nr, Helgaker1992-ph, Christiansen1998-pe}
+perturbed wave function parameters $\vect{\eta}^{[n]}$, are obtained by differentiating the
+time-averaged quasienergy variational condition Eq.
+\eqref{eq:var-cond} with respect to the perturbation parameters:\autocite{Olsen1985-nr, Helgaker1992-ph, Christiansen1998-pe}
 \begin{equation}
  R_{\sigma;x_1\cdots x_n}^{[n]} = \left[\prod_{i=1}^n \deriv{}{\epsilon_{x_i}}\right]
-\left. \deriv{\aveQ}{\lambda_\sigma}\right|_{\vect{\epsilon}=0} = 0.
+\left. \deriv{\aveQ}{\eta_\sigma}\right|_{\vect{\epsilon}=0} = 0,
 \end{equation}
-The determination of second-order properties requires the knowledge of
-the first-order response of the wave function:
+followed by evaluation at zero perturbation strength.
+The first order response of the wave function $\vect{\eta}^{[1]}$,
+needed for the determination of second order properties
+$Q_{ab}^{[2]}$, can be computed by solving the corresponding response
+equation:
 \begin{subequations}
@@
  R_{\sigma}^{[0]} &= Q_{0;\sigma}^{[1]} = 0 \\
- R_{\sigma;a}^{[1]} &= \sum_{\tau}Q_{0;\sigma\tau}^{[2]}\lambda_{\tau;a}^{[1]} + E_{a;\sigma}^{[1]} = 0.
+ R_{\sigma;a}^{[1]} &= \sum_{\tau}Q_{0;\sigma\tau}^{[2]}\eta_{\tau;a}^{[1]} + E_{a;\sigma}^{[1]} = 0.
  \label{eq:1st-order-response}
@@
 
-Let us consider in more detail the form of the linear response function
-Eq. \eqref{eq:linres}. The formal solution to the response equation Eq.
+\arrowthought{Let us consider in more detail the form of the linear
+response function Eq. \eqref{eq:linres}.}
+The formal solution to the response equation Eq.
 \eqref{eq:1st-order-response} can be obtained by inverting the
-\emph{Hessian}. Eq. \eqref{eq:linres} can then be rewritten as:
+\emph{Hessian}, $\mat{Q}_0^{[2]}$. Eq. \eqref{eq:linres} can then be rewritten as:
 \begin{equation}
@@
 \begin{equation}
+  \begin{aligned}
   \response{V_A}{V_B}{\omega_b}
   =
-  -\sum_{n>0}\left[
-  \frac{\braket{0|V_A|n}\braket{n|V_B|0}}{\omega_{n0}-\omega_b}
-  +
+  &-\sum_{n>0}
+  \bigg[
+  \frac{\braket{0|V_A|n}\braket{n|V_B|0}}{\omega_{n0}-\omega_b} \\
+  &+
   \frac{\braket{0|V_B|n}\braket{n|V_A|0}}{\omega_{n0}-\omega_a}
-  \right]\delta_{\omega_a+\omega_b}.
+  \bigg]
+  \delta_{\omega_a+\omega_b}.
+\end{aligned}
 \end{equation}
@@
 
-The derivation above holds for an exact state, a situation that is never
+\newthought{The derivation above holds for an exact state,} a situation that is never
 realized in common practice. However, response theory is
@@
 open-ended \acrshort{AO} density matrix-based approach presented by
-\citeauthor{Thorvaldsen2008-sg}.\autocite{Thorvaldsen2008-sg}
-By formulating the perturbation-strenght derivative of the time-averaged
+\citeauthor{Thorvaldsen2008-sg}\autocite{Thorvaldsen2008-sg}
+By formulating the perturbation-strength derivative of the time-averaged
 quasienergy Lagrangian in terms of the \acrshort{AO} density matrix,
@@
 equations can be obtained by differentiation of the \gls{TDSCF}
-equation, the stationarity condition on the density matrix.
+equation.
 The approach lends itself to an efficient \acrshort{AO}-based
@@
 leverage different response parameters elimination
-schemes,\autocite{Thorvaldsen2008-sg, Kristensen2008-hv} and is amenable
+schemes\autocite{Thorvaldsen2008-sg, Kristensen2008-hv} and is amenable
 to an open-ended, recursive implementation.\autocite{Ringholm2014-gx,
@@
 
-\section{SCF Response Theory for Quantum/Classical Polarizable Hamiltonians}\label{sec:csm-response}
+\section[Quantum/Classical Polarizable Response Theory]{SCF Response Theory for Quantum/Classical Polarizable Hamiltonians}\label{sec:csm-response}
 
 The development of response theory for quantum/classical polarizable
-Hamiltonians has moved almost at the same pace as the polarizable
+Hamiltonians has moved almost at the same pace as the
 models themselves.
@@
 
-Solvent relaxation is an important point when formulating response
-properties. Solvation dynamics occurs on timescales and through
+\arrowthought{Solvent relaxation is an important point when formulating response
+properties.} Solvation dynamics occurs on timescales and through
 relaxation modes that are typical of the solvent molecular structure and of the
@@
 of the polarization vector:
-The polarization vector will then be (approximately) a sum of two
-components:
 \begin{equation}
@@
 
-The variational setting introduced in Section \ref{sec:variational},
+\newthought{The variational setting introduced in Section
+\ref{sec:variational},}
 offers a series of advantages also in the formulation of response
@@
  \frac{\mathrm{i}}{2}\pertmat{T}\right]\pertmat{D}
- + \tilde{E}_\mathrm{xc}[\tilde{\rho}(\pertmat{D})] + h_\mathrm{nuc}
- + \frac{1}{2}{}^t\tilde{\p}\V\tilde{\p}
+ + \tilde{E}_\mathrm{xc}[\tilde{\rho}(\pertmat{D})] + h_\mathrm{nuc} \\
+ &+ \frac{1}{2}{}^t\tilde{\p}\V\tilde{\p}
  + {}^t\tilde{\p}\tilde{\mathbfup{s}}\pertmat{D},
@@
 \begin{equation}\label{eq:differential}
+  \begin{aligned}
   \mathcal{G}^{mn, abc}
-  =
+  &=
   \frac{\partial^{m+n+3}\mathcal{G}}{\partial(\mat{D}^T)^m\partial\p^n\partial\epsilon_a\partial\epsilon_b\partial\epsilon_c}
-  =
+  \\
+  &=
     \mathcal{E}^{m, abc} +
   \frac{\partial^{m+n+3}U_\mathrm{pol}}{\partial(\mat{D}^T)^m\partial\p^n\partial\epsilon_a\partial\epsilon_b\partial\epsilon_c} \text{.}
+  \end{aligned}
 \end{equation}
@@
     {}^t\tilde{\p}\Tr({\tilde{\mathbfup{s}}^a\pertmat{D}}) \\
-  &\eqtr
-  (\pertmat{h}^a + \pertmat{V}^{t,a} + \frac{1}{2}\pertmat{G}^{\gamma, a}(\pertmat{D}) + \pertmat{F}^{\Omega_a}_\mathrm{xc}
-  -\frac{\mathrm{i}}{2}\pertmat{T}^a)\pertmat{D}
-  + h_\mathrm{nuc}^a + {}^t\tilde{\p}\tilde{\mathbfup{s}}^a\pertmat{D}.
+  \eqtr
+  (\pertmat{h}^a + \pertmat{V}^{t,a} &+ \frac{1}{2}\pertmat{G}^{\gamma, a}(\pertmat{D}) + \pertmat{F}^{\Omega_a}_\mathrm{xc}
+  -\frac{\mathrm{i}}{2}\pertmat{T}^a)\pertmat{D} \\
+  &+ h_\mathrm{nuc}^a + {}^t\tilde{\p}\tilde{\mathbfup{s}}^a\pertmat{D}.
   \end{aligned}
 \end{equation}
-$\pertmat{F}^{\Omega_a}_\mathrm{xc}$ is the functional derivative matrix
-defined in terms of the perturbed overlap distributions
-$\tilde{\Omega}^a$.
+The integral derivative matrices involved in the previous expression
+were introduced in \noparcite[ref.][]{Thorvaldsen2008-sg}.
 Further perturbation-strength differentiations and evaluation at zero
@@
     + \mathcal{G}^{10,a}\mat{D}^{b}
-    + \mathcal{G}^{01,a}\p^{b}
-    - \mat{S}^{ab}\mat{W}
+    + \mathcal{G}^{01,a}\p^{b} \nonumber \\
+    &- \mat{S}^{ab}\mat{W}
     - \mat{S}^{a}\mat{W}^{b} \label{eq:second-order} \\
@@
 obtained in complete analogy with the handling of pertubed density
-matrices described in \noparcite[ref.][]{Thorvaldsen2008-sg}
-Denoting by the superscript $b_N$ a tuple of simulatenous perturbations,
-the polarization can be decomposed into frequency components.
+matrices described in \noparcite[ref.][]{Thorvaldsen2008-sg}.
+Denoting by the superscript $b_N$ a tuple of simultaneous perturbations,
+the polarization can be decomposed into frequency components $\p_\omega^{b_N}$.
 Similarly to the treatment of the perturbed density matrices,
 we introduce a partition of the frequency components of the
-polarization $\p_\omega^{b_N}$ into homogeneous (H) and particular (P)
+polarization into homogeneous (H) and particular (P)
 terms:
@@
 \begin{equation}\label{eq:response-equation}
- \left[\genHessian - \omega_{b_N}\genMetric \right]\rspParam{b_N} = \RHS{b_N}
+ \left[\genHessian - \omega_{b_N}\genMetric \right]\rspParam{b_N} =
+ \RHS{b_N},
 \end{equation}
 where the generalized Hessian $\genHessian$ and metric $\genMetric$
-matrices were introduced and are defined by their transformations on the
-response parameters $\rspParam{b_N}$:\autocite{Larsen2000-hj,
+matrices were introduced and are defined by the transformations of the
+response parameters $\rspParam{b_N}$ they induce:\autocite{Larsen2000-hj,
 Kjaergaard2008-hy}
@@
 The last two terms in Eq. \ref{eq:E2-lintra} are \emph{explicit} terms,
-as they explicitly involve the $N$-th order homogenous polarization
+as they explicitly involve the $N$-th order homogeneous polarization
 variational density.
@@
 dealing with frequency-dependent properties,\autocite{Cammi1999-rb, Tomasi2005-vm}
-will affect the explicit terms in Eq. \ref{eq:E2-lintra} which are
+will affect the explicit terms in Eq. \eqref{eq:E2-lintra}. These are
 evaluated by making use of the optical permittivity $\dield$ in the
-polarization equations \eqref{eq:part-polarization-eq}.
+polarization equations \eqref{eq:homo-polarization-eq} and \eqref{eq:part-polarization-eq}.
 The \acrlong{RHS} in the response equation only includes terms that

### 2016-10-04 Chapters/SolvationAndCorrelation.tex
Adjust equations

@@
 \chapter{Electron Correlation and Solvation}\label{ch:solvation-correlation}
-
-\begin{epigraphs}
-\qitem{Potius sero quam numquam}{--- \textsc{Titus Livius}, \textit{Ab Urbe Condita}}
-\end{epigraphs}
+\footnotetext{Potius sero quam numquam \\
+\hspace*{15pt} --- \textsc{Titus Livius}, \\ \hspace*{15pt}
+\textit{Ab Urbe Condita}}
 
@@
 with prof. T. Daniel Crawford (Virginia Tech) and Dr. Andrew
-C. Simmonett (NIH)
+C. Simmonett (NIH).
 
@@
 In Section \ref{sec:pcm-cc-models} I will derive the governing equation
-for the effective Lagrangian and show how previous work on
+for the effective Lagrangian, showing how previous work on
 continuum\autocite{Christiansen1999-tj, Cammi2009-gu, Caricato2011-tx}
@@
 \begin{equation}
+  \begin{aligned}
   \overline{O}
-  =
+  &=
+  O + \tilde{O} \\
+  &=
   O + \BCHfirst{O}{T} + \BCHsecond{O}{T} + \ldots
-  =
-  O + \tilde{O},
+  \end{aligned}
   \tag{\ref{eq:bch} from Chapter \ref{ch:QM}}
@@
   + \braket{\mathrm{HF} | \tilde{\s} | \mathrm{HF}}
-  + \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \overline{\s} | \mathrm{HF}}
+  + \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \overline{\s} |
+  \mathrm{HF}}.
   \end{aligned}
@@
 The \acrshort{BCH} expansions of the source term has been rewritten as:
-\begin{equation}
+\begin{equation}\label{eq:source-BCH}
   \overline{\s} = \s + \tilde{\s} = \s + \commutator{\s}{T}
-  + \frac{1}{2}\commutator{\commutator{\s}{T}}{T}
+  + \frac{1}{2}\commutator{\commutator{\s}{T}}{T},
 \end{equation}
@@
     \braket{\mathrm{HF} | \overline{\s} - \s_\mathrm{HF} | \mathrm{HF}}
-  + \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \overline{\s} - \s_\mathrm{HF} | \mathrm{HF}}
+  + \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \overline{\s} -
+  \s_\mathrm{HF} | \mathrm{HF}}.
   \end{aligned}
@@
   &=
-  \frac{1}{2}({}^t\p_\mathrm{HF}+{}^t\p)\V(\p_\mathrm{HF}+\p)
-  + ({}^t\p_\mathrm{HF}+{}^t\p)(\s_\mathrm{HF}+\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}) \\
+  \frac{1}{2}({}^t\p_\mathrm{HF}+{}^t\p)\V(\p_\mathrm{HF}+\p) \\
+  &+ ({}^t\p_\mathrm{HF}+{}^t\p)(\s_\mathrm{HF}+\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}) \\
   &=
@@
   + \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}
-  + {}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}
+  + {}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}.
   \end{aligned}
@@
   &+
-  \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}
-  + {}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}
+  \frac{1}{2}{}^t\p\V\p +
+  {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M} \\
+  &+ {}^t\p_\mathrm{HF}\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}
   + U_\mathrm{pol}^\mathrm{ref}
@@
   {}^\mathrm{PTE}\lag(\tamp{}, \tbar{})_\mathcal{M}
-  + \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M}
+  + \frac{1}{2}{}^t\p\V\p +
+  {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathcal{M},
 \end{equation}
@@
   + \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \overline{H} | \mathrm{HF}}
-  + U_\mathrm{pol}^\mathrm{ref}
+  + U_\mathrm{pol}^\mathrm{ref}.
 \end{equation}
@@
 the \acrshort{PTE} Hamiltonian is computed in \acrshort{MP}-partitioned form by
-employing the ``solvated'' orbitals and Fock matrix:
+employing the \enquote{solvated} orbitals and Fock matrix:
 \begin{equation}
@@
 \begin{equation}
+  \begin{aligned}
   {}^\mathrm{PTE}\lag(\tamp{}, \tbar{})_\mathcal{M}
-  =
+  &=
   G_0
   + \sum_{u=1}^{\mathcal{M}}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
-  + \braket{\mathrm{HF} | \overline{\Phi} | \mathrm{HF}}
-  + \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \overline{\Phi} | \mathrm{HF}}
+  + \braket{\mathrm{HF} | \overline{\Phi} | \mathrm{HF}} \\
+  &+ \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \overline{\Phi} | \mathrm{HF}}
+\end{aligned}
 \end{equation}
@@
 
-Before delving into the details of the governing equations for
-amplitudes, multipliers and polarization, we need to take a closer look
-at the source term in classical energy functional. This is the purpose
-of the next Section.
-
 \section{Source Terms in the Classical Energy Functional}
@@
 
-In the supermatrix formalism, the similarity-transformed terms will be
-written as:
-\begin{equation}
-  \overline{\s} = \s + \tilde{\s} = \s +
-  \commutator{\s}{T} +
-  \frac{1}{2}\commutator{\commutator{\s}{T}}{T}
-\end{equation}
+Let us take a closer look at the source term in the classical energy
+functional.
 It is important to single out which terms contribute in the various
-expectation values where the similarity-transformed sources will be
-involved.
+expectation values where the similarity-transformed sources Eq.
+\eqref{eq:source-BCH} are involved.
 This analysis is carried out in detail in Appendix \ref{sec:CC-expval}
@@
   \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCSD}
-  + \braket{\tbar{2} | \BCHfirst{\s}{T_3} | \mathrm{HF}}
-  + \braket{\tbar{3} | \BCHfirst{\s}{T_3} | \mathrm{HF}} \\
+  + \braket{\tbar{2} | \BCHfirst{\s}{T_3} | \mathrm{HF}} \\
+  &+ \braket{\tbar{3} | \BCHfirst{\s}{T_3} | \mathrm{HF}} \\
   &+ \frac{1}{2}
   \braket{\tbar{3} | \commutator{\commutator{\s}{T_2}}{T_2} |
+  \mathrm{HF}} \\
+  &+ \braket{\tbar{3} | \commutator{\commutator{\s}{T_1}}{T_3} |
   \mathrm{HF}}
-  + \braket{\tbar{3} | \commutator{\commutator{\s}{T_1}}{T_3} |
-  \mathrm{HF}}
   \end{aligned}
@@
 
-Differentiation of the effective Lagrangian in Eq. \eqref{eq:pted-cc}
+\begarrowthought{Differentiation of the effective Lagrangian in Eq.
+\eqref{eq:pted-cc}}
 with respect to the variational parameters $\tamp{\mu_i}$,
@@
     &+
-    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1} \nonumber \\
-    &+
-    {}^t\p\sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}
-    +
+    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
+    +
+    {}^t\p\sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
+    \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}} \nonumber \\
+    &+
     {}^t\p\sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} | \commutator{\commutator{\s}{T}}{\cluster{\mu_q}} | \mathrm{HF}}
@@
 
-It is thus convenient to devise approximations that are able to simplify the
+\newthought{It is thus convenient to devise approximations} that are able to simplify the
 Lagrangian and the governing equations by breaking or weakening the coupling.
@@
 \begin{equation}\label{eq:ptes-cc}
-  {}^{\mathrm{PTES}}\lag(\tamp{}, \tbar{}, \p)_\mathcal{M} =
-  {}^\mathrm{PTE}\lag(\tamp{}, \tbar{})_\mathcal{M}
-  + \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},
+  \begin{aligned}
+  \hspace{-20pt}  {}^{\mathrm{PTES}}\lag(\tamp{}, \tbar{},
+  \p)_\mathcal{M} &=
+  {}^\mathrm{PTE}\lag(\tamp{}, \tbar{})_\mathcal{M} \\
+  &+ \frac{1}{2}{}^t\p\V\p + {}^t\p\s_{\mathrm{N}}(\tamp{},
   \tbar{})_\mathrm{CCS}
+\end{aligned}
 \end{equation}

### 2016-10-04 FrontBackmatter/Acknowledgments.tex
Adjust equations

@@
 
-It has not been one day, it's been four years.\footnotetext{
-Just a perfect day \\
+It has not been one day, it's been four years.\footnotetext{Just a perfect day \\
 You made me forget myself \\
@@
 
-in no particular order:\footnotetext{
-And all of my friends were there, \\
+in no particular order:\footnotetext{And all of my friends were there, \\
 and no one cared \\
@@
 \end{itemize}
+
+\textitalian{A Grazia, Graziella e al Toblerone, qualcuno sa perché.}

### 2016-10-04 FrontBackmatter/Introduction.tex
Adjust equations

@@
 
-\section*{What Are the Electrons Really Doing in Molecules?}\footnotetext{
-[...] \textnorsk{og fordi jeg alltid har hatt en dragning mot det skjulte og hemmelige.}
+\section*{What Are the Electrons Really Doing in Molecules?}\footnotetext{[...] \textnorsk{og fordi jeg alltid har hatt en dragning mot det skjulte og hemmelige.}
 \\
 \hspace*{15pt}
---- \textnorsk{\textsc{K. O. Knausgård}, \textit{Om Høsten}}
+--- \textnorsk{\textsc{Karl Ove Knausgård}, \\ \hspace*{25pt} \textit{Om Høsten}}
 }

### 2016-10-04 FrontBackmatter/Acknowledgments.tex
Written acknowledgements

@@
 la pazienza infinita che dimostrate ogni giorno.}
-
-in no particular order:\footnotetext{And all of my friends were there, \\
+To my advisor, Luca Frediani, for leaving me ample freedom of exploring
+and experimenting in these past four years.
+It's been fun working with you. You have made me feel like a part of
+your family and this is also for Emil and Emma.
+To my co-advisor, Benedetta Mennucci, for always taking the time to
+answer my questions.
+To my friends from high school, for the good times we've had and for
+those we will have.\footnotetext{And all of my friends were there, \\
 and no one cared \\
@@
 --- \textsc{The Kinks}}
-\begin{itemize}
-  \item To \emph{Radovan Bast}, \emph{il miglior fabbro},
-  \item To \emph{Davide Michetti}, the inventor of nicknames,
-  \item To \emph{The Little Lebowski Urban Achievers}, Erlend, Ana and
-    Torgeir,
-  \item To the \emph{Schrodinger killed my cat!}
-\end{itemize}
+To the \enquote{Pisa diaspora} and in particular to Stefano, Federica,
+Laura and Marco. Remember, this is almost your last chance to visit me in Tromsø
+before I disinherit you!
+A special mention to Simone, my numerical analysis guru.
+To the \enquote{Repubblica di Cisanello}: Giacomo, Emilio and Romualdo.
+For the fun and conversations we've had. And for the, now legendary,
+late-night couch moving session across town with Daniele and Matteo.
+To all the MoLEcoLabbers, thank you for the great environment and
+hospitality whenever I visit Pisa.
+To all the members of CTCC, for the stimulating environment, the
+interesting discussions, the excellent cakes and all the wine bottles I
+have won.
+To Radovan, \emph{il miglior fabbro}, for all I have learnt
+from you.
+To my flatmates, Davide, Thibaud, Eivind, Audun, Jenn and JB. For making
+my daily life in Tromsø less lonely and more fun.
+Ah, yeah, sorry for the loud blasts of noise I like to listen to!
+To Alex, that always has a word of advice and encouragement.
+To Maarten and Anna, for their precious friendship and for pushing me to
+speak Norwegian more.
+To Laura, the Finnish lady who thought it would be fun to bring me along
+on a camping-and-bouldering trip to Lofoten. And it actually was!
+Despite the rain. And me not actually bouldering. But that's another
+story.
+To the Italian \enquote{colony} in Tromsø, thank you. Tromsø feels like
+home also because of you.
+To Laura, for not giving up on me, despite the evenings spent at the
+filmklubb.
+To Lindon, the coolest organic chemist. Thanks for all the music tips
+and the amazing time at the Rock Your Head Festival.\footnotetext{Oh, I rely all of my life \\
+All of my life \\
+On my Ffunny Ffrends \\
+--- \textsc{Unknown Mortal Orchestra}}
+To Glenn, for being an awesome friend.
+To Ravna and Lili, for your kindness and the hours spent at the swimming
+pool together.
+To \emph{The Little Lebowski Urban Achievers}, Erlend, Ana and Torgeir.
+One day we'll leave \emph{The Monroeville Zombies} in the dust, I am
+sure.
+To T. Daniel Crawford, for his hospitality, enthusiasm and patience with
+my slow progress.
+My visits to Blacksburg would not have been possible without the warm
+hospitality of Laurie, Charles and Alec. This is dedicated to you too!
+To Marco, because the world really is small and finding family where you
+least expect it is always a great surprise.
+To Yang, Töph, Pablo, Tommaso and Lorenzo, for the great fun we've had at the
+schools.
 

### 2016-10-04 Chapters/ContributionsSummary.tex
Going through typos and typography
DONE

@@
 A shorter version of the material in Section \ref{sec:software} has been
-accepted for publication in the proceedings of the \emph{Producing High
+published in the proceedings of the \emph{Producing High
 Performance and Sustainable Software for Molecular Simulation} workshop
@@
 
-The growing complexity of quantum chemical program packages requires that an
+\begarrowthought{The growing complexity of quantum chemical program
+  packages} requires that an
 appropriate strategy be devised to implement new features.
@@
 
-The idea of a \emph{modular programming paradigm} can be traced back to
-the dawn of computer science. \citeauthor{Dijkstra1968-zp}
+\arrowthought{The idea of a \emph{modular programming paradigm} can be traced back to
+the dawn of computer science.} \citeauthor{Dijkstra1968-zp}
 successfully used it in the development of the THE operating
@@
 cloud-based, free, code development services, such as hosted
-\acp{DVCS},\autocite{github, gitlab} continuous
-integration,\autocite{travis-ci, magnum-ci, appveyor-ci} code
-coverage analysis,\autocite{coveralls} static and dynamic code
-analyses,\autocite{coverity-scan} nightly regression testing, public
+\acp{DVCS},\footnote{\url{https://github.com/} \\
+\url{https://gitlab.com/}}
+continuous integration,\footnote{\url{https://travis-ci.org/} \\
+\url{https://www.appveyor.com/} \\
+\url{https://magnum-ci.com}}
+code coverage analysis,\footnote{\url{https://coveralls.io/}}
+static and dynamic code analyses,\footnote{\url{https://scan.coverity.com/}}
+nightly regression testing, public
 issue tracking, code review and so forth: adoption of a modern code
@@
 
-The \acrshort{PCM} is an ideal candidate for the creation of a solvation
+\arrowthought{The \acrshort{PCM} is an ideal candidate} for the creation of a solvation
 \acrshort{API}. Consider for example the schematic representation of a
@@
 implementation of \acrshort{AD}.\autocite{Bartholomew-Biggs2000-db, libtaylor}
-The adoption of \git as \acrshort{DVCS}, \cmake for cross-platform builds and
-automatic documentation deployment on \readthedocs simplify
+The adoption of git\footnote{\url{https://git-scm.com/}} as \acrshort{DVCS},
+CMake\footnote{\url{https://cmake.org/}} for cross-platform builds and
+automatic documentation deployment on ReadTheDocs\footnote{\url{https://readthedocs.org/}} simplify
 extensibility of the module and promote third-party contributions to the
@@
 
-Careful usage of the C++ object-oriented paradigm is the key to this
+Judicious usage of the C++ object-oriented paradigm is the key to this
 wide spectrum of functionalities. The library is in itself made of
@@
 The \acrshort{API} is implemented in ISO C99 as a \emph{context-aware} set of
-functions.\autocite{library-ronacher, context-api-example}
+functions.\autocite{library-ronacher}\footnote{\url{https://github.com/bast/context-api-example}}
 Upon initialization of the library, a \emph{context} object containing
@@
 proceed through manipulations of the context object.
+A context-aware \acrshort{API} offers a simple strategy for parallelization.
+The host program could spawn contexts on a per-process basis,
+limiting ownership and access to a context just to its parent process.
+Race conditions are trivially avoided in such a scenario.
+
 \pcmsolver introduces the concept of surface functions:
@@
 \pcmsolver \acrshort{API} functions might look like in an actual C or Fortran
-program, respectively.
+host program, respectively.
 
@@
                 ]{snippets/C_host.c}
-
-A context-aware \acrshort{API} offers a simple strategy for parallelization.
-The host program could spawn contexts on a per-process basis,
-limiting ownership and access to a context just to its parent process.
-Race conditions are trivially avoided in such a scenario.
 
@@
 
-Programming effort in this dissertation has not only been directed at the
+\arrowthought{Programming effort in this dissertation} has not only been directed at the
 creation of the \pcmsolver library, but also at interfacing it with a
@@
 programming strategy. We showed that it is entirely possible to adopt the same
-\acrshort{PCM} code and implementation "checkpoints" across altogether different
+\acrshort{PCM} code and implementation \enquote{checkpoints} across altogether different
 problem domains, Fig.~\ref{fig:algorithm}.
@@
   \item \acrshort{MEP} integrals for four-component wave functions,
-  \item the additional Fock matrix contributions and
+  \item the additional Fock matrix contributions, and
   \item the additional terms in the response equations.
@@
 
-Already~\citeauthor{Weijo2010-hy} had shown that using \gls{PWC} bases can lead
+Already~\citeauthor{Weijo2010-hy} had shown that using \gls{PWC} wavelet bases can lead
 to superior accuracy and convergence in the calculation of quantum mechanical
@@
 In this work we showed that even faster convergence can be achieved when
-\gls{PWL} bases are used instead.
+\gls{PWL} wavelet bases are used instead.
 Moreover, the same holds for the calculation of static electric properties.
@@
 \begin{equation}
-  \left[ \bi{S}_\mathrm{e}\left(2\pi + \bi{D}^\dagger_\mathrm{i}\right)
-  +
-  \left(2\pi - \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}
-  \right]\sigma =
-  -\left[\left(2\pi-\bi{D}_\mathrm{e}\right)
+  \begin{aligned}
+  \big[ \bi{S}_\mathrm{e}\big(2\pi + \bi{D}^\dagger_\mathrm{i}\big)
+  &+
+  \big(2\pi - \bi{D}_\mathrm{e}\big)\bi{S}_\mathrm{i}
+  \big]\sigma = \\
+  &-\big[\big(2\pi-\bi{D}_\mathrm{e}\big)
   -\bi{S}_\mathrm{e}\bi{S}_\mathrm{i}^{-1}
-  \left(2\pi-\bi{D}_\mathrm{i}\right)
-  \right]\esp,
+  \big(2\pi-\bi{D}_\mathrm{i}\big)
+  \big]\esp,
+  \end{aligned}
   \tag{\ref{eq:full-IEF} from Chapter \ref{ch:CSM}}
@@
 spherically symmetric interfaces and peculiar properties may arise.
-In this work, we analyzed both effect on the water-vapor and oil-water transfer
+In this work, we analyzed both effects on the water-vapor and oil-water transfer
 of \ce{Li+}, \ce{Br-}, acetone, \emph{para}-nitroaniline and the L0 dye.
@@
 Ville Weijo and Hui Cao. In particular, I derived the separation of
-the Coulomb singularity in its final form.
+the Coulomb singularity in its final form.\footnote{
+Equation (40) for $G_\mathrm{img}(\vect{r}, \vect{r}^\prime)$ in the
+final published version of the paper contains a misprint. The
+parentheses should not appear.}
 Moreover, I contributed the implementation and testing of the Green's function code.
@@
   \\
-  In preparation for submission to \textit{Phys. Chem. Chem. Phys.}
+  Submitted to \textit{Phys. Chem. Chem. Phys.}
   }
@@
 the measured signal.
-Our group has newly developed an open-ended methodology for the computation of
+Our group has recently developed an open-ended methodology for the computation of
 \acrshort{SCF} response functions\autocite{Thorvaldsen2008-sg, Ringholm2014-gx} and
-their residues.\autocite{Friese2015-kb}
+their single residues.\autocite{Friese2015-kb}
 These developments offer a route towards a synergistic experimental and
@@
 \paper{V} grafts a classical polarizable continuum approach to solvation on top
-of the open-ended methodology of \citeauthor{Thorvaldsen2008-sg}.
+of the open-ended methodology of \citeauthor{Thorvaldsen2008-sg}
 Still nowadays, continuum models represent a cost-effective methodology for the
@@
 \DALTON.\autocite{Cammi2003-qy, Frediani2005-nc, Ferrighi2010-pm}
-Together with coauthors Maarten T.~P.~Beerepoot and Yann Cornaton, I carried out
+Together with coauthors Maarten T. P. Beerepoot and Yann Cornaton, I carried out
 the multiphoton absorption calculations presented in the paper. I contributed

### 2016-10-04 Chapters/MathematicalResults.tex
DONE

@@
 \chapter{Some Mathematical Results}\label{app:mathematical-results}
-
-\begin{epigraphs}
-\qitem{\textitalian{Non temerai i terrori della notte non temerai il terrore}}{--- \textsc{CCCP}}
-\end{epigraphs}
 
@@
 No proofs or examples are given, the interested reader is referred to the
-monographs by \citeauthor{Ern2004-oo}, \citeauthor{Hsiao2008-xb} and
-\citeauthor{Sauter2011-an}\autocite{Ern2004-oo, Hsiao2008-xb, Sauter2011-an}
+monographs by \citeauthor{Ern2004-oo},\autocite{Ern2004-oo} \citeauthor{Hsiao2008-xb}\autocite{Hsiao2008-xb} and
+\citeauthor{Sauter2011-an}\autocite{Sauter2011-an}
 
@@
 expressions.
-For the \acrshort{CCS} model, this is straightforward.
+\arrowthought{For the \acrshort{CCS} model, this is straightforward.}
 The Lagrangian is:
@@
 \begin{equation}
-  \braket{\mu_n |
+\begin{aligned}
+  &\braket{\mu_n |
   \commutator{\commutator{\commutator{O}{T_{n_1}}}{\ldots}}{T_{n_k}} | \mathrm{HF}} \neq 0
-  \Longleftrightarrow
-  n - m_O + k \leq \sum_{i=1}^k n_i \leq n + m_O
+  \Longleftrightarrow \\
+  &n - m_O + k \leq \sum_{i=1}^k n_i \leq n + m_O
+\end{aligned}
 \end{equation}
@@
 
-The \acrshort{CCSD} Lagrangian is then:
+\arrowthought{The \acrshort{CCSD} Lagrangian is then:}
 \begin{equation}\label{eq:ccsd-lagrangian}
@@
  E_0
- + \sum_{u=1}^{2}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
- + \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
+ + \sum_{u=1}^{2}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u} \\
+ &+ \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
  &+ \braket{\tbar{1} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
- &+ \braket{\tbar{2} | \check{\Phi} + \commutator{\check{\Phi}}{T_2}
-   +\frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} | \mathrm{HF}}
+ &+ \braket{\tbar{2} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
+ &+ \braket{\tbar{2} |
+   \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} | \mathrm{HF}}
   \end{aligned}
@@
 
-The \acrshort{CCSDT} Lagrangian is built in the same exact manner:
+\arrowthought{The \acrshort{CCSDT} Lagrangian is built in the same exact manner:}
 \begin{equation}\label{eq:ccsdt-lagrangian}
@@
  E_0
- + \sum_{u=1}^{3}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
- + \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
+ + \sum_{u=1}^{3}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u} \\
+ &+ \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
  &+ \braket{\tbar{1} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} + \commutator{\check{\Phi}}{T_3} | \mathrm{HF}} \\
- &+ \braket{\tbar{2} | \check{\Phi} + \commutator{\check{\Phi}}{T_2}
-   +\frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{T_3} | \mathrm{HF}} \\
+ &+ \braket{\tbar{2} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
+ &+ \braket{\tbar{2} |
+   \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{T_3} | \mathrm{HF}} \\
  &+ \braket{\tbar{3} | \commutator{\check{\Phi}}{T_2}
-   + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{T_3}
-   + \commutator{\commutator{\check{\Phi}}{T_2}}{T_3}
+   + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2}
+   | \mathrm{HF}} \\
+ &+ \braket{\tbar{3} | \commutator{\check{\Phi}}{T_3} + \commutator{\commutator{\check{\Phi}}{T_2}}{T_3}
    | \mathrm{HF}}
@@
              = 0 \label{eq:ccsdt-singles-amplitudes}\\
-   \denom{\mu_2}\tamp{\mu_2} &+ \braket{\mu_2 |
+   \denom{\mu_2}\tamp{\mu_2}
+   &+ \braket{\mu_2 |
    \check{\Phi} + \commutator{\check{\Phi}}{T_2}
-   + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{T_3}
+   | \mathrm{HF}} \nonumber \\
+   &+ \braket{\mu_2 |
+    \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{T_3}
    | \mathrm{HF}}
              = 0 \label{eq:ccsdt-doubles-amplitudes} \\
-   \denom{\mu_3}\tamp{\mu_3} &+ \braket{\mu_3 |
+   \denom{\mu_3}\tamp{\mu_3} &+
+   \braket{\mu_3 |
    \commutator{\check{\Phi}}{T_2}
-   + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{T_3}
+   + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2}
+   | \mathrm{HF}} \nonumber \\
+   &+ \braket{\mu_3 |
+     \commutator{\check{\Phi}}{T_3}
    + \commutator{\commutator{\check{\Phi}}{T_2}}{T_3}
@@
     + \frac{1}{2}\commutator{\commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}}{T_2}
-    + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_3}
+    | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{3} |
+     \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_3}
     | \mathrm{HF}}
@@
     +  \commutator{\commutator{\check{\Phi}}{\cluster{\mu_2}}}{T_2}
-    +  \commutator{\commutator{\check{\Phi}}{\cluster{\mu_2}}}{T_3}
+    | \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{3} |
+      \commutator{\commutator{\check{\Phi}}{\cluster{\mu_2}}}{T_3}
     | \mathrm{HF}}
@@
 
-For the \acrshort{CCS} model, the expectation value of a one-electron operator
-is simply:
+\arrowthought{For the \acrshort{CCS} model, the expectation value of a one-electron operator
+is simply:}
 \begin{equation}\label{eq:ccs-expectation-value}
@@
   &=
-  \braket{\mathrm{HF} | O + \commutator{O}{T_1} | \mathrm{HF} }
-  + \braket{\tbar{1} | O + \commutator{O}{T_1}
+  \braket{\mathrm{HF} | O + \commutator{O}{T_1} | \mathrm{HF} } \\
+  &+ \braket{\tbar{1} | O + \commutator{O}{T_1}
   + \frac{1}{2}\commutator{\commutator{O}{T_1}}{T_1} | \mathrm{HF} }
@@
     O(\tamp{}, \tbar{})_\mathrm{CCS}
-  + \braket{\tbar{1} | \commutator{\check{O}}{T_2} | \mathrm{HF} }
-  + \braket{\tbar{2} | \commutator{\check{O}}{T_2} | \mathrm{HF} } \\
+  + \braket{\tbar{1} | \commutator{\check{O}}{T_2} | \mathrm{HF} } \\
+  &+ \braket{\tbar{2} | \commutator{\check{O}}{T_2} | \mathrm{HF} } \\
   &=
     O(\tamp{}, \tbar{})_\mathrm{CCS}
-  + \braket{\tbar{1} | \commutator{O}{T_2} | \mathrm{HF} }
-  + \braket{\tbar{2} | \commutator{O}{T_2}
+  + \braket{\tbar{1} | \commutator{O}{T_2} | \mathrm{HF} } \\
+  &+ \braket{\tbar{2} | \commutator{O}{T_2}
       + \commutator{\commutator{O}{T_1}}{T_2}
@@
     + \commutator{\commutator{O}{T_1}}{T_3}
-    + \frac{1}{2}\commutator{\commutator{O}{T_2}}{T_2}
+  | \mathrm{HF} } \\
+  &+ \braket{\tbar{3} |
+     \frac{1}{2}\commutator{\commutator{O}{T_2}}{T_2}
   | \mathrm{HF} }
@@
 \section{Selected Results in Functional Analysis}
+
+\begin{defin}[Continuity]\label{def:continuity}
+  A bilinear form on a normed vector space $V$ is \emph{bounded}, or
+  \emph{continuous}, if there is a constant $C$ such that
+  $\forall u, v \in V$:
+  \[
+  a(u, v) \leq C\lVert u\rVert\lVert v\rVert
+  \]
+\end{defin}
+
+\begin{defin}[Coercivity]\label{def:coercivity}
+  A bilinear form on a normed vector space $V$ is \emph{coercive},
+  or \emph{elliptic}, if there is a constant $\alpha >0$ such that
+  $\forall u \in V$:
+  \[
+  a(u, u) \geq \alpha\lVert u\rVert^2
+  \]
+\end{defin}
+Coercivity implies that no eigenvalue of the linear operator associated to the
+bilinear form can be zero, hence its invertibility.\autocite{Ern2004-oo}
 
@@
 \left\lbrace
-u \in L^p(\Omega) |
+u \in L^p(\Omega) \Bigg|
 \frac{u(\vect{r}) - u(\vect{r}^\prime)}{\| \vect{r} - \vect{r}^\prime\|^{s+\frac{d}{p}}}
@@
 part of $s$, $W^{s,p}(\Omega)$ is defined as
-\begin{equation}
+\begin{equation*}
 W^{s, p}(\Omega) =
@@
 \right\rbrace.
-\end{equation}
+\end{equation*}
 When $p=2$, we denote $H^s(\Omega) = W^{s,2}(\Omega)$.
 \end{defin}
-
-\begin{defin}[Continuity]\label{def:continuity}
-  A bilinear form on a normed vector space $V$ is \emph{bounded}, or
-  \emph{continuous}, if there is a constant $C$ such that
-  $\forall u, v \in V$:
-  \[
-  a(u, v) \leq C\lVert u\rVert\lVert v\rVert
-  \]
-\end{defin}
-
-\begin{defin}[Coercivity]\label{def:coercivity}
-  A bilinear form on a normed vector space $V$ is \emph{coercive},
-  or \emph{elliptic}, if there is a constant $\alpha >0$ such that
-  $\forall u \in V$:
-  \[
-  a(u, u) \geq \alpha\lVert u\rVert^2
-  \]
-\end{defin}
-Coercivity implies that no eigenvalue of the linear operator associated to the
-bilinear form can be zero, hence its invertibility.\autocite{Ern2004-oo}
 
@@
 u = \bi{S}_\mathrm{i}(\partiali u)
-- \bi{D}_\mathrm{i}(\ui) + \int_{\Omegai} \Gi f_\mathrm{i}\diff \vect{r}^\prime
+- \bi{D}_\mathrm{i}(\ui) + \int_{\Omegai}\diff \vect{r}^\prime \Gi f_\mathrm{i}
 \end{equation}
@@
 u = -\bi{S}_\mathrm{e}(\partiale u)
-+ \bi{D}_\mathrm{e}(\ue) + \int_{\Omegae}\Ge f_\mathrm{e}\diff \vect{r}^\prime
++ \bi{D}_\mathrm{e}(\ue) + \int_{\Omegae}\diff \vect{r}^\prime\Ge f_\mathrm{e}
 \end{equation}
@@
     \frac{1}{2}\ui = \bi{S}_\mathrm{i}(\partiali u)
-    - \bi{D}_\mathrm{i}(\ui) + \int_{\Omegai}\Gi f_\mathrm{i}\diff \vect{r}^\prime
+    - \bi{D}_\mathrm{i}(\ui) + \int_{\Omegai}\diff \vect{r}^\prime\Gi f_\mathrm{i}
   \end{equation}
@@
     \frac{1}{2}\ue = -\bi{S}_\mathrm{e}(\partiale u)
-    + \bi{D}_\mathrm{e}(\ue) + \int_{\Omegae}\Ge f_\mathrm{e}\diff \vect{r}^\prime
+    + \bi{D}_\mathrm{e}(\ue) + \int_{\Omegae}\diff \vect{r}^\prime\Ge f_\mathrm{e}
   \end{equation}
@@
 \begin{empheq}[left={h(\vect{r}) = \empheqlbrace}]{align}
-    &\int_{\mathbb{R}^3}\Gi\rhoi(\vect{r}^\prime)\diff \vect{r}^\prime  \quad \vect{r} \in \Omegai \\
-    &\int_{\mathbb{R}^3}\Ge\rhoi(\vect{r}^\prime)\diff \vect{r}^\prime  \quad \vect{r} \in \Omegae
+    &\int_{\mathbb{R}^3}\diff \vect{r}^\prime\Gi\rhoi(\vect{r}^\prime)  \quad \vect{r} \in \Omegai \\
+    &\int_{\mathbb{R}^3}\diff \vect{r}^\prime\Ge\rhoi(\vect{r}^\prime)  \quad \vect{r} \in \Omegae
 \end{empheq}
@@
 \begin{equation}
-  \phi(\vect{r}) = (\bi{N}\rhoi)(\vect{r}) = \int_{\mathbb{R}^3}\Gi\rhoi(\vect{r}^\prime)\diff \vect{r}^\prime
+  \phi(\vect{r}) = (\bi{N}\rhoi)(\vect{r}) = \int_{\mathbb{R}^3}\diff \vect{r}^\prime\Gi\rhoi(\vect{r}^\prime)
    = \left.h(\vect{r})\right|_{\Omegai}

### 2016-10-04 Chapters/SolvationAndCorrelation.tex
DONE

@@
 \chapter{Electron Correlation and Solvation}\label{ch:solvation-correlation}
-\footnotetext{Potius sero quam numquam \\
-\hspace*{15pt} --- \textsc{Titus Livius}, \\ \hspace*{15pt}
-\textit{Ab Urbe Condita}}
+\epigraph{Potius sero quam numquam}{
+--- \textsc{Titus Livius}, \textit{Ab Urbe Condita}}
 
@@
 
-The first approximate scheme truncates the source term to its
+The first approximate scheme\footnote{Thanks to prof. Marco Caricato
+(University of Kansas) for clarifying some details of his original
+derivation of the \gls{PTES} scheme.} truncates the source term to its
 \acrshort{CCS} expectation value, yielding the effective
@@
    \tampEq{\mu_q}(\tamp{}, \tbar{}, \p)  &=
-   \denom{\mu_q}\tamp{\mu_q} + \braket{\mu_q | \overline{\Phi} | \mathrm{HF}}
-   + {}^t\p\braket{\mu_q | \check{\s} | \mathrm{HF}}\delta_{\mu_q\mu_1}
+   \denom{\mu_q}\tamp{\mu_q} + \braket{\mu_q | \overline{\Phi} | \mathrm{HF}} \nonumber \\
+   &+ {}^t\p\braket{\mu_q | \check{\s} | \mathrm{HF}}\delta_{\mu_q\mu_1}
              = 0 \\
@@
    {}^{\mathrm{PTES}}\tampEq{\mu_q}(\tamp{}, \tbar{}, \p)  &=
-   \denom{\mu_q}\tamp{\mu_q} + \braket{\mu_q | \overline{\Phi} | \mathrm{HF}}
-   + {}^t\p\braket{\mu_q | \check{\s} | \mathrm{HF}}\delta_{\mu_q\mu_1}
+   \denom{\mu_q}\tamp{\mu_q} + \braket{\mu_q | \overline{\Phi} |
+   \mathrm{HF}} \nonumber \\
+   &+ {}^t\p\braket{\mu_q | \check{\s} | \mathrm{HF}}\delta_{\mu_q\mu_1}
              = 0 \label{eq:ptes-cc-amplitudes} \\
@@
 where no truncation for the source term is introduced.
-Both the amplitudes and multipliers equations are formally identical to
+\newthought{Both the amplitudes and multipliers equations} are formally identical to
 the \acrshort{PTED} equations \eqref{eq:pted} but are coupled to
@@
 \begin{equation}\label{eq:pte-s-cc}
-    {}^{\mathrm{PTE(S)}}\lag(\tamp{}, \tbar{}, \p)_\mathcal{M} =
-  {}^\mathrm{PTE}\lag(\tamp{}, \tbar{})_\mathcal{M}
-  + \frac{1}{2}{}^t\p\V\p + {}^t\p\braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}}
+  \begin{aligned}
+  \hspace{-20pt}  {}^{\mathrm{PTE(S)}}\lag(\tamp{}, \tbar{},
+  \p)_\mathcal{M} &=
+  {}^\mathrm{PTE}\lag(\tamp{}, \tbar{})_\mathcal{M} \\
+  &+ \frac{1}{2}{}^t\p\V\p + {}^t\p\braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}}
+\end{aligned}
 \end{equation}
@@
     \denom{\mu_q}\tbar{\mu_q} +
-    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}} +
-    \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
+    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}}
+    \nonumber \\
+    &+\sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
     \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}}
@@
 We will now develop a perturbative many-body expansion of the effective
-\acrshort{PTED}-\acrshort{CC} Lagrangian \eqref{eq:pted-cc}
+\acrshort{PTED}-\acrshort{CC} Lagrangian.
 The use of a \acrshort{CC} ansatz for the wave function ensures term-by-term
@@
 \begin{equation}%\label{eq:1st-order-mult}
-  \bar{\Omega}_{\mu_q}^{[1]} =
+  \begin{aligned}
+  \bar{\Omega}_{\mu_q}^{[1]} &=
     \denom{\mu_q}\tbar{\mu_q}^{[1]}
     + \braket{\mathrm{HF} | \commutator{\Phi}{\cluster{\mu_q}} | \mathrm{HF}}
-    + {}^t\p^{[1]}\braket{\mathrm{HF} |
+    \nonumber \\
+    &+ {}^t\p^{[1]}\braket{\mathrm{HF} |
     \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1} =0
+  \end{aligned}
 \end{equation}
@@
 quantum/classical coupling.
-Moreover, as is the case \emph{in vacuo}, there are no triples contributions to
-the first order amplitudes and multipliers: $\tamp{\mu_3}^{[1]} = \tbar{\mu_3}^{[1]} = 0$.
+Moreover, as is the case \emph{in vacuo}, there are neither triples
+amplitudes nor multipliers
+to first order: $\tamp{\mu_3}^{[1]} = \tbar{\mu_3}^{[1]} = 0$.
 
@@
   \Omega_{\mu_q}^{[2]} &= \denom{\mu_q}\tamp{\mu_q}^{[2]}
-  + \braket{\mu_q | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
-  + {}^t\p^{[2]}\braket{\mu_q | \s | \mathrm{HF}}\delta_{\mu_q\mu_1} \\
+  + \braket{\mu_q | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}} \\
+  &+ {}^t\p^{[2]}\braket{\mu_q | \s | \mathrm{HF}}\delta_{\mu_q\mu_1} \\
   &+ {}^t\p^{[1]}\braket{\mu_q | \commutator{\s}{T^{[1]}} |
@@
   \Omega_{\mu_2}^{[2]} &= \denom{\mu_2}\tamp{\mu_2}^{[2]}
-  + \braket{\mu_2 | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
-  + {}^t\p^{[1]}\braket{\mu_2 | \commutator{\s}{T_2^{[1]}} |
+  + \braket{\mu_2 | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}} \nonumber
+  \\
+  &+ {}^t\p^{[1]}\braket{\mu_2 | \commutator{\s}{T_2^{[1]}} |
   \mathrm{HF}}
@@
     \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
-    + {}^t\p^{[1]}
+    \nonumber \\
+    &+ {}^t\p^{[1]}
     \sum_{u=1}^\mathcal{M}\braket{\tbar{u}^{[1]} |
@@
 to second order are given as:
+
 \begin{subequations}
@@
       &+ {}^t\p^{[2]}\braket{\mathrm{HF} |
-      \commutator{\s}{\cluster{\mu_1}} | \mathrm{HF}}
-      + {}^t\p^{[1]}
+      \commutator{\s}{\cluster{\mu_1}} | \mathrm{HF}} \nonumber \\
+      &+ {}^t\p^{[1]}
       \braket{\tbar{1}^{[1]} |
@@
       \commutator{\Phi}{\cluster{\mu_2}} | \mathrm{HF}}
-      + {}^t\p^{[1]}
+      \nonumber \\
+      &+ {}^t\p^{[1]}
       \sum_{u=1}^\mathcal{M}\braket{\tbar{u}^{[1]} |
@@
 \end{subequations}
-\citeauthor{Koch1997-nm} showed that, assuming real wave functions, the
+\noindent\citeauthor{Koch1997-nm} showed that, assuming real wave functions, the
 \emph{in vacuo} second order amplitudes and multipliers equations are
@@
 \begin{equation}\tag{\ref{eq:2nd-order} from Chapter \ref{ch:QM}}
+  \begin{aligned}
   \denom{\mu_3}\tamp{\mu_3}^{[2]}
-  = - \braket{\mu_i | \commutator{\Phi}{T_2^{[1]}} | \mathrm{HF}}
-  = - \braket{\tbar{2}^{[1]} | \commutator{\Phi}{\cluster{\mu_3}} |
+  &= - \braket{\mu_i | \commutator{\Phi}{T_2^{[1]}} | \mathrm{HF}} \\
+  &= - \braket{\tbar{2}^{[1]} | \commutator{\Phi}{\cluster{\mu_3}} |
   \mathrm{HF}}
   = \denom{\mu_3}\tbar{\mu_3}^{[2]},
+\end{aligned}
 \end{equation}
@@
 \begin{equation}\label{eq:pted-mbpt-1st}
-  G^{[1]} = \braket{\mathrm{HF} | \Phi | \mathrm{HF}}
+  G^{[1]} = \braket{\mathrm{HF} | \Phi | \mathrm{HF}}.
 \end{equation}
@@
 \begin{equation}
-  G_\mathrm{HF} = G^{[0]} + G^{[1]}
+  G_\mathrm{HF} = G^{[0]} + G^{[1]},
 \end{equation}
@@
 
-The second order energy correction will be formed including first order
+\arrowthought{The second order energy correction will be formed} including first order
 amplitudes and polarization, while using zeroth order multipliers:
@@
   G^{[2]} &=
-  \braket{\mathrm{HF} | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}}
-  + \frac{1}{2}{}^t\p^{[1]}\V\p^{[1]} +
+  \braket{\mathrm{HF} | \commutator{\Phi}{T^{[1]}} | \mathrm{HF}} \\
+  &+ \frac{1}{2}{}^t\p^{[1]}\V\p^{[1]} +
   {}^t\p^{[1]}\s_{\mathrm{N}}^{[1]}(\tamp{}^{[1]}, \tbar{}^{[0]}) \\
@@
   +
-  {}^t\p^{[1]}\braket{\mathrm{HF} | \commutator{\s}{T^{[1]}_1} | \mathrm{HF}}
-  = E^{[2]}
+  {}^t\p^{[1]}\braket{\mathrm{HF} | \commutator{\s}{T^{[1]}_1} |
+  \mathrm{HF}} \\
+  &= E^{[2]}
   \end{aligned}
@@
 
-The third order energy correction will be formed including first order
+\arrowthought{The third order energy correction will be formed} including first order
 amplitudes, multipliers and polarization:
@@
     + {}^t\p^{[1]}
-    \left\lbrace
       \braket{\tbar{1}^{[1]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
-    + \braket{\tbar{2}^{[1]} | \commutator{\s}{T^{[1]}_2} | \mathrm{HF}}
-    \right\rbrace
-  \end{aligned}
-\end{equation}
-
-The fourth order energy corrections will be formed including second
+      \\
+    &+ {}^t\p^{[1]}
+     \braket{\tbar{2}^{[1]} | \commutator{\s}{T^{[1]}_2} | \mathrm{HF}}
+  \end{aligned}
+\end{equation}
+
+\arrowthought{The fourth order energy corrections will be formed} including second
 order amplitudes and polarization, while using first order multipliers:
-\begin{equation}\label{eq:pted-mbpt-4th}
+\begin{equation}
   \begin{aligned}
     G^{[4]} &=
-    \sum_{u=1}^2\braket{\tbar{u}^{[1]} | \commutator{\Phi}{T^{[2]}} | \mathrm{HF}}
-    +
+    \sum_{u=1}^2\braket{\tbar{u}^{[1]} | \commutator{\Phi}{T^{[2]}} |
+    \mathrm{HF}} \\
+    &+
     \sum_{u=1}^2\braket{\tbar{u}^{[1]} |
     \frac{1}{2}\commutator{\commutator{\Phi}{T^{[1]}}}{T^{[1]}} | \mathrm{HF}}
-    \\
-    &+\frac{1}{2}{}^t\p^{[2]}\V\p^{[2]}
-    + {}^t\p^{[2]}\s_{\mathrm{N}}^{[2]}(\tamp{}^{[2]}, \tbar{}^{[1]})
-    + {}^t\p^{[1]}\s_{\mathrm{N}}^{[3]}(\tamp{}^{[2]}, \tbar{}^{[1]}) \\
-    &= E^{[4]}
-    + \frac{1}{2}{}^t\p^{[2]}\V\p^{[2]} \\
-    &+{}^t\p^{[2]}
-    \left\lbrace
-    \braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}} | \mathrm{HF}}
-    +
-    \sum_{u=1}^2\braket{\tbar{u}^{[1]}|
-    \commutator{\s}{T^{[1]}} | \mathrm{HF} }
-    \right\rbrace \\
-    &+{}^t\p^{[1]}
-    \left\lbrace
-    \sum_{u=1}^2\braket{\tbar{u}^{[1]}|
-    \commutator{\s}{T^{[2]}} | \mathrm{HF}}
-    +
-    \sum_{u=1}^2\braket{\tbar{u}^{[1]}|
-    \frac{1}{2}
-    \commutator{\commutator{\s}{T^{[1]}}}{T^{[1]}}
-    |
-    \mathrm{HF}}
-    \right\rbrace \\
+    +\frac{1}{2}{}^t\p^{[2]}\V\p^{[2]}
+    \\
+    &+ {}^t\p^{[2]}\s_{\mathrm{N}}^{[2]}(\tamp{}^{[2]}, \tbar{}^{[1]})
+    + {}^t\p^{[1]}\s_{\mathrm{N}}^{[3]}(\tamp{}^{[2]}, \tbar{}^{[1]})
+  \end{aligned}
+\end{equation}
+The first two terms are recognized to be equal to $E^{[4]}$, so that:
+\begin{equation}
+  \begin{aligned}
+    G^{[4]}
   &= E^{[4]}
@@
     \braket{\tbar{1}^{[1]} | \commutator{\s}{T_1^{[2]} + T_2^{[2]}} | \mathrm{HF}}
-    +
+    \\
+    &+
     {}^t\p^{[1]}
@@
     \mathrm{HF}}
-    +
+    \\
+    &+
     {}^t\p^{[1]}
@@
 second order polarization can be simplified further:
-\begin{equation}
+\begin{equation}\label{eq:pted-mbpt-4th}
   \begin{aligned}
@@
       \braket{\mathrm{HF} | \commutator{\s}{T_1^{[2]}} | \mathrm{HF}}
-    - \frac{1}{2}{}^t\p^{[2]}
-       \braket{\tbar{1}^{[2]}| \s | \mathrm{HF}} \\
-    &+ \frac{1}{2}{}^t\p^{[2]}
+    % Mmm check?
+    %- \frac{1}{2}{}^t\p^{[2]}
+    %   \braket{\tbar{1}^{[2]}| \s | \mathrm{HF}}
+    \\
+    &+ \frac{1}{2}{}^t\p^{[2]}\left(
       \braket{\tbar{1}^{[1]}| \commutator{\s}{T^{[1]}} | \mathrm{HF} }
-    + \frac{1}{2}{}^t\p^{[2]}
-       \braket{\tbar{2}^{[1]}| \commutator{\s}{T_2^{[1]}} | \mathrm{HF} } \\
-    &+ \frac{1}{2}{}^t\p^{[2]}\braket{\tbar{1}^{[1]}| \commutator{\s}{T^{[1]}} | \mathrm{HF} }
-    + \frac{1}{2}{}^t\p^{[2]}\braket{\tbar{2}^{[1]}| \commutator{\s}{T_2^{[1]}} | \mathrm{HF} } \\
+    +
+       \braket{\tbar{2}^{[1]}| \commutator{\s}{T_2^{[1]}} | \mathrm{HF}
+       }\right) \\
+    &+ \frac{1}{2}{}^t\p^{[2]}\left(\braket{\tbar{1}^{[1]}| \commutator{\s}{T^{[1]}} | \mathrm{HF} }
+    + \braket{\tbar{2}^{[1]}|
+    \commutator{\s}{T_2^{[1]}} | \mathrm{HF} } \right)\\
     &+
     {}^t\p^{[1]}
-    \braket{\tbar{1}^{[1]} | \commutator{\s}{T_1^{[2]} + T_2^{[2]}} | \mathrm{HF}}
-    +
+    \braket{\tbar{1}^{[1]} | \commutator{\s}{T_1^{[2]} + T_2^{[2]}} |
+    \mathrm{HF}} \\
+    &+
     \highlight{brewerRed}{
@@
     |
-    \mathrm{HF}}
-    +
+    \mathrm{HF}} \\
+    &+
     {}^t\p^{[1]}
@@
 
-Eventually, the fifth order energy correction is formed including second order
+\arrowthought{Eventually, the fifth order energy correction is formed} including second order
 amplitudes, multipliers and polarization:
-\begin{equation}\label{eq:pted-mbpt-5th}
+\begin{equation}
   \begin{aligned}
@@
     \commutator{\commutator{\Phi}{T^{[2]}}}{T^{[2]}}
-    | \mathrm{HF}}
-    +
+    | \mathrm{HF}} \\
+    &+
     \sum_{u=1}^2\braket{\tbar{u}^{[1]}|
@@
       \commutator{\Phi}{T^{[2]}}
-    | \mathrm{HF}}
-    +
+    | \mathrm{HF}} \\
+    &+
     \sum_{u=1}^3\braket{\tbar{u}^{[2]}|
@@
     &+ {}^t\p^{[2]}\s_{\mathrm{N}}^{[3]}(\tamp{}^{[2]}, \tbar{}^{[2]})
-    + {}^t\p^{[1]}\s_{\mathrm{N}}^{[4]}(\tamp{}^{[2]}, \tbar{}^{[2]}) \\
-    &= E^{[5]}
-    + {}^t\p^{[2]}
+    + {}^t\p^{[1]}\s_{\mathrm{N}}^{[4]}(\tamp{}^{[2]}, \tbar{}^{[2]})
+  \end{aligned}
+\end{equation}
+This can be rewritten in terms of $E^{[5]}$ and a number of polarization
+terms:
+\begin{equation}\label{eq:pted-mbpt-5th}
+  \begin{aligned}
+    G^{[5]} &=
+    E^{[5]}
+    + {}^t\p^{[2]} \Bigg(
     \sum_{u=1}^2\braket{\tbar{u}^{[1]} |
     \commutator{\s}{T^{[2]}} | \mathrm{HF}} \\
-    &+ {}^t\p^{[2]}
+    &+
     \sum_{u=1}^2\braket{\tbar{u}^{[1]} |
     \frac{1}{2}\commutator{\commutator{\s}{T^{[1]}}}{T^{[1]}}
-    | \mathrm{HF}}
-    + {}^t\p^{[2]}
+    | \mathrm{HF}} \\
+    &+
     \sum_{u=1}^3\braket{ \tbar{u}^{[2]}
                                  | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
-    \\
-    &+ {}^t\p^{[1]}
+                                 \Bigg)
+    \\
+    &+ {}^t\p^{[1]} \Bigg(
     \sum_{u=1}^3\braket{ \tbar{u}^{[2]}
-    \commutator{\s}{T^{[2]}} | \mathrm{HF}}
-    + {}^t\p^{[1]}
+    \commutator{\s}{T^{[2]}} | \mathrm{HF}} \\
+    &+
     \sum_{u=1}^3\braket{ \tbar{u}^{[2]}
@@
     | \mathrm{HF}} \\
-    &+ {}^t\p^{[1]}
+    &+
     \sum_{u=1}^2\braket{\tbar{u}^{[1]} |
@@
     | \mathrm{HF}}
-    \\
-  &= E^{[5]}
-    + {}^t\p^{[2]}\braket{\tbar{1}^{[1]} | \commutator{\s}{T_1^{[2]} +
+              \Bigg)
+  \end{aligned}
+\end{equation}
+Finally, consideration of the excitation ranks involved in the
+quantum/classical coupling terms yields
+\begin{equation}\label{eq:pted-mbpt-5th}
+  \begin{aligned}
+    G^{[5]} &=
+     E^{[5]}
+    + {}^t\p^{[2]}\Bigg(\braket{\tbar{1}^{[1]} | \commutator{\s}{T_1^{[2]} +
     T_2^{[2]}} | \mathrm{HF}} \\
     &+ \highlight{brewerRed}{
-    {}^t\p^{[2]}\braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[2]} + T_3^{[2]}} | \mathrm{HF}}
+    \braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[2]} + T_3^{[2]}} | \mathrm{HF}}
     } \\
-    &+ {}^t\p^{[2]}\braket{\tbar{1}^{[2]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
-    + {}^t\p^{[2]}
+    &+ \braket{\tbar{1}^{[2]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
+    +
     \braket{\tbar{1}^{[1]} |
@@
     | \mathrm{HF}} \\
-    &+ {}^t\p^{[2]}
+    &+
     \braket{\tbar{2}^{[1]} |
     \commutator{\commutator{\s}{T_1^{[1]}}}{T_2^{[1]}}
-    | \mathrm{HF}}
-    +
-    {}^t\p^{[1]}
+    | \mathrm{HF}} \\
+    &+
+    {}^t\p^{[1]} \Bigg(
     \braket{\tbar{1}^{[2]} | \commutator{\s}{T_1^{[2]} + T_2^{[2]}} |
@@
     \highlight{brewerRed}{
-    {}^t\p^{[1]}
     \braket{\tbar{2}^{[2]} | \commutator{\s}{T_2^{[2]} + T_3^{[2]}} | \mathrm{HF}}
@@
     \highlight{brewerRed}{
-    {}^t\p^{[1]}
     \braket{\tbar{3}^{[2]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
@@
     &+
-    {}^t\p^{[1]}
     \braket{\tbar{1}^{[2]} |
     \frac{1}{2}\commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[1]}}
-    | \mathrm{HF}}
-    +
-    {}^t\p^{[1]}
+    | \mathrm{HF}} \\
+    &+
     \braket{\tbar{2}^{[2]} |
@@
     \highlight{brewerRed}{
-    {}^t\p^{[1]}
     \braket{\tbar{3}^{[2]} |
@@
     } \\
-    &+ {}^t\p^{[1]}
+    &+
     \braket{\tbar{1}^{[1]} |
     \commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[2]}}
-    | \mathrm{HF}}
-    + {}^t\p^{[1]}
+    | \mathrm{HF}} \\
+    &+
     \braket{\tbar{2}^{[1]} |
@@
     | \mathrm{HF}} \\
-    &+ {}^t\p^{[1]}
+    &+
     \braket{\tbar{2}^{[1]} |
@@
     | \mathrm{HF}}
+    \Bigg)
   \end{aligned}
@@
     \commutator{\commutator{\Phi}{T_2^{[1]}}}{T_2^{[1]}}
-    | \mathrm{HF}}
-  +
+    | \mathrm{HF}} \\
+  &+
   \highlight{brewerYellow}{
@@
   + {}^t\p^{[1]}\braket{\tbar{2}^{[2]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}} \\
-  &+ {}^t\p^{[1]}\braket{\tbar{3}^{[2]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
-  + {}^t\p^{[1]}\braket{\tbar{3}^{[2]} |
+  &+ {}^t\p^{[1]}\braket{\tbar{3}^{[2]} | \commutator{\s}{T_3^{[2]}} |
+  \mathrm{HF}} \\
+  &+ {}^t\p^{[1]}\braket{\tbar{3}^{[2]} |
     \frac{1}{2}\commutator{\commutator{\s}{T_2^{[1]}}}{T_2^{[1]}}
@@
 
-In the \acrshort{PTE(S)} scheme, the amplitudes equations are uncoupled
+\arrowthought{In the \acrshort{PTE(S)} scheme,} the amplitudes equations are uncoupled
 from the multipliers and polarization equations. \acrshort{MBPT} for
@@
     &=
-    - \left(
-      \braket{\tbar{2}^{[1]} | \commutator{\Phi}{\cluster{\mu_q}} | \mathrm{HF}}
-      + {}^t\p^{[2]}\braket{\mathrm{HF} |
+    -
+      \braket{\tbar{2}^{[1]} | \commutator{\Phi}{\cluster{\mu_q}} |
+      \mathrm{HF}} \\
+    &- {}^t\p^{[2]}\braket{\mathrm{HF} |
       \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
-    \right)
     = 0
@@
 
-In the \acrshort{PTES} scheme, two different polarizations are
+\arrowthought{In the \acrshort{PTES} scheme,} two different polarizations are
 introduced to achieve an approximate decoupling of the governing
@@
     &=
-    - \braket{\tbar{2}^{[1]} | \commutator{\Phi}{\cluster{\mu_q}} | \mathrm{HF}}
-    - {}^t\p^{[2]}\braket{\mathrm{HF} |
+    - \braket{\tbar{2}^{[1]} | \commutator{\Phi}{\cluster{\mu_q}} |
+    \mathrm{HF}} \\
+    &- {}^t\p^{[2]}\braket{\mathrm{HF} |
       \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
@@
   | \mathrm{HF}}
-  + \braket{\tbar{2} | \check{\Phi} | \mathrm{HF}}
-  + \frac{1}{2}{}^t\p\V\p +
+  + \braket{\tbar{2} | \check{\Phi} | \mathrm{HF}} \\
+  &+ \frac{1}{2}{}^t\p\V\p +
   {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathrm{CC2}
@@
 \eqref{eq:cc2-expectation-value}
-The multipliers equations now become:
+The amplitudes equations now become:
 \begin{subequations}\label{eq:pted-cc2-amplitudes}
@@
     \braket{\mu_1 | \check{\Phi} +
-    \commutator{\check{\Phi}}{Q_2} | \mathrm{HF}}
-    + {}^t\p\braket{\mu_1 | \check{\s} +
+    \commutator{\check{\Phi}}{Q_2} | \mathrm{HF}} \\
+    &+ {}^t\p\braket{\mu_1 | \check{\s} +
     \commutator{\check{\s}}{Q_2} | \mathrm{HF}}
@@
     \denom{\mu_1}\tbar{\mu_1} &+
-     \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}}
-   + \braket{\tbar{1} |
+     \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} |
+     \mathrm{HF}} \nonumber \\
+   &+ \braket{\tbar{1} |
        \commutator{\check{\Phi}}{\cluster{\mu_1}}
@@
    &+ {}^t\p\braket{\tbar{1} | \commutator{\check{\s}}{\cluster{\mu_1}} | \mathrm{HF}}
-   + {}^t\p\braket{\tbar{2} | \commutator{\commutator{\check{\s}}{\cluster{\mu_1}}}{Q_2} | \mathrm{HF}}
+   \nonumber \\
+   &+ {}^t\p\braket{\tbar{2} | \commutator{\commutator{\check{\s}}{\cluster{\mu_1}}}{Q_2} | \mathrm{HF}}
    = 0 \label{eq:pted-cc2-singles-multipliers} \\
@@
   E_0
-  + \sum_{u=1}^{3}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
-  + \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
+  + \sum_{u=1}^{3}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u} \\
+  &+ \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
   &+ \braket{\tbar{1} |
@@
   \check{\Phi} + \commutator{\check{\Phi}}{T_2}
-  + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{Q_3}
   | \mathrm{HF}} \\
-  &+ \braket{\tbar{3} | \commutator{\check{\Phi}}{T_2} | \mathrm{HF}}
-  + \frac{1}{2}{}^t\p\V\p +
+  &+ \braket{\tbar{2} |
+   \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{Q_3}
+  | \mathrm{HF}} \\
+  &+ \braket{\tbar{3} | \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
+  &+ \frac{1}{2}{}^t\p\V\p +
   {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathrm{CC3}
@@
     \commutator{\check{\Phi}}{T_2}
-    | \mathrm{HF}}
-    + {}^t\p\braket{\mu_3 | \commutator{\check{\s}}{Q_3}
+    | \mathrm{HF}} \\
+    &+ {}^t\p\braket{\mu_3 | \commutator{\check{\s}}{Q_3}
      + \frac{1}{2}\commutator{\commutator{\check{\s}}{T_2}}{T_2}
@@
 
-We can construct the \emph{noniterative}
+\arrowthought{We can construct the \emph{noniterative}}
 $\Lambda\text{CCSD}(\text{T})$\autocite{Kucharski1998-qq,
@@
 inserting the \acrshort{MBPT} expansion of the converged \acrshort{CCSD}
-parameters:\autocite{Stanton1997-lx, Koch1997-nm,
+parameters.\autocite{Stanton1997-lx, Koch1997-nm,
 Eriksen2014-gy, Eriksen2015-il, Kristensen2016-od}
@@
 \begin{equation}
-  \denom{\mu_3}\tamp{\mu_3}^{*} =
-  - \braket{\mu_3 |\commutator{\Phi}{T_2^{[1]}} | \mathrm{HF}}
-  - \braket{\mu_3 | \commutator{\Phi}{T_2^{[2]}} | \mathrm{HF}}
+  \begin{aligned}
+  \denom{\mu_3}\tamp{\mu_3}^{*} &=
+  - \braket{\mu_3 |\commutator{\Phi}{T_2^{[1]}} | \mathrm{HF}} \\
+  &- \braket{\mu_3 | \commutator{\Phi}{T_2^{[2]}} | \mathrm{HF}}
   + O(3)
+\end{aligned}
 \end{equation}
@@
 
-The converged \acrshort{PTED}-\acrshort{PTED} parameters are correct up
+The converged \acrshort{PTED}-\acrshort{CCSD} parameters are correct up
 to second order in the fluctuation potential:
@@
   \braket{\tbar{3}^{[2]} | \commutator{\Phi}{T_2^{[2]}} | \mathrm{HF}}
-  +
+  \\
+  &+
   \sum_{u=1}^2 \braket{\tbar{u}^{[2]} | \commutator{\Phi}{T_3^{[2]}} | \mathrm{HF}}
@@
   +
-  {}^t\p^{[2]} \braket{\tbar{2}^{[1]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
-  + O(6)
+  {}^t\p^{[2]} \braket{\tbar{2}^{[1]} | \commutator{\s}{T_3^{[2]}} |
+  \mathrm{HF}} \\
+  &+ O(6)
   \end{aligned}
@@
 
-Let us first consider the noniterative models for the inclusion of
-connected triples.
+\newthought{Let us first consider the noniterative models for the inclusion of
+connected triples.}
 The \acrshort{PTE} scheme is particularly simple: the equations are
@@
 
-The theory and implementation for iterative \acrshort{CC} methods with
+\arrowthought{The theory and implementation for iterative \acrshort{CC}
+methods} with
 quantum/classical polarizable Hamiltonians has already been presented in
@@
       \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
-     | \mathrm{HF}}
-   + {}^t\p\braket{\mathrm{HF} |
+     | \mathrm{HF}} \nonumber \\
+   &+ {}^t\p\braket{\mathrm{HF} |
    \commutator{\check{\s}_\mathrm{N}}{\cluster{\mu_1}} | \mathrm{HF}}
@@
      + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_2}}}{T_2}
-     | \mathrm{HF}}
-  + \braket{\tbar{3} |
+     | \mathrm{HF}} \nonumber \\
+  &+ \braket{\tbar{3} |
        \commutator{\check{\Phi}}{\cluster{\mu_2}}
@@
     + \commutator{\check{\Phi}}{Q_3}
-    | \mathrm{HF}}
-    + {}^t\p\braket{\mu_1 | \check{\s} | \mathrm{HF}}
+    | \mathrm{HF}} \nonumber \\
+    &+ {}^t\p\braket{\mu_1 | \check{\s} | \mathrm{HF}}
     = 0 \label{eq:ptes-cc3-singles-amplitudes} \\
-    \denom{\mu_2}\tamp{\mu_2} &+ \braket{\mu_2 |
+    \denom{\mu_2}\tamp{\mu_2}
+    &+ \braket{\mu_2 |
     \check{\Phi} + \commutator{\check{\Phi}}{T_2}
+    | \mathrm{HF}} \nonumber \\
+    &+ \braket{\mu_2 |
     + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2}

### 2016-10-04 FrontBackmatter/Acknowledgments.tex
DONE

@@
 
-It has not been one day, it's been four years.\footnotetext{Just a perfect day \\
-You made me forget myself \\
-I thought I was \\
-Someone else, someone good \\
-\hspace*{15pt}
+\epigraph{Just a perfect day
+
+You made me forget myself
+
+I thought I was
+
+Someone else, someone good
+}{
 --- \textsc{Lou Reed}}
+
+It has not been one day, it's been four years.
 And it's not always been perfect.
@@
 To my friends from high school, for the good times we've had and for
-those we will have.\footnotetext{And all of my friends were there, \\
-and no one cared \\
-\hspace*{15pt}
---- \textsc{The Kinks}}
+those we will have.
 To the \enquote{Pisa diaspora} and in particular to Stefano, Federica,
@@
 To Lindon, the coolest organic chemist. Thanks for all the music tips
-and the amazing time at the Rock Your Head Festival.\footnotetext{Oh, I rely all of my life \\
-All of my life \\
-On my Ffunny Ffrends \\
---- \textsc{Unknown Mortal Orchestra}}
+and the amazing time at the Rock Your Head Festival.
 To Glenn, for being an awesome friend.

### 2016-10-04 FrontBackmatter/Introduction.tex
DONE

--- 2016-10-03 FrontBackmatter/Introduction.tex
@@
 
+\epigraph{
+[...] \textnorsk{og fordi jeg alltid har hatt en dragning mot det skjulte og hemmelige.}
+}{
+--- \textnorsk{\textsc{Karl Ove Knausgård}, \textit{Om Høsten}}
+}
+
 \renewcommand{\thefigure}{\Alph{figure}}
 
-\section*{What Are the Electrons Really Doing in Molecules?}\footnotetext{
-[...] \textnorsk{og fordi jeg alltid har hatt en dragning mot det skjulte og hemmelige.}
-\\
-\hspace*{15pt}
---- \textnorsk{\textsc{K. O. Knausgård}, \textit{Om Høsten}}
-}
+\section*{What Are the Electrons Really Doing in Molecules?}
 

### 2016-10-05 RobertoDiRemigioContributionsSummary.tex
Thesis ready for submission

--- 2016-10-03 Chapters/ContributionsSummary.tex
@@
-% mainfile: ../RobertoDiRemigioPhDThesis.tex
-%************************************************
-\chapter{Summary of Contributions}\label{ch:contributions-summary}
+\documentclass[notitlepage,a4paper,11pt,dvipsnames]{article}
+\usepackage{amsmath, amssymb}
+\usepackage[inline]{showlabels}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage[english]{babel}
+\usepackage[dvips]{graphicx}
+\usepackage[style=phys,
+maxcitenames=3,
+maxbibnames=5,
+minbibnames=3,
+firstinits=true,
+url=false,
+isbn=false,
+eprint=false,
+texencoding=utf8,
+bibencoding=utf8,
+autocite=superscript,
+backend=biber
+]{biblatex}
+\usepackage{booktabs}
+\usepackage{caption}
+\usepackage{subcaption}
+\captionsetup{format=hang, font=footnotesize,labelfont=bf,labelsep=space}
+\captionsetup[subfigure]{format=hang, font=footnotesize,labelfont=bf,labelsep=space}
+\usepackage{colortbl}
+\usepackage{tcolorbox}
+\usepackage[version=4]{mhchem}
 
-\epigraph{\textonehalf\,\,research \kern 20pt and \textonehalf \kern 20pt \textgreek{Tέχνη}
+\usepackage[nohyperlinks,nolist]{acronym}
+\usepackage{csquotes}
+\usepackage{braket}
+\usepackage{bm}
+\usepackage{mathtools}
+\usepackage{hyperref}
+\hypersetup{colorlinks,
+            breaklinks,
+            urlcolor=blue,
+            linkcolor=blue,
+            citecolor=blue} % Set link colors
 
-          \textonehalf\,\,observation, \kern 21pt \textonehalf \kern 20pt \textgreek{Tέχνη}
+\newcommand*{\paper}[1]{\textbf{Paper #1}}
 
-          \textonehalf\,\,training, \kern 35pt \textonehalf \kern 20pt \textgreek{Tέχνη}
-          }{
-  --- \textsc{Ezra Pound}, \textit{Canto LXXXV}}
+\title{Summary of Contributions to the Scientific Articles Included in the Doctoral Dissertation \\ {\itshape The Polarizable Continuum Model Goes Viral!}}
 
-This final Chapter provides a brief overview of the motivations, results and
-conclusions of the papers this dissertation is based on.
-All publications included in this dissertation have involved some programming effort.
-In Section \ref{sec:software} I will describe the principles we have
-striven to follow in developing our \acrshort{PCM} software library and its
-management philosophy.
-I will also give some details on the interfaces we have developed to different
-quantum chemistry codes.
-A shorter version of the material in Section \ref{sec:software} has been
-accepted for publication in the proceedings of the \emph{Producing High
-Performance and Sustainable Software for Molecular Simulation} workshop
-held at the 2015 Supercomputing Conference.\autocite{%SC15, Trott2016-pf,
-DiRemigio2016-es}
-Sections \ref{sec:relapcm}--\ref{sec:pcmopenrsp} contain very short
-summaries of the papers and a list of my contributions.
-My coauthors have all read and approved the description of contributions.
+\author{Roberto Di Remigio}
+\date{\today}
 
-\section{Software}\label{sec:software}
+\begin{document}
+\maketitle
 
-The growing complexity of quantum chemical program packages requires that an
-appropriate strategy be devised to implement new features.
-Scalability is of paramount importance, but it has become clear that
-maintainability and extensibility of the code play an equally important role in
-managing software complexity\autocite{Wilson2014-vh, ssi, nsf-s2i2,
-netherlands-escience, coderefinery, molssi} and ensuring
-scientific reproducibility.\autocite{Hatton1994-us, Hatton1997-ml,
-Hatton1997-bb, Ioannidis2005-aj, Merali2010-uy, Prinz2011-kz}
+The doctoral dissertation {\itshape The Polarizable Continuum Model Goes
+Viral!} contains 5 Chapters, 1 Appendix and 5 scientific papers. Of these, 3
+have been published, 1 has been accepted for publication and 1 has been
+submitted for publication.
 
-The idea of a \emph{modular programming paradigm} can be traced back to
-the dawn of computer science. \citeauthor{Dijkstra1968-zp}
-successfully used it in the development of the THE operating
-system in 1968,\autocite{Dijkstra1968-zp} while
-\citeauthor{Parnas1972-im} advocated it as a best practice already in
-1972.\autocite{Parnas1972-im}
-Modularity is nothing more than a \emph{divide et impera} strategy
-applied to programming and it has been recognized as beneficial in many
-other scientific computing contexts.
-New features are isolated into libraries that can be accessed by host
-programs through a well-defined \gls{API}.
+My contributions to the scientific papers are summarized in the following.
 
-In this paradigm, computational tasks are implemented into separate,
-independent and interchangeable modules.
-Modules are to be thought as computational black boxes \emph{developed},
-\emph{tested}, \emph{packaged} and \emph{distributed} independently of
-the particular host program in which they will be used.
-A well-defined \gls{API} clearly delimits the boundaries of the functionality
-offered, effectively enforcing a programming style and standardisation of the
-functionality, eventually.\autocite{Reddy2011-sd}
-
-Recently, a debate over \emph{open-source}
-licensing\autocite{St_Laurent2004-fa} and \emph{open data} practices has
-sparked in the quantum chemistry community.\autocite{Gezelter2015-gz, Krylov2015-fs,
-Jacob2016-oq}
-Open-source licensing enables an open code review process that is, in
-our opinion, essential in guaranteeing scientific
-reproducibility.\autocite{Gezelter2015-gz}
-The points raised \emph{against} wholesale adoption of open-source
-licensing models,\autocite{Krylov2015-fs} especially the concerns about
-code quality and maintenance, do partially resonate with us.
-However, we still think that a modular approach to programming can only
-have significant impact, in terms of improved reproducibility and
-overall code quality, when an open-source licensing model is adopted.
-Open-source development can fully leverage the benefits of widespread,
-cloud-based, free, code development services, such as hosted
-\acp{DVCS},\autocite{github, gitlab} continuous
-integration,\autocite{travis-ci, magnum-ci, appveyor-ci} code
-coverage analysis,\autocite{coveralls} static and dynamic code
-analyses,\autocite{coverity-scan} nightly regression testing, public
-issue tracking, code review and so forth: adoption of a modern code
-development workflow is easily within reach.
-
-It is, of course, true that the above mentioned services are not
-exclusive prerogatives of open-source projects. However, an open review
-process of scientific software can often help to establish
-reproducibility, extensibility and sustainability of the software
-ecosystem.
-Open-source software, modular development of new functionalities and
-full-fledged exploitation of \acsp{DVCS} ensure a much larger scientific
-impact. Third-parties can easily contribute to the project: by improving
-the documentation, by reporting bugs or by actively extending the
-codebase with new functionality.
-
-The \acrshort{PCM} is an ideal candidate for the creation of a solvation
-\acrshort{API}. Consider for example the schematic representation of a
-\acrshort{PCM}-\acrshort{SCF} algorithm given in Figure \ref{fig:algorithm}.
-The input to and from the \acrshort{PCM} library is limited and well-defined,
-as are the components that host quantum chemistry codes need to
-implement.
-This provides a natural \acrshort{API} design: the \acrshort{API} functions can be
-compared \emph{vis-à-vis} with the working equations derived for the
-different quantum chemical methods.
-
-\begin{figure}[tb]
-\centering
-\scalebox{0.7}{\input{gfx/algorithm.tikz}}
-\caption[Modular approach to programming a \acrshort{PCM} functionality into an existing \acrshort{SCF} code.]{
-Schematic view of a \acrshort{PCM}-\acrshort{SCF} algorithm. Computations/data in
-\textcolor{brewerBlue}{blue} are implemented by the \acrshort{PCM} \acrshort{API}.
-Computations/data in \textcolor{Green}{green} are implemented by the
-host quantum chemistry code.
-  }
-\label{fig:algorithm}
-\end{figure}
-
-The \pcmsolver library has been developed to fulfill these requirements and
-offer the community an easy-to-use implementation of the
-\acrshort{PCM}.\autocite{PCMSolver}
-\pcmsolver is written in C++ with Fortran, C and Python components.
-Architectural and conceptual organization of the code is provided by the
-C++ layer. To ensure the largest possible impact, the code is C++03 ISO standard
-compliant, with external dependencies kept to a minimum.
-Tuples, container algorithms and functional programming tools are not
-part of the standard we chose to comply to.
-Parts of the \Boost C++ libraries are used to alleviate these
-shortcomings.\autocite{boost}
-Linear algebra operations are managed through the \Eigen C++ template
-library.\autocite{eigen}
-The Python layer provides input reading and parsing facilities,
-as implemented by the \GetKw library.\autocite{libgetkw}
-Cavity construction and discretization is handled either by the Fortan
-or by the C layer, depending on which computational \emph{backend} is
-needed.
-Evaluation of the Green's functions, needed to set up the representation
-of the boundary integral operators, leverages \gls{AD}. We use the \Taylor library, a template C++
-implementation of \acrshort{AD}.\autocite{Bartholomew-Biggs2000-db, libtaylor}
-The adoption of \git as \acrshort{DVCS}, \cmake for cross-platform builds and
-automatic documentation deployment on \readthedocs simplify
-extensibility of the module and promote third-party contributions to the
-code base. Continuous integration and nightly testing offer an
-invaluable level of confidence in the code.
-
-Careful usage of the C++ object-oriented paradigm is the key to this
-wide spectrum of functionalities. The library is in itself made of
-modules, communicating by means of composition at the outermost
-level of design.
-This is an indispensable feature to achieve frictionless language
-mixing.
-Library internal classes balance the dynamic polymorphism offered by
-inheritance and the static polymorphism offered by template programming.
-\autocite{Alexandrescu2001-bp, Sutter2004-nt, Langr2012-js}
-Preeminent use of composition over inheritance keeps the coupling
-between different submodules as low as possible.\autocite{Gamma1995-fd}
-
-The library is released under the terms of the
-version 3 \gls{LGPL}, a standard open-source license.\autocite{LGPLv3}
-We strongly believe that open-source software is the key to a larger
-impact for scientific software and a standard license means lower
-thresholds to adoption and third-party contributions.\autocite{LGPLv3}
-The library serves as our development platform for the \acrshort{PCM}.
-The interfaces to different quantum chemistry codes allow us to explore
-new methodologies, significantly cutting down development times.
-Use of our \acrshort{API} \emph{significantly} limits coding effort on the
-side of the host: continuum solvation at the \acrshort{SCF} level of theory
-can be implemented in the host program almost out-of-the-box.
-
-The library was designed to allow host-\acrshort{API} communication through
-\emph{pure} functions, \ie{} free from side-effects.
-Such an implementation hinges on the theoretical realization that the
-\acrshort{PCM} layer is independent of the \acrshort{AO} or \acrshort{MO} spaces
-defined in the quantum chemical layer.
-Passing and modifying large data structures, such as the Fock and
-density matrices, can be completely avoided.
-As schematically depicted in Figure \ref{fig:algorithm},
-computation of the \acrshort{ASC} only depends on
-the electrostatic potential sampled at the cavity boundary:
-\begin{equation}
- \esp(\vect{s}) =
- \sum_A \frac{Z_A}{|\vect{R}_A-\vect{s}|} +  \sum_{\kappa\lambda} D_{\lambda\kappa}
- \left[\int\diff\vect{r}\frac{-\Omega_{\kappa\lambda}(\vect{r})}{|\vect{r}-\vect{s}|}\right],
-\end{equation}
-which is the contraction of charge attraction integrals with a, possibly
-perturbed, density matrix, a task handled by the host program.
-Formation of the solvent contribution to the Fock matrix requires the
-computation of the scalar product
-$\scalprod{\sigma(\vect{s})}{\esp_{\kappa\lambda}(\vect{s})}$
-between the \acrshort{ASC} and the charge attraction integral matrix, again
-a task easily accomplished by the host program.
-As shown in Figure \ref{fig:pcmsolver-scheme}, data transfer between
-\pcmsolver and the host is limited to the communication of
-$\esp(\vect{s})$ and $\sigma(\vect{s})$ and can be implemented
-\emph{without} storing any quantity to disk, avoiding possibly costly
-I/O operations.
-
-\begin{figure}[tb]
-  \centering
-  \scalebox{0.7}{\input{gfx/pcmsolver-scheme.tikz}}
-  \caption[Schematic view of the relationship between a host quantum chemistry
-  program and the \pcmsolver library.]{
-  Schematic view of the relationship between a host quantum chemistry
-  program and the \pcmsolver library.
-  }
-  \label{fig:pcmsolver-scheme}
-\end{figure}
-
-The \acrshort{API} is implemented in ISO C99 as a \emph{context-aware} set of
-functions.\autocite{library-ronacher, context-api-example}
-Upon initialization of the library, a \emph{context} object containing
-the state of the \acrshort{PCM} \acrshort{API} is constructed and an
-opaque handle to it (a pointer in C++ parlance) is returned to the host
-program. Further transactions between \pcmsolver and the host program
-proceed through manipulations of the context object.
-\pcmsolver introduces the concept of surface functions:
-labelled instances of any quantity defined on the cavity surface.
-Bookkeeping of a possibly arbitrary number of such quantities is managed
-through a map storing labels and pointers to the contents of the
-functions.
-Listings \ref{lst:C-host} and \ref{lst:Fortran-host} show how calls to
-\pcmsolver \acrshort{API} functions might look like in an actual C or Fortran
-program, respectively.
-
-\lstinputlisting[language=C,
-                 caption={Relevant calls to \pcmsolver in a C host program.
-                 Ellipses in the functions' argument list stand for omitted parameters.
-                 For the full, functional example visit
-                 \url{http://pcmsolver.readthedocs.io/en/v1.1.6/users/C-example.html}},
-                 label={lst:C-host}
-                ]{snippets/C_host.c}
-
-A context-aware \acrshort{API} offers a simple strategy for parallelization.
-The host program could spawn contexts on a per-process basis,
-limiting ownership and access to a context just to its parent process.
-Race conditions are trivially avoided in such a scenario.
-
-\lstinputlisting[language=Fortran,
-                 caption={Relevant calls to \pcmsolver in a Fortran host program.
-                 Notice the use of the standard Fortan ISO C bindings.
-                 Ellipses in the functions' argument list stand for omitted parameters.
-                 For the full, functional example visit
-                 \url{http://pcmsolver.readthedocs.io/en/v1.1.6/users/fortran-example.html}},
-                 label={lst:Fortran-host}
-                 ]{snippets/Fortran_host.F90}
-
-\pcmsolver currently offers an implementation of the traditional collocation
-solvers for isotropic media,\autocite{Tomasi2005-vm}
-together with some unique functionalities: the wavelet Galerkin solvers on smooth
-molecular surfaces,\autocite{Weijo2010-hy, Bugeanu2015-tp} the Green’s
-functions for spherical diffuse interfaces\autocite{DiRemigio2016-nn} and
-the delayed solvers for real-time \acrlong{TDSCF}
-simulations.\autocite{Corni2015-pe}
-
-Programming effort in this dissertation has not only been directed at the
-creation of the \pcmsolver library, but also at interfacing it with a
-number of quantum chemistry codes.
-The first interface was implemented and released within the \DIRAC code
-and is described in \paper{I}.\autocite{DIRAC15, DiRemigio2015-ou}.
-Interfaces to the \LSDALTON\autocite{LSDALTON16, Aidas2013-rp} and
-\DALTON\autocite{LSDALTON16, Aidas2013-rp} codes have followed, as
-described in \paper{II} and \paper{V}, respectively.\autocite{Bugeanu2015-tp, pcm-openrsp}
-Details about the interface to \ReSpect\autocite{ReSpect-3.5.0} can be
-found in \paper{IV}.\autocite{pcm-respect}
-Coupling with \psicode\autocite{Turney2012-de} has allowed us to implement the
-\acrshort{PCM}-\acrshort{CC} developments described in this dissertation.
-Finally, development of an interface with the KOALA code is underway.\autocite{Hofener2014-ex, Hofener2016-qz}
-
-\section{Continuum Solvation in the Relativistic Regime}\label{sec:relapcm}
+\section*{\paper{I}}
 
@@
 
-Systems containing heavy elements are notoriously challenging for quantum chemistry.
-One has to properly include the effect of special relativity in the
-quantum chemical description in order to achieve at least qualitative
-agreement with experiment.
-Additional complications arise if one needs to also include environment effects.
-Continuum models arguably represent a cost-effective strategy to achieve a first
-approximation of these effects.
-In \paper{I} we presented the first derivation and implementation of the \acrshort{PCM}
-coupled to a \acrshort{SCF} description of the solute in the fully relativistic, four-component
-regime.
-Our preliminary calculations on the group 16 dihydrides \ce{H2X} (\ce{X} =
-\ce{O}, \ce{S}, \ce{Se}, \ce{Te}, \ce{Po}) have shown that the method predicts
-a noticeable interplay of relativistic and solvent effects when heavier
-elements are involved.
-The main point of \paper{I} was, however, the adoption of a fully modular
-programming strategy. We showed that it is entirely possible to adopt the same
-\acrshort{PCM} code and implementation "checkpoints" across altogether different
-problem domains, Fig.~\ref{fig:algorithm}.
-
-As a by-product of the four-component implementation, we were able to obtain and visualize \acrshort{MEP} maps
-from four-component \acrshort{SCF} wave functions. These add yet another interpretive tool to the toolbox available
-in the four-component relativistic regime.
-The interface to \DIRAC was first released in the 2014 version,
-providing \acrshort{PCM} capabilities to the software package.
-
 I contributed the theoretical derivation of the quantum/classical polarizable
-terms in a four-component \acrshort{SCF} framework, for energies and linear response
-properties. I devised the coupling of the four-component program \DIRAC with
-\pcmsolver by providing the implementation and testing of:
+terms in a four-component self-consistent field framework, for energies and linear response
+properties. I devised the coupling of the four-component program DIRAC with
+PCMSolver by providing the implementation and testing of:
 \begin{enumerate}
-  \item \acrshort{MEP} integrals for four-component wave functions,
-  \item the additional Fock matrix contributions and
+  \item molecular electrostatic potential integrals for four-component wave functions,
+  \item the additional Fock matrix contributions, and
   \item the additional terms in the response equations.
@@
 all subsequent versions.
+My coauthors have all read and approved the description of contributions.
 
-\section{The Wavelet Galerkin Boundary Element Method for PCM}\label{sec:wemlin}
+\section*{\paper{II}}
 
@@
 
-Continuum solvation models are inherently parametrized. Apart from solvent permittivities,
-the atomic radii and molecular surface definition play a crucial role
-in determining the performance of the models.
-However, the numerical accuracy of the \acrshort{BEM} procedure used to numerically solve the
-underlying \acrshort{BIE} is a not-so-often studied aspect of these models.
-Traditionally, collocation methods have been used, but these require parametrization of some of the necessary surface integrals.
-Galerkin methods do not suffer from such a limitation and additionally preserve symmetry of the underlying
-boundary integral operators.
-The use of biorthogonal wavelet bases as finite elements achieves sparsity in
-the \acrshort{BEM} procedure, due to the intrinsical hierarchy and
-the existence of \emph{a priori} and \emph{a posteriori} matrix compression
-estimates.
-Thus, wavelet Galerkin \acrshort{BEM} represents a valid alternative to traditional
-collocation methods, both to achieve a better computational scaling and to
-provide accurate, benchmark results.\autocite{Harbrecht2004-uo,
-Harbrecht2006-ug, Dahmen2006-pj}
-
-Already~\citeauthor{Weijo2010-hy} had shown that using \gls{PWC} bases can lead
-to superior accuracy and convergence in the calculation of quantum mechanical
-molecular solvation energies.
-In this work we showed that even faster convergence can be achieved when
-\gls{PWL} bases are used instead.
-Moreover, the same holds for the calculation of static electric properties.
-Notably, the traditional collocation solver cannot guarantee the same accuracy,
-even for very large finite element bases. This suggests that, in some cases,
-\acrshort{BEM} collocation methodologies might slow down or even prevent the
-convergence of the quantum mechanical response equations solvers.
-
-For \paper{II}, I provided template interface and test sets for the cavity
-generator\autocite{Harbrecht2009-no, Harbrecht2011-dk} and wavelet Galerkin
-\acrshort{BEM} solver\autocite{Harbrecht2004-uo, Harbrecht2006-ug} with \pcmsolver.
+I provided template interface and test sets for the cavity
+generator and wavelet Galerkin
+boundary element method solver with PCMSolver.
 These were used to interface with the new C++ implementation of the wavelet
 Galerkin solvers of Monica Bugeanu.
-I implemented the interface between the \LSDALTON quantum chemistry software
-package and the \pcmsolver software library. The interface allows to run \acrshort{HF} and
-\acrshort{KS}-\acrshort{DFT} single-point and linear response calculations.
+I implemented the interface between the LSDALTON quantum chemistry software
+package and the PCMSolver software library. The interface allows to run Hartree--Fock (HF) and
+Kohn--Shame density-functional theory (KS-DFT)
+single-point and linear response calculations.
 Together with coauthor Krzysztof Mozgawa, I performed the benchmark quantum
@@
 
-Finally, the interface to \LSDALTON was later released in the 2016 version,
-providing \acrshort{PCM} capabilities to the software package.
+Finally, the interface to LSDALTON was later released in the 2016 version,
+providing polarizable continuum model (PCM) capabilities to the software package.
+My coauthors have all read and approved the description of contributions.
 
-\section{Non Homogeneous Environments}\label{sec:spherical}
+\section*{\paper{III}}
 
@@
 
-Continuum solvation models offer a simple route to the treatment of non
-homogeneous environments.
-The general integral equation formulation is in fact transparent with respect
-to the definition of the Green's function for the space portion exterior to the
-cavity.
-As outlined in Chapter \ref{ch:CSM}, the boundary integral operators in the
-\acrshort{IEF} equation:
-\begin{equation}
-  \left[ \bi{S}_\mathrm{e}\left(2\pi + \bi{D}^\dagger_\mathrm{i}\right)
-  +
-  \left(2\pi - \bi{D}_\mathrm{e}\right)\bi{S}_\mathrm{i}
-  \right]\sigma =
-  -\left[\left(2\pi-\bi{D}_\mathrm{e}\right)
-  -\bi{S}_\mathrm{e}\bi{S}_\mathrm{i}^{-1}
-  \left(2\pi-\bi{D}_\mathrm{i}\right)
-  \right]\esp,
-  \tag{\ref{eq:full-IEF} from Chapter \ref{ch:CSM}}
-\end{equation}
-can be set up once the Green's functions $\Gi$ and $\Ge$ are known.\autocite{Cances1998-og}
-\citeauthor{Frediani2004-er} showed that a numerical representation of the
-Green's function is sufficient to obtain the boundary integral operators in the
-\acrshort{PCM} integral equation.
-The authors introduced a numerical integration procedure to calculate the
-Green's function for an environment characterized by spatially varying, yet
-cilindrically symmetric, permittivity functions: a model for planar diffuse interfaces.
-
-In this work, a similar procedure was introduced to tackle diffuse interfaces
-in spherical symmetry.
-In contrast to previously existing work, our implementation offers a more robust
-treatment of thin interfaces, with a rather generic functional form for the
-permittivity profile.
-We thoroughly analyzed the necessity for the \emph{a posteriori} removal of the
-Coulomb singularity from the computed Green's function and its efficient
-implementation.
-Interface width and curvature influence the transfer of ions and molecules across
-spherically symmetric interfaces and peculiar properties may arise.
-In this work, we analyzed both effect on the water-vapor and oil-water transfer
-of \ce{Li+}, \ce{Br-}, acetone, \emph{para}-nitroaniline and the L0 dye.
-Nonelectrostatic interactions were not included in our implementation, although
-they have been proved to be crucial for non homogeneous
-environments.\autocite{Mozgawa2014-ad}
-Nevertheless our implementation represents a first significant step in the continuum treatment
-of such nontrivial environments.
-
 I contributed the theoretical work for \paper{III}, based on earlier drafts from coauthors
@@
 Moreover, I contributed the implementation and testing of the Green's function code.
-The interface to the \LSDALTON program package, developed within
+The interface to the LSDALTON program package, developed within
 \paper{II}, was
@@
 I wrote the first draft of the paper and coordinated all subsequent editing stages.
+My coauthors have all read and approved the description of contributions.
 
-\section{Relativistic Calculation of EPR and pNMR Parameters in
-Solution}\label{sec:pcmepr}
+\section*{\paper{IV}}
 
@@
 
-\paper{IV} is a step further in our exploration of the interplay between
-relativistic and solvent effects initiated with \paper{I}.
-Whereas \paper{I} presented the essential framework for the coupling of
-four-component \acrshort{SCF} wave functions with a classical polarizable continuum,
-in this paper we explored the calculation of first-order magnetic properties:
-\gls{EPR} and \gls{pNMR} parameters.\autocite{Repisky2010-ls, Malkin2011-nm,
-Komorovsky2013-xa, Cherry2016-ij}
-The two works are thus complementary since they explore two different classes
-of properties and present implementations in two algorithmically different
-relativistic quantum chemistry codes.
-In the relativistic framework, spin-orbit interactions are included from the
-outset in the variational optimization of the wave function.
-Hence, \acrshort{EPR} and \acrshort{pNMR} parameters are formulated as expectation
-values, by virtue of the Hellmann--Feynman theorem.\autocite{Konishi2009-zb,
-Helgaker2000-tz}
-Moreover, the \ReSpect code can exploit the \emph{Kramers unrestricted} formalism,
-allowing for spin polarization and thus granting facile access to the computation
-of spin-dependent properties.\autocite{Dyall2007-tu}
-The same modular programming strategy was adopted in crafting an interface between the
-relativistic four-component code \ReSpect\autocite{ReSpect-3.5.0} and \pcmsolver.
-
 My contributions to this paper include prototyping the interface between the
-\pcmsolver library and the \ReSpect quantum chemistry code.
+PCMSolver library and the ReSpect quantum chemistry code.
 The interface is maintained in collaboration with coauthor Michal Repisky, who also
@@
 I tested the interface against one-component and four-component results obtained with
-the \LSDALTON and \DIRAC codes, respectively.
+the LSDALTON and DIRAC codes, respectively.
 I helped coauthors Michal Repisky, Stanislav Komorovsky and Peter Hrobarik
-with setting up the \acrshort{PCM} calculations described in the paper.
+with setting up the PCM calculations described in the paper.
 Finally, I provided the first draft for Section 2 of the paper and took part in all editing stages.
-The interface to \ReSpect will be released in the next public version of the software
-package, providing \acrshort{PCM} and \acrshort{COSMO} capabilities.
+The interface to ReSpect will be released in the next public version of the software
+package, providing PCM and COSMO capabilities.
+My coauthors have all read and approved the description of contributions.
 
-\section{Open-Ended Self-Consistent Field Response Theory in
-Solution}\label{sec:pcmopenrsp}
+\section*{\paper{IV}}
 
@@
   \\
-  In preparation for submission to \textit{Phys. Chem. Chem. Phys.}
+  Submitted to \textit{Phys. Chem. Chem. Phys.}
   }
 \end{tcolorbox}
-
-In recent years, the availability of strong lasers has allowed to
-design and carry out experiments where the high-order response of
-molecular materials can be routinely probed.
-The more intense the light source, the more complicated the interpretation of
-the measured signal.
-Our group has newly developed an open-ended methodology for the computation of
-\acrshort{SCF} response functions\autocite{Thorvaldsen2008-sg, Ringholm2014-gx} and
-their residues.\autocite{Friese2015-kb}
-These developments offer a route towards a synergistic experimental and
-theoretical approach to high-order absorption spectroscopies.
-\paper{V} grafts a classical polarizable continuum approach to solvation on top
-of the open-ended methodology of \citeauthor{Thorvaldsen2008-sg}.
-Still nowadays, continuum models represent a cost-effective methodology for the
-approximate inclusion of solvent effects, albeit their known limitations with respect
-to specific solute-solvent interactions.
-
-I developed the theoretical framework for the open-ended \acrshort{SCF} formulation
+I developed the theoretical framework for the open-ended SCF formulation
 of molecular response properties when a quantum/classical polarizable continuum
-Hamiltonian is used.\autocite{Thorvaldsen2008-sg, Lipparini2010-be}
-I provided its implementation within the \DALTON code, by interfacing the
-\pcmsolver library and the open-ended \acrshort{SCF} response code of
-\citeauthor{Ringholm2014-gx}\autocite{Ringholm2014-gx, Friese2015-kb}
+Hamiltonian is used.
+I provided its implementation within the DALTON code, by interfacing the
+PCMSolver library and the open-ended SCF response code of Ringholm \emph{et al}.
 I performed extensive testing of the code by comparing with previously
-published implementations of the \acrshort{PCM}-\acrshort{SCF} response functions within
-\DALTON.\autocite{Cammi2003-qy, Frediani2005-nc, Ferrighi2010-pm}
-Together with coauthors Maarten T.~P.~Beerepoot and Yann Cornaton, I carried out
+published implementations of the PCM-SCF response functions within
+DALTON.
+Together with coauthors Maarten T. P. Beerepoot and Yann Cornaton, I carried out
 the multiphoton absorption calculations presented in the paper. I contributed
@@
 editing stages with coauthor Maarten T. P. Beerepoot.
+My coauthors have all read and approved the description of contributions.
+
+\begin{flushright}
+    \begin{tabular}{m{5cm}}
+      \textit{Troms{\o}, \today} \\
+        {\small \textbf{Candidate}} \\
+      \includegraphics[width=0.35\textwidth]{gfx/signature.png} \\
+        \hline
+        {\small Roberto \textsc{Di Remigio}} \\
+    \end{tabular}
+    \\
+    \begin{tabular}{m{5cm}}
+        {\small \textbf{Advisor} }
+        \\
+      \includegraphics[width=0.3\textwidth]{gfx/LucaSignature.png} \\
+        \hline
+        {\small Luca \textsc{Frediani}} \\
+    \end{tabular}
+\end{flushright}
+
+\end{document}

### 2016-10-19 FrontBackmatter/Publications.tex
Going through typos and typography
Correct typo in Arnfinn's name
Update publications

@@
   \\
-  DOI: \url{10.1063/1.4943782}
+  DOI: \href{http://dx.doi.org/10.1063/1.4943782}{10.1063/1.4943782}
   }
@@
   \\
-  DOI: \url{10.1039/C5CP03410H}
+  DOI: \href{http://dx.doi.org/10.1039/C5CP03410H}{10.1039/C5CP03410H}
   }
@@
   Four-Component Relativistic Density Functional Theory with the
-  Polarizable Continuum Model: Application to EPR Parameters
+  Polarisable Continuum Model: Application to EPR Parameters
   and Paramagnetic NMR Shifts
@@
   \\
-  DOI: \url{10.1080/00268976.2016.1239846}
+  DOI: \href{http://dx.doi.org/10.1080/00268976.2016.1239846}{10.1080/00268976.2016.1239846}
   }
@@
   \textbf{R. Di Remigio}, M. T. P. Beerepoot, Y. Cornaton, M. Ringholm,
-  A. H. S. Steindal, K. Ruud, and L. Frediani
+  A. H. Steindal, K. Ruud, and L. Frediani
   \\
-  In preparation for submission to \textit{Phys. Chem. Chem. Phys.}
+  Submitted to \textit{Phys. Chem. Chem. Phys.}
   }

### 2016-12-05 FrontBackmatter/Errata.tex
Add Errata Corrige to the submitted and approved version of the dissertation

\pagestyle{empty}

\pdfbookmark[1]{Errata Corrige}{errata}
\chapter*{Errata Corrige}

\begin{itemize}
\item On page 23. The labelling of the cavity boundary in Figure 2.1 is
inconsistent. \enquote{$\Gamma\equiv\Omega_\mathrm{i}$} should be replaced by
\enquote{$\Gamma\equiv\partial\Omega_\mathrm{i}$}, as in the following Figure.
  \begin{figure*}[!h]
    \centering
    \includegraphics[width=.5\textwidth]{IEF-correct.pdf}
  \end{figure*}
\item On page viii, the name of one of the coauthors was not abbreviated
correctly: \enquote{A.~H.~S.~Steindal} should be replaced by \enquote{A.~H.~Steindal}.
\item On page 96, the name of one of the coauthors was not abbreviated
correctly: \enquote{A.~H.~S.~Steindal} should be replaced by \enquote{A.~H.~Steindal}.
\item On page 210, the name of one of the coauthors was not abbreviated
correctly: \enquote{A.~H.~S.~Steindal} should be replaced by \enquote{A.~H.~Steindal}.
\item On page vii, the spelling of the title of Paper IV has been changed to
reflect the publisher's editorial policy. \enquote{Polarizable} should be
replaced by \enquote{Polarisable}.
\item On page 94, the spelling of the title of Paper IV has been changed to
reflect the publisher's editorial policy. \enquote{Polarizable} should be
replaced by \enquote{Polarisable}.
\item On page 186, the spelling of the title of Paper IV has been changed to
reflect the publisher's editorial policy. \enquote{Polarizable} should be
replaced by \enquote{Polarisable}.
\item On page 126, there is a misprint in the title for reference 225.
\enquote{RNa} and \enquote{DNa} should be replaced by \enquote{RNA} and
\enquote{DNA}, respectively.
\item On page viii, \enquote{Submitted to \textit{Phys. Chem. Chem. Phys.}}
should be replaced by \enquote{\textit{Phys. Chem. Chem. Phys.}, 2017, DOI:
\url{10.1039/C6CP06814F}} to reflect the fact that the paper was published in
final form on the journal's website.
\item On page 96, \enquote{Submitted to \textit{Phys. Chem. Chem. Phys.}}
should be replaced by \enquote{\textit{Phys. Chem. Chem. Phys.}, 2017, DOI:
\url{10.1039/C6CP06814F}} to reflect the fact that the paper was published in
final form on the journal's website.
\item On page 210, \enquote{Submitted to \textit{Phys. Chem. Chem. Phys.}}
should be replaced by \enquote{\textit{Phys. Chem. Chem. Phys.}, 2017, DOI:
\url{10.1039/C6CP06814F}} to reflect the fact that the paper was published in
final form on the journal's website.
\item Paper IV was accepted for publication on September 16, 2016. The paper
was published in final form on the journal's website shortly thereafter. The
\emph{verbatim} copy included in the dissertation reflects the accepted version
of Paper IV.
\item Paper V was accepted for publication on November 15, 2016. The paper was
published in final form on the journal's website shortly thereafter. The
\emph{verbatim} copy included in the dissertation reflects the submitted
version of Paper V, which underwent minor revisions after the peer-review
process.
\end{itemize}

### 2017-01-27 Chapters/ContributionsSummary.tex
Correct typo in Arnfinn's name
Update info for last two papers
This is the final printed version

@@
   \textsf{Four-Component Relativistic Density Functional Theory with the
-  Polarizable Continuum Model: Application to EPR Parameters
+  Polarisable Continuum Model: Application to EPR Parameters
   and Paramagnetic NMR Shifts
@@
   \textbf{R. Di Remigio}, M. T. P. Beerepoot, Y. Cornaton, M. Ringholm,
-  A. H. S. Steindal, K. Ruud, and L. Frediani
+  A. H. Steindal, K. Ruud, and L. Frediani
   \\

### 2017-01-27 FrontBackmatter/Publications.tex
Update info for last two papers
This is the final printed version

@@
   \\
-  DOI: \href{http://dx.doi.org/10.1063/1.4943782}{10.1063/1.4943782}
+  DOI: \url{10.1063/1.4943782}
   }
@@
   \\
-  DOI: \href{http://dx.doi.org/10.1039/C5CP03410H}{10.1039/C5CP03410H}
+  DOI: \url{10.1039/C5CP03410H}
   }
@@
   \\
-  DOI: \href{http://dx.doi.org/10.1080/00268976.2016.1239846}{10.1080/00268976.2016.1239846}
+  DOI: \url{10.1080/00268976.2016.1239846}
   }
@@
 \renewcommand{\descriptionlabel}[1]{\hspace*{\labelsep}\spacedlowsmallcaps{#1}}
+
+\section*{Notes Added in Proof}
+
+\begin{itemize}
+\item \paper{IV} was published online in final form on October 18, 2016.
+  The \emph{verbatim} copy included in the dissertation reflects the
+  accepted version of \paper{IV}.
+\item \paper{V} was accepted for publication on November 15, 2016 and
+  assigned the DOI \url{10.1039/C6CP06814F}.
+  The paper was published online in final form on December 1, 2016.
+  The \emph{verbatim} copy included in the dissertation reflects the
+  submitted version of \paper{V}, which underwent minor revisions after
+  the peer-review process.
+\end{itemize}
